### 2024-08-20 Paper.qmd
Preparing manuscript.

---
title: "Child Signals of Need in Utila, Honduras"
format: html
editor: visual
bibliography: signaling.bib
---

```{r setup, include=FALSE}
source("analysisforpaper.R", local = knitr::knit_global())

```

# Introduction

Despite the complexity of childhood and adolescence and their importance to adult outcomes, the behavioral strategies of children are often overlooked, especially when it comes to how they elicit support from parents and other caretakers [@gottlieb2000; @wells2003; @hirschfeld2002]. To gain support, children use signals to communicate their desires to parents. Infant signals of need, particularly infant cries, have been extensively studied in a range of populations [@barr1991; @fouts2005; @konner1974; @kruger2010; @kushnick2006; @meehan2017; @müller2021; @tronick1987], yet the full range of potential child signals of need, which likely extend beyond cries, has never been investigated in children of all ages. In particular, if and when children signal need with costly signals [@grafen1990; @spence1973; @zahavi1975a] has rarely been investigated (for one exception, see [@syme2023]).

We argue that older children (i.e., those older than infants) have evolved strategies to tailor their signals in response to 1) their valuation of a given resource, 2) the extent of conflicts of interest with caretakers, and 3) the presence of existing information about a child's need or potential to benefit from a resource so that they can credibly convey they are worth additional investment. Such strategies can lead to some combination of verbal requests for more support, facial sadness, crying, withdrawal, temper tantrums and self-harm, all of which may be effective in certain circumstances and are sometimes successful in eliciting more support from caretakers, despite a child's low status and formidability.

## Interdependence and investment within the family

Throughout human evolutionary history, individuals often found themselves in situations in which they received fitness-relevant benefits from others that could not easily be replaced [@apicella2012; @apicella2019; @hill2011; @tooby1996]. In these instances, natural selection is expected to have favored those who provided support to others if 1) such help meant one could expect to receive fitness benefits from those helped in the future and 2) the costs of providing help were outweighed by the benefits to the helper on average[@andré2010; @hamilton1964; @tooby1996; @trivers1971]. At the very least, this would have taken the form of mothers providing investment to their children due to direct fitness benefits.

The fitness interdependence of offspring and at least one parent is seen in many animals [@clutton-brock1991; @webb1999]. However, humans are a uniquely interdependent species among the living primates [@tomasello2012], with one’s fitness prospects depending greatly on a network of kin and other social allies [@apicella2019]. Although also true with adults, this need for others is most intense with children, who are born highly altricial and require many years of investment to develop to the point where they can obtain enough food to survive on their own and develop the social skills needed to succeed in their communities[@kaplan2009; @rosenberg2021; @schniter2015]. Much of this investment comes from parents[@kramer2018], with most human reproduction occurring in the context of biparental investment and long-term pair bonds [@apostolou2007; @schacht2019]. However, in many societies, other adults like grandparents, aunts, uncles, and non-relatives have also been found to provide significant investment in children [@kramer2018; @sear2008], something which was likely favored due to indirect fitness benefits, reciprocity, or reputation management [@gurven2000] + (Schniter Preprint (permission needed). Children also often serve as alloparents, helping to raise their younger siblings or provide for them through food preparation or collection [@hrdy2009; @meehan2009; @sear2008], despite still requiring substantial amounts of investment themselves [@kramer2018a].

## Conflicts of interest within the family

Although the fitness prospects of both children and familial caretakers are largely intertwined, evolutionary theory predicts that disagreements about the amount of investment from parents will be ubiquitous [@fouts2005; @godfray1995; @godfray1992; @trivers1974]. Most resources that translate to fitness benefits are limited so that one unit of a resource invested in one area cannot be used in another. In response to this basic fact of life, humans, like other organisms, are expected to have evolved to efficiently allocate resources between somatic effort (i.e., growth and maintenance) and reproductive effort (i.e., mating and parenting effort) [@daly1983], something which often pits the fitness interests of parents against the fitness interests of offspring.

Conflicts of interest within families are largely structured by the likelihood an individual shares their genes with another. As a sexually reproducing species, humans have a 50% chance of passing down a given gene to each child. Each child, however, shares more genes with themselves (100%) than with their full siblings (50% on average), half-siblings (25% on average), and stepsiblings (0%). Due to this asymmetry, children are often expected to attempt to elicit more investment from parents than is optimal from the perspective of the parent, while parents are expected to attempt to invest less than is optimal for each child in order to invest in other siblings, whom parents are expected to value identically, all else being equal [@trivers1974].

For the same reason, siblings also often find themselves in conflicts of interest with one another, with each being expected to prefer more resources for themself than others. Sibling competition is evident across taxa and can range from relatively benign forms like strong begging competition in chicks [@caro2016] to siblicide in birds, mammals, and sharks [@salmon2014]. In humans, sibling competition is largely estimated by counts of siblings, with large numbers of siblings having been associated with decreases in growth [@hadley2011; @hagen2006; @salmon2014], survival [@lawson2012; @meij2009; @strassmann2011], and reproductive success [@gibson2011; @mace1996] in multiple populations, highlighting the likely significance of sibling competition for resources, even if this does not always manifest with behavioral conflict [@bateson1994].

While parent-offspring conflict and sibling competition are likely ubiquitous, the resources being contested will vary across human groups. In relatively egalitarian and food constrained populations, conflict over resources is likely to be centered on food, which probably explains the negative impacts siblings have on child age-specific heights and weights among the Shuar of Ecuador [@hagen2009], although similar effects have also been found in industrialized populations [@lawson2008]. In societies with greater inequality, competition is more likely to be over wealth, as is the case with older brothers in pastoralist communities like the Gabbra and Arsi Oromo limiting the marriage prospects of laterborn sons [@gibson2011; @mace1996]. Compared to large-scale investments like bride price or inheritance, little is understood about smaller transfers of wealth to younger children despite the potential importance of such transfers to 1) facilitate the development of embodied capital outside of growth and fat reserves or 2) help children succeed in their peer groups.

Conflicts of interest between children and parents may also involve how much children invest in the household. As previously mentioned, each individual child is expected to prefer greater investment in oneself than in one’s siblings, while parents, who are equally related to all their children, should often be expected to favor a more even distribution [@trivers1974]. This can lead to conflict with older children who might prefer to spend their time in ways which are likely to result in future direct fitness benefits (e.g., finding romantic partners or forming strong bonds with friends) rather than taking care of their siblings or acquiring or preparing food. Such a dynamic has been found among the Chuukese, with both male and female children reporting conflict with their families over children’s desires to invest financial resources in their own status enhancing opportunities rather than in the household [@lowe2003]. Similarly, Syme and Hagen [@syme2023] found greater numbers of siblings were associated with increased likelihoods of severe conflict in a retrospective study with Chuukese-Americans. This may also be consistent with parents pressuring older children to invest in the household; however, there are other ways in which limited resources can result in behavioral conflict [@syme2023].

Conflicts of interest over investment in the household also appear to extend to when a child starts having their own children [@kushnick2023; @emlen1995]. Due to the benefits children may provide their siblings, parents may be motivated to delay or subvert the reproductive careers of certain offspring [@flinn1989; @emlen1992], a conflict which is likely enhanced in families with step-parents due to the likelihood of future children being step-siblings rather than full siblings, thereby reducing the potential fitness benefits of alloparenting [@moya2014; @schaffnit2021]. Situations can also arise in which it might pay parents for their children to start having their own children sooner than would be ideal for the adult child. For example, parents may be motivated for a child to leave the household and begin having children earlier than is ideal from the child’s perspective if their presence in the household leads to them taking investment from other children who caretakers deem more worthy of investment [@kushnick2023].

Studies testing the existence of familial conflict over the timing of child reproduction are rare and direct tests on the existence of parent-offspring conflict over when a child should start their own reproductive career versus help existing siblings in humans are needed [@schaffnit2021]. The bulk of evidence for the existence of such conflicts comes from work with non-human animals. For example, the cooperative breeding white-fronted bee-eaters (*Merops bullockoides*) will interrupt the courtship of their offspring, seemingly to continue to benefit from their undiluted assistance in digging nests, incubating the clutch, finding food for nestlings, and defending the nest [@emlen1992]. When it comes to humans, Flinn found that rural Trinidadian females contributed substantial amounts of childcare and those with siblings under 10 years old had their first child at a later age [@flinn1989], both of which would be expected if there is a conflict of interest over investing in one’s siblings or one’s own reproduction. However, a study among Karo Batak women found no evidence of delayed reproduction for alloparenting benefits [@kushnick2006].

## Signaling

In order to elicit more support, children often use signals. Signals are a key source of information for individuals in social species and are commonly used to indicate one's value as a mate, one’s formidability to competitors, or one’s worthiness of investment to parents and other social partners [@caro2016; @maynard-smith2004; @preuschoft2000; @setchell2005; @zahavi1975a]. Signals differ from cues in that signals were favored by natural selection due to their ability to alter the behavior or perceptions of others in ways that benefit the signaler, whereas cues are pieces of information organisms use to influence their behavior or physiological state that did not evolve for communicative effect [@maynardsmith1995]. Crying is likely a signal of need[\[1\]](#_ftn1), for instance, while bleeding would be a cue of need.

When conflicts of interest over investment exist, there  may be much to gain from deception (i.e., exaggerating or completely fabricating one’s quality or need) so that one can receive the benefits that come from being seen as having a particular quality without actually doing so [@zahavi1993]. Although the risks of being deceived can be substantial, signal receivers may still benefit from paying attention to signals so long as the information is credible enough to inform one’s behavior in ways that are adaptive on average [@bliegebird2001; @grafen1990]. These benefits coupled with the likelihood of strong selection against individuals who respond to fake signals results in the expectation that signals within species will often need to benefit both signaler and receiver to persist evolutionarily [@hasson1994; @maynard-smith2004].

Credible signals[\[2\]](#_ftn2) can evolve and be maintained in multiple ways. Despite reasonable expectations that such circumstances will be rare [@zahavi1993], signals can be produced and trusted in instances when there are no conflicts of interest over what information is conveyed [@guilford1995; @maynardsmith1995]. They may also arise when signals cannot be faked (i.e., indexical signals) or in instances when there are greater net payoffs to signalers who have the quality being signaled to a greater extent than others [@maynardsmith1995]. Although the latter can arise through differential benefits [@penn2020; @tiokhin2016] or a combination of differential costs and benefits (Bergstrom preprint), more attention has been paid to the importance of differential costs alone, with costly signaling being a possible pathway to credible signaling whenever signals are prohibitively costly for those who might fake the signal but still worth making for those with the underlying trait [@maynard-smith2004].

Regardless of the mechanism behind the credibility of a signal, credible signals work by revealing private information about one’s qualities to others. In the context of child signals of need, a child has access to information about their internal state, which is unavailable to others, that allows them to estimate their valuation of additional investment. By signaling in a credible way, the child reveals their resource valuation to potential caretakers who may be skeptical of the child’s claims due to the common conflicts of interest described above. With the new information revealed by the signal, caretakers should generally be more likely to provide additional support due to their fitness interest in the child’s wellbeing. When this is the case, all parties are better off than they would be without the signal (so long as the benefits outweigh the costs of the signal to child and parent), with caretakers benefitting from the revelation of fitness-relevant information and the child benefitting from the increased likelihood of a beneficial response.

However, even if a child was completely successful in conveying their needs to caretakers, it may not be enough to result in a beneficial response. This is because belief in one’s child’s need is not the only factor determining the form and amount of support given by parents. Instead, parents are expected to weigh the costs and benefits of support with other investment options [@daly1987; @daly1983]. Ancestrally, this likely often meant evaluating the viability of the child as well as how much a particular resource benefited one child relative to other children or future children [@clutton-brock1991; @daly1983; @dawkins1976; @hagen1999]. The former is thought to have resulted in children having to demonstrate their vigor early in life [@soltis2004], while the latter is expected to result in context-specific support [@beaulieu2008; @frankenhuis2022]. For example, one unit of a resource is often expected to be worth more to a newborn than an adolescent, favoring that investment flowing to the newborn all else being equal. However, if resources are in short supply, parents may benefit more from investing more in older children as they are closer to reaching adulthood and have higher reproductive value [@daly1983]. Additionally, a parent may also decide against investing in a child despite deeming investment in a child worthwhile if time is not right (e.g., if a mother preferred to breastfeed during rest rather than foraging) [@altmann1980]. When this is the case, outward conflicts may exaggerate the extent of a conflict of interest and involve parents teaching children ways to behave that benefit them both [@altmann1980].

## Crying

Crying is the most studied way by which children may signal need. Crying is a human universal [@gracanin2018] with few suspected non-signaling functions outside of the role vocal cries may play in the development of vocalization [@wells2003] and the functions tears have in terms of cleaning, protecting, and lubricating the eye [@murube2009; @sznycer]. Hunger, pain, anger, fear, and separation are all common causes of crying [@vingerhoets2013; @wells2003], although tear production also often happens with joyful events [@sznycer; @vingerhoets2013; @zickfeld].

Infant cries have been found to result in increased amounts of care, resources, and attention they receive and increased proximity to parents [@lin2012]. For example, Karo Batak infants who cried more were breastfed more, and the length of crying bouts predicted the time mothers breastfed [@kushnick2006]. Similarly, !Kung San mothers frequently respond to crying infants by providing comfort and breastfeeding access, with comforting behaviors, in general, being far more common than one’s meant to restrict infant behavior [@kruger2010]. Although this study found mothers were the most common responders, siblings, fathers, and other adults frequently stepped in so that 88% of crying bouts were responded to, with most of the cries not responded to lasting under ten seconds [@kruger2010]. Similar responsiveness has also been reported among the Aka [@hewlett1998], Mbendjele hunter-gatherers [@chaudhary2024], the Efe [@tronick1987], and the Ache, among whom young children have been described as “win\[ning\] most conflicts with parents simply by crying and whining” [@kaplan1987]. Studies examining the potential for long-term benefits are rare and difficult to achieve; however, during a time of high infant mortality in a sample of 13 Massai infants, those with easy temperaments were more likely to die (5/7) than those with difficult temperaments (1/6), something which is consistent with signals resulting in beneficial treatment [@devries1987].

Crying does not always succeed in eliciting benefits, however. Compared to the previously mentioned groups, the Ngandu [@hewlett1998] and the Hadza [@blurtonjones1993] have been found to be more likely to ignore the cries of infants and young children, and concerns over the risk of spoiling children by being too responsive seem to be reasonably common in some Western populations [@barnett2010; @burchinal2010; @chaudhary2024; @frankenhuis2022; @maute2018]. Crying can also lead to negative responses beyond being ignored [@barr1990; @soltis2004; @zeifman2017]. It has been found to increase parental distress, the likelihood of maternal depression, and the child’s risk of maltreatment ~~especially if the child is in poor condition~~ [@frodi1985; @fujiwara2011; @howard2006; @lin2012; @soltis2004]. For example, in a sample of 3,259 parents of Dutch infants under 6 months old, 5.6% self-reported smothering, slapping, or shaking their babies in response to their crying, outcomes which were more likely among parents who were single, unemployed, or deemed their infant’s crying to be excessive [@reijneveld2004]. Excessive infant crying, i.e., colic, has also been found to be a risk factor for multiple negative outcomes [@zeevenhooven2018], including abuse, shaken baby syndrome [@kurth2011], and earlier cessation of breastfeeding [@howard2006].

Although widely accepted as a signal, there is no consensus on why emotional crying may work to elicit positive responses. One possibility is that crying is physiologically constrained in ways that prevent or limit faking. Evidence consistent with this comes from findings that crying is hard to do on command and findings individuals often cry while alone, something which is not predicted by signaling theory [@vingerhoets2001]. However, the likelihood of crying is often clearly related to presence of supporters, something which is seen in findings that lonely individuals cry less than those with more support options despite reporting lower well-being [@vingerhoets2013] and the existence of delayed crying—times when crying is inhibited during the cause of negative affect only for it to be expressed when one is near potential supporters [@gracanin2018]. A more likely possibility is that certain acoustic elements of infant cries are tied to their physiological state in ways that make faking certain states impossible, with crying serving as a graded signal in most other instances [@soltis2004; @zeifman2017].

It is likely that the functions of crying and the mechanisms behind its ability to elicit pro-signaler responses change over one’s life. Compared to adults and older children, the physiological costs of crying are greater for newborns and infants, among whom crying can result in a substantial increase in energy expenditure [@rao1997; @thureen1998]. Due to these costs, crying at certain intensities or for long periods of time may be prohibitively costly for infants with low survival prospects, making it a potential credible signal of vigor [@furlow1997; @lummaa1998; @soltis2004; @wells2003]. Such signals are thought to be important due to parental adaptations which are suspected to take the viability of offspring into account when making decisions about whether it is worth investing in a specific child or withdrawing investment to instead invest in other children who are more likely to survive to reproductive age and be more competitive adults [@hagen1999; @lummaa1998]. In line with this reasoning, most infanticides cross-culturally occur when a child has poor survival prospects, due to either health conditions or resource constraints [@daly1984; @soltis2004].

Another often suggested means by which crying could result in credible signals of need is if the increased risk of predation that comes from loud vocalizations allows for signaling to be worthwhile for those in relatively high need but not for those with low need [@lummaa1998; @wells2003; @zahavi1977]. Evidence of increased predation has been found in studies on birds [@leech1997; @lima2009; @redondo1992; @ryan1982], among whom louder broods in general have also been found to be more conspicuous to predators [@haskell1997; @leech1997]. It is unclear how significant the risk of alerting predators would be ancestrally. However, there are ethnographic reports of mothers keeping infants and young children close in dangerous environments [@hewlett1998; @kaplan1987; @konner1972], something which likely lessens this risk[\[MG10\]](#_msocom_10) .

As children age, their reproductive value increases due to reduced odds of mortality, leading to a decreased risk of infanticide (Hagen 2004)[\[MG11\]](#_msocom_11) . For this reason, there is likely a transition period in which crying begins to convey information about how much a child values further investment. Many models have been put forward which attempt to explain human crying and begging in other species as credible signals of need due to the differential costs and benefits of crying and analogous vocalizations in other species as deterring the exaggeration of need [@godfray1991; @kilner1997; @parker2002]. Although differing in their details, a common theme is that the indirect fitness costs signalers incur from imposing costs on that parents and siblings through begging (be them energetic costs, survival costs, or reduced likelihood of future siblings) creates a floor to possible deception but falls short of preventing it [@daly1990; @wells2003].

When it comes to older children, the relatively insignificant energetic costs of crying are unlikely to be enough to result in credible signals in times of substantial conflicts of interest. For this reason, crying among older children may be more similar to adult crying, something which has been found to result in increased support and empathy but sometimes come at the cost of reputation and the potential for stigmatization [@cornelius2001; @cretser1982; @hendriks2006; @jesser1989; @vingerhoets2016]. Such costs may be a means by which crying can still be a credible signal [@sznycer] with moderate conflicts of interest if they dissuade potential fakers from making the signal.

## Withdrawal

Withdrawal has been suggested to be an underappreciated way those with low social power may credibly signal need [@syme2023]. Due to their lower status, children have limited ways to change how others treat them. However, as children 1) are their parents’ path to direct fitness and 2) can help family members through allocare and other forms of labor, they provide benefits that are valuable and hard to replace. By withholding these benefits or putting the future of these benefits into question, a child simultaneously delays the benefits they get from this cooperation while imposing costs on those with a fitness interest in the child. As these costs are likely lower for those truly in need than those who are not, withdrawal may be a credible way to signal need if the costs to faking the signal make it unprofitable. In these cases, the end result is analogous to a labor strike, in which the signaler conveys to others their situation demands more support, and if they want to continue receiving benefits from the child, they need to upregulate their support of the child.

One of the key proximate mechanisms behind withdrawal is sadness or low mood [@scherer1994]\[\[3\]\](#\_ftn3). Sadness is a type of negative affect that is generally thought to arise from loss [@horwitz2017; @levenson1999; @smith1993], especially when one has little power to reverse or mitigate the loss on one’s own [@emanuel2023; @gunderson2021]. Sadness and low mood, like other forms of psychological pain, probably evolved to motivate victims of adversity to shift their attention to the causes of adversity to mitigate its negative fitness consequences and to learn to solve or avoid future such adverse events [@andrews2009; @nesse1990; @thornhill1989]\[\[MG12\]\](#\_msocom_12). Low mood, for example, is often thought to be adaptive in situations when activity is dangerous or unlikely to result in benefits, with anhedonia, pessimism, and fatigue all helping reduce behavior that may not be worth the costs or that may cause additional problems [@nesse2019; @nettle2009]. It may also encourage the reappraisal of one’s circumstances after new information comes to light that one’s current strategy is not working [@nesse2009].

Through the reduction in behavior and attention in one’s usual interests that can accompany sadness and low-mood [@nettle2009], one’s relatives and other social partners would likely gain some indication, in the form of cues, that the sad individual was experiencing adversity and could benefit from increased support. In addition to this cue-based information, sadness also results in facial expressions and crying, both of which probably evolved to indicate need [@balsters2013; @bowlby1980; @reed2017; @darwin1872]. Such expressions seem to be important to conveying this information as is seen by responses to those who cannot produce certain elements of facial expressions like tear production [@hasson2009; @nesse2019] or certain facial movements [@bogart2012; @bogart2010]. However, facial expressions, along with tears, may not be enough to credibly signal need in times of significant conflicts of interest. In these instances, depression and suicidal behavior have been suggested to function as ways to credibly signal need due to the expectation that such states are costlier for those facing adversity than those who are not (for more detail on these approaches see: [@hagen2003; @rosenthal1993; @syme2016; @watson2002].

## Temper tantrums and self-harm

Temper tantrums may also be adaptive strategies by which children can elicit more support from parents. They are common in children from 18 months to 4 years old [@potegal2003], and similar behaviors have also been reported in other young animals, often surrounding feeding or weaning conflict [@altmann1980; @goodall1986; @schaller1964]. Temper tantrums vary in their severity, ranging from crying, shrieking, and screaming to violent bodily movements, aggression, and self-harm [@potegal2003]. They have been reported in many societies with different subsistence systems, political organization, and norms surrounding anger and parenting [@albino1956; @banisalameh2021; @fouts2005; @henderson1966; @hilger1957; @konner1974; @ward1970], suggesting they may be a human universal. As with the other proposed signals of need, they are capable of eliciting a wide range of responses, including pro-child responses, neutral responses (e.g., parents ignoring the child) [@hilger1957; @konner1974], and negative responses like abuse [@schmitt1987].

As with the vocalizations that come with crying, temper tantrums might motivate caretakers to upregulate their support of the signaler due to the energetic costs of tantrums or the increased risk of predation stemming from the noisy display, both of which might result in credible signals if they are costly enough to not be worth faking [@lummaa1998]. In addition, tantrums also impose direct costs on caretakers who may be less able to do fitness enhancing tasks [@altmann1980] like finding or preparing food or taking care of other children.

Another possibility is that temper tantrums carry the risk or implication of offspring self-harm. For example, Utilian mothers reported in focus groups that some temper tantrums involved children banging their heads on things when they do not get what they want, something that has been reported in other populations [@belden2008; @burrows1953; @daniels2012; @dennis1940; @vandenakker2022]. Similarly, young chimpanzees (*Pan troglodytes*) have been observed to beat their heads on the ground when refused access to breastmilk [@goodall1986], and American white pelican chicks (*Pelecanus erythrorhynchos*) have been found to throw themselves on the ground, flail violently, and bite at their own wings when begging for food [@schaller1964]. Such behaviors have been suggested to follow similar logic to depression and suicide attempts as they are likely relatively less costly to those who are truly in dire need compared to those who are not to the point faking is not worth the costs [@hagen2008] (Hagen et al. 2008). In other words, the risk of injury from violent temper tantrums carries greater fitness costs to a child who is receiving enough support to develop high levels of embodied capital than it does to a child who is neglected to the point where their survival prospects are low. In this scenario, the risk of harm is likely not worth the potential benefits of increased support to the child receiving substantial investment, but it may be for the neglected child who may have few other options to compel others to invest in them.

## Study Aims

This study was designed to help establish the natural history of child signaling behavior in Utila, Honduras, by quantifying the frequencies of common child signals of need and various types of parental responses. From this, we sought to examine 1) the causes of child signals of need, 2) how and why signal costs might vary, 3) how parents respond to indications of child need, 4) how signaling behavior changes with age and sex, 5) and if there were differences in child signaling and parental responses among two sub-populations.

A major challenge when studying signals of need is the fact that such signaling is expected to 1) be more common in times when individuals can benefit from increased investment (e.g., when they are facing major adversity or when a fitness-relevant opportunity arises) and 2) result in benefits to the signaler on average. This hypothisized causal, bi-directional relationship between investment and signaling makes it difficult to isolate the effects of signaling behavior on investment in children when using standard measures of child welfare (i.e., BMI, health metrics, and height for age). When experiments are not appropriate, longitudinal studies with carefully selected variables can help address this (cite). However, our small sample size with signaling data and anthropometric data at two time periods prevents this for this study. For this reason, all analyses for this study were exploratory and meant to inform future studies of child signaling.

# Methods

## *Study population*

This study was done in Utila, Honduras, one of the country’s three major Bay Islands in the Western Caribbean. Participants in the study largely fit into two groups who face different levels of economic and social adversity, 1) a long-standing English-speaking community made up of expats and their descendants, and 2) Spanish-speaking immigrants from the mainland. Compared to those born on Utila, immigrants from the mainland [@davis2014] are more likely to live in crowded and unsanitary conditions, face active discrimination [@garcia2017], and experience economic uncertainty due the predominance of low paying, seasonal jobs primarily related to the island’s diving tourism industry [@garcia2018; @stonich2000]. Such inequality appears to be increasing in response to recent growth in the tourism industry, with the majority of wealth from tourism being enjoyed by a small portion of expat islanders and westerners [@davis2014].

During focus groups held in July 2022, parents indicated high levels of sibling competition and various ways children might signal need. Both Spanish-speaking immigrants and English-speaking island natives reported that fighting between siblings was extremely common, with fights over toys, gifts, and phone time being the most reported causes. In both populations, crying and various forms of withdrawal were listed as the most common responses to a child not receiving investment or support, followed by temper tantrums and making messes in their home. Spanish-speaking mothers reported that refusing to do homework was a common tactic, while English-speaking mothers emphasized running away and self-harm (e.g., a child slamming their head on the ground) to a greater degree. In response to these hypothesized signals, discipline was common. Although the mothers generally reported that children were treated equally, needier children were described as getting more resources from parents, something that may be partially mediated by differences in signaling among children in addition to information caretakers gained from cues.

## *Sampling*

Participants for this study were recruited by a variety of means. JH recruited early participants through convenience and snowball sampling. Participants were also recruited via Facebook posts, and fliers were sent home with children the Utila's Methodist school.

In all forms of recruitment, adult participants were informed of the purpose of the study, what they would be asked to do, and the inclusion criteria, which included being over 18, residing on Utila, and having a child between 6 and 18 years old.

XXXXXXXXCHIld RecruitmentXXXXXXXXXXX

## *Participants*

All participants for the survey portion of the study met the inclusion criteria. Most participants were mothers of at least one biological child in the study, with the participant being the biological mother of the child in 357 of 375 caregiver/child dyads (our imputation).

Child participants for the anthropometric portion of the study were at least 5 years of age. In the majority of instances, child participants had a primary caretaker who took the survey portion of the study first, although some children participated before parents. To have their data included in this study, a caretaker in their household must have taken the survey.

## *Ethics*

All adult participants provided informed consent. Surveys were designed to last about hour plus 20 minutes for each additional child. Adults were paid \$XXXXX for the survey portion of the study.

xxxxxxxxxxxxxChild consent infoxxxxxxxxxxxxxx.

This study was approved by the xxxxxxxIRBINFOxxxxxxxxxxxxx.

## *Survey*

The survey data for this study come from a larger project run by the Utila Child Health Project. Data was collected by Jess Hlay, Izabel Rodriguez, Daniela Mejia, Angela Garcia, or Caroline Smith who read the survey questions to participants and recorded their answers. Participants could choose to do the survey in English or Spanish.

### *Frequency of child signaling, alloparenting contributions, and conflict with caretaker*

We used 6-point Likert scales to get caregiver reports of 1) child signaling, 2) interpersonal conflict with each participating child, 3) and child alloparenting behavior. See Table 1 for exact wording.

|                                 |                                                                                                                   |
|-------------------------|-----------------------------------------------|
| Variable                        | Survey question wording                                                                                           |
| Crying Frequency                | How often does your child cry?                                                                                    |
| Sad Frequency                   | How often is your child sad?                                                                                      |
| Tantrum Frequency               | How often does your child throw temper tantrums? (e.g., scream, throws things, or throws themselves on the floor) |
| Running Away Frequency          | How often does your child leave the house without telling you when they get mad?                                  |
| Parent/Child Conflict Frequency | How often do you have conflicts with your child?                                                                  |
| Child Alloparenting Frequency   | How often does your child take care of their other siblings?                                                      |

**Table 1**. Table showing variable names and question wording for measures about the frequency of behaviors. All questions had the following labels attached to each point on its Likert scale: Never; Once a month or less; More than once a month but less than once a week; More than once a week but not daily; Daily; Multiple times per day.

As these responses reflect frequencies of actual behavior, we converted these to counts which reflected the number of times per month a child engaged in these behaviors. The original wording of the Likert scale and the frequency conversion are shown in Table 2.

|                                                  |                               |
|------------------------------------------------|-----------------------------------------------|
| Likert scale wording                             | Count value (times per month) |
| Never                                            | 0                             |
| Once a month or less                             | 1                             |
| More than once a month but less than once a week | 3                             |
| More than once a week but not daily              | 8                             |
| Daily                                            | 30                            |
| Multiple times per day                           | 60                            |

**Table 2.** Table showing conversions from Likert scale to count value for the frequencies of child behaviors included in this study. Variables created with this scheme include SadFreqN, CryFreqN, TantrumFreqN, RunawayFreqN, ConflictFreqN, and AlloparentingFreqN.

From these count variables, we created three additional variables to be used, primarily, as our main dependent variables in analyses. SignalFreq was the sum of the frequencies of child sadness, crying, and temper tantrums. SignalFreqMax was the frequency of the most common of these three signals. SignalCost was the sum of the frequencies of these three signals weighted by their relative costliness (i.e., SignalCost = SadFreqN + CryFreqN\*2 + TantrumFreqN\*3). The frequency of running away was not included in the creation of any of these variables due to uncertainty over whether it was likely to be a credible signal due to the relative safety of Utila compared to other areas in which running away has been found to be a successful strategy for children.

### *Relative neediness of children and relative investment in children within the family*

We also used 6-point Likert scales to get parental reports of 1) the relative neediness of each participating child and 2) the relative investment they provide to each participating child compared to the rest of their children. See Table 3 for exact wording.

|                     |                                                                                                  |
|-------------------------|-----------------------------------------------|
| Variable            | Survey question wording                                                                          |
| Relative Need       | How much time, effort, and resources does this child need compared to your other children?       |
| Relative Investment | How much time, effort, and resources do you put into this child compared to your other children? |

**Table 3.** Table showing variable names and question wording for measures about relative need and investment. Both questions had the following labels attached to each point on its Likert scale: Much less; Less; The same amount; More; Much more.

### *Last instance of child signaling*

We also asked caretakers about the last time each participating child was sad or cried. This involved providing parents with a free response text box to describe the cause of the emotion that lead to the display. We then asked how the parent responded to their child. The researcher giving the survey would then bin responses into one of 31 categories which included positive, negative, and neutral responses. For the full list of options see Table S1.

### *Child anthropometrics*

All anthropometric measurements were collected in accordance with anthropometric standards [@lohman1988] and involved taking the mean of two measurements. Before being measured, participants were asked to remove their shoes and heavy articles of clothing. Height was measured to the nearest millimeter using an anthropometer, and weight was measured to the nearest 100 g with a digital scale (model name) that used within a building with a concrete foundation. All participants were wearing lightweight clothing, so we made no adjustments for clothing weight. Grip strength was measured using a dynamometer (model name), with participants seated/standing? and told to xxxxxxxx. XXXX time was passed between the first and second measurements.

We computed each child's height, weight, and BMI Zscores with respect to the WHO child growth standards (cite).

### *Perceptions of child immune function (or current susceptibility to illness)*

Caretakers were also asked three questions using a five point Likert scale about their child's current health which involved the likelihood of them becoming sick. We then inverted the scores for two measures so that a score of 5 corresponded with the lowest susceptibility to illness before averaging the scores of these to create the CurrentHealthMean variable. Means were still calculated if 1-2 values had an NA (only xxx cases), with the numerator being the number of variables with data. See Table 4 for the wording of these variables.

|                |                                                                                       |
|-------------------------|-----------------------------------------------|
| Variable       | Survey question wording                                                               |
| Currenthealth1 | If an illness is 'going around', my child will get it.                                |
| Currenthealth2 | In general, my child is very susceptible to colds, flu, and other infectious diseases |
| Currenthealth3 | My child’s immune system protects him/her from most illnesses that other kids get.    |

: **Table 4.** Table showing variable names and question wording for measures about a child's current likelihood of becoming sick. All three questions were asked using a Likert scale from 1-5 where 1 was labeled "Strongly disagree" and 5 was labeled "Strongly agree." The scores for Currenthealth1 and Currenthealth2 were inverted so that 5 corresponded with the lowest susceptibility to illness.

### *Demographic questions*

The survey also included many demographic questions about the survey taker and their households. Those used for this study included the (1) ages of those in the household, (2) sexes of those in the household, (3) partnership status of the primary caregiver, (4) number of adults who provide allocare within the household, (5) highest level of education of the primary caregiver, (5) immigration status (in regards to Utila), and (6) monthly household income. This data allowed us to calculate (1) the number of children within the household and (2) the number of adults within the household.

CHANGE TO TOTAL CHILDCARE

DECIDE on alloparentingFreqN2

We also calculated the number of individuals who engaged in childcare within the household, by combining the number of adults who provide care within the household with data from child alloparenting frequencies. Due to the fact the adult childcare variable was a count of helpers and the child alloparenting variable was frequency based, we created 3 total alloparenting variables as a robustness check. "OtherAlloparents" is the sum of the number of adults who provide childcare plus the sum of other children who engaged in alloprarenting at least "more than once a week but not daily." The second and third involved summing the alloparenting frequency (AlloparentingFreqN) of other children within the household. We then summed this with the frequency of adult childcare, which either scored each adult who was listed as providing childcare as helping 30 times per month (AlloparentingFreqNa) or 60 times per month (AlloparentingFreqNb).

## *Statistical analyses*

We fit generalized linear mixed models for exploratory analyses with signal frequency and summed signal cost as dependent variables. As our signaling variables invovled over-dispersed count data, we used negative binomial regression. This was done in R with the GLMM package [@brooks2017], with householdID as a random effect.

For the models examining 1) parental responses to the last reported time each participating child was sad or cried, 2) the relative neediness of the child, and 3) the investment a child received compared to the other children in the household, we fit ordinal logistic regression models using two methods. The first used the polr function in the MASS package [@venables2002]. The results were then interpreted with clustered standard errors for householdID. Due to the risk of overfitting that comes with having too few clusters [@cameron2015], we also fit the same models using BRMS with householdID as a random effect.

## *Data availability*

The anonymized data used in the study are available at xxxxxxxLINKxxxxxxx. Study code is available at xxxxxxLINKxxxxxxxxxx.

# Results

```{r sumtab_children}
#| echo: false
#| label: sumtab_children
#| fig-cap: "Summary statistics for child study variables."

vars_c <- c(
  ChildAge = "Age (years)",
  SadFreqN = "Sad frequency (times per month)",
  CryFreqN = "Crying frequency (times per month)",
  TantrumFreqN = "Temper tantrum frequency (times per month)",
  AlloparentingFreqN = "Frequency of child's own alloparetning effort (times per month)",
  ConflictFreqN = "Frequency of conflict with focal caretaker (times per month)",
  SignalFreq = "Frequency of signals summed (times per month)",
  SignalCost = "Signal cost",
  SignalFreqMax = "Frequency of most common signal (times per month)",
  Currenthealth1 = "Currenthealth1",
  Currenthealth2 = "Currenthealth2",
  Currenthealth3 = "Currenthealth3",
  CurrentHealthMean = "Mean of current health measures"
)

kable(custom.summarize(d2, vars_c)) %>% kable_styling()

#kable(custom.summarize(d2, vars), caption = "Summary statistics for child study variables.") %>% kable_styling()

```

```{r sumtab_adults}
#| echo: false
#| label: sumtab_cadults
#| fig-cap: "Summary statistics for adult and household study variables."

vars_c <- c(
  CaregiverAge = "Age (years)",
  EducationLevelYears = "Education level (years)",
  IncomeCategoryN = "Household income (Lempira)"
)

kable(custom.summarize(d2, vars_c)) %>% kable_styling()

#kable(custom.summarize(d2, vars), caption = "Summary statistics for child study variables.") %>% kable_styling()

```

# Discussion

*Negative association between age and child signaling (all signaling measures)*

As seen in Western cultures [@zeifman2013; @potegal2003; @koch2003] (check for others), the frequencies of crying and temper tantrums decreased with age, something which is likely due to numerous reasons. One possibility is that older children have greater ability to satisfy one's desires and behave in ways which match cultural expectations of older children and adults. Another is that as children survive to the point where they would have been unlikely to die young in ancestral environments, conflicts of interest with parents decrease due to children passing this important milestone and requiring less investment to reach adulthood than younger children, current or future (cite daly and wilson). For this reason, less costly signals may be more likely to work to increase investment, all else being equal\[\[4\]\]. The decrease in crying and tantrum frequencies with age may also result from the potential for greater reputational costs, although this may also increase the credibility of such signals among older children as previously mentioned. It is also possible that parents may see older children less than younger children. However, the commonality of decreased signals of need with age, suggest that this is unlikely to have a large effect with this population.

There was no significant association between the frequency of child sadness and age. Part of this likely results from the fact that, unlike crying, sadness is thought to have important nonsignaling functions including drawing one’s attention to a problem, motivating deeper analysis of one’s situation, and causing withdrawal from situations with low fitness prospects.

The frequency of running away increased with age. This may be due to younger children being observed more closely or due to older children experiencing less danger from leaving the home alone (more on this below).

*Positive association between conflict with parent and child signaling (all signaling measures)*

Consistent with signaling models which emphasize the role of conflicts of interest in favoring the use of costlier signals (cite this lab and others), there were positive associations between caretakers' perceptions of conflict frequency and 1) the summed costliness of children's signaling, 2) the frequency of child signals of need, and 3) the frequency of children's most common need signaling strategy. Although conflicts of interest do not always lead to behavioral conflict [@altmann1980; @bateson1994], it is likely many, if not most, conflicts are ultimately caused by conflicts of interest, making it a useful proxy for the underlying conflicts of interest that may increase the benefits of deception and the costs of being deceived\[\[5\]\]. If this is the case, the association between signaling and conflict may represent children's attempts at credibly signaling their value of additional investment or the costliness of punishment to parents who would have good reason to doubt the trustworthiness of simple verbal request (or the trustworthiness of facial sadness and crying for those resorting to temper tantrums).

When viewed in isolation, the frequencies of sadness, crying, and throwing temper tantrums were also positively associated with conflict frequency between child and caretaker. However, the frequency of running away was not significantly associated with conflicts with parents. A retrospective study of parent-offspring conflict among Chon Chuuk immigrants in the northwest United States found that running away was the most common response to long-standing conflicts with parents and often resulted in beneficial outcomes for the child [@syme2023]. The threat running away poses to child wellbeing and, therefore, its ability to credibly signal information about one’s current situation is likely highly dependent on one’s socioecology and age. Among the Chon Chuuk, running away is often seen as a suicide threat [@fischer1950; @syme2023], likely due to xxxxxxxxx. Similar risks may also occur in environments with large predators, venomous animals, or high degrees of human conflict. In these instances, even one night away can involve substantial risk as is seen among the !Kung where a woman running away and spending the night outside of camp can be used to communicate “the strength of her feelings” about a an unhappy relationship [@shostak1981]\[\[6\]\]. When it comes to Utila, mothers in focus groups reported that running away on Utila was much less risky than it would be on mainland Honduras, with a few mothers mentioning that they warned their children about the difference in danger that came with running away on the mainland. Despite the relative safety of Utila, young children may still face risks out on their own due to proximity to the ocean, large dogs, and fast-moving motorcycles, golf carts, and tuk-tuks on the main road. For these reasons, running away may be too costly for young children but not costly enough to credibly signal need for older children.

*Sex and focal child alloparenting effort interact (postive for females, negative/flat for males)*

One of the most reliable findings across the models tested was an interaction between how much a given child contributed allocare within the household and child sex, whereby the slope of the relationship was postitive or flat for girls and flat or negative for boys. In many societies, girls are often expected to do more alloparenting within the household and often end up doing so (cite). While girls in this sample were also reported to give more allocare, the difference was not significant. However, Utilian mothers reported in focus groups (confirm) that girls tend to provide more help for their siblings than boys. If this difference was the outcome of caretakers pressuring girls to do more (or resulted in the expectation of more alloparenting), the increase in signaling with alloparenting effort seen among girls may be the result of a strategy to signal the costliness of their efforts in order to invest more in romantic relationships and status seeking behaviors outside the household.

Evidence for conflicts similar to the one hypothesized here has been reported by multiple researchers working with the Chuukese. Lowe found that both male and female children reported conflict with their families over their desires to invest financial resources in their own status enhancing opportunities rather than in the household [@lowe2003]. In an analysis of conflict among Chuukese-Americans, Syme and Hagen found that severe conflicts were more likely with greater numbers of siblings and when conflicts involved a child spending time away from home or evading family responsibilities [@syme2023], findings which would be consistent with parents pressuring older children into investing more in the household among other explanations. In response to such conflicts, participants reported using a wide range strategies as children including, deception, withholding cooperation, running away, and seeking the assistance of others. Through these means, participants were often able to resolve conflicts in their favor despite their lower status in the household and community.

*Lack of other sex differences*

Add

*In most models, there is a negative association between education and child signaling?*

One unexpected finding was a negative association between caretakers’ education level (years in school) and many of our child signaling measures despite our models controlling for household income and partner status. One possible reason for this may be that more educated caretakers see their children less during the day due to greater ability to afford childcare or due to them having jobs that are less compatible with child care (test model of social support predicted by education). Unfortunately, we do not have data on how much time caretakers spent with children to test this, and it seems equally likely that less educated parents may see their kids less often if they need to work more hours to support their families. Another possibility is that lower levels of education are associated with greater adversity for both parents and children, something which would increase the benefits of signaling need for children and the difficulty of parents to respond. It is also possible that education influences parenting norms in ways that cause differences in child signaling behavior.

*Negative association between caretaker estimates of child health and child signaling for some measures. This does not extend to the costlier signals.*

A negative association between caretaker estimates of child immune function and child signaling was seen, suggesting that signaling behavior was influenced by adversity. Perceptions of greater immune susceptibility were associated with both the summed costliness of signaling and the frequency of the most common child signal of need, while the frequency of child signaling was marginally significant. When examining the effect of perceived immune function on each signal, perceptions of greater immune susceptibility predicted the frequency of sadness but not the frequencies of crying and temper tantrums.

One explanation for this may be that parents can easily detect child illness and that conflicts of interest tend to be less significant when it comes to illness care than other areas. Pathogens have been a strong selection pressure for life on earth for much of its existence and have greatly shaped evolutionary history (cite). In response to this, humans have evovled many adapatations for limiting the harm we experience from pathogens, including the ability to identify relaible indicators of illness in others. In some contexts, this activates disgust and motivates avoidance of sick inidivudals. However, it can also lead to increased care of relatives and allies (cite), with there being some suggestion that we have even evolved to upregulate some of our symptoms beyond what would arise from an immune response/pathogen interaction to better convince others that we are sick (cite). Due to this evolutionary history, parents may be better equipped to identify the possible benefits to investing in a sick child when compared to investment aimed at helping a child in status competition (e.g., buying new or popular clothing), something which may be exacerbated by culture change in what is popular among children. In other words, there is less private information for a child to reveal. This coupled with the likelihood that conflicts of interest between children and caretakers are smaller when it comes to illness care compared to other areas means there is less to gain from costlier signals.

## **Limitations**

A major challenge when studying signals of need is the fact that such signaling is expected to 1) be more common in times when individuals can benefit from increased investment (e.g., when they are facing major adversity or when a fitness-relevant opportunity arises) and 2) result in benefits to the signaler on average. In other words, higher levels of signaling may stem from greater levels of need while simultaneously eliciting more investment from others. This hypothisized causal, bi-directional relationship makes it difficult to isolate the effects of signaling behavior on investment in children when using standard measures of child welfare (i.e., BMI, health metrics, and height for age). Longitudinal studies can help address this. However, our small sample size of participants with signaling data and anthropometric data at two time periods prevents this for this study.

Due to the fact the data used here were a part of survey which covered many topics and often multiple children, we sometimes faced a tradeoff between survey time and question quality. Asking adult participants about their relative investment in children in their household was likely not as effective as more in depth, single child focused questions due to both the relative lack of detail it provides and the potential for social desirability bias [@grimm2010], which could lead adult participants to be be more likely to report equal investment across children. Our questions about the last time the child signaled may have also been more effective if they focused on specific signals due to our methods combining crying, which is unambiguous as a signal, with sadness, which has both signaling and non-signaling functions.

Despite seeking data from both English-speaking participants born on the island Spanish-speaking immigrants from the mainland, our ability to detect differences between the populations may limited by our smaller sample of English-speaking Utilians. It is possible that differences in wealth and status between the groups translate to differences in child signaling due to children facing differerent types adversity or weighting different types of investment differently. It is also likely that the two groups differ in how caretakers could respond to child signaling, and caretakers between the two populations may also differ in their cultural models of parenting.

**Footnotes**

[\[1\]](#_ftnref1) We expect “signals of need” to actually be signals of resource valuation. However, we use “signal of need” here to distinguish signals that involve attempts at eliciting support through revealing private information about one’s vulnerabilities from the more commonly studied signals which reveal private information about one’s positive qualities (i.e., those related to partner choice and intrasexual competition).

[\[2\]](#_ftnref2) Our use of “credible” signaling is largely synonymous with the more common term, honest signaling. However, we prefer the term credible signals since signals do not need to be completely honest for natural selection to have favored individuals who paid attention to them.

[\[3\]](#_ftnref3) Researchers often distinguish between sadness and low mood based on the length of negative affect. We remain agnostic on how different sadness and low mood are from one another but use the terms researchers used when describing their research.

[\[4\]](#_ftnref4) All else is often not equal when it comes to the costliness of investment as older children may often cost more (e.g., the costs education and food are often higher among older children). However, parents would still have to pay similar costs to younger children as they grew, something which natural selection has likely acted on when it comes to our adaptations involved with differential child investment.

[\[5\]](#_ftnref5) Even in instances where conflicts begin due to mistakes, the initial conflict can lead to increased conflicts of interest. This may be seen with findings that individuals who accidentally harm another often grow to dislike them (cite), something which has been suggested to be adaptive to the accidental perpetrator as the unintended victim is likely to value than they would have before, potentially to the point of wishing them harm.

[\[6\]](#_ftnref6) Despite the risk, Shostak notes there were no known recent incidents of harm to women who ran away to spend the night in the Bush [@shostak1981].

# Supplementary Information

List of options for research assistant to classify participant responses to the question of how they responded when their child was last sad or cried.

Positive: Give resources (e.g., toys, food, money)\
Positive: Promise resources\
Positive: Give attention/affection\
Positive: Promise attention/affection\
Positive: Give opportunities (e.g., loosen restrictions, new privileges, time with friends, fun activities)\
Positive: Promise opportunities\
Positive: Remove responsibilities (e.g., fewer chores)\
Positive: Promise removal of responsibilities\
Positive: Take child's side in conflict (e.g., taking child's side against sibling, intervening against bully)\
Positive: Give advice\
Positive: Other \[Text Box\]\
Neutral: Distract\
Neutral: Do nothing\
Neutral: Explain (e.g., go over why the child cannot have or do something)\
Neutral: Other \[Text Box\]\
Negative: Threaten removal of resources\
Negative: Remove resources (e.g., toys, food, money)\
Negative: Threaten withholding of attention/affection\
Negative: Withhold attention/affection (e.g., silent treatment, ignoring)\
Negative: Threaten loss of opportunities (e.g., freedom to move around the island, beach time, timeout, time with friends)\
Negative: Loss of opportunities\
Negative: Threaten more responsibilities (e.g., chores, alloparenting, homework)\
Negative: Add responsibilities\
Negative: Support adversary (e.g., taking sibling's side)\
Negative: Threaten physical punishment\
Negative: Punish physically\
Negative: Punish verbally (includes yelling)\
Negative: Downplay adversity\
Negative: Other \[Text Box\]\
NA (Can't remember, refuse to answer, etc)\
Other: Other \[Text Box\]

[\[MG1\]](#_msoanchor_1)Alternative start: Although exceptions exist (cite),

[\[MG2\]](#_msoanchor_2)Lumping together various models for the evolution of cooperation with either delayed benefits or inclusive fitness benefits. Not sure if this would confuse readers.

[\[MG3\]](#_msoanchor_3)Probably worth mentioning studies which fail to find evidence for it and why that might me (largely alloparenting benefits of other children)

[\[MG4\]](#_msoanchor_4)Find and cite others, if we decide this much detail is worth including. I am not sure it is.

[\[MG5\]](#_msoanchor_5)Removed paragraph over conflict over marriage partners due to it not seeming highly relevant given our data.

[\[MG6\]](#_msoanchor_6)Probably worth adding (see xxxxx for exceptions)

[\[MG7\]](#_msoanchor_7)Might also be worth mentioning signals between siblings (e.g., Roulin 2001 or the negotiating need section in Wells 2003).

[\[MG8\]](#_msoanchor_8)Consider reordering so that the adaptive problem precedes crying?

[\[MG9\]](#_msoanchor_9)Not sure if this paragraph is worth the space.

[\[MG10\]](#_msoanchor_10)Hadza children appear to be an exception  (Blurton Jones paper)

[\[MG11\]](#_msoanchor_11)Comment in Soltis. Also worth fitting in the possibility that children may use crying to annoy parents until they give in.

[\[MG12\]](#_msoanchor_12)Del Giudice, 2018 too?

### 2024-08-20 analysisforpaper.R

@@ -1,0 +1,1098 @@
+#' ---
+#' title: Child signaling
+#' ---
+#' # Child signaling
+#' **Data**: interviews with caretakers (usually mothers) in 157 families, with data
+#' on 375 children (more complete data on ~242)
+#'
+#' **Principle outcome variable**: child signaling frequency (Likert scale converted to times per month)
+#'
+#' **Secondary outcome variable**: child signaling cost (sad = low, crying = moderate, tantrums = high)
+#' (SadFreq + CryFreq\*2 + TantrumFreq\*3)
+#'
+#' **Tertiary outcome variable**: parental response to the most recent instance of child signaling
+#'
+#' **Predictor variables**:
+#'
+#' * ChildAge
+#' * Sex
+#' * Other children in the household (OtherChildrenHH)
+#' * LogIncome (median value of range for each option on the Likert Scale)
+#' * Number of adults, including the caretaker taking the survey (number_adults)
+#' * Partner status of caretaker (PartnerStatus: Partnered/Unpartnered)
+#' * Frequency of conflict between child and caretaker (Likert scale converted to times per month)
+#' * Frequency of the child's alloparenting contributions as a Likert scale converted to times per month (AlloparentingFreqN)
+#' * Caretaker's education level in years (EducationLevelYears)
+#' * Mean of caretaker's responses to three child health measures (CurrentHealthMean):
+#'
+#'   1) If an illness is 'going around', my child will get it.
+#'
+#'   2) My child’s immune system protects him/her from most illnesses that other kids get.
+#'
+#'   3) In general, my child is very susceptible to colds, flu, and other infectious diseases.
+#' * Random effects grouping variable (householdID)
+#'
+#' **Results**:
+#'
+#' * Negative association between age and child signaling (all signaling measures)
+#' * Positive association between conflict with parent and child signaling (all signaling measures)
+#' * Sex and focal child alloparenting effort interact (postive for females, negative/flat for males)
+#' * In most models, there is a negative association between education and child signaling?
[…]

### 2024-08-20 recode.R

@@ -1,0 +1,496 @@
+# unclear how long some punishments are
+
+# HRAF style coding
+# Vars:
+# within family conflict; conflict with family over peers; purely conflict outside the family; death, household adversity, investment desire, punishment
+# loss of privileges
+com1 <- c(
+  "Asked for play station, told to work for it" = "Low", # not in prereg
+  "Because he's bored" = "Low", # not in prereg
+  "Because her mom was sad" = "Low", # not in prereg
+  "Because mom broke her phone" = "Moderate",
+  "Because of hisndad" = "Unclear",
+  "Because of the family financial struggle" = "severity",
+  "Because she doesn't have a good relationship with her dad" = "severity",
+  "Because she had to get a shower" = "Low",
+  "Because she was not allowed to use the phone" = "Low",
+  "Because she was sad" = "Unclear",
+  "because sister was crying" = "Low",
+  "Because they didn't celebrate her birthday" = "severity",
+  "benched during baseball game" = "Low",
+  "Brother took phone away" = "Low",
+  "can't remember" = "Unclear",
+  "Casi nunca esta triste o llorando" = "Unclear",
+  "castigo" = "Low", # could be more severe
+  "Couldn't go swimming" = "Low",
+  "Dad promised something and he couldn't get it" = "Low", # not in prereg
+  "Dad scared her with a gecko" = "Low",
+  "Dad yelled at her" = "Low",
+  "didn't bring her to the interview" = "Low",
+  "Doesn't remember" = "Unclear",
+  "Doesn't want to clean" = "Low",
+  "DR" = "Unclear",
+  "el quiere que hacer una cosa como salir y su mama dice que no puede" = "Low",
+  "Emotional moment between them" = "Unclear",
+  "felt like mom wasn't giving him enough attention" = "Low",
+  "Fighting with brother" = "Low", # not clear if physical
+  "Financial struggles of the family" = "severity",
+  "For being messy and getting scolded" = "Low",
+  "For being punished" = "Low",
+  "Fought her sister" = "Low",  # not clear if physical
+  "Fought with brother" = "Low", # not clear if physical
+  "got grounded" = "severity", # not clear if long-term
+  "Got grounded" = "severity",  # not clear if long-term
+  "Got in a fight with siblings" = "Low", # not clear if physical
+  "Got in a fight with siblings or when he wants something" = "Low", # not clear if physical
+  "Got into a fight with younger brother" = "Low", # not clear if physical
[…]

### 2024-08-23 Paper.qmd
Pared down analysis script and started adding tables and figures to the paper.

@@
 title: "Child Signals of Need in Utila, Honduras"
+execute: 
+  echo: false
+  warning: false
 format: html
 editor: visual
+tbl-cap-location: bottom
 bibliography: signaling.bib
@@
 We used 6-point Likert scales to get caregiver reports of 1) child signaling, 2) interpersonal conflict with each participating child, 3) and child alloparenting behavior.
-See Table 1 for exact wording.
+See @tbl-freqvars for exact wording.
 
@@
 
-**Table 1**.
-Table showing variable names and question wording for measures about the frequency of behaviors.
+: Table showing variable names and question wording for measures about the frequency of behaviors.
 All questions had the following labels attached to each point on its Likert scale: Never; Once a month or less; More than once a month but less than once a week; More than once a week but not daily; Daily; Multiple times per day.
+{#tbl-freqvars}
 
 As these responses reflect frequencies of actual behavior, we converted these to counts which reflected the number of times per month a child engaged in these behaviors.
-The original wording of the Likert scale and the frequency conversion are shown in Table 2.
+The original wording of the Likert scale and the frequency conversion are shown in @tbl-convertedfreqs.
 
@@
 
-**Table 2.** Table showing conversions from Likert scale to count value for the frequencies of child behaviors included in this study.
+: Table showing conversions from Likert scale to count value for the frequencies of child behaviors included in this study.
 Variables created with this scheme include SadFreqN, CryFreqN, TantrumFreqN, RunawayFreqN, ConflictFreqN, and AlloparentingFreqN.
+{#tbl-convertedfreqs}
 
@@
 We also used 6-point Likert scales to get parental reports of 1) the relative neediness of each participating child and 2) the relative investment they provide to each participating child compared to the rest of their children.
-See Table 3 for exact wording.
+See Table @tbl-relvars for exact wording.
 
@@
 
-**Table 3.** Table showing variable names and question wording for measures about relative need and investment.
+: Table showing variable names and question wording for measures about relative need and investment.
 Both questions had the following labels attached to each point on its Likert scale: Much less; Less; The same amount; More; Much more.
+{#tbl-relvars}
 
@@
 Means were still calculated if 1-2 values had an NA (only xxx cases), with the numerator being the number of variables with data.
-See Table 4 for the wording of these variables.
+See Table @tbl-healthvars for the wording of these variables.
 
@@
 
-: **Table 4.** Table showing variable names and question wording for measures about a child's current likelihood of becoming sick.
+: Table showing variable names and question wording for measures about a child's current likelihood of becoming sick.
 All three questions were asked using a Likert scale from 1-5 where 1 was labeled "Strongly disagree" and 5 was labeled "Strongly agree." The scores for Currenthealth1 and Currenthealth2 were inverted so that 5 corresponded with the lowest susceptibility to illness.
+{#tbl-healthvars}
 
@@
 ```{r sumtab_children}
-#| echo: false
-#| label: sumtab_children
-#| fig-cap: "Summary statistics for child study variables."
-
+#| label: tbl-sumtab_children
+#| tbl-cap: Summary statistics for child study variables.
 vars_c <- c(
@@
 ```{r sumtab_adults}
-#| echo: false
-#| label: sumtab_cadults
-#| fig-cap: "Summary statistics for adult and household study variables."
+#| label: tbl-sumtab_adults
+#| tbl-cap: Summary statistics for adult and household study variables.
 
@@
 ```
+
+## Distributions of signaling behavior, alloparenting effort, and conflict
+
+:::{#fig-barplot_SignalFreq}
+```{r}
+barplot_SignalFreq
+```
+
+caption
+:::
+
+:::{#fig-barplot_ConflictFreq}
+```{r}
+barplot_conflict
+```
+
+caption
+:::
+
+:::{#fig-barplot_AlloparentingFreq}
+```{r}
+barplot_alloparenting
+```
+
+caption
+:::
+
+## Signaling behavior is negatively associated with age
+
+```{r}
+#| label: tbl-main_models
+#| tbl-cap: Regression tables for our main exploratory models.
+r_vars <- c(
+  ChildAge = "Age (years)",
+  SexMale = "Sex (male)",
+  OtherChildrenHH = "Other children",
+  LogIncome = "Income (log)",
+  number_adults = "Adults in home",
+  PartnerStatusUnpartnered = "Partner status (single)",
+  EducationLevelYears = "Parent ed. (years)",
+  ConflictFreqN = "Conflict freq.",
+  SignalFreq = "Signal freq.",
+  SignalCost = "Signal cost",
+  SignalFreqMax = "Max signal freq.",
+  CurrentHealthMean = "Percieved immunity",
+  SadFreqN = "Sadness freq.",
+  CryFreqN = "Crying freq.",
+  TantrumFreqN = "Tantrum freq.",
+  AlloparentingFreqN = "Child allocare"
+)
+modelsummary(list(Sadness = msadH, Crying = mcryH, Tantrum = mtantrumH, SignalFreq = msfH, SignalCost = mscH, SignalFreqMax = mmsfH), gof_map = "nobs", coef_rename = r_vars, stars = TRUE, coef_omit = "(Intercept)")
+```
+
+:::{#fig-effect1}
+```{r}
+plot_predictions(msadH, condition = "ChildAge", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Frequency of sadness (per month)") + xlab("Child age (years)") + xlim(5,20)
+```
+
+The effect of the age on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior.
+Bars are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
+Each dot is 1 child.
+Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate).
+See @tbl-main_models for the corresponding regression table.
+:::
+
+:::{#fig-effect2}
+```{r}
+plot_predictions(mcryH, condition = "ChildAge", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Crying frequency (per month)") + xlab("Child age (years)") + xlim(5,20)
+```
+
+combine with other age plots and use caption 1
+:::
+
+:::{#fig-effect3}
+```{r}
+plot_predictions(mtantrumH, condition = "ChildAge", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Tantrum frequency (per month)") + xlab("Child age (years)") + xlim(5,20)
+```
+
+combine with other age plots and use caption 1
+:::
+
+:::{#fig-effect4}
+```{r}
+plot_predictions(mscH, condition = "ChildAge", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Summed signal cost") + xlab("Child age (years)") + xlim(5,20)
+```
+
+combine with other age plots and use caption 1
+:::
+
+:::{#fig-effect5}
+```{r}
+plot_predictions(msfH, condition = "ChildAge", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Summed frequency of signals") + xlab("Child age (years)") + xlim(5,20)
+```
+
+combine with other age plots and use caption 1
+:::
+
+:::{#fig-effect6}
+```{r}
+plot_predictions(mmsfH, condition = "ChildAge", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Frequency of most common signal") + xlab("Child age (years)") + xlim(5,20)
+```
+
+combine with other age plots and use caption 1
+:::
+
+## Signaling behavior is positively associated with conflict with primary caretakers
+
+:::{#fig-effect1c}
+```{r}
+plot_predictions(msadH, condition = "ConflictFreqN", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Frequency of sadness (per month)") + xlab("Frequency of conflict (per month)") 
+```
+
+The effect of the frequency of conflict with one's primary caretaker on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior.
+Bars are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
+Each dot is 1 child.
+Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate).
+See @tbl-main_models for the corresponding regression table.
+:::
+
+:::{#fig-effect2c}
+```{r}
+plot_predictions(mcryH, condition = "ConflictFreqN", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Crying frequency (per month)") + xlab("Frequency of conflict (per month)")
+```
+
+combine with other confict plots and use caption 1
+:::
+
+:::{#fig-effect3c}
+```{r}
+plot_predictions(mtantrumH, condition = "ConflictFreqN", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Tantrum frequency (per month)") + xlab("Frequency of conflict (per month)")
+```
+
+combine with other conflict plots and use caption 1
+:::
+
+:::{#fig-effect4c}
+```{r}
+plot_predictions(mscH, condition = "ConflictFreqN", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Summed signal cost") + xlab("Frequency of conflict (per month)")
+```
+
+combine with other conflict plots and use caption 1
+:::
+
+:::{#fig-effect5c}
+```{r}
+plot_predictions(msfH, condition = "ConflictFreqN", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Summed frequency of signals") + xlab("Frequency of conflict (per month)") 
+```
+
+combine with other conflict plots and use caption 1
+:::
+
+:::{#fig-effect6c}
+```{r}
+plot_predictions(mmsfH, condition = "ConflictFreqN", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Frequency of most common signal") + xlab("Frequency of conflict (per month)")
+```
+
+combine with other conflict plots and use caption 1
+:::
+
+## Interaction between child sex and child alloparenting
+
+## Summed signal cost and tantrum frequency are negatively associated with body fat
+
+```{r}
+#| label: tbl-bodyfat_models
+#| tbl-cap: Regression tables for the models examining body fat measures.
+r_vars <- c(
+  ChildAge = "Age (years)",
+  SexMale = "Sex (male)",
+  OtherChildrenHH = "Other children",
+  LogIncome = "Income (log)",
+  number_adults = "Adults in home",
+  PartnerStatusUnpartnered = "Partner status (single)",
+  EducationLevelYears = "Parent ed. (years)",
+  ConflictFreqN = "Conflict freq.",
+  SignalFreq = "Signal freq.",
+  SignalCost = "Signal cost",
+  SignalFreqMax = "Max signal freq.",
+  CurrentHealthMean = "Percieved immunity",
+  SadFreqN = "Sadness freq.",
+  CryFreqN = "Crying freq.",
+  TantrumFreqN = "Tantrum freq.",
+  AlloparentingFreqN = "Child allocare",
+  TricepR = "Tricep skinfold (residuals)",
+  SubscapR = "Subscapular skinfold (residuals)",
+  BodyFat = "BodyFat (composite)"
+)
+modelsummary(list(SignalCost = mBFc, Tantrum = mBFtantrum, SignalCost = msubC, Tantrum = msubTantrum, SignalCost = mtriC, Tantrum = mtriTantrum), gof_map = "nobs", coef_rename = r_vars, stars = TRUE, coef_omit = "(Intercept)")
+```
+
+:::{#fig-cost_bodyfat}
+```{r}
+plot_predictions(mBFc, condition = "BodyFat", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Summed signal cost") + xlab("Body fat composite") + xlim(0,10) + ylim(0,200)
+```
+
+combine with other age plots and use caption 1
+:::
+
+:::{#fig-tantrum_bodyfat}
+```{r}
+plot_predictions(mBFtantrum, condition = "BodyFat", vcov = TRUE, points = .25) + theme_minimal(15) + ylab("Frequency of tantrums (per month)") + xlab("Body fat composite") + xlim(0,10)
+```
+
+combine with other body fat plots and use caption 1
+:::
+
+## Cargivers percieve older children as relatively needy
+
+## Parental responses to last signal of need
+
+## Sociodemographics
 

### 2024-08-23 analysisforpaper.R

@@ -1,61 +1,8 @@
-#' ---
-#' title: Child signaling
-#' ---
-#' # Child signaling
-#' **Data**: interviews with caretakers (usually mothers) in 157 families, with data
-#' on 375 children (more complete data on ~242)
-#'
-#' **Principle outcome variable**: child signaling frequency (Likert scale converted to times per month)
-#'
-#' **Secondary outcome variable**: child signaling cost (sad = low, crying = moderate, tantrums = high)
-#' (SadFreq + CryFreq\*2 + TantrumFreq\*3)
-#'
-#' **Tertiary outcome variable**: parental response to the most recent instance of child signaling
-#'
-#' **Predictor variables**:
-#'
-#' * ChildAge
-#' * Sex
-#' * Other children in the household (OtherChildrenHH)
-#' * LogIncome (median value of range for each option on the Likert Scale)
-#' * Number of adults, including the caretaker taking the survey (number_adults)
-#' * Partner status of caretaker (PartnerStatus: Partnered/Unpartnered)
-#' * Frequency of conflict between child and caretaker (Likert scale converted to times per month)
-#' * Frequency of the child's alloparenting contributions as a Likert scale converted to times per month (AlloparentingFreqN)
-#' * Caretaker's education level in years (EducationLevelYears)
-#' * Mean of caretaker's responses to three child health measures (CurrentHealthMean):
-#'
-#'   1) If an illness is 'going around', my child will get it.
-#'
-#'   2) My child’s immune system protects him/her from most illnesses that other kids get.
-#'
-#'   3) In general, my child is very susceptible to colds, flu, and other infectious diseases.
-#' * Random effects grouping variable (householdID)
-#'
-#' **Results**:
-#'
-#' * Negative association between age and child signaling (all signaling measures)
-#' * Positive association between conflict with parent and child signaling (all signaling measures)
-#' * Sex and focal child alloparenting effort interact (postive for females, negative/flat for males)
[…]

### 2024-08-26 analysisforpaper.R
Commented out a ggalluvial plot

@@ -502,6 +502,14 @@ ggplot(sfdfsum2, aes(axis1 = Sadness, axis2 = Crying, axis3 = Tantrums, y = Freq
   theme_minimal(20) +
   theme(axis.title.y = element_text(angle = 0, hjust = 1))
 
+# ggplot(sfdfsum2, aes(axis1 = Tantrums, axis2 = Crying, axis3 = Sadness, y = Freq)) +
+#   geom_alluvium(aes(fill = Tantrums), color = "black", show.legend = F) +
+#   geom_stratum() +
+#   geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
+#   scale_x_discrete(limits = c("Tantrums", "Crying", "Sadness"), expand = c(.2, .05)) +
+#   ylab("Number of\nchildren") +
+#   theme_minimal(20) +
+#   theme(axis.title.y = element_text(angle = 0, hjust = 1))
 
 signaldf_long <-
   sfdf |>

### 2024-08-26 Paper.qmd
Added text to results.

@@
 
+Child sadness, crying, and temper tantrums were reasonably common, with caretakers reporting over xx% of children engaging in each at least xxxTimexxx.
+A small subset of children engaged in these signals/states at least daily (stats).
+However, many children were reported to never engage in these signals/states (stat) or engage in them once a month or less (cite).
+
+Conflict between children and caretakers was also common, with `r sum(na.omit(d2$ConflictFreqN >= 8))` children (`r signif(sum((na.omit(d2$ConflictFreqN >= 8)) / sum(!is.na(d2$ConflictFreqN)) * 100), 2)`%) having conflicts with parents more than once a week.
+As with the frequencies of various types of signaling, many children (`r sum(na.omit(d2$ConflictFreqN == 0))` [`r signif(sum((na.omit(d2$ConflictFreqN == 0)) / sum(!is.na(d2$ConflictFreqN)) * 100), 2)`%]) were described as never having conflicts with parents (stat).
+
+# note these figures do not match plot currently (do we want to only include those with other data or all kids mentioned in households?)
+
+The frequency of alloparenting effort was bimodally distributed, with `r sum(na.omit(d2$AlloparentingFreqN == 0))` children (`r signif(sum((na.omit(d2$AlloparentingFreqN == 0)) / sum(!is.na(d2$AlloparentingFreqN)) * 100), 2)`%) being reported to never engage in it (`r sum(d2$NumberOfChildren==1)` instances were due to a child being an only child) and `r sum(na.omit(d2$AlloparentingFreqN >= 30))` children (`r signif(sum((na.omit(d2$AlloparentingFreqN >= 30)) / sum(!is.na(d2$AlloparentingFreqN)) * 100), 2)`%) being reported to help with childcare at least once a day.
+
 :::{#fig-barplot_SignalFreq}
@@
 caption
+:::
+
+:::{#fig-signal_alluvial_plot}
+```{r}
+signal_alluvial_plot
+```
+
+Place next to stacked barplot?
 :::
@@
 ## Signaling behavior is negatively associated with age
+
+For every signaling measure except sadness (which also involves non-signaling functions), there was a statistically significant negative relationship between the age of the child and the frequency of signaling when controlling for the sex of the child, household demographics, the frequency of conflict with one's primary caretaker, and an alloparenting effort by sex interaction.
+See @tbl-main_models for regression statistics.
 
@@
 
-The effect of the age on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior.
+The effect of age on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior.
 Bars are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
@@
 ## Signaling behavior is positively associated with conflict with primary caretakers
+
+There was a statistically significant positive relationship between the frequency of conflict between a child and their primary caretaker and the frequency of signaling when controlling for the sex of the child, the age of the child, household demographics, and an alloparenting effort by sex interaction.
+See @tbl-main_models for regression statistics.
 
@@
 
+For each signal, there was also a significant interaction between how often a child cared for their siblings and the sex of the child, whereby the slope of the relationship was positive or flat for girls and flat or negative for boys.
+In other words, our models predict that as children provide more allocare within their families, the frequency of signaling would remain at the same rate or, more often, increase with alloparenting effort for girls and remain at the same rate or decrease with alloparenting effort for boys.
+
 ## Summed signal cost and tantrum frequency are negatively associated with body fat
+
+(Alternative way, show just composite measure and all signaling outcomes)
+There were significant negative relationships between summed signal cost and tantrum frequency and 1) the residuals for subscapular skinfold thickness (stat), 2) triscapular skinflold thickness (stat), 3) and a scaled composite of the two measures (stat) when controlling for the sex of the child, the age of the child, the frequency of conflict with one's primary caretaker, household demographics, and an alloparenting effort by sex interaction.
+See @tbl-bodyfat_models for regression statistics.
+
+Height Z-scores, weight Z-scores, BMI Z-scores, and body fat percentage residuals were not significant predictors of signaling behavior in models with the same control variables see SI plot for regression table.
 
@@
 
+There were significant positive relationships between perceived relative neediness within the family and both the summed frequency of child signaling and child age when controlling for child sex, household demographics, the frequency of conflict between child and caretaker, and the frequency of a child providing allocare.
+There were also positive relationships between summed signal cost and perceived relative neediness within the family when controlling for the same variables; howerver, these associations were not significant.
+
 ## Parental responses to last signal of need
+
+There were no significant relationships between, child age, child sex, the number of children in the household, and log family income and the valence (positive, neutral, or negative) of parental reponses from the child's perspective.
+However, primary caretakers were significantly more likely to respond respond positively towards a child when they attributed the cause of signaling to discomfort, pain, injury, or illness (stat), and significantly less likely to respond positively when they attrivuted the cause of signaling to child punishment (stat).
+See fig xxxxx and SI table xxx.
 

### 2024-08-26 analysisforpaper.R

@@ -493,7 +493,7 @@ sfdfsum2 <-
   dplyr::filter(Freq > 2)
 
 
-ggplot(sfdfsum2, aes(axis1 = Sadness, axis2 = Crying, axis3 = Tantrums, y = Freq)) +
+signal_alluvial_plot <- ggplot(sfdfsum2, aes(axis1 = Sadness, axis2 = Crying, axis3 = Tantrums, y = Freq)) +
   geom_alluvium(aes(fill = Sadness), color = "black", show.legend = F) +
   geom_stratum() +
   geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
@@ -501,7 +501,7 @@ ggplot(sfdfsum2, aes(axis1 = Sadness, axis2 = Crying, axis3 = Tantrums, y = Freq
   ylab("Number of\nchildren") +
   theme_minimal(20) +
   theme(axis.title.y = element_text(angle = 0, hjust = 1))
-
+signal_alluvial_plot
 
 signaldf_long <-
   sfdf |>
@@ -511,8 +511,8 @@ signaldf_long <-
   ) |>
   na.omit()
 
-barplot_SignalFreq <- ggplot(signaldf_long, aes(y=Freq, fill=Signal)) + geom_bar(position = "dodge")
-barplot_SignalFreq
+# barplot_SignalFreq <- ggplot(signaldf_long, aes(y=Freq, fill=Signal)) + geom_bar(position = "dodge")
+# barplot_SignalFreq
 
 ggplot(signaldf_long, aes(Signal, Freq, group = uniqueID)) + geom_line(show.legend = FALSE, alpha = .25 , position = position_jitter(h = .1, w = .1))
 
@@ -520,7 +520,8 @@ signaldf_long_sum <- signaldf_long |>
   group_by(Signal, Freq) |>
   summarise(N = n())
 
-ggplot(signaldf_long_sum, aes(N, Signal, fill = Freq)) + geom_col(position = "stack")
+barplot_SignalFreq <- ggplot(signaldf_long_sum, aes(N, Signal, fill = Freq)) + geom_col(position = "stack")
+barplot_SignalFreq
 
 ggplot(signaldf_long_sum, aes(Freq, N, color = Signal, group = Signal)) + geom_line()
 
@@ -537,14 +538,19 @@ ggplot(x, aes(axis1 = Var1, axis2 = Var2, y = Freq)) +
 d2_conflict_filter <- d2 |>
[…]

### 2024-08-29 Paper.qmd
Added to and edited the methods, results, and discussion sections of the paper.
Cleaned up the analysis script.

--- 2024-08-20 Paper.qmd
@@
 
-We argue that older children (i.e., those older than infants) have evolved strategies to tailor their signals in response to 1) their valuation of a given resource, 2) the extent of conflicts of interest with caretakers, and 3) the presence of existing information about a child's need or potential to benefit from a resource so that they can credibly convey they are worth additional investment.
+We argue that older children (i.e., those older than infants) have evolved strategies to tailor their signals to 1) their valuation of a given resource, 2) the extent of conflicts of interest with caretakers, and 3) the presence of existing information about a child's need or potential to benefit from a resource so that they can credibly convey they are worth additional investment.
 Such strategies can lead to some combination of verbal requests for more support, facial sadness, crying, withdrawal, temper tantrums and self-harm, all of which may be effective in certain circumstances and are sometimes successful in eliciting more support from caretakers, despite a child's low status and formidability.
@@
 Each child, however, shares more genes with themselves (100%) than with their full siblings (50% on average), half-siblings (25% on average), and stepsiblings (0%).
-Due to this asymmetry, children are often expected to attempt to elicit more investment from parents than is optimal from the perspective of the parent, while parents are expected to attempt to invest less than is optimal for each child in order to invest in other siblings, whom parents are expected to value identically, all else being equal [@trivers1974].
+Due to this asymmetry, children are often expected to attempt to elicit more investment from parents than is optimal from the perspective of the parent, while parents are expected to attempt to invest less than is optimal for each child in order to invest in their other children, whom parents are expected to value identically, all else being equal [@trivers1974].
 
@@
 
-## Crying
+### Crying
 
@@
 
-## Withdrawal
+### Withdrawal
 
@@
 
-## Temper tantrums and self-harm
+### Temper tantrums and self-harm
 
@@
 
-## *Study population*
+## Study population
 
 This study was done in Utila, Honduras, one of the country’s three major Bay Islands in the Western Caribbean.
-Participants in the study largely fit into two groups who face different levels of economic and social adversity, 1) a long-standing English-speaking community made up of expats and their descendants, and 2) Spanish-speaking immigrants from the mainland.
+Participants in the study largely fit into two groups who face different levels of economic and social adversity, 1) a long-standing English-speaking community primarily made up of the descendants of expats and 2) a relatively new community of Spanish-speaking immigrants from the mainland.
 Compared to those born on Utila, immigrants from the mainland [@davis2014] are more likely to live in crowded and unsanitary conditions, face active discrimination [@garcia2017], and experience economic uncertainty due the predominance of low paying, seasonal jobs primarily related to the island’s diving tourism industry [@garcia2018; @stonich2000].
@@
 During focus groups held in July 2022, parents indicated high levels of sibling competition and various ways children might signal need.
-Both Spanish-speaking immigrants and English-speaking island natives reported that fighting between siblings was extremely common, with fights over toys, gifts, and phone time being the most reported causes.
+Both Spanish-speaking and English-speaking moms reported that fighting between siblings was extremely common, with fights over toys, gifts, and phone time being the most reported causes.
 In both populations, crying and various forms of withdrawal were listed as the most common responses to a child not receiving investment or support, followed by temper tantrums and making messes in their home.
@@
 
-## *Sampling*
+## Sampling
 
@@
 JH recruited early participants through convenience and snowball sampling.
-Participants were also recruited via Facebook posts, and fliers were sent home with children the Utila's Methodist school.
+Participants were also recruited via Facebook posts and fliers which were sent home with children the Utila's Methodist school.
 
@@
 
-## *Participants*
+## Participants
 
 All participants for the survey portion of the study met the inclusion criteria.
-Most participants were mothers of at least one biological child in the study, with the participant being the biological mother of the child in 357 of 375 caregiver/child dyads (our imputation).
+Most participants were mothers of at least one biological child in the study, with the participant being the biological mother of the child in xxx out of xxxx caregiver/child dyads.
 
@@
 
-## *Ethics*
+## Ethics
 
-All adult participants provided informed consent. Surveys were designed to last about hour plus 20 minutes for each additional child. Adults were paid \$XXXXX for the survey portion of the study.
+All adult participants provided informed consent. Surveys were designed to last about an hour plus 20 minutes for each additional child. Adults were paid \$XXXXX for the survey portion of the study.
 
@@
 
-## *Survey*
+## Survey
 
@@
 
-### *Frequency of child signaling, alloparenting contributions, and conflict with caretaker*
+### Frequency of child signaling, alloparenting contributions, and conflict with caretaker
 
-We used 6-point Likert scales to get caregiver reports of 1) child signaling, 2) interpersonal conflict with each participating child, 3) and child alloparenting behavior.
-See Table 1 for exact wording.
+We used 6-point Likert scales to get caregiver reports on the frequencies of 1) various types of child signaling, 2) interpersonal conflict with each participating child, 3) and child alloparenting behavior.
+See @tbl-freqvars for exact wording.
+Caretakers were often not asked about child alloparenting behavior if a child was an only child.
+For this reason, we gave children an alloparenting frequency of never if they were the only child within a household.
 
 |                                 |                                                                                                                   |
-|-------------------------|-----------------------------------------------|
+|-------------------------|----------------------------------------------|
 | Variable                        | Survey question wording                                                                                           |
@@
 
-**Table 1**.
-Table showing variable names and question wording for measures about the frequency of behaviors.
+: Table showing variable names and question wording for measures about the frequency of behaviors.
 All questions had the following labels attached to each point on its Likert scale: Never; Once a month or less; More than once a month but less than once a week; More than once a week but not daily; Daily; Multiple times per day.
+{#tbl-freqvars}
 
-As these responses reflect frequencies of actual behavior, we converted these to counts which reflected the number of times per month a child engaged in these behaviors.
-The original wording of the Likert scale and the frequency conversion are shown in Table 2.
+As these responses reflect frequencies of behavior, we converted these to counts which reflected the number of times per month a child engaged in these behaviors for analyses.
+The original wording of the Likert scale and the frequency conversion are shown in @tbl-convertedfreqs.
 
 |                                                  |                               |
-|------------------------------------------------|-----------------------------------------------|
+|------------------------------------|------------------------------------|
 | Likert scale wording                             | Count value (times per month) |
@@
 
-**Table 2.** Table showing conversions from Likert scale to count value for the frequencies of child behaviors included in this study.
+: Table showing conversions from Likert scale to count value for the frequencies of child behaviors included in this study.
 Variables created with this scheme include SadFreqN, CryFreqN, TantrumFreqN, RunawayFreqN, ConflictFreqN, and AlloparentingFreqN.
+{#tbl-convertedfreqs}
 
-From these count variables, we created three additional variables to be used, primarily, as our main dependent variables in analyses.
-SignalFreq was the sum of the frequencies of child sadness, crying, and temper tantrums.
-SignalFreqMax was the frequency of the most common of these three signals.
-SignalCost was the sum of the frequencies of these three signals weighted by their relative costliness (i.e., SignalCost = SadFreqN + CryFreqN\*2 + TantrumFreqN\*3).
-The frequency of running away was not included in the creation of any of these variables due to uncertainty over whether it was likely to be a credible signal due to the relative safety of Utila compared to other areas in which running away has been found to be a successful strategy for children.
+From these count variables, we created three additional signaling variables to be used, primarily, as our main dependent variables in analyses.
+The frequency of signaling was the sum of the frequencies of child sadness, crying, and temper tantrums.
+Maximum signal frequency was the frequency of the most common of these three signals.
+Signal cost was the sum of the frequencies of these three signals weighted by their relative costliness (i.e., Signal cost = frequency of sadness + frequency of crying * 2 + frequency of temper tantrums * 3).
+Although we collected data on the frequency of running away, it was not included in the creation of any of these variables due to uncertainty over whether it was likely to be a credible signal due to the relative safety of Utila compared to other areas in which running away has been found to be a successful strategy for children.
 
-### *Relative neediness of children and relative investment in children within the family*
+### Relative neediness of children and relative investment in children within the family
 
 We also used 6-point Likert scales to get parental reports of 1) the relative neediness of each participating child and 2) the relative investment they provide to each participating child compared to the rest of their children.
-See Table 3 for exact wording.
+See Table @tbl-relvars for exact wording.
 
 |                     |                                                                                                  |
-|-------------------------|-----------------------------------------------|
+|-------------------------|----------------------------------------------|
 | Variable            | Survey question wording                                                                          |
@@
 
-**Table 3.** Table showing variable names and question wording for measures about relative need and investment.
+: Table showing variable names and question wording for measures about relative need and investment.
 Both questions had the following labels attached to each point on its Likert scale: Much less; Less; The same amount; More; Much more.
+{#tbl-relvars}
 
-### *Last instance of child signaling*
+### Causes and responses to the last instance of child signaling
 
 We also asked caretakers about the last time each participating child was sad or cried.
-This involved providing parents with a free response text box to describe the cause of the emotion that lead to the display.
+This involved asking what the caretaker thought caused the emotion or display and recording their answer in a text box.
 We then asked how the parent responded to their child.
-The researcher giving the survey would then bin responses into one of 31 categories which included positive, negative, and neutral responses.
+The researcher giving the survey would then bin responses into at least one of 31 categories which included positive, negative, and neutral responses (from the child's perspective).
 For the full list of options see Table S1.
 
-### *Child anthropometrics*
+From this data, we created a parental response variable, which was coded as 1 for a positve response, 0 for a neutral response, and -1 for a negative response (all from the child's perspective).
+MG also coded each cause for the presence or absence of 17 various elements.
+Of these 4 were use in analyses: whether or not the cause involved 1) the child being asked to do an unwanted task, 2) punishment, 3) conflict within the family, and 4) discomfort, pain, injury, or illness.
+
+### Child anthropometrics
 
@@
 
-We computed each child's height, weight, and BMI Zscores with respect to the WHO child growth standards (cite).
+We computed each child's height, weight, and BMI Zscores with respect to the CDC child growth standards (cite).
 
-### *Perceptions of child immune function (or current susceptibility to illness)*
+### Perceptions of child immune function (or current susceptibility to illness)
 
-Caretakers were also asked three questions using a five point Likert scale about their child's current health which involved the likelihood of them becoming sick.
-We then inverted the scores for two measures so that a score of 5 corresponded with the lowest susceptibility to illness before averaging the scores of these to create the CurrentHealthMean variable.
-Means were still calculated if 1-2 values had an NA (only xxx cases), with the numerator being the number of variables with data.
-See Table 4 for the wording of these variables.
+Caretakers were also asked three questions using a five point Likert scale about their child's susceptibility to illness.
+We inverted the scores on two measures so that a score of 5 corresponded with the lowest susceptibility to illness before we averaged the scores of these to create the CurrentHealthMean variable.
+In the `r sum((!is.na(modeldf$Currenthealth2)) & is.na(modeldf$Currenthealth3))` instance in which we only data for 2 of the measures, we still calculated the mean a numerator of 2.
+See Table @tbl-healthvars for the wording of these variables.
 
 |                |                                                                                       |
-|-------------------------|-----------------------------------------------|
+|-------------------------|----------------------------------------------|
 | Variable       | Survey question wording                                                               |
@@
 
-: **Table 4.** Table showing variable names and question wording for measures about a child's current likelihood of becoming sick.
+: Table showing variable names and question wording for measures about a child's current likelihood of becoming sick.
 All three questions were asked using a Likert scale from 1-5 where 1 was labeled "Strongly disagree" and 5 was labeled "Strongly agree." The scores for Currenthealth1 and Currenthealth2 were inverted so that 5 corresponded with the lowest susceptibility to illness.
+{#tbl-healthvars}
 
-### *Demographic questions*
+### Demographic questions
 
@@
 
-CHANGE TO TOTAL CHILDCARE
-
-DECIDE on alloparentingFreqN2
+OLD AND CURRENTLY NOT USING
 
@@
 
-## *Statistical analyses*
+## Statistical analyses
 
-We fit generalized linear mixed models for exploratory analyses with signal frequency and summed signal cost as dependent variables.
+We fit generalized linear mixed models for exploratory analyses with signal frequencies and summed signal cost as dependent variables.
 As our signaling variables invovled over-dispersed count data, we used negative binomial regression.
@@
 
-## *Data availability*
+## Data availability
 
@@
 
-kable(custom.summarize(d2, vars_c)) %>% kable_styling()
+kable(custom.summarize(modeldf, vars_c)) %>% kable_styling()
 
@@
 
+# CONFIRM: should this be household level?
 ```{r sumtab_adults}
-#| echo: false
-#| label: sumtab_cadults
-#| fig-cap: "Summary statistics for adult and household study variables."
+#| label: tbl-sumtab_adults
+#| tbl-cap: Summary statistics for adult and household study variables.
 
@@
 ```
+
+## Distributions of signaling behavior, alloparenting effort, and conflict
+
+Child sadness, crying, and temper tantrums were reasonably common.
+Caregivers reported that (`r sum(na.omit(modeldf$SadFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$SadFreqN >= 8)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\]) of children were sad at least once a week, (`r sum(na.omit(modeldf$CryFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$CryFreqN >= 8)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\]) cried at least more than once a week, and (`r sum(na.omit(modeldf$TantrumFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$TantrumFreqN >= 8)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]) threw temper tantrums at least once a week.
+A small subset of children engaged in these signals/states at least daily (sadness = `r sum(na.omit(modeldf$SadFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$SadFreqN >= 30)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\], crying = `r sum(na.omit(modeldf$CryFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$CryFreqN >= 30)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\], and tantrums = `r sum(na.omit(modeldf$TantrumFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$TantrumFreqN >= 30)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]).
+However, many children were reported to never engage in these signals/states (sadness = `r sum(na.omit(modeldf$SadFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$SadFreqN == 00)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\], crying = `r sum(na.omit(modeldf$CryFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$CryFreqN == 0)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\], and tantrums = `r sum(na.omit(modeldf$TantrumFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$TantrumFreqN == 0)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\])
+
+Conflict between children and caretakers was also common, with `r sum(na.omit(modeldf$ConflictFreqN >= 8))` children (`r signif(sum((na.omit(modeldf$ConflictFreqN >= 8)) / sum(!is.na(modeldf$ConflictFreqN)) * 100), 2)`%) having conflicts with parents more than once a week.
+As with the frequencies of various types of signaling, many children (`r sum(na.omit(modeldf$ConflictFreqN == 0))` \[`r signif(sum((na.omit(modeldf$ConflictFreqN == 0)) / sum(!is.na(modeldf$ConflictFreqN)) * 100), 2)`%\]) were described as never having conflicts with parents (stat).
+
+The frequency of alloparenting effort was bimodally distributed, with `r sum(na.omit(modeldf$AlloparentingFreqN == 0))` children (`r signif(sum((na.omit(modeldf$AlloparentingFreqN == 0)) / sum(!is.na(modeldf$AlloparentingFreqN)) * 100), 2)`%) being reported to never engage in it (`r sum(na.omit(modeldf$NumberOfChildren==1))` instances were due to a child being an only child) and `r sum(na.omit(modeldf$AlloparentingFreqN >= 30))` children (`r signif(sum((na.omit(modeldf$AlloparentingFreqN >= 30)) / sum(!is.na(modeldf$AlloparentingFreqN)) * 100), 2)`%) being reported to help with childcare at least once a day.
+
+::: {#fig-barplot_SignalFreq}
+```{r}
+barplot_SignalFreq
+```
+
+Stacked bar plot showing the frequencies of child sadness, crying, and temper tantrums per month as reported by primary caretaker.
+Frequency key: 0 corresponds to "never," <=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and >30 corresponds to "multiple times per day."
+:::
+
+::: {#fig-signal_alluvial_plot}
+```{r}
+signal_alluvial_plot
+```
+
+Alluvial plot showing the combinations of signaling frequencies.
+Alluvium width equals the number of individiuals with the same frequencies for the three signals/states.
+Stratum height equals the count of individuals at a given frequency.
+Frequency key: Low includes "never" and "once a month or less," mediums includes "more than once a month but less than once a week" and "more than once a week but not daily," high includes "daily," and  "multiple times per day."
+:::
+
+::: {#fig-barplot_ConflictFreq}
+```{r}
+barplot_conflict
+```
+
+Stacked bar plot showing the frequency of conflict between child and primary caregiver as reported by primary caretaker.
+Color corresponds to sex of the child.
+Frequency key: 0 corresponds to "never," <=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and >30 corresponds to "multiple times per day."
+:::
+
+::: {#fig-barplot_AlloparentingFreq}
+```{r}
+barplot_alloparenting
+```
+
+Stacked bar plot showing the frequency of conflict a child's alloparenting effort as reported by primary caretaker.
+Color corresponds to sex of the child.
+Frequency key: 0 corresponds to "never," <=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and >30 corresponds to "multiple times per day."
+:::
+
+## Signaling behavior is negatively associated with age
+
+For every signaling measure except sadness (which also involves non-signaling functions), there was a statistically significant negative relationship between the age of the child and the frequency of signaling when controlling for the sex of the child, household demographics, the frequency of conflict with one's primary caretaker, caretaker perceptions of child immune function, and an alloparenting effort by sex interaction.
+See @tbl-main_models for regression statistics and @fig-effect_age for effects plots.
+
+Age had largest effect on crying (`r stats$mcryH$ChildAge$str`), with the model predicting crying frequency dropping to half of the age 5 value by around age 8 (visual inspection).
+The effect of age on temper tantrums (`r stats$mtantrumH$ChildAge$str`) was smaller with a more linear slope.
+
+```{r}
+#| label: tbl-main_models
+#| tbl-cap: Regression tables for our main exploratory models.
+r_vars <- c(
+  ChildAge = "Age (years)",
+  SexMale = "Sex (male)",
+  OtherChildrenHH = "Other children",
+  LogIncome = "Income (log)",
+  number_adults = "Adults in home",
+  PartnerStatusUnpartnered = "Partner status (single)",
+  EducationLevelYears = "Parent ed. (years)",
+  ConflictFreqN = "Conflict freq.",
+  SignalFreq = "Signal freq.",
+  SignalCost = "Signal cost",
+  SignalFreqMax = "Max signal freq.",
+  CurrentHealthMean = "Percieved immunity",
+  SadFreqN = "Sadness freq.",
+  CryFreqN = "Crying freq.",
+  TantrumFreqN = "Tantrum freq.",
+  AlloparentingFreqN = "Child allocare"
+)
+modelsummary(list(Sadness = msadH, Crying = mcryH, Tantrum = mtantrumH, SignalFreq = msfH, SignalCost = mscH, SignalFreqMax = mmsfH), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+```
+
+::: {#fig-effect_age}
+```{r}
+signaling_data_plot_age
+```
+
+The effect of age on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior.
+Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
+Each dot is 1 child.
+Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate).
+See @tbl-main_models for the corresponding regression table.
+:::
+
+## Signaling behavior is positively associated with conflict with primary caretakers
+
+There were statistically significant positive relationships between the frequency of conflict between a child and their primary caretaker and all of our signaling measures when controlling for the sex of the child, the age of the child, household demographics, caretaker perceptions of child immune function, and an alloparenting effort by sex interaction.
+See @tbl-main_models for regression statistics @fig-effect_conflict for effects plots.
+
+# FIX: y axis extends farther for tantrum model (reducing this manually breaks error shading)
+
+::: {#fig-effect_conflict}
+```{r}
+signaling_data_plot_conflict
+```
+
+The effect of the frequency of conflict with one's primary caretaker on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior.
+Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
+Each dot is 1 child.
+Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate).
+See @tbl-main_models for the corresponding regression table.
+:::
+
+## Interaction between child sex and child alloparenting
+
+::: {#fig-effect_allo_by_sex}
+```{r}
+signaling_data_plot_allo_by_sex
+```
+
+The effect of an alloparenting effort by sex interaction on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior.
+Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
+Each dot is 1 child.
+Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate).
+See @tbl-main_models for the corresponding regression table.
+:::
+
+For each signaling measure, there was also a significant interaction between how often a child cared for their siblings and the sex of the child when controlling for the age of the child, the sex of the child, household demographics, the frequency of conflict with one's primary caretaker, and caretaker perceptions of child immune function.
+The slope of these relationships were curvilinear and positive for girls and flat or slightly negative for boys.
+In other words, our models predict that as children provide more allocare within their families, the frequency of signaling would increase with alloparenting effort for girls and remain at the same rate or decrease slightly with alloparenting effort for boys.
+See @tbl-main_models for the corresponding regression table and @fig-effect_allo_by_sex for effects plots.
+
+## Summed signal cost, crying frequency, and tantrum frequency are negatively associated with body fat
+
+There were significant negative relationships between summed signal cost (`r stats$mBFc$BodyFat$str`) crying frequency (`r stats$mBFcry$BodyFat$str`), and tantrum frequency (`r stats$mBFtantrum$BodyFat$str`) and a scaled composite of subscapular and tricep skinfold thicknesses when controlling for the sex of the child, the age of the child, the frequency of conflict with one's primary caretaker, household demographics,caretaker perceptions of child immune function, and an alloparenting effort by sex interaction.
+See @tbl-bodyfat_models for regression statistics and @fig-effect_bodyfat for effects plots.
+Subscapular and tricep skinfold thicknesses were also significant predictors of summed signal cost and tantrum frequency when included in the same models as the sole measure of body fat.
+See xxxxSI TABLESxxxx for regression tables.
+
+Height Z-scores, weight Z-scores, BMI Z-scores, and body fat percentage residuals were not significant predictors of signaling behavior in models with the same control variables.
+See xxxxSI TABLESxxxx for regression tables.
+
+```{r}
+#| label: tbl-bodyfat_models
+#| tbl-cap: Regression tables for the models examining the effect of body fat on signaling.
+r_vars <- c(
+  ChildAge = "Age (years)",
+  SexMale = "Sex (male)",
+  OtherChildrenHH = "Other children",
+  LogIncome = "Income (log)",
+  number_adults = "Adults in home",
+  PartnerStatusUnpartnered = "Partner status (single)",
+  EducationLevelYears = "Parent ed. (years)",
+  ConflictFreqN = "Conflict freq.",
+  SignalFreq = "Signal freq.",
+  SignalCost = "Signal cost",
+  SignalFreqMax = "Max signal freq.",
+  CurrentHealthMean = "Percieved immunity",
+  SadFreqN = "Sadness freq.",
+  CryFreqN = "Crying freq.",
+  TantrumFreqN = "Tantrum freq.",
+  AlloparentingFreqN = "Child allocare",
+  TricepR = "Tricep skinfold (residuals)",
+  SubscapR = "Subscapular skinfold (residuals)",
+  BodyFat = "BodyFat (composite)"
+)
+modelsummary(list(Sadness = mBFsad, Crying = mBFcry, Tantrum = mBFtantrum, SignalFreq = mBFf, SignalCost = mBFc, SignalFreqMax = mBFfm), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+```
+
+::: {#fig-effect_bodyfat}
+```{r}
+signaling_data_plot_bodyfat
+```
+
+The effect of body fat (scaled composite measure of subscapular and tricep skinfold thicknesses) on A) the summed cost of signaling behavior, B) the frequency of crying, and C) the frequency of temper tantrums.
+Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
+Each dot is 1 child.
+Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate).
+See @tbl-bodyfat_models for the corresponding regression table.
+:::
+
+## Positive associations between child age and signaling frequency and child neediness
+
+Due to the rarity of primary caretakers describing a child as "much less" (`r sum(na.omit(modeldf$RelativeNeed == "Much less"))` \[`r signif((sum(na.omit(modeldf$RelativeNeed == "Much less"))/ sum(!is.na(modeldf$RelativeNeed)) * 100), 2)`%\]) or "much more" (`r sum(na.omit(modeldf$RelativeNeed == "Much more"))` \[`r signif((sum(na.omit(modeldf$RelativeNeed == "Much more"))/ sum(!is.na(modeldf$RelativeNeed)) * 100), 2)`%\]) needy compared to their other children, we collapsed "less" and "much less" into one "less" factor and "more" and "much more" into one "more" factor.
+The end result was `r rel_need_table[1]` less needy children, `r rel_need_table[2]` children who were as needy as their siblings, and `r rel_need_table[3]` more needy children.
+
+There were significant positive relationships between perceived relative neediness within the family and both the summed frequency of child signaling and child age when controlling for child sex, household demographics, the frequency of conflict between child and caretaker, and the frequency of a child providing allocare.
+There was also a positive relationship between summed signal cost and perceived relative neediness within the family when controlling for the same variables; however, this association was not significant.
+See @tbl-need_models for the corresponding regression table.
+
+#FIX: how to use coeftest with this
+```{r}
+#| label: tbl-need_models
+#| tbl-cap: Regression tables for the models examining relative perceptions of child need within the family as a dependent variable.
+r_vars <- c(
+  ChildAge = "Age (years)",
+  SexMale = "Sex (male)",
+  OtherChildrenHH = "Other children",
+  LogIncome = "Income (log)",
+  number_adults = "Adults in home",
+  PartnerStatusUnpartnered = "Partner status (single)",
+  EducationLevelYears = "Parent ed. (years)",
+  ConflictFreqN = "Conflict freq.",
+  SignalFreq = "Signal freq.",
+  SignalCost = "Signal cost",
+  SignalFreqMax = "Max signal freq.",
+  CurrentHealthMean = "Percieved immunity",
+  SadFreqN = "Sadness freq.",
+  CryFreqN = "Crying freq.",
+  TantrumFreqN = "Tantrum freq.",
+  AlloparentingFreqN = "Child allocare",
+  TricepR = "Tricep skinfold (residuals)",
+  SubscapR = "Subscapular skinfold (residuals)",
+  BodyFat = "BodyFat (composite)",
+  RelativeNeed3 = "Relative need (w/in household)"
+)
+modelsummary(list(SignalFreq = mN1, SignalCost = mN2), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+```
+
+::: {#fig-effect_relativeneed}
+```{r, fig.width= 10}
+signaling_plot_need
+```
+
+Plots showing the likelihood of child being more needy, as needy, or less needy than a caretaker's other children.
+A: Effect of child age on relative neediness B: Effect of summed signaling frequency on relative neediness.
+:::
+
+## Parental responses to last signal of need
+
+There were `r response_overlap[2,2]` instances in which a caregiver's response involved acts which were both negative and positive from the child's perspective.
+These were excluded from analyses due to their rarity.
+
+There were no significant relationships between, child age, child sex, the number of children in the household, and log family income and the valence of parental reponses (positive, neutral, or negative) from the child's perspective.
+However, primary caretakers were significantly more likely to respond positively towards a child when they attributed the cause of signaling to discomfort, pain, injury, or illness (stat), and significantly less likely to respond positively when they attributed the cause of signaling to child punishment (stat).
+See @fig-effect_last_signal_response and @tbl-last_signal_model.
+
+#FIX: table is not properly dealing with p values
+```{r}
+#| label: tbl-last_signal_model
+#| tbl-cap: Regression table for the model examining responses to the last time a child was sad or cried.
+r_vars <- c(
+  ChildAge = "Age (years)",
+  SexMale = "Sex (male)",
+  OtherChildrenHH = "Other children",
+  LogIncome = "Income (log)",
+  number_adults = "Adults in home",
+  PartnerStatusUnpartnered = "Partner status (single)",
+  EducationLevelYears = "Parent ed. (years)",
+  ConflictFreqN = "Conflict freq.",
+  SignalFreq = "Signal freq.",
+  SignalCost = "Signal cost",
+  SignalFreqMax = "Max signal freq.",
+  CurrentHealthMean = "Percieved immunity",
+  SadFreqN = "Sadness freq.",
+  CryFreqN = "Crying freq.",
+  TantrumFreqN = "Tantrum freq.",
+  AlloparentingFreqN = "Child allocare",
+  TricepR = "Tricep skinfold (residuals)",
+  SubscapR = "Subscapular skinfold (residuals)",
+  BodyFat = "BodyFat (composite)",
+  RelativeNeed3 = "Relative need (w/in household)",
+  UnwantedTask = "Unwanted task",
+  Punishment21 = "Punishment",
+  Family21 = "Family conflict",
+  DiscomfortPainInjuryIllness = "Discomfort/Pain/Injury/Illness",
+  CaregiverResponse = "Caregiver Response"
+)
+modelsummary(list(CaregiverResponse = mpr), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+```
+
+::: {#fig-effect_last_signal_response}
+```{r, fig.width= 8}
+signaling_plot_last_signal_response
+```
+
+Plots showing the likelihood of positive, neutral, and negative responses to the last time a child became sad or cried.
+A: Effect of punishment being implicated in the cause (0 = no punishment, 1 = punishment.) B: Effect of discomfort, pain, injury, or illness being implicated in the cause (0 = discomfort, pain, injury, or illness, 1 = discomfort, pain, injury, or illness.)
+:::
+
+## Sociodemographics
 
@@
 One possibility is that older children have greater ability to satisfy one's desires and behave in ways which match cultural expectations of older children and adults.
-Another is that as children survive to the point where they would have been unlikely to die young in ancestral environments, conflicts of interest with parents decrease due to children passing this important milestone and requiring less investment to reach adulthood than younger children, current or future (cite daly and wilson).
+Another is that as children survive to the point where they would have been unlikely to die young in ancestral environments, the intensity conflicts of interest with parents decrease due to children passing this important milestone and requiring less investment to reach adulthood than younger children, current or future (cite daly and wilson).
 For this reason, less costly signals may be more likely to work to increase investment, all else being equal\[\[4\]\].
-The decrease in crying and tantrum frequencies with age may also result from the potential for greater reputational costs, although this may also increase the credibility of such signals among older children as previously mentioned.
-It is also possible that parents may see older children less than younger children.
-However, the commonality of decreased signals of need with age, suggest that this is unlikely to have a large effect with this population.
+The decrease in crying and tantrum frequencies with age may also result from the potential for greater reputational costs, something which may also increase the credibility of such signals among older children as previously mentioned.
+It is also possible that parents see older children less than younger children.
+However, the negative assocition between need with age observed in other societies and the fact that all children in this study were of school age suggest that this is unlikely to be the sole reason responsible for this.
 
 There was no significant association between the frequency of child sadness and age.
-Part of this likely results from the fact that, unlike crying, sadness is thought to have important nonsignaling functions including drawing one’s attention to a problem, motivating deeper analysis of one’s situation, and causing withdrawal from situations with low fitness prospects.
+Part of this likely results from the fact that, unlike crying, sadness is thought to have important non-signaling functions including drawing one’s attention to a problem, motivating deeper analysis of one’s situation, and causing withdrawal from situations with low fitness prospects (cite), from which others can still obtain cue-based information about one's emotional state, well-being, and desires.
+By asking about sadness as a whole, our measure was unable to disentangle information caretakers gleaned from cues from information that came from signals.
+For this reason, a change in non-crying signals that arise from sadness (e.g., facial sadness and some, but all, forms of withdrawal) may have been harder to detect with our sample size even if young children expressed facial sadness more often.
+It is also possible that there is no effect of age on sadness frequency among the age ranges examined it.
+add in depression over the lifecourse
 
-The frequency of running away increased with age.
-This may be due to younger children being observed more closely or due to older children experiencing less danger from leaving the home alone (more on this below).
+Negative association with signal cost too
 
@@
 
-One of the most reliable findings across the models tested was an interaction between how much a given child contributed allocare within the household and child sex, whereby the slope of the relationship was postitive or flat for girls and flat or negative for boys.
+We found evidence of an interaction between how often a child cared for their siblings and the sex of the child in our models examining signal frequencies and cost, whereby there was a positive relationship between the frequency of and summed costliness of child signals of need and alloparenting effort for girls but no relationship or a slightly negative one for boys.
 In many societies, girls are often expected to do more alloparenting within the household and often end up doing so (cite).
@@
 
+*Body fat discussion*
+
 *Lack of other sex differences*
 
-Add
+*Positive associations between child age and signaling frequency and child neediness*
+
+There was a positive association between the summed frequency of child signaling and caretaker perceptions of child neediness within the family, something which would be consistent with sadness, crying, and temper tantrums being signals (or involving signals in the case of crying) if parents took this to mean how much they value additional investment.
+
+*Parental responses to last signal of need*
+
+Parents were more likely to respond positively and less likely to respond negatively to the last time a child signaled when the cause involved discomfort, pain, injury, or illness than when it did not.
+One potential causal explanation for this is that parents may often have information for sources outside of child signaling to corroborate a child's need (e.g., a fever, a diagnosis, or witness a child hurt themselves) so that there is less private information about a child's current state.
+Simultaneously, conflicts of interest may be less significant in these times due to either 1) both sides benefiting similarly from their child returning to normal (e.g., with an illness) or 2) the cause involving something which is minor and would quickly resolve itself (e.g.
+a skinned knee).
+It may also be that many of the typical positive responses to pain and illness are not costly (e.g., providing comfort).
+
+When the cause involved punishment, however, parents were more likely to respond negatively and less likely to respond positively.
+In these circumstances, children may be signaling about the severity of a punishment (cite).
+However, parents may be reluctant to reduce punishments if the behavior that warranted the punsishment was repeated or severe.
+They may also react with scolding or further punishment.
+That said, our model predicts positive responses in xxxxx cases, suggesting that children are successful in negotiating better outcomes at least some of the time.
+- costly apolgy model?
+
+Although not a significant relationship, the negative effect of the family conflict being a reported cause of signaling on the likelihood of a positive response may reflect tradeoffs within the family, whereby investment in one child cannot go to another.
+Caretakers may have also responded more negatively when the cause involved family conflict, if it involved a child imposing costs on other family members.
 
@@
 
-*Negative association between caretaker estimates of child health and child signaling for some measures. This does not extend to the costlier signals.*
+*CUT Negative association between caretaker estimates of child health and child signaling for some measures. This does not extend to the costlier signals.*
 
@@
 Pathogens have been a strong selection pressure for life on earth for much of its existence and have greatly shaped evolutionary history (cite).
-In response to this, humans have evovled many adapatations for limiting the harm we experience from pathogens, including the ability to identify relaible indicators of illness in others.
-In some contexts, this activates disgust and motivates avoidance of sick inidivudals.
+In response to this, humans have evolved many adaptations for limiting the harm we experience from pathogens, including the ability to identify relaible indicators of illness in others.
+In some contexts, this activates disgust and motivates avoidance of sick individuals.
 However, it can also lead to increased care of relatives and allies (cite), with there being some suggestion that we have even evolved to upregulate some of our symptoms beyond what would arise from an immune response/pathogen interaction to better convince others that we are sick (cite).
@@
 
-## **Limitations**
+## Limitations
 

### 2024-08-29 analysisforpaper.R

@@ -1,24 +1,18 @@
 # loading libraries -------------------------------------------------------
 
-#+ message=F, warning=F, fig.height=9, fig.width=10
-
-# note some are no longer used
+# note: list is cleaned up but some may no longer used
 library(utiladata2023) # data package
 library(tidyverse)
-# library(lme4)
 library(effects)
 library(visreg)
 library(emmeans)
 library(car)
 library(haven)
 library(MASS)
-library(glmmTMB) # try this negative binomial (random effect for neighborhood)
-#library(gee)
-#library(geepack)
+library(glmmTMB)
 library(sandwich)
 library(lmtest)
 library(ggcorrplot)
-#library(geeM)
 library(marginaleffects)
 library(ordinal) # clmm # polr
 library(brms)
@@ -28,9 +22,9 @@ library(knitr)
 library(kableExtra)
 library(modelsummary)
 library(ggalluvial)
+library(patchwork)
 
 source("recode.R")
-
 
 # functions used ----------------------------------------------------------
 
@@ -58,7 +52,6 @@ mean2 <- function(..., na.rm = F){
   if(all(is.na(v))) return(NA)
   mean(v, na.rm = na.rm)
 }
-
 
 # dataframe prep ----------------------------------------------------------
 
@@ -173,7 +166,7 @@ d <- children |>
       #.default = ConflictFreq
       ),
     AlloparentingFreq0 = case_when(
-      is.na(AlloparentingFreq) & ChildAge < 5 ~ "Never",
+      is.na(AlloparentingFreq) & ChildAge < 5 ~ "Never", #does nothing
       is.na(AlloparentingFreq) & NumberOfChildren == 1 ~ "Never",
       is.na(AlloparentingFreq) & Siblings == 0 ~ "Never", # check later
       is.na(AlloparentingFreq) & YoungerKids == 0 ~ "Never",
@@ -239,12 +232,12 @@ d <- children |>
     NeutralResponseF = as.factor(NeutralResponseBinary),
     NegativeResponseBinary = ifelse(NegativeResponse >= 1, 1, 0),
     NegativeResponseF = as.factor(NegativeResponseBinary),
-    ParentalResponse = case_when(
+    CaregiverResponse = case_when(
[…]

### 2024-08-31 Paper.qmd
Analysis script: moved some recodings to data package, updated and fixed plots and updated alloparenting (dependent variable) model.
Paper: Added to results and methods. Also updated and added some citations.

@@
 format: html
+lightbox: true
 editor: visual
@@
 Infant signals of need, particularly infant cries, have been extensively studied in a range of populations [@barr1991; @fouts2005; @konner1974; @kruger2010; @kushnick2006; @meehan2017; @müller2021; @tronick1987], yet the full range of potential child signals of need, which likely extend beyond cries, has never been investigated in children of all ages.
-In particular, if and when children signal need with costly signals [@grafen1990; @spence1973; @zahavi1975a] has rarely been investigated (for one exception, see [@syme2023]).
+In particular, if and when children signal need with costly signals [@grafen1990; @spence1973; @zahavi1975a] has rarely been investigated [for one exception, see, @syme2023a].
 
@@
 Such a dynamic has been found among the Chuukese, with both male and female children reporting conflict with their families over children’s desires to invest financial resources in their own status enhancing opportunities rather than in the household [@lowe2003].
-Similarly, Syme and Hagen [@syme2023] found greater numbers of siblings were associated with increased likelihoods of severe conflict in a retrospective study with Chuukese-Americans.
-This may also be consistent with parents pressuring older children to invest in the household; however, there are other ways in which limited resources can result in behavioral conflict [@syme2023].
+Similarly, Syme and Hagen [@syme2023a] found greater numbers of siblings were associated with increased likelihoods of severe conflict in a retrospective study with Chuukese-Americans.
+This may also be consistent with parents pressuring older children to invest in the household; however, there are other ways in which limited resources can result in behavioral conflict [@syme2023a].
 
@@
 
-Withdrawal has been suggested to be an underappreciated way those with low social power may credibly signal need [@syme2023].
+Withdrawal has been suggested to be an underappreciated way those with low social power may credibly signal need [@syme2023a].
 Due to their lower status, children have limited ways to change how others treat them.
@@
 |                                 |                                                                                                                   |
-|-------------------------|----------------------------------------------|
+|--------------------------|----------------------------------------------|
 | Variable                        | Survey question wording                                                                                           |
@@
 Maximum signal frequency was the frequency of the most common of these three signals.
-Signal cost was the sum of the frequencies of these three signals weighted by their relative costliness (i.e., Signal cost = frequency of sadness + frequency of crying * 2 + frequency of temper tantrums * 3).
+Signal cost was the sum of the frequencies of these three signals weighted by their relative costliness (i.e., Signal cost = frequency of sadness + frequency of crying \* 2 + frequency of temper tantrums \* 3).
 Although we collected data on the frequency of running away, it was not included in the creation of any of these variables due to uncertainty over whether it was likely to be a credible signal due to the relative safety of Utila compared to other areas in which running away has been found to be a successful strategy for children.
+
+We also calculated the summed frequency of child alloparenting contributions by other children within the household.
 
@@
 |                     |                                                                                                  |
-|-------------------------|----------------------------------------------|
+|--------------------------|----------------------------------------------|
 | Variable            | Survey question wording                                                                          |
@@
 
-We also asked caretakers about the last time each participating child was sad or cried.
+We asked caretakers about the last time each participating child was sad or cried.
 This involved asking what the caretaker thought caused the emotion or display and recording their answer in a text box.
@@
 
-From this data, we created a parental response variable, which was coded as 1 for a positve response, 0 for a neutral response, and -1 for a negative response (all from the child's perspective).
-MG also coded each cause for the presence or absence of 17 various elements.
-Of these 4 were use in analyses: whether or not the cause involved 1) the child being asked to do an unwanted task, 2) punishment, 3) conflict within the family, and 4) discomfort, pain, injury, or illness.
-
-### Child anthropometrics
-
-All anthropometric measurements were collected in accordance with anthropometric standards [@lohman1988] and involved taking the mean of two measurements.
-Before being measured, participants were asked to remove their shoes and heavy articles of clothing.
-Height was measured to the nearest millimeter using an anthropometer, and weight was measured to the nearest 100 g with a digital scale (model name) that used within a building with a concrete foundation.
-All participants were wearing lightweight clothing, so we made no adjustments for clothing weight.
-Grip strength was measured using a dynamometer (model name), with participants seated/standing?
-and told to xxxxxxxx.
-XXXX time was passed between the first and second measurements.
-
-We computed each child's height, weight, and BMI Zscores with respect to the CDC child growth standards (cite).
+From this data, we created a parental response variable, which was coded as 1 for a positive response, 0 for a neutral response, and -1 for a negative response (all from the child's perspective).
+MG also coded each case for whether or not the cause had involved: 1) the child being asked to do an unwanted task, 2) punishment, 3) conflict within the family, and 4) discomfort, pain, injury, or illness.
 
@@
 |                |                                                                                       |
-|-------------------------|----------------------------------------------|
+|--------------------------|----------------------------------------------|
 | Variable       | Survey question wording                                                               |
@@
 Those used for this study included the (1) ages of those in the household, (2) sexes of those in the household, (3) partnership status of the primary caregiver, (4) number of adults who provide allocare within the household, (5) highest level of education of the primary caregiver, (5) immigration status (in regards to Utila), and (6) monthly household income.
-This data allowed us to calculate (1) the number of children within the household and (2) the number of adults within the household.
+This data allowed us to calculate the number of children (total, younger, and older) and adults within the household.
 
-OLD AND CURRENTLY NOT USING
+## Child anthropometrics
 
-We also calculated the number of individuals who engaged in childcare within the household, by combining the number of adults who provide care within the household with data from child alloparenting frequencies.
-Due to the fact the adult childcare variable was a count of helpers and the child alloparenting variable was frequency based, we created 3 total alloparenting variables as a robustness check.
-"OtherAlloparents" is the sum of the number of adults who provide childcare plus the sum of other children who engaged in alloprarenting at least "more than once a week but not daily." The second and third involved summing the alloparenting frequency (AlloparentingFreqN) of other children within the household.
-We then summed this with the frequency of adult childcare, which either scored each adult who was listed as providing childcare as helping 30 times per month (AlloparentingFreqNa) or 60 times per month (AlloparentingFreqNb).
+xxxxxxxxxxDescribe camps brieflyxxxxxxxxxxxxxx - include dates and how we handled those
+
+All anthropometric measurements were collected in accordance with anthropometric standards [@lohman1988] and involved taking the mean of two measurements.
+Before being measured, participants were asked to remove their shoes and heavy articles of clothing.
+Height was measured to the nearest millimeter using an anthropometer, and weight was measured to the nearest 100 g with a digital scale (model name) that used within a building with a concrete foundation.
+All participants were wearing lightweight clothing, so we made no adjustments for clothing weight.
+Grip strength was measured using a dynamometer (model name), with participants seated/standing?
+and told to xxxxxxxx.
+XXXX time was passed between the first and second measurements.
+
+Triceps and subscapular skinfold thicknesses were measured using skinfold calipers.
+Measurements were taken directly on the skin.
+The triceps measurement occurred at the midpoint of the upper arm.
+The subscapular measurement occured immediately below the inferior angle of the scapula, being picked up diagonally at the natural fold line of the skin [@antóna].
+
+We computed each child's height, weight, and BMI Zscores with respect to the CDC child growth standards [@kuczmarski2002].
 
@@
 As our signaling variables invovled over-dispersed count data, we used negative binomial regression.
-This was done in R with the GLMM package [@brooks2017], with householdID as a random effect.
+This was done in R with the GLMM package [@brooks2017], with household as a random effect.
+Control variables differed in some models, but always include child age and sex as controls.
+Specific control variables are mentioned in the results
 
@@
 The first used the polr function in the MASS package [@venables2002].
-The results were then interpreted with clustered standard errors for householdID.
-Due to the risk of overfitting that comes with having too few clusters [@cameron2015], we also fit the same models using BRMS with householdID as a random effect.
+The results were then interpreted with clustered standard errors for household.
+Due to the risk of overfitting that comes with having too few clusters [@cameron2015], we also fit the same models using BRMS with household as a random effect.
 
@@
 # Results
+
+`r dplyr::n_distinct(modeldf$householdID)` primary caretakers took the survey and provided signaling data for at least one child.
+`r nrow(modeldf)` children had frequency data for at least 1 signal, with `r nrow(modeldf_FULLSIG)` having data for all 3.
+Among those with at least one record of signaling frequency `r sex_table[1]` were female and `r sex_table[2]` were male.
+Anthropometric and signaling data exist for `r nrow(modeldf_FULLANTHROPOMETRICS)` children, of whom `r sex_table2[1]` were female and `r sex_table2[2]` were male.
+For summary statistics of child-level variables see @tbl-sumtab_children.
+For summary statistics of caretaker and household-level variables see @tbl-sumtab_adults.
 
@@
   Currenthealth3 = "Currenthealth3",
-  CurrentHealthMean = "Mean of current health measures"
+  CurrentHealthMean = "Mean of current health measures",
+  HeightZ = "Height Z-score (CDC)",
+  WeightZ = "Weight Z-score (CDC)",
+  BMIZ = "BMI Z-score (CDC)",
+  SubscapMean = "Subscapular skinfold thickness",
+  SubscapR = "Subscapular skinfold residuals",
+  TricepMean = "Tricep skinfold thickness",
+  TricepR = "Tricep skinfold residuals",
+  BodyFat =  "Body fat composite"
 )
@@
 
-# CONFIRM: should this be household level?
 ```{r sumtab_adults}
@@
   EducationLevelYears = "Education level (years)",
-  IncomeCategoryN = "Household income (Lempira)"
+  IncomeCategoryN = "Monthly household income (Lempira)",
+  number_adults = "Adults in household",
+  AdultsChildcare = "Adults who provide childcare"
 )
 
-kable(custom.summarize(modeldf, vars_c)) %>% kable_styling()
+kable(custom.summarize(caregivers, vars_c)) %>% kable_styling()
 
@@
 Child sadness, crying, and temper tantrums were reasonably common.
-Caregivers reported that (`r sum(na.omit(modeldf$SadFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$SadFreqN >= 8)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\]) of children were sad at least once a week, (`r sum(na.omit(modeldf$CryFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$CryFreqN >= 8)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\]) cried at least more than once a week, and (`r sum(na.omit(modeldf$TantrumFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$TantrumFreqN >= 8)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]) threw temper tantrums at least once a week.
+Caregivers reported that (`r sum(na.omit(modeldf$SadFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$SadFreqN >= 8)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\]) children were sad at least once a week, (`r sum(na.omit(modeldf$CryFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$CryFreqN >= 8)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\]) cried at least more than once a week, and (`r sum(na.omit(modeldf$TantrumFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$TantrumFreqN >= 8)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]) threw temper tantrums at least once a week.
 A small subset of children engaged in these signals/states at least daily (sadness = `r sum(na.omit(modeldf$SadFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$SadFreqN >= 30)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\], crying = `r sum(na.omit(modeldf$CryFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$CryFreqN >= 30)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\], and tantrums = `r sum(na.omit(modeldf$TantrumFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$TantrumFreqN >= 30)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]).
-However, many children were reported to never engage in these signals/states (sadness = `r sum(na.omit(modeldf$SadFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$SadFreqN == 00)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\], crying = `r sum(na.omit(modeldf$CryFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$CryFreqN == 0)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\], and tantrums = `r sum(na.omit(modeldf$TantrumFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$TantrumFreqN == 0)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\])
+However, many children were reported to never engage in these signals/states (sadness = `r sum(na.omit(modeldf$SadFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$SadFreqN == 00)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\], crying = `r sum(na.omit(modeldf$CryFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$CryFreqN == 0)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\], and tantrums = `r sum(na.omit(modeldf$TantrumFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$TantrumFreqN == 0)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]).
+See @fig-barplot_SignalFreq for plot showing signaling frequencies and @fig-signal_alluvial_plot for plot showing the number of individuals who engaged in specific combinations of signal frequencies..
 
 Conflict between children and caretakers was also common, with `r sum(na.omit(modeldf$ConflictFreqN >= 8))` children (`r signif(sum((na.omit(modeldf$ConflictFreqN >= 8)) / sum(!is.na(modeldf$ConflictFreqN)) * 100), 2)`%) having conflicts with parents more than once a week.
-As with the frequencies of various types of signaling, many children (`r sum(na.omit(modeldf$ConflictFreqN == 0))` \[`r signif(sum((na.omit(modeldf$ConflictFreqN == 0)) / sum(!is.na(modeldf$ConflictFreqN)) * 100), 2)`%\]) were described as never having conflicts with parents (stat).
+As with the frequencies of various types of signaling, many children (`r sum(na.omit(modeldf$ConflictFreqN == 0))` \[`r signif(sum((na.omit(modeldf$ConflictFreqN == 0)) / sum(!is.na(modeldf$ConflictFreqN)) * 100), 2)`%\]) were described as never having conflicts with parents.
+See @fig-barplot_ConflictFreq for plot showing the frequencies of conflict between caretakers and children.
 
 The frequency of alloparenting effort was bimodally distributed, with `r sum(na.omit(modeldf$AlloparentingFreqN == 0))` children (`r signif(sum((na.omit(modeldf$AlloparentingFreqN == 0)) / sum(!is.na(modeldf$AlloparentingFreqN)) * 100), 2)`%) being reported to never engage in it (`r sum(na.omit(modeldf$NumberOfChildren==1))` instances were due to a child being an only child) and `r sum(na.omit(modeldf$AlloparentingFreqN >= 30))` children (`r signif(sum((na.omit(modeldf$AlloparentingFreqN >= 30)) / sum(!is.na(modeldf$AlloparentingFreqN)) * 100), 2)`%) being reported to help with childcare at least once a day.
+See @fig-barplot_AlloparentingFreq for plot showing the frequencies child alloparenting effort.
 
@@
 Stacked bar plot showing the frequencies of child sadness, crying, and temper tantrums per month as reported by primary caretaker.
-Frequency key: 0 corresponds to "never," <=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and >30 corresponds to "multiple times per day."
+Frequency key: 0 corresponds to "never," \<=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and \>30 corresponds to "multiple times per day."
 :::
@@
 Color corresponds to sex of the child.
-Frequency key: 0 corresponds to "never," <=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and >30 corresponds to "multiple times per day."
+Frequency key: 0 corresponds to "never," \<=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and \>30 corresponds to "multiple times per day."
 :::
@@
 
-# FIX: y axis extends farther for tantrum model (reducing this manually breaks error shading)
-
 ::: {#fig-effect_conflict}
@@
 
+## Signaling behavior is negatively associated with caregiver education
+
+There were statistically significant negative relationships between years of caregiver education and all of our signaling measures except sadness (`r stats$msadH$EducationLevelYears$str`) when controlling for the sex of the child, the age of the child, household demographics, caretaker perceptions of child immune function, and an alloparenting effort by sex interaction.
+See @tbl-main_models for regression statistics @fig-effect_edu for effects plots.
+
+::: {#fig-effect_edu}
+```{r}
+signaling_data_plot_edu
+```
+
+The effect of years of caregiver education on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior.
+Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
+See @tbl-main_models for the corresponding regression table.
+:::
+
+## Predictors of conflict between children and caretakers
+
+The frequency of conflict between children and caretakers is negatively associated with child age (`r stats$mconflict$ChildAge$str`) and the number of adults who provide childcare in a household (`r stats$mconflict$AdultsChildcare$str`) when controlling for the sex of the child, household demographics, the number of adults who provide and do not contribute to childcare, an alloparenting effort by sex interaction, and the summed frequency of alloparenting from all other children in the household.
+See @tbl-conflict_dep_var for the corresponding regression table and @fig-effect_conflict_dep_var for effects plots.
+
+There was also a negative relationship between conflict frequency and the number of older children, however, this was not statistically significant (`r stats$mconflict$OlderKids$str`).
+Similarly, the model predicted conflict to be more common between caretakers and males, but this also failed to reach statistical significance (`r stats$mconflict$SexMale$str`).
+
+::: {#fig-effect_conflict_dep_var}
+```{r}
+signaling_data_plot_conflict_dep_var
+```
+
+The effect of A) child age and B) the number of adults in the household that contribute to childrcare on the frequency of conflict between caretaker and child.
+Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
+Each dot is 1 child.
+Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate).
+See @tbl-conflict_dep_var for the corresponding regression table.
+:::
+
+```{r}
+#| label: tbl-conflict_dep_var
+#| tbl-cap: Regression table for the model examining the frequency of conflict between child and caretaker as the dependent variable.
+r_vars <- c(
+  ChildAge = "Age (years)",
+  SexMale = "Sex (male)",
+  OtherChildrenHH = "Other children",
+  LogIncome = "Income (log)",
+  number_adults = "Adults in home",
+  PartnerStatusUnpartnered = "Partner status (single)",
+  EducationLevelYears = "Parent ed. (years)",
+  ConflictFreqN = "Conflict freq.",
+  SignalFreq = "Signal freq.",
+  SignalCost = "Signal cost",
+  SignalFreqMax = "Max signal freq.",
+  CurrentHealthMean = "Percieved immunity",
+  SadFreqN = "Sadness freq.",
+  CryFreqN = "Crying freq.",
+  TantrumFreqN = "Tantrum freq.",
+  AlloparentingFreqN = "Child allocare",
+  TricepR = "Tricep skinfold (residuals)",
+  SubscapR = "Subscapular skinfold (residuals)",
+  BodyFat = "BodyFat (composite)",
+  RelativeNeed3 = "Relative need (w/in household)",
+  UnwantedTask = "Unwanted task",
+  Punishment21 = "Punishment",
+  Family21 = "Family conflict",
+  DiscomfortPainInjuryIllness = "Discomfort/Pain/Injury/Illness",
+  CaregiverResponse = "Caregiver Response",
+  OlderKids = "Older children",
+  YoungerKids = "Younger children",
+  AdultsNoChildcare = "Adults that do not provide childcare",
+  AdultsChildcare = "Adults that provide childcare",
+  OtherChildAlloparentingFreqN = "Other household child alloparetning freq."
+)
+modelsummary(mconflict, gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+```
+
 ## Interaction between child sex and child alloparenting
+
+#FIX coords do not work for this
 
@@
 
-The effect of an alloparenting effort by sex interaction on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior.
+The effect of A) child age and B) the number of adults that contribute to childrecare within the household on the frequency of conflict between child and caretaker.
+Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
+Rug corresponds to actual values of child alloparenting frequency with a small amount of jitter to reveal overlapped values.
+See @tbl-conflict_dep_var for the corresponding regression table.
+:::
+
+For each signaling measure, there was also a significant interaction between how often a child cared for their siblings and the sex of the child when controlling for the age of the child, the sex of the child, household demographics, the frequency of conflict with one's primary caretaker, and caretaker perceptions of child immune function.
+The slope of these relationships were positive for girls and flat or slightly negative for boys.
+In other words, our models predict that as children provide more allocare within their families, the frequency of signaling would increase with alloparenting effort for girls and remain at the same rate or decrease slightly with alloparenting effort for boys.
+See @tbl-main_models for the corresponding regression table and @fig-effect_allo_by_sex for effects plots.
+
+## Positive assoication between child allopaenting effort and number of older children
+
+There was a statistically significant negative relationship between the number of older children in a household and the frequency of child alloparenting effort `r stats$malloparenting$OlderKids$str` when controlling for the sex of the child, the number of younger children, household income (log), the number of older children, the number of adults who provide childcare within the household, and the summed frequency of alloparenting from all other children in the household.
+As our model could not handle high values of the summed frequency of other child alloparenting, we removed `r sum(d2$OtherChildAlloparentingFreqN > 121, na.rm = TRUE)` outliers (`r sum(d2$OtherChildAlloparentingFreqN == 360, na.rm = TRUE)` of which had a value of 360 times per month and `r sum(d2$OtherChildAlloparentingFreqN == 420, na.rm = TRUE)` of which had a value of 420 timers per month) despite expecting these values to not be outliers in the broader Utilian community.
+See @tbl-alloparenting_dep_var for the corresponding regression table and @fig-effect_alloparenting_dep_var_plot for effects plots.
+
+::: {#fig-effect_alloparenting_dep_var_plot}
+```{r}
+alloparenting_dep_var_plot
+```
+
+The effect of the number of younger children in the household on the frequency of a child's alloparenting effort.
 Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
@@
 Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate).
-See @tbl-main_models for the corresponding regression table.
+See @tbl-alloparenting_dep_var for the corresponding regression table.
 :::
 
-For each signaling measure, there was also a significant interaction between how often a child cared for their siblings and the sex of the child when controlling for the age of the child, the sex of the child, household demographics, the frequency of conflict with one's primary caretaker, and caretaker perceptions of child immune function.
-The slope of these relationships were curvilinear and positive for girls and flat or slightly negative for boys.
-In other words, our models predict that as children provide more allocare within their families, the frequency of signaling would increase with alloparenting effort for girls and remain at the same rate or decrease slightly with alloparenting effort for boys.
-See @tbl-main_models for the corresponding regression table and @fig-effect_allo_by_sex for effects plots.
+```{r}
+#| label: tbl-alloparenting_dep_var
+#| tbl-cap: Regression table for the model examining the frequency of conflict between child and caretaker as the dependent variable.
+r_vars <- c(
+  ChildAge = "Age (years)",
+  SexMale = "Sex (male)",
+  OtherChildrenHH = "Other children",
+  LogIncome = "Income (log)",
+  number_adults = "Adults in home",
+  PartnerStatusUnpartnered = "Partner status (single)",
+  EducationLevelYears = "Parent ed. (years)",
+  ConflictFreqN = "Conflict freq.",
+  SignalFreq = "Signal freq.",
+  SignalCost = "Signal cost",
+  SignalFreqMax = "Max signal freq.",
+  CurrentHealthMean = "Percieved immunity",
+  SadFreqN = "Sadness freq.",
+  CryFreqN = "Crying freq.",
+  TantrumFreqN = "Tantrum freq.",
+  AlloparentingFreqN = "Child allocare",
+  TricepR = "Tricep skinfold (residuals)",
+  SubscapR = "Subscapular skinfold (residuals)",
+  BodyFat = "BodyFat (composite)",
+  RelativeNeed3 = "Relative need (w/in household)",
+  UnwantedTask = "Unwanted task",
+  Punishment21 = "Punishment",
+  Family21 = "Family conflict",
+  DiscomfortPainInjuryIllness = "Discomfort/Pain/Injury/Illness",
+  CaregiverResponse = "Caregiver Response",
+  OlderKids = "Older children",
+  YoungerKids = "Younger children",
+  AdultsNoChildcare = "Adults that do not provide childcare",
+  AdultsChildcare = "Adults that provide childcare",
+  OtherChildAlloparentingFreqN = "Other household child alloparetning freq."
+)
+modelsummary(malloparenting, gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+```
 
@@
 
-#FIX: how to use coeftest with this
 ```{r}
@@
 )
-modelsummary(list(SignalFreq = mN1, SignalCost = mN2), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+modelsummary(list(SignalFreq = mN1, SignalCost = mN2), vcov = vcovCL, type = "HC0", cluster = ~householdID, gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
 ```
+
+# FIX: want y axes to extend from 0-1.00
 
@@
 
-## Parental responses to last signal of need
+## Caregiver responses to last signal of need
 
@@
 
-#FIX: table is not properly dealing with p values
 ```{r}
@@
 )
-modelsummary(list(CaregiverResponse = mpr), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+modelsummary(mpr, vcov = vcovCL, type = "HC0", cluster = ~householdID, gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
 ```
+
+# FIX: want y axes to extend from 0-1.00
 
@@
 Plots showing the likelihood of positive, neutral, and negative responses to the last time a child became sad or cried.
-A: Effect of punishment being implicated in the cause (0 = no punishment, 1 = punishment.) B: Effect of discomfort, pain, injury, or illness being implicated in the cause (0 = discomfort, pain, injury, or illness, 1 = discomfort, pain, injury, or illness.)
+A: Effect of punishment being implicated in the cause (0 = no punishment, 1 = punishment.) B: Effect of discomfort, pain, injury, or illness being implicated in the cause (0 = discomfort, pain, injury, or illness, 1 = discomfort, pain, injury, or illness.) Legend key: -1 = negative response, 0 = neutral response, 1 = positive response (all from child's perspective).
 :::
@@
 However, the frequency of running away was not significantly associated with conflicts with parents.
-A retrospective study of parent-offspring conflict among Chon Chuuk immigrants in the northwest United States found that running away was the most common response to long-standing conflicts with parents and often resulted in beneficial outcomes for the child [@syme2023].
+A retrospective study of parent-offspring conflict among Chon Chuuk immigrants in the northwest United States found that running away was the most common response to long-standing conflicts with parents and often resulted in beneficial outcomes for the child [@syme2023a].
 The threat running away poses to child wellbeing and, therefore, its ability to credibly signal information about one’s current situation is likely highly dependent on one’s socioecology and age.
-Among the Chon Chuuk, running away is often seen as a suicide threat [@fischer1950; @syme2023], likely due to xxxxxxxxx.
+Among the Chon Chuuk, running away is often seen as a suicide threat [@fischer1950; @syme2023a], likely due to xxxxxxxxx.
 Similar risks may also occur in environments with large predators, venomous animals, or high degrees of human conflict.
@@
 Lowe found that both male and female children reported conflict with their families over their desires to invest financial resources in their own status enhancing opportunities rather than in the household [@lowe2003].
-In an analysis of conflict among Chuukese-Americans, Syme and Hagen found that severe conflicts were more likely with greater numbers of siblings and when conflicts involved a child spending time away from home or evading family responsibilities [@syme2023], findings which would be consistent with parents pressuring older children into investing more in the household among other explanations.
+In an analysis of conflict among Chuukese-Americans, Syme and Hagen found that severe conflicts were more likely with greater numbers of siblings and when conflicts involved a child spending time away from home or evading family responsibilities [@syme2023a], findings which would be consistent with parents pressuring older children into investing more in the household among other explanations.
 In response to such conflicts, participants reported using a wide range strategies as children including, deception, withholding cooperation, running away, and seeking the assistance of others.
@@
 *Body fat discussion*
+
+*More conflict with younger children but despite percieving older children as needy*
 

### 2024-08-31 analysisforpaper.R

@@ -84,7 +84,7 @@ anthropometricMeans$GripR <- residuals(m)
 m2 <- mgcv::gam(TricepMean ~ Sex+s(AgeAtMeasurement, by = as.factor(Sex)), data = anthropometricMeans, na.action = na.exclude)
 anthropometricMeans$TricepR <- residuals(m2)
 
-m3 <- mgcv::gam(SubscapMean ~ Sex+s(AgeAtMeasurement, by = as.factor(Sex), k = 2), data = anthropometricMeans, na.action = na.exclude)
+m3 <- mgcv::gam(SubscapMean ~ Sex+s(AgeAtMeasurement, by = as.factor(Sex), k = 3), data = anthropometricMeans, na.action = na.exclude)
 anthropometricMeans$SubscapR <- residuals(m3)
 
 m4 <- mgcv::gam(BodyFatPercentageMean ~ Sex+s(AgeAtMeasurement, by = as.factor(Sex), k = 3), data = anthropometricMeans, na.action = na.exclude)
@@ -203,6 +203,7 @@ d <- children |>
     RelativeNeedRecode2 = case_when(
       RelativeNeed == "Much more" ~ "More",
       RelativeNeed == "Much less" ~ "Less",
+      #RelativeNeed == "The same amount" ~ "The same",
       .default = RelativeNeed),
     RelativeMaternalInvestment2 = factor( # NOTE this is not always the mom
       RelativeMaternalInvestment2,
@@ -215,7 +216,7 @@ d <- children |>
     RelativeNeed3 = factor(
       RelativeNeedRecode2,
       ordered = TRUE,
-      levels = c("Less", "The same amount", "More")),
+      levels = c("Less", "The same amount", "More")), # change back to "The same" here
     Currenthealth1 = invert(Currenthealth1),
     Currenthealth2 = invert(Currenthealth2),
     Currenthealth3 = Currenthealth3 + 0, # to remove labels
@@ -251,8 +252,8 @@ d <- children |>
   left_join(anthropometricMeans, by = "ChildID")
 
 # add household data to d2
-d <- left_join(d, dplyr::select(caregivers, householdID, CPRatio, Neighborhood, ImmigrateUtila, IncomeCategory,
-                                EducationLevel, AdultsMoney, AdultsHousework, AdultsChildcare, CaregiverAge,
+d <- left_join(d, dplyr::select(caregivers, householdID, CPRatio, Neighborhood, ImmigrateUtila, IncomeCategory, IncomeCategoryN,
[…]

### 2024-09-02 Paper.qmd
Added more to the discussion.

@@
 
-## Positive assoication between child allopaenting effort and number of older children
+## Positive assoication between child alloparenting effort and number of older children
 
@@
 
-The effect of the number of younger children in the household on the frequency of a child's alloparenting effort.
+The effect of the number of older children in the household on the frequency of a child's alloparenting effort.
 Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
@@
 
-There were significant negative relationships between summed signal cost (`r stats$mBFc$BodyFat$str`) crying frequency (`r stats$mBFcry$BodyFat$str`), and tantrum frequency (`r stats$mBFtantrum$BodyFat$str`) and a scaled composite of subscapular and tricep skinfold thicknesses when controlling for the sex of the child, the age of the child, the frequency of conflict with one's primary caretaker, household demographics,caretaker perceptions of child immune function, and an alloparenting effort by sex interaction.
+There were significant negative relationships between summed signal cost (`r stats$mBFc$BodyFat$str`) crying frequency (`r stats$mBFcry$BodyFat$str`), and tantrum frequency (`r stats$mBFtantrum$BodyFat$str`) and a scaled composite of subscapular and tricep skinfold thicknesses when controlling for the sex of the child, the age of the child, the frequency of conflict with one's primary caretaker, household demographics,caretaker perceptions of child immune function, an alloparenting effort by sex interaction and a body fat by sex interaction.
 See @tbl-bodyfat_models for regression statistics and @fig-effect_bodyfat for effects plots.
@@
 See xxxxSI TABLESxxxx for regression tables.
+
+Although not significant, our models predicted a negative slope for the relationship between body fat and summed signal cost, crying frequency, and tantrum frequency for girls but no relationship for boys.
 
@@
 
+We found no significant associations between sex and child signaling behavior outside of the previously mentioned interactions, including when sex was included in models as interaction with age.
+See table SI xxxxxxx for regression tables which included the sex by age interaction.
+Monthly household income also was not a significant predictor of signaling behavior in any model nor was population when using language, immigration status, and neighborhood as proxies for it.
+See table SI xxxxxxx.
+
 # Discussion
@@
 There was no significant association between the frequency of child sadness and age.
-Part of this likely results from the fact that, unlike crying, sadness is thought to have important non-signaling functions including drawing one’s attention to a problem, motivating deeper analysis of one’s situation, and causing withdrawal from situations with low fitness prospects (cite), from which others can still obtain cue-based information about one's emotional state, well-being, and desires.
+Part of this likely results from the fact that, unlike crying, sadness is thought to have important non-signaling functions including drawing one’s attention to a problem, motivating deeper analysis of one’s situation, and causing withdrawal from situations with low fitness prospects (cite), from which, others can still obtain cue-based information about one's emotional state, well-being, and desires.
 By asking about sadness as a whole, our measure was unable to disentangle information caretakers gleaned from cues from information that came from signals.
-For this reason, a change in non-crying signals that arise from sadness (e.g., facial sadness and some, but all, forms of withdrawal) may have been harder to detect with our sample size even if young children expressed facial sadness more often.
+For this reason, a change in signals that arise from sadness outside of crying (e.g., facial sadness and some, but all, forms of withdrawal) may have been harder to detect with our sample size even if young children expressed facial sadness more often.
 It is also possible that there is no effect of age on sadness frequency among the age ranges examined it.
-add in depression over the lifecourse
+
+- add in depression over the lifecourse
 
@@
 Although conflicts of interest do not always lead to behavioral conflict [@altmann1980; @bateson1994], it is likely many, if not most, conflicts are ultimately caused by conflicts of interest, making it a useful proxy for the underlying conflicts of interest that may increase the benefits of deception and the costs of being deceived\[\[5\]\].
-If this is the case, the association between signaling and conflict may represent children's attempts at credibly signaling their value of additional investment or the costliness of punishment to parents who would have good reason to doubt the trustworthiness of simple verbal request (or the trustworthiness of facial sadness and crying for those resorting to temper tantrums).
-
-When viewed in isolation, the frequencies of sadness, crying, and throwing temper tantrums were also positively associated with conflict frequency between child and caretaker.
+If this is the case, the association between signaling and conflict may represent children's attempts at credibly signaling their value of additional investment or the costliness of punishment (cite) to parents who would have good reason to doubt the trustworthiness of simple verbal request (or the trustworthiness of facial sadness and crying for those resorting to temper tantrums).
+
+Probably cut
+When viewed in isolation, the frequencies of sadness, crying, and temper tantrums were also positively associated with conflict frequency between child and caretaker.
 However, the frequency of running away was not significantly associated with conflicts with parents.
@@
 
-*Sex and focal child alloparenting effort interact (postive for females, negative/flat for males)*
+*Sex and focal child alloparenting effort interact (positive for females, negative/flat for males)*
 
@@
 However, Utilian mothers reported in focus groups (confirm) that girls tend to provide more help for their siblings than boys.
-If this difference was the outcome of caretakers pressuring girls to do more (or resulted in the expectation of more alloparenting), the increase in signaling with alloparenting effort seen among girls may be the result of a strategy to signal the costliness of their efforts in order to invest more in romantic relationships and status seeking behaviors outside the household.
+If this difference was the outcome of caretakers pressuring girls to do more or resulted in the expectation of more alloparenting, the increase in signaling with alloparenting effort seen among girls may be the result of a strategy to signal the costliness of their efforts in order to invest more in romantic relationships and status seeking behaviors outside the household.
 
@@
 
+*Alloparenting effort model*
+Not sure on interpretation here yet
+
 *Body fat discussion*
 
-*More conflict with younger children but despite percieving older children as needy*
-
-*Lack of other sex differences*
+Children with less body fat cried more, threw more temper tantrums, and had higher summed signal costs.
+In other words, needier children signaled more often.
+Although the body fat by sex interaction was only marginally significant, our models predict a negative relationship between body fat and signaling for girls and no relationship for boys.
+Female reproduction is highly dependent on body fat, with conception requiring about xx% body fat and mothers investing a significant amount of calories in both children in the womb and breastfed infants (cite).
+For these reasons, low body fat is likely a be better predictor of future reproductive success in older girls than boys, potentially making signaling for increased investment a more viable strategy in instances when an individual has low body fat.
 
@@
 There was a positive association between the summed frequency of child signaling and caretaker perceptions of child neediness within the family, something which would be consistent with sadness, crying, and temper tantrums being signals (or involving signals in the case of crying) if parents took this to mean how much they value additional investment.
+Caretakers also perceived older children as needier, despite having more conflicts with younger children.
+One reason for this could be that caretakers weighted monetary investment highly when evaluating child need.
 
@@
 a skinned knee).
-It may also be that many of the typical positive responses to pain and illness are not costly (e.g., providing comfort).
+It may also be that many of the typical positive responses to pain and illness are not very costly (e.g., providing comfort).
 
@@
 
-One unexpected finding was a negative association between caretakers’ education level (years in school) and many of our child signaling measures despite our models controlling for household income and partner status.
+We found a negative association between caretakers’ education level (years in school) and many of our child signaling measures despite our models controlling for household income and partner status.
 One possible reason for this may be that more educated caretakers see their children less during the day due to greater ability to afford childcare or due to them having jobs that are less compatible with child care (test model of social support predicted by education).
 Unfortunately, we do not have data on how much time caretakers spent with children to test this, and it seems equally likely that less educated parents may see their kids less often if they need to work more hours to support their families.
-Another possibility is that lower levels of education are associated with greater adversity for both parents and children, something which would increase the benefits of signaling need for children and the difficulty of parents to respond.
+Another possibility is that lower levels of education are associated with greater levels of adversity for both parents and children, something which would increase the benefits of signaling need for children and the difficulty of parents to respond.
 It is also possible that education influences parenting norms in ways that cause differences in child signaling behavior.
@@
 
+*Lack of evidence for demographic effects*
+
+Outside of the sex by alloparenting effort interaction, we found little evidence of sex differences in child signaling behavior.
+As adults, females cry more than males, a finding that has been reported across cultures (cite).
+A similar sex difference is also seen with depression and suicide attempts, which some of us have argued to often be attempts at signaling need in other papers (cite), with females experiencing depression being about twice as often (cite) and young females being xxxx as likely to attempt suicide (normal citaiton).
+The most likely reason for this difference is that there is less of a sex difference in adversity, conflict, and cultural evaluations of signals of need children until adolescence and young adulthood.
+Although we do have data on adolescents, our small sample for this age range may have prevented us from detecting this.
+
+Monthly income was also not found to be a significant predictor of signaling behavior.
+Many models designed to explain signals of need (e.g., the begging of young birds or withdrawal in humans) predict that offspring who are worse off or are receiving less investment have more to gain from signaling than those with more (cite).
+However, it is possible that this can be true and household level resources have little influence on child signaling frequencies between families.
+As signaling behavior is costly to child, caretakers, and siblings, children may gain direct and indirect fitness benefits from limiting their signaling behavior in times of resource stress (i.e., when the chances of gaining investment or the size of investment are small).
+Similarly, children in high resource families may also be motivated to signal at higher than expected frequencies due to the potential for larger pieces of investment.
+We know of no empirical explorations of this in humans (check more and confirm).
+However, multiple game theoretical models of sibling competition and have found that parental resources have a small impact on offspring signaling, especially with high rates of monogamy [Bossan + Mock and Parker + Check for newer sources].
+Although counterintuitive, such an outcome fits the fact that fitness is relative and that natural selection should be expected to have favored those who maximized their own inclusive fitness, whether one found themselves in a great situation or a dire one.
+
+Despite population level differences in wealth, job certainty, discrimination, and living conditions [@davis2014; @garcia2017; @garcia2018; @stonich2000], we generally did not find an effect of population on child signaling behavior.
+Although we sought data from both English-speaking participants born on the island and Spanish-speaking immigrants from the mainland, our smaller sample of English-speaking Utilians may have limited our ability to detect real differences between the populations.
+It is possible that differences in wealth and status between the groups translate to differences in child signaling due to children facing different types adversity or weighting different types of investment differently.
+It is also likely that the two groups differ in how caretakers could respond to child signaling, and caretakers between the two populations may also differ in their cultural models of parenting.
+However, as previously mentioned, it is also conceivable that household level differences have little effect on child signaling behavior due to the costs of signaling and a child's fitness interest in their relatedness.
+
 ## Limitations
@@
 Due to the fact the data used here were a part of survey which covered many topics and often multiple children, we sometimes faced a tradeoff between survey time and question quality.
-Asking adult participants about their relative investment in children in their household was likely not as effective as more in depth, single child focused questions due to both the relative lack of detail it provides and the potential for social desirability bias [@grimm2010], which could lead adult participants to be be more likely to report equal investment across children.
+Asking adult participants about their relative investment in children in their household was likely not as effective as more in depth, questions focused on specific questions.
+Framing the question this way may have also increased the likelihood of social desirability bias [@grimm2010], which could lead adult participants to be be more likely to report equal investment across children.
 Our questions about the last time the child signaled may have also been more effective if they focused on specific signals due to our methods combining crying, which is unambiguous as a signal, with sadness, which has both signaling and non-signaling functions.
-
-Despite seeking data from both English-speaking participants born on the island Spanish-speaking immigrants from the mainland, our ability to detect differences between the populations may limited by our smaller sample of English-speaking Utilians.
-It is possible that differences in wealth and status between the groups translate to differences in child signaling due to children facing differerent types adversity or weighting different types of investment differently.
-It is also likely that the two groups differ in how caretakers could respond to child signaling, and caretakers between the two populations may also differ in their cultural models of parenting.
 

### 2024-09-04 Paper.qmd
Updated plots to better show the effect of predictors.
Added to limitations section and edited the introduction, condensing and removing sections.

@@
 ---
-title: "Child Signals of Need in Utila, Honduras"
+title: "Child Signals of Need in Utila, Honduras: An Exploratory Study"
 execute: 
@@
 
-We argue that older children (i.e., those older than infants) have evolved strategies to tailor their signals to 1) their valuation of a given resource, 2) the extent of conflicts of interest with caretakers, and 3) the presence of existing information about a child's need or potential to benefit from a resource so that they can credibly convey they are worth additional investment.
+In this paper, we use exploratory analysis on child signaling data from Utila Honduras to examine the hypothesis that children have evolved strategies to tailor their signals of need to 1) their valuation of a given resource, 2) the extent of conflicts of interest with caretakers, and 3) the presence of existing information about a child's need or potential to benefit from a resource so that they can credibly convey they are worth additional investment.
 Such strategies can lead to some combination of verbal requests for more support, facial sadness, crying, withdrawal, temper tantrums and self-harm, all of which may be effective in certain circumstances and are sometimes successful in eliciting more support from caretakers, despite a child's low status and formidability.
@@
 Throughout human evolutionary history, individuals often found themselves in situations in which they received fitness-relevant benefits from others that could not easily be replaced [@apicella2012; @apicella2019; @hill2011; @tooby1996].
-In these instances, natural selection is expected to have favored those who provided support to others if 1) such help meant one could expect to receive fitness benefits from those helped in the future and 2) the costs of providing help were outweighed by the benefits to the helper on average[@andré2010; @hamilton1964; @tooby1996; @trivers1971].
-At the very least, this would have taken the form of mothers providing investment to their children due to direct fitness benefits.
+In these instances, natural selection is expected to have favored those who provided support to others if 1) such help meant one could expect to receive fitness benefits from those helped in the future and 2) the costs of providing help were outweighed by the benefits to the helper on average [@andré2010; @hamilton1964; @tooby1996; @trivers1971].
+At its simplest, this would have taken the form of mothers providing investment to their children.
 
 The fitness interdependence of offspring and at least one parent is seen in many animals [@clutton-brock1991; @webb1999].
-However, humans are a uniquely interdependent species among the living primates [@tomasello2012], with one’s fitness prospects depending greatly on a network of kin and other social allies [@apicella2019].
-Although also true with adults, this need for others is most intense with children, who are born highly altricial and require many years of investment to develop to the point where they can obtain enough food to survive on their own and develop the social skills needed to succeed in their communities[@kaplan2009; @rosenberg2021; @schniter2015].
-Much of this investment comes from parents[@kramer2018], with most human reproduction occurring in the context of biparental investment and long-term pair bonds [@apostolou2007; @schacht2019].
+However, humans are a uniquely interdependent species among the living primates [@tomasello2012] in that one’s fitness prospects depend greatly on a network of kin and other social allies [@apicella2019].
+Although also true with adults, this need for others is most intense with children, who are born highly altricial and require many years of investment to develop to the point where they can obtain enough food to survive on their own and develop the social skills needed to succeed in their communities [@kaplan2009; @rosenberg2021; @schniter2015].
+Much of this investment comes from parents [@kramer2018], with most human reproduction occurring in the context of biparental investment and long-term pair bonds [@apostolou2007; @schacht2019].
 However, in many societies, other adults like grandparents, aunts, uncles, and non-relatives have also been found to provide significant investment in children [@kramer2018; @sear2008], something which was likely favored due to indirect fitness benefits, reciprocity, or reputation management [@gurven2000] + (Schniter Preprint (permission needed).
@@
 As a sexually reproducing species, humans have a 50% chance of passing down a given gene to each child.
-Each child, however, shares more genes with themselves (100%) than with their full siblings (50% on average), half-siblings (25% on average), and stepsiblings (0%).
+Each child, however, shares more genes through descent with themselves (100%) than with their full siblings (50% on average), half-siblings (25% on average), and stepsiblings (0%).
 Due to this asymmetry, children are often expected to attempt to elicit more investment from parents than is optimal from the perspective of the parent, while parents are expected to attempt to invest less than is optimal for each child in order to invest in their other children, whom parents are expected to value identically, all else being equal [@trivers1974].
@@
 Sibling competition is evident across taxa and can range from relatively benign forms like strong begging competition in chicks [@caro2016] to siblicide in birds, mammals, and sharks [@salmon2014].
-In humans, sibling competition is largely estimated by counts of siblings, with large numbers of siblings having been associated with decreases in growth [@hadley2011; @hagen2006; @salmon2014], survival [@lawson2012; @meij2009; @strassmann2011], and reproductive success [@gibson2011; @mace1996] in multiple populations, highlighting the likely significance of sibling competition for resources, even if this does not always manifest with behavioral conflict [@bateson1994].
-
-While parent-offspring conflict and sibling competition are likely ubiquitous, the resources being contested will vary across human groups.
+In humans, sibling competition is largely estimated by counts of siblings, and large numbers of siblings have been associated with decreased growth [@hadley2011; @hagen2006; @salmon2014], survival [@lawson2012; @meij2009; @strassmann2011], and reproductive success [@gibson2011; @mace1996] in multiple populations, highlighting the likely significance of sibling competition for resources, even if this does not always manifest with behavioral conflict [@bateson1994].
+
+Potentially cut this paragraph: While parent-offspring conflict and sibling competition are likely ubiquitous, the resources being contested will vary across human groups.
 In relatively egalitarian and food constrained populations, conflict over resources is likely to be centered on food, which probably explains the negative impacts siblings have on child age-specific heights and weights among the Shuar of Ecuador [@hagen2009], although similar effects have also been found in industrialized populations [@lawson2008].
@@
 As previously mentioned, each individual child is expected to prefer greater investment in oneself than in one’s siblings, while parents, who are equally related to all their children, should often be expected to favor a more even distribution [@trivers1974].
-This can lead to conflict with older children who might prefer to spend their time in ways which are likely to result in future direct fitness benefits (e.g., finding romantic partners or forming strong bonds with friends) rather than taking care of their siblings or acquiring or preparing food.
-Such a dynamic has been found among the Chuukese, with both male and female children reporting conflict with their families over children’s desires to invest financial resources in their own status enhancing opportunities rather than in the household [@lowe2003].
-Similarly, Syme and Hagen [@syme2023a] found greater numbers of siblings were associated with increased likelihoods of severe conflict in a retrospective study with Chuukese-Americans.
-This may also be consistent with parents pressuring older children to invest in the household; however, there are other ways in which limited resources can result in behavioral conflict [@syme2023a].
-
-Conflicts of interest over investment in the household also appear to extend to when a child starts having their own children [@kushnick2023; @emlen1995].
-Due to the benefits children may provide their siblings, parents may be motivated to delay or subvert the reproductive careers of certain offspring [@flinn1989; @emlen1992], a conflict which is likely enhanced in families with step-parents due to the likelihood of future children being step-siblings rather than full siblings, thereby reducing the potential fitness benefits of alloparenting [@moya2014; @schaffnit2021].
-Situations can also arise in which it might pay parents for their children to start having their own children sooner than would be ideal for the adult child.
-For example, parents may be motivated for a child to leave the household and begin having children earlier than is ideal from the child’s perspective if their presence in the household leads to them taking investment from other children who caretakers deem more worthy of investment [@kushnick2023].
-
-Studies testing the existence of familial conflict over the timing of child reproduction are rare and direct tests on the existence of parent-offspring conflict over when a child should start their own reproductive career versus help existing siblings in humans are needed [@schaffnit2021].
-The bulk of evidence for the existence of such conflicts comes from work with non-human animals. For example, the cooperative breeding white-fronted bee-eaters (*Merops bullockoides*) will interrupt the courtship of their offspring, seemingly to continue to benefit from their undiluted assistance in digging nests, incubating the clutch, finding food for nestlings, and defending the nest [@emlen1992].
-When it comes to humans, Flinn found that rural Trinidadian females contributed substantial amounts of childcare and those with siblings under 10 years old had their first child at a later age [@flinn1989], both of which would be expected if there is a conflict of interest over investing in one’s siblings or one’s own reproduction.
-However, a study among Karo Batak women found no evidence of delayed reproduction for alloparenting benefits [@kushnick2006].
+This can lead to conflict with older children who might prefer to spend their time in ways which are likely to result in future direct fitness benefits (e.g., finding romantic partners or forming strong bonds with friends) rather than taking care of their siblings or acquiring or preparing food [@lowe2003; @syme2023a].
+Conflicts of interest over investment in the household also appear to extend to when a child starts having their own children [@kushnick2023; @emlen1995], with parents potentially benefiting from delaying the reproductive careers of certain offspring in order to gain more care for their other children [@flinn1989; @emlen1992] or speeding up a child’s departure from the household if the child’s presence limits how much caretakers can invest in other children they deem more worthy of investment [@kushnick2023].
 
@@
 Similarly, !Kung San mothers frequently respond to crying infants by providing comfort and breastfeeding access, with comforting behaviors, in general, being far more common than one’s meant to restrict infant behavior [@kruger2010].
-Although this study found mothers were the most common responders, siblings, fathers, and other adults frequently stepped in so that 88% of crying bouts were responded to, with most of the cries not responded to lasting under ten seconds [@kruger2010].
+Although this study found mothers were the most common responders, siblings, fathers, and other adults frequently stepped in so that 88% of crying bouts were responded to, with most of the cries not responded to lasting under ten seconds.
 Similar responsiveness has also been reported among the Aka [@hewlett1998], Mbendjele hunter-gatherers [@chaudhary2024], the Efe [@tronick1987], and the Ache, among whom young children have been described as “win\[ning\] most conflicts with parents simply by crying and whining” [@kaplan1987].
@@
 Crying can also lead to negative responses beyond being ignored [@barr1990; @soltis2004; @zeifman2017].
-It has been found to increase parental distress, the likelihood of maternal depression, and the child’s risk of maltreatment ~~especially if the child is in poor condition~~ [@frodi1985; @fujiwara2011; @howard2006; @lin2012; @soltis2004].
+It has been found to increase parental distress, the likelihood of maternal depression, and the child’s risk of maltreatment [@frodi1985; @fujiwara2011; @howard2006; @lin2012; @soltis2004].
 For example, in a sample of 3,259 parents of Dutch infants under 6 months old, 5.6% self-reported smothering, slapping, or shaking their babies in response to their crying, outcomes which were more likely among parents who were single, unemployed, or deemed their infant’s crying to be excessive [@reijneveld2004].
@@
 Although widely accepted as a signal, there is no consensus on why emotional crying may work to elicit positive responses.
-One possibility is that crying is physiologically constrained in ways that prevent or limit faking.
-Evidence consistent with this comes from findings that crying is hard to do on command and findings individuals often cry while alone, something which is not predicted by signaling theory [@vingerhoets2001].
-However, the likelihood of crying is often clearly related to presence of supporters, something which is seen in findings that lonely individuals cry less than those with more support options despite reporting lower well-being [@vingerhoets2013] and the existence of delayed crying—times when crying is inhibited during the cause of negative affect only for it to be expressed when one is near potential supporters [@gracanin2018].
-A more likely possibility is that certain acoustic elements of infant cries are tied to their physiological state in ways that make faking certain states impossible, with crying serving as a graded signal in most other instances [@soltis2004; @zeifman2017].
-
-It is likely that the functions of crying and the mechanisms behind its ability to elicit pro-signaler responses change over one’s life.
-Compared to adults and older children, the physiological costs of crying are greater for newborns and infants, among whom crying can result in a substantial increase in energy expenditure [@rao1997; @thureen1998].
-Due to these costs, crying at certain intensities or for long periods of time may be prohibitively costly for infants with low survival prospects, making it a potential credible signal of vigor [@furlow1997; @lummaa1998; @soltis2004; @wells2003].
-Such signals are thought to be important due to parental adaptations which are suspected to take the viability of offspring into account when making decisions about whether it is worth investing in a specific child or withdrawing investment to instead invest in other children who are more likely to survive to reproductive age and be more competitive adults [@hagen1999; @lummaa1998].
-In line with this reasoning, most infanticides cross-culturally occur when a child has poor survival prospects, due to either health conditions or resource constraints [@daly1984; @soltis2004].
-
-Another often suggested means by which crying could result in credible signals of need is if the increased risk of predation that comes from loud vocalizations allows for signaling to be worthwhile for those in relatively high need but not for those with low need [@lummaa1998; @wells2003; @zahavi1977].
-Evidence of increased predation has been found in studies on birds [@leech1997; @lima2009; @redondo1992; @ryan1982], among whom louder broods in general have also been found to be more conspicuous to predators [@haskell1997; @leech1997].
-It is unclear how significant the risk of alerting predators would be ancestrally.
-However, there are ethnographic reports of mothers keeping infants and young children close in dangerous environments [@hewlett1998; @kaplan1987; @konner1972], something which likely lessens this risk[\[MG10\]](#_msocom_10) .
-
-As children age, their reproductive value increases due to reduced odds of mortality, leading to a decreased risk of infanticide (Hagen 2004)[\[MG11\]](#_msocom_11) .
-For this reason, there is likely a transition period in which crying begins to convey information about how much a child values further investment.
+One possibility is that the physiological costs of crying for newborns and infants [@rao1997; @thureen1998] may be prohibitively costly for faking to pay off.
+When this is the case, crying may serve as a credible of vigor [@furlow1997; @lummaa1998; @soltis2004; @wells2003], something which may be particularly important early in life due to increased risk of infanticide for children who have has poor survival prospects, due to either health conditions or resource constraints [@daly1984; @soltis2004].
+
+Potentially combine and condense with paragraph below: As children age, their reproductive value increases due to reduced odds of mortality, leading to a decreased risk of infanticide (Hagen 2004).
+For this reason, there is likely a transition period in which crying is primarily meant to provide information about how much a child values further investment.
 Many models have been put forward which attempt to explain human crying and begging in other species as credible signals of need due to the differential costs and benefits of crying and analogous vocalizations in other species as deterring the exaggeration of need [@godfray1991; @kilner1997; @parker2002].
@@
 
-When it comes to older children, the relatively insignificant energetic costs of crying are unlikely to be enough to result in credible signals in times of substantial conflicts of interest.
+For older children, the relatively insignificant energetic costs of crying may not be enough to result in credible signals in times of substantial conflicts of interest.
 For this reason, crying among older children may be more similar to adult crying, something which has been found to result in increased support and empathy but sometimes come at the cost of reputation and the potential for stigmatization [@cornelius2001; @cretser1982; @hendriks2006; @jesser1989; @vingerhoets2016].
@@
 Due to their lower status, children have limited ways to change how others treat them.
-However, as children 1) are their parents’ path to direct fitness and 2) can help family members through allocare and other forms of labor, they provide benefits that are valuable and hard to replace.
+However, as children 1) are a major component of parents' inclusive fitness and 2) can help family members through allocare and other forms of labor, they provide benefits that are valuable and hard to replace.
 By withholding these benefits or putting the future of these benefits into question, a child simultaneously delays the benefits they get from this cooperation while imposing costs on those with a fitness interest in the child.
@@
 However, facial expressions, along with tears, may not be enough to credibly signal need in times of significant conflicts of interest.
-In these instances, depression and suicidal behavior have been suggested to function as ways to credibly signal need due to the expectation that such states are costlier for those facing adversity than those who are not (for more detail on these approaches see: [@hagen2003; @rosenthal1993; @syme2016; @watson2002].
+In these instances, depression and suicidal behavior have been suggested to function as ways to credibly signal need due to the expectation that such states are costlier for those facing adversity than those who are not (for more detail on these approaches see: [@hagen2003; @rosenthal1993; @syme2016; @watson2002; @gaffney2022].
 
@@
 For example, Utilian mothers reported in focus groups that some temper tantrums involved children banging their heads on things when they do not get what they want, something that has been reported in other populations [@belden2008; @burrows1953; @daniels2012; @dennis1940; @vandenakker2022].
-Similarly, young chimpanzees (*Pan troglodytes*) have been observed to beat their heads on the ground when refused access to breastmilk [@goodall1986], and American white pelican chicks (*Pelecanus erythrorhynchos*) have been found to throw themselves on the ground, flail violently, and bite at their own wings when begging for food [@schaller1964].
-Such behaviors have been suggested to follow similar logic to depression and suicide attempts as they are likely relatively less costly to those who are truly in dire need compared to those who are not to the point faking is not worth the costs [@hagen2008] (Hagen et al.
-2008).
+Similarly, young chimpanzees (*Pan troglodytes*) have been observed to beat their heads on the ground when refused access to breast milk [@goodall1986], and American white pelican chicks (*Pelecanus erythrorhynchos*) have been found to throw themselves on the ground, flail violently, and bite at their own wings when begging for food [@schaller1964].
+Such behaviors have been suggested to follow similar logic to depression and suicide attempts as they are likely relatively less costly to those who are truly in dire need compared to those who are not to the point faking is not worth the costs [@hagen2008].
 In other words, the risk of injury from violent temper tantrums carries greater fitness costs to a child who is receiving enough support to develop high levels of embodied capital than it does to a child who is neglected to the point where their survival prospects are low.
@@
 This study was designed to help establish the natural history of child signaling behavior in Utila, Honduras, by quantifying the frequencies of common child signals of need and various types of parental responses.
-From this, we sought to examine 1) the causes of child signals of need, 2) how and why signal costs might vary, 3) how parents respond to indications of child need, 4) how signaling behavior changes with age and sex, 5) and if there were differences in child signaling and parental responses among two sub-populations.
+From this, we sought to use signaling theory to examine 1) the causes of child signals of need, 2) how and why signal costs might vary, 3) how parents respond to indications of child need, 4) how signaling behavior changes with age and sex, 5) and if there were differences in child signaling and parental responses two groups who face different levels of social and economic adversity.
 
@@
 When experiments are not appropriate, longitudinal studies with carefully selected variables can help address this (cite).
-However, our small sample size with signaling data and anthropometric data at two time periods prevents this for this study.
+However, our small sample size with signaling data and anthropometric data at two time periods limits our ability to do this in this study.
 For this reason, all analyses for this study were exploratory and meant to inform future studies of child signaling.
@@
 The subscapular measurement occured immediately below the inferior angle of the scapula, being picked up diagonally at the natural fold line of the skin [@antóna].
+Add flexed bicep measurementxxxxxx.
 
@@
 
+To model the effects of skinfold thickness, flexed bicep circumference, and grip strength on signaling behavior, we used the residuals from generalized additive models which modeled the effect of sex and age on each measurement using the mgcv package [@wood2011].
+This resulted in variables which represented how each child differed from the average child for their given sex and age.
+
 ## Statistical analyses
+
+All statistical analyses were exploratory and involved fitting multiple models while being cognizant of the risk of overfitting the data, but with the payoff of idenftifying theoretically relavent effecrs that should be tested in the future.
 
@@
 
-Alluvial plot showing the combinations of signaling frequencies.
+Alluvial plot showing the combinations of signaling frequencies per month.
 Alluvium width equals the number of individiuals with the same frequencies for the three signals/states.
@@
 
-Stacked bar plot showing the frequency of conflict between child and primary caregiver as reported by primary caretaker.
+Stacked bar plot showing the frequency of conflict (per month) between child and primary caregiver as reported by primary caretaker.
 Color corresponds to sex of the child.
@@
 
-Stacked bar plot showing the frequency of conflict a child's alloparenting effort as reported by primary caretaker.
+Stacked bar plot showing the frequency of a child's alloparenting effort (per month) as reported by primary caretaker.
 Color corresponds to sex of the child.
@@
 
-The effect of age on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior.
+The effect of age on specific signal frequencies (per month), summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior.
 Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
@@
 
-The effect of the frequency of conflict with one's primary caretaker on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior.
+The effect of the frequency of conflict with one's primary caretaker (per month) on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior.
 Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
@@
 ```{r}
-signaling_data_plot_edu
+signaling_data_plot_education
 ```
@@
 
-The effect of A) child age and B) the number of adults in the household that contribute to childrcare on the frequency of conflict between caretaker and child.
+The effect of A) child age and B) the number of adults in the household that contribute to childcare on the frequency of conflict between caretaker and child.
 Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
@@
 
-#FIX coords do not work for this
-
 ::: {#fig-effect_allo_by_sex}
@@
 
-The effect of A) child age and B) the number of adults that contribute to childrecare within the household on the frequency of conflict between child and caretaker.
+The effect of an alloparenting effort by sex interaction on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior.
 Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
@@
 ```{r}
-alloparenting_dep_var_plot
+# alloparenting_dep_var_plot
 ```
@@
 
-# FIX: want y axes to extend from 0-1.00
+# FIX: error ribbons cross 1 and 0
 
@@
 
-# FIX: want y axes to extend from 0-1.00
-
 ::: {#fig-effect_last_signal_response}
@@
 
-Probably cut
-When viewed in isolation, the frequencies of sadness, crying, and temper tantrums were also positively associated with conflict frequency between child and caretaker.
+Probably cut When viewed in isolation, the frequencies of sadness, crying, and temper tantrums were also positively associated with conflict frequency between child and caretaker.
 However, the frequency of running away was not significantly associated with conflicts with parents.
@@
 
-*Alloparenting effort model*
-Not sure on interpretation here yet
+*Alloparenting effort model* Not sure on interpretation here yet
 
@@
 We know of no empirical explorations of this in humans (check more and confirm).
-However, multiple game theoretical models of sibling competition and have found that parental resources have a small impact on offspring signaling, especially with high rates of monogamy [Bossan + Mock and Parker + Check for newer sources].
+However, multiple game theoretical models of sibling competition and have found that parental resources have a small impact on offspring signaling, especially with high rates of monogamy \[Bossan + Mock and Parker + Check for newer sources\].
 Although counterintuitive, such an outcome fits the fact that fitness is relative and that natural selection should be expected to have favored those who maximized their own inclusive fitness, whether one found themselves in a great situation or a dire one.
@@
 
-A major challenge when studying signals of need is the fact that such signaling is expected to 1) be more common in times when individuals can benefit from increased investment (e.g., when they are facing major adversity or when a fitness-relevant opportunity arises) and 2) result in benefits to the signaler on average.
-In other words, higher levels of signaling may stem from greater levels of need while simultaneously eliciting more investment from others.
-This hypothisized causal, bi-directional relationship makes it difficult to isolate the effects of signaling behavior on investment in children when using standard measures of child welfare (i.e., BMI, health metrics, and height for age).
-Longitudinal studies can help address this.
-However, our small sample size of participants with signaling data and anthropometric data at two time periods prevents this for this study.
-
-Due to the fact the data used here were a part of survey which covered many topics and often multiple children, we sometimes faced a tradeoff between survey time and question quality.
+The main limitation of this study is that it is completely exploratory, something which results in a high risk of our models overfitting the data and indicating associations which might not exist in the real world.
+Our use of convenience and snowball sampling is also a limitation as it resulted in a non-probability sample which may not be representative of the broader population.
+This could happen if there was selection bias so that those who participated differed from the rest of the population in how they interacted with their children or how they assessed, remembered, and reported child behavior.
+
+Outside of child anthropometric measurements, all data come from caregiver reports, and therefore, reflect the biases of caretakers, whatever they may be.
+Due to the fact the data used here were a part of survey which covered many topics and often multiple children, we also faced a tradeoff between survey time and question depth.
 Asking adult participants about their relative investment in children in their household was likely not as effective as more in depth, questions focused on specific questions.
@@
 Our questions about the last time the child signaled may have also been more effective if they focused on specific signals due to our methods combining crying, which is unambiguous as a signal, with sadness, which has both signaling and non-signaling functions.
+
+Another limitation is that our study was cross-sectional.
+A major challenge when studying signals of need is the fact that such signaling is expected to 1) be more common in times when individuals can benefit from increased investment (e.g., when they are facing major adversity or when a fitness-relevant opportunity arises) and 2) result in benefits to the signaler on average.
+In other words, higher levels of signaling may stem from greater levels of need while simultaneously eliciting more investment from others.
+This hypothesized causal, bi-directional relationship makes it difficult to isolate the effects of signaling behavior on investment in children when using standard measures of child welfare (i.e., BMI, health metrics, and height for age).
+Longitudinal studies can help address this.
+However, our small sample size of participants with signaling data and anthropometric data at two time periods prevents this for this study.
 
@@
 [\[MG12\]](#_msoanchor_12)Del Giudice, 2018 too?
+
+# Cut from paper:
+
+## Intro
+
+### Conflicts of interest within the family
+
+Conflicts of interest between children and parents may also involve how much children invest in the household.
+As previously mentioned, each individual child is expected to prefer greater investment in oneself than in one’s siblings, while parents, who are equally related to all their children, should often be expected to favor a more even distribution [@trivers1974].
+This can lead to conflict with older children who might prefer to spend their time in ways which are likely to result in future direct fitness benefits (e.g., finding romantic partners or forming strong bonds with friends) rather than taking care of their siblings or acquiring or preparing food.
+Such a dynamic has been found among the Chuukese, with both male and female children reporting conflict with their families over children’s desires to invest financial resources in their own status enhancing opportunities rather than in the household [@lowe2003].
+Similarly, Syme and Hagen [@syme2023a] found greater numbers of siblings were associated with increased likelihoods of severe conflict in a retrospective study with Chuukese-Americans.
+This may also be consistent with parents pressuring older children to invest in the household; however, there are other ways in which limited resources can result in behavioral conflict [@syme2023a].
+
+Conflicts of interest over investment in the household also appear to extend to when a child starts having their own children [@kushnick2023; @emlen1995].
+Due to the benefits children may provide their siblings, parents may be motivated to delay or subvert the reproductive careers of certain offspring [@flinn1989; @emlen1992], a conflict which is likely enhanced in families with step-parents due to the likelihood of future children being step-siblings rather than full siblings, thereby reducing the potential fitness benefits of alloparenting [@moya2014; @schaffnit2021].
+Situations can also arise in which it might pay parents for their children to start having their own children sooner than would be ideal for the adult child.
+For example, parents may be motivated for a child to leave the household and begin having children earlier than is ideal from the child’s perspective if their presence in the household leads to them taking investment from other children who caretakers deem more worthy of investment [@kushnick2023].
+
+Studies testing the existence of familial conflict over the timing of child reproduction are rare and direct tests on the existence of parent-offspring conflict over when a child should start their own reproductive career versus help existing siblings in humans are needed [@schaffnit2021].
+The bulk of evidence for the existence of such conflicts comes from work with non-human animals. For example, the cooperative breeding white-fronted bee-eaters (*Merops bullockoides*) will interrupt the courtship of their offspring, seemingly to continue to benefit from their undiluted assistance in digging nests, incubating the clutch, finding food for nestlings, and defending the nest [@emlen1992].
+When it comes to humans, Flinn found that rural Trinidadian females contributed substantial amounts of childcare and those with siblings under 10 years old had their first child at a later age [@flinn1989], both of which would be expected if there is a conflict of interest over investing in one’s siblings or one’s own reproduction.
+However, a study among Karo Batak women found no evidence of delayed reproduction for alloparenting benefits [@kushnick2006].
+
+### Crying
+
+One possibility is that crying is physiologically constrained in ways that prevent or limit faking.
+Evidence consistent with this comes from findings that crying is hard to do on command and findings individuals often cry while alone, something which is not predicted by signaling theory [@vingerhoets2001].
+However, the likelihood of crying is often clearly related to presence of supporters, something which is seen in findings that lonely individuals cry less than those with more support options despite reporting lower well-being [@vingerhoets2013] and the existence of delayed crying—times when crying is inhibited during the cause of negative affect only for it to be expressed when one is near potential supporters [@gracanin2018].
+A more likely possibility is that certain acoustic elements of infant cries are tied to their physiological state in ways that make faking certain states impossible, with crying serving as a graded signal in most other instances [@soltis2004; @zeifman2017].
+
+It is likely that the functions of crying and the mechanisms behind its ability to elicit pro-signaler responses change over one’s life.
+Compared to adults and older children, the physiological costs of crying are greater for newborns and infants, among whom crying can result in a substantial increase in energy expenditure [@rao1997; @thureen1998].
+Due to these costs, crying at certain intensities or for long periods of time may be prohibitively costly for infants with low survival prospects, making it a potential credible signal of vigor [@furlow1997; @lummaa1998; @soltis2004; @wells2003].
+Such signals are thought to be important due to parental adaptations which are suspected to take the viability of offspring into account when making decisions about whether it is worth investing in a specific child or withdrawing investment to instead invest in other children who are more likely to survive to reproductive age and be more competitive adults [@hagen1999; @lummaa1998].
+In line with this reasoning, most infanticides cross-culturally occur when a child has poor survival prospects, due to either health conditions or resource constraints [@daly1984; @soltis2004].
+
+Another often suggested means by which crying could result in credible signals of need is if the increased risk of predation that comes from loud vocalizations allows for signaling to be worthwhile for those in relatively high need but not for those with low need [@lummaa1998; @wells2003; @zahavi1977].
+Evidence of increased predation has been found in studies on birds [@leech1997; @lima2009; @redondo1992; @ryan1982], among whom louder broods in general have also been found to be more conspicuous to predators [@haskell1997; @leech1997].
+It is unclear how significant the risk of alerting predators would be ancestrally.
+However, there are ethnographic reports of mothers keeping infants and young children close in dangerous environments [@hewlett1998; @kaplan1987; @konner1972], something which likely lessens this risk[\[MG10\]](#_msocom_10) .
+
+As children age, their reproductive value increases due to reduced odds of mortality, leading to a decreased risk of infanticide (Hagen 2004)[\[MG11\]](#_msocom_11) .
+For this reason, there is likely a transition period in which crying begins to convey information about how much a child values further investment.
+Many models have been put forward which attempt to explain human crying and begging in other species as credible signals of need due to the differential costs and benefits of crying and analogous vocalizations in other species as deterring the exaggeration of need [@godfray1991; @kilner1997; @parker2002].
+Although differing in their details, a common theme is that the indirect fitness costs signalers incur from imposing costs on that parents and siblings through begging (be them energetic costs, survival costs, or reduced likelihood of future siblings) creates a floor to possible deception but falls short of preventing it [@daly1990; @wells2003].
+
+When it comes to older children, the relatively insignificant energetic costs of crying are unlikely to be enough to result in credible signals in times of substantial conflicts of interest.
+For this reason, crying among older children may be more similar to adult crying, something which has been found to result in increased support and empathy but sometimes come at the cost of reputation and the potential for stigmatization [@cornelius2001; @cretser1982; @hendriks2006; @jesser1989; @vingerhoets2016].
+Such costs may be a means by which crying can still be a credible signal [@sznycer] with moderate conflicts of interest if they dissuade potential fakers from making the signal.

### 2024-09-04 analysisforpaper.R

@@ -55,7 +55,7 @@ mean2 <- function(..., na.rm = F){
 
 # dataframe prep ----------------------------------------------------------
 
-anthropometricMeans <- anthropometrics |>
+anthropometricMeans0 <- anthropometrics |>
   dplyr::select(
     ChildID,
     AgeAtMeasurement = Age,
@@ -68,11 +68,23 @@ anthropometricMeans <- anthropometrics |>
     .by = ChildID
   ) |>
   mutate(
-    BMI = WeightMean_KG / (HeightMean / 100 )^2
+    BMI = WeightMean_KG / (HeightMean / 100 )^2,
+    Year = year(Date.of.measurement)
   ) |>
+  relocate(measurements, .after = Year)
+
+anthropometricMeans <- anthropometricMeans0 |>
   dplyr::filter(
     measurements == 1 | year(Date.of.measurement) == 2023
   )
+
+# moved below
+# anthropometricMeansWide <- anthropometricMeans0 |>
+#   dplyr::filter(
+#     measurements == 2
+#   ) |>
+#   pivot_wider(names_from = "Year", values_from = BicepMean:BMI, id_cols = ChildID) |>
+#   left_join(d2, by = "ChildID")
 
 anthropometricMeans$WeightZ <- growthRef(AgeAtMeasurement, WeightMean_KG, Sex, anthropometricMeans, type = "Weight", pop = "CDC")
 anthropometricMeans$HeightZ <- growthRef(AgeAtMeasurement, HeightMean, Sex, anthropometricMeans, type = "Height", pop = "CDC")
@@ -89,6 +101,9 @@ anthropometricMeans$SubscapR <- residuals(m3)
 
 m4 <- mgcv::gam(BodyFatPercentageMean ~ Sex+s(AgeAtMeasurement, by = as.factor(Sex), k = 3), data = anthropometricMeans, na.action = na.exclude)
 anthropometricMeans$BodyFatPercentageR <- residuals(m4)
+
+m5 <- mgcv::gam(FlexedMean ~ Sex+s(AgeAtMeasurement, by = as.factor(Sex), k = 3), data = anthropometricMeans, na.action = na.exclude)
+anthropometricMeans$FlexedR <- residuals(m5)
 
 anthropometricMeans$Sex <- NULL
 
@@ -496,7 +511,8 @@ signal_alluvial_plot <- ggplot(sfdfsum2, aes(axis1 = Sadness, axis2 = Crying, ax
[…]

### 2024-09-06 Paper.qmd
Latest edits.

@@
 In these instances, natural selection is expected to have favored those who provided support to others if 1) such help meant one could expect to receive fitness benefits from those helped in the future and 2) the costs of providing help were outweighed by the benefits to the helper on average [@andré2010; @hamilton1964; @tooby1996; @trivers1971].
-At its simplest, this would have taken the form of mothers providing investment to their children.
-
-The fitness interdependence of offspring and at least one parent is seen in many animals [@clutton-brock1991; @webb1999].
+
+Care motivated by a high degree of fitness interdependence between offspring and at least one parent is seen in many animals [@clutton-brock1991; @webb1999].
 However, humans are a uniquely interdependent species among the living primates [@tomasello2012] in that one’s fitness prospects depend greatly on a network of kin and other social allies [@apicella2019].
-Although also true with adults, this need for others is most intense with children, who are born highly altricial and require many years of investment to develop to the point where they can obtain enough food to survive on their own and develop the social skills needed to succeed in their communities [@kaplan2009; @rosenberg2021; @schniter2015].
-Much of this investment comes from parents [@kramer2018], with most human reproduction occurring in the context of biparental investment and long-term pair bonds [@apostolou2007; @schacht2019].
-However, in many societies, other adults like grandparents, aunts, uncles, and non-relatives have also been found to provide significant investment in children [@kramer2018; @sear2008], something which was likely favored due to indirect fitness benefits, reciprocity, or reputation management [@gurven2000] + (Schniter Preprint (permission needed).
+This need for others is most intense with children, who are born highly altricial and require many years of investment to develop to the point where they can obtain enough food to survive on their own and develop the social skills needed to succeed in their communities [@kaplan2009; @rosenberg2021; @schniter2015].
+Much of this investment comes from parents [@kramer2018], However, other adults like grandparents, aunts, uncles, and non-relatives have also been found to provide significant investment in children in many societies [@kramer2018; @sear2008], something which was likely favored due to a combination of indirect fitness benefits, reciprocity, or reputation management [@gurven2000] + (Schniter Preprint (permission needed).
 Children also often serve as alloparents, helping to raise their younger siblings or provide for them through food preparation or collection [@hrdy2009; @meehan2009; @sear2008], despite still requiring substantial amounts of investment themselves [@kramer2018a].
@@
 Although the fitness prospects of both children and familial caretakers are largely intertwined, evolutionary theory predicts that disagreements about the amount of investment from parents will be ubiquitous [@fouts2005; @godfray1995; @godfray1992; @trivers1974].
-Most resources that translate to fitness benefits are limited so that one unit of a resource invested in one area cannot be used in another.
+Most resources relevant to fitness outcomes are limited so that one unit of a resource invested in one area cannot be used in another.
 In response to this basic fact of life, humans, like other organisms, are expected to have evolved to efficiently allocate resources between somatic effort (i.e., growth and maintenance) and reproductive effort (i.e., mating and parenting effort) [@daly1983], something which often pits the fitness interests of parents against the fitness interests of offspring.
@@
 Conflicts of interest between children and parents may also involve how much children invest in the household.
-As previously mentioned, each individual child is expected to prefer greater investment in oneself than in one’s siblings, while parents, who are equally related to all their children, should often be expected to favor a more even distribution [@trivers1974].
+As previously mentioned, each child is expected to prefer greater investment in oneself than in one’s siblings, while parents, who are equally related to all their children, should often be expected to favor a more even distribution [@trivers1974].
 This can lead to conflict with older children who might prefer to spend their time in ways which are likely to result in future direct fitness benefits (e.g., finding romantic partners or forming strong bonds with friends) rather than taking care of their siblings or acquiring or preparing food [@lowe2003; @syme2023a].
@@
 Although the risks of being deceived can be substantial, signal receivers may still benefit from paying attention to signals so long as the information is credible enough to inform one’s behavior in ways that are adaptive on average [@bliegebird2001; @grafen1990].
-These benefits coupled with the likelihood of strong selection against individuals who respond to fake signals results in the expectation that signals within species will often need to benefit both signaler and receiver to persist evolutionarily [@hasson1994; @maynard-smith2004].
+These benefits coupled with the likelihood of strong selection against individuals who respond to fake signals results in the expectation that signals within species will often need to benefit both signaler and receiver to persist evolutionary [@hasson1994; @maynard-smith2004].
 
@@
 
-Regardless of the mechanism behind the credibility of a signal, credible signals work by revealing private information about one’s qualities to others.
+Credible signals work by revealing private information about one’s qualities to others.
 In the context of child signals of need, a child has access to information about their internal state, which is unavailable to others, that allows them to estimate their valuation of additional investment.
 By signaling in a credible way, the child reveals their resource valuation to potential caretakers who may be skeptical of the child’s claims due to the common conflicts of interest described above.
-With the new information revealed by the signal, caretakers should generally be more likely to provide additional support due to their fitness interest in the child’s wellbeing.
-When this is the case, all parties are better off than they would be without the signal (so long as the benefits outweigh the costs of the signal to child and parent), with caretakers benefitting from the revelation of fitness-relevant information and the child benefitting from the increased likelihood of a beneficial response.
+With the information revealed by the signal, caretakers should generally be more likely to provide additional support due to their fitness interest in the child’s wellbeing.
+When this is the case, all parties are better off than they would be without the signal, with caretakers benefitting from the revelation of fitness-relevant information and the child benefitting from the increased likelihood of a beneficial response as long as the benefits outweigh the costs of the signal to child and parent.
 
@@
 
-Potentially combine and condense with paragraph below: As children age, their reproductive value increases due to reduced odds of mortality, leading to a decreased risk of infanticide (Hagen 2004).
+Potentially combine and condense with paragraph below: As children age, their reproductive value increases due to reduced odds of mortality, leading to a decreased risk of infanticide (Hagen 2004 comment).
 For this reason, there is likely a transition period in which crying is primarily meant to provide information about how much a child values further investment.
@@
 The survey data for this study come from a larger project run by the Utila Child Health Project.
-Data was collected by Jess Hlay, Izabel Rodriguez, Daniela Mejia, Angela Garcia, or Caroline Smith who read the survey questions to participants and recorded their answers.
+Data was collected by either Jess Hlay, Izabel Rodriguez, Daniela Mejia, Angela Garcia, or Caroline Smith who read the survey questions to participants and recorded their answers.
 Participants could choose to do the survey in English or Spanish.
@@
 We inverted the scores on two measures so that a score of 5 corresponded with the lowest susceptibility to illness before we averaged the scores of these to create the CurrentHealthMean variable.
-In the `r sum((!is.na(modeldf$Currenthealth2)) & is.na(modeldf$Currenthealth3))` instance in which we only data for 2 of the measures, we still calculated the mean a numerator of 2.
+In the `r sum((!is.na(modeldf$Currenthealth2)) & is.na(modeldf$Currenthealth3))` instance in which we only data for 2 of the measures, we still calculated the mean but used a numerator of 2.
 See Table @tbl-healthvars for the wording of these variables.
@@
 The survey also included many demographic questions about the survey taker and their households.
-Those used for this study included the (1) ages of those in the household, (2) sexes of those in the household, (3) partnership status of the primary caregiver, (4) number of adults who provide allocare within the household, (5) highest level of education of the primary caregiver, (5) immigration status (in regards to Utila), and (6) monthly household income.
+Those used for this study included the (1) ages of those in the household, (2) sexes of those in the household, (3) partnership status of the primary caregiver, (4) number of adults who provide childcare within the household, (5) highest level of education of the primary caregiver, (5) immigration status (in regards to Utila), and (6) monthly household income.
 This data allowed us to calculate the number of children (total, younger, and older) and adults within the household.
@@
 
-To model the effects of skinfold thickness, flexed bicep circumference, and grip strength on signaling behavior, we used the residuals from generalized additive models which modeled the effect of sex and age on each measurement using the mgcv package [@wood2011].
-This resulted in variables which represented how each child differed from the average child for their given sex and age.
+For our independent variables of skinfold thickness, flexed bicep circumference, and grip strength on signaling behavior, we used the residuals from generalized additive models which modeled the effect of sex and age on each measurement using the mgcv package [@wood2011].
+The resulting variables which represent how each child differed from the average child for their given sex and age.
+
+Summed body fat
 
@@
 
-All statistical analyses were exploratory and involved fitting multiple models while being cognizant of the risk of overfitting the data, but with the payoff of idenftifying theoretically relavent effecrs that should be tested in the future.
+All statistical analyses were exploratory and involved fitting many models with different combinations of independent and dependent variables while being cognizant of the risk of overfitting the data.
+The payoff to this is the identification of theoretically relevant effects that should be tested in the future.
 
 We fit generalized linear mixed models for exploratory analyses with signal frequencies and summed signal cost as dependent variables.
-As our signaling variables invovled over-dispersed count data, we used negative binomial regression.
+As our signaling variables involved over-dispersed count data, we used negative binomial regression.
 This was done in R with the GLMM package [@brooks2017], with household as a random effect.
 Control variables differed in some models, but always include child age and sex as controls.
-Specific control variables are mentioned in the results
+Specific control variables are mentioned in the results.
 
@@
 `r nrow(modeldf)` children had frequency data for at least 1 signal, with `r nrow(modeldf_FULLSIG)` having data for all 3.
-Among those with at least one record of signaling frequency `r sex_table[1]` were female and `r sex_table[2]` were male.
+Among those with at least one record of signaling frequency, `r sex_table[1]` were female and `r sex_table[2]` were male.
 Anthropometric and signaling data exist for `r nrow(modeldf_FULLANTHROPOMETRICS)` children, of whom `r sex_table2[1]` were female and `r sex_table2[2]` were male.
@@
   BMIZ = "BMI Z-score (CDC)",
-  SubscapMean = "Subscapular skinfold thickness",
+  SubscapMean = "Subscapular skinfold thickness (mm)",
   SubscapR = "Subscapular skinfold residuals",
-  TricepMean = "Tricep skinfold thickness",
+  TricepMean = "Tricep skinfold thickness (mm)",
   TricepR = "Tricep skinfold residuals",
@@
 Stratum height equals the count of individuals at a given frequency.
-Frequency key: Low includes "never" and "once a month or less," mediums includes "more than once a month but less than once a week" and "more than once a week but not daily," high includes "daily," and "multiple times per day."
+Frequency key: Low includes "never" and "once a month or less," medium includes "more than once a month but less than once a week" and "more than once a week but not daily," high includes "daily," and "multiple times per day."
 :::
@@
 
-The effect of age on specific signal frequencies (per month), summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior.
+The effect of age on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior, all per month.
 Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
-Each dot is 1 child.
-Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate).
+Rug corresponds to actual values of child age with a small amount of jitter to reveal overlapped values.
 See @tbl-main_models for the corresponding regression table.
@@
 
-The effect of the frequency of conflict with one's primary caretaker (per month) on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior.
+The effect of the frequency of conflict with one's primary caretaker on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior, all per month.
 Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
-Each dot is 1 child.
-Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate).
+Rug corresponds to actual values of the frequency of conflict between child and primary caretaker with a small amount of jitter to reveal overlapped values.
 See @tbl-main_models for the corresponding regression table.
@@
 
-The effect of years of caregiver education on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior.
+The effect of years of caregiver education on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior, all per month.
 Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
+Rug corresponds to actual values of primary caretaker education (years) with a small amount of jitter to reveal overlapped values.
 See @tbl-main_models for the corresponding regression table.
@@
 Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
-Each dot is 1 child.
-Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate).
+Rug corresponds to actual values of A) child age and B) the number of adults in the household that contribute to childcare with a small amount of jitter to reveal overlapped values.
 See @tbl-conflict_dep_var for the corresponding regression table.
@@
 
-The effect of an alloparenting effort by sex interaction on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior.
+The effect of an alloparenting effort by sex interaction on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior, all per month.
 Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
@@
 
-## Positive assoication between child alloparenting effort and number of older children
+## Positive association between child alloparenting effort and number of older children
 
@@
 ```{r}
-signaling_data_plot_bodyfat
+signaling_data_plot_bodyfat2
 ```
@@
 
-Plots showing the likelihood of child being more needy, as needy, or less needy than a caretaker's other children.
+The likelihood of child being more needy, as needy, or less needy than a caretaker's other children.
 A: Effect of child age on relative neediness B: Effect of summed signaling frequency on relative neediness.
@@
 
-Plots showing the likelihood of positive, neutral, and negative responses to the last time a child became sad or cried.
+The likelihood of positive, neutral, and negative responses to the last time a child became sad or cried.
 A: Effect of punishment being implicated in the cause (0 = no punishment, 1 = punishment.) B: Effect of discomfort, pain, injury, or illness being implicated in the cause (0 = discomfort, pain, injury, or illness, 1 = discomfort, pain, injury, or illness.) Legend key: -1 = negative response, 0 = neutral response, 1 = positive response (all from child's perspective).
@@
 As seen in Western cultures [@zeifman2013; @potegal2003; @koch2003] (check for others), the frequencies of crying and temper tantrums decreased with age, something which is likely due to numerous reasons.
-One possibility is that older children have greater ability to satisfy one's desires and behave in ways which match cultural expectations of older children and adults.
-Another is that as children survive to the point where they would have been unlikely to die young in ancestral environments, the intensity conflicts of interest with parents decrease due to children passing this important milestone and requiring less investment to reach adulthood than younger children, current or future (cite daly and wilson).
+One possibility is that older children have greater ability to satisfy one's desires and behave in ways which match cultural expectations of teenagers and adults.
+Another is that since one's reproductive value increases as one survives periods of high mortality risk in childhood, the intensity conflicts of interest with parents decreases since these children are likely to reach adulthood and require less investment to reach adulthood than younger children, current or future [@daly1995].
 For this reason, less costly signals may be more likely to work to increase investment, all else being equal\[\[4\]\].
+That said, all else may often not be equal if investment in older children is costlier, something which may be reflected in caretakers seeing older children as relatively needy.
 The decrease in crying and tantrum frequencies with age may also result from the potential for greater reputational costs, something which may also increase the credibility of such signals among older children as previously mentioned.
-It is also possible that parents see older children less than younger children.
-However, the negative assocition between need with age observed in other societies and the fact that all children in this study were of school age suggest that this is unlikely to be the sole reason responsible for this.
+~~It is also possible that parents see older children less than younger children.
+However, the negative association between need with age observed in other societies and the fact that all children in this study were of school age suggest that this is unlikely to be the sole reason responsible for this.~~
 
 There was no significant association between the frequency of child sadness and age.
-Part of this likely results from the fact that, unlike crying, sadness is thought to have important non-signaling functions including drawing one’s attention to a problem, motivating deeper analysis of one’s situation, and causing withdrawal from situations with low fitness prospects (cite), from which, others can still obtain cue-based information about one's emotional state, well-being, and desires.
+Part of this likely results from the fact that, unlike crying, sadness is thought to have important non-signaling functions including drawing one’s attention to a problem, motivating deeper analysis of one’s situation, and causing withdrawal from situations with low fitness prospects [@nesse2019; @nettle2009; @allen2003; @hagen2011], from which, others can still obtain cue-based information about one's emotional state, well-being, and desires.
 By asking about sadness as a whole, our measure was unable to disentangle information caretakers gleaned from cues from information that came from signals.
@@
 
--   add in depression over the lifecourse
+-   add in depression over the lifecourse?
 
@@
 
-Probably cut When viewed in isolation, the frequencies of sadness, crying, and temper tantrums were also positively associated with conflict frequency between child and caretaker.
-However, the frequency of running away was not significantly associated with conflicts with parents.
-A retrospective study of parent-offspring conflict among Chon Chuuk immigrants in the northwest United States found that running away was the most common response to long-standing conflicts with parents and often resulted in beneficial outcomes for the child [@syme2023a].
-The threat running away poses to child wellbeing and, therefore, its ability to credibly signal information about one’s current situation is likely highly dependent on one’s socioecology and age.
-Among the Chon Chuuk, running away is often seen as a suicide threat [@fischer1950; @syme2023a], likely due to xxxxxxxxx.
-Similar risks may also occur in environments with large predators, venomous animals, or high degrees of human conflict.
-In these instances, even one night away can involve substantial risk as is seen among the !Kung where a woman running away and spending the night outside of camp can be used to communicate “the strength of her feelings” about a an unhappy relationship [@shostak1981]\[\[6\]\].
-When it comes to Utila, mothers in focus groups reported that running away on Utila was much less risky than it would be on mainland Honduras, with a few mothers mentioning that they warned their children about the difference in danger that came with running away on the mainland.
-Despite the relative safety of Utila, young children may still face risks out on their own due to proximity to the ocean, large dogs, and fast-moving motorcycles, golf carts, and tuk-tuks on the main road.
-For these reasons, running away may be too costly for young children but not costly enough to credibly signal need for older children.
-
 *Sex and focal child alloparenting effort interact (positive for females, negative/flat for males)*
@@
 We found evidence of an interaction between how often a child cared for their siblings and the sex of the child in our models examining signal frequencies and cost, whereby there was a positive relationship between the frequency of and summed costliness of child signals of need and alloparenting effort for girls but no relationship or a slightly negative one for boys.
-In many societies, girls are often expected to do more alloparenting within the household and often end up doing so (cite).
+In many societies, girls are often expected to do more alloparenting within the household and often end up doing so [@helfrecht2016 @weisner1977; @barry1971; @turke1988].
 While girls in this sample were also reported to give more allocare, the difference was not significant.
@@
 Although the body fat by sex interaction was only marginally significant, our models predict a negative relationship between body fat and signaling for girls and no relationship for boys.
-Female reproduction is highly dependent on body fat, with conception requiring about xx% body fat and mothers investing a significant amount of calories in both children in the womb and breastfed infants (cite).
-For these reasons, low body fat is likely a be better predictor of future reproductive success in older girls than boys, potentially making signaling for increased investment a more viable strategy in instances when an individual has low body fat.
+Female reproduction is highly dependent on body fat, with conception requiring about xx% body fat and mothers investing a significant amount of calories in both children in the womb and breastfed infants [@jasienska2009].
+For these reasons, low body fat is likely a be better predictor of future reproductive success in adolescent (and older) girls than boys, potentially making signaling for increased investment a more viable strategy in instances when an individual has low body fat.
 
@@
 a skinned knee).
-It may also be that many of the typical positive responses to pain and illness are not very costly (e.g., providing comfort).
+It may also be that many of the typical positive responses to pain and illness are not very costly to caretakers (e.g., providing comfort), thereby increasing their frequency.
 
 When the cause involved punishment, however, parents were more likely to respond negatively and less likely to respond positively.
-In these circumstances, children may be signaling about the severity of a punishment (cite).
-However, parents may be reluctant to reduce punishments if the behavior that warranted the punsishment was repeated or severe.
+In these circumstances, children may be signaling about the severity of a punishment (cite syme costly apology).
+However, parents may be reluctant to reduce punishments if the behavior that warranted the punishment was repeated or severe.
 They may also react with scolding or further punishment.
 That said, our model predicts positive responses in xxxxx cases, suggesting that children are successful in negotiating better outcomes at least some of the time.
-- costly apolgy model?
+- costly apology model?
 
@@
 
-*CUT Negative association between caretaker estimates of child health and child signaling for some measures. This does not extend to the costlier signals.*
-
-A negative association between caretaker estimates of child immune function and child signaling was seen, suggesting that signaling behavior was influenced by adversity.
-Perceptions of greater immune susceptibility were associated with both the summed costliness of signaling and the frequency of the most common child signal of need, while the frequency of child signaling was marginally significant.
-When examining the effect of perceived immune function on each signal, perceptions of greater immune susceptibility predicted the frequency of sadness but not the frequencies of crying and temper tantrums.
-
-One explanation for this may be that parents can easily detect child illness and that conflicts of interest tend to be less significant when it comes to illness care than other areas.
-Pathogens have been a strong selection pressure for life on earth for much of its existence and have greatly shaped evolutionary history (cite).
-In response to this, humans have evolved many adaptations for limiting the harm we experience from pathogens, including the ability to identify relaible indicators of illness in others.
-In some contexts, this activates disgust and motivates avoidance of sick individuals.
-However, it can also lead to increased care of relatives and allies (cite), with there being some suggestion that we have even evolved to upregulate some of our symptoms beyond what would arise from an immune response/pathogen interaction to better convince others that we are sick (cite).
-Due to this evolutionary history, parents may be better equipped to identify the possible benefits to investing in a sick child when compared to investment aimed at helping a child in status competition (e.g., buying new or popular clothing), something which may be exacerbated by culture change in what is popular among children.
-In other words, there is less private information for a child to reveal.
-This coupled with the likelihood that conflicts of interest between children and caretakers are smaller when it comes to illness care compared to other areas means there is less to gain from costlier signals.
-
 *Lack of evidence for demographic effects*
@@
 Monthly income was also not found to be a significant predictor of signaling behavior.
-Many models designed to explain signals of need (e.g., the begging of young birds or withdrawal in humans) predict that offspring who are worse off or are receiving less investment have more to gain from signaling than those with more (cite).
+Many models designed to explain signals of need (e.g., the begging of young birds or withdrawal in humans) predict that offspring who are worse off or are receiving less investment have more to gain from signaling than those with more [@mock2011; @godfray1991; @caro2016a; @godfray2000].
 However, it is possible that this can be true and household level resources have little influence on child signaling frequencies between families.
 As signaling behavior is costly to child, caretakers, and siblings, children may gain direct and indirect fitness benefits from limiting their signaling behavior in times of resource stress (i.e., when the chances of gaining investment or the size of investment are small).
-Similarly, children in high resource families may also be motivated to signal at higher than expected frequencies due to the potential for larger pieces of investment.
+Similarly, children in high resource families may also be motivated to signal at high frequencies due to the potential for larger pieces of investment.
 We know of no empirical explorations of this in humans (check more and confirm).
-However, multiple game theoretical models of sibling competition and have found that parental resources have a small impact on offspring signaling, especially with high rates of monogamy \[Bossan + Mock and Parker + Check for newer sources\].
-Although counterintuitive, such an outcome fits the fact that fitness is relative and that natural selection should be expected to have favored those who maximized their own inclusive fitness, whether one found themselves in a great situation or a dire one.
+However, multiple game theoretical models examining parent-offspring conflict have found that parental resources have a modest impact on the evolutionary stable value for offspring demand for resources [@bossan2013; @mock1997 [confirm]].
+~~Although counterintuitive, such an outcome fits the fact that fitness is relative and that natural selection should be expected to have favored those who maximized their own inclusive fitness, whether one found themselves in a great situation or a dire one.~~ Compared to overall effort in eliciting need, resource availability may have a larger impact on how children signal for more investment based on the finding that birds in stable and unusually favorable environments are more likely to compete with siblings through signals thought to reveal information about need, while those in relatively unpredictable and unusually poor environments have been found to compete more through signals thought to reveal information about quality [@caro2016a].
 
@@
 
-Another limitation is that our study was cross-sectional.
+Our study was also predominantly cross-sectional, which limis our ability to make causal inferences .
 A major challenge when studying signals of need is the fact that such signaling is expected to 1) be more common in times when individuals can benefit from increased investment (e.g., when they are facing major adversity or when a fitness-relevant opportunity arises) and 2) result in benefits to the signaler on average.
@@
 It is unclear how significant the risk of alerting predators would be ancestrally.
-However, there are ethnographic reports of mothers keeping infants and young children close in dangerous environments [@hewlett1998; @kaplan1987; @konner1972], something which likely lessens this risk[\[MG10\]](#_msocom_10) .
-
-As children age, their reproductive value increases due to reduced odds of mortality, leading to a decreased risk of infanticide (Hagen 2004)[\[MG11\]](#_msocom_11) .
+However, there are ethnographic reports of mothers keeping infants and young children close in dangerous environments [@hewlett1998; @kaplan1987; @konner1972], something which likely lessens this risk.
+
+As children age, their reproductive value increases due to reduced odds of mortality, leading to a decreased risk of infanticide (Hagen 2004).
 For this reason, there is likely a transition period in which crying begins to convey information about how much a child values further investment.
@@
 Such costs may be a means by which crying can still be a credible signal [@sznycer] with moderate conflicts of interest if they dissuade potential fakers from making the signal.
+
+## Discussion
+
+### Running away
+
+Probably cut When viewed in isolation, the frequencies of sadness, crying, and temper tantrums were also positively associated with conflict frequency between child and caretaker.
+However, the frequency of running away was not significantly associated with conflicts with parents.
+A retrospective study of parent-offspring conflict among Chon Chuuk immigrants in the northwest United States found that running away was the most common response to long-standing conflicts with parents and often resulted in beneficial outcomes for the child [@syme2023a].
+The threat running away poses to child wellbeing and, therefore, its ability to credibly signal information about one’s current situation is likely highly dependent on one’s socioecology and age.
+Among the Chon Chuuk, running away is often seen as a suicide threat [@fischer1950; @syme2023a], likely due to xxxxxxxxx.
+Similar risks may also occur in environments with large predators, venomous animals, or high degrees of human conflict.
+In these instances, even one night away can involve substantial risk as is seen among the !Kung where a woman running away and spending the night outside of camp can be used to communicate “the strength of her feelings” about a an unhappy relationship [@shostak1981]\[\[6\]\].
+When it comes to Utila, mothers in focus groups reported that running away on Utila was much less risky than it would be on mainland Honduras, with a few mothers mentioning that they warned their children about the difference in danger that came with running away on the mainland.
+Despite the relative safety of Utila, young children may still face risks out on their own due to proximity to the ocean, large dogs, and fast-moving motorcycles, golf carts, and tuk-tuks on the main road.
+For these reasons, running away may be too costly for young children but not costly enough to credibly signal need for older children.
+
+### Perceived immune function
+
+A negative association between caretaker estimates of child immune function and child signaling was seen, suggesting that signaling behavior was influenced by adversity.
+Perceptions of greater immune susceptibility were associated with both the summed costliness of signaling and the frequency of the most common child signal of need, while the frequency of child signaling was marginally significant.
+When examining the effect of perceived immune function on each signal, perceptions of greater immune susceptibility predicted the frequency of sadness but not the frequencies of crying and temper tantrums.
+
+One explanation for this may be that parents can easily detect child illness and that conflicts of interest tend to be less significant when it comes to illness care than other areas.
+Pathogens have been a strong selection pressure for life on earth for much of its existence and have greatly shaped evolutionary history (cite).
+In response to this, humans have evolved many adaptations for limiting the harm we experience from pathogens, including the ability to identify relaible indicators of illness in others.
+In some contexts, this activates disgust and motivates avoidance of sick individuals.
+However, it can also lead to increased care of relatives and allies (cite), with there being some suggestion that we have even evolved to upregulate some of our symptoms beyond what would arise from an immune response/pathogen interaction to better convince others that we are sick (cite).
+Due to this evolutionary history, parents may be better equipped to identify the possible benefits to investing in a sick child when compared to investment aimed at helping a child in status competition (e.g., buying new or popular clothing), something which may be exacerbated by culture change in what is popular among children.
+In other words, there is less private information for a child to reveal.
+This coupled with the likelihood that conflicts of interest between children and caretakers are smaller when it comes to illness care compared to other areas means there is less to gain from costlier signals.
+
+### Cross-sectional study
+Another limitation is that our study was cross-sectional.
+A major challenge when studying signals of need is the fact that such signaling is expected to 1) be more common in times when individuals can benefit from increased investment (e.g., when they are facing major adversity or when a fitness-relevant opportunity arises) and 2) result in benefits to the signaler on average.
+In other words, higher levels of signaling may stem from greater levels of need while simultaneously eliciting more investment from others.
+This hypothesized causal, bi-directional relationship makes it difficult to isolate the effects of signaling behavior on investment in children when using standard measures of child welfare (i.e., BMI, health metrics, and height for age).
+Longitudinal studies can help address this.
+However, our small sample size of participants with signaling data and anthropometric data at two time periods prevents this for this study.

### 2024-09-06 analysisforpaper.R

@@ -784,6 +784,22 @@ m2 <- glmmTMB(WeightMean_KG_2024 ~ WeightMean_KG_2023 + CryFreqN*ChildAge + (1|h
 summary(m2)
 plot(allEffects(m2))
 
+p <- plot_predictions(m2, condition = c("CryFreqN", "ChildAge"), vcov = TRUE) +
+  geom_rug(data = m2$frame, aes(x = CryFreqN , y = 0), position = position_jitter(width = 1, seed = 123), sides = "b") +
+  scale_color_viridis_d(option = "A", end = .8) +
+  guides(color = guide_legend(reverse = TRUE, title = "Child age (years)"), fill = guide_legend(reverse = TRUE, title = "Child age (years)")) +
+  xlab("Crying freq. (per month)") +
+  ylab("Weight (2024)") +
+  ylim(40, 55) +
+  theme_minimal(15)
+p$layers[[2]]$aes_params$linewidth <- 2
+p$layers[[1]]$aes_params$alpha <- 0
+p
+
+# causality going the other direction (weight causes more crying?)
+
+out <- avg_comparisons(m2, variables = list("CryFreqN" = c(0, 30), "ChildAge" = c(5,14)), cross = TRUE)
+out <- avg_comparisons(m2, variables = list("CryFreqN" = c(0, 8), "ChildAge" = c(5,14)), cross = TRUE)
 # ConflictFreqN model -----------------------------------------------------------
 
 d2$AdultsNoChildcare <- d2$number_adults - d2$AdultsChildcare
@@ -954,6 +970,8 @@ mN1 <- polr(RelativeNeed3 ~ SignalFreq + ChildAge + Sex + OtherChildrenHH + LogI
 summary(mN1)
 coeftest(mN1, vcov = vcovCL, type = "HC0", cluster = ~householdID)
 
+
+
 p_need_age_sf <- plot_predictions(mN1, condition = c("ChildAge", "group"), type = "probs", vcov = ~householdID) +
   ylab("Relative need") +
   xlab("Child age (years)") +
@@ -1023,6 +1041,10 @@ table(modeldf$PositiveResponse, modeldf$NegativeResponse)
 
 mpr <- polr(CaregiverResponse ~ ChildAge + Sex + OtherChildrenHH + LogIncome + number_adults + UnwantedTask + Punishment2 + Family2 + DiscomfortPainInjuryIllness, d2)
 summary(mpr)
[…]

### 2024-09-09 Paper.qmd
Paper edits and addition of code to plot effects of crying on weight and height gain.

@@
 This hypothisized causal, bi-directional relationship between investment and signaling makes it difficult to isolate the effects of signaling behavior on investment in children when using standard measures of child welfare (i.e., BMI, health metrics, and height for age).
-When experiments are not appropriate, longitudinal studies with carefully selected variables can help address this (cite).
+When experiments are not appropriate, longitudinal studies with carefully selected variables can help address this [@arjas2004].
 However, our small sample size with signaling data and anthropometric data at two time periods limits our ability to do this in this study.
@@
 
-: Table showing variable names and question wording for measures about the frequency of behaviors.
+: Variable names and question wording for measures about the frequency of behaviors.
 All questions had the following labels attached to each point on its Likert scale: Never; Once a month or less; More than once a month but less than once a week; More than once a week but not daily; Daily; Multiple times per day.
@@
 
-: Table showing conversions from Likert scale to count value for the frequencies of child behaviors included in this study.
-Variables created with this scheme include SadFreqN, CryFreqN, TantrumFreqN, RunawayFreqN, ConflictFreqN, and AlloparentingFreqN.
-{#tbl-convertedfreqs}
+: The conversions from Likert scale to count value for the frequencies of child behaviors included in this study.
+Variables created with this scheme include the frequencies of sadness, crying, temper tantrums, conflict with primary caretakers, and child alloparenting effort.
 
@@
 
-: Table showing variable names and question wording for measures about relative need and investment.
+: Variable names and question wording for measures about relative need and investment.
 Both questions had the following labels attached to each point on its Likert scale: Much less; Less; The same amount; More; Much more.
@@
 
-### Perceptions of child immune function (or current susceptibility to illness)
+### Perceptions of child susceptibility to illness (or current susceptibility to illness)
 
 Caretakers were also asked three questions using a five point Likert scale about their child's susceptibility to illness.
-We inverted the scores on two measures so that a score of 5 corresponded with the lowest susceptibility to illness before we averaged the scores of these to create the CurrentHealthMean variable.
-In the `r sum((!is.na(modeldf$Currenthealth2)) & is.na(modeldf$Currenthealth3))` instance in which we only data for 2 of the measures, we still calculated the mean but used a numerator of 2.
+We inverted the scores on two measures so that a score of 5 corresponded with the lowest susceptibility to illness before we averaged the scores of these to create a mean illness susceptibility variable.
+In the `r sum((!is.na(modeldf$IllnessSusceptibility2)) & is.na(modeldf$IllnessSusceptibility3))` instance in which we only data for 2 of the measures, we still calculated the mean but used a numerator of 2.
 See Table @tbl-healthvars for the wording of these variables.
@@
 | Variable               | Survey question wording                                                               |
-| Currenthealth1 | If an illness is 'going around', my child will get it.                                |
-| Currenthealth2 | In general, my child is very susceptible to colds, flu, and other infectious diseases |
-| Currenthealth3 | My child’s immune system protects him/her from most illnesses that other kids get.    |
-
-: Table showing variable names and question wording for measures about a child's current likelihood of becoming sick.
-All three questions were asked using a Likert scale from 1-5 where 1 was labeled "Strongly disagree" and 5 was labeled "Strongly agree." The scores for Currenthealth1 and Currenthealth2 were inverted so that 5 corresponded with the lowest susceptibility to illness.
+| IllnessSusceptibility1 | If an illness is 'going around', my child will get it.                                |
+| IllnessSusceptibility2 | In general, my child is very susceptible to colds, flu, and other infectious diseases |
+| IllnessSusceptibility3 | My child’s immune system protects him/her from most illnesses that other kids get.    |
+
+: Variable names and question wording for measures about a child's current likelihood of becoming sick.
+All three questions were asked using a Likert scale from 1-5 where 1 was labeled "Strongly disagree" and 5 was labeled "Strongly agree." The scores for IllnessSusceptibility1 and IllnessSusceptibility2 were inverted so that 5 corresponded with the lowest susceptibility to illness.
 {#tbl-healthvars}
@@
 The subscapular measurement occured immediately below the inferior angle of the scapula, being picked up diagonally at the natural fold line of the skin [@antóna].
-Add flexed bicep measurementxxxxxx.
+Flexed bicep circumference was measured perpendicular to the long axis of the upper arm.
+For this measure participants were asked to sit/stand?
+an position their upper arm perpendicular to their body with their elbow being help at a 90° angle, before flexing their clenched fist towards themselves as much as they could.
 
@@
 The resulting variables which represent how each child differed from the average child for their given sex and age.
-
-Summed body fat
+We also centered and scaled the skinfold thickness residuals before summing them to create an overall body fat variable.
 
@@
 
-All statistical analyses were exploratory and involved fitting many models with different combinations of independent and dependent variables while being cognizant of the risk of overfitting the data.
-The payoff to this is the identification of theoretically relevant effects that should be tested in the future.
+All statistical analyses were exploratory and involved fitting many models with different combinations of independent and dependent variables.
+While were cognizant of the risk of overfitting the data, it is still more likely than if the study was entirely confirmatory.
+The payoff to this approach is the identification of theoretically relevant effects that should be tested in the future.
 
@@
 This was done in R with the GLMM package [@brooks2017], with household as a random effect.
-Control variables differed in some models, but always include child age and sex as controls.
+Control variables differed in some models, but child age and sex were always included.
 Specific control variables are mentioned in the results.
@@
 The results were then interpreted with clustered standard errors for household.
-Due to the risk of overfitting that comes with having too few clusters [@cameron2015], we also fit the same models using BRMS with household as a random effect.
+Due to the risk of overfitting that comes with having too few clusters [@cameron2015], we also fit the same models using BRMS [@bürkner2017] with household as a random effect.
 
@@
   SignalFreqMax = "Frequency of most common signal (times per month)",
-  Currenthealth1 = "Currenthealth1",
-  Currenthealth2 = "Currenthealth2",
-  Currenthealth3 = "Currenthealth3",
-  CurrentHealthMean = "Mean of current health measures",
+  IllnessSusceptibility1 = "IllnessSusceptibility1",
+  IllnessSusceptibility2 = "IllnessSusceptibility2",
+  IllnessSusceptibility3 = "IllnessSusceptibility3",
+  IllnessSusceptibilityMean = "Mean of current health measures",
   HeightZ = "Height Z-score (CDC)",
@@
 
-Child sadness, crying, and temper tantrums were reasonably common.
-Caregivers reported that (`r sum(na.omit(modeldf$SadFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$SadFreqN >= 8)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\]) children were sad at least once a week, (`r sum(na.omit(modeldf$CryFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$CryFreqN >= 8)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\]) cried at least more than once a week, and (`r sum(na.omit(modeldf$TantrumFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$TantrumFreqN >= 8)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]) threw temper tantrums at least once a week.
+Child sadness, crying, and temper tantrums were common.
+Caregivers reported that `r sum(na.omit(modeldf$SadFreqN >= 8))` (`r signif(sum((na.omit(modeldf$SadFreqN >= 8)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%) children were sad at least once a week, `r sum(na.omit(modeldf$CryFreqN >= 8))` (`r signif(sum((na.omit(modeldf$CryFreqN >= 8)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%) cried at least more than once a week, and `r sum(na.omit(modeldf$TantrumFreqN >= 8))` (`r signif(sum((na.omit(modeldf$TantrumFreqN >= 8)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%) threw temper tantrums at least once a week.
 A small subset of children engaged in these signals/states at least daily (sadness = `r sum(na.omit(modeldf$SadFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$SadFreqN >= 30)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\], crying = `r sum(na.omit(modeldf$CryFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$CryFreqN >= 30)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\], and tantrums = `r sum(na.omit(modeldf$TantrumFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$TantrumFreqN >= 30)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]).
@@
 
-For every signaling measure except sadness (which also involves non-signaling functions), there was a statistically significant negative relationship between the age of the child and the frequency of signaling when controlling for the sex of the child, household demographics, the frequency of conflict with one's primary caretaker, caretaker perceptions of child immune function, and an alloparenting effort by sex interaction.
+For every signaling measure except sadness (which also involves non-signaling functions), there was a statistically significant negative relationship between the age of the child and the frequency of signaling when controlling for the sex of the child, household demographics, the frequency of conflict with one's primary caretaker, caretaker perceptions of child susceptibility to illness, and an alloparenting effort by sex interaction.
 See @tbl-main_models for regression statistics and @fig-effect_age for effects plots.
@@
   SignalFreqMax = "Max signal freq.",
-  CurrentHealthMean = "Percieved immunity",
+  IllnessSusceptibilityMean = "Percieved illness susceptibility",
   SadFreqN = "Sadness freq.",
@@
 
-There were statistically significant positive relationships between the frequency of conflict between a child and their primary caretaker and all of our signaling measures when controlling for the sex of the child, the age of the child, household demographics, caretaker perceptions of child immune function, and an alloparenting effort by sex interaction.
+There were statistically significant positive relationships between the frequency of conflict between a child and their primary caretaker and all of our signaling measures when controlling for the sex of the child, the age of the child, household demographics, caretaker perceptions of child susceptibility to illness, and an alloparenting effort by sex interaction.
 See @tbl-main_models for regression statistics @fig-effect_conflict for effects plots.
@@
 
-There were statistically significant negative relationships between years of caregiver education and all of our signaling measures except sadness (`r stats$msadH$EducationLevelYears$str`) when controlling for the sex of the child, the age of the child, household demographics, caretaker perceptions of child immune function, and an alloparenting effort by sex interaction.
+There were statistically significant negative relationships between years of caregiver education and all of our signaling measures except sadness (`r stats$msadH$EducationLevelYears$str`) when controlling for the sex of the child, the age of the child, household demographics, caretaker perceptions of child susceptibility to illness, and an alloparenting effort by sex interaction.
 See @tbl-main_models for regression statistics @fig-effect_edu for effects plots.
@@
 
+For each signaling measure, there was also a significant interaction between how often a child cared for their siblings and the sex of the child when controlling for the age of the child, the sex of the child, household demographics, the frequency of conflict with one's primary caretaker, and caretaker perceptions of child susceptibility to illness.
+The slope of these relationships were positive for girls and flat or slightly negative for boys.
+In other words, our models predict that as children provide more allocare within their families, the frequency of signaling would increase with alloparenting effort for girls and remain at the same rate or decrease slightly with alloparenting effort for boys.
+See @tbl-main_models for the corresponding regression table and @fig-effect_allo_by_sex for effects plots.
+
 ::: {#fig-effect_allo_by_sex}
@@
 :::
-
-For each signaling measure, there was also a significant interaction between how often a child cared for their siblings and the sex of the child when controlling for the age of the child, the sex of the child, household demographics, the frequency of conflict with one's primary caretaker, and caretaker perceptions of child immune function.
-The slope of these relationships were positive for girls and flat or slightly negative for boys.
-In other words, our models predict that as children provide more allocare within their families, the frequency of signaling would increase with alloparenting effort for girls and remain at the same rate or decrease slightly with alloparenting effort for boys.
-See @tbl-main_models for the corresponding regression table and @fig-effect_allo_by_sex for effects plots.
 
@@
 
-There were significant negative relationships between summed signal cost (`r stats$mBFc$BodyFat$str`) crying frequency (`r stats$mBFcry$BodyFat$str`), and tantrum frequency (`r stats$mBFtantrum$BodyFat$str`) and a scaled composite of subscapular and tricep skinfold thicknesses when controlling for the sex of the child, the age of the child, the frequency of conflict with one's primary caretaker, household demographics,caretaker perceptions of child immune function, an alloparenting effort by sex interaction and a body fat by sex interaction.
+There were significant negative relationships between summed signal cost (`r stats$mBFc$BodyFat$str`) crying frequency (`r stats$mBFcry$BodyFat$str`), and tantrum frequency (`r stats$mBFtantrum$BodyFat$str`) and a scaled composite of subscapular and tricep skinfold thicknesses when controlling for the sex of the child, the age of the child, the frequency of conflict with one's primary caretaker, household demographics,caretaker perceptions of child susceptibility to illness, an alloparenting effort by sex interaction and a body fat by sex interaction.
 See @tbl-bodyfat_models for regression statistics and @fig-effect_bodyfat for effects plots.
@@
 See @tbl-bodyfat_models for the corresponding regression table.
+:::
+
+## Crying frequency is positively associated with greater height and weight gain in a small longitudinal sample
+
+Signaling data and longitudinal data on child height and weight (2023 and 2024) exist for `r sum(table(height_long_cry$frame$HeightMean_2024))` children.
+In this sample, there was a significant effect of a child age by crying frequency interaction on 2024 child height and weight whereby more crying was associated with more growth when controlling for 2023 height and weight.
+
+See @tbl-longitudinal for the corresponding regression table.
+
+# Check: SD (Observations) in the table
+```{r}
+#| label: tbl-longitudinal
+#| tbl-cap: Regression tables for the models examining height and weight growth.
+r_vars <- c(
+  ChildAge = "Age (years)",
+  SexMale = "Sex (male)",
+  OtherChildrenHH = "Other children",
+  LogIncome = "Income (log)",
+  number_adults = "Adults in home",
+  PartnerStatusUnpartnered = "Partner status (single)",
+  EducationLevelYears = "Parent ed. (years)",
+  ConflictFreqN = "Conflict freq.",
+  SignalFreq = "Signal freq.",
+  SignalCost = "Signal cost",
+  SignalFreqMax = "Max signal freq.",
+  IllnessSusceptibilityMean = "Percieved illness susceptibility",
+  SadFreqN = "Sadness freq.",
+  CryFreqN = "Crying freq.",
+  TantrumFreqN = "Tantrum freq.",
+  AlloparentingFreqN = "Child allocare",
+  TricepR = "Tricep skinfold (residuals)",
+  SubscapR = "Subscapular skinfold (residuals)",
+  BodyFat = "BodyFat (composite)",
+  RelativeNeed3 = "Relative need (w/in household)",
+  HeightMean_2024 = "2024 height",
+  HeightMean_2023 = "2023 height",
+  WeightMean_KG_2024 = "2024 weight",
+  WeightMean_KG_2023 = "2023 weight"
+)
+
+modelsummary(list(Weight2024 = weight_long_cry, Height2024 = height_long_cry), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+```
+
+::: {#fig-effect_longitudinal_plot}
+```{r}
+longitudinal_plot
+```
+
+The effect of a child crying frequency by age on child weight and height in 2024, controlling for the corresponding measure in 2023.
+Rug corresponds to actual values of crying frequency per month with a small amount of jitter to reveal overlapped values.
+See @tbl-longitudinal for the corresponding regression table.
 :::
@@
 One possibility is that older children have greater ability to satisfy one's desires and behave in ways which match cultural expectations of teenagers and adults.
-Another is that since one's reproductive value increases as one survives periods of high mortality risk in childhood, the intensity conflicts of interest with parents decreases since these children are likely to reach adulthood and require less investment to reach adulthood than younger children, current or future [@daly1995].
+Another is that since one's reproductive value increases as one survives periods of high mortality risk in childhood, the intensity of conflicts of interest with parents decreases since these children are likely to reach adulthood and require less investment to reach adulthood than younger children, current or future [@daly1995].
 For this reason, less costly signals may be more likely to work to increase investment, all else being equal\[\[4\]\].
@@
 
-Consistent with signaling models which emphasize the role of conflicts of interest in favoring the use of costlier signals (cite this lab and others), there were positive associations between caretakers' perceptions of conflict frequency and 1) the summed costliness of children's signaling, 2) the frequency of child signals of need, and 3) the frequency of children's most common need signaling strategy.
+Consistent with signaling models which emphasize the role of conflicts of interest in favoring the use of costlier signals, there were positive associations between caretakers' perceptions of conflict frequency and 1) the summed costliness of children's signaling, 2) the frequency of child signals of need, and 3) the frequency of children's most common need signaling strategy.
 Although conflicts of interest do not always lead to behavioral conflict [@altmann1980; @bateson1994], it is likely many, if not most, conflicts are ultimately caused by conflicts of interest, making it a useful proxy for the underlying conflicts of interest that may increase the benefits of deception and the costs of being deceived\[\[5\]\].
-If this is the case, the association between signaling and conflict may represent children's attempts at credibly signaling their value of additional investment or the costliness of punishment (cite) to parents who would have good reason to doubt the trustworthiness of simple verbal request (or the trustworthiness of facial sadness and crying for those resorting to temper tantrums).
+If this is the case, the association between signaling and conflict may represent children's attempts at credibly signaling their value of additional investment or the costliness of punishment [@syme2019] to parents who would have good reason to doubt the trustworthiness of simple verbal requests (or the trustworthiness of facial sadness and crying for those resorting to temper tantrums).
 
@@
 
-We found evidence of an interaction between how often a child cared for their siblings and the sex of the child in our models examining signal frequencies and cost, whereby there was a positive relationship between the frequency of and summed costliness of child signals of need and alloparenting effort for girls but no relationship or a slightly negative one for boys.
+We found evidence of an interaction between how often a child cared for their siblings and the sex of the child in our models examining signal frequencies and cost, whereby there was a positive relationship between the frequency and summed costliness of child signals of need and alloparenting effort for girls but no relationship or a slightly negative one for boys.
 In many societies, girls are often expected to do more alloparenting within the household and often end up doing so [@helfrecht2016 @weisner1977; @barry1971; @turke1988].
@@
 Although the body fat by sex interaction was only marginally significant, our models predict a negative relationship between body fat and signaling for girls and no relationship for boys.
-Female reproduction is highly dependent on body fat, with conception requiring about xx% body fat and mothers investing a significant amount of calories in both children in the womb and breastfed infants [@jasienska2009].
-For these reasons, low body fat is likely a be better predictor of future reproductive success in adolescent (and older) girls than boys, potentially making signaling for increased investment a more viable strategy in instances when an individual has low body fat.
+Female reproduction requires a significant caloric invest in children, both in the womb and when breastfeeding [@jasienska2009], costs which males do not have to pay.
+For this reason, low body fat, which can prevent ovulation in extreme cases [@ellison2003], is likely a be better predictor of future reproductive success in adolescent (and older) girls than boys, potentially making signaling for increased investment a more viable strategy in instances when an individual has low body fat.
 
@@
 One reason for this could be that caretakers weighted monetary investment highly when evaluating child need.
+
+Signaling behavior did not predict (try more models?) differences in relative caregiver investment between children within households.
+However, crying frequency was  associated with both height and weight gain for older children.
+If older children face greater reputation costs from crying, crying may be a relatively credible credible signal compared to crying among younger children, with it being a less viable strategy to exaggerate one's value of investment or the costliness of adversity.
 
@@
 When the cause involved punishment, however, parents were more likely to respond negatively and less likely to respond positively.
-In these circumstances, children may be signaling about the severity of a punishment (cite syme costly apology).
+In these circumstances, children may be signaling about the severity of a punishment [@syme2019].
 However, parents may be reluctant to reduce punishments if the behavior that warranted the punishment was repeated or severe.
 They may also react with scolding or further punishment.
-That said, our model predicts positive responses in xxxxx cases, suggesting that children are successful in negotiating better outcomes at least some of the time.
-- costly apology model?
+That said, our model predicts positive responses in `r signif(out_mpr_full$estimate[3] * 100, 2)`% of cases (when all other predictors are held at their mean values), suggesting that children are successful in negotiating better outcomes at least some of the time.
 
@@
 Unfortunately, we do not have data on how much time caretakers spent with children to test this, and it seems equally likely that less educated parents may see their kids less often if they need to work more hours to support their families.
-Another possibility is that lower levels of education are associated with greater levels of adversity for both parents and children, something which would increase the benefits of signaling need for children and the difficulty of parents to respond.
+Another possibility is that lower levels of education are associated with greater levels of adversity for both parents and children, something which appears to be the case for other measures of SES [@pepper2017] and would increase the benefits of signaling need for children and the difficulty of parents to respond.
 It is also possible that education influences parenting norms in ways that cause differences in child signaling behavior.
@@
 Outside of the sex by alloparenting effort interaction, we found little evidence of sex differences in child signaling behavior.
-As adults, females cry more than males, a finding that has been reported across cultures (cite).
-A similar sex difference is also seen with depression and suicide attempts, which some of us have argued to often be attempts at signaling need in other papers (cite), with females experiencing depression being about twice as often (cite) and young females being xxxx as likely to attempt suicide (normal citaiton).
-The most likely reason for this difference is that there is less of a sex difference in adversity, conflict, and cultural evaluations of signals of need children until adolescence and young adulthood.
-Although we do have data on adolescents, our small sample for this age range may have prevented us from detecting this.
+As adults, females cry more than males, a finding that has been reported across cultures [@vanhemert2011; @becht2001].
+One reason why our study may have failed to find significant effects of sex is that sex differences in the causes and reputational effects of signaling need are small or nonexistent until adolescence and young adulthood, with our small sample size for this age range preventing us from detecting a real effect should it exist.
+It is also possible that the sex difference seen in rates of crying in adults may reflect the lesser ability of females to bargain for better treatment through aggression or the threat of it when faced with adversity [@hagen2016].
+Such a sex difference may not exist among preadolescencents due to a lack of difference in strength or athletic ability beofre puberty [@handelsman2018] or due to their relative lack of formidability compared to adults.
 
 Monthly income was also not found to be a significant predictor of signaling behavior.
-Many models designed to explain signals of need (e.g., the begging of young birds or withdrawal in humans) predict that offspring who are worse off or are receiving less investment have more to gain from signaling than those with more [@mock2011; @godfray1991; @caro2016a; @godfray2000].
+Many models designed to explain signals of need (e.g., the begging of young birds) predict that offspring who are worse off or are receiving less investment have more to gain from signaling than those with more [@mock2011; @godfray1991; @caro2016a; @godfray2000].
 However, it is possible that this can be true and household level resources have little influence on child signaling frequencies between families.
@@
 It is also likely that the two groups differ in how caretakers could respond to child signaling, and caretakers between the two populations may also differ in their cultural models of parenting.
-However, as previously mentioned, it is also conceivable that household level differences have little effect on child signaling behavior due to the costs of signaling and a child's fitness interest in their relatedness.
+However, as previously mentioned, it is  conceivable that household level differences have little effect on child signaling behavior.
 
@@
 The main limitation of this study is that it is completely exploratory, something which results in a high risk of our models overfitting the data and indicating associations which might not exist in the real world.
+Our study was also predominantly cross-sectional, further limiting our ability to make causal inferences.
+
 Our use of convenience and snowball sampling is also a limitation as it resulted in a non-probability sample which may not be representative of the broader population.
@@
 Our questions about the last time the child signaled may have also been more effective if they focused on specific signals due to our methods combining crying, which is unambiguous as a signal, with sadness, which has both signaling and non-signaling functions.
-
-Our study was also predominantly cross-sectional, which limis our ability to make causal inferences .
-A major challenge when studying signals of need is the fact that such signaling is expected to 1) be more common in times when individuals can benefit from increased investment (e.g., when they are facing major adversity or when a fitness-relevant opportunity arises) and 2) result in benefits to the signaler on average.
-In other words, higher levels of signaling may stem from greater levels of need while simultaneously eliciting more investment from others.
-This hypothesized causal, bi-directional relationship makes it difficult to isolate the effects of signaling behavior on investment in children when using standard measures of child welfare (i.e., BMI, health metrics, and height for age).
-Longitudinal studies can help address this.
-However, our small sample size of participants with signaling data and anthropometric data at two time periods prevents this for this study.
 
@@
 ### Cross-sectional study
+
 Another limitation is that our study was cross-sectional.

### 2024-09-09 analysisforpaper.R

@@ -232,9 +232,9 @@ d <- children |>
       RelativeNeedRecode2,
       ordered = TRUE,
       levels = c("Less", "The same amount", "More")), # change back to "The same" here
-    Currenthealth1 = invert(Currenthealth1),
-    Currenthealth2 = invert(Currenthealth2),
-    Currenthealth3 = Currenthealth3 + 0, # to remove labels
+    IllnessSusceptibility1 = invert(Currenthealth1),
+    IllnessSusceptibility2 = invert(Currenthealth2),
+    IllnessSusceptibility3 = Currenthealth3 + 0, # to remove labels
     Sex = as.factor(Sex),
     ChildAge = as.numeric(ChildAge),
     Age0_2.5 = ifelse(ChildAge < 2.5, 1, 0), #Helfrecht and Meehan 2015 style risk periods
@@ -260,8 +260,8 @@ d <- children |>
     SignalFreq = sum2(SadFreqN, CryFreqN, TantrumFreqN, na.rm = TRUE), # RunawayFreqN removed because interpretation not clear
     SignalFreqMax = max2(SadFreqN, CryFreqN, TantrumFreqN, na.rm = TRUE), # RunawayFreqN removed because interpretation not clear
     SignalCost = sum2(SadFreqN, CryFreqN * 2, TantrumFreqN * 3, na.rm = TRUE), # RunawayFreqN removed because interpretation not clear
-    CurrentHealthSum = sum2(Currenthealth1, Currenthealth2, Currenthealth3, na.rm = TRUE),
-    CurrentHealthMean = mean2(Currenthealth1, Currenthealth2, Currenthealth3, na.rm = TRUE)
+    IllnessSusceptibilitySum = sum2(IllnessSusceptibility1, IllnessSusceptibility2, IllnessSusceptibility3, na.rm = TRUE),
+    IllnessSusceptibilityMean = mean2(IllnessSusceptibility1, IllnessSusceptibility2, IllnessSusceptibility3, na.rm = TRUE)
   ) |>
   ungroup() |>
   left_join(anthropometricMeans, by = "ChildID")
@@ -394,10 +394,10 @@ modeldf <- d2 |>
     ConflictFreqN,
     AlloparentingFreqN,
     EducationLevelYears,
-    Currenthealth1,
-    Currenthealth2,
-    Currenthealth3,
-    CurrentHealthMean,
+    IllnessSusceptibility1,
[…]

### 2024-09-10 Paper.qmd

@@
 
-xxxxxxxxxxDescribe camps brieflyxxxxxxxxxxxxxx - include dates and how we handled those
+xxxxxxxxxxDescribe camps brieflyxxxxxxxxxxxxxx - include dates and how we handled those.
+For cross-sectional analyses, we used the 2023 anthropometric values due to them being closer in time to the caretaker surveys.
 
@@
 Measurements were taken directly on the skin.
-The triceps measurement occurred at the midpoint of the upper arm.
-The subscapular measurement occured immediately below the inferior angle of the scapula, being picked up diagonally at the natural fold line of the skin [@antóna].
+The triceps measurement was taken at the midpoint of the upper arm.
+The subscapular measurement occurred immediately below the inferior angle of the scapula, being picked up diagonally at the natural fold line of the skin [@antóna].
 Flexed bicep circumference was measured perpendicular to the long axis of the upper arm.
@@
 They may also react with scolding or further punishment.
-That said, our model predicts positive responses in `r signif(out_mpr_full$estimate[3] * 100, 2)`% of cases (when all other predictors are held at their mean values), suggesting that children are successful in negotiating better outcomes at least some of the time.
+That said, our model predicts positive responses in `r signif(updated_out$estimate[3] * 100, 2)`% of cases (when all other predictors are held at their mean values), suggesting that children are successful in negotiating better outcomes at least some of the time.
 
@@
 Other: Other \[Text Box\]
-
-[\[MG1\]](#_msoanchor_1)Alternative start: Although exceptions exist (cite),
-
-[\[MG2\]](#_msoanchor_2)Lumping together various models for the evolution of cooperation with either delayed benefits or inclusive fitness benefits. Not sure if this would confuse readers.
-
-[\[MG3\]](#_msoanchor_3)Probably worth mentioning studies which fail to find evidence for it and why that might me (largely alloparenting benefits of other children)
-
-[\[MG4\]](#_msoanchor_4)Find and cite others, if we decide this much detail is worth including. I am not sure it is.
-
-[\[MG5\]](#_msoanchor_5)Removed paragraph over conflict over marriage partners due to it not seeming highly relevant given our data.
-
-[\[MG6\]](#_msoanchor_6)Probably worth adding (see xxxxx for exceptions)
-
-[\[MG7\]](#_msoanchor_7)Might also be worth mentioning signals between siblings (e.g., Roulin 2001 or the negotiating need section in Wells 2003).
-
-[\[MG8\]](#_msoanchor_8)Consider reordering so that the adaptive problem precedes crying?
-
-[\[MG9\]](#_msoanchor_9)Not sure if this paragraph is worth the space.
-
-[\[MG10\]](#_msoanchor_10)Hadza children appear to be an exception  (Blurton Jones paper)
-
-[\[MG11\]](#_msoanchor_11)Comment in Soltis. Also worth fitting in the possibility that children may use crying to annoy parents until they give in.
-
-[\[MG12\]](#_msoanchor_12)Del Giudice, 2018 too?
-
-# Cut from paper:
-
-## Intro
-
-### Conflicts of interest within the family
-
-Conflicts of interest between children and parents may also involve how much children invest in the household.
-As previously mentioned, each individual child is expected to prefer greater investment in oneself than in one’s siblings, while parents, who are equally related to all their children, should often be expected to favor a more even distribution [@trivers1974].
-This can lead to conflict with older children who might prefer to spend their time in ways which are likely to result in future direct fitness benefits (e.g., finding romantic partners or forming strong bonds with friends) rather than taking care of their siblings or acquiring or preparing food.
-Such a dynamic has been found among the Chuukese, with both male and female children reporting conflict with their families over children’s desires to invest financial resources in their own status enhancing opportunities rather than in the household [@lowe2003].
-Similarly, Syme and Hagen [@syme2023a] found greater numbers of siblings were associated with increased likelihoods of severe conflict in a retrospective study with Chuukese-Americans.
-This may also be consistent with parents pressuring older children to invest in the household; however, there are other ways in which limited resources can result in behavioral conflict [@syme2023a].
-
-Conflicts of interest over investment in the household also appear to extend to when a child starts having their own children [@kushnick2023; @emlen1995].
-Due to the benefits children may provide their siblings, parents may be motivated to delay or subvert the reproductive careers of certain offspring [@flinn1989; @emlen1992], a conflict which is likely enhanced in families with step-parents due to the likelihood of future children being step-siblings rather than full siblings, thereby reducing the potential fitness benefits of alloparenting [@moya2014; @schaffnit2021].
-Situations can also arise in which it might pay parents for their children to start having their own children sooner than would be ideal for the adult child.
-For example, parents may be motivated for a child to leave the household and begin having children earlier than is ideal from the child’s perspective if their presence in the household leads to them taking investment from other children who caretakers deem more worthy of investment [@kushnick2023].
-
-Studies testing the existence of familial conflict over the timing of child reproduction are rare and direct tests on the existence of parent-offspring conflict over when a child should start their own reproductive career versus help existing siblings in humans are needed [@schaffnit2021].
-The bulk of evidence for the existence of such conflicts comes from work with non-human animals. For example, the cooperative breeding white-fronted bee-eaters (*Merops bullockoides*) will interrupt the courtship of their offspring, seemingly to continue to benefit from their undiluted assistance in digging nests, incubating the clutch, finding food for nestlings, and defending the nest [@emlen1992].
-When it comes to humans, Flinn found that rural Trinidadian females contributed substantial amounts of childcare and those with siblings under 10 years old had their first child at a later age [@flinn1989], both of which would be expected if there is a conflict of interest over investing in one’s siblings or one’s own reproduction.
-However, a study among Karo Batak women found no evidence of delayed reproduction for alloparenting benefits [@kushnick2006].
-
-### Crying
-
-One possibility is that crying is physiologically constrained in ways that prevent or limit faking.
-Evidence consistent with this comes from findings that crying is hard to do on command and findings individuals often cry while alone, something which is not predicted by signaling theory [@vingerhoets2001].
-However, the likelihood of crying is often clearly related to presence of supporters, something which is seen in findings that lonely individuals cry less than those with more support options despite reporting lower well-being [@vingerhoets2013] and the existence of delayed crying—times when crying is inhibited during the cause of negative affect only for it to be expressed when one is near potential supporters [@gracanin2018].
-A more likely possibility is that certain acoustic elements of infant cries are tied to their physiological state in ways that make faking certain states impossible, with crying serving as a graded signal in most other instances [@soltis2004; @zeifman2017].
-
-It is likely that the functions of crying and the mechanisms behind its ability to elicit pro-signaler responses change over one’s life.
-Compared to adults and older children, the physiological costs of crying are greater for newborns and infants, among whom crying can result in a substantial increase in energy expenditure [@rao1997; @thureen1998].
-Due to these costs, crying at certain intensities or for long periods of time may be prohibitively costly for infants with low survival prospects, making it a potential credible signal of vigor [@furlow1997; @lummaa1998; @soltis2004; @wells2003].
-Such signals are thought to be important due to parental adaptations which are suspected to take the viability of offspring into account when making decisions about whether it is worth investing in a specific child or withdrawing investment to instead invest in other children who are more likely to survive to reproductive age and be more competitive adults [@hagen1999; @lummaa1998].
-In line with this reasoning, most infanticides cross-culturally occur when a child has poor survival prospects, due to either health conditions or resource constraints [@daly1984; @soltis2004].
-
-Another often suggested means by which crying could result in credible signals of need is if the increased risk of predation that comes from loud vocalizations allows for signaling to be worthwhile for those in relatively high need but not for those with low need [@lummaa1998; @wells2003; @zahavi1977].
-Evidence of increased predation has been found in studies on birds [@leech1997; @lima2009; @redondo1992; @ryan1982], among whom louder broods in general have also been found to be more conspicuous to predators [@haskell1997; @leech1997].
-It is unclear how significant the risk of alerting predators would be ancestrally.
-However, there are ethnographic reports of mothers keeping infants and young children close in dangerous environments [@hewlett1998; @kaplan1987; @konner1972], something which likely lessens this risk.
-
-As children age, their reproductive value increases due to reduced odds of mortality, leading to a decreased risk of infanticide (Hagen 2004).
-For this reason, there is likely a transition period in which crying begins to convey information about how much a child values further investment.
-Many models have been put forward which attempt to explain human crying and begging in other species as credible signals of need due to the differential costs and benefits of crying and analogous vocalizations in other species as deterring the exaggeration of need [@godfray1991; @kilner1997; @parker2002].
-Although differing in their details, a common theme is that the indirect fitness costs signalers incur from imposing costs on that parents and siblings through begging (be them energetic costs, survival costs, or reduced likelihood of future siblings) creates a floor to possible deception but falls short of preventing it [@daly1990; @wells2003].
-
-When it comes to older children, the relatively insignificant energetic costs of crying are unlikely to be enough to result in credible signals in times of substantial conflicts of interest.
-For this reason, crying among older children may be more similar to adult crying, something which has been found to result in increased support and empathy but sometimes come at the cost of reputation and the potential for stigmatization [@cornelius2001; @cretser1982; @hendriks2006; @jesser1989; @vingerhoets2016].
-Such costs may be a means by which crying can still be a credible signal [@sznycer] with moderate conflicts of interest if they dissuade potential fakers from making the signal.
-
-## Discussion
-
-### Running away
-
-Probably cut When viewed in isolation, the frequencies of sadness, crying, and temper tantrums were also positively associated with conflict frequency between child and caretaker.
-However, the frequency of running away was not significantly associated with conflicts with parents.
-A retrospective study of parent-offspring conflict among Chon Chuuk immigrants in the northwest United States found that running away was the most common response to long-standing conflicts with parents and often resulted in beneficial outcomes for the child [@syme2023a].
-The threat running away poses to child wellbeing and, therefore, its ability to credibly signal information about one’s current situation is likely highly dependent on one’s socioecology and age.
-Among the Chon Chuuk, running away is often seen as a suicide threat [@fischer1950; @syme2023a], likely due to xxxxxxxxx.
-Similar risks may also occur in environments with large predators, venomous animals, or high degrees of human conflict.
-In these instances, even one night away can involve substantial risk as is seen among the !Kung where a woman running away and spending the night outside of camp can be used to communicate “the strength of her feelings” about a an unhappy relationship [@shostak1981]\[\[6\]\].
-When it comes to Utila, mothers in focus groups reported that running away on Utila was much less risky than it would be on mainland Honduras, with a few mothers mentioning that they warned their children about the difference in danger that came with running away on the mainland.
-Despite the relative safety of Utila, young children may still face risks out on their own due to proximity to the ocean, large dogs, and fast-moving motorcycles, golf carts, and tuk-tuks on the main road.
-For these reasons, running away may be too costly for young children but not costly enough to credibly signal need for older children.
-
-### Perceived immune function
-
-A negative association between caretaker estimates of child immune function and child signaling was seen, suggesting that signaling behavior was influenced by adversity.
-Perceptions of greater immune susceptibility were associated with both the summed costliness of signaling and the frequency of the most common child signal of need, while the frequency of child signaling was marginally significant.
-When examining the effect of perceived immune function on each signal, perceptions of greater immune susceptibility predicted the frequency of sadness but not the frequencies of crying and temper tantrums.
-
-One explanation for this may be that parents can easily detect child illness and that conflicts of interest tend to be less significant when it comes to illness care than other areas.
-Pathogens have been a strong selection pressure for life on earth for much of its existence and have greatly shaped evolutionary history (cite).
-In response to this, humans have evolved many adaptations for limiting the harm we experience from pathogens, including the ability to identify relaible indicators of illness in others.
-In some contexts, this activates disgust and motivates avoidance of sick individuals.
-However, it can also lead to increased care of relatives and allies (cite), with there being some suggestion that we have even evolved to upregulate some of our symptoms beyond what would arise from an immune response/pathogen interaction to better convince others that we are sick (cite).
-Due to this evolutionary history, parents may be better equipped to identify the possible benefits to investing in a sick child when compared to investment aimed at helping a child in status competition (e.g., buying new or popular clothing), something which may be exacerbated by culture change in what is popular among children.
-In other words, there is less private information for a child to reveal.
-This coupled with the likelihood that conflicts of interest between children and caretakers are smaller when it comes to illness care compared to other areas means there is less to gain from costlier signals.
-
-### Cross-sectional study
-
-Another limitation is that our study was cross-sectional.
-A major challenge when studying signals of need is the fact that such signaling is expected to 1) be more common in times when individuals can benefit from increased investment (e.g., when they are facing major adversity or when a fitness-relevant opportunity arises) and 2) result in benefits to the signaler on average.
-In other words, higher levels of signaling may stem from greater levels of need while simultaneously eliciting more investment from others.
-This hypothesized causal, bi-directional relationship makes it difficult to isolate the effects of signaling behavior on investment in children when using standard measures of child welfare (i.e., BMI, health metrics, and height for age).
-Longitudinal studies can help address this.
-However, our small sample size of participants with signaling data and anthropometric data at two time periods prevents this for this study.

### 2024-09-19 Paper.qmd
Added new plots and added householdID to anthropometrics for use in the GAMs.

@@
 Signals differ from cues in that signals were favored by natural selection due to their ability to alter the behavior or perceptions of others in ways that benefit the signaler, whereas cues are pieces of information organisms use to influence their behavior or physiological state that did not evolve for communicative effect [@maynardsmith1995].
-Crying is likely a signal of need[\[1\]](#_ftn1), for instance, while bleeding would be a cue of need.
+Crying is likely a signal of need[^1], for instance, while bleeding would be a cue of need.
+
+[^1]: We expect “signals of need” to actually be signals of resource valuation.
+However, we use “signal of need” here to distinguish signals that involve attempts at eliciting support through revealing private information about one’s vulnerabilities from the more commonly studied signals which reveal private information about one’s positive qualities (i.e., those related to partner choice and intrasexual competition).
 
@@
 All statistical analyses were exploratory and involved fitting many models with different combinations of independent and dependent variables.
-While were cognizant of the risk of overfitting the data, it is still more likely than if the study was entirely confirmatory.
+While were cognizant of the risk of overfitting the data, it is still more likely than if the study was entirely confirmatory [@tukey1977].
 The payoff to this approach is the identification of theoretically relevant effects that should be tested in the future.
@@
 The first used the polr function in the MASS package [@venables2002].
-The results were then interpreted with clustered standard errors for household.
+The results were then interpreted with clustered standard errors for household [@zeileis2002].
 Due to the risk of overfitting that comes with having too few clusters [@cameron2015], we also fit the same models using BRMS [@bürkner2017] with household as a random effect.
@@
 # Check: SD (Observations) in the table
+
 ```{r}

### 2024-09-19 analysisforpaper.R

@@ -55,6 +55,10 @@ mean2 <- function(..., na.rm = F){
 
 # dataframe prep ----------------------------------------------------------
 
+children2 <- children |>
+  dplyr::filter(ChildID != "") |>
+  dplyr::select(ChildID, householdID)
+
 anthropometricMeans0 <- anthropometrics |>
   dplyr::select(
     ChildID,
@@ -76,7 +80,8 @@ anthropometricMeans0 <- anthropometrics |>
 anthropometricMeans <- anthropometricMeans0 |>
   dplyr::filter(
     measurements == 1 | year(Date.of.measurement) == 2023
-  )
+  ) |>
+  left_join(children2) # note that we filter out householdID based on it's column number
 
 # moved below
 # anthropometricMeansWide <- anthropometricMeans0 |>
@@ -89,21 +94,25 @@ anthropometricMeans <- anthropometricMeans0 |>
 anthropometricMeans$WeightZ <- growthRef(AgeAtMeasurement, WeightMean_KG, Sex, anthropometricMeans, type = "Weight", pop = "CDC")
 anthropometricMeans$HeightZ <- growthRef(AgeAtMeasurement, HeightMean, Sex, anthropometricMeans, type = "Height", pop = "CDC")
 anthropometricMeans$BMIZ <- growthRef(AgeAtMeasurement, BMI, Sex, anthropometricMeans, type = "BMI", pop = "CDC")
+anthropometricMeans$Sex2 <- factor(anthropometricMeans$Sex)
 
-m <- mgcv::gam(GripStrengthMean ~ Sex+s(AgeAtMeasurement, by = as.factor(Sex), k = 4), data = anthropometricMeans)
+m <- mgcv::gam(GripStrengthMean ~ Sex+s(AgeAtMeasurement, by = Sex2, k = 4), data = anthropometricMeans)
 anthropometricMeans$GripR <- residuals(m)
 
-m2 <- mgcv::gam(TricepMean ~ Sex+s(AgeAtMeasurement, by = as.factor(Sex)), data = anthropometricMeans, na.action = na.exclude)
+m2 <- mgcv::gam(TricepMean ~ Sex+s(AgeAtMeasurement, by = Sex2, k = 3), data = anthropometricMeans, na.action = na.exclude)
 anthropometricMeans$TricepR <- residuals(m2)
 
[…]

### 2024-09-19 Paper.qmd
Removed some vars from summary table, and split table by sex

@@
   SignalFreqMax = "Frequency of most common signal (times per month)",
-  IllnessSusceptibility1 = "IllnessSusceptibility1",
-  IllnessSusceptibility2 = "IllnessSusceptibility2",
-  IllnessSusceptibility3 = "IllnessSusceptibility3",
-  IllnessSusceptibilityMean = "Mean of current health measures",
+  # IllnessSusceptibility1 = "Illness Susceptibility1",
+  # IllnessSusceptibility2 = "Illness Susceptibility2",
+  # IllnessSusceptibility3 = "Illness Susceptibility3",
+  IllnessSusceptibilityMean = "Illness avoidance index",
   HeightZ = "Height Z-score (CDC)",
@@
 
-kable(custom.summarize(modeldf, vars_c)) %>% kable_styling()
+kable(custom.summarize(modeldf, vars_c, facvar = "Sex")) %>% kable_styling()
 

### 2024-09-25 Paper.qmd
added toc. tweaked axis labels. added code to simplify stat reporting.

@@
   warning: false
-format: html
+format:
+  html:
+    toc: true
 lightbox: true
@@
 vars_c <- c(
-  CaregiverAge = "Age (years)",
-  EducationLevelYears = "Education level (years)",
+  CaregiverAge = "Caregiver age (years)",
+  EducationLevelYears = "Caregiver education level (years)",
   IncomeCategoryN = "Monthly household income (Lempira)",
   number_adults = "Adults in household",
-  AdultsChildcare = "Adults who provide childcare"
+  AdultsChildcare = "Household adults who provide childcare"
 )
@@
 
+```{r}
+fmt_sum <- function(v){
+  v <- na.omit(v)
+  n <- length(v)
+  thesum <- sum(v)
+  thepct <- 100*signif(thesum/n, 2)
+  str_glue("{thesum} children ({thepct}%)")
+}
+```
+
 Child sadness, crying, and temper tantrums were common.
-Caregivers reported that `r sum(na.omit(modeldf$SadFreqN >= 8))` (`r signif(sum((na.omit(modeldf$SadFreqN >= 8)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%) children were sad at least once a week, `r sum(na.omit(modeldf$CryFreqN >= 8))` (`r signif(sum((na.omit(modeldf$CryFreqN >= 8)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%) cried at least more than once a week, and `r sum(na.omit(modeldf$TantrumFreqN >= 8))` (`r signif(sum((na.omit(modeldf$TantrumFreqN >= 8)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%) threw temper tantrums at least once a week.
-A small subset of children engaged in these signals/states at least daily (sadness = `r sum(na.omit(modeldf$SadFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$SadFreqN >= 30)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\], crying = `r sum(na.omit(modeldf$CryFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$CryFreqN >= 30)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\], and tantrums = `r sum(na.omit(modeldf$TantrumFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$TantrumFreqN >= 30)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]).
-However, many children were reported to never engage in these signals/states (sadness = `r sum(na.omit(modeldf$SadFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$SadFreqN == 00)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\], crying = `r sum(na.omit(modeldf$CryFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$CryFreqN == 0)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\], and tantrums = `r sum(na.omit(modeldf$TantrumFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$TantrumFreqN == 0)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]).
-See @fig-barplot_SignalFreq for plot showing signaling frequencies and @fig-signal_alluvial_plot for plot showing the number of individuals who engaged in specific combinations of signal frequencies..
-
-Conflict between children and caretakers was also common, with `r sum(na.omit(modeldf$ConflictFreqN >= 8))` children (`r signif(sum((na.omit(modeldf$ConflictFreqN >= 8)) / sum(!is.na(modeldf$ConflictFreqN)) * 100), 2)`%) having conflicts with parents more than once a week.
-As with the frequencies of various types of signaling, many children (`r sum(na.omit(modeldf$ConflictFreqN == 0))` \[`r signif(sum((na.omit(modeldf$ConflictFreqN == 0)) / sum(!is.na(modeldf$ConflictFreqN)) * 100), 2)`%\]) were described as never having conflicts with parents.
-See @fig-barplot_ConflictFreq for plot showing the frequencies of conflict between caretakers and children.
-
-The frequency of alloparenting effort was bimodally distributed, with `r sum(na.omit(modeldf$AlloparentingFreqN == 0))` children (`r signif(sum((na.omit(modeldf$AlloparentingFreqN == 0)) / sum(!is.na(modeldf$AlloparentingFreqN)) * 100), 2)`%) being reported to never engage in it (`r sum(na.omit(modeldf$NumberOfChildren==1))` instances were due to a child being an only child) and `r sum(na.omit(modeldf$AlloparentingFreqN >= 30))` children (`r signif(sum((na.omit(modeldf$AlloparentingFreqN >= 30)) / sum(!is.na(modeldf$AlloparentingFreqN)) * 100), 2)`%) being reported to help with childcare at least once a day.
-See @fig-barplot_AlloparentingFreq for plot showing the frequencies child alloparenting effort.
+Caregivers reported that `{r} fmt_sum(modeldf$SadFreqN >= 8)`  were sad at least once a week, `{r} fmt_sum(modeldf$CryFreqN >= 8)` cried at least once a week, and `{r} fmt_sum(modeldf$TantrumFreqN >= 8)` threw temper tantrums at least once a week.
+A small subset of children engaged in these signals/states at least daily; sadness: `{r} fmt_sum(modeldf$SadFreqN >= 30)`; crying: `{r} fmt_sum(modeldf$CryFreqN >= 30)`;  and tantrums: `{r} fmt_sum(modeldf$TantrumFreqN >= 30)`.
+However, many children were reported to never engage in these signals/states; sadness: `{r} fmt_sum(modeldf$SadFreqN == 0)`; crying; `{r} fmt_sum(modeldf$CryFreqN == 0)`; and tantrums: `{r} fmt_sum(modeldf$TantrumFreqN == 0)`.
+See @fig-barplot_SignalFreq and @fig-signal_alluvial_plot.
+
+Conflict between children and caretakers was also common, with `{r} fmt_sum(modeldf$ConflictFreqN >= 8)` having conflicts with parents more than once a week.
+As with the frequencies of various types of signaling, `{r} fmt_sum(modeldf$ConflictFreqN == 0)` were described as never having conflicts with parents.
+See @fig-barplot_ConflictFreq.
+
+The frequency of alloparenting effort was bimodally distributed, with `{r} fmt_sum(modeldf$AlloparentingFreqN == 0)` being reported to never engage in it (`r sum(na.omit(modeldf$NumberOfChildren==1))` instances were due to a child being an only child) and `{r} fmt_sum(modeldf$AlloparentingFreqN >= 30)` being reported to help with childcare at least once a day.
+See @fig-barplot_AlloparentingFreq.
 
@@
 ```{r}
-barplot_SignalFreq
+barplot_SignalFreq + guides(fill = guide_legend(title = 'Times/month', nrow = 1, position = 'top', reverse = T))
 ```

### 2024-09-25 agesignaling.R

@@ -1,0 +1,21 @@
+
+library(mgcv)
+library(lmerTest)
+library(marginaleffects)
+library(gratia)
+
+signals <-
+  d2 |>
+  dplyr::select(uniqueID, ChildAge, SadFreqN, CryFreqN, TantrumFreqN) |>
+  pivot_longer(SadFreqN:TantrumFreqN, names_to = 'Type', values_to = 'Freq') |>
+  mutate(uniqueID = factor(uniqueID), Type = factor(Type)) |>
+  na.omit()
+
+m <- gam(Freq ~ s(ChildAge, by = Type) + s(uniqueID, bs = 're'), family = nb, data = signals)
+summary(m)
+draw(m, select = 1:3, fun = exp) # Correct?
+
+plot_predictions(m, condition = c('ChildAge', 'Type'), type = 'link', transform = exp)
+plot_predictions(m, condition = c('ChildAge', 'Type'), type = 'response')
+
+m2 <- lmer()

### 2024-09-25 Paper.qmd
Analysis script: Added corrplots. Fixed plots for the main models with signaling as the outcome. Added code exploring alloparenting.
Paper:
Added code to source dictionary. Minor dictionary changes.

--- 2024-09-19 Paper.qmd
@@
 ```{r setup, include=FALSE}
-source("analysisforpaper.R", local = knitr::knit_global())
+source("dictionaries.R")
+source("analysisforpaper.R", local = knitr::knit_global()) # experiment with taking this away
 
@@
 |                                 |                                                                                                                   |
-|--------------------------|----------------------------------------------|
+|--------------------------|---------------------------------------------|
 | Variable                        | Survey question wording                                                                                           |
@@
 Variables created with this scheme include the frequencies of sadness, crying, temper tantrums, conflict with primary caretakers, and child alloparenting effort.
+{#tbl-convertedfreqs}
 
@@
 |                     |                                                                                                  |
-|--------------------------|----------------------------------------------|
+|--------------------------|---------------------------------------------|
 | Variable            | Survey question wording                                                                          |
@@
 |                        |                                                                                       |
-|--------------------------|----------------------------------------------|
+|--------------------------|---------------------------------------------|
 | Variable               | Survey question wording                                                               |
@@
   TantrumFreqN = "Temper tantrum frequency (times per month)",
-  AlloparentingFreqN = "Frequency of child's own alloparetning effort (times per month)",
+  AlloparentingFreqN = "Frequency of child's own alloparenting effort (times per month)",
   ConflictFreqN = "Frequency of conflict with focal caretaker (times per month)",
@@
   SignalFreqMax = "Frequency of most common signal (times per month)",
-  IllnessSusceptibility1 = "IllnessSusceptibility1",
-  IllnessSusceptibility2 = "IllnessSusceptibility2",
-  IllnessSusceptibility3 = "IllnessSusceptibility3",
-  IllnessSusceptibilityMean = "Mean of current health measures",
+  # IllnessSusceptibility1 = "IllnessSusceptibility1",
+  # IllnessSusceptibility2 = "IllnessSusceptibility2",
+  # IllnessSusceptibility3 = "IllnessSusceptibility3",
+  IllnessSusceptibilityMean = "Illness avoidance index",
   HeightZ = "Height Z-score (CDC)",
@@
   AdultsChildcare = "Adults that provide childcare",
-  OtherChildAlloparentingFreqN = "Other household child alloparetning freq."
+  OtherChildAlloparentingFreqN = "Other household child alloparenting freq."
 )

### 2024-09-25 analysisforpaper.R

@@ -23,6 +23,7 @@ library(kableExtra)
 library(modelsummary)
 library(ggalluvial)
 library(patchwork)
+library(gratia)
 
 source("recode.R")
 
@@ -99,8 +100,29 @@ anthropometricMeans$Sex2 <- factor(anthropometricMeans$Sex)
 m <- mgcv::gam(GripStrengthMean ~ Sex+s(AgeAtMeasurement, by = Sex2, k = 4), data = anthropometricMeans)
 anthropometricMeans$GripR <- residuals(m)
 
+# Result for bodyfat predicitng crying freq changes if k is unspecified
 m2 <- mgcv::gam(TricepMean ~ Sex+s(AgeAtMeasurement, by = Sex2, k = 3), data = anthropometricMeans, na.action = na.exclude)
 anthropometricMeans$TricepR <- residuals(m2)
+
+# anthropometricMeans$Sex3 <- as.numeric(anthropometricMeans$Sex2)
+
+# anthropometricMeans_female <- na.omit(anthropometricMeans[anthropometricMeans$Sex == "F", c("ChildID", "AgeAtMeasurement", "Sex", "TricepMean")])
+# anthropometricMeans_male <- na.omit(anthropometricMeans[anthropometricMeans$Sex == "M", c("ChildID", "AgeAtMeasurement", "Sex", "TricepMean")])
+#
+# m2bf <- loess(TricepMean ~ AgeAtMeasurement, data = anthropometricMeans_female, span = 5)
+# plot_predictions(m2bf, condition = c("AgeAtMeasurement"))
+# visreg(m2bf)
+# anthropometricMeans_female$TricepR <- m2bf$residuals
+#
+# m2bm <- loess(TricepMean ~ AgeAtMeasurement, data = anthropometricMeans_male, span = 5)
+# plot_predictions(m2bm, condition = c("AgeAtMeasurement"))
+# visreg(m2bm)
+# anthropometricMeans_male$TricepR <- m2bm$residuals
+#
+# temp <- bind_rows(anthropometricMeans_female, anthropometricMeans_male) |>
+#   dplyr::select(ChildID, TricepR)
+#
+# anthropometricMeans <- dplyr::left_join(anthropometricMeans, temp)
 
 m3 <- mgcv::gam(SubscapMean ~ Sex+s(AgeAtMeasurement, by = Sex2, k = 3), data = anthropometricMeans, na.action = na.exclude)
 anthropometricMeans$SubscapR <- residuals(m3)
[…]

### 2024-09-25 dictionaries.R

@@ -1,0 +1,48 @@
+# c(
+#   "IllnessSusceptibilityMean" = "",
+#   "EducationLevelYears" = "",
+#   "OtherChildrenHH" = "",
+#   "LogIncome" = "",
+#   "ConflictFreqN" = "",
+#   "number_adults" = "",
+#   "PartnerStatus" = "",
+#   "AlloparentingFreqN" = "",
+#   "Sex" = "",
+#   "ChildAge" = "",
+#   "RelativeNeed3" = "",
+#   "RelativeMaternalInvestment2" = "",
+#   "SadFreqN" = "",
+#   "CryFreqN" = "",
+#   "TantrumFreqN" = "",
+#   "SignalFreq" = "",
+#   "SignalFreqMax" = "",
+#   "SignalCost" = "",
+# )
+
+r_vars <- c(
+  ChildAge = "Age (years)",
+  SexMale = "Sex (male)",
+  OtherChildrenHH = "Other children",
+  LogIncome = "Income (log)",
+  number_adults = "Adults in home",
+  PartnerStatusUnpartnered = "Partner status (single)",
+  EducationLevelYears = "Parent ed. (years)",
+  ConflictFreqN = "Conflict freq.",
+  SignalFreq = "Signal freq.",
+  SignalCost = "Signal cost",
+  SignalFreqMax = "Max signal freq.",
+  IllnessSusceptibilityMean = "Percieved illness susceptibility",
+  SadFreqN = "Sadness freq.",
+  CryFreqN = "Crying freq.",
+  TantrumFreqN = "Tantrum freq.",
+  AlloparentingFreqN = "Child allocare",
+  TricepR = "Tricep skinfold (residuals)",
+  SubscapR = "Subscapular skinfold (residuals)",
+  BodyFat = "BodyFat (composite)",
+  RelativeNeed3 = "Relative need (w/in household)",
+  UnwantedTask = "Unwanted task",
+  Punishment21 = "Punishment",
+  Family21 = "Family conflict",
+  DiscomfortPainInjuryIllness = "Discomfort/Pain/Injury/Illness",
+  CaregiverResponse = "Caregiver Response"
+)

### 2024-09-27 Paper.qmd
Adding relative need predicts relative investment section to the paper.

@@
 ```{r setup, include=FALSE}
-source("dictionaries.R")
+# source("dictionaries.R")
 source("analysisforpaper.R", local = knitr::knit_global()) # experiment with taking this away
@@
 
+## Relative need predicts relative investment
+
+Signals are capable of eliciting better treatment because the information they provide alters the behavior of signal targets in ways the benefit the signaler.
+For this reason, we tested the association between caretaker perceptions of relative child need within the household and the caretaker's relative investment within the household.
+Consistent with signaling theory, our model predicts a greater proportion of caretakers investing more in children who they deem needier and a greater proportion of caretakers investing less in children they deem less needy, controlling for the age of the child, and household demographics.
+See @fig-plot_relative_investment.
+
+::: {#fig-plot_relative_investment}
+```{r, fig.width= 8}
+plot_relative_investment
+```
+
+The proportion of children reported to be given more investment, the same amount of investment, or less investment than other children in the household by the caretaker.
+Effect of perceived need (in terms of time, effort, and resources).
+Outcome variable question wording: How much time, effort, and resources do you put into this child compared to your other children?
+:::
+
 ## Caregiver responses to last signal of need

### 2024-09-27 analysisforpaper.R

@@ -552,12 +552,12 @@ sfdfsum2 <-
   ) |>
   mutate(
     CostCategory = cut(
-      MeanSignalCost, 
+      MeanSignalCost,
       quantile(MeanSignalCost)[-3], # Bottom 1/4, Middle 1/2, Top 1/4
-      labels = c("Low", "Medium", "High"), 
+      labels = c("Low", "Medium", "High"),
       include.lowest = T
     )
-  ) |> 
+  ) |>
   arrange(desc(Freq)) |>
   dplyr::filter(Freq > 2)
 
@@ -896,7 +896,7 @@ draw(m)
 anthropometricMeansWide$heightR <- residuals(m)
 
 ggplot(anthropometricMeansWide, aes(SadFreqN, heightR, colour = Sex2)) +
-  geom_jitter(size = 3) + 
+  geom_jitter(size = 3) +
   geom_smooth(method = 'lm') +
   scale_color_binary()
 
@@ -1297,7 +1297,7 @@ coeftest(mpr, vcov = vcovCL, type = "HC0", cluster = ~householdID)
 
 mpr_plot_data <- plot_predictions(mpr, condition = c("RelativeNeed3", "group"), type = "probs", vcov = ~householdID, draw = FALSE)
 
-ggplot(mpr_plot_data, aes(RelativeNeed3, estimate, color = group, group = group, ymin = conf.low, ymax = conf.high)) +
+plot_relative_investment <- ggplot(mpr_plot_data, aes(RelativeNeed3, estimate, color = group, group = group, ymin = conf.low, ymax = conf.high)) +
   geom_line(position = position_dodge(width = .2), linewidth = 1.5) +
   geom_pointrange(position = position_dodge(width = .2)) +
   theme_minimal(15) +
@@ -1308,11 +1308,40 @@ ggplot(mpr_plot_data, aes(RelativeNeed3, estimate, color = group, group = group,
   guides(color = guide_legend(title = "Relative investment:")) &
   theme(legend.position = "top")
 
+p_need_age_sf <- plot_predictions(mN1, condition = c("ChildAge", "group"), type = "probs", vcov = ~householdID) +
+  ylab("Relative need") +
+  xlab("Child age (years)") +
+  scale_color_viridis_d(option = "B", end = .8) +
+  scale_fill_viridis_d(option = "B", end = .8) +
+  scale_linewidth_ordinal(range = c(5,10)) +
[…]

### 2024-09-27 Paper.qmd
Updated dictionaries.R and updated paper to use these dictionaries.

@@
 ```{r setup, include=FALSE}
-# source("dictionaries.R")
-source("analysisforpaper.R", local = knitr::knit_global()) # experiment with taking this away
+source("dictionaries.R", local = knitr::knit_global())
+source("analysisforpaper.R", local = knitr::knit_global())
 
@@
 
-kable(custom.summarize(modeldf, vars_c, facvar = "Sex")) %>% kable_styling()
+#kable(custom.summarize(modeldf, vars_c, facvar = "Sex")) %>% kable_styling()
+kable(custom.summarize(modeldf, longform_dict[1:18])) %>% kable_styling()
 
@@
 
-kable(custom.summarize(caregivers, vars_c)) %>% kable_styling()
+kable(custom.summarize(caregivers, longform_dict[19:23])) %>% kable_styling()
 
@@
 )
-modelsummary(list(Sadness = msadH, Crying = mcryH, Tantrum = mtantrumH, SignalFreq = msfH, SignalCost = mscH, SignalFreqMax = mmsfH), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+# modelsummary(list(Sadness = msadH, Crying = mcryH, Tantrum = mtantrumH, SignalFreq = msfH, SignalCost = mscH, SignalFreqMax = mmsfH), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+
+modelsummary(list(Sadness = msadH, Crying = mcryH, Tantrum = mtantrumH, SignalFreq = msfH, SignalCost = mscH, SignalFreqMax = mmsfH), gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
 ```
@@
 )
-modelsummary(mconflict, gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+modelsummary(mconflict, gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
 ```
@@
 )
-modelsummary(malloparenting, gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+modelsummary(malloparenting, gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
 ```
@@
 )
-modelsummary(list(Sadness = mBFsad, Crying = mBFcry, Tantrum = mBFtantrum, SignalFreq = mBFf, SignalCost = mBFc, SignalFreqMax = mBFfm), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+modelsummary(list(Sadness = mBFsad, Crying = mBFcry, Tantrum = mBFtantrum, SignalFreq = mBFf, SignalCost = mBFc, SignalFreqMax = mBFfm), gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
 ```
@@
 
-modelsummary(list(Weight2024 = weight_long_cry, Height2024 = height_long_cry), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+modelsummary(list(Weight2024 = weight_long_cry, Height2024 = height_long_cry), gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
 ```
@@
 )
-modelsummary(list(SignalFreq = mN1, SignalCost = mN2), vcov = vcovCL, type = "HC0", cluster = ~householdID, gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+modelsummary(list(SignalFreq = mN1, SignalCost = mN2), vcov = vcovCL, type = "HC0", cluster = ~householdID, gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
 ```
@@
 )
-modelsummary(mpr, vcov = vcovCL, type = "HC0", cluster = ~householdID, gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+modelsummary(mpr, vcov = vcovCL, type = "HC0", cluster = ~householdID, gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
 ```

### 2024-09-27 analysisforpaper.R

@@ -866,6 +866,7 @@ plot(allEffects(weight_long_cry))
 
 anthropometricMeansWide$weightR <- residuals(weight_long_cry)
 ggplot(anthropometricMeansWide, aes(ChildAge, CryFreqN, colour = weightR)) + geom_jitter(size = 3) + scico::scale_color_scico(palette = 'vik', midpoint = 0)
+ggplot(anthropometricMeansWide, aes(ChildAge, weightR, colour = CryFreqN)) + geom_jitter(size = 3) + scico::scale_color_scico(palette = 'vik', midpoint = 0)
 
 p <- plot_predictions(weight_long_cry, condition = c("CryFreqN", "ChildAge"), vcov = TRUE) +
   geom_rug(data = weight_long_cry$frame, aes(x = CryFreqN , y = 0), position = position_jitter(width = 1, seed = 123), sides = "b") +
@@ -879,12 +880,13 @@ p$layers[[2]]$aes_params$linewidth <- 2
 p$layers[[1]]$aes_params$alpha <- 0
 p
 
-height_long_cry <- glmmTMB(HeightMean_2024 ~ HeightMean_2023 + CryFreqN*ChildAge + (1|householdID), data = anthropometricMeansWide, family = gaussian, na.action = na.exclude)
+height_long_cry <- glmmTMB(HeightMean_2024 ~ HeightMean_2023 + CryFreqN*ChildAge + Sex + (1|householdID), data = anthropometricMeansWide, family = gaussian, na.action = na.exclude)
 summary(height_long_cry)
 plot(allEffects(height_long_cry))
 
 anthropometricMeansWide$heightR <- residuals(height_long_cry)
 ggplot(anthropometricMeansWide, aes(ChildAge, CryFreqN, colour = heightR)) + geom_jitter(size = 3) + scico::scale_color_scico(palette = 'vik', midpoint = 0)
+ggplot(anthropometricMeansWide, aes(ChildAge, heightR, colour = CryFreqN)) + geom_jitter(size = 3) + scico::scale_color_scico(palette = 'vik', midpoint = 0)
 
 m <- glmmTMB(HeightMean_2024 ~ HeightMean_2023 + Sex2 + (1|householdID), data = anthropometricMeansWide, family = gaussian, na.action = na.exclude)
 summary(m)
@@ -1872,9 +1874,11 @@ summary(m)
 draw(m)
[…]

### 2024-09-27 dictionaries.R

@@ -1,48 +1,70 @@
-# c(
-#   "IllnessSusceptibilityMean" = "",
-#   "EducationLevelYears" = "",
-#   "OtherChildrenHH" = "",
-#   "LogIncome" = "",
-#   "ConflictFreqN" = "",
-#   "number_adults" = "",
-#   "PartnerStatus" = "",
-#   "AlloparentingFreqN" = "",
-#   "Sex" = "",
-#   "ChildAge" = "",
-#   "RelativeNeed3" = "",
-#   "RelativeMaternalInvestment2" = "",
-#   "SadFreqN" = "",
-#   "CryFreqN" = "",
-#   "TantrumFreqN" = "",
-#   "SignalFreq" = "",
-#   "SignalFreqMax" = "",
-#   "SignalCost" = "",
-# )
+# long form dictionary (for use in summary tables)
+longform_dict <-
+  c(
+    ChildAge = "Age (years)",
+    SadFreqN = "Sad frequency (times per month)",
+    CryFreqN = "Crying frequency (times per month)",
+    TantrumFreqN = "Temper tantrum frequency (times per month)",
+    AlloparentingFreqN = "Frequency of child's own alloparenting effort (times per month)",
+    ConflictFreqN = "Frequency of conflict with focal caretaker (times per month)",
+    SignalFreq = "Frequency of signals summed (times per month)",
+    SignalCost = "Signal cost",
+    SignalFreqMax = "Frequency of most common signal (times per month)",
+    IllnessSusceptibilityMean = "Illness avoidance index",
+    HeightZ = "Height Z-score (CDC)",
+    WeightZ = "Weight Z-score (CDC)",
+    BMIZ = "BMI Z-score (CDC)",
+    SubscapMean = "Subscapular skinfold thickness (mm)",
+    SubscapR = "Subscapular skinfold residuals",
+    TricepMean = "Tricep skinfold thickness (mm)",
+    TricepR = "Tricep skinfold residuals",
+    BodyFat =  "Body fat composite", # child summary stats end here (1:18)
+    CaregiverAge = "Caregiver age (years)",
+    EducationLevelYears = "Caregiver education level (years)",
+    IncomeCategoryN = "Monthly household income (Lempira)",
+    number_adults = "Adults in household",
+    AdultsChildcare = "Household adults who provide childcare" # adults summary table (19:23)
+  )
 
-r_vars <- c(
-  ChildAge = "Age (years)",
[…]

### 2024-10-03 Paper.qmd
Paper: Added accepted changes from current paper draft. Updated plots, tables, and in-text calls of model values to fit with new scheme.
Analysis script: Many changes.

--- 2024-09-27 Paper.qmd
@@
 ---
-title: "Child Signals of Need in Utila, Honduras: An Exploratory Study"
+title: "The Natural History of Child Signals of Need in Utila, Honduras: An Exploratory Study"
 execute:
@@
 bibliography: signaling.bib
+crossref:
+  custom:
+    - kind: float
+      reference-prefix: Figure S
+      key: sfig
+      space-before-numbering: false
+    - kind: float
+      reference-prefix: Table S
+      key: stbl
+      space-before-numbering: false
 ---
@@
 
-Despite the complexity of childhood and adolescence and their importance to adult outcomes, the behavioral strategies of children are often overlooked, especially when it comes to how they elicit support from parents and other caretakers [@gottlieb2000; @wells2003; @hirschfeld2002].
-To gain support, children use signals to communicate their desires to parents.
-Infant signals of need, particularly infant cries, have been extensively studied in a range of populations [@barr1991; @fouts2005; @konner1974; @kruger2010; @kushnick2006; @meehan2017; @müller2021; @tronick1987], yet the full range of potential child signals of need, which likely extend beyond cries, has never been investigated in children of all ages.
-In particular, if and when children signal need with costly signals [@grafen1990; @spence1973; @zahavi1975a] has rarely been investigated [for one exception, see, @syme2023a].
-
-In this paper, we use exploratory analysis on child signaling data from Utila Honduras to examine the hypothesis that children have evolved strategies to tailor their signals of need to 1) their valuation of a given resource, 2) the extent of conflicts of interest with caretakers, and 3) the presence of existing information about a child's need or potential to benefit from a resource so that they can credibly convey they are worth additional investment.
-Such strategies can lead to some combination of verbal requests for more support, facial sadness, crying, withdrawal, temper tantrums and self-harm, all of which may be effective in certain circumstances and are sometimes successful in eliciting more support from caretakers, despite a child's low status and formidability.
-
-## Interdependence and investment within the family
-
-Throughout human evolutionary history, individuals often found themselves in situations in which they received fitness-relevant benefits from others that could not easily be replaced [@apicella2012; @apicella2019; @hill2011; @tooby1996].
-In these instances, natural selection is expected to have favored those who provided support to others if 1) such help meant one could expect to receive fitness benefits from those helped in the future and 2) the costs of providing help were outweighed by the benefits to the helper on average [@andré2010; @hamilton1964; @tooby1996; @trivers1971].
-
-Care motivated by a high degree of fitness interdependence between offspring and at least one parent is seen in many animals [@clutton-brock1991; @webb1999].
-However, humans are a uniquely interdependent species among the living primates [@tomasello2012] in that one’s fitness prospects depend greatly on a network of kin and other social allies [@apicella2019].
-This need for others is most intense with children, who are born highly altricial and require many years of investment to develop to the point where they can obtain enough food to survive on their own and develop the social skills needed to succeed in their communities [@kaplan2009; @rosenberg2021; @schniter2015].
-Much of this investment comes from parents [@kramer2018], However, other adults like grandparents, aunts, uncles, and non-relatives have also been found to provide significant investment in children in many societies [@kramer2018; @sear2008], something which was likely favored due to a combination of indirect fitness benefits, reciprocity, or reputation management [@gurven2000] + (Schniter Preprint (permission needed).
-Children also often serve as alloparents, helping to raise their younger siblings or provide for them through food preparation or collection [@hrdy2009; @meehan2009; @sear2008], despite still requiring substantial amounts of investment themselves [@kramer2018a].
+Compared to chimpanzee mothers, mothers in hunter-gatherer societies have shorter interbirth intervals and produce offspring that require many more years of parental care [@Davison_2021; @Davison_2022].
+Unlike chimpanzee offspring, human offspring in ancestral populations therefore often grew up with multiple older and/or younger siblings who were potential competitors for parental investment and yet who could also provide (or require) allocare [@xx, @yy].
 
@@
 
-Although the fitness prospects of both children and familial caretakers are largely intertwined, evolutionary theory predicts that disagreements about the amount of investment from parents will be ubiquitous [@fouts2005; @godfray1995; @godfray1992; @trivers1974].
-Most resources relevant to fitness outcomes are limited so that one unit of a resource invested in one area cannot be used in another.
-In response to this basic fact of life, humans, like other organisms, are expected to have evolved to efficiently allocate resources between somatic effort (i.e., growth and maintenance) and reproductive effort (i.e., mating and parenting effort) [@daly1983], something which often pits the fitness interests of parents against the fitness interests of offspring.
-
-Conflicts of interest within families are largely structured by the likelihood an individual shares their genes with another.
-As a sexually reproducing species, humans have a 50% chance of passing down a given gene to each child.
-Each child, however, shares more genes through descent with themselves (100%) than with their full siblings (50% on average), half-siblings (25% on average), and stepsiblings (0%).
-Due to this asymmetry, children are often expected to attempt to elicit more investment from parents than is optimal from the perspective of the parent, while parents are expected to attempt to invest less than is optimal for each child in order to invest in their other children, whom parents are expected to value identically, all else being equal [@trivers1974].
+Although the fitness prospects of children and their familial caretakers are largely intertwined, evolutionary theory predicts that disagreements about the amount of investment from parents and other close kin will be ubiquitous [@fouts2005; @godfray1995; @godfray1992; @trivers1974; @bossan2013] because each child shares more genes through descent with themselves (100%) than with their full siblings (50% on average), half-siblings (25% on average), and stepsiblings (0%).
+Due to this asymmetry, children are often expected to attempt to elicit more investment from parents than is optimal from the perspective of the parent, while parents are expected to attempt to invest less than is optimal for each child in order to invest in their other children, whom parents are expected to value identically, all else being equal [@trivers1974; @bossan2013].
 
@@
 
-Potentially cut this paragraph: While parent-offspring conflict and sibling competition are likely ubiquitous, the resources being contested will vary across human groups.
+While parent-offspring conflict and sibling competition are likely ubiquitous, the resources being contested will vary across human groups.
 In relatively egalitarian and food constrained populations, conflict over resources is likely to be centered on food, which probably explains the negative impacts siblings have on child age-specific heights and weights among the Shuar of Ecuador [@hagen2009], although similar effects have also been found in industrialized populations [@lawson2008].
@@
 
-Conflicts of interest between children and parents may also involve how much children invest in the household.
-As previously mentioned, each child is expected to prefer greater investment in oneself than in one’s siblings, while parents, who are equally related to all their children, should often be expected to favor a more even distribution [@trivers1974].
-This can lead to conflict with older children who might prefer to spend their time in ways which are likely to result in future direct fitness benefits (e.g., finding romantic partners or forming strong bonds with friends) rather than taking care of their siblings or acquiring or preparing food [@lowe2003; @syme2023a].
-Conflicts of interest over investment in the household also appear to extend to when a child starts having their own children [@kushnick2023; @emlen1995], with parents potentially benefiting from delaying the reproductive careers of certain offspring in order to gain more care for their other children [@flinn1989; @emlen1992] or speeding up a child’s departure from the household if the child’s presence limits how much caretakers can invest in other children they deem more worthy of investment [@kushnick2023].
-
-## Signaling
-
-In order to elicit more support, children often use signals.
-Signals are a key source of information for individuals in social species and are commonly used to indicate one's value as a mate, one’s formidability to competitors, or one’s worthiness of investment to parents and other social partners [@caro2016; @maynard-smith2004; @preuschoft2000; @setchell2005; @zahavi1975a].
-Signals differ from cues in that signals were favored by natural selection due to their ability to alter the behavior or perceptions of others in ways that benefit the signaler, whereas cues are pieces of information organisms use to influence their behavior or physiological state that did not evolve for communicative effect [@maynardsmith1995].
-Crying is likely a signal of need[^1], for instance, while bleeding would be a cue of need.
-
-[^1]: We expect “signals of need” to actually be signals of resource valuation.
-However, we use “signal of need” here to distinguish signals that involve attempts at eliciting support through revealing private information about one’s vulnerabilities from the more commonly studied signals which reveal private information about one’s positive qualities (i.e., those related to partner choice and intrasexual competition).
-
-When conflicts of interest over investment exist, there may be much to gain from deception (i.e., exaggerating or completely fabricating one’s quality or need) so that one can receive the benefits that come from being seen as having a particular quality without actually doing so [@zahavi1993].
-Although the risks of being deceived can be substantial, signal receivers may still benefit from paying attention to signals so long as the information is credible enough to inform one’s behavior in ways that are adaptive on average [@bliegebird2001; @grafen1990].
-These benefits coupled with the likelihood of strong selection against individuals who respond to fake signals results in the expectation that signals within species will often need to benefit both signaler and receiver to persist evolutionary [@hasson1994; @maynard-smith2004].
-
-Credible signals[\[2\]](#_ftn2) can evolve and be maintained in multiple ways.
-Despite reasonable expectations that such circumstances will be rare [@zahavi1993], signals can be produced and trusted in instances when there are no conflicts of interest over what information is conveyed [@guilford1995; @maynardsmith1995].
-They may also arise when signals cannot be faked (i.e., indexical signals) or in instances when there are greater net payoffs to signalers who have the quality being signaled to a greater extent than others [@maynardsmith1995].
-Although the latter can arise through differential benefits [@penn2020; @tiokhin2016] or a combination of differential costs and benefits (Bergstrom preprint), more attention has been paid to the importance of differential costs alone, with costly signaling being a possible pathway to credible signaling whenever signals are prohibitively costly for those who might fake the signal but still worth making for those with the underlying trait [@maynard-smith2004].
-
-Credible signals work by revealing private information about one’s qualities to others.
+## Cooperation within the family
+
+Humans are a uniquely interdependent species among the living primates [@tomasello2012] in that one’s fitness prospects depend greatly on a network of close kin and other social allies [@apicella2019].
+This need for others is most intense with children, who are born highly altricial and require many years of investment to develop to the point where they can obtain enough food to survive on their own and develop the social skills needed to succeed in their communities [@kaplan2009; @rosenberg2021; @schniter2015].
+Much of this investment comes from mothers and fathers [@kramer2018].
+However, grandparents, aunts, uncles, and non-relatives also provide significant investment in children in many societies [@kramer2018; @sear2008], which was likely favored due to a combination of inclusive fitness benefits, reciprocity, and reputation management [@gurven2000] + (Schniter Preprint (permission needed).
+Children also often serve as alloparents, helping to raise their younger siblings via direct care, food collection and preparation, and other family tasks [@hrdy2009; @meehan2009; @sear2008] despite still requiring substantial amounts of investment themselves [@kramer2018a].
+
+Children’s potential to invest in their families can also cause conflicts with their parents.
+Older children might prefer to spend their time investing in romantic or other social relationships, or learning valuable skills, whereas parents might prefer them to care for their siblings, acquire or prepare food, or complete other household chores [@lowe2003; @syme2023a].
+Conflicts of interest over investment in the family also appear to extend to when a child starts having their own children [@kushnick2023; @emlen1995], with parents potentially benefiting from delaying the reproductive careers of certain offspring in order to gain more allocare for their other children [@flinn1989; @emlen1992] or speeding up a child’s departure from the household so as to provide more parental investment to other children [@kushnick2023].
+
+## Signals of needs and wants
+
+In order to elicit more support, children often use signals of needs and wants.
+Across the tree of life, signals evolve to reveal agents’ private information to other agents when doing so benefits both the signaler and the receiver [@caro2016; @maynard-smith2004; @preuschoft2000; @setchell2005; @zahavi1975a].
+Children often have needs and wants, where needs involve avoiding potential fitness costs, such as poor nutrition, injuries, or threats to status, and wants involve acquiring potential fitness benefits, such as highly nutritious foods and valuable artifacts.
+Needs and wants are often not apparent to caregivers, however, i.e., they are private information, and hence children must signal them to receive help.
+Signals differ from cues in that signals were favored by natural selection due to their ability to alter the behavior or perceptions of others in ways that benefit the signaler, on average, whereas cues are information organisms use to influence their behavior or physiological state that did not evolve for communicative effect [@maynardsmith1995].
+Crying when injured is likely a signal of need[\[1\]](#_ftn1), for instance, whereas bleeding would be a cue of need.
+
+When conflicts of interest over investment exist, there may be much to gain from deception (e.g., crying despite the lack of need) so that one could receive the benefits that come from being seen as needy [@zahavi1993].
+Unchecked, deceptive signals could result in receivers ignoring all signals.
+
+Credible signals can nevertheless evolve and be maintained in multiple ways.
+Signals can be trusted in instances when there are no conflicts of interest over what information is conveyed [@guilford1995; @maynardsmith1995].
+They can also arise when signals cannot be faked (i.e., indexical signals).
+Credible signals can also evolve when the benefits of sending a signal outweigh the costs for individuals with the underlying quality or state, but otherwise the costs outweigh the benefits [@spence1973; @zahavi75; @bergstrom2024].
+Such signals are often termed “costly”, although what is costly is deception, not necessarily signaling per se [for discussion, see @bergstrom2024].
+
 In the context of child signals of need, a child has access to information about their internal state, which is unavailable to others, that allows them to estimate their valuation of additional investment.
@@
 With the information revealed by the signal, caretakers should generally be more likely to provide additional support due to their fitness interest in the child’s wellbeing.
-When this is the case, all parties are better off than they would be without the signal, with caretakers benefitting from the revelation of fitness-relevant information and the child benefitting from the increased likelihood of a beneficial response as long as the benefits outweigh the costs of the signal to child and parent.
-
-However, even if a child was completely successful in conveying their needs to caretakers, it may not be enough to result in a beneficial response.
-This is because belief in one’s child’s need is not the only factor determining the form and amount of support given by parents.
-Instead, parents are expected to weigh the costs and benefits of support with other investment options [@daly1987; @daly1983].
-Ancestrally, this likely often meant evaluating the viability of the child as well as how much a particular resource benefited one child relative to other children or future children [@clutton-brock1991; @daly1983; @dawkins1976; @hagen1999].
+When this is the case, all parties are better off than they would be without the signal, with caretakers benefitting from the revelation of fitness-relevant information and the child benefitting from the increased likelihood of a beneficial response.
+
+However, even credible signals of needs or wants might not result in a beneficial response because belief in a child’s need is not the only factor determining the form and amount of support given by caregivers.
+Instead, caregivers are expected to weigh the costs and benefits of support against other options [@daly1987; @daly1983].
+Ancestrally, this likely often meant evaluating the viability of the child as well as how much a particular resource benefited one child relative to other children, current or future[@clutton-brock1991; @daly1983; @dawkins1976; @hagen1999].
 The former is thought to have resulted in children having to demonstrate their vigor early in life [@soltis2004], while the latter is expected to result in context-specific support [@beaulieu2008; @frankenhuis2022].
@@
 However, if resources are in short supply, parents may benefit more from investing more in older children as they are closer to reaching adulthood and have higher reproductive value [@daly1983].
-Additionally, a parent may also decide against investing in a child despite deeming investment in a child worthwhile if time is not right (e.g., if a mother preferred to breastfeed during rest rather than foraging) [@altmann1980].
+Additionally, a parent may also decide against investing in a child, despite deeming that investment worthwhile, if time is not right (e.g., if a mother preferred to breastfeed during rest rather than foraging) [@altmann1980].
 When this is the case, outward conflicts may exaggerate the extent of a conflict of interest and involve parents teaching children ways to behave that benefit them both [@altmann1980].
+Finally, caregivers often have better information about the costs and benefits of various items and actions, and therefore can make more beneficial decisions for a child than the child can herself (e.g., preventing a child from playing with matches).
+
+Saved for now: Although differing in their details, a common theme is that the indirect fitness costs signalers incur from imposing costs on their that parents and siblings through begging (be them energetic costs, survival costs, or reduced likelihood of future siblings) creates a floor to possible deception but falls short of preventing it [@daly1990; @wells2003].
+
+Throughout human evolutionary history, individuals often found themselves in situations in which they received fitness-relevant benefits from others that could not easily be replaced [@apicella2012; @apicella2019; @hill2011; @tooby1996].
+In these instances, natural selection is expected to have favored those who provided support to others if 1) such help meant one could expect to receive fitness benefits from those helped in the future and 2) the costs of providing help were outweighed by the benefits to the helper on average [@andré2010; @hamilton1964; @tooby1996; @trivers1971].
 
@@
 
-Crying is the most studied way by which children may signal need.
-Crying is a human universal [@gracanin2018] with few suspected non-signaling functions outside of the role vocal cries may play in the development of vocalization [@wells2003] and the functions tears have in terms of cleaning, protecting, and lubricating the eye [@murube2009; @sznycer].
+Crying is the most studied way children signal need.
+Crying is a human universal [@gracanin2018] with few suspected non-signaling functions outside of the role vocal cries might play in the development of vocalization [@wells2003] and the functions tears have in terms of cleaning, protecting, and lubricating the eye [@murube2009; @sznycer].
 Hunger, pain, anger, fear, and separation are all common causes of crying [@vingerhoets2013; @wells2003], although tear production also often happens with joyful events [@sznycer; @vingerhoets2013; @zickfeld].
 
-Infant cries have been found to result in increased amounts of care, resources, and attention they receive and increased proximity to parents [@lin2012].
+Infant cries increase amounts of care, resources, and attention infants receive, as well as increase their proximity to parents [@lin2012].
 For example, Karo Batak infants who cried more were breastfed more, and the length of crying bouts predicted the time mothers breastfed [@kushnick2006].
-Similarly, !Kung San mothers frequently respond to crying infants by providing comfort and breastfeeding access, with comforting behaviors, in general, being far more common than one’s meant to restrict infant behavior [@kruger2010].
-Although this study found mothers were the most common responders, siblings, fathers, and other adults frequently stepped in so that 88% of crying bouts were responded to, with most of the cries not responded to lasting under ten seconds.
-Similar responsiveness has also been reported among the Aka [@hewlett1998], Mbendjele hunter-gatherers [@chaudhary2024], the Efe [@tronick1987], and the Ache, among whom young children have been described as “win\[ning\] most conflicts with parents simply by crying and whining” [@kaplan1987].
-Studies examining the potential for long-term benefits are rare and difficult to achieve; however, during a time of high infant mortality in a sample of 13 Massai infants, those with easy temperaments were more likely to die (5/7) than those with difficult temperaments (1/6), something which is consistent with signals resulting in beneficial treatment [@devries1987].
+Similarly, !Kung San mothers frequently respond to crying infants by providing comfort and breastfeeding access, with comforting behaviors being far more common than behaviors restricting infants [@kruger2010].
+Although this study found mothers were the most common responders, siblings, fathers, and other adults frequently stepped in so that 88% of crying bouts were responded to, with the majority of the cries ignored lasting under ten seconds.
+Similar responsiveness has also been reported among the Aka [@hewlett1998], the Mbendjele [@chaudhary2024], the Efe [@tronick1987], and the Ache, among whom young children have been described as “win\[ning\] most conflicts with parents simply by crying and whining” [@kaplan1987].
+Studies examining the potential for long-term benefits are rare and difficult to achieve; however, during a time of high infant mortality in a sample of 13 Massai infants, those with easy temperaments were more likely to die (5/7) than those with difficult temperaments (1/6), an outcome that is consistent with signals eliciting beneficial treatment [@devries1987].
 
 Crying does not always succeed in eliciting benefits, however.
-Compared to the previously mentioned groups, the Ngandu [@hewlett1998] and the Hadza [@blurtonjones1993] have been found to be more likely to ignore the cries of infants and young children, and concerns over the risk of spoiling children by being too responsive seem to be reasonably common in some Western populations [@barnett2010; @burchinal2010; @chaudhary2024; @frankenhuis2022; @maute2018].
+Compared to the previously mentioned groups, the Ngandu [@hewlett1998] and the Hadza [@blurtonjones1993] have been found to be more likely to ignore the cries of infants and young children, and concerns that being too responsive will spoil children are reasonably common in some Western populations [@barnett2010; @burchinal2010; @chaudhary2024; @frankenhuis2022; @maute2018].
 Crying can also lead to negative responses beyond being ignored [@barr1990; @soltis2004; @zeifman2017].
 It has been found to increase parental distress, the likelihood of maternal depression, and the child’s risk of maltreatment [@frodi1985; @fujiwara2011; @howard2006; @lin2012; @soltis2004].
-For example, in a sample of 3,259 parents of Dutch infants under 6 months old, 5.6% self-reported smothering, slapping, or shaking their babies in response to their crying, outcomes which were more likely among parents who were single, unemployed, or deemed their infant’s crying to be excessive [@reijneveld2004].
+For example, in a sample of 3,259 parents of Dutch infants under 6 months old, 5.6% self-reported smothering, slapping, or shaking their babies in response to their crying, outcomes that were more likely among parents who were single, unemployed, or deemed their infant’s crying to be excessive [@reijneveld2004].
 Excessive infant crying, i.e., colic, has also been found to be a risk factor for multiple negative outcomes [@zeevenhooven2018], including abuse, shaken baby syndrome [@kurth2011], and earlier cessation of breastfeeding [@howard2006].
 
-Although widely accepted as a signal, there is no consensus on why emotional crying may work to elicit positive responses.
-One possibility is that the physiological costs of crying for newborns and infants [@rao1997; @thureen1998] may be prohibitively costly for faking to pay off.
-When this is the case, crying may serve as a credible of vigor [@furlow1997; @lummaa1998; @soltis2004; @wells2003], something which may be particularly important early in life due to increased risk of infanticide for children who have has poor survival prospects, due to either health conditions or resource constraints [@daly1984; @soltis2004].
-
-Potentially combine and condense with paragraph below: As children age, their reproductive value increases due to reduced odds of mortality, leading to a decreased risk of infanticide (Hagen 2004 comment).
-For this reason, there is likely a transition period in which crying is primarily meant to provide information about how much a child values further investment.
-Many models have been put forward which attempt to explain human crying and begging in other species as credible signals of need due to the differential costs and benefits of crying and analogous vocalizations in other species as deterring the exaggeration of need [@godfray1991; @kilner1997; @parker2002].
-Although differing in their details, a common theme is that the indirect fitness costs signalers incur from imposing costs on that parents and siblings through begging (be them energetic costs, survival costs, or reduced likelihood of future siblings) creates a floor to possible deception but falls short of preventing it [@daly1990; @wells2003].
-
-For older children, the relatively insignificant energetic costs of crying may not be enough to result in credible signals in times of substantial conflicts of interest.
-For this reason, crying among older children may be more similar to adult crying, something which has been found to result in increased support and empathy but sometimes come at the cost of reputation and the potential for stigmatization [@cornelius2001; @cretser1982; @hendriks2006; @jesser1989; @vingerhoets2016].
-Such costs may be a means by which crying can still be a credible signal [@sznycer] with moderate conflicts of interest if they dissuade potential fakers from making the signal.
-
-### Withdrawal
-
-Withdrawal has been suggested to be an underappreciated way those with low social power may credibly signal need [@syme2023a].
-Due to their lower status, children have limited ways to change how others treat them.
-However, as children 1) are a major component of parents' inclusive fitness and 2) can help family members through allocare and other forms of labor, they provide benefits that are valuable and hard to replace.
-By withholding these benefits or putting the future of these benefits into question, a child simultaneously delays the benefits they get from this cooperation while imposing costs on those with a fitness interest in the child.
-As these costs are likely lower for those truly in need than those who are not, withdrawal may be a credible way to signal need if the costs to faking the signal make it unprofitable.
-In these cases, the end result is analogous to a labor strike, in which the signaler conveys to others their situation demands more support, and if they want to continue receiving benefits from the child, they need to upregulate their support of the child.
-
-One of the key proximate mechanisms behind withdrawal is sadness or low mood [@scherer1994]\[\[3\]\](#\_ftn3).
+Although widely accepted as a signal, there is no consensus on why emotional crying might elicit positive responses.
+As the costs and benefits of signaling and gaining investment change with age, it is likely the strategic logic of crying changes too.
+It could be vital for newborns to signal health and vigor to reduce the risk of neglect or infanticide experienced by children with poor survival prospects [@daly1984; @soltis2004].
+As crying comes with physiological costs [@rao1997; @thureen1998], it could serve as a credible signal of vigor if these costs are too high for unhealthy infants to bear [@furlow1997; @lummaa1998; @soltis2004; @wells2003].
+Alternatively, crying might impose costs on caregivers by, e.g., attracting predators or interfering with important activities, motivating caregivers to invest more to stop the crying [so-called blackmail or bargaining models; @xx, @yy].
+More generally, contact calls between parents and offspring have evolved in numerous species, so cries might simply be a means for infants to elicit attention from caregivers [@mehr2021].
+For older children, the relatively insignificant energetic costs of crying might not be enough to ensure it is a credible signal when there are conflicts of interest.
+For this reason, crying among older children might be more similar to adult crying, which can increase support and empathy but sometimes also reduces reputation and increases stigmatization [@cornelius2001; @cretser1982; @hendriks2006; @jesser1989; @vingerhoets2016].
+Such costs could be a means by which crying can still be a credible signal of need if they dissuade potential fakers from making the signal [@sznycer] when there are moderate conflicts of interest.
+
+### Sadness
+
 Sadness is a type of negative affect that is generally thought to arise from loss [@horwitz2017; @levenson1999; @smith1993], especially when one has little power to reverse or mitigate the loss on one’s own [@emanuel2023; @gunderson2021].
-Sadness and low mood, like other forms of psychological pain, probably evolved to motivate victims of adversity to shift their attention to the causes of adversity to mitigate its negative fitness consequences and to learn to solve or avoid future such adverse events [@andrews2009; @nesse1990; @thornhill1989]\[\[MG12\]\](#\_msocom_12).
+Sadness and low mood, like other forms of psychological pain, probably evolved to motivate victims of adversity to shift their attention to the causes of adversity to mitigate its negative fitness consequences and to learn to solve or avoid future such adverse events [@andrews2009; @nesse1990; @thornhill1989].
 Low mood, for example, is often thought to be adaptive in situations when activity is dangerous or unlikely to result in benefits, with anhedonia, pessimism, and fatigue all helping reduce behavior that may not be worth the costs or that may cause additional problems [@nesse2019; @nettle2009].
@@
 Such expressions seem to be important to conveying this information as is seen by responses to those who cannot produce certain elements of facial expressions like tear production [@hasson2009; @nesse2019] or certain facial movements [@bogart2012; @bogart2010].
-However, facial expressions, along with tears, may not be enough to credibly signal need in times of significant conflicts of interest.
-In these instances, depression and suicidal behavior have been suggested to function as ways to credibly signal need due to the expectation that such states are costlier for those facing adversity than those who are not (for more detail on these approaches see: [@hagen2003; @rosenthal1993; @syme2016; @watson2002; @gaffney2022].
-
-### Temper tantrums and self-harm
+
+### Temper tantrums
 
@@
 
-As with the vocalizations that come with crying, temper tantrums might motivate caretakers to upregulate their support of the signaler due to the energetic costs of tantrums or the increased risk of predation stemming from the noisy display, both of which might result in credible signals if they are costly enough to not be worth faking [@lummaa1998].
-In addition, tantrums also impose direct costs on caretakers who may be less able to do fitness enhancing tasks [@altmann1980] like finding or preparing food or taking care of other children.
-
-Another possibility is that temper tantrums carry the risk or implication of offspring self-harm.
-For example, Utilian mothers reported in focus groups that some temper tantrums involved children banging their heads on things when they do not get what they want, something that has been reported in other populations [@belden2008; @burrows1953; @daniels2012; @dennis1940; @vandenakker2022].
-Similarly, young chimpanzees (*Pan troglodytes*) have been observed to beat their heads on the ground when refused access to breast milk [@goodall1986], and American white pelican chicks (*Pelecanus erythrorhynchos*) have been found to throw themselves on the ground, flail violently, and bite at their own wings when begging for food [@schaller1964].
-Such behaviors have been suggested to follow similar logic to depression and suicide attempts as they are likely relatively less costly to those who are truly in dire need compared to those who are not to the point faking is not worth the costs [@hagen2008].
-In other words, the risk of injury from violent temper tantrums carries greater fitness costs to a child who is receiving enough support to develop high levels of embodied capital than it does to a child who is neglected to the point where their survival prospects are low.
-In this scenario, the risk of harm is likely not worth the potential benefits of increased support to the child receiving substantial investment, but it may be for the neglected child who may have few other options to compel others to invest in them.
+As with the vocalizations that come with crying, temper tantrums might motivate caretakers to upregulate their support of the signaler due to the energetic costs of tantrums or the increased risk of predation stemming from the noisy display, both of which might result in credible signals if they are costly enough to deter faking [@lummaa1998].
+In addition, tantrums also impose direct costs on caretakers who may be less able to complete important tasks [@altmann1980] like finding or preparing food or taking care of other children.
+
+Another possibility is that temper tantrums carry the risk or implication of self-harm.
+For example, Utilian mothers reported in focus groups that some temper tantrums involved children banging their heads on things when they do not get what they want, behaviors that have been reported in other populations [@belden2008; @burrows1953; @daniels2012; @dennis1940; @vandenakker2022].
+Similarly, young chimpanzees have been observed to beat their heads on the ground when refused access to breast milk [@goodall1986], and American white pelican chicks have been found to throw themselves on the ground, flail violently, and bite at their own wings when begging for food [@schaller1964].
+Such behaviors s could credibly signal need as they are likely relatively less costly to those who are truly in dire need compared to those who are not, thus deterring deceptive signaling [@hagen2008].
 
@@
 
-This study was designed to help establish the natural history of child signaling behavior in Utila, Honduras, by quantifying the frequencies of common child signals of need and various types of parental responses.
-From this, we sought to use signaling theory to examine 1) the causes of child signals of need, 2) how and why signal costs might vary, 3) how parents respond to indications of child need, 4) how signaling behavior changes with age and sex, 5) and if there were differences in child signaling and parental responses two groups who face different levels of social and economic adversity.
+Although infant signals of need, particularly infant cries, have been extensively studied in many populations [@barr1991; @fouts2005; @konner1974; @kruger2010; @kushnick2006; @meehan2017; @müller2021; @tronick1987], the full range of potential child signals of need, which likely extend beyond cries, has never been investigated in children of all ages.
+In particular, how and when older children signal need has rarely been investigated [for one exception, see @syme2023a].
+
+This study aimed to establish the natural history of child signaling in Utila, Honduras, by quantifying the frequencies of common child signals of need and various types of caregiver responses.
+From these, we used signaling theory to explore 1) the causes of child signals of need, 2) how and why signal costs might vary, 3) how parents respond to indications of child need, 4) how signaling behavior changes with age and sex, 5) and if there were differences in child signaling and parental responses in two subpopulations that face different levels of social and economic adversity.
+More specifically, we hypothesized that children have evolved strategies to tailor their signals of need to 1) their valuation of a given resource, 2) the presence of existing information about a child's needs and wants (i.e., potential to benefit from a resource) and 3) the extent of conflicts of interest with caretakers, which should influence signal cost.
+Such strategies can lead to some combination of verbal requests for more support, facial sadness, crying, withdrawal, temper tantrums and self-harm, all of which may be effective in certain circumstances and are sometimes successful in eliciting more support from caretakers, despite a child's low status and formidability.
 
 A major challenge when studying signals of need is the fact that such signaling is expected to 1) be more common in times when individuals can benefit from increased investment (e.g., when they are facing major adversity or when a fitness-relevant opportunity arises) and 2) result in benefits to the signaler on average.
-This hypothisized causal, bi-directional relationship between investment and signaling makes it difficult to isolate the effects of signaling behavior on investment in children when using standard measures of child welfare (i.e., BMI, health metrics, and height for age).
-When experiments are not appropriate, longitudinal studies with carefully selected variables can help address this [@arjas2004].
-However, our small sample size with signaling data and anthropometric data at two time periods limits our ability to do this in this study.
-For this reason, all analyses for this study were exploratory and meant to inform future studies of child signaling.
+This hypothesized causal, bi-directional relationship between investment and signaling makes it difficult to isolate the effects of signaling behavior on investment in children, especially given the lack of prior studies on signaling by older children in low income settings.For this reason, all analyses of these unique data were exploratory and meant to lay the foundation for future studies of child signaling.
 
@@
 
-This study was done in Utila, Honduras, one of the country’s three major Bay Islands in the Western Caribbean.
-Participants in the study largely fit into two groups who face different levels of economic and social adversity, 1) a long-standing English-speaking community primarily made up of the descendants of expats and 2) a relatively new community of Spanish-speaking immigrants from the mainland.
+This study of children aged `{r} paste(range(modeldf$ChildAge), collapse="-")` and their parents and other primary caretakers, was conducted in Utila, Honduras, one of the country’s small Bay Islands in the Western Caribbean.
+Participants in the study were sampled from two populations who face different levels of economic and social adversity: 1) a long-standing English-speaking community primarily comprising descendants of British immigrants who arrived on the island in the 1800’s, and 2) a more recent community of Spanish-speakers, mostly from the mainland.
 Compared to those born on Utila, immigrants from the mainland [@davis2014] are more likely to live in crowded and unsanitary conditions, face active discrimination [@garcia2017], and experience economic uncertainty due the predominance of low paying, seasonal jobs primarily related to the island’s diving tourism industry [@garcia2018; @stonich2000].
-Such inequality appears to be increasing in response to recent growth in the tourism industry, with the majority of wealth from tourism being enjoyed by a small portion of expat islanders and westerners [@davis2014].
-
-During focus groups held in July 2022, parents indicated high levels of sibling competition and various ways children might signal need.
+Such inequality appears to be increasing in response to recent growth in the tourism industry, with the majority of wealth from tourism being enjoyed by a small portion of native English-speaking islanders and other Westerners [@davis2014].
+
+## Focus group interviews
+
+To inform development of our survey questionnaire on child signaling, focus groups of Utilan parents were held in July 2022.
+Parents indicated frequent sibling competition and various ways children might signal need.
 Both Spanish-speaking and English-speaking moms reported that fighting between siblings was extremely common, with fights over toys, gifts, and phone time being the most reported causes.
@@
 In response to these hypothesized signals, discipline was common.
-Although the mothers generally reported that children were treated equally, needier children were described as getting more resources from parents, something that may be partially mediated by differences in signaling among children in addition to information caretakers gained from cues.
+Although the mothers generally reported that children were treated equally, needier children were described as getting more resources from parents, perhaps due to differences in signaling among children as well as information caretakers gained from cues of need.
 
@@
 
-Participants for this study were recruited by a variety of means.
-JH recruited early participants through convenience and snowball sampling.
-Participants were also recruited via Facebook posts and fliers which were sent home with children the Utila's Methodist school.
-
-In all forms of recruitment, adult participants were informed of the purpose of the study, what they would be asked to do, and the inclusion criteria, which included being over 18, residing on Utila, and having a child between 6 and 18 years old.
+Participants for this study were recruited through convenience sampling.
+JH recruited early participants with the help of primary informants and the use of snowball sampling.
+Participants were also recruited via Facebook posts and fliers which were sent home with children attending Utila's Methodist school.
+
+In all forms of recruitment, adult participants were informed of the purpose of the study, what they would be asked to do, and the inclusion criteria, such as being 18 or older, having permanent residence on Utila, and having a child between 6 and 18 years old.
 
@@
 All participants for the survey portion of the study met the inclusion criteria.
-Most participants were mothers of at least one biological child in the study, with the participant being the biological mother of the child in xxx out of xxxx caregiver/child dyads.
-
-Child participants for the anthropometric portion of the study were at least 5 years of age.
-In the majority of instances, child participants had a primary caretaker who took the survey portion of the study first, although some children participated before parents.
-To have their data included in this study, a caretaker in their household must have taken the survey.
+Most adult participants were the biological mothers of one or more children in the study.
+All children in the study, whose ages ranged from 5-20, were cared for by one of the adult participants.
 
@@
 
-All adult participants provided informed consent. Surveys were designed to last about an hour plus 20 minutes for each additional child. Adults were paid \$XXXXX for the survey portion of the study.
+All adult participants provided informed consent.
+The surveys, which included many topics outside of child signaling, were designed to last about an hour, plus 20 minutes for each additional child.
+Adults were paid \$XXXXX for the survey portion of the study.
 
@@
 
-The survey data for this study come from a larger project run by the Utila Child Health Project.
-Data was collected by either Jess Hlay, Izabel Rodriguez, Daniela Mejia, Angela Garcia, or Caroline Smith who read the survey questions to participants and recorded their answers.
+The data for this study were collected as part of the Utila Child Health Project.
 Participants could choose to do the survey in English or Spanish.
+All questions were read to them and recorded by a research team member.
 
@@
 
-We used 6-point Likert scales to get caregiver reports on the frequencies of 1) various types of child signaling, 2) interpersonal conflict with each participating child, 3) and child alloparenting behavior.
+We used 6-point Likert scales to get caregiver reports on the frequencies of 1) various types of child signaling, such as sadness and crying, 2) conflict between the caretaker and each participating child, 3) and child allocare of other children in their household.
 See @tbl-freqvars for exact wording.
 Caretakers were often not asked about child alloparenting behavior if a child was an only child.
-For this reason, we gave children an alloparenting frequency of never if they were the only child within a household.
+For this reason, we gave children an alloparenting frequency of “never” if they were the only child within a household.
 
 |  |  |
-|--------------------------|---------------------------------------------|
+|--------------------------|----------------------------------------------|
 | Variable | Survey question wording |
@@
 
-: Variable names and question wording for measures about the frequency of behaviors.
+: Variable names and question wording for measures about the frequency of signaling, allocare, and conflict.
 All questions had the following labels attached to each point on its Likert scale: Never; Once a month or less; More than once a month but less than once a week; More than once a week but not daily; Daily; Multiple times per day.
+For analyses, we converted these categorical variables to numeric frequencies per month as follows: never = 0; once a month or less = 1; more than once a month = 3; more than once a week = 8; daily = 30; multiple times per day = 60.
 {#tbl-freqvars}
 
-As these responses reflect frequencies of behavior, we converted these to counts which reflected the number of times per month a child engaged in these behaviors for analyses.
-The original wording of the Likert scale and the frequency conversion are shown in @tbl-convertedfreqs.
+From these count variables we created three composite signaling variables to be used as outcome variables.
+The *frequency of signaling* was the sum of the frequencies of child sadness, crying, and temper tantrums.
+*Maximum signal frequency* was the frequency of the most common of these three signals.
+*Signal cost* was the sum of the frequencies of these three signals weighted by their relative costliness (i.e., *Signal cost* = *frequency of sadness* + 2 $\times$ *frequency of crying* + 3 $\times$ *frequency of temper tantrums*).
+
+We also calculated the summed frequency of child alloparenting contributions by other children within the household.
+
+### Relative neediness of children and relative investment in children within the family
+
+We used 6-point Likert scales for parental reports of 1) the relative neediness of each participating child, and 2) the relative investment they provide to each participating child compared to the rest of their children.
+See Table @stbl-relvars for exact wording.
 
 |  |  |
-|------------------------------------|------------------------------------|
-| Likert scale wording                             | Count value (times per month) |
-| Never                                            | 0                             |
-| Once a month or less                             | 1                             |
-| More than once a month but less than once a week | 3                             |
-| More than once a week but not daily              | 8                             |
-| Daily                                            | 30                            |
-| Multiple times per day                           | 60                            |
-
-: The conversions from Likert scale to count value for the frequencies of child behaviors included in this study.
-Variables created with this scheme include the frequencies of sadness, crying, temper tantrums, conflict with primary caretakers, and child alloparenting effort.
-{#tbl-convertedfreqs}
-
-From these count variables, we created three additional signaling variables to be used, primarily, as our main dependent variables in analyses.
-The frequency of signaling was the sum of the frequencies of child sadness, crying, and temper tantrums.
-Maximum signal frequency was the frequency of the most common of these three signals.
-Signal cost was the sum of the frequencies of these three signals weighted by their relative costliness (i.e., Signal cost = frequency of sadness + frequency of crying \* 2 + frequency of temper tantrums \* 3).
-Although we collected data on the frequency of running away, it was not included in the creation of any of these variables due to uncertainty over whether it was likely to be a credible signal due to the relative safety of Utila compared to other areas in which running away has been found to be a successful strategy for children.
-
-We also calculated the summed frequency of child alloparenting contributions by other children within the household.
-
-### Relative neediness of children and relative investment in children within the family
-
-We also used 6-point Likert scales to get parental reports of 1) the relative neediness of each participating child and 2) the relative investment they provide to each participating child compared to the rest of their children.
-See Table @tbl-relvars for exact wording.
-
-|                     |                                                                                                  |
-|--------------------------|---------------------------------------------|
+|--------------------------|----------------------------------------------|
 | Variable | Survey question wording |
@@
 Both questions had the following labels attached to each point on its Likert scale: Much less; Less; The same amount; More; Much more.
-{#tbl-relvars}
-
-### Causes and responses to the last instance of child signaling
+{#stbl-relvars}
+
+### Causes of, and responses to, the last instance of child signaling of need
 
 We asked caretakers about the last time each participating child was sad or cried.
-This involved asking what the caretaker thought caused the emotion or display and recording their answer in a text box.
-We then asked how the parent responded to their child.
-The researcher giving the survey would then bin responses into at least one of 31 categories which included positive, negative, and neutral responses (from the child's perspective).
+Caretakers provided a free response of the reason(s) for the emotional display, and their responses to it.
+The researcher then binned responses into at least one of 31 categories which included positive, negative, and neutral responses (from the child's perspective).
 For the full list of options see Table S1.
-
-From this data, we created a parental response variable, which was coded as 1 for a positive response, 0 for a neutral response, and -1 for a negative response (all from the child's perspective).
-MG also coded each case for whether or not the cause had involved: 1) the child being asked to do an unwanted task, 2) punishment, 3) conflict within the family, and 4) discomfort, pain, injury, or illness.
-
-### Perceptions of child susceptibility to illness (or current susceptibility to illness)
+We then created a *caregiver response* variable, which was coded as 1 for a positive response (e.g., helping child), 0 for a neutral response (e.g., not responding), and -1 for a negative response (e.g., punishment; all from the child's perspective).
+MG also coded the causes of each as follows: 1) the child was asked to do an *unwanted task*, 2) *punishment*, 3) *family conflict*, and 4) *discomfort, pain, injury, or illness*.
+
+### Caretaker perceptions of child’s ability to avoid illness
 
 Caretakers were also asked three questions using a five point Likert scale about their child's susceptibility to illness.
-We inverted the scores on two measures so that a score of 5 corresponded with the lowest susceptibility to illness before we averaged the scores of these to create a mean illness susceptibility variable.
-In the `r sum((!is.na(modeldf$IllnessSusceptibility2)) & is.na(modeldf$IllnessSusceptibility3))` instance in which we only data for 2 of the measures, we still calculated the mean but used a numerator of 2.
-See Table @tbl-healthvars for the wording of these variables.
-
-|                        |                                                                                       |
-|--------------------------|---------------------------------------------|
-| Variable               | Survey question wording                                                               |
-| IllnessSusceptibility1 | If an illness is 'going around', my child will get it.                                |
-| IllnessSusceptibility2 | In general, my child is very susceptible to colds, flu, and other infectious diseases |
-| IllnessSusceptibility3 | My child’s immune system protects him/her from most illnesses that other kids get.    |
-
-: Variable names and question wording for measures about a child's current likelihood of becoming sick.
-All three questions were asked using a Likert scale from 1-5 where 1 was labeled "Strongly disagree" and 5 was labeled "Strongly agree." The scores for IllnessSusceptibility1 and IllnessSusceptibility2 were inverted so that 5 corresponded with the lowest susceptibility to illness.
-{#tbl-healthvars}
+We reversed the scores on two measures so that a score of 5 corresponded with the lowest susceptibility to illness before we averaged the scores of these to create a mean *illness avoidance* variable.
+In the `{r} sum((!is.na(modeldf$IllnessSusceptibility2)) & is.na(modeldf$IllnessSusceptibility3))` instance in which we only had data for 2 of the measures, we still calculated the mean but used a numerator of 2.
+See @stbl-healthvars for the wording of these variables.
 
@@
 
-The survey also included many demographic questions about the survey taker and their households.
-Those used for this study included the (1) ages of those in the household, (2) sexes of those in the household, (3) partnership status of the primary caregiver, (4) number of adults who provide childcare within the household, (5) highest level of education of the primary caregiver, (5) immigration status (in regards to Utila), and (6) monthly household income.
-This data allowed us to calculate the number of children (total, younger, and older) and adults within the household.
+The survey also included many demographic questions about the caretakers and their households.
+Those used in this study included the (1) ages of those in the household, (2) sexes of those in the household, (3) partnership status of the primary caregiver, (4) number of adults who provide childcare within the household, (5) highest level of education of the primary caregiver, (5) immigration status (in regards to Utila), and (6) monthly household income.
+These data allowed us to calculate the number of children (total, younger, and older) and adults within the household.
 
@@
 
-All anthropometric measurements were collected in accordance with anthropometric standards [@lohman1988] and involved taking the mean of two measurements.
+All anthropometric measurements were taken twice, in accordance with anthropometric standards [@lohman1988], and then averaged.
 Before being measured, participants were asked to remove their shoes and heavy articles of clothing.
-Height was measured to the nearest millimeter using an anthropometer, and weight was measured to the nearest 100 g with a digital scale (model name) that used within a building with a concrete foundation.
+Height was measured to the nearest millimeter using an anthropometer, and weight was measured to the nearest 100 g with a digital scale (model name) that was used in a building with a concrete foundation.
 All participants were wearing lightweight clothing, so we made no adjustments for clothing weight.
@@
 Triceps and subscapular skinfold thicknesses were measured using skinfold calipers.
-Measurements were taken directly on the skin.
 The triceps measurement was taken at the midpoint of the upper arm.
@@
 For this measure participants were asked to sit/stand?
-an position their upper arm perpendicular to their body with their elbow being help at a 90° angle, before flexing their clenched fist towards themselves as much as they could.
-
-We computed each child's height, weight, and BMI Zscores with respect to the CDC child growth standards [@kuczmarski2002].
+and position their upper arm perpendicular to their body with their elbow being held at a 90° angle, before flexing their clenched fist towards themselves as much as they could.
+
+We computed each child's height, weight, and BMI Z-scores with respect to the CDC child growth standards [@kuczmarski2002].
 
 For our independent variables of skinfold thickness, flexed bicep circumference, and grip strength on signaling behavior, we used the residuals from generalized additive models which modeled the effect of sex and age on each measurement using the mgcv package [@wood2011].
-The resulting variables which represent how each child differed from the average child for their given sex and age.
-We also centered and scaled the skinfold thickness residuals before summing them to create an overall body fat variable.
+The resulting variables represent how each child differed from the average child for their given sex and age.
+We also centered and scaled the skinfold thickness residuals before summing them to create a body fat index.
 
@@
 All statistical analyses were exploratory and involved fitting many models with different combinations of independent and dependent variables.
-While were cognizant of the risk of overfitting the data, it is still more likely than if the study was entirely confirmatory [@tukey1977].
-The payoff to this approach is the identification of theoretically relevant effects that should be tested in the future.
-
-We fit generalized linear mixed models for exploratory analyses with signal frequencies and summed signal cost as dependent variables.
-As our signaling variables involved over-dispersed count data, we used negative binomial regression.
-This was done in R with the GLMM package [@brooks2017], with household as a random effect.
-Control variables differed in some models, but child age and sex were always included.
-Specific control variables are mentioned in the results.
-
-For the models examining 1) parental responses to the last reported time each participating child was sad or cried, 2) the relative neediness of the child, and 3) the investment a child received compared to the other children in the household, we fit ordinal logistic regression models using two methods.
-The first used the polr function in the MASS package [@venables2002].
-The results were then interpreted with clustered standard errors for household [@zeileis2002].
-Due to the risk of overfitting that comes with having too few clusters [@cameron2015], we also fit the same models using BRMS [@bürkner2017] with household as a random effect.
+Whereas confirmatory studies test prespecified hypotheses, exploratory analyses prioritize discovery: they risk overfitting the data, generating false positives, but can reveal unexpected yet potentially important relationships that warrant future investigation [@tukey1977].
+
+Child signaling frequencies and signaling costs, which were overdispersed count variables, were analyzed using generalized linear mixed models with a negative binomial error family, fit in R with the glmmTMB package [@brooks2017].
+Unless otherwise noted, explanatory variables included child age, child sex,, child illness avoidance, caretaker education, caretaker partner status, household income, the number of other children in the household, the number of the adults in the household, the frequency of conflict between the primary caretaker and the child, and an interaction between the frequency of child alloparenting effort and child sex..
+A household random effect was included in all models.
+
+Parental responses to child signaling, their perceptions of child need, and their reported investment in the child were coded as three-level ordinal variables, either negative, neutral or positive; or less than other children, same as other children, or more than other children.
+We analyzed these outcomes using ordinal logistic regression, fit in R using the polr function in the MASS package [@venables2002], with clustered standard errors by household [@zeileis2002].
 
@@
 
-`r dplyr::n_distinct(modeldf$householdID)` primary caretakers took the survey and provided signaling data for at least one child.
-`r nrow(modeldf)` children had frequency data for at least 1 signal, with `r nrow(modeldf_FULLSIG)` having data for all 3.
-Among those with at least one record of signaling frequency, `r sex_table[1]` were female and `r sex_table[2]` were male.
-Anthropometric and signaling data exist for `r nrow(modeldf_FULLANTHROPOMETRICS)` children, of whom `r sex_table2[1]` were female and `r sex_table2[2]` were male.
+`{r} dplyr::n_distinct(modeldf$householdID)` primary caretakers (females = `{r} table(caregivers$CaregiverSex)[1]`, males = `{r} table(caregivers$CaregiverSex)[2]`, non-binary = `{r} table(caregivers$CaregiverSex)[3]`) took the survey and provided signaling data for at least one child.
+`{r} nrow(modeldf)` children had frequency data for at least one signal, with `{r} nrow(modeldf_FULLSIG)` having data for all three.
+Among those with at least one record of signaling frequency, `{r} sex_table[1]` were female and `{r} sex_table[2]` were male.
+Anthropometric and signaling data exist for `{r} nrow(modeldf_FULLANTHROPOMETRICS)` children, of whom `{r} sex_table2[1]` were female and `{r} sex_table2[2]` were male.
 For summary statistics of child-level variables see @tbl-sumtab_children.
@@
 #| label: tbl-sumtab_children
-#| tbl-cap: Summary statistics for child study variables.
-vars_c <- c(
-  ChildAge = "Age (years)",
-  SadFreqN = "Sad frequency (times per month)",
-  CryFreqN = "Crying frequency (times per month)",
-  TantrumFreqN = "Temper tantrum frequency (times per month)",
-  AlloparentingFreqN = "Frequency of child's own alloparenting effort (times per month)",
-  ConflictFreqN = "Frequency of conflict with focal caretaker (times per month)",
-  SignalFreq = "Frequency of signals summed (times per month)",
-  SignalCost = "Signal cost",
-  SignalFreqMax = "Frequency of most common signal (times per month)",
-  # IllnessSusceptibility1 = "Illness Susceptibility1",
-  # IllnessSusceptibility2 = "Illness Susceptibility2",
-  # IllnessSusceptibility3 = "Illness Susceptibility3",
-  IllnessSusceptibilityMean = "Illness avoidance index",
-  HeightZ = "Height Z-score (CDC)",
-  WeightZ = "Weight Z-score (CDC)",
-  BMIZ = "BMI Z-score (CDC)",
-  SubscapMean = "Subscapular skinfold thickness (mm)",
-  SubscapR = "Subscapular skinfold residuals",
-  TricepMean = "Tricep skinfold thickness (mm)",
-  TricepR = "Tricep skinfold residuals",
-  BodyFat =  "Body fat composite"
-)
-
-kable(custom.summarize(modeldf, vars_c, facvar = "Sex")) %>% kable_styling()
-
-#kable(custom.summarize(d2, vars), caption = "Summary statistics for child study variables.") %>% kable_styling()
-
+#| tbl-cap: Summary statistics for child variables.
+
+kable(custom.summarize(modeldf, longform_dict[1:18])) %>% kable_styling()
 ```
@@
 #| label: tbl-sumtab_adults
-#| tbl-cap: Summary statistics for adult and household study variables.
-
-vars_c <- c(
-  CaregiverAge = "Caregiver age (years)",
-  EducationLevelYears = "Caregiver education level (years)",
-  IncomeCategoryN = "Monthly household income (Lempira)",
-  number_adults = "Adults in household",
-  AdultsChildcare = "Household adults who provide childcare"
-)
-
-kable(custom.summarize(caregivers, vars_c)) %>% kable_styling()
-
-#kable(custom.summarize(d2, vars), caption = "Summary statistics for child study variables.") %>% kable_styling()
-
+#| tbl-cap: Summary statistics for adult and household variables.
+
+kable(custom.summarize(caregivers, longform_dict[19:23])) %>% kable_styling()
 ```
@@
 Conflict between children and caretakers was also common, with `{r} fmt_sum(modeldf$ConflictFreqN >= 8)` having conflicts with parents more than once a week.
-As with the frequencies of various types of signaling, `{r} fmt_sum(modeldf$ConflictFreqN == 0)` were described as never having conflicts with parents.
+However, `{r} fmt_sum(modeldf$ConflictFreqN == 0)` were described as never having conflicts with parents.
 See @fig-barplot_ConflictFreq.
 
-The frequency of alloparenting effort was bimodally distributed, with `{r} fmt_sum(modeldf$AlloparentingFreqN == 0)` being reported to never engage in it (`r sum(na.omit(modeldf$NumberOfChildren==1))` instances were due to a child being an only child) and `{r} fmt_sum(modeldf$AlloparentingFreqN >= 30)` being reported to help with childcare at least once a day.
+The frequency of alloparenting effort was bimodally distributed, with `{r} fmt_sum(modeldf$AlloparentingFreqN == 0)` reported to never engage in it (`r sum(na.omit(modeldf$NumberOfChildren==1))` instances were due to a child being an only child) and `{r} fmt_sum(modeldf$AlloparentingFreqN >= 30)` reported to help with childcare at least once a day.
 See @fig-barplot_AlloparentingFreq.
@@
 
-Stacked bar plot showing the frequencies of child sadness, crying, and temper tantrums per month as reported by primary caretaker.
+Frequencies of child sadness, crying, and temper tantrums per month as reported by the primary caretaker.
 Frequency key: 0 corresponds to "never," \<=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and \>30 corresponds to "multiple times per day."
@@
 
+::: {#fig-signal_corrplot}
+```{r}
+signal_corrplot
+```
+:::
+
 ::: {#fig-signal_alluvial_plot}
@@
 
-Alluvial plot showing the combinations of signaling frequencies per month.
-Alluvium width equals the number of individiuals with the same frequencies for the three signals/states.
+Combinations of signaling frequencies per month.
+Alluvium width equals the number of individuals with the same frequencies for the three signals/states.
 Stratum height equals the count of individuals at a given frequency.
@@
 
-Stacked bar plot showing the frequency of conflict (per month) between child and primary caregiver as reported by primary caretaker.
+Frequency of conflict (per month) between child and primary caregiver as reported by primary caretaker.
 Color corresponds to sex of the child.
@@
 
-Stacked bar plot showing the frequency of a child's alloparenting effort (per month) as reported by primary caretaker.
+Frequency of a child's alloparenting effort (per month) as reported by primary caretaker.
 Color corresponds to sex of the child.
@@
 
-For every signaling measure except sadness (which also involves non-signaling functions), there was a statistically significant negative relationship between the age of the child and the frequency of signaling when controlling for the sex of the child, household demographics, the frequency of conflict with one's primary caretaker, caretaker perceptions of child susceptibility to illness, and an alloparenting effort by sex interaction.
-See @tbl-main_models for regression statistics and @fig-effect_age for effects plots.
-
-Age had largest effect on crying (`r stats$mcryH$ChildAge$str`), with the model predicting crying frequency dropping to half of the age 5 value by around age 8 (visual inspection).
-The effect of age on temper tantrums (`r stats$mtantrumH$ChildAge$str`) was smaller with a more linear slope.
-
-```{r}
-#| label: tbl-main_models
-#| tbl-cap: Regression tables for our main exploratory models.
-r_vars <- c(
-  ChildAge = "Age (years)",
-  SexMale = "Sex (male)",
-  OtherChildrenHH = "Other children",
-  LogIncome = "Income (log)",
-  number_adults = "Adults in home",
-  PartnerStatusUnpartnered = "Partner status (single)",
-  EducationLevelYears = "Parent ed. (years)",
-  ConflictFreqN = "Conflict freq.",
-  SignalFreq = "Signal freq.",
-  SignalCost = "Signal cost",
-  SignalFreqMax = "Max signal freq.",
-  IllnessSusceptibilityMean = "Percieved illness susceptibility",
-  SadFreqN = "Sadness freq.",
-  CryFreqN = "Crying freq.",
-  TantrumFreqN = "Tantrum freq.",
-  AlloparentingFreqN = "Child allocare"
-)
-modelsummary(list(Sadness = msadH, Crying = mcryH, Tantrum = mtantrumH, SignalFreq = msfH, SignalCost = mscH, SignalFreqMax = mmsfH), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
-```
+As children mature, they are more capable of fulfilling their own needs, more able to ask others for what they need, and are less vulnerable to threats.
+We therefore expected signaling frequencies to decrease with age.
+For every signaling measure except sadness (which also has non-signaling functions), there was a statistically significant negative relationship between the age of the child and the frequency of signaling.
+See @stbl-main-models and @fig-effect_age.
+
+Age had the largest effect on crying (`{r} stats$mcryH$ChildAge$str`), with the model predicting that by age 8 crying frequency drops to half of its value at age 5 (visual inspection).
+The effect of age on temper tantrums (`{r} stats$mtantrumH$ChildAge$str`) was smaller with a more linear slope.
 
@@
 Rug corresponds to actual values of child age with a small amount of jitter to reveal overlapped values.
-See @tbl-main_models for the corresponding regression table.
+See @stbl-main-models for the corresponding regression table.
 :::
@@
 
-There were statistically significant positive relationships between the frequency of conflict between a child and their primary caretaker and all of our signaling measures when controlling for the sex of the child, the age of the child, household demographics, caretaker perceptions of child susceptibility to illness, and an alloparenting effort by sex interaction.
-See @tbl-main_models for regression statistics @fig-effect_conflict for effects plots.
+Costly signals of need are necessary when there are conflicts of interest.
+There were statistically significant positive relationships between the frequency of conflict between a child and their primary caretaker and all of our signaling measures (although frequent signaling might cause caretaker perceptions of conflict; see the Discussion section).
+See @stbl-main-models and @fig-effect_conflict.
 
@@
 
-The effect of the frequency of conflict with one's primary caretaker on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior, all per month.
+The effect of the frequency of conflict with one's primary caretaker on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and signal cost, all per month.
 Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
 Rug corresponds to actual values of the frequency of conflict between child and primary caretaker with a small amount of jitter to reveal overlapped values.
-See @tbl-main_models for the corresponding regression table.
+See @stbl-main-models for the corresponding regression table.
 :::
@@
 ## Signaling behavior is negatively associated with caregiver education
-
-There were statistically significant negative relationships between years of caregiver education and all of our signaling measures except sadness (`r stats$msadH$EducationLevelYears$str`) when controlling for the sex of the child, the age of the child, household demographics, caretaker perceptions of child susceptibility to illness, and an alloparenting effort by sex interaction.
-See @tbl-main_models for regression statistics @fig-effect_edu for effects plots.
+[delete this section if we remove conflict from the main model? or mention on significant when controlling for conflict]
+
+We did not predict relationships between child signaling and our demographic control variables, but we did find statistically significant negative relationships between years of caregiver education and all of our signaling measures except sadness (`{r} stats$msadH$EducationLevelYears$str`).
+See @stbl-main-models s and @fig-effect_edu.
 
@@
 
-The effect of years of caregiver education on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior, all per month.
+The effect of years of caregiver education on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and signal cost, all per month.
 Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
 Rug corresponds to actual values of primary caretaker education (years) with a small amount of jitter to reveal overlapped values.
-See @tbl-main_models for the corresponding regression table.
+See @stbl-main-models for the corresponding regression table.
+:::
+
+## Signaling behavior and child alloparenting effort by sex
+
+High alloparenting effort by children is associated with conflict with their parents (Syme and Hagen, xx).
+Across cultures, girls do more alloparenting than boys (xx, yy).
+We therefore investigated the association of child alloparenting effort and child signaling, with an interaction term for child sex.
+For each signaling measure, there was also a significant interaction between how often a child cared for their siblings and the sex of the child.
+The slope of these relationships were positive for girls and flat or slightly negative for boys.
+In other words, our models predict that as children provide more allocare within their families, the girls’ frequency of signaling increases with increases in their alloparenting effort, and boys’ frequency of signaling remains at the same rate or decreases slightly with increases in their alloparenting effort.
+See @stbl-main-models and @fig-effect_allo_by_sex.
+
+::::::::::: {#fig-effect_allo_by_sex}
+```{r}
+signaling_data_plot_allo_by_sex
+```
+
+The effect of an alloparenting effort by sex interaction on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior, all per month.
+Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
+Rug corresponds to actual values of child alloparenting frequency with a small amount of jitter to reveal overlapped values.
+See @stbl-conflict-dep-var for the corresponding regression table.
 :::
@@
 
-The frequency of conflict between children and caretakers is negatively associated with child age (`r stats$mconflict$ChildAge$str`) and the number of adults who provide childcare in a household (`r stats$mconflict$AdultsChildcare$str`) when controlling for the sex of the child, household demographics, the number of adults who provide and do not contribute to childcare, an alloparenting effort by sex interaction, and the summed frequency of alloparenting from all other children in the household.
-See @tbl-conflict_dep_var for the corresponding regression table and @fig-effect_conflict_dep_var for effects plots.
-
-There was also a negative relationship between conflict frequency and the number of older children, however, this was not statistically significant (`r stats$mconflict$OlderKids$str`).
-Similarly, the model predicted conflict to be more common between caretakers and males, but this also failed to reach statistical significance (`r stats$mconflict$SexMale$str`).
+Because caretaker perceptions of conflict predicted child signaling frequency, we investigated predictors of conflict.
+The frequency of conflict between children and caretakers is negatively associated with child age (`{r} stats$mconflict$ChildAge$str`) and the number of adults who provide childcare in a household (`{r} stats$mconflict$AdultsChildcare$str`) when controlling for the sex of the child, household demographics, the number of adults who provide and do not contribute to childcare, an alloparenting effort by sex interaction, and the summed frequency of alloparenting from all other children in the household.
+See @stbl-conflict-dep-var and @fig-effect_conflict_dep_var.
 
@@
 Rug corresponds to actual values of A) child age and B) the number of adults in the household that contribute to childcare with a small amount of jitter to reveal overlapped values.
-See @tbl-conflict_dep_var for the corresponding regression table.
-:::
-
-```{r}
-#| label: tbl-conflict_dep_var
-#| tbl-cap: Regression table for the model examining the frequency of conflict between child and caretaker as the dependent variable.
-r_vars <- c(
-  ChildAge = "Age (years)",
-  SexMale = "Sex (male)",
-  OtherChildrenHH = "Other children",
-  LogIncome = "Income (log)",
-  number_adults = "Adults in home",
-  PartnerStatusUnpartnered = "Partner status (single)",
-  EducationLevelYears = "Parent ed. (years)",
-  ConflictFreqN = "Conflict freq.",
-  SignalFreq = "Signal freq.",
-  SignalCost = "Signal cost",
-  SignalFreqMax = "Max signal freq.",
-  IllnessSusceptibilityMean = "Percieved illness susceptibility",
-  SadFreqN = "Sadness freq.",
-  CryFreqN = "Crying freq.",
-  TantrumFreqN = "Tantrum freq.",
-  AlloparentingFreqN = "Child allocare",
-  TricepR = "Tricep skinfold (residuals)",
-  SubscapR = "Subscapular skinfold (residuals)",
-  BodyFat = "BodyFat (composite)",
-  RelativeNeed3 = "Relative need (w/in household)",
-  UnwantedTask = "Unwanted task",
-  Punishment21 = "Punishment",
-  Family21 = "Family conflict",
-  DiscomfortPainInjuryIllness = "Discomfort/Pain/Injury/Illness",
-  CaregiverResponse = "Caregiver Response",
-  OlderKids = "Older children",
-  YoungerKids = "Younger children",
-  AdultsNoChildcare = "Adults that do not provide childcare",
-  AdultsChildcare = "Adults that provide childcare",
-  OtherChildAlloparentingFreqN = "Other household child alloparenting freq."
-)
-modelsummary(mconflict, gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
-```
-
-## Interaction between child sex and child alloparenting
-
-For each signaling measure, there was also a significant interaction between how often a child cared for their siblings and the sex of the child when controlling for the age of the child, the sex of the child, household demographics, the frequency of conflict with one's primary caretaker, and caretaker perceptions of child susceptibility to illness.
-The slope of these relationships were positive for girls and flat or slightly negative for boys.
-In other words, our models predict that as children provide more allocare within their families, the frequency of signaling would increase with alloparenting effort for girls and remain at the same rate or decrease slightly with alloparenting effort for boys.
-See @tbl-main_models for the corresponding regression table and @fig-effect_allo_by_sex for effects plots.
-
-::: {#fig-effect_allo_by_sex}
-```{r}
-signaling_data_plot_allo_by_sex
-```
-
-The effect of an alloparenting effort by sex interaction on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior, all per month.
-Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
-Rug corresponds to actual values of child alloparenting frequency with a small amount of jitter to reveal overlapped values.
-See @tbl-conflict_dep_var for the corresponding regression table.
-:::
-
-## Positive association between child alloparenting effort and number of older children
-
-There was a statistically significant negative relationship between the number of older children in a household and the frequency of child alloparenting effort `r stats$malloparenting$OlderKids$str` when controlling for the sex of the child, the number of younger children, household income (log), the number of older children, the number of adults who provide childcare within the household, and the summed frequency of alloparenting from all other children in the household.
-As our model could not handle high values of the summed frequency of other child alloparenting, we removed `r sum(d2$OtherChildAlloparentingFreqN > 121, na.rm = TRUE)` outliers (`r sum(d2$OtherChildAlloparentingFreqN == 360, na.rm = TRUE)` of which had a value of 360 times per month and `r sum(d2$OtherChildAlloparentingFreqN == 420, na.rm = TRUE)` of which had a value of 420 timers per month) despite expecting these values to not be outliers in the broader Utilian community.
-See @tbl-alloparenting_dep_var for the corresponding regression table and @fig-effect_alloparenting_dep_var_plot for effects plots.
+See @stbl-conflict-dep-var for the corresponding regression table.
+:::
+
+## Negative association between child alloparenting effort and number of older children
+
+\[DELETE THIS SECTION OR MOVE TO SI???\]
+
+Because child allparenting effort was associated with child signaling, we investigated predictors of child alloparenting effort.
+We expected that children’s alloparenting effort would be positively associated with their age, female sex, and the number of younger siblings, and negatively with the number of older siblings and the total alloparenting effort of others in the household.
+Initial linear modeling of child alloparenting effort revealed that a few households with exceptionally high levels of total child alloparenting effort produced models that predicted extremely unrealistic levels of child alloparenting effort.
+We therefore removed `{r}`r sum(d2$OtherChildAlloparentingFreqN > 121, na.rm = TRUE)` outliers (`{r} `r sum(d2$OtherChildAlloparentingFreqN == 360, na.rm = TRUE)`of which had a value of 360 times per month and`{r} `r sum(d2$OtherChildAlloparentingFreqN == 420, na.rm = TRUE)` of which had a value of 420 timers per month) despite expecting these values to not be outliers in the broader Utilian community.
 
@@
 ```{r}
-# alloparenting_dep_var_plot
+alloparenting_dep_var_plot
 ```
@@
 Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate).
-See @tbl-alloparenting_dep_var for the corresponding regression table.
-:::
-
-```{r}
-#| label: tbl-alloparenting_dep_var
-#| tbl-cap: Regression table for the model examining the frequency of conflict between child and caretaker as the dependent variable.
-r_vars <- c(
-  ChildAge = "Age (years)",
-  SexMale = "Sex (male)",
-  OtherChildrenHH = "Other children",
-  LogIncome = "Income (log)",
-  number_adults = "Adults in home",
-  PartnerStatusUnpartnered = "Partner status (single)",
-  EducationLevelYears = "Parent ed. (years)",
-  ConflictFreqN = "Conflict freq.",
-  SignalFreq = "Signal freq.",
-  SignalCost = "Signal cost",
-  SignalFreqMax = "Max signal freq.",
-  IllnessSusceptibilityMean = "Percieved illness susceptibility",
-  SadFreqN = "Sadness freq.",
-  CryFreqN = "Crying freq.",
-  TantrumFreqN = "Tantrum freq.",
-  AlloparentingFreqN = "Child allocare",
-  TricepR = "Tricep skinfold (residuals)",
-  SubscapR = "Subscapular skinfold (residuals)",
-  BodyFat = "BodyFat (composite)",
-  RelativeNeed3 = "Relative need (w/in household)",
-  UnwantedTask = "Unwanted task",
-  Punishment21 = "Punishment",
-  Family21 = "Family conflict",
-  DiscomfortPainInjuryIllness = "Discomfort/Pain/Injury/Illness",
-  CaregiverResponse = "Caregiver Response",
-  OlderKids = "Older children",
-  YoungerKids = "Younger children",
-  AdultsNoChildcare = "Adults that do not provide childcare",
-  AdultsChildcare = "Adults that provide childcare",
-  OtherChildAlloparentingFreqN = "Other household child alloparenting freq."
-)
-modelsummary(malloparenting, gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
-```
-
-## Summed signal cost, crying frequency, and tantrum frequency are negatively associated with body fat
-
-There were significant negative relationships between summed signal cost (`r stats$mBFc$BodyFat$str`) crying frequency (`r stats$mBFcry$BodyFat$str`), and tantrum frequency (`r stats$mBFtantrum$BodyFat$str`) and a scaled composite of subscapular and tricep skinfold thicknesses when controlling for the sex of the child, the age of the child, the frequency of conflict with one's primary caretaker, household demographics,caretaker perceptions of child susceptibility to illness, an alloparenting effort by sex interaction and a body fat by sex interaction.
-See @tbl-bodyfat_models for regression statistics and @fig-effect_bodyfat for effects plots.
-Subscapular and tricep skinfold thicknesses were also significant predictors of summed signal cost and tantrum frequency when included in the same models as the sole measure of body fat.
-See xxxxSI TABLESxxxx for regression tables.
-
-Although not significant, our models predicted a negative slope for the relationship between body fat and summed signal cost, crying frequency, and tantrum frequency for girls but no relationship for boys.
-
-Height Z-scores, weight Z-scores, BMI Z-scores, and body fat percentage residuals were not significant predictors of signaling behavior in models with the same control variables.
-See xxxxSI TABLESxxxx for regression tables.
-
-```{r}
-#| label: tbl-bodyfat_models
-#| tbl-cap: Regression tables for the models examining the effect of body fat on signaling.
-r_vars <- c(
-  ChildAge = "Age (years)",
-  SexMale = "Sex (male)",
-  OtherChildrenHH = "Other children",
-  LogIncome = "Income (log)",
-  number_adults = "Adults in home",
-  PartnerStatusUnpartnered = "Partner status (single)",
-  EducationLevelYears = "Parent ed. (years)",
-  ConflictFreqN = "Conflict freq.",
-  SignalFreq = "Signal freq.",
-  SignalCost = "Signal cost",
-  SignalFreqMax = "Max signal freq.",
-  IllnessSusceptibilityMean = "Percieved illness susceptibility",
-  SadFreqN = "Sadness freq.",
-  CryFreqN = "Crying freq.",
-  TantrumFreqN = "Tantrum freq.",
-  AlloparentingFreqN = "Child allocare",
-  TricepR = "Tricep skinfold (residuals)",
-  SubscapR = "Subscapular skinfold (residuals)",
-  BodyFat = "BodyFat (composite)"
-)
-modelsummary(list(Sadness = mBFsad, Crying = mBFcry, Tantrum = mBFtantrum, SignalFreq = mBFf, SignalCost = mBFc, SignalFreqMax = mBFfm), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
-```
+See @stbl-alloparenting-dep-var for the corresponding regression table.
+:::
+
+## Anthropometric predictors of child signaling
+
+There were negative relationships between our Body Fat Index and signal cost (`{r} stats$mBFc$BodyFat$str`), crying frequency (`{r} stats$mBFcry$BodyFat$str`), and tantrum frequency (`{r} stats$mBFtantrum$BodyFat$str`).
+See @stbl-bodyfat-models and @fig-effect_bodyfat.
+However, in all models examining the effect of body fat on signaling, two children with Body Fat Index values greater than 6 and low signaling frequencies had a large effect on model predictions, and after removing them from the models, our body fat index was no longer a significant predictor.
+
+Although interactions of Body Fat Index and sex were not significant, our models predicted a negative relationship between body fat and summed signal cost, crying frequency, and tantrum frequency for girls but no relationship for boys.
+
+Height Z-scores, weight Z-scores, BMI Z-scores, and body fat percentage residuals were not significant predictors of signaling behavior. See xxxxSI TABLESxxxx for regression tables.
 
@@
 Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate).
-See @tbl-bodyfat_models for the corresponding regression table.
-:::
-
-## Crying frequency is positively associated with greater height and weight gain in a small longitudinal sample
-
-Signaling data and longitudinal data on child height and weight (2023 and 2024) exist for `r sum(table(height_long_cry$frame$HeightMean_2024))` children.
-In this sample, there was a significant effect of a child age by crying frequency interaction on 2024 child height and weight whereby more crying was associated with more growth when controlling for 2023 height and weight.
-
-See @tbl-longitudinal for the corresponding regression table.
-
-# Check: SD (Observations) in the table
-
-```{r}
-#| label: tbl-longitudinal
-#| tbl-cap: Regression tables for the models examining height and weight growth.
-r_vars <- c(
-  ChildAge = "Age (years)",
-  SexMale = "Sex (male)",
-  OtherChildrenHH = "Other children",
-  LogIncome = "Income (log)",
-  number_adults = "Adults in home",
-  PartnerStatusUnpartnered = "Partner status (single)",
-  EducationLevelYears = "Parent ed. (years)",
-  ConflictFreqN = "Conflict freq.",
-  SignalFreq = "Signal freq.",
-  SignalCost = "Signal cost",
-  SignalFreqMax = "Max signal freq.",
-  IllnessSusceptibilityMean = "Percieved illness susceptibility",
-  SadFreqN = "Sadness freq.",
-  CryFreqN = "Crying freq.",
-  TantrumFreqN = "Tantrum freq.",
-  AlloparentingFreqN = "Child allocare",
-  TricepR = "Tricep skinfold (residuals)",
-  SubscapR = "Subscapular skinfold (residuals)",
-  BodyFat = "BodyFat (composite)",
-  RelativeNeed3 = "Relative need (w/in household)",
-  HeightMean_2024 = "2024 height",
-  HeightMean_2023 = "2023 height",
-  WeightMean_KG_2024 = "2024 weight",
-  WeightMean_KG_2023 = "2023 weight"
-)
-
-modelsummary(list(Weight2024 = weight_long_cry, Height2024 = height_long_cry), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
-```
-
-::: {#fig-effect_longitudinal_plot}
-```{r}
-longitudinal_plot
-```
-
-The effect of a child crying frequency by age on child weight and height in 2024, controlling for the corresponding measure in 2023.
-Rug corresponds to actual values of crying frequency per month with a small amount of jitter to reveal overlapped values.
-See @tbl-longitudinal for the corresponding regression table.
-:::
-
-## Positive associations between child age and signaling frequency and child neediness
-
-Due to the rarity of primary caretakers describing a child as "much less" (`r sum(na.omit(modeldf$RelativeNeed == "Much less"))` \[`r signif((sum(na.omit(modeldf$RelativeNeed == "Much less"))/ sum(!is.na(modeldf$RelativeNeed)) * 100), 2)`%\]) or "much more" (`r sum(na.omit(modeldf$RelativeNeed == "Much more"))` \[`r signif((sum(na.omit(modeldf$RelativeNeed == "Much more"))/ sum(!is.na(modeldf$RelativeNeed)) * 100), 2)`%\]) needy compared to their other children, we collapsed "less" and "much less" into one "less" factor and "more" and "much more" into one "more" factor.
-The end result was `r rel_need_table[1]` less needy children, `r rel_need_table[2]` children who were as needy as their siblings, and `r rel_need_table[3]` more needy children.
-
-There were significant positive relationships between perceived relative neediness within the family and both the summed frequency of child signaling and child age when controlling for child sex, household demographics, the frequency of conflict between child and caretaker, and the frequency of a child providing allocare.
-There was also a positive relationship between summed signal cost and perceived relative neediness within the family when controlling for the same variables; however, this association was not significant.
-See @tbl-need_models for the corresponding regression table.
-
-```{r}
-#| label: tbl-need_models
-#| tbl-cap: Regression tables for the models examining relative perceptions of child need within the family as a dependent variable.
-r_vars <- c(
-  ChildAge = "Age (years)",
-  SexMale = "Sex (male)",
-  OtherChildrenHH = "Other children",
-  LogIncome = "Income (log)",
-  number_adults = "Adults in home",
-  PartnerStatusUnpartnered = "Partner status (single)",
-  EducationLevelYears = "Parent ed. (years)",
-  ConflictFreqN = "Conflict freq.",
-  SignalFreq = "Signal freq.",
-  SignalCost = "Signal cost",
-  SignalFreqMax = "Max signal freq.",
-  IllnessSusceptibilityMean = "Percieved illness susceptibility",
-  SadFreqN = "Sadness freq.",
-  CryFreqN = "Crying freq.",
-  TantrumFreqN = "Tantrum freq.",
-  AlloparentingFreqN = "Child allocare",
-  TricepR = "Tricep skinfold (residuals)",
-  SubscapR = "Subscapular skinfold (residuals)",
-  BodyFat = "BodyFat (composite)",
-  RelativeNeed3 = "Relative need (w/in household)"
-)
-modelsummary(list(SignalFreq = mN1, SignalCost = mN2), vcov = vcovCL, type = "HC0", cluster = ~householdID, gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
-```
+See @stbl-bodyfat-models for the corresponding regression table.
+:::
+
+## Positive associations between child age and signaling frequency and perceptions of child neediness
+
+[add sad freq2]
+Primary caretakers rarely described a child as “much less” `{r} fmt_sum(modeldf$RelativeNeed == "Much less")` or “much more” `{r} fmt_sum(modeldf$RelativeNeed == "Much more")` needy compared to their other children.
+Therefore, we collapsed “less” and “much less” into a “less” factor and “more” and “much more” into a “more” factor.
+The end result was `{r} rel_need_table[1]` less needy children, `{r} rel_need_table[2]` children who were as needy as their siblings, and `{r} rel_need_table[3]` more needy children.
+
+There were significant positive relationships between perceived relative neediness within the family and both the summed frequency of child signaling and child age.
+See @fig-effect_relativeneed and @stbl-need-models.
+There was also a positive relationship between summed signal cost and perceived relative neediness within the family, controlling for the same variables; see @stbl-need-models and @sfig-effect-relativeneed.
 
@@
 ```{r, fig.width= 10}
-signaling_plot_need
-```
-
-The likelihood of child being more needy, as needy, or less needy than a caretaker's other children.
+signaling_plot_need_freq
+```
+
+The proportion of children perceived by caretakers as being more needy, as needy, or less needy than a caretaker's other children.
 A: Effect of child age on relative neediness B: Effect of summed signaling frequency on relative neediness.
+Outcome variable question wording: How much time, effort, and resources does this child need compared to your other children?
 :::
@@
 
-Signals are capable of eliciting better treatment because the information they provide alters the behavior of signal targets in ways the benefit the signaler.
+Signals are capable of eliciting better treatment because the information they provide alters the behavior of signal targets in ways that benefit the signaler.
 For this reason, we tested the association between caretaker perceptions of relative child need within the household and the caretaker's relative investment within the household.
 Consistent with signaling theory, our model predicts a greater proportion of caretakers investing more in children who they deem needier and a greater proportion of caretakers investing less in children they deem less needy, controlling for the age of the child, and household demographics.
-See @fig-plot_relative_investment.
+See @fig-plot_relative_investment and @stbl-need-investment.
 
@@
 
-The proportion of children reported to be given more investment, the same amount of investment, or less investment than other children in the household by the caretaker.
-Effect of perceived need (in terms of time, effort, and resources).
+The effect of perceived need (in terms of time, effort, and resources) relative investment within the household.
 Outcome variable question wording: How much time, effort, and resources do you put into this child compared to your other children?
@@
 
-## Caregiver responses to last signal of need
-
-There were `r response_overlap[2,2]` instances in which a caregiver's response involved acts which were both negative and positive from the child's perspective.
+## Caregiver responses to most recent signal of need
+
+There were `{r} response_overlap[2,2]` instances in which a caregiver's response involved acts which were both negative and positive from the child's perspective.
 These were excluded from analyses due to their rarity.
@@
 However, primary caretakers were significantly more likely to respond positively towards a child when they attributed the cause of signaling to discomfort, pain, injury, or illness (stat), and significantly less likely to respond positively when they attributed the cause of signaling to child punishment (stat).
-See @fig-effect_last_signal_response and @tbl-last_signal_model.
-
-```{r}
-#| label: tbl-last_signal_model
-#| tbl-cap: Regression table for the model examining responses to the last time a child was sad or cried.
-r_vars <- c(
-  ChildAge = "Age (years)",
-  SexMale = "Sex (male)",
-  OtherChildrenHH = "Other children",
-  LogIncome = "Income (log)",
-  number_adults = "Adults in home",
-  PartnerStatusUnpartnered = "Partner status (single)",
-  EducationLevelYears = "Parent ed. (years)",
-  ConflictFreqN = "Conflict freq.",
-  SignalFreq = "Signal freq.",
-  SignalCost = "Signal cost",
-  SignalFreqMax = "Max signal freq.",
-  IllnessSusceptibilityMean = "Percieved illness susceptibility",
-  SadFreqN = "Sadness freq.",
-  CryFreqN = "Crying freq.",
-  TantrumFreqN = "Tantrum freq.",
-  AlloparentingFreqN = "Child allocare",
-  TricepR = "Tricep skinfold (residuals)",
-  SubscapR = "Subscapular skinfold (residuals)",
-  BodyFat = "BodyFat (composite)",
-  RelativeNeed3 = "Relative need (w/in household)",
-  UnwantedTask = "Unwanted task",
-  Punishment21 = "Punishment",
-  Family21 = "Family conflict",
-  DiscomfortPainInjuryIllness = "Discomfort/Pain/Injury/Illness",
-  CaregiverResponse = "Caregiver Response"
-)
-modelsummary(mpr, vcov = vcovCL, type = "HC0", cluster = ~householdID, gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
-```
+See @fig-effect_last_signal_response and @stbl-last-signal-model.
 
@@
 # Supplementary Information
+
+|  |  |
+|--------------------------|----------------------------------------------|
+| Variable | Survey question wording |
+| IllnessAvoidance1 | If an illness is 'going around', my child will get it. |
+| IllnessAvoidance2 | In general, my child is very susceptible to colds, flu, and other infectious diseases |
+| IllnessAvoidance3 | My child’s immune system protects him/her from most illnesses that other kids get. |
+
+: Variable names and question wording for measures about a child's current likelihood of resisting illness.
+All three questions were asked using a Likert scale from 1-5 where 1 was labeled "Strongly disagree" and 5 was labeled "Strongly agree." The scores for IllnessAvoidance1 and IllnessAvoidance2 were inverted so that 5 corresponded with the lowest susceptibility to illness, i.e., greatest *illness avoidance*.
+{#stbl-healthvars}
 
@@
 Other: Other \[Text Box\]
+
+## Figures
+
+::: {#sfig-effect-relativeneed}
+```{r, fig.width= 10}
+p_need_age_cost
+```
+
+The proportion of children perceived by caretakers as being more needy, as needy, or less needy than a caretaker's other children.
+A: Effect of child age on relative neediness B: Effect of signal cost on relative neediness.
+Outcome variable question wording: How much time, effort, and resources does this child need compared to your other children?
+:::
+
+## Regression tables
+
+### Main models
+
+::: {#stbl-main-models}
+```{r}
+modelsummary(list(
+  Sadness = glmmTMBmodels$Model$SadFreqNillness, 
+  Crying = glmmTMBmodels$Model$CryFreqNillness, 
+  Tantrum = glmmTMBmodels$Model$TantrumFreqNillness,
+  SignalFreq = glmmTMBmodels$Model$SignalFreqillness,
+  SignalFreqMax = glmmTMBmodels$Model$SignalFreqMaxillness,
+  SignalCost = glmmTMBmodels$Model$SignalCostillness),
+  gof_map = "nobs", 
+  coef_rename = shortform_dict, 
+  stars = c('*' = .05, '**' = .01, '***' = .001), 
+  coef_omit = "(Intercept)")
+```
+
+Regression tables for our main exploratory models.
+:::
+
+### Conflict frequency model
+
+::: {#stbl-conflict-dep-var}
+```{r}
+modelsummary(glmmTMBmodels$Model$ConflictFreqN, gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+```
+
+Regression table for the model examining the frequency of conflict between child and caretaker as the outcome variable.
+:::
+
+### b
+
+::: {#stbl-alloparenting-dep-var}
+```{r}
+modelsummary(malloparenting, gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+```
+
+Regression table for the model examining the frequency of child alloparenting effort as the dependent variable.
+:::
+
+### Models of signaling predicted by body fat index
+
+::: {#stbl-bodyfat-models}
+```{r}
+modelsummary(list(
+  Sadness = glmmTMBmodels$Model$SadFreqNbodyfat, 
+  Crying = glmmTMBmodels$Model$CryFreqNbodyfat,
+  Tantrum = glmmTMBmodels$Model$TantrumFreqNbodyfat,
+  SignalFreq = glmmTMBmodels$Model$SignalFreqbodyfat,
+  SignalFreqMax = glmmTMBmodels$Model$SignalFreqMaxbodyfat,
+  SignalCost = glmmTMBmodels$Model$SignalCostbodyfat),
+  gof_map = "nobs",
+  coef_rename = shortform_dict,
+  stars = c('*' = .05, '**' = .01, '***' = .001),
+  coef_omit = "(Intercept)")
+```
+
+Regression tables for the models examining the effect of body fat on signaling.
+:::
+
+### d
+
+::: {#stbl-need-models}
+```{r}
+modelsummary(list(SignalFreq = mN1, SignalCost = mN2), vcov = vcovCL, type = "HC0", cluster = ~householdID, gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+```
+
+Regression tables for the models examining relative perceptions of child need within the family as a dependent variable.
+:::
+
+:::
+
+### Model of relative investment within the household
+
+::: {#stbl-need-investment}
+```{r}
+modelsummary(polrModels$Model$Invest, vcov = vcovCL, type = "HC0", cluster = ~householdID, gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+```
+
+:::
+
+### Model of caregiver response to last signal of need
+
+::: {#stbl-last-signal-model}
+```{r}
+modelsummary(polrModels$Model$Response, vcov = vcovCL, type = "HC0", cluster = ~householdID, gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
+```
+
+Regression table for the model examining responses to the last time a child was sad or cried.
+:::

### 2024-10-03 analysisforpaper.R

@@ -29,6 +29,7 @@ library(mgcv)
 library(gratia)
 
 source("recode.R")
+source("dictionaries.R")
 
 # functions used ----------------------------------------------------------
 
@@ -303,7 +304,7 @@ d <-
   left_join(anthropometricMeans[-21], by = "ChildID") |>  # removing householdID column
   left_join(dplyr::select(caregivers, householdID, CPRatio, Neighborhood, ImmigrateUtila, IncomeCategory, IncomeCategoryN,
                              EducationLevel, EducationLevelYears, AdultsMoney, AdultsHousework, AdultsChildcare, AdultsNoChildcare, CaregiverAge,
-                             number_children2, number_adults, UserLanguage, contains("CaregiverMarital"),
+                             number_children2, number_adults, UserLanguage, contains("CaregiverMarital"), NeighborhoodQuality, HouseQuality,
                              contains("SocialSupport"), contains("FoodSecurity")), by = "householdID") |>
   left_join(dplyr::select(causes, -householdID, -childHHid, -ChildID, -Sad1), by = "uniqueID")
 
@@ -465,10 +466,20 @@ modeldf <- d2 |>
     TricepMean,
     TricepR,
     BodyFat,
-    FlexedR
-
+    FlexedR,
+    CaregiverAge,
+    Neighborhood2,
+    UserLanguage,
+    ImmigrateUtila,
+    FoodSecurity,
+    PartnerStatus,
+    NeighborhoodQuality,
+    HouseQuality
   ) |>
-  dplyr::filter(!(is.na(CryFreqN) & is.na(SadFreqN) & is.na(TantrumFreqN)))
+  dplyr::filter(!(is.na(CryFreqN) & is.na(SadFreqN) & is.na(TantrumFreqN))) |>
+  mutate(
+    CaregiverAge = ifelse(is.na(CaregiverAge), mean(CaregiverAge, na.rm = TRUE), CaregiverAge) # imputing age for 1 caretaker where it is missing
+  )
 
 # raw data plots  --------------------------------------------------
 
@@ -673,16 +684,136 @@ ggplot(caregivers2, aes(y=Neighborhood, fill=ImmigrateUtila)) +
 
 # Correlation matrix of predictor and outcome variables
 
[…]

### 2024-10-03 dictionaries.R

@@ -30,6 +30,7 @@ longform_dict <-
 shortform_dict <-
   c(
     ChildAge = "Age (years)",
+    Sex = "Sex",
     SexMale = "Sex (male)",
     OtherChildrenHH = "Other children",
     LogIncome = "Income (log)",
@@ -64,6 +65,8 @@ shortform_dict <-
     WeightMean_KG_2024 = "2024 weight (kg)",
     WeightMean_KG_2023 = "2023 weight (kg)"
   )
+
+# shortform_dict[c(1,2,9,10,11,12,14,15,16)]
 
 # kable(custom.summarize(modeldf, c(dict1[1,3], dict2[1,3,5]), facvar = "Sex")) %>% kable_styling()
 

### 2024-10-04 regularlize.R
Added dataprep.R and regularize.R scripts

@@ -1,0 +1,66 @@
+
+library(glmnet)
+library(hagenutils)
+library(patchwork)
+
+out <- skim(modeldf)
+nms <- out$skim_variable[out$skim_type == 'numeric' & out$complete_rate > 0.9]
+nms <- c(nms, "Sex", "UserLanguage", "ImmigrateUtila", "PartnerStatus")
+mdf2 <-
+  modeldf |>
+  dplyr::select(
+    all_of(nms),
+    -householdID,
+    -childHHid,
+    -NegativeResponse,
+    -PositiveResponse,
+    -IncomeCategoryN,
+    -ConflictFreqN,
+    -OtherChildrenHH, # omit or keep?
+    -contains('IllnessSusceptibility')
+    ) |>
+  na.omit() |>
+  mutate(
+    Sex = ifelse(Sex == "Female", 0, 1),
+    UserLanguage = ifelse(UserLanguage == "EN", 0, 1),
+    ImmigrateUtila = ifelse(ImmigrateUtila == "No", 0, 1),
+    PartnerStatus = ifelse(PartnerStatus == "Unpartnered", 0, 1),
+    across(-c(1:6), \(x) c(scale(x))),
+    AlloparentingXsex = AlloparentingFreqN * Sex,
+    ChildAgeXsex = ChildAge * Sex
+  )
+
+# PCA ---------------------------------------------------------------------
+
+m <- prcomp(mdf2[-c(4:6)], scale. = T) # Remove composite signaling vars
+plot_loadings <- pca_loadings_plot(m, 1:2) # + theme(legend.position = 'top', legend.title = element_blank())
+plot_biplot <- pca_biplot(m) + theme_minimal(15)
+
+plot_pca <- plot_loadings + plot_biplot + plot_layout(widths = c(1,2))
+plot_pca
+
+# Regularized regression --------------------------------------------------
+
+nvar <- ncol(mdf2)
+
+glmnetSignal <- function(signal, alpha = 0, relax = T, s = "lambda.min"){
+  m <- cv.glmnet(
+    as.matrix(mdf2[7:nvar]),
+    mdf2[[signal]],
+    alpha = alpha, # alpha = 0: ridge; alpha = 1: lasso
+    relax = relax,
+    standardize = T,
+    family =  quasipoisson() # MASS::negative.binomial(3) #
+  )
+
+  coefs <- coef(m, s=s)[-1,]
[…]

### 2024-10-04 Paper.qmd
Added new dataframe for responses to last child signal and added information from it to the paper.

@@
 
-Child sadness, crying, and temper tantrums were common.
+Child sadness, crying, and temper tantrums were common and highly correlated with one another.
+See @fig-barplot_SignalFreq and @fig-signal_corrplot.
 Caregivers reported that `{r} fmt_sum(modeldf$SadFreqN >= 8)` were sad at least once a week, `{r} fmt_sum(modeldf$CryFreqN >= 8)` cried at least once a week, and `{r} fmt_sum(modeldf$TantrumFreqN >= 8)` threw temper tantrums at least once a week.
@@
 However, many children were reported to never engage in these signals/states; sadness: `{r} fmt_sum(modeldf$SadFreqN == 0)`; crying; `{r} fmt_sum(modeldf$CryFreqN == 0)`; and tantrums: `{r} fmt_sum(modeldf$TantrumFreqN == 0)`.
-See @fig-barplot_SignalFreq and @fig-signal_alluvial_plot.
+For a visualization of how signal frequencies varied within children see @sfig-signal_alluvial_plot.
 
@@
 ```
-:::
-
-::: {#fig-signal_alluvial_plot}
-```{r}
-signal_alluvial_plot
-```
-
-Combinations of signaling frequencies per month.
-Alluvium width equals the number of individuals with the same frequencies for the three signals/states.
-Stratum height equals the count of individuals at a given frequency.
-Frequency key: Low includes "never" and "once a month or less," medium includes "more than once a month but less than once a week" and "more than once a week but not daily," high includes "daily," and "multiple times per day."
+Correlation matrix of the frequencies of signals/states covered in the study with sex, child age, and the frequency of conflict between children and primary caretakers.
 :::
@@
 We did not predict relationships between child signaling and our demographic control variables, but we did find statistically significant negative relationships between years of caregiver education and all of our signaling measures except sadness (`{r} stats$msadH$EducationLevelYears$str`).
-See @stbl-main-models s and @fig-effect_edu.
+See @stbl-main-models and @fig-effect_edu.
 
@@
 
+[do we want to mentioned distributions of the actual data?]
+Caretakers were much more likely to respond positively than neutrally or negatively to children when caretakers perceived the cause of signaling to involve a child experiencing discomfort, pain, injury, or illness (`r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$PositiveResponseBinary > 0)` positive responses vs.
+`r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$NeutralResponseBinary > 0)` neutral responses vs.
+`r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$NegativeResponseBinary > 0)` negative response).
+However, they were also more likely to respond in a neutral or negative manner rather than a positive one when caretakers believed signaling came in response to the child being punished (`r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$PositiveResponseBinary > 0)` positive responses vs.
+`r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$NeutralResponseBinary > 0)` neutral responses vs.
+`r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$NegativeResponseBinary > 0)` negative responses).
+
 There were `{r} response_overlap[2,2]` instances in which a caregiver's response involved acts which were both negative and positive from the child's perspective.
@@
 There were no significant relationships between, child age, child sex, the number of children in the household, and log family income and the valence of parental reponses (positive, neutral, or negative) from the child's perspective.
-However, primary caretakers were significantly more likely to respond positively towards a child when they attributed the cause of signaling to discomfort, pain, injury, or illness (stat), and significantly less likely to respond positively when they attributed the cause of signaling to child punishment (stat).
+However, primary caretakers were significantly more likely to respond positively towards a child when they attributed the cause of signaling to discomfort, pain, injury, or illness (any statistic report here?), and significantly less likely to respond positively when they attributed the cause of signaling to child punishment (any statistic report here?).
 See @fig-effect_last_signal_response and @stbl-last-signal-model.
@@
 
+::: {#sfig-signal_alluvial_plot}
+```{r}
+signal_alluvial_plot
+```
+
+Combinations of signaling frequencies per month.
+Alluvium width equals the number of individuals with the same frequencies for the three signals/states.
+Stratum height equals the count of individuals at a given frequency.
+Frequency key: Low includes "never" and "once a month or less," medium includes "more than once a month but less than once a week" and "more than once a week but not daily," high includes "daily," and "multiple times per day."
+:::
+
 ::: {#sfig-effect-relativeneed}

### 2024-10-04 analysisforpaper.R

@@ -2006,6 +2006,26 @@ modeldf_FULLANTHROPOMETRICS <- modeldf |>
 
 sex_table2 <- table(modeldf_FULLANTHROPOMETRICS$Sex)
 
+response_last_signal <- d2 |>
+  dplyr::select(
+    householdID,
+    childHHid,
+    CaregiverResponse,
+    PositiveResponseBinary,
+    NeutralResponseBinary,
+    NegativeResponseBinary,
+    DiscomfortPainInjuryIllness,
+    Punishment2
+  ) |>
+  mutate(
+    NeutralResponseBinary2 = ifelse((NeutralResponseBinary == 1) & ((PositiveResponseBinary == 1) | (NegativeResponseBinary == 1)), 0, NeutralResponseBinary)
+    # NeutralResponseBinary2 = ifelse(((NeutralResponseBinary = 1) & (PositiveResponseBinary = 1)) | ((NeutralResponseBinary = 1) & (NegativeResponseBinary = 1)), 0, NegativeResponseBinary)
+  ) |>
+  dplyr::filter(!((PositiveResponseBinary == 0) & (NeutralResponseBinary == 0) & (NegativeResponseBinary == 0)))
+
+last_signal_response_tbl_hurt <- table(d2$CaregiverResponse, d2$DiscomfortPainInjuryIllness)
+last_signal_response_tbl_punish <- table(d2$CaregiverResponse, d2$Punishment2)
+
 # model statistics --------------------------------------------------------
 
 library (glue)

### 2024-10-07 dataprep.R
Added barplot for caregiver response to last child signal.

@@ -249,7 +249,9 @@ d <-
     SignalFreqMax = max2(SadFreqN, CryFreqN, TantrumFreqN, na.rm = TRUE), # RunawayFreqN removed because interpretation not clear
     SignalCost = sum2(SadFreqN, CryFreqN * 2, TantrumFreqN * 3, na.rm = TRUE), # RunawayFreqN removed because interpretation not clear
     IllnessSusceptibilitySum = sum2(IllnessSusceptibility1, IllnessSusceptibility2, IllnessSusceptibility3, na.rm = TRUE),
-    IllnessSusceptibilityMean = mean2(IllnessSusceptibility1, IllnessSusceptibility2, IllnessSusceptibility3, na.rm = TRUE)
+    IllnessSusceptibilityMean = mean2(IllnessSusceptibility1, IllnessSusceptibility2, IllnessSusceptibility3, na.rm = TRUE),
+    MedicalProblemsSum = sum2(SchoolAges1, SchoolAges2, na.rm = TRUE),
+    MedicalProblemsMean = mean2(SchoolAges1, SchoolAges2, na.rm = TRUE) # 1 child has data for only 1
   ) |>
   ungroup() |>
   left_join(anthropometricMeans[-21], by = "ChildID") |>  # removing householdID column

### 2024-10-07 regularize.R

@@ -2,6 +2,7 @@
 library(glmnet)
 library(hagenutils)
 library(patchwork)
+library(skimr)
 
 out <- skim(modeldf)
 nms <- out$skim_variable[out$skim_type == 'numeric' & out$complete_rate > 0.9]

### 2024-10-08 Paper.qmd
Added barplot for caregiver response to last child signal.
New stuff in dataprep and regularize. Minor changes caused by the visual editor.

@@
 |  |  |
-|--------------------------|----------------------------------------------|
+|----|----|
 | Variable | Survey question wording |
@@
 ```
+
 Correlation matrix of the frequencies of signals/states covered in the study with sex, child age, and the frequency of conflict between children and primary caretakers.
@@
 ## Signaling behavior is negatively associated with caregiver education
-[delete this section if we remove conflict from the main model? or mention on significant when controlling for conflict]
+
+\[delete this section if we remove conflict from the main model? or mention on significant when controlling for conflict\]
 
@@
 
-::::::::::: {#fig-effect_allo_by_sex}
+::: {#fig-effect_allo_by_sex}
 ```{r}
@@
 
-[add sad freq2]
-Primary caretakers rarely described a child as “much less” `{r} fmt_sum(modeldf$RelativeNeed == "Much less")` or “much more” `{r} fmt_sum(modeldf$RelativeNeed == "Much more")` needy compared to their other children.
+\[add sad freq2\] Primary caretakers rarely described a child as “much less” `{r} fmt_sum(modeldf$RelativeNeed == "Much less")` or “much more” `{r} fmt_sum(modeldf$RelativeNeed == "Much more")` needy compared to their other children.
 Therefore, we collapsed “less” and “much less” into a “less” factor and “more” and “much more” into a “more” factor.
@@
 
-[do we want to mentioned distributions of the actual data?]
-Caretakers were much more likely to respond positively than neutrally or negatively to children when caretakers perceived the cause of signaling to involve a child experiencing discomfort, pain, injury, or illness (`r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$PositiveResponseBinary > 0)` positive responses vs.
+\[do we want to mentioned distributions of the actual data?\] Caretakers were much more likely to respond positively than neutrally or negatively to children when caretakers perceived the cause of signaling to involve a child experiencing discomfort, pain, injury, or illness (`r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$PositiveResponseBinary > 0)` positive responses vs.
 `r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$NeutralResponseBinary > 0)` neutral responses vs.
@@
 ```
-
 :::

### 2024-10-08 dataprep.R

@@ -100,6 +100,10 @@ d <-
   mutate(
     YoungerKids = map_int(ChildAge, \(x) sum(x > ChildAge)),
     OlderKids = map_int(ChildAge, \(x) sum(x < ChildAge)),
+    OlderGirls = map_int(ChildAge, \(x) sum(x < ChildAge & Sex == 'Female')),
+    OlderBoys = map_int(ChildAge, \(x) sum(x < ChildAge & Sex == 'Male')),
+    OneOlderGirl = OlderGirls > 0,
+    OldestChild = ChildAge == max(ChildAge),
     .by = householdID
   ) |>
   mutate(
@@ -326,7 +330,9 @@ d2 <-
 # dataframe for plots -----------------------------------------------------
 
 # filters based on having at least one value of SadFreqN, CryFreqN, or TantrumFreqN
-modeldf <- d2 |>
+modeldf <-
+  d2 |>
+  dplyr::filter(!(is.na(CryFreqN) & is.na(SadFreqN) & is.na(TantrumFreqN))) |>
   dplyr::select(
     householdID,
     childHHid,
@@ -339,6 +345,9 @@ modeldf <- d2 |>
     SignalFreqMax,
     ChildAge,
     OtherChildrenHH,
+    YoungerKids,
+    OlderGirls,
+    OldestChild,
     LogIncome,
     number_adults,
     ConflictFreqN,
@@ -379,7 +388,6 @@ modeldf <- d2 |>
     NeighborhoodQuality,
     HouseQuality
   ) |>
-  dplyr::filter(!(is.na(CryFreqN) & is.na(SadFreqN) & is.na(TantrumFreqN))) |>
   mutate(
     CaregiverAge = ifelse(is.na(CaregiverAge), mean(CaregiverAge, na.rm = TRUE), CaregiverAge) # imputing age for 1 caretaker where it is missing
   )

### 2024-10-08 regularlize.R

@@ -1,11 +1,15 @@
 
 library(glmnet)
+library(glmnetUtils)
 library(hagenutils)
 library(patchwork)
+library(skimr)
+library(pvclust)
+library(ggcorrplot)
 
 out <- skim(modeldf)
 nms <- out$skim_variable[out$skim_type == 'numeric' & out$complete_rate > 0.9]
-nms <- c(nms, "Sex", "UserLanguage", "ImmigrateUtila", "PartnerStatus")
+nms <- c(nms, "Sex", "UserLanguage", "ImmigrateUtila", "PartnerStatus", "OldestChild")
 mdf2 <-
   modeldf |>
   dplyr::select(
@@ -15,7 +19,7 @@ mdf2 <-
     -NegativeResponse,
     -PositiveResponse,
     -IncomeCategoryN,
-    -ConflictFreqN,
+    # -ConflictFreqN,
     -OtherChildrenHH, # omit or keep?
     -contains('IllnessSusceptibility')
     ) |>
@@ -27,7 +31,8 @@ mdf2 <-
     PartnerStatus = ifelse(PartnerStatus == "Unpartnered", 0, 1),
     across(-c(1:6), \(x) c(scale(x))),
     AlloparentingXsex = AlloparentingFreqN * Sex,
-    ChildAgeXsex = ChildAge * Sex
+    ChildAgeXsex = ChildAge * Sex,
+    OldestXsex = OldestChild * Sex
   )
 
 # PCA ---------------------------------------------------------------------
@@ -42,6 +47,23 @@ plot_pca
 # Regularized regression --------------------------------------------------
 
 nvar <- ncol(mdf2)
+
+# Cross validate both lambda and alpha
+
+m <- cva.glmnet(
+  as.matrix(mdf2[7:nvar]),
+  mdf2[["TantrumFreqN"]],
+  standardize = T,
+  family =  quasipoisson() # MASS::negative.binomial(3) #
+)
+plot(m)
+minlossplot(m, cv.type = 'min')
+
+# Sad: alpha = 1
+# Cry: alpha = 1
+# Tantrum: alpha = 0
+
+# Now fit with those alpha values
 
 glmnetSignal <- function(signal, alpha = 0, relax = T, s = "lambda.min"){
   m <- cv.glmnet(
@@ -59,8 +81,42 @@ glmnetSignal <- function(signal, alpha = 0, relax = T, s = "lambda.min"){
[…]

### 2024-10-09 Paper.qmd
Minor changes for knitting.

@@
 
+::: {#fig-PCA}
+```{r}
+plot_pca
+```
+
+A) Biplot map of PC1 and PC2. Each point is one child. Arrows represent the loadings of PC1 and PC2 on each variable. B) Loadings of PC1 and PC2 on each variable.
+:::
+
 ## Distributions of signaling behavior, alloparenting effort, and conflict
@@
 
-\[do we want to mentioned distributions of the actual data?\] Caretakers were much more likely to respond positively than neutrally or negatively to children when caretakers perceived the cause of signaling to involve a child experiencing discomfort, pain, injury, or illness (`r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$PositiveResponseBinary > 0)` positive responses vs.
+```{r}
+# \[do we want to mentioned distributions of the actual data?\] Caretakers were much more likely to respond positively than neutrally or negatively to children when caretakers perceived the cause of signaling to involve a child experiencing discomfort, pain, injury, or illness (`r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$PositiveResponseBinary > 0)` positive responses vs.
 `r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$NeutralResponseBinary > 0)` neutral responses vs.
@@
 `r sum(response_last_signal[response_last_signal$Punishment2 == 1,]$NegativeResponseBinary > 0)` negative responses).
+```
+
+::: {#fig-effect_last_signal_all_responses}
+```{r}
+barplot_CaregiverResponse
+```
+
+Cap to add.
+:::
 

### 2024-10-09 analysisforpaper.R

@@ -29,11 +29,13 @@ library(mgcv)
 library(gratia)
 library(pvclust)
 library(nnet)
+library(skimr)
+library(pvclust)
 
 source("recode.R")
 source("dictionaries.R")
 source("dataprep.R")
-source("regularize.R") # typo in name
+# source("regularize.R")
 
 # raw data plots  --------------------------------------------------
 
@@ -392,7 +394,7 @@ plot(allEffects(mmsfH))
 
 # Signal specific models
 
-msadH <- glmmTMB(SadFreqN ~ ChildAge + Sex + OtherChildrenHH + LogIncome + number_adults + PartnerStatus + ConflictFreqN + AlloparentingFreqN*Sex + EducationLevelYears + IllnessSusceptibilityMean + (1|householdID),data = d2, family = nbinom2)
+msadH <- glmmTMB(SadFreqN ~ ChildAge + Sex + OtherChildrenHH + LogIncome + number_adults + PartnerStatus + AlloparentingFreqN*Sex + EducationLevelYears + IllnessSusceptibilityMean + (1|householdID),data = d2, family = nbinom2)
 summary(msadH)
 plot(allEffects(msadH))
 
@@ -1782,3 +1784,41 @@ ggplot(anthropometricMeansWide, aes(SadFreqN, WeightR, colour = Sex2)) +
   scale_color_binary()
 
   # New comment to test ZED
+
+# PCA
+out <- skim(modeldf)
+nms <- out$skim_variable[out$skim_type == 'numeric' & out$complete_rate > 0.9]
+nms <- c(nms, "Sex", "UserLanguage", "ImmigrateUtila", "PartnerStatus", "OldestChild")
+mdf2 <-
+  modeldf |>
+  dplyr::select(
+    all_of(nms),
+    -householdID,
+    -childHHid,
+    -NegativeResponse,
+    -PositiveResponse,
+    -IncomeCategoryN,
+    # -ConflictFreqN,
+    -OtherChildrenHH, # omit or keep?
+    -contains('IllnessSusceptibility')
+  ) |>
+  na.omit() |>
+  mutate(
+    Sex = ifelse(Sex == "Female", 0, 1),
+    UserLanguage = ifelse(UserLanguage == "EN", 0, 1),
+    ImmigrateUtila = ifelse(ImmigrateUtila == "No", 0, 1),
+    PartnerStatus = ifelse(PartnerStatus == "Unpartnered", 0, 1),
[…]

### 2024-10-09 dataprep.R

@@ -239,11 +239,7 @@ d <-
     NeutralResponseBinary = ifelse(NeutralResponse >= 1, 1, 0),
     NeutralResponseF = as.factor(NeutralResponseBinary),
     NegativeResponseBinary = ifelse(NegativeResponse >= 1, 1, 0),
-<<<<<<< HEAD
     NegativeResponseF = as.factor(NegativeResponseBinary)
-=======
-    NegativeResponseF = as.factor(NegativeResponseBinary),
->>>>>>> 15646fc16f77da6ca697aa502e1607a7ae8b6673
   ) |>
   rowwise() |>
   mutate(

### 2024-10-09 Paper.qmd

@@
 
+<!-- [NOTE: Add categorical variables to the table; combine child and adult variables into 1 table] -->
 ```{r sumtab_children}
@@
 
+A PCA of primary study variables found that sociodemographic and ecological variables loaded primarily on PC1, whereas the child signaling, conflict, and child age variables loaded primarily on PC2.
+Child signaling variables modestly load on PC1 as well in a direction suggesting that greater signaling covaries with more adverse environments.
+See @fig-PCA.
+
 ::: {#fig-PCA}
 ```{r}
+#| fig-width: 11
 plot_pca
@@
 
-A) Biplot map of PC1 and PC2. Each point is one child. Arrows represent the loadings of PC1 and PC2 on each variable. B) Loadings of PC1 and PC2 on each variable.
+A) Loadings of PC1 and PC2 on each variable. B) Biplot map of PC1 and PC2. Each point is one child. Arrows represent the loadings of PC1 and PC2 on each variable. 
 :::
@@
 See @fig-barplot_ConflictFreq.
+Conflict was also associated with higher signaling frequencies.
+See @fig-PCA and @fig-signal_corrplot.
 
-The frequency of alloparenting effort was bimodally distributed, with `{r} fmt_sum(modeldf$AlloparentingFreqN == 0)` reported to never engage in it (`r sum(na.omit(modeldf$NumberOfChildren==1))` instances were due to a child being an only child) and `{r} fmt_sum(modeldf$AlloparentingFreqN >= 30)` reported to help with childcare at least once a day.
+The frequency of child alloparenting effort, which previous research has linked to parent-child conflict, was bimodally distributed, with `{r} fmt_sum(modeldf$AlloparentingFreqN == 0)` reported to never engage in it (`r sum(na.omit(modeldf$NumberOfChildren==1))` instances were due to a child being an only child) and `{r} fmt_sum(modeldf$AlloparentingFreqN >= 30)` reported to help with childcare at least once a day.
 See @fig-barplot_AlloparentingFreq.

### 2024-10-10 analysisforpaper.R
Added dataframes for glmnet models. Minor fixes for knitting.

@@ -632,7 +632,7 @@ summary(m)
 
 m <- gam(HeightMean_2024 ~ Sex2 * SadFreqN + s(HeightMean_2023) + s(ChildAge, by = Sex2, k = 3), data = anthropometricMeansWide, na.action = na.exclude)
 summary(m)
-plot(m, all.terms = T)
+# plot(m, all.terms = T)
 plot_predictions(m, condition = c("SadFreqN", "Sex2"), points = 1) +
   theme_minimal(15)
 

### 2024-10-10 dataprep.R

@@ -256,8 +256,8 @@ d <-
   left_join(anthropometricMeans[-21], by = "ChildID") |>  # removing householdID column
   left_join(dplyr::select(caregivers, householdID, CPRatio, Neighborhood, ImmigrateUtila, IncomeCategory, IncomeCategoryN,
                           EducationLevel, EducationLevelYears, AdultsMoney, AdultsHousework, AdultsChildcare, AdultsNoChildcare, CaregiverAge,
-                          number_children2, number_adults, UserLanguage, contains("CaregiverMarital"), NeighborhoodQuality, HouseQuality,
-                          contains("SocialSupport"), contains("FoodSecurity")), by = "householdID") |>
+                          number_children2, number_adults, UserLanguage, CurrentJob, contains("CaregiverMarital"), NeighborhoodQuality, HouseQuality,
+                          contains("SocialSupport"), contains("FoodSecurity"), contains("HomeInstability_"), contains("Reality_")), by = "householdID") |>
   left_join(dplyr::select(causes, -householdID, -childHHid, -ChildID, -Sad1), by = "uniqueID")
 
 # prepare household variables for analyses
@@ -265,6 +265,8 @@ d2 <-
   d |>
   mutate(
     LogIncome = log10(IncomeCategoryN),
+    CaregiverAge = ifelse(is.na(CaregiverAge), mean(CaregiverAge, na.rm = TRUE), CaregiverAge), # imputing age for 1 caretaker where it is missing (affects 2 children). Also imputes caregiver ages for 7 other children. But these children do not have signaling data.
+    StayAtHomeMom = ifelse(str_detect(CurrentJob, "home"), 1, 0),
     BodyFat = c(scale(TricepR) + scale(SubscapR)),
     RecentSignalSeverity = com1[as.character(Sad1)],
     RecentSignalCause = com2[as.character(Sad1)],
@@ -293,6 +295,7 @@ d2 <-
     FoodSecurityMean = mean2(FoodSecurity_1, FoodSecurity_2, FoodSecurity_3, FoodSecurity_4, FoodSecurity_5, FoodSecurity_6, na.rm = TRUE),
[…]

### 2024-10-10 regularize.R

@@ -6,6 +6,8 @@ library(patchwork)
 library(skimr)
 library(pvclust)
 library(ggcorrplot)
+library(tidymodels)
+library(poissonreg)
 
 out <- skim(modeldf)
 nms <- out$skim_variable[out$skim_type == 'numeric' & out$complete_rate > 0.9]
@@ -92,7 +94,7 @@ glmnetSignal <- function(signal, alpha = 0, relax = T, s = "lambda.min"){
 
 p1 <- glmnetSignal("SadFreqN", alpha = 0.0, relax = T, s = "lambda.min")
 p2 <- glmnetSignal("CryFreqN", alpha = 1.0, relax = F, s = "lambda.min")
-p2p3 <- glmnetSignal("TantrumFreqN", alpha = 0.0, relax = T, s = "lambda.min")
+p3 <- glmnetSignal("TantrumFreqN", alpha = 0.0, relax = T, s = "lambda.min")
 
 p1/p2/p3 + plot_layout(axes = 'collect')
 

### 2024-10-10 regularize.R
Working on elasticnet stuff

@@ -6,6 +6,8 @@ library(patchwork)
 library(skimr)
 library(pvclust)
 library(ggcorrplot)
+library(tidymodels)
+library(poissonreg)
 
 out <- skim(modeldf)
 nms <- out$skim_variable[out$skim_type == 'numeric' & out$complete_rate > 0.9]
@@ -92,49 +94,51 @@ glmnetSignal <- function(signal, alpha = 0, relax = T, s = "lambda.min"){
 
 p1 <- glmnetSignal("SadFreqN", alpha = 0.0, relax = T, s = "lambda.min")
 p2 <- glmnetSignal("CryFreqN", alpha = 1.0, relax = F, s = "lambda.min")
-p2p3 <- glmnetSignal("TantrumFreqN", alpha = 0.0, relax = T, s = "lambda.min")
+p3 <- glmnetSignal("TantrumFreqN", alpha = 0.0, relax = T, s = "lambda.min")
 
 p1/p2/p3 + plot_layout(axes = 'collect')
 
 
 # Plot predictions
 
-m <- cv.glmnet(
-  as.matrix(mdf2[8:nvar]),
-  mdf2[["CryFreqN"]],
-  alpha = 1, # alpha = 0: ridge; alpha = 1: lasso
-  relax = F,
-  standardize = T,
-  family =  quasipoisson() # MASS::negative.binomial(3) #
-)
-
-
-m2 <- glmnet(
-  as.matrix(mdf2[8:nvar]),
-  mdf2[["CryFreqN"]],
-  alpha = 1, # alpha = 0: ridge; alpha = 1: lasso
-  lambda = m$lambda.min,
-  relax = F,
-  standardize = T,
-  family =  quasipoisson() # MASS::negative.binomial(3) #
-)
-
 # library(tidymodels)
 
-fit <-
-  poisson_reg(mixture = 1, penalty = m$lambda.min) %>%
-  set_engine("glmnet", family = quasipoisson) %>%
-  fit(CryFreqN ~ ., data = mdf2[c(2, 8:22)])
+glmnet2 <- function(d, signal, indices, alpha = 1){
+  lambdas <- c()
+  for (i in 1:20){
+    m <- cv.glmnet(
+      as.matrix(d[indices]),
+      d[[signal]],
+      alpha = alpha, # 0: ridge; 1: lasso
+      relax = F,
+      standardize = T,
+      family = quasipoisson() # MASS::negative.binomial(0.4) #quasipoisson()
+    )
[…]

### 2024-10-14 regularize.R
glmnet code, coef plot code

@@ -10,6 +10,103 @@ library(tidymodels)
 library(poissonreg)
 library(marginaleffects)
 
+
+# Regularized regression --------------------------------------------------
+
+glmnet2 <- function(d, outcome, indices, alpha = 1){
+  # Scale numeric vars by 2 SD, leave binary vars alone
+  # d[indices] <- apply(d[indices], MARGIN = 2, FUN = \(x) if (length(unique(x)) == 2) return (x) else return(c(scale(x)/2)))
+  d[indices] <- scale(d[indices]) # Scale all vars
+  cvm <- list()
+  for (i in 1:20){
+    m <- cv.glmnet(
+      as.matrix(d[indices]),
+      d[[outcome]],
+      alpha = alpha, # 0: ridge; 1: lasso
+      relax = F,
+      standardize = F,
+      family = quasipoisson() # MASS::negative.binomial(0.4) #quasipoisson()
+    )
+    cvm <- c(cvm, list(m$cvm))
+  }
+  
+  cvm.mean <- rowMeans(matrix(list_c(cvm), nrow = length(cvm[[1]])))
+  # return(tibble(lambda = m$lambda, cvm = rowMeans(cvm.mat)))
+  
+  lambda.min <- m$lambda[which.min(cvm.mean)]
+
+  m2 <-
+    poisson_reg(mixture = alpha, penalty = lambda.min) |> 
+    set_engine("glmnet", family = quasipoisson) |> # MASS::negative.binomial(0.4)
+    fit_xy(d[indices], d[[outcome]])
+
+  coefs <- tidy(m2)
+  lmbda <- coefs$lambda[which.min(abs(coefs$lambda - lambda.min))]
+  # print(lmbda)
+  coefs <- coefs |> dplyr::filter(lambda == lmbda)
+  names(coefs$estimate) <- shortform_dict[coefs$term]
+
+  p <-
+    hagenutils::ggdotchart(coefs$estimate[-1]) +
+    geom_vline(xintercept = 0, linetype = 'dotted') +
+    xlab("Coefficients") +
+    ggtitle(shortform_dict[outcome])
+
+  return(list(data = d, model = m2, coefs = coefs, coefplot = p))
+}
+
+signalparams <- expand_grid(Outcome = c("SadFreqN", "CryFreqN", "TantrumFreqN", "SignalFreq", "SignalCost"), alpha = c(0, 1))
[…]

### 2024-10-14 Paper.qmd
Added dataframes for glmnet models. Minor fixes for knitting.
Paper: edited discussion.
Rest: minor changes.

@@
 These data allowed us to calculate the number of children (total, younger, and older) and adults within the household.
+We also imputed data for caregiver age by setting the missing value to the mean.
+For the data used in analyses, this amounted to imputing data for one mother which affected two children.
 
@@
 
-As seen in Western cultures [@zeifman2013; @potegal2003; @koch2003] (check for others), the frequencies of crying and temper tantrums decreased with age, something which is likely due to numerous reasons.
+As seen in Western cultures [@zeifman2013; @potegal2003; @koch2003] (check for others), the frequencies of crying and temper tantrums decreased with age.
 One possibility is that older children have greater ability to satisfy one's desires and behave in ways which match cultural expectations of teenagers and adults.
@@
 The decrease in crying and tantrum frequencies with age may also result from the potential for greater reputational costs, something which may also increase the credibility of such signals among older children as previously mentioned.
-~~It is also possible that parents see older children less than younger children.
-However, the negative association between need with age observed in other societies and the fact that all children in this study were of school age suggest that this is unlikely to be the sole reason responsible for this.~~
-
-There was no significant association between the frequency of child sadness and age.
-Part of this likely results from the fact that, unlike crying, sadness is thought to have important non-signaling functions including drawing one’s attention to a problem, motivating deeper analysis of one’s situation, and causing withdrawal from situations with low fitness prospects [@nesse2019; @nettle2009; @allen2003; @hagen2011], from which, others can still obtain cue-based information about one's emotional state, well-being, and desires.
+
+There was also a negative association between child sadness and age.
+However, interpretation of this finding is more difficult due to the fact that, unlike crying and temper tantrums, sadness is thought to have important non-signaling functions including drawing one’s attention to a problem, motivating deeper analysis of one’s situation, and causing withdrawal from situations with low fitness prospects [@nesse2019; @nettle2009; @allen2003; @hagen2011].
 By asking about sadness as a whole, our measure was unable to disentangle information caretakers gleaned from cues from information that came from signals.
-For this reason, a change in signals that arise from sadness outside of crying (e.g., facial sadness and some, but all, forms of withdrawal) may have been harder to detect with our sample size even if young children expressed facial sadness more often.
-It is also possible that there is no effect of age on sadness frequency among the age ranges examined it.
+- difference in sadness or a difference in facial sadness?
 
@@
 
-Consistent with signaling models which emphasize the role of conflicts of interest in favoring the use of costlier signals, there were positive associations between caretakers' perceptions of conflict frequency and 1) the summed costliness of children's signaling, 2) the frequency of child signals of need, and 3) the frequency of children's most common need signaling strategy.
-Although conflicts of interest do not always lead to behavioral conflict [@altmann1980; @bateson1994], it is likely many, if not most, conflicts are ultimately caused by conflicts of interest, making it a useful proxy for the underlying conflicts of interest that may increase the benefits of deception and the costs of being deceived\[\[5\]\].
-If this is the case, the association between signaling and conflict may represent children's attempts at credibly signaling their value of additional investment or the costliness of punishment [@syme2019] to parents who would have good reason to doubt the trustworthiness of simple verbal requests (or the trustworthiness of facial sadness and crying for those resorting to temper tantrums).
+Consistent with signaling models which emphasize the role of conflicts of interest in favoring the use of costlier signals, conflict and signaling behavior were highly correlated.
+Although conflicts of interest do not always lead to behavioral conflict [@altmann1980; @bateson1994], it is likely many, if not most, conflicts are ultimately caused by conflicts of interest, potentially making it a useful proxy for the underlying conflicts of interest that may increase the benefits of deception and the costs of being deceived\[\[5\]\].
+If this is the case, the correlation between signaling and conflict may represent children's attempts at credibly signaling their value of additional investment or the costliness of punishment [@syme2019] to parents who would have good reason to doubt the trustworthiness of simple verbal requests (or the trustworthiness of facial sadness and crying for those resorting to temper tantrums).
+However, it is also likely that parents view aversive forms of child signaling as conflict in and of itself.
 
@@
 In many societies, girls are often expected to do more alloparenting within the household and often end up doing so [@helfrecht2016 @weisner1977; @barry1971; @turke1988].
-While girls in this sample were also reported to give more allocare, the difference was not significant.
-However, Utilian mothers reported in focus groups (confirm) that girls tend to provide more help for their siblings than boys.
+Our models predict that girls in this sample are more likely to provide allocare (confirm), something which matches findins from focus groups of Utilian mothers.
 If this difference was the outcome of caretakers pressuring girls to do more or resulted in the expectation of more alloparenting, the increase in signaling with alloparenting effort seen among girls may be the result of a strategy to signal the costliness of their efforts in order to invest more in romantic relationships and status seeking behaviors outside the household.
@@
 
-*Alloparenting effort model* Not sure on interpretation here yet
+However, child alloparenting effort was negatively associated with caregiver reports of child conflict.
+One potential reason for this is since parents presumably value the child alloparenting within the family, they see their relationships with children who invest more in the household as less conflictual.
+In line with this reasoning, we found evidence of a child alloparenting effort by child age interaction, with our model predicting increased levels of conflict for older children with greater alloparenting.
+Ideally, we could control for signaling behavior to test this.
+However, the bidirectional relationship between child signaling and conflict with caretakers prevents this (explain better).
+It is also possible that our hypothesis about conflict over investment in the household is incorrect.
 
@@
 :::
+
+### OLD Discussion
+
+*Negative association between age and child signaling (all signaling measures)*
+
+As seen in Western cultures [@zeifman2013; @potegal2003; @koch2003] (check for others), the frequencies of crying and temper tantrums decreased with age, something which is likely due to numerous reasons.
+One possibility is that older children have greater ability to satisfy one's desires and behave in ways which match cultural expectations of teenagers and adults.
+Another is that since one's reproductive value increases as one survives periods of high mortality risk in childhood, the intensity of conflicts of interest with parents decreases since these children are likely to reach adulthood and require less investment to reach adulthood than younger children, current or future [@daly1995].
+For this reason, less costly signals may be more likely to work to increase investment, all else being equal\[\[4\]\].
+That said, all else may often not be equal if investment in older children is costlier, something which may be reflected in caretakers seeing older children as relatively needy.
+The decrease in crying and tantrum frequencies with age may also result from the potential for greater reputational costs, something which may also increase the credibility of such signals among older children as previously mentioned.
+~~It is also possible that parents see older children less than younger children.
+However, the negative association between need with age observed in other societies and the fact that all children in this study were of school age suggest that this is unlikely to be the sole reason responsible for this.~~
+
+There was no significant association between the frequency of child sadness and age.
+Part of this likely results from the fact that, unlike crying, sadness is thought to have important non-signaling functions including drawing one’s attention to a problem, motivating deeper analysis of one’s situation, and causing withdrawal from situations with low fitness prospects [@nesse2019; @nettle2009; @allen2003; @hagen2011], from which, others can still obtain cue-based information about one's emotional state, well-being, and desires.
+By asking about sadness as a whole, our measure was unable to disentangle information caretakers gleaned from cues from information that came from signals.
+For this reason, a change in signals that arise from sadness outside of crying (e.g., facial sadness and some, but all, forms of withdrawal) may have been harder to detect with our sample size even if young children expressed facial sadness more often.
+It is also possible that there is no effect of age on sadness frequency among the age ranges examined it.
+
+-   add in depression over the lifecourse?
+
+Negative association with signal cost too
+
+*Positive association between conflict with parent and child signaling (all signaling measures)*
+
+Consistent with signaling models which emphasize the role of conflicts of interest in favoring the use of costlier signals, there were positive associations between caretakers' perceptions of conflict frequency and 1) the summed costliness of children's signaling, 2) the frequency of child signals of need, and 3) the frequency of children's most common need signaling strategy.
+Although conflicts of interest do not always lead to behavioral conflict [@altmann1980; @bateson1994], it is likely many, if not most, conflicts are ultimately caused by conflicts of interest, making it a useful proxy for the underlying conflicts of interest that may increase the benefits of deception and the costs of being deceived\[\[5\]\].
+If this is the case, the association between signaling and conflict may represent children's attempts at credibly signaling their value of additional investment or the costliness of punishment [@syme2019] to parents who would have good reason to doubt the trustworthiness of simple verbal requests (or the trustworthiness of facial sadness and crying for those resorting to temper tantrums).
+
+*Sex and focal child alloparenting effort interact (positive for females, negative/flat for males)*
+
+We found evidence of an interaction between how often a child cared for their siblings and the sex of the child in our models examining signal frequencies and cost, whereby there was a positive relationship between the frequency and summed costliness of child signals of need and alloparenting effort for girls but no relationship or a slightly negative one for boys.
+In many societies, girls are often expected to do more alloparenting within the household and often end up doing so [@helfrecht2016 @weisner1977; @barry1971; @turke1988].
+While girls in this sample were also reported to give more allocare, the difference was not significant.
+However, Utilian mothers reported in focus groups (confirm) that girls tend to provide more help for their siblings than boys.
+If this difference was the outcome of caretakers pressuring girls to do more or resulted in the expectation of more alloparenting, the increase in signaling with alloparenting effort seen among girls may be the result of a strategy to signal the costliness of their efforts in order to invest more in romantic relationships and status seeking behaviors outside the household.
+
+Evidence for conflicts similar to the one hypothesized here has been reported by multiple researchers working with the Chuukese.
+Lowe found that both male and female children reported conflict with their families over their desires to invest financial resources in their own status enhancing opportunities rather than in the household [@lowe2003].
+In an analysis of conflict among Chuukese-Americans, Syme and Hagen found that severe conflicts were more likely with greater numbers of siblings and when conflicts involved a child spending time away from home or evading family responsibilities [@syme2023a], findings which would be consistent with parents pressuring older children into investing more in the household among other explanations.
+In response to such conflicts, participants reported using a wide range strategies as children including, deception, withholding cooperation, running away, and seeking the assistance of others.
+Through these means, participants were often able to resolve conflicts in their favor despite their lower status in the household and community.
+
+*Alloparenting effort model* Not sure on interpretation here yet
+
+*Body fat discussion*
+
+Children with less body fat cried more, threw more temper tantrums, and had higher summed signal costs.
+In other words, needier children signaled more often.
+Although the body fat by sex interaction was only marginally significant, our models predict a negative relationship between body fat and signaling for girls and no relationship for boys.
+Female reproduction requires a significant caloric invest in children, both in the womb and when breastfeeding [@jasienska2009], costs which males do not have to pay.
+For this reason, low body fat, which can prevent ovulation in extreme cases [@ellison2003], is likely a be better predictor of future reproductive success in adolescent (and older) girls than boys, potentially making signaling for increased investment a more viable strategy in instances when an individual has low body fat.
+
+*Positive associations between child age and signaling frequency and child neediness*
+
+There was a positive association between the summed frequency of child signaling and caretaker perceptions of child neediness within the family, something which would be consistent with sadness, crying, and temper tantrums being signals (or involving signals in the case of crying) if parents took this to mean how much they value additional investment.
+Caretakers also perceived older children as needier, despite having more conflicts with younger children.
+One reason for this could be that caretakers weighted monetary investment highly when evaluating child need.
+
+Signaling behavior did not predict (try more models?) differences in relative caregiver investment between children within households.
+However, crying frequency was associated with both height and weight gain for older children.
+If older children face greater reputation costs from crying, crying may be a relatively credible credible signal compared to crying among younger children, with it being a less viable strategy to exaggerate one's value of investment or the costliness of adversity.
+
+*Parental responses to last signal of need*
+
+Parents were more likely to respond positively and less likely to respond negatively to the last time a child signaled when the cause involved discomfort, pain, injury, or illness than when it did not.
+One potential causal explanation for this is that parents may often have information for sources outside of child signaling to corroborate a child's need (e.g., a fever, a diagnosis, or witness a child hurt themselves) so that there is less private information about a child's current state.
+Simultaneously, conflicts of interest may be less significant in these times due to either 1) both sides benefiting similarly from their child returning to normal (e.g., with an illness) or 2) the cause involving something which is minor and would quickly resolve itself (e.g.
+a skinned knee).
+It may also be that many of the typical positive responses to pain and illness are not very costly to caretakers (e.g., providing comfort), thereby increasing their frequency.
+
+When the cause involved punishment, however, parents were more likely to respond negatively and less likely to respond positively.
+In these circumstances, children may be signaling about the severity of a punishment [@syme2019].
+However, parents may be reluctant to reduce punishments if the behavior that warranted the punishment was repeated or severe.
+They may also react with scolding or further punishment.
+That said, our model predicts positive responses in `r signif(updated_out$estimate[3] * 100, 2)`% of cases (when all other predictors are held at their mean values), suggesting that children are successful in negotiating better outcomes at least some of the time.
+
+Although not a significant relationship, the negative effect of the family conflict being a reported cause of signaling on the likelihood of a positive response may reflect tradeoffs within the family, whereby investment in one child cannot go to another.
+Caretakers may have also responded more negatively when the cause involved family conflict, if it involved a child imposing costs on other family members.
+
+*In most models, there is a negative association between education and child signaling?*
+
+We found a negative association between caretakers’ education level (years in school) and many of our child signaling measures despite our models controlling for household income and partner status.
+One possible reason for this may be that more educated caretakers see their children less during the day due to greater ability to afford childcare or due to them having jobs that are less compatible with child care (test model of social support predicted by education).
+Unfortunately, we do not have data on how much time caretakers spent with children to test this, and it seems equally likely that less educated parents may see their kids less often if they need to work more hours to support their families.
+Another possibility is that lower levels of education are associated with greater levels of adversity for both parents and children, something which appears to be the case for other measures of SES [@pepper2017] and would increase the benefits of signaling need for children and the difficulty of parents to respond.
+It is also possible that education influences parenting norms in ways that cause differences in child signaling behavior.
+
+*Lack of evidence for demographic effects*
+
+Outside of the sex by alloparenting effort interaction, we found little evidence of sex differences in child signaling behavior.
+As adults, females cry more than males, a finding that has been reported across cultures [@vanhemert2011; @becht2001].
+One reason why our study may have failed to find significant effects of sex is that sex differences in the causes and reputational effects of signaling need are small or nonexistent until adolescence and young adulthood, with our small sample size for this age range preventing us from detecting a real effect should it exist.
+It is also possible that the sex difference seen in rates of crying in adults may reflect the lesser ability of females to bargain for better treatment through aggression or the threat of it when faced with adversity [@hagen2016].
+Such a sex difference may not exist among preadolescencents due to a lack of difference in strength or athletic ability beofre puberty [@handelsman2018] or due to their relative lack of formidability compared to adults.
+
+Monthly income was also not found to be a significant predictor of signaling behavior.
+Many models designed to explain signals of need (e.g., the begging of young birds) predict that offspring who are worse off or are receiving less investment have more to gain from signaling than those with more [@mock2011; @godfray1991; @caro2016a; @godfray2000].
+However, it is possible that this can be true and household level resources have little influence on child signaling frequencies between families.
+As signaling behavior is costly to child, caretakers, and siblings, children may gain direct and indirect fitness benefits from limiting their signaling behavior in times of resource stress (i.e., when the chances of gaining investment or the size of investment are small).
+Similarly, children in high resource families may also be motivated to signal at high frequencies due to the potential for larger pieces of investment.
+We know of no empirical explorations of this in humans (check more and confirm).
+However, multiple game theoretical models examining parent-offspring conflict have found that parental resources have a modest impact on the evolutionary stable value for offspring demand for resources [@bossan2013; @mock1997 [confirm]].
+~~Although counterintuitive, such an outcome fits the fact that fitness is relative and that natural selection should be expected to have favored those who maximized their own inclusive fitness, whether one found themselves in a great situation or a dire one.~~ Compared to overall effort in eliciting need, resource availability may have a larger impact on how children signal for more investment based on the finding that birds in stable and unusually favorable environments are more likely to compete with siblings through signals thought to reveal information about need, while those in relatively unpredictable and unusually poor environments have been found to compete more through signals thought to reveal information about quality [@caro2016a].
+
+Despite population level differences in wealth, job certainty, discrimination, and living conditions [@davis2014; @garcia2017; @garcia2018; @stonich2000], we generally did not find an effect of population on child signaling behavior.
+Although we sought data from both English-speaking participants born on the island and Spanish-speaking immigrants from the mainland, our smaller sample of English-speaking Utilians may have limited our ability to detect real differences between the populations.
+It is possible that differences in wealth and status between the groups translate to differences in child signaling due to children facing different types adversity or weighting different types of investment differently.
+It is also likely that the two groups differ in how caretakers could respond to child signaling, and caretakers between the two populations may also differ in their cultural models of parenting.
+However, as previously mentioned, it is conceivable that household level differences have little effect on child signaling behavior.
+
+#### OLD Limitations
+
+The main limitation of this study is that it is completely exploratory, something which results in a high risk of our models overfitting the data and indicating associations which might not exist in the real world.
+Our study was also predominantly cross-sectional, further limiting our ability to make causal inferences.
+
+Our use of convenience and snowball sampling is also a limitation as it resulted in a non-probability sample which may not be representative of the broader population.
+This could happen if there was selection bias so that those who participated differed from the rest of the population in how they interacted with their children or how they assessed, remembered, and reported child behavior.
+
+Outside of child anthropometric measurements, all data come from caregiver reports, and therefore, reflect the biases of caretakers, whatever they may be.
+Due to the fact the data used here were a part of survey which covered many topics and often multiple children, we also faced a tradeoff between survey time and question depth.
+Asking adult participants about their relative investment in children in their household was likely not as effective as more in depth, questions focused on specific questions.
+Framing the question this way may have also increased the likelihood of social desirability bias [@grimm2010], which could lead adult participants to be be more likely to report equal investment across children.
+Our questions about the last time the child signaled may have also been more effective if they focused on specific signals due to our methods combining crying, which is unambiguous as a signal, with sadness, which has both signaling and non-signaling functions.

### 2024-10-17 dictionaries.R
Code for effects and graph plots

@@ -93,7 +93,9 @@ shortform_dict <-
     AdultsHousework = "Adults helping",
     Fullsibs = "Full-siblings",
     Halfsibs = "Half-siblings",
-    Stepsibs = "Step-siblings"
+    Stepsibs = "Step-siblings",
+    MeanChildRelatedness = "Child relatedness",
+    MeanSiblingRelatedness = "Sibling relatedness"
   )
 
 # shortform_dict[c(1,2,9,10,11,12,14,15,16)]

### 2024-10-17 regularize.R

@@ -98,6 +98,41 @@ plotpredictions <- function(params, outcome, condition, title = NULL){
 plotpredictions(signalparams, "CryFreqN1", c("ChildAge", "LifestyleReality_5"), "Crying")
 plotpredictions(conflictparams, "ConflictFreqN1", c("ChildAge", "Sex"))
 
+
+plotpredictions2 <- function(params, cond, title = NULL){
+  # model <- params$out[[outcome]]$model
+  # data <- params$out[[outcome]]$data
+
+  d <-
+    map(params$out, \(x) plot_predictions(x$model, condition = cond, newdata = x$data, draw = F)) |>
+    list_rbind(names_to = 'Outcome') |>
+    dplyr::filter(str_ends(Outcome, '1') & Outcome != 'SignalCost1') |>
+    mutate(Outcome = shortform_dict[str_remove(Outcome, '1')])
+
+  ggplot(d, aes_string(cond, 'estimate', color = 'Outcome')) +
+    geom_line(linewidth = 2) +
+    ylim(0, 40) +
+    labs(x = shortform_dict[cond], y = "Frequency per month") +
+    theme_minimal(15)
+}
+
+# Protective factors (negative coefficients)
+signal_effecs_plot <-
+  plotpredictions2(signalparams, 'ChildAge', 'Adjusted signaling frequencies') +
+  plotpredictions2(signalparams, 'NeighborhoodQuality', 'Adjusted signaling frequencies') +
+  plotpredictions2(signalparams, 'AdultsChildcare', 'Adjusted signaling frequencies') +
+  plotpredictions2(signalparams, 'LogIncome', 'Adjusted signaling frequencies') +
+  plotpredictions2(signalparams, 'EducationLevelYears', 'Adjusted signaling frequencies') +
+  plotpredictions2(signalparams, 'MedicalProblemsMean', 'Adjusted signaling frequencies') +
+  plot_layout(ncol = 2, guides = 'collect', axes = 'collect_y') + plot_annotation(title = "Adjusted signaling frequencies")
+signal_effecs_plot
+
+# Need to plot this by sex
+plotpredictions2(signalparams, 'AlloparentingXsex', 'Adjusted signaling frequencies')
+
+# Risk factors for signaling
+
+
[…]

### 2024-10-17 Paper.qmd
Added code to present minimum spanning tree in paper.

@@
 Correlation matrix of the frequencies of signals/states covered in the study with sex, child age, and the frequency of conflict between children and primary caretakers.
+:::
+
+::: {#fig-mst}
+```{r}
+include_graphics("Figures/plot_mst.pdf")
+```
+
+Minimum spanning tree of child signaling and variables thought to cause differences in signaling strategies.
+Blue edges = negative [do we want more detail here?
+e.g., partial correlations or w/e the proper Bayesian way to word that is] correlations between variables.
+Red edges = positive correlations between variables.
+Edge lengths represent [check that this is right and if more specificity is needed] distances between variables.
+Line weights represent(1 - correlation coefficient)
 :::

### 2024-10-17 analysisforpaper.R

@@ -377,9 +377,15 @@ summary(mscH)
 plot(allEffects(mscH))
 # OtherChildAlloparents*Sex
 
-mscH2 <- glmmTMB(SignalCost ~ ChildAge + Sex + OtherChildrenHH + LogIncome + number_adults + PartnerStatus  + AlloparentingFreqN*Sex + EducationLevelYears + Neighborhood2 + (1|householdID),data = d2, family = nbinom2)
+mscH2 <- glmmTMB(SignalCost ~ ChildAge + Sex + OtherChildrenHH + LogIncome*MeanChildRelatedness + number_adults + PartnerStatus  + AlloparentingFreqN*Sex + EducationLevelYears + Neighborhood2 + (1|householdID),data = d2, family = nbinom2)
 summary(mscH2)
 plot(allEffects(mscH2))
+
+mscH3 <- glmmTMB(SignalCost ~ ChildAge + Sex + Fullsibs + Halfsibs + NonSibs + LogIncome + number_adults + PartnerStatus  + AlloparentingFreqN*Sex + EducationLevelYears + Neighborhood2 + (1|householdID),data = d2, family = nbinom2)
+summary(mscH3)
+plot(allEffects(mscH3))
+
+
 # signal frequency
 
 msfH <- glmmTMB(SignalFreq ~ ChildAge + Sex + OtherChildrenHH + LogIncome + number_adults + PartnerStatus + ConflictFreqN + AlloparentingFreqN*Sex + EducationLevelYears + IllnessSusceptibilityMean + (1|householdID),data = d2, family = nbinom2)
@@ -666,6 +672,14 @@ mconflict <- glmmTMB(ConflictFreqN ~ ChildAge + Sex + OlderKids + YoungerKids + 
 summary(mconflict)
 plot(allEffects(mconflict))
 
+mc2 <- glmmTMB(ConflictFreqN ~ ChildAge + Sex + Fullsibs + Halfsibs + NonSibs + LogIncome + number_adults + PartnerStatus + AlloparentingFreqN*Sex + EducationLevelYears + IllnessSusceptibilityMean + (1|householdID),data = d2, family = nbinom2)
+summary(mc2)
+plot(allEffects(mc2))
+
[…]

### 2024-10-17 regularize.R

@@ -427,3 +427,5 @@ plot_mst <-
   theme_graph()
 plot_mst
 ggsave('plot_mst.svg', plot_mst)
+
+ggsave("Figures/plot_mst.pdf", plot_mst, width = 9, height = 9)

### 2024-10-18 Paper.qmd
Editing results for new figures. More exploratory analyses of anthropometrics

--- 2024-10-14 Paper.qmd
@@
 source("analysisforpaper.R", local = knitr::knit_global())
-
+# source("regularize.R")
 ```
@@
 |  |  |
-|----|----|
+|------------------------------------|------------------------------------|
 | Variable | Survey question wording |
@@
 
+The overall structure of study variables was analyzed with principle components analysis (PCA) and Bayesian graphical models (BGM) and minimum spanning trees (MST).
+
 Child signaling frequencies and signaling costs, which were overdispersed count variables, were analyzed using generalized linear mixed models with a negative binomial error family, fit in R with the glmmTMB package [@brooks2017].
@@
 <!-- [NOTE: Add categorical variables to the table; combine child and adult variables into 1 table] -->
+
 ```{r sumtab_children}
@@
 ::: {#fig-PCA}
-```{r}
-#| fig-width: 11
-plot_pca
-```
+![](Figures/plot_pca.svg)
 
 A)  Loadings of PC1 and PC2 on each variable. B) Biplot map of PC1 and PC2. Each point is one child. Arrows represent the loadings of PC1 and PC2 on each variable.
+:::
+
+We then estimated a BGM for all study variables, extracting the MST, which revealed that signaling variables clustered together, conflict was positively associated with signal cost, and child age negatively associated with crying frequency.
+See @fig-mst.
+
+::: {#fig-mst}
+![](Figures/plot_mst_uw.svg)
+
+Minimum spanning tree of child signaling and variables thought to cause differences in signaling strategies.
+Blue edges = negative \[do we want more detail here?
+e.g., partial correlations or w/e the proper Bayesian way to word that is\] correlations between variables.
+Red edges = positive correlations between variables.
+Edge lengths represent \[check that this is right and if more specificity is needed\] distances between variables.
+Line weights represent(1 - correlation coefficient)
 :::
@@
 
-## Signaling behavior is negatively associated with age
-
-As children mature, they are more capable of fulfilling their own needs, more able to ask others for what they need, and are less vulnerable to threats.
-We therefore expected signaling frequencies to decrease with age.
-For every signaling measure except sadness (which also has non-signaling functions), there was a statistically significant negative relationship between the age of the child and the frequency of signaling.
-See @stbl-main-models and @fig-effect_age.
-
-Age had the largest effect on crying (`{r} stats$mcryH$ChildAge$str`), with the model predicting that by age 8 crying frequency drops to half of its value at age 5 (visual inspection).
-The effect of age on temper tantrums (`{r} stats$mtantrumH$ChildAge$str`) was smaller with a more linear slope.
-
-::: {#fig-effect_age}
-```{r}
-signaling_data_plot_age
-```
-
-The effect of age on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior, all per month.
-Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
-Rug corresponds to actual values of child age with a small amount of jitter to reveal overlapped values.
-See @stbl-main-models for the corresponding regression table.
-:::
-
-## Signaling behavior is positively associated with conflict with primary caretakers
-
-Costly signals of need are necessary when there are conflicts of interest.
-There were statistically significant positive relationships between the frequency of conflict between a child and their primary caretaker and all of our signaling measures (although frequent signaling might cause caretaker perceptions of conflict; see the Discussion section).
-See @stbl-main-models and @fig-effect_conflict.
-
-::: {#fig-effect_conflict}
-```{r}
-signaling_data_plot_conflict
-```
-
-The effect of the frequency of conflict with one's primary caretaker on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and signal cost, all per month.
-Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
-Rug corresponds to actual values of the frequency of conflict between child and primary caretaker with a small amount of jitter to reveal overlapped values.
-See @stbl-main-models for the corresponding regression table.
-:::
-
-## Signaling behavior is negatively associated with caregiver education
-
-\[delete this section if we remove conflict from the main model? or mention on significant when controlling for conflict\]
-
-We did not predict relationships between child signaling and our demographic control variables, but we did find statistically significant negative relationships between years of caregiver education and all of our signaling measures except sadness (`{r} stats$msadH$EducationLevelYears$str`).
-See @stbl-main-models and @fig-effect_edu.
-
-::: {#fig-effect_edu}
-```{r}
-signaling_data_plot_education
-```
-
-The effect of years of caregiver education on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and signal cost, all per month.
-Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
-Rug corresponds to actual values of primary caretaker education (years) with a small amount of jitter to reveal overlapped values.
-See @stbl-main-models for the corresponding regression table.
-:::
-
-## Signaling behavior and child alloparenting effort by sex
-
-High alloparenting effort by children is associated with conflict with their parents (Syme and Hagen, xx).
-Across cultures, girls do more alloparenting than boys (xx, yy).
-We therefore investigated the association of child alloparenting effort and child signaling, with an interaction term for child sex.
-For each signaling measure, there was also a significant interaction between how often a child cared for their siblings and the sex of the child.
-The slope of these relationships were positive for girls and flat or slightly negative for boys.
-In other words, our models predict that as children provide more allocare within their families, the girls’ frequency of signaling increases with increases in their alloparenting effort, and boys’ frequency of signaling remains at the same rate or decreases slightly with increases in their alloparenting effort.
-See @stbl-main-models and @fig-effect_allo_by_sex.
-
-::: {#fig-effect_allo_by_sex}
-```{r}
-signaling_data_plot_allo_by_sex
-```
-
-The effect of an alloparenting effort by sex interaction on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior, all per month.
-Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
-Rug corresponds to actual values of child alloparenting frequency with a small amount of jitter to reveal overlapped values.
-See @stbl-conflict-dep-var for the corresponding regression table.
-:::
-
-## Predictors of conflict between children and caretakers
-
-Because caretaker perceptions of conflict predicted child signaling frequency, we investigated predictors of conflict.
-The frequency of conflict between children and caretakers is negatively associated with child age (`{r} stats$mconflict$ChildAge$str`) and the number of adults who provide childcare in a household (`{r} stats$mconflict$AdultsChildcare$str`) when controlling for the sex of the child, household demographics, the number of adults who provide and do not contribute to childcare, an alloparenting effort by sex interaction, and the summed frequency of alloparenting from all other children in the household.
-See @stbl-conflict-dep-var and @fig-effect_conflict_dep_var.
-
-::: {#fig-effect_conflict_dep_var}
-```{r}
-signaling_data_plot_conflict_dep_var
-```
-
-The effect of A) child age and B) the number of adults in the household that contribute to childcare on the frequency of conflict between caretaker and child.
-Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
-Rug corresponds to actual values of A) child age and B) the number of adults in the household that contribute to childcare with a small amount of jitter to reveal overlapped values.
-See @stbl-conflict-dep-var for the corresponding regression table.
-:::
-
-## Negative association between child alloparenting effort and number of older children
-
-\[DELETE THIS SECTION OR MOVE TO SI???\]
-
-Because child allparenting effort was associated with child signaling, we investigated predictors of child alloparenting effort.
-We expected that children’s alloparenting effort would be positively associated with their age, female sex, and the number of younger siblings, and negatively with the number of older siblings and the total alloparenting effort of others in the household.
-Initial linear modeling of child alloparenting effort revealed that a few households with exceptionally high levels of total child alloparenting effort produced models that predicted extremely unrealistic levels of child alloparenting effort.
-We therefore removed `{r}`r sum(d2$OtherChildAlloparentingFreqN > 121, na.rm = TRUE)` outliers (`{r} `r sum(d2$OtherChildAlloparentingFreqN == 360, na.rm = TRUE)`of which had a value of 360 times per month and`{r} `r sum(d2$OtherChildAlloparentingFreqN == 420, na.rm = TRUE)` of which had a value of 420 timers per month) despite expecting these values to not be outliers in the broader Utilian community.
-
-::: {#fig-effect_alloparenting_dep_var_plot}
-```{r}
-alloparenting_dep_var_plot
-```
-
-The effect of the number of older children in the household on the frequency of a child's alloparenting effort.
-Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
-Each dot is 1 child.
-Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate).
-See @stbl-alloparenting-dep-var for the corresponding regression table.
+## Predictors of signaling frequency and conflict frequency
+
+Using lasso regression, we explored all potential predictors of each child signal frequency, as well as of conflict frequency. See @fig-lasso.
+
+::: {#fig-lasso}
+![](Figures/plot_lasso.svg)
+
+Lasso regression models of each child signaling frequency and conflict frequency. Variables whose coefficient values were shrunk to 0 are not displayed. Models fit uins using quasi-Poisson regression.
+:::
+
+The effects of particularly large risk or protective factors for child signaling are shown in @fig-signal-effects.
+
+::: {#fig-signal-effects}
+![](Figures/signal_effects_plot.svg)
+
+Selected predictors of child signaling frequencies.
+:::
+
+The effects of particularly large risk or protective factors for conflict frequency are shown in @fig-conflict-effects.
+
+::: {#fig-conflict-effects}
+![](Figures/conflict_effects_plot.svg)
+
+Selected predictors of child-parent conflict frequencies.
 :::

### 2024-10-18 dataprep.R

@@ -469,6 +469,42 @@ SignalVars <- d2 |>
   ) |>
   na.omit()
 
+SignalVarsAnthro0 <-
+  anthropometricMeans0 |>
+  dplyr::filter(
+    measurements == 2
+  ) |>
+  pivot_wider(names_from = "Year", values_from = BicepMean:BMI, id_cols = ChildID) |>
+  left_join(d2[c("householdID", "childHHid", "ChildID")], by = 'ChildID') |>
+  dplyr::filter(!is.na(householdID) & !is.na(childHHid))
+
+SignalVarsAnthro <-
+  SignalVars |>
+  left_join(SignalVarsAnthro0) |>
+  dplyr::filter(!is.na(HeightMean_2024))
+
+# out <- skim(anthropometricMeans)
+# nms <- out$skim_variable[out$skim_type == 'numeric' & out$complete_rate > 0.9]
+
+tmp <- left_join(anthropometricMeans, d2[c("childHHid", "ChildID")], by = 'ChildID')
+SignalVarsAnthro2 <-
+  left_join(SignalVars, tmp) |>
+  dplyr::select(
+    -AgeAtMeasurement,
+    -Year,
+    -measurements,
+    -Date.of.measurement,
+    -Sex2,
+    -ChildID,
+    -FlexedRb,
+    -HeightMean,
+    -WeightMean_KG,
+    -BMI,
+    -BodyFatPercentageMean,
+    -GripStrengthMean
+    ) |>
+  na.omit()
+
 conflict_vars_temp <-
   "ChildAge:Younger children have more needs they cannot address on their own, something which can lead to conflict. Older children are more likely to have conflicts with parents over investment within vs. outside the family.X
   Sex:The sexes differ in the adversity they face and their ability to bargain for increased support outside of signals of need. These differences can lead to differences in conflict frequency. These differences may be small or nonexistent in young children.X
@@ -489,39 +525,39 @@ conflict_vars_temp <-
   ImmigrateUtila:There may be cultural differences in child signaling or parental responsiveness. Differences in resources by group influences the payoffs to signaling. All of these might affect the frequency of conflict.X
   UserLanguage:There may be cultural differences in child signaling or parental responsiveness. Differences in resources by group influences the payoffs to signaling. All of these might affect the frequency of conflict."
 
-conflictvars <- data.frame(conflict_vars_temp) |>
-  separate_rows(conflict_vars_temp, sep = "X") |>
[…]

### 2024-10-18 dictionaries.R

@@ -95,7 +95,8 @@ shortform_dict <-
     Halfsibs = "Half-siblings",
     Stepsibs = "Step-siblings",
     MeanChildRelatedness = "Child relatedness",
-    MeanSiblingRelatedness = "Sibling relatedness"
+    MeanSiblingRelatedness = "Sibling relatedness",
+    `Possession score` = "Possession score"
   )
 
 # shortform_dict[c(1,2,9,10,11,12,14,15,16)]

### 2024-10-18 regularize.R

@@ -12,11 +12,16 @@ library(marginaleffects)
 library(easybgm)
 library(tidygraph)
 library(ggraph)
+library(BDgraph)
 
 # Regularized regression --------------------------------------------------
 
-glmnet2 <- function(d, outcome, indices, alpha = 1){
-  # Scale numeric vars by 2 SD, leave binary vars alone
+glmnet2 <- function(d, outcome, indices, alpha = 1, fam = 'quasipoisson'){
+
+  reg <- ifelse(fam == 'quasipoisson', poisson_reg, linear_reg)
+  fam <- ifelse(fam == 'quasipoisson', quasipoisson, gaussian)
+
+    # Scale numeric vars by 2 SD, leave binary vars alone
   # d[indices] <- apply(d[indices], MARGIN = 2, FUN = \(x) if (length(unique(x)) == 2) return (x) else return(c(scale(x)/2)))
   d[indices] <- scale(d[indices]) # Scale all vars
   cvm <- list()
@@ -27,7 +32,7 @@ glmnet2 <- function(d, outcome, indices, alpha = 1){
       alpha = alpha, # 0: ridge; 1: lasso
       relax = F,
       standardize = F,
-      family = quasipoisson() # MASS::negative.binomial(0.4) #quasipoisson()
+      family = fam # MASS::negative.binomial(0.4) #quasipoisson()
     )
     cvm <- c(cvm, list(m$cvm))
   }
@@ -38,8 +43,8 @@ glmnet2 <- function(d, outcome, indices, alpha = 1){
   lambda.min <- m$lambda[which.min(cvm.mean)]
 
   m2 <-
-    poisson_reg(mixture = alpha, penalty = lambda.min) |>
-    set_engine("glmnet", family = quasipoisson) |> # MASS::negative.binomial(0.4)
+    reg(mixture = alpha, penalty = lambda.min) |>
+    set_engine("glmnet", family = fam) |> # MASS::negative.binomial(0.4)
     fit_xy(d[indices], d[[outcome]])
 
   coefs <- tidy(m2)
@@ -71,43 +76,18 @@ plot_lasso <-
   signalparams$out$SignalCost1$coefplot + conflictparams$out$ConflictFreqN1$coefplot +
   plot_layout(axis_titles = 'collect', ncol = 2)
 plot_lasso
+ggsave("Figures/plot_lasso.pdf", plot_lasso, width = 12, height = 12)
[…]

### 2024-10-22 regularize.R
penalized ordinal regression

@@ -294,20 +294,110 @@ anthroparams$out <- map2(anthroparams$Outcome, anthroparams$alpha, \(x, y) glmne
 
 # Predicting signals
 
-anthro2params <- expand_grid(Outcome = c("SadFreqN", "CryFreqN", "TantrumFreqN", "SignalFreq", "SignalCost"), alpha = c(0, 1))
-names(anthro2params$Outcome) <- str_c(anthro2params$Outcome, anthro2params$alpha)
-anthro2params$out <- map2(anthro2params$Outcome, anthro2params$alpha, \(x, y) glmnet2(SignalVarsAnthro2, x, 9:ncol(SignalVarsAnthro2), alpha = y), .progress = T)
+# anthro2params <- expand_grid(Outcome = c("SadFreqN", "CryFreqN", "TantrumFreqN", "SignalFreq", "SignalCost"), alpha = c(0, 1))
+# names(anthro2params$Outcome) <- str_c(anthro2params$Outcome, anthro2params$alpha)
+# anthro2params$out <- map2(anthro2params$Outcome, anthro2params$alpha, \(x, y) glmnet2(SignalVarsAnthro2, x, 9:ncol(SignalVarsAnthro2), alpha = y), .progress = T)
 
-conflictAnthroparams <- expand_grid(Outcome = c("ConflictFreqN"), alpha = c(0, 1))
+conflictAnthroparams <- expand_grid(Outcome = c("ConflictFreqN"), alpha = c(0, 0.5, 1))
 names(conflictAnthroparams$Outcome) <- str_c(conflictAnthroparams$Outcome, conflictAnthroparams$alpha)
 conflictAnthroparams$out <- map2(conflictAnthroparams$Outcome, conflictAnthroparams$alpha, \(x, y) glmnet2(SignalVarsAnthro2, x, 9:ncol(SignalVarsAnthro2), alpha = y), .progress = T)
 
-plot_anthro_ridge <-
-  conflictAnthroparams$out$SadFreqN0$coefplot + conflictAnthroparams$out$CryFreqN0$coefplot +
-  conflictAnthroparams$out$TantrumFreqN0$coefplot + conflictAnthroparams$out$SignalFreq0$coefplot +
-  conflictAnthroparams$out$SignalCost0$coefplot + conflictAnthroparams$out$ConflictFreqN0$coefplot +
-  plot_layout(axis_titles = 'collect', ncol = 2)
-plot_anthro_lasso
+
[…]

### 2024-10-22 Paper.qmd
Paper: worked on discussion.
Discussion: fixed typo.

@@
 
-As seen in Western cultures [@zeifman2013; @potegal2003; @koch2003] (check for others), the frequencies of crying and temper tantrums decreased with age.
-One possibility is that older children have greater ability to satisfy one's desires and behave in ways which match cultural expectations of teenagers and adults.
+As seen in Western cultures [@zeifman2013; @potegal2003; @koch2003] (check for others), the frequencies of crying and temper tantrums decreased with age, as did the frequency of child sadness, summed signal frequency, and signal cost.
+One possible reason for this is that as children mature, they are more capable of fulfilling their own needs, more able to ask others for what they need, and are less vulnerable to threats.
 Another is that since one's reproductive value increases as one survives periods of high mortality risk in childhood, the intensity of conflicts of interest with parents decreases since these children are likely to reach adulthood and require less investment to reach adulthood than younger children, current or future [@daly1995].
@@
 
-There was also a negative association between child sadness and age.
+The effect of age was smallest on child sadness.
 However, interpretation of this finding is more difficult due to the fact that, unlike crying and temper tantrums, sadness is thought to have important non-signaling functions including drawing one’s attention to a problem, motivating deeper analysis of one’s situation, and causing withdrawal from situations with low fitness prospects [@nesse2019; @nettle2009; @allen2003; @hagen2011].
 By asking about sadness as a whole, our measure was unable to disentangle information caretakers gleaned from cues from information that came from signals.
-- difference in sadness or a difference in facial sadness?
-
--   add in depression over the lifecourse?
-
-Negative association with signal cost too
 
@@
 Consistent with signaling models which emphasize the role of conflicts of interest in favoring the use of costlier signals, conflict and signaling behavior were highly correlated.
-Although conflicts of interest do not always lead to behavioral conflict [@altmann1980; @bateson1994], it is likely many, if not most, conflicts are ultimately caused by conflicts of interest, potentially making it a useful proxy for the underlying conflicts of interest that may increase the benefits of deception and the costs of being deceived\[\[5\]\].
+Although conflicts of interest do not always lead to behavioral conflict [@altmann1980; @bateson1994], it is likely many, if not most, conflicts are ultimately caused by conflicts of interest, potentially making it a useful proxy for the underlying conflicts of interest that may increase the benefits of deception for children and the benefits of greater skepticism for parents\[\[5\]\].
 If this is the case, the correlation between signaling and conflict may represent children's attempts at credibly signaling their value of additional investment or the costliness of punishment [@syme2019] to parents who would have good reason to doubt the trustworthiness of simple verbal requests (or the trustworthiness of facial sadness and crying for those resorting to temper tantrums).
-However, it is also likely that parents view aversive forms of child signaling as conflict in and of itself.
+However, it is also likely that parents find some forms of child signaling aversive, and therefore see it as conflict in and of itself.
 
@@
 
-We found evidence of an interaction between how often a child cared for their siblings and the sex of the child in our models examining signal frequencies and cost, whereby there was a positive relationship between the frequency and summed costliness of child signals of need and alloparenting effort for girls but no relationship or a slightly negative one for boys.
-In many societies, girls are often expected to do more alloparenting within the household and often end up doing so [@helfrecht2016 @weisner1977; @barry1971; @turke1988].
-Our models predict that girls in this sample are more likely to provide allocare (confirm), something which matches findins from focus groups of Utilian mothers.
+We found evidence of an interaction between how often a child cared for their siblings and the sex of the child in our lasso models examining signal frequencies and cost, whereby there was a positive relationship between the frequency and summed costliness of child signals of need and alloparenting effort for girls but no relationship or a slightly negative one for boys.
+In many societies, girls do more alloparenting than boys [@helfrecht2016 @weisner1977; @barry1971; @turke1988], something which matches findings from focus groups of Utilian mothers [add info from an analysis here?].
 If this difference was the outcome of caretakers pressuring girls to do more or resulted in the expectation of more alloparenting, the increase in signaling with alloparenting effort seen among girls may be the result of a strategy to signal the costliness of their efforts in order to invest more in romantic relationships and status seeking behaviors outside the household.
@@
 
+[delete based on lasso coefficients?]
 However, child alloparenting effort was negatively associated with caregiver reports of child conflict.
@@
 
+[delete now?]
 *Body fat discussion*
@@
 
+[delete?]
 *In most models, there is a negative association between education and child signaling?*
@@
 
-*Lack of evidence for demographic effects*
+*Demographic effects*
 
@@
 It is also possible that the sex difference seen in rates of crying in adults may reflect the lesser ability of females to bargain for better treatment through aggression or the threat of it when faced with adversity [@hagen2016].
-Such a sex difference may not exist among preadolescencents due to a lack of difference in strength or athletic ability beofre puberty [@handelsman2018] or due to their relative lack of formidability compared to adults.
-
-Monthly income was also not found to be a significant predictor of signaling behavior.
+Such a sex difference may not exist among preadolescencents due to a lack of difference in strength or athletic ability before puberty [@handelsman2018] or due to their relative lack of formidability compared to adults.
+
+Monthly income had a modest effect on crying frequency and summed signal frequency in our lasso regressions.
 Many models designed to explain signals of need (e.g., the begging of young birds) predict that offspring who are worse off or are receiving less investment have more to gain from signaling than those with more [@mock2011; @godfray1991; @caro2016a; @godfray2000].
@@
 However, multiple game theoretical models examining parent-offspring conflict have found that parental resources have a modest impact on the evolutionary stable value for offspring demand for resources [@bossan2013; @mock1997 [confirm]].
-~~Although counterintuitive, such an outcome fits the fact that fitness is relative and that natural selection should be expected to have favored those who maximized their own inclusive fitness, whether one found themselves in a great situation or a dire one.~~ Compared to overall effort in eliciting need, resource availability may have a larger impact on how children signal for more investment based on the finding that birds in stable and unusually favorable environments are more likely to compete with siblings through signals thought to reveal information about need, while those in relatively unpredictable and unusually poor environments have been found to compete more through signals thought to reveal information about quality [@caro2016a].
-
+Compared to overall effort in eliciting need, resource availability may have a larger impact on how children signal for more investment based on the finding that birds in stable and unusually favorable environments are more likely to compete with siblings through signals thought to reveal information about need, while those in relatively unpredictable and unusually poor environments have been found to compete more through signals thought to reveal information about quality [@caro2016a].
+
+Despite the differences in wealth, job certainty, discrimination, and living conditions [@davis2014; @garcia2017; @garcia2018; @stonich2000] between recent immigrants from the mainland and native Utilians, primary language, immigration status, and residence in Camponado did not predict child signaling in our lasso regressions.
+However, neighborhood quality was preserved in the lasso regressions predicting every signaling measure but sadness and was among the variables with the largest negative coefficients.
+One reason for this might that neighborhood quality affects the amount of adversity children face, thereby increasing their neediness [no evidence of this when neighborhood quality is included in models predicting relative child need.
+]
+
+[old]
 Despite population level differences in wealth, job certainty, discrimination, and living conditions [@davis2014; @garcia2017; @garcia2018; @stonich2000], we generally did not find an effect of population on child signaling behavior.
@@
 However, as previously mentioned, it is conceivable that household level differences have little effect on child signaling behavior.
+
+possessions
 

### 2024-10-22 dictionaries.R

@@ -65,7 +65,7 @@ shortform_dict <-
     WeightMean_KG_2024 = "2024 weight (kg)",
     WeightMean_KG_2023 = "2023 weight (kg)",
     FoodSecurity = "Food insecurity",
-    Neighborhood2 = "Campanado",
+    Neighborhood2 = "Camponado",
     UserLanguage = "Primary language",
     NumberOfChildren = " Number of children",
     ImmigrateUtila = "Immigrant",

### 2024-10-22 Paper.qmd
Final work on ordinal models

--- 2024-10-18 Paper.qmd
@@
 ```{r, fig.width= 10}
-signaling_plot_need_freq
+# signaling_plot_need_freq
+plot_need_age / plot_need_sad + ggtitle("") + plot_layout(guides = 'collect') + plot_annotation(tag_levels = "A")
 ```
@@
 ```{r, fig.width= 8}
-plot_relative_investment
+# plot_relative_investment
+plot_invest_need
 ```
@@
 ```{r, fig.width= 8}
-signaling_plot_last_signal_response
+# signaling_plot_last_signal_response
+plot_caregiver_punish / plot_caregiver_pain + ggtitle("") + plot_layout(guides = 'collect')
 ```

### 2024-10-22 dictionaries.R

@@ -51,7 +51,9 @@ shortform_dict <-
     BodyFat = "Body fat index",
     RelativeNeed3 = "Relative need (w/in household)",
     UnwantedTask = "Unwanted task",
+    Punishment2 = "Punishment",
     Punishment21 = "Punishment",
+    Punishment = "Punishment",
     Family21 = "Family conflict",
     DiscomfortPainInjuryIllness = "Discomfort/Pain/Injury/Illness",
     CaregiverResponse = "Caregiver Response",

### 2024-10-22 regularize.R

@@ -307,6 +307,121 @@ conflictAnthroparams$out <- map2(conflictAnthroparams$Outcome, conflictAnthropar
 
 library(ordinalNet)
 
+# For binary predictors
+ordinal_plot <- function(fit, predictor, data, title){
+  x <- predict(fit, type = 'response')
+  d <- data |>
+    mutate(
+      Negative = x[,1],
+      Neutral = x[,2],
+      Positive = x[,3],
+      {{predictor}} := ifelse({{predictor}} < 0, 0, 1)
+    ) |>
+    dplyr::select(
+      {{predictor}}, Negative:Positive
+    ) |>
+    pivot_longer(Negative:Positive)
+  ggplot(d, aes({{predictor}}, value, colour = name)) +
+    geom_count(position = position_dodge(width = 0.1)) +
+    geom_smooth(method='lm', se = F, position = position_dodge(width = 0.1)) +
+    scale_x_continuous(breaks = c(0, 1)) +
+    scale_color_viridis_d(option = 'B', end = 0.8) +
+    guides(color = guide_legend("Response", reverse = T)) +
+    ylim(0, NA) +
+    labs(title = title, x = shortform_dict[deparse(substitute(predictor))], y = "Probability")+
+    theme_minimal(15)
+}
+
+# For continuous predictors
+ordinal_plot2 <- function(fit, predictor, data, title){
+  x <- predict(fit, type = 'response')
+  d <- data |>
+    mutate(
+      Less = x[,1],
+      Equal = x[,2],
+      More = x[,3]
+      # {{predictor}} := ifelse({{predictor}} < 0, 0, 1)
+    ) |>
+    summarise(
+      Less = mean(Less),
+      Equal = mean(Equal),
+      More = mean(More),
+      .by = {{predictor}}
+      # {{predictor}} := ifelse({{predictor}} < 0, 0, 1)
+    ) |>
+    pivot_longer(Less:More) |>
+    mutate(name = factor(name, levels = c("Less", "Equal", "More")))
+
+  ggplot(d, aes({{predictor}}, value, colour = name)) +
+    # geom_smooth(se = F, linewidth = 2) +
+    geom_line(linewidth = 2) +
+    scale_color_viridis_d(option = 'B', end = 0.8) +
[…]

### 2024-10-24 Paper.qmd
Major edits to paper to reflect switch to Lasso models from glmmTMB.

--- 2024-10-22 Paper.qmd
@@
 title: "The Natural History of Child Signals of Need in Utila, Honduras: An Exploratory Study"
+author: 
+  - name: Michael R. Gaffney
+    corresponding: false
+    affiliation: 
+      - name: Washington State University
+        department: Anthropology
+  - name: Edward H. Hagen
+    orcid: 0000-0002-5758-9093
+    email: edhagen@wsu.edu
+    corresponding: true
+    affiliation: 
+      - name: Washington State University
+        department: Anthropology
 execute:
@@
     toc: true
+    toc-expand: 2
+    number-sections: true
 lightbox: true
@@
 |  |  |
-|------------------------------------|------------------------------------|
+|--------------------|----------------------------------------------------|
 | Variable | Survey question wording |
@@
 
-We used 6-point Likert scales for parental reports of 1) the relative neediness of each participating child, and 2) the relative investment they provide to each participating child compared to the rest of their children.
+We used 6-point Likert scales for parental reports of 1) the relative neediness of children within the household, and 2) the relative investment they provide to each participating child compared to the rest of their children.
+Both questions were framed need as involving time, effort, and resources.
 See Table @stbl-relvars for exact wording.
-
-|  |  |
-|------------------------------------|------------------------------------|
-| Variable | Survey question wording |
-| Relative Need | How much time, effort, and resources does this child need compared to your other children? |
-| Relative Investment | How much time, effort, and resources do you put into this child compared to your other children? |
-
-: Variable names and question wording for measures about relative need and investment.
-Both questions had the following labels attached to each point on its Likert scale: Much less; Less; The same amount; More; Much more.
-{#stbl-relvars}
 
@@
 The researcher then binned responses into at least one of 31 categories which included positive, negative, and neutral responses (from the child's perspective).
-For the full list of options see Table S1.
+For the full list of options see @stbl-response-valence.
 We then created a *caregiver response* variable, which was coded as 1 for a positive response (e.g., helping child), 0 for a neutral response (e.g., not responding), and -1 for a negative response (e.g., punishment; all from the child's perspective).
-MG also coded the causes of each as follows: 1) the child was asked to do an *unwanted task*, 2) *punishment*, 3) *family conflict*, and 4) *discomfort, pain, injury, or illness*.
+MG and KS also coded the causes for the presence or absence of 17 elements (e.g., whether the caused involved the child was asked to do an *unwanted task*, *punishment*, *family conflict*, and/or *discomfort, pain, injury, or illness*).
 
-### Caretaker perceptions of child’s ability to avoid illness
+### Caretaker perceptions of child health and ability to avoid illness
+
+Caretakers were asked about the frequency of their child missing school and their child's history of medical problems in general terms (i.e., the extent to which they experienced a "variety of medical issues").
+Both involved the use of five point Likert scales, the scores of which were averaged for use in analyses.
+See @stbl-medical-record for the wording of these variables.
 
@@
 
+### Caretaker perceptions of house quality, neighborhood quality, home instability, and food security
+
+Caregivers were asked various questions about their home, which we used to create a home quality index.
+This involved scoring the following summing the affrimatives to the following conditions: having an indoor bathroom, having a bathroom that is not shared with anyone outside of the household, having a flush toilet, having running water to wash hands in the bathroom, and having a home which did not flood in the last year.
+
+Caregiver perceptions of neighborhood quality were measured using ten questions on a five point Likert scale relating to the costs and benefits associated with living in one's neighborhood.
+The mean of these values was used in analyses as a neighborhood quality index.
+See @stbl-Neighborhood-Quality for variable wording.
+
+Levels of home instability were measured with four questions covering the degree to which the caregiver's family had/has to move homes and whether people move in and out of their homes frequently.
+All questions invovled five point Likert scales.
+See @stbl-Home-Instability for variable wording.
+
+We used a modified version of the USDA Food Security Questionnaire [@blumberg1999] to measure food insecurity, with the most significant changes being converting of the measures to a five point Likert scale and averaging them for a food insecurity index.
+See @stbl-Food-Security for the wording of these variables.
+
 ### Demographic questions
@@
 The survey also included many demographic questions about the caretakers and their households.
-Those used in this study included the (1) ages of those in the household, (2) sexes of those in the household, (3) partnership status of the primary caregiver, (4) number of adults who provide childcare within the household, (5) highest level of education of the primary caregiver, (5) immigration status (in regards to Utila), and (6) monthly household income.
-These data allowed us to calculate the number of children (total, younger, and older) and adults within the household.
+Those used in this study included the (1) ages of those in the household, (2) sexes of those in the household, (3) kinship ties within households, (4) partnership status of the primary caregiver, (5) number of adults who provide childcare within the household, (6) highest level of education of the primary caregiver, (7) immigration status (in regards to Utila), (8) monthly household income, (9) neighborhood of residence, (10) current job of the primary caretaker, and (11) ownership of various possesions (e.g., homes, washing machines, or cellphones; see SIXXXXXX for full list).
+These data allowed us to calculate the number of children (total, younger, and older) and adults within the household as well as the genetic relatedness between siblings.
 We also imputed data for caregiver age by setting the missing value to the mean.
@@
 
-## Child anthropometrics
+\[mention Campanado here and in ethnographic portion?\]
+
+\[I think this gets deleted or do we mention no findings worth reporting?\] \## Child anthropometrics
 
@@
 
-All statistical analyses were exploratory and involved fitting many models with different combinations of independent and dependent variables.
+All statistical analyses were exploratory.
 Whereas confirmatory studies test prespecified hypotheses, exploratory analyses prioritize discovery: they risk overfitting the data, generating false positives, but can reveal unexpected yet potentially important relationships that warrant future investigation [@tukey1977].
@@
 
-Child signaling frequencies and signaling costs, which were overdispersed count variables, were analyzed using generalized linear mixed models with a negative binomial error family, fit in R with the glmmTMB package [@brooks2017].
-Unless otherwise noted, explanatory variables included child age, child sex,, child illness avoidance, caretaker education, caretaker partner status, household income, the number of other children in the household, the number of the adults in the household, the frequency of conflict between the primary caretaker and the child, and an interaction between the frequency of child alloparenting effort and child sex..
-A household random effect was included in all models.
+To investigate the associations between potential causes of child signaling and conflict with caretakers and caregiver reports of child behavior, we used elasticnet regression models.
+Elasticnet models (e.g., lasso and ridge regression) are well suited to instances when the number of predictor variables are large relative to the number of observations due their use of penalty terms and cross-validation to limit the risk of overfitting the data.
+
+Analyses were run in R using the glmnet package [@friedman2010] with a quasipoisson error family due to our outcome variables being overdispered counts.
+Predictors includeddemographic, household, health, and neighborhood variables that might cause signaling behavior or conflict between children and caretakers.
+All variables were centered and standardized by one standard deviation prior to fitting.
 
 Parental responses to child signaling, their perceptions of child need, and their reported investment in the child were coded as three-level ordinal variables, either negative, neutral or positive; or less than other children, same as other children, or more than other children.
-We analyzed these outcomes using ordinal logistic regression, fit in R using the polr function in the MASS package [@venables2002], with clustered standard errors by household [@zeileis2002].
+We analyzed these outcomes using ordinal logistic regression with elastic net penalties, fit in R using the ordinalNet package [@wurm].
 
@@
 kable(custom.summarize(modeldf, longform_dict[1:18])) %>% kable_styling()
+# dx <- modeldf
+# names(dx) <- shortform_dict[names(modeldf)] # set-names??? line 1844
+# datasummary_skim(dx, type = "categorical)
 ```
@@
 
-## Anthropometric predictors of child signaling
-
-There were negative relationships between our Body Fat Index and signal cost (`{r} stats$mBFc$BodyFat$str`), crying frequency (`{r} stats$mBFcry$BodyFat$str`), and tantrum frequency (`{r} stats$mBFtantrum$BodyFat$str`).
-See @stbl-bodyfat-models and @fig-effect_bodyfat.
-However, in all models examining the effect of body fat on signaling, two children with Body Fat Index values greater than 6 and low signaling frequencies had a large effect on model predictions, and after removing them from the models, our body fat index was no longer a significant predictor.
-
-Although interactions of Body Fat Index and sex were not significant, our models predicted a negative relationship between body fat and summed signal cost, crying frequency, and tantrum frequency for girls but no relationship for boys.
-
-Height Z-scores, weight Z-scores, BMI Z-scores, and body fat percentage residuals were not significant predictors of signaling behavior. See xxxxSI TABLESxxxx for regression tables.
-
-::: {#fig-effect_bodyfat}
-```{r}
-signaling_data_plot_bodyfat2
-```
-
-The effect of body fat (scaled composite measure of subscapular and tricep skinfold thicknesses) on A) the summed cost of signaling behavior, B) the frequency of crying, and C) the frequency of temper tantrums.
-Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account).
-Each dot is 1 child.
-Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate).
-See @stbl-bodyfat-models for the corresponding regression table.
-:::
-
 ## Positive associations between child age and signaling frequency and perceptions of child neediness
 
-\[add sad freq2\] Primary caretakers rarely described a child as “much less” `{r} fmt_sum(modeldf$RelativeNeed == "Much less")` or “much more” `{r} fmt_sum(modeldf$RelativeNeed == "Much more")` needy compared to their other children.
+Primary caretakers rarely described a child as “much less” `{r} fmt_sum(modeldf$RelativeNeed == "Much less")` or “much more” `{r} fmt_sum(modeldf$RelativeNeed == "Much more")` needy compared to their other children.
 Therefore, we collapsed “less” and “much less” into a “less” factor and “more” and “much more” into a “more” factor.
@@
 
-There were significant positive relationships between perceived relative neediness within the family and both the summed frequency of child signaling and child age.
-See @fig-effect_relativeneed and @stbl-need-models.
-There was also a positive relationship between summed signal cost and perceived relative neediness within the family, controlling for the same variables; see @stbl-need-models and @sfig-effect-relativeneed.
-
-# FIX: error ribbons cross 1 and 0
+There were positive relationships between perceived relative neediness within the family and both child age and the frequency of child sadness. See @fig-effect_relativeneed.
 
 ::: {#fig-effect_relativeneed}
+![](Figures/plot_need_combined.svg)
+
+The proportion of children perceived by caretakers as being more needy, as needy, or less needy than a caretaker's other children.
+A: Effect of child age on relative neediness B: Effect of summed signaling frequency on relative neediness.
+Outcome variable question wording: How much time, effort, and resources does this child need compared to your other children?
+:::
+
+::: {#fig-effect_relativeneed2}
 ```{r, fig.width= 10}
 signaling_plot_need_freq
+# plot_need_age / plot_need_sad + ggtitle("") + plot_layout(guides = 'collect') + plot_annotation(tag_levels = "A")
 ```
@@
 Consistent with signaling theory, our model predicts a greater proportion of caretakers investing more in children who they deem needier and a greater proportion of caretakers investing less in children they deem less needy, controlling for the age of the child, and household demographics.
-See @fig-plot_relative_investment and @stbl-need-investment.
+See @fig-plot_relative_investment.
 
 ::: {#fig-plot_relative_investment}
+![](Figures/plot_invest_need.svg)
+
+The effect of perceived need (in terms of time, effort, and resources) relative investment within the household.
+Outcome variable question wording: How much time, effort, and resources do you put into this child compared to your other children?
+:::
+
+::: {#fig-plot_relative_investment2}
 ```{r, fig.width= 8}
 plot_relative_investment
+#plot_invest_need
 ```
@@
 ```{r}
-# \[do we want to mentioned distributions of the actual data?\] Caretakers were much more likely to respond positively than neutrally or negatively to children when caretakers perceived the cause of signaling to involve a child experiencing discomfort, pain, injury, or illness (`r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$PositiveResponseBinary > 0)` positive responses vs.
+# \[do we want to mention distributions of the actual data?\] Caretakers were much more likely to respond positively than neutrally or negatively to children when caretakers perceived the cause of signaling to involve a child experiencing discomfort, pain, injury, or illness (`r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$PositiveResponseBinary > 0)` positive responses vs.
 `r sum(response_last_signal[response_last_signal$DiscomfortPainInjuryIllness == 1,]$NeutralResponseBinary > 0)` neutral responses vs.
@@
 
-There were no significant relationships between, child age, child sex, the number of children in the household, and log family income and the valence of parental reponses (positive, neutral, or negative) from the child's perspective.
-However, primary caretakers were significantly more likely to respond positively towards a child when they attributed the cause of signaling to discomfort, pain, injury, or illness (any statistic report here?), and significantly less likely to respond positively when they attributed the cause of signaling to child punishment (any statistic report here?).
-See @fig-effect_last_signal_response and @stbl-last-signal-model.
+Our models found no evidence of relationships between, child age, child sex, the number of children in the household, and log family income and the valence of parental reponses (positive, neutral, or negative) from the child's perspective.
+However, primary caretakers were  more likely to respond positively towards a child when they attributed the cause of signaling to discomfort, pain, injury, or illness, and less likely to respond positively when they attributed the cause of signaling to child punishment (any statistic report here?).
+See @fig-effect_last_signal_response.
 
 ::: {#fig-effect_last_signal_response}
+![](Figures/plot_caregiver_response_combined.svg)
+
+The likelihood of positive, neutral, and negative responses to the last time a child became sad or cried.
+A: Effect of punishment being implicated in the cause (0 = no punishment, 1 = punishment.) B: Effect of discomfort, pain, injury, or illness being implicated in the cause (0 = discomfort, pain, injury, or illness, 1 = discomfort, pain, injury, or illness.) Legend key: -1 = negative response, 0 = neutral response, 1 = positive response (all from child's perspective).
+:::
+
+::: {#fig-effect_last_signal_response2}
 ```{r, fig.width= 8}
 signaling_plot_last_signal_response
+# plot_caregiver_punish / plot_caregiver_pain + ggtitle("") + plot_layout(guides = 'collect')
 ```
@@
 
-We found no significant associations between sex and child signaling behavior outside of the previously mentioned interactions, including when sex was included in models as interaction with age.
-See table SI xxxxxxx for regression tables which included the sex by age interaction.
+[needs updating. worth having a section just for this?]
+We found no evidence of associations between sex and child signaling behavior outside of the previously mentioned interactions.
 Monthly household income also was not a significant predictor of signaling behavior in any model nor was population when using language, immigration status, and neighborhood as proxies for it.
-See table SI xxxxxxx.
 
@@
 We found evidence of an interaction between how often a child cared for their siblings and the sex of the child in our lasso models examining signal frequencies and cost, whereby there was a positive relationship between the frequency and summed costliness of child signals of need and alloparenting effort for girls but no relationship or a slightly negative one for boys.
-In many societies, girls do more alloparenting than boys [@helfrecht2016 @weisner1977; @barry1971; @turke1988], something which matches findings from focus groups of Utilian mothers [add info from an analysis here?].
+In many societies, girls do more alloparenting than boys [@helfrecht2016 @weisner1977; @barry1971; @turke1988], something which matches findings from focus groups of Utilian mothers \[add info from an analysis here?\].
 If this difference was the outcome of caretakers pressuring girls to do more or resulted in the expectation of more alloparenting, the increase in signaling with alloparenting effort seen among girls may be the result of a strategy to signal the costliness of their efforts in order to invest more in romantic relationships and status seeking behaviors outside the household.
@@
 
-[delete based on lasso coefficients?]
-However, child alloparenting effort was negatively associated with caregiver reports of child conflict.
+\[delete based on lasso coefficients?\] However, child alloparenting effort was negatively associated with caregiver reports of child conflict.
 One potential reason for this is since parents presumably value the child alloparenting within the family, they see their relationships with children who invest more in the household as less conflictual.
@@
 
-[delete now?]
-*Body fat discussion*
-
-Children with less body fat cried more, threw more temper tantrums, and had higher summed signal costs.
-In other words, needier children signaled more often.
-Although the body fat by sex interaction was only marginally significant, our models predict a negative relationship between body fat and signaling for girls and no relationship for boys.
-Female reproduction requires a significant caloric invest in children, both in the womb and when breastfeeding [@jasienska2009], costs which males do not have to pay.
-For this reason, low body fat, which can prevent ovulation in extreme cases [@ellison2003], is likely a be better predictor of future reproductive success in adolescent (and older) girls than boys, potentially making signaling for increased investment a more viable strategy in instances when an individual has low body fat.
-
 *Positive associations between child age and signaling frequency and child neediness*
@@
 There was a positive association between the summed frequency of child signaling and caretaker perceptions of child neediness within the family, something which would be consistent with sadness, crying, and temper tantrums being signals (or involving signals in the case of crying) if parents took this to mean how much they value additional investment.
-Caretakers also perceived older children as needier, despite having more conflicts with younger children.
-One reason for this could be that caretakers weighted monetary investment highly when evaluating child need.
 
+Caretakers perceived older children as needier, despite having more conflicts with younger children.
+One possible reason for this might be that caretakers see the types of investment older children seek as more costly.
+In line with this reasoning, multiple caretakers mentioned that older kids need more money and advice and get into trouble more often during surveys.
+It may also be that caretakers see investing in older children as relatively costly as they seek to withdraw investment in order to invest in younger children.
+
+[still needs updating]
 Signaling behavior did not predict (try more models?) differences in relative caregiver investment between children within households.
@@
 
-[delete?]
-*In most models, there is a negative association between education and child signaling?*
+\[delete?\] *In most models, there is a negative association between education and child signaling?*
 
@@
 However, neighborhood quality was preserved in the lasso regressions predicting every signaling measure but sadness and was among the variables with the largest negative coefficients.
-One reason for this might that neighborhood quality affects the amount of adversity children face, thereby increasing their neediness [no evidence of this when neighborhood quality is included in models predicting relative child need.
-]
+One reason for this might that neighborhood quality affects the amount of adversity children face, thereby increasing their neediness \[no evidence of this when neighborhood quality is included in models predicting relative child need.
+\]
 
-[old]
-Despite population level differences in wealth, job certainty, discrimination, and living conditions [@davis2014; @garcia2017; @garcia2018; @stonich2000], we generally did not find an effect of population on child signaling behavior.
+\[old\] Despite population level differences in wealth, job certainty, discrimination, and living conditions [@davis2014; @garcia2017; @garcia2018; @stonich2000], we generally did not find an effect of population on child signaling behavior.
 Although we sought data from both English-speaking participants born on the island and Spanish-speaking immigrants from the mainland, our smaller sample of English-speaking Utilians may have limited our ability to detect real differences between the populations.
@@
 
+## Additional information for survey measures
+
+::: {#stbl-relvars}
+
 |  |  |
-|------------------------------------|------------------------------------|
+|----------------------|--------------------------------------------------|
+| Variable | Survey question wording |
+| Relative Need | How much time, effort, and resources does this child need compared to your other children? |
+| Relative Investment | How much time, effort, and resources do you put into this child compared to your other children? |
+
+Variable names and question wording for measures about relative need and investment.
+Both questions had the following labels attached to each point on its Likert scale: Much less; Less; The same amount; More; Much more.
+:::
+
+|  |  |
+|------------------|------------------------------------------------------|
+| Variable | Survey question wording |
+| SchoolAges1 | My child has missed a lot of school due to illness. |
+| SchoolAges2 | My child has had to deal with a variety of medical issues in their life so far. |
+
+Variable names and question wording for measures about a child's health history.
+Both questions were asked using a Likert scale from 1-5 where 1 was labeled "Strongly disagree" and 5 was labeled "Strongly agree." {#stbl-medical-record}
+
+|  |  |
+|-----------------------|-------------------------------------------------|
 | Variable | Survey question wording |
@@
 
-List of options for research assistant to classify participant responses to the question of how they responded when their child was last sad or cried.
+|  |  |
+|------------------|------------------------------------------------------|
+| Variable | Survey question wording |
+| FoodSecurity_1 | In the past year, the food that we bought often didn’t last, and we didn’t have money to get more. |
+| FoodSecurity_2 | In the past year, we couldn’t afford to eat balanced meals. |
+| FoodSecurity_3 | In the past year, I (or other adults in my household) sometimes cut the size of my meals or skipped meals because there wasn't enough money for food. |
+| FoodSecurity_4 | In the past year, I sometimes ate less than I felt I should because there wasn't enough money for food. |
+| FoodSecurity_5 | In the past year, I was sometimes hungry but didn’t eat because there wasn’t enough money for food. |
+| FoodSecurity_6 | In the past year, I ate lower quality food because I couldn't afford better food. |
+
+: Variable names and question wording for the modified version of the USDA Food Security Questionnaire [@blumberg1999].
+All questions were asked using a Likert scale from 1-5 where 1 was labeled “Strongly disagree” and 5 was labeled “Strongly agree.” {#stbl-Food-Security}
+
+|  |  |
+|--------------------|----------------------------------------------------|
+| Variable | Survey question wording |
+| HomeInstability_1 | We move our home a lot. |
+| HomeInstability_2 | I wish our household felt more stable. |
+| HomeInstability_3 | I worry that we will be forced to move before we are ready. |
+| HomeInstability_4 | People move in and move out of our home frequently. |
+
+: Variable names and question wording for the home instability questions.
+All questions were asked using a Likert scale from 1-5 where 1 was labeled “Disagree strongly” and 5 was labeled “Agree Strongly.” {#stbl-Home-Instability}
+
+|  |  |
+|-----------------------|-------------------------------------------------|
+| Variable | Survey question wording |
+| NeighborhoodQuality_1 | As a place to live, my neighborhood is good quality. |
+| NeighborhoodQuality_2 | As a place to raise children, my neighborhood is good quality. |
+| NeighborhoodQuality_3 | My neighborhood quality has gotten worse over time. |
+| NeighborhoodQuality_4 | My neighborhood is more dangerous than other parts of the island. |
+| NeighborhoodQuality_5 | Sometimes I hear about a fight with a weapon in my neighborhood. |
+| NeighborhoodQuality_6 | Sometimes I hear about illegal drug activity in my neighborhood. |
+| NeighborhoodQuality_7 | Sometimes I hear about someone being badly hurt in my neighborhood. |
+| NeighborhoodQuality_8 | Sometimes I hear about someone's home being robbed in my neighborhood. |
+| NeighborhoodQuality_9 | Sometimes I hear about someone being mugged in my neighborhood. |
+| NeighborhoodQuality_10 | Sometimes I hear about someone being assaulted in my neighborhood. |
+
+: Variable names and question wording for the perceptions of neighborhood quality questions.
+All questions were asked using a Likert scale from 1-5 where 1 was labeled “Strongly disagree” and 5 was labeled “Strongly agree.” {#stbl-Neighborhood-Quality}
 
@@
 
+: List of options for research assistant to classify participant responses to the question of how they responded when their child was last sad or cried. {#stbl-response-valence}
+
+List of possessions owned (yes or no).
+
+Motorized vehicle (scooter, 4-wheeler, etc.) Bicylce Television (TV) with cable Smartphone Washing machine Refrigerator A home that you own Credit card
+
 ## Figures
@@
 :::
-
-## Regression tables
-
-### Main models
-
-::: {#stbl-main-models}
-```{r}
-modelsummary(list(
-  Sadness = glmmTMBmodels$Model$SadFreqNillness, 
-  Crying = glmmTMBmodels$Model$CryFreqNillness, 
-  Tantrum = glmmTMBmodels$Model$TantrumFreqNillness,
-  SignalFreq = glmmTMBmodels$Model$SignalFreqillness,
-  SignalFreqMax = glmmTMBmodels$Model$SignalFreqMaxillness,
-  SignalCost = glmmTMBmodels$Model$SignalCostillness),
-  gof_map = "nobs", 
-  coef_rename = shortform_dict, 
-  stars = c('*' = .05, '**' = .01, '***' = .001), 
-  coef_omit = "(Intercept)")
-```
-
-Regression tables for our main exploratory models.
-:::
-
-### Conflict frequency model
-
-::: {#stbl-conflict-dep-var}
-```{r}
-modelsummary(glmmTMBmodels$Model$ConflictFreqN, gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
-```
-
-Regression table for the model examining the frequency of conflict between child and caretaker as the outcome variable.
-:::
-
-### b
-
-::: {#stbl-alloparenting-dep-var}
-```{r}
-modelsummary(malloparenting, gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
-```
-
-Regression table for the model examining the frequency of child alloparenting effort as the dependent variable.
-:::
-
-### Models of signaling predicted by body fat index
-
-::: {#stbl-bodyfat-models}
-```{r}
-modelsummary(list(
-  Sadness = glmmTMBmodels$Model$SadFreqNbodyfat, 
-  Crying = glmmTMBmodels$Model$CryFreqNbodyfat,
-  Tantrum = glmmTMBmodels$Model$TantrumFreqNbodyfat,
-  SignalFreq = glmmTMBmodels$Model$SignalFreqbodyfat,
-  SignalFreqMax = glmmTMBmodels$Model$SignalFreqMaxbodyfat,
-  SignalCost = glmmTMBmodels$Model$SignalCostbodyfat),
-  gof_map = "nobs",
-  coef_rename = shortform_dict,
-  stars = c('*' = .05, '**' = .01, '***' = .001),
-  coef_omit = "(Intercept)")
-```
-
-Regression tables for the models examining the effect of body fat on signaling.
-:::
-
-### d
-
-::: {#stbl-need-models}
-```{r}
-modelsummary(list(SignalFreq = mN1, SignalCost = mN2), vcov = vcovCL, type = "HC0", cluster = ~householdID, gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
-```
-
-Regression tables for the models examining relative perceptions of child need within the family as a dependent variable.
-:::
-
-:::
-
-### Model of relative investment within the household
-
-::: {#stbl-need-investment}
-```{r}
-modelsummary(polrModels$Model$Invest, vcov = vcovCL, type = "HC0", cluster = ~householdID, gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
-```
-:::
-
-### Model of caregiver response to last signal of need
-
-::: {#stbl-last-signal-model}
-```{r}
-modelsummary(polrModels$Model$Response, vcov = vcovCL, type = "HC0", cluster = ~householdID, gof_map = "nobs", coef_rename = shortform_dict, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
-```
-
-Regression table for the model examining responses to the last time a child was sad or cried.
-:::
-
-### OLD Discussion
-
-*Negative association between age and child signaling (all signaling measures)*
-
-As seen in Western cultures [@zeifman2013; @potegal2003; @koch2003] (check for others), the frequencies of crying and temper tantrums decreased with age, something which is likely due to numerous reasons.
-One possibility is that older children have greater ability to satisfy one's desires and behave in ways which match cultural expectations of teenagers and adults.
-Another is that since one's reproductive value increases as one survives periods of high mortality risk in childhood, the intensity of conflicts of interest with parents decreases since these children are likely to reach adulthood and require less investment to reach adulthood than younger children, current or future [@daly1995].
-For this reason, less costly signals may be more likely to work to increase investment, all else being equal\[\[4\]\].
-That said, all else may often not be equal if investment in older children is costlier, something which may be reflected in caretakers seeing older children as relatively needy.
-The decrease in crying and tantrum frequencies with age may also result from the potential for greater reputational costs, something which may also increase the credibility of such signals among older children as previously mentioned.
-~~It is also possible that parents see older children less than younger children.
-However, the negative association between need with age observed in other societies and the fact that all children in this study were of school age suggest that this is unlikely to be the sole reason responsible for this.~~
-
-There was no significant association between the frequency of child sadness and age.
-Part of this likely results from the fact that, unlike crying, sadness is thought to have important non-signaling functions including drawing one’s attention to a problem, motivating deeper analysis of one’s situation, and causing withdrawal from situations with low fitness prospects [@nesse2019; @nettle2009; @allen2003; @hagen2011], from which, others can still obtain cue-based information about one's emotional state, well-being, and desires.
-By asking about sadness as a whole, our measure was unable to disentangle information caretakers gleaned from cues from information that came from signals.
-For this reason, a change in signals that arise from sadness outside of crying (e.g., facial sadness and some, but all, forms of withdrawal) may have been harder to detect with our sample size even if young children expressed facial sadness more often.
-It is also possible that there is no effect of age on sadness frequency among the age ranges examined it.
-
--   add in depression over the lifecourse?
-
-Negative association with signal cost too
-
-*Positive association between conflict with parent and child signaling (all signaling measures)*
-
-Consistent with signaling models which emphasize the role of conflicts of interest in favoring the use of costlier signals, there were positive associations between caretakers' perceptions of conflict frequency and 1) the summed costliness of children's signaling, 2) the frequency of child signals of need, and 3) the frequency of children's most common need signaling strategy.
-Although conflicts of interest do not always lead to behavioral conflict [@altmann1980; @bateson1994], it is likely many, if not most, conflicts are ultimately caused by conflicts of interest, making it a useful proxy for the underlying conflicts of interest that may increase the benefits of deception and the costs of being deceived\[\[5\]\].
-If this is the case, the association between signaling and conflict may represent children's attempts at credibly signaling their value of additional investment or the costliness of punishment [@syme2019] to parents who would have good reason to doubt the trustworthiness of simple verbal requests (or the trustworthiness of facial sadness and crying for those resorting to temper tantrums).
-
-*Sex and focal child alloparenting effort interact (positive for females, negative/flat for males)*
-
-We found evidence of an interaction between how often a child cared for their siblings and the sex of the child in our models examining signal frequencies and cost, whereby there was a positive relationship between the frequency and summed costliness of child signals of need and alloparenting effort for girls but no relationship or a slightly negative one for boys.
-In many societies, girls are often expected to do more alloparenting within the household and often end up doing so [@helfrecht2016 @weisner1977; @barry1971; @turke1988].
-While girls in this sample were also reported to give more allocare, the difference was not significant.
-However, Utilian mothers reported in focus groups (confirm) that girls tend to provide more help for their siblings than boys.
-If this difference was the outcome of caretakers pressuring girls to do more or resulted in the expectation of more alloparenting, the increase in signaling with alloparenting effort seen among girls may be the result of a strategy to signal the costliness of their efforts in order to invest more in romantic relationships and status seeking behaviors outside the household.
-
-Evidence for conflicts similar to the one hypothesized here has been reported by multiple researchers working with the Chuukese.
-Lowe found that both male and female children reported conflict with their families over their desires to invest financial resources in their own status enhancing opportunities rather than in the household [@lowe2003].
-In an analysis of conflict among Chuukese-Americans, Syme and Hagen found that severe conflicts were more likely with greater numbers of siblings and when conflicts involved a child spending time away from home or evading family responsibilities [@syme2023a], findings which would be consistent with parents pressuring older children into investing more in the household among other explanations.
-In response to such conflicts, participants reported using a wide range strategies as children including, deception, withholding cooperation, running away, and seeking the assistance of others.
-Through these means, participants were often able to resolve conflicts in their favor despite their lower status in the household and community.
-
-*Alloparenting effort model* Not sure on interpretation here yet
-
-*Body fat discussion*
-
-Children with less body fat cried more, threw more temper tantrums, and had higher summed signal costs.
-In other words, needier children signaled more often.
-Although the body fat by sex interaction was only marginally significant, our models predict a negative relationship between body fat and signaling for girls and no relationship for boys.
-Female reproduction requires a significant caloric invest in children, both in the womb and when breastfeeding [@jasienska2009], costs which males do not have to pay.
-For this reason, low body fat, which can prevent ovulation in extreme cases [@ellison2003], is likely a be better predictor of future reproductive success in adolescent (and older) girls than boys, potentially making signaling for increased investment a more viable strategy in instances when an individual has low body fat.
-
-*Positive associations between child age and signaling frequency and child neediness*
-
-There was a positive association between the summed frequency of child signaling and caretaker perceptions of child neediness within the family, something which would be consistent with sadness, crying, and temper tantrums being signals (or involving signals in the case of crying) if parents took this to mean how much they value additional investment.
-Caretakers also perceived older children as needier, despite having more conflicts with younger children.
-One reason for this could be that caretakers weighted monetary investment highly when evaluating child need.
-
-Signaling behavior did not predict (try more models?) differences in relative caregiver investment between children within households.
-However, crying frequency was associated with both height and weight gain for older children.
-If older children face greater reputation costs from crying, crying may be a relatively credible credible signal compared to crying among younger children, with it being a less viable strategy to exaggerate one's value of investment or the costliness of adversity.
-
-*Parental responses to last signal of need*
-
-Parents were more likely to respond positively and less likely to respond negatively to the last time a child signaled when the cause involved discomfort, pain, injury, or illness than when it did not.
-One potential causal explanation for this is that parents may often have information for sources outside of child signaling to corroborate a child's need (e.g., a fever, a diagnosis, or witness a child hurt themselves) so that there is less private information about a child's current state.
-Simultaneously, conflicts of interest may be less significant in these times due to either 1) both sides benefiting similarly from their child returning to normal (e.g., with an illness) or 2) the cause involving something which is minor and would quickly resolve itself (e.g.
-a skinned knee).
-It may also be that many of the typical positive responses to pain and illness are not very costly to caretakers (e.g., providing comfort), thereby increasing their frequency.
-
-When the cause involved punishment, however, parents were more likely to respond negatively and less likely to respond positively.
-In these circumstances, children may be signaling about the severity of a punishment [@syme2019].
-However, parents may be reluctant to reduce punishments if the behavior that warranted the punishment was repeated or severe.
-They may also react with scolding or further punishment.
-That said, our model predicts positive responses in `r signif(updated_out$estimate[3] * 100, 2)`% of cases (when all other predictors are held at their mean values), suggesting that children are successful in negotiating better outcomes at least some of the time.
-
-Although not a significant relationship, the negative effect of the family conflict being a reported cause of signaling on the likelihood of a positive response may reflect tradeoffs within the family, whereby investment in one child cannot go to another.
-Caretakers may have also responded more negatively when the cause involved family conflict, if it involved a child imposing costs on other family members.
-
-*In most models, there is a negative association between education and child signaling?*
-
-We found a negative association between caretakers’ education level (years in school) and many of our child signaling measures despite our models controlling for household income and partner status.
-One possible reason for this may be that more educated caretakers see their children less during the day due to greater ability to afford childcare or due to them having jobs that are less compatible with child care (test model of social support predicted by education).
-Unfortunately, we do not have data on how much time caretakers spent with children to test this, and it seems equally likely that less educated parents may see their kids less often if they need to work more hours to support their families.
-Another possibility is that lower levels of education are associated with greater levels of adversity for both parents and children, something which appears to be the case for other measures of SES [@pepper2017] and would increase the benefits of signaling need for children and the difficulty of parents to respond.
-It is also possible that education influences parenting norms in ways that cause differences in child signaling behavior.
-
-*Lack of evidence for demographic effects*
-
-Outside of the sex by alloparenting effort interaction, we found little evidence of sex differences in child signaling behavior.
-As adults, females cry more than males, a finding that has been reported across cultures [@vanhemert2011; @becht2001].
-One reason why our study may have failed to find significant effects of sex is that sex differences in the causes and reputational effects of signaling need are small or nonexistent until adolescence and young adulthood, with our small sample size for this age range preventing us from detecting a real effect should it exist.
-It is also possible that the sex difference seen in rates of crying in adults may reflect the lesser ability of females to bargain for better treatment through aggression or the threat of it when faced with adversity [@hagen2016].
-Such a sex difference may not exist among preadolescencents due to a lack of difference in strength or athletic ability beofre puberty [@handelsman2018] or due to their relative lack of formidability compared to adults.
-
-Monthly income was also not found to be a significant predictor of signaling behavior.
-Many models designed to explain signals of need (e.g., the begging of young birds) predict that offspring who are worse off or are receiving less investment have more to gain from signaling than those with more [@mock2011; @godfray1991; @caro2016a; @godfray2000].
-However, it is possible that this can be true and household level resources have little influence on child signaling frequencies between families.
-As signaling behavior is costly to child, caretakers, and siblings, children may gain direct and indirect fitness benefits from limiting their signaling behavior in times of resource stress (i.e., when the chances of gaining investment or the size of investment are small).
-Similarly, children in high resource families may also be motivated to signal at high frequencies due to the potential for larger pieces of investment.
-We know of no empirical explorations of this in humans (check more and confirm).
-However, multiple game theoretical models examining parent-offspring conflict have found that parental resources have a modest impact on the evolutionary stable value for offspring demand for resources [@bossan2013; @mock1997 [confirm]].
-~~Although counterintuitive, such an outcome fits the fact that fitness is relative and that natural selection should be expected to have favored those who maximized their own inclusive fitness, whether one found themselves in a great situation or a dire one.~~ Compared to overall effort in eliciting need, resource availability may have a larger impact on how children signal for more investment based on the finding that birds in stable and unusually favorable environments are more likely to compete with siblings through signals thought to reveal information about need, while those in relatively unpredictable and unusually poor environments have been found to compete more through signals thought to reveal information about quality [@caro2016a].
-
-Despite population level differences in wealth, job certainty, discrimination, and living conditions [@davis2014; @garcia2017; @garcia2018; @stonich2000], we generally did not find an effect of population on child signaling behavior.
-Although we sought data from both English-speaking participants born on the island and Spanish-speaking immigrants from the mainland, our smaller sample of English-speaking Utilians may have limited our ability to detect real differences between the populations.
-It is possible that differences in wealth and status between the groups translate to differences in child signaling due to children facing different types adversity or weighting different types of investment differently.
-It is also likely that the two groups differ in how caretakers could respond to child signaling, and caretakers between the two populations may also differ in their cultural models of parenting.
-However, as previously mentioned, it is conceivable that household level differences have little effect on child signaling behavior.
-
-#### OLD Limitations
-
-The main limitation of this study is that it is completely exploratory, something which results in a high risk of our models overfitting the data and indicating associations which might not exist in the real world.
-Our study was also predominantly cross-sectional, further limiting our ability to make causal inferences.
-
-Our use of convenience and snowball sampling is also a limitation as it resulted in a non-probability sample which may not be representative of the broader population.
-This could happen if there was selection bias so that those who participated differed from the rest of the population in how they interacted with their children or how they assessed, remembered, and reported child behavior.
-
-Outside of child anthropometric measurements, all data come from caregiver reports, and therefore, reflect the biases of caretakers, whatever they may be.
-Due to the fact the data used here were a part of survey which covered many topics and often multiple children, we also faced a tradeoff between survey time and question depth.
-Asking adult participants about their relative investment in children in their household was likely not as effective as more in depth, questions focused on specific questions.
-Framing the question this way may have also increased the likelihood of social desirability bias [@grimm2010], which could lead adult participants to be be more likely to report equal investment across children.
-Our questions about the last time the child signaled may have also been more effective if they focused on specific signals due to our methods combining crying, which is unambiguous as a signal, with sadness, which has both signaling and non-signaling functions.

### 2024-10-24 regularize.R

@@ -298,9 +298,9 @@ anthroparams$out <- map2(anthroparams$Outcome, anthroparams$alpha, \(x, y) glmne
 # names(anthro2params$Outcome) <- str_c(anthro2params$Outcome, anthro2params$alpha)
 # anthro2params$out <- map2(anthro2params$Outcome, anthro2params$alpha, \(x, y) glmnet2(SignalVarsAnthro2, x, 9:ncol(SignalVarsAnthro2), alpha = y), .progress = T)
 
-conflictAnthroparams <- expand_grid(Outcome = c("ConflictFreqN"), alpha = c(0, 0.5, 1))
-names(conflictAnthroparams$Outcome) <- str_c(conflictAnthroparams$Outcome, conflictAnthroparams$alpha)
-conflictAnthroparams$out <- map2(conflictAnthroparams$Outcome, conflictAnthroparams$alpha, \(x, y) glmnet2(SignalVarsAnthro2, x, 9:ncol(SignalVarsAnthro2), alpha = y), .progress = T)
+# conflictAnthroparams <- expand_grid(Outcome = c("ConflictFreqN"), alpha = c(0, 0.5, 1))
+# names(conflictAnthroparams$Outcome) <- str_c(conflictAnthroparams$Outcome, conflictAnthroparams$alpha)
+# conflictAnthroparams$out <- map2(conflictAnthroparams$Outcome, conflictAnthroparams$alpha, \(x, y) glmnet2(SignalVarsAnthro2, x, 9:ncol(SignalVarsAnthro2), alpha = y), .progress = T)
 
 
 # Ordinal regressions -----------------------------------------------------
@@ -391,6 +391,10 @@ plot_need_coefs
 plot_need_age <- ordinal_plot2(out$fit, ChildAge, SignalVars4, 'Relative need')
 plot_need_sad <- ordinal_plot2(out$fit, SadFreqN, SignalVars4, 'Relative need')
 
+plot_need_combined <- plot_need_age / plot_need_sad + ggtitle("") + plot_layout(guides = 'collect') + plot_annotation(tag_levels = "A")
+ggsave("Figures/plot_need_combined.pdf", plot_need_combined, width = 12, height = 12)
+ggsave("Figures/plot_need_combined.svg", plot_need_combined, width = 12, height = 12)
+
 # Relative investment
 
 SignalVars5 <-
@@ -418,7 +422,8 @@ colMeans(summary(out))
[…]

### 2024-10-24 Paper.qmd
Minor commit to pull.

@@
 
-Caregivers were asked various questions about their home, which we used to create a home quality index.
-This involved scoring the following summing the affrimatives to the following conditions: having an indoor bathroom, having a bathroom that is not shared with anyone outside of the household, having a flush toilet, having running water to wash hands in the bathroom, and having a home which did not flood in the last year.
+Affirmative caregiver responses to the following questions about their home were summed to create a home quality index: having an indoor bathroom, having a bathroom that is not shared with anyone outside of the household, having a flush toilet, having running water to wash hands in the bathroom, and having a home which did not flood in the last year.
 
-Caregiver perceptions of neighborhood quality were measured using ten questions on a five point Likert scale relating to the costs and benefits associated with living in one's neighborhood.
+Caregiver perceptions of neighborhood quality were measured using ten questions relating to the costs and benefits associated with living in one's neighborhood.
 The mean of these values was used in analyses as a neighborhood quality index.
-See @stbl-Neighborhood-Quality for variable wording.
-
 Levels of home instability were measured with four questions covering the degree to which the caregiver's family had/has to move homes and whether people move in and out of their homes frequently.
-All questions invovled five point Likert scales.
-See @stbl-Home-Instability for variable wording.
+For wording of the these items, which used five point Likert scales, see @stbl-Neighborhood-Quality and @stbl-Home-Instability.
 
@@
 All statistical analyses were exploratory.
-Whereas confirmatory studies test prespecified hypotheses, exploratory analyses prioritize discovery: they risk overfitting the data, generating false positives, but can reveal unexpected yet potentially important relationships that warrant future investigation [@tukey1977].
+Whereas confirmatory studies test prespecified hypotheses generated blind to the data, exploratory analyses are data-driven and prioritize discovery: they risk overfitting the data, generating false positives, but can reveal unexpected yet potentially important relationships that warrant future investigation [@tukey1977].
 
-The overall structure of study variables was analyzed with principle components analysis (PCA) and Bayesian graphical models (BGM) and minimum spanning trees (MST).
+The overall structure of our data was analyzed with principle components analysis (PCA), Bayesian graphical models (BGM), and minimum spanning trees (MST).
+PCA is a linear transformation of the data to a new orthonormal coordinate system in which the the first principle component (PC1) corresponds to the axis of maximum variance, the second principle component (PC2) corresponds to the axis perpendicular to the first of the maximum of the remaining variance, and so forth.
+Among other uses, PCA identifies the linear combinations of variables that account for most of the variance, thereby providing a lower dimensional representation of high dimensional data.
+In BGA, variables are nodes in an undirected graph, and edges represent their conditional dependencies.
+Specifically, an edge indicates that two variables are dependent, conditional on the other variables; edge weights indicate the strength and sign of the relationship.
+Importantly, the absence of an edge indicates that two variables are conditionally independent [@huth2023].
+The graph can therefore identify clusters of dependent variables that are independent of other variables.
+From the graphical models identified with BGM, we extracted an MST, i.e., a subset of the graph in which each node (variable) is connected without any cycles and with the minimum total "distance" between the nodes (more strongly dependent variables, i.e., those with larger absolute weights, having lower "distances").
+The MST is therefore typically a dramatically simplified representation of the original graph that often identifies the most important edges among the nodes, which in our case are the most important dependencies among the variables.
 
 To investigate the associations between potential causes of child signaling and conflict with caretakers and caregiver reports of child behavior, we used elasticnet regression models.
-Elasticnet models (e.g., lasso and ridge regression) are well suited to instances when the number of predictor variables are large relative to the number of observations due their use of penalty terms and cross-validation to limit the risk of overfitting the data.
+Elasticnet models are useful in cases where the number of predictor variables (p) is large relative to the number of observations (n).
+These models, often referred to as penalized or regularized regression (e.g., lasso and ridge regression), aim to limit overfitting the data (variance) by allowing some bias, i.e., by optimizing a bias-variance tradeoff.
+In particular, lasso regression sets the coefficients of uninformative variables to exactly 0, thereby serving as a form of variable selection [@zou2005].
 
-Analyses were run in R using the glmnet package [@friedman2010] with a quasipoisson error family due to our outcome variables being overdispered counts.
-Predictors includeddemographic, household, health, and neighborhood variables that might cause signaling behavior or conflict between children and caretakers.
+Elasticnet regression models of signaling and conflict frequency, which were overdispered count data, were fit using the glmnet package [@friedman2010; @wurm2021] and the quasipoisson error family.
+Predictors included demographic, household, health, and neighborhood variables that might cause signaling behavior or conflict between children and caretakers.
 All variables were centered and standardized by one standard deviation prior to fitting.
@@
 Parental responses to child signaling, their perceptions of child need, and their reported investment in the child were coded as three-level ordinal variables, either negative, neutral or positive; or less than other children, same as other children, or more than other children.
-We analyzed these outcomes using ordinal logistic regression with elastic net penalties, fit in R using the ordinalNet package [@wurm].
+We analyzed these outcomes using ordinal logistic regression with elastic net penalties, fit with the ordinalNet package [@wurm2021].
+
+All analyses were conducted with `{r} R.version.string`.
 
@@
 
-The anonymized data used in the study are available at xxxxxxxLINKxxxxxxx. Study code is available at xxxxxxLINKxxxxxxxxxx.
+The anonymized data used in the study are available at https://github.com/grasshoppermouse/utiladata2023. Study code is available at https://github.com/michaelrgaffney/utilasignaling.
 

### 2024-10-30 Paper.qmd
Paper: primarily changes to discussion plus added figures to SI.
Dataprep + dictionary: code changes for summary tables.

Regularize: saved figures for paper.

@@
   - name: Michael R. Gaffney
+    orcid: 0000-0002-7552-6942
     corresponding: false
@@
 source("analysisforpaper.R", local = knitr::knit_global())
+
+library(modelsummary)
 # source("regularize.R")
@@
 
-Compared to chimpanzee mothers, mothers in hunter-gatherer societies have shorter interbirth intervals and produce offspring that require many more years of parental care [@Davison_2021; @Davison_2022].
+Compared to chimpanzee mothers, mothers in hunter-gatherer societies have shorter interbirth intervals and produce offspring that require many more years of parental care [@davison2021; @davison2022].
 Unlike chimpanzee offspring, human offspring in ancestral populations therefore often grew up with multiple older and/or younger siblings who were potential competitors for parental investment and yet who could also provide (or require) allocare [@xx, @yy].
@@
 In order to elicit more support, children often use signals of needs and wants.
-Across the tree of life, signals evolve to reveal agents’ private information to other agents when doing so benefits both the signaler and the receiver [@caro2016; @maynard-smith2004; @preuschoft2000; @setchell2005; @zahavi1975a].
+Across the tree of life, signals evolve to reveal agents’ private information to other agents when doing so benefits both the signaler and the receiver [@caro2016; @maynard-smith2004; @preuschoft2000; @setchell2005; @zahavi1975].
 Children often have needs and wants, where needs involve avoiding potential fitness costs, such as poor nutrition, injuries, or threats to status, and wants involve acquiring potential fitness benefits, such as highly nutritious foods and valuable artifacts.
@@
 They can also arise when signals cannot be faked (i.e., indexical signals).
-Credible signals can also evolve when the benefits of sending a signal outweigh the costs for individuals with the underlying quality or state, but otherwise the costs outweigh the benefits [@spence1973; @zahavi75; @bergstrom2024].
-Such signals are often termed “costly”, although what is costly is deception, not necessarily signaling per se [for discussion, see @bergstrom2024].
+Credible signals can also evolve when the benefits of sending a signal outweigh the costs for individuals with the underlying quality or state, but otherwise the costs outweigh the benefits [@spence1973; @zahavi1975; @chisausky2023].
+Such signals are often termed “costly”, although what is costly is deception, not necessarily signaling per se [for discussion, see @chisausky2023].
 
@@
 |  |  |
-|--------------------|----------------------------------------------------|
+|-----------------------|-------------------------------------------------|
 | Variable | Survey question wording |
@@
 Anthropometric and signaling data exist for `{r} nrow(modeldf_FULLANTHROPOMETRICS)` children, of whom `{r} sex_table2[1]` were female and `{r} sex_table2[2]` were male.
-For summary statistics of child-level variables see @tbl-sumtab_children.
-For summary statistics of caretaker and household-level variables see @tbl-sumtab_adults.
+For summary statistics of study variables see @tbl-sumtab_combined and @tbl-sumtab_factors.
 
@@
 
-```{r sumtab_children}
-#| label: tbl-sumtab_children
-#| tbl-cap: Summary statistics for child variables.
+```{r sumtab_combined}
+#| label: tbl-sumtab_combined
+#| tbl-cap: Summary statistics for continuous variables.
 
-kable(custom.summarize(modeldf, longform_dict[1:18])) %>% kable_styling()
-# dx <- modeldf
-# names(dx) <- shortform_dict[names(modeldf)] # set-names??? line 1844
-# datasummary_skim(dx, type = "categorical)
+kable(custom.summarize(modeldf, longform_dict[1:31])) %>% kable_styling()
 ```
 
-```{r sumtab_adults}
-#| label: tbl-sumtab_adults
-#| tbl-cap: Summary statistics for adult and household variables.
+```{r sumtab_factors}
+#| label: tbl-sumtab_factors
+#| tbl-cap: Summary statistics for factor variables.
 
-kable(custom.summarize(caregivers, longform_dict[19:23])) %>% kable_styling()
+# convert to factor to omit NAs in datasummary_skim
+dx <- modeldf |> 
+  dplyr::select(any_of(names(longform_dict))) |> 
+  mutate(
+    across(starts_with(c("Lifestyle", "ImmigrateUtila")), as.factor)
+  )
+
+names(dx) <- longform_dict[names(dx)]
+datasummary_skim(data = dx, type = "categorical")
 ```
@@
 Correlation matrix of the frequencies of signals/states covered in the study with sex, child age, and the frequency of conflict between children and primary caretakers.
+For a more complete correlation matrix of study variables see SIxxxxxxx.
 :::
@@
 
-Lasso regression models of each child signaling frequency and conflict frequency. Variables whose coefficient values were shrunk to 0 are not displayed. Models fit uins using quasi-Poisson regression.
+Lasso regression models of each child signaling frequency and conflict frequency. Variables whose coefficient values were shrunk to 0 are not displayed. Models fit using quasi-Poisson regression.
 :::
@@
 
-Selected predictors of child signaling frequencies.
+Effects of selected predictors of child signaling frequencies when holding continuous variables at their mean and factor variables at their mode.
 :::
@@
 
-Selected predictors of child-parent conflict frequencies.
+Effects of selected predictors of child-parent conflict frequencies when holding continuous variables at their mean and factor variables at their mode.
 :::
@@
 
-Primary caretakers rarely described a child as “much less” `{r} fmt_sum(modeldf$RelativeNeed == "Much less")` or “much more” `{r} fmt_sum(modeldf$RelativeNeed == "Much more")` needy compared to their other children.
+Primary caretakers rarely described a child as “much less” `{r} fmt_sum(modeldf$RelativeNeed == "Much less")` \[fix in word doc?\] or “much more” `{r} fmt_sum(modeldf$RelativeNeed == "Much more")` needy compared to their other children.
 Therefore, we collapsed “less” and “much less” into a “less” factor and “more” and “much more” into a “more” factor.
@@
 The proportion of children perceived by caretakers as being more needy, as needy, or less needy than a caretaker's other children.
-A: Effect of child age on relative neediness B: Effect of summed signaling frequency on relative neediness.
-Outcome variable question wording: How much time, effort, and resources does this child need compared to your other children?
-:::
-
-::: {#fig-effect_relativeneed2}
-```{r, fig.width= 10}
-signaling_plot_need_freq
-# plot_need_age / plot_need_sad + ggtitle("") + plot_layout(guides = 'collect') + plot_annotation(tag_levels = "A")
-```
-
-The proportion of children perceived by caretakers as being more needy, as needy, or less needy than a caretaker's other children.
-A: Effect of child age on relative neediness B: Effect of summed signaling frequency on relative neediness.
+A: Effect of child age on relative neediness B: Effect child sadness frequency on relative neediness.
 Outcome variable question wording: How much time, effort, and resources does this child need compared to your other children?
@@
 ![](Figures/plot_invest_need.svg)
-
-The effect of perceived need (in terms of time, effort, and resources) relative investment within the household.
-Outcome variable question wording: How much time, effort, and resources do you put into this child compared to your other children?
-:::
-
-::: {#fig-plot_relative_investment2}
-```{r, fig.width= 8}
-plot_relative_investment
-#plot_invest_need
-```
 
@@
 
-There were `{r} response_overlap[2,2]` instances in which a caregiver's response involved acts which were both negative and positive from the child's perspective.
+There were `{r} response_overlap[1,1]` instances in which a caregiver's response involved acts which were both negative and positive from the child's perspective.
 These were excluded from analyses due to their rarity.
@@
 
-::: {#fig-effect_last_signal_response2}
-```{r, fig.width= 8}
-signaling_plot_last_signal_response
-# plot_caregiver_punish / plot_caregiver_pain + ggtitle("") + plot_layout(guides = 'collect')
-```
-
-The likelihood of positive, neutral, and negative responses to the last time a child became sad or cried.
-A: Effect of punishment being implicated in the cause (0 = no punishment, 1 = punishment.) B: Effect of discomfort, pain, injury, or illness being implicated in the cause (0 = discomfort, pain, injury, or illness, 1 = discomfort, pain, injury, or illness.) Legend key: -1 = negative response, 0 = neutral response, 1 = positive response (all from child's perspective).
-:::
-
-## Sociodemographics
-
-[needs updating. worth having a section just for this?]
-We found no evidence of associations between sex and child signaling behavior outside of the previously mentioned interactions.
-Monthly household income also was not a significant predictor of signaling behavior in any model nor was population when using language, immigration status, and neighborhood as proxies for it.
-
 # Discussion
@@
 As seen in Western cultures [@zeifman2013; @potegal2003; @koch2003] (check for others), the frequencies of crying and temper tantrums decreased with age, as did the frequency of child sadness, summed signal frequency, and signal cost.
-One possible reason for this is that as children mature, they are more capable of fulfilling their own needs, more able to ask others for what they need, and are less vulnerable to threats.
-Another is that since one's reproductive value increases as one survives periods of high mortality risk in childhood, the intensity of conflicts of interest with parents decreases since these children are likely to reach adulthood and require less investment to reach adulthood than younger children, current or future [@daly1995].
-For this reason, less costly signals may be more likely to work to increase investment, all else being equal\[\[4\]\].
-That said, all else may often not be equal if investment in older children is costlier, something which may be reflected in caretakers seeing older children as relatively needy.
-The decrease in crying and tantrum frequencies with age may also result from the potential for greater reputational costs, something which may also increase the credibility of such signals among older children as previously mentioned.
+One possible reason for this is that as children mature, they are more capable of fulfilling their own desires, more able to ask others for what they want, and less vulnerable to threats.
+The effects of age on signaling are also likely due to how age influences the costliness of signaling need.
+For example, while crying may be costly among the youngest children through its energetic costs, these costs are likely insignificant among older children, who might instead face costs in the form of reputational harm from crying in some contexts but not others.
+In times when the older child's valuation of potential increased support is small, such costs may be prohibitive.
+However, in times when the child places more value on potential support, those same costs could allow for a child to credibly signal need if they are too high to be worth it for deceptive signaling.
+- Increased use of signals not mentioned here (.e.g, depression, suicality, others?)
 
-The effect of age was smallest on child sadness.
-However, interpretation of this finding is more difficult due to the fact that, unlike crying and temper tantrums, sadness is thought to have important non-signaling functions including drawing one’s attention to a problem, motivating deeper analysis of one’s situation, and causing withdrawal from situations with low fitness prospects [@nesse2019; @nettle2009; @allen2003; @hagen2011].
-By asking about sadness as a whole, our measure was unable to disentangle information caretakers gleaned from cues from information that came from signals.
+~~Another is that since older children require less investment to reach adulthood than younger children, current or future, they may experience less severe conflicts of interest with  parents.
+For this reason, less costly signals may be more likely to work to increase investment, all else being equal\[\[4\]\].~~
 
@@
 Although conflicts of interest do not always lead to behavioral conflict [@altmann1980; @bateson1994], it is likely many, if not most, conflicts are ultimately caused by conflicts of interest, potentially making it a useful proxy for the underlying conflicts of interest that may increase the benefits of deception for children and the benefits of greater skepticism for parents\[\[5\]\].
-If this is the case, the correlation between signaling and conflict may represent children's attempts at credibly signaling their value of additional investment or the costliness of punishment [@syme2019] to parents who would have good reason to doubt the trustworthiness of simple verbal requests (or the trustworthiness of facial sadness and crying for those resorting to temper tantrums).
+If this is the case, the positive correlation between signaling effort and conflict frequency may represent children's attempts at credibly signaling their value of additional investment or the costliness of punishment [@syme2019] to parents who would have good reason to doubt the trustworthiness of simple verbal requests (or the trustworthiness of facial sadness and crying for those resorting to temper tantrums).
 However, it is also likely that parents find some forms of child signaling aversive, and therefore see it as conflict in and of itself.
@@
 
+\[still need the plot for this\]
+
 We found evidence of an interaction between how often a child cared for their siblings and the sex of the child in our lasso models examining signal frequencies and cost, whereby there was a positive relationship between the frequency and summed costliness of child signals of need and alloparenting effort for girls but no relationship or a slightly negative one for boys.
-In many societies, girls do more alloparenting than boys [@helfrecht2016 @weisner1977; @barry1971; @turke1988], something which matches findings from focus groups of Utilian mothers \[add info from an analysis here?\].
-If this difference was the outcome of caretakers pressuring girls to do more or resulted in the expectation of more alloparenting, the increase in signaling with alloparenting effort seen among girls may be the result of a strategy to signal the costliness of their efforts in order to invest more in romantic relationships and status seeking behaviors outside the household.
+In many societies, girls do more alloparenting than boys [@helfrecht2016; @weisner1977; @barry1971; @turke1988], something which matches findings from focus groups of Utilian mothers \[add info from an analysis here?\].
+If this difference was the outcome of caretakers pressuring girls to do more or resulted in the expectation of more alloparenting, the increase in signaling with alloparenting effort seen among girls may be the result of a strategy to signal the costliness of their efforts in order to invest more in relationships and status seeking behaviors outside the household.
 
@@
 
-\[delete based on lasso coefficients?\] However, child alloparenting effort was negatively associated with caregiver reports of child conflict.
+However, child alloparenting effort was negatively associated with caregiver reports of child conflict.
 One potential reason for this is since parents presumably value the child alloparenting within the family, they see their relationships with children who invest more in the household as less conflictual.
-In line with this reasoning, we found evidence of a child alloparenting effort by child age interaction, with our model predicting increased levels of conflict for older children with greater alloparenting.
+In line with this reasoning, we found evidence of a child alloparenting effort by child age interaction, with our model predicting increased levels of conflict for older children with greater alloparenting \[if we say this we should report the results of that lasso too\].
 Ideally, we could control for signaling behavior to test this.
@@
 
-There was a positive association between the summed frequency of child signaling and caretaker perceptions of child neediness within the family, something which would be consistent with sadness, crying, and temper tantrums being signals (or involving signals in the case of crying) if parents took this to mean how much they value additional investment.
+\[Am I missing anything about the significance of this being a relative measure. Could faking drive up the costs in ways everyone signals similarly\]
+
+There was a positive association between the frequency of child sadness and caretaker perceptions of child neediness within the family.
+However, the other signaling measures examined had coefficients set to 0 in the same lasso model despite the expectation that their greater costliness would make them more credible sources of information to parents.
+One reason for this might be that costs of sadness, while minor short-term, can be significant long-term.
+In line with this reasoning, depression, which often involves an exteneded period of a lack of interest in one's normal activities and decreased investment in oneself and one's cooperative relationships, has been suggested to capable of credibly signal need in times of substantial conflicts of interest [@watson2002; @hagen2003].
+
+Another possibility is that parents are using cues from sadness (i.e., withdrawal and rumination) to evaluate the relative neediness of children.
+As cues are not created for the purpose of communication, they are completely reliable if perceived and interpreted correctly.
+Therefore, if some cues are hard to impossible to fake, they may be more reliable than most signals.
+That said, it is likely that some symptoms of sadness are simultaneously the result of both signals and cues, something which may have been favored by natural selection if the threat of deception meant cue-level symptoms were insufficient in eliciting support without upregulation, something which has been suggested to be the case with various illness symptoms [@tiokhin2016; @steinkopf2015].
 
 Caretakers perceived older children as needier, despite having more conflicts with younger children.
-One possible reason for this might be that caretakers see the types of investment older children seek as more costly.
+One likely reason for this might be that caretakers percieve the types of investment older children seek as more costly.
 In line with this reasoning, multiple caretakers mentioned that older kids need more money and advice and get into trouble more often during surveys.
@@
 
-[still needs updating]
-Signaling behavior did not predict (try more models?) differences in relative caregiver investment between children within households.
-However, crying frequency was associated with both height and weight gain for older children.
-If older children face greater reputation costs from crying, crying may be a relatively credible credible signal compared to crying among younger children, with it being a less viable strategy to exaggerate one's value of investment or the costliness of adversity.
+Signaling behavior did not predict differences in relative caregiver investment between children within households.
+However, greater perceptions of need among caregivers were associated with increased likelihood of a child being reported to receive more investment in terms of time, effort, and resources relative to other children within the household.
+As previously mentioned, credible signals are capable of leading to pro-signaler responses due to their ability to reveal private information to others who would be motivated to increase their investment in another should they have that information.
+In other words, much of the effect of signaling on levels of support is expected to be mediated by the change in the beliefs of others that one could benefit from increased support that it causes (e.g., that a child is truly in need due to adversity) rather than direct effects.
+In line with this reasoning, an experimental vignette study which found that depression and suicidality increased the likelihood of costly support from participants in times of uncertainty and substantial conflicts of interest also found that between 67% and 78% of the effect of signal type on the likelihood of costly support was mediated by the effect of signaling on belief in need [@gaffney2022].
 
@@
 
-Monthly income had a modest effect on crying frequency and summed signal frequency in our lasso regressions.
-Many models designed to explain signals of need (e.g., the begging of young birds) predict that offspring who are worse off or are receiving less investment have more to gain from signaling than those with more [@mock2011; @godfray1991; @caro2016a; @godfray2000].
-However, it is possible that this can be true and household level resources have little influence on child signaling frequencies between families.
-As signaling behavior is costly to child, caretakers, and siblings, children may gain direct and indirect fitness benefits from limiting their signaling behavior in times of resource stress (i.e., when the chances of gaining investment or the size of investment are small).
-Similarly, children in high resource families may also be motivated to signal at high frequencies due to the potential for larger pieces of investment.
-We know of no empirical explorations of this in humans (check more and confirm).
-However, multiple game theoretical models examining parent-offspring conflict have found that parental resources have a modest impact on the evolutionary stable value for offspring demand for resources [@bossan2013; @mock1997 [confirm]].
-Compared to overall effort in eliciting need, resource availability may have a larger impact on how children signal for more investment based on the finding that birds in stable and unusually favorable environments are more likely to compete with siblings through signals thought to reveal information about need, while those in relatively unpredictable and unusually poor environments have been found to compete more through signals thought to reveal information about quality [@caro2016a].
+Despite the differences in wealth, job certainty, discrimination, and living conditions [@davis2014; @garcia2017; @garcia2018; @stonich2000] between recent immigrants from the mainland and native Utilians, primary language, immigration status, and residence in Camponado did not predict child signaling in our lasso regressions.
+However, neighborhood quality, which was strongly negaltively correlated with residence in Camponado, (check others) was preserved in the lasso regressions predicting every signaling measure but sadness and was among the variables with the largest negative coefficients.
+Monthly income was also negatively associated with crying frequency,~~ tantrum frequency (too small of a coefficient to mention here?),~~ summed signal frequency, and signal cost albeit with smaller coefficients.
 
-Despite the differences in wealth, job certainty, discrimination, and living conditions [@davis2014; @garcia2017; @garcia2018; @stonich2000] between recent immigrants from the mainland and native Utilians, primary language, immigration status, and residence in Camponado did not predict child signaling in our lasso regressions.
-However, neighborhood quality was preserved in the lasso regressions predicting every signaling measure but sadness and was among the variables with the largest negative coefficients.
-One reason for this might that neighborhood quality affects the amount of adversity children face, thereby increasing their neediness \[no evidence of this when neighborhood quality is included in models predicting relative child need.
-\]
+One reason for this might be that neighborhood quality affects the amount of adversity children face, thereby increasing their neediness.
 
-\[old\] Despite population level differences in wealth, job certainty, discrimination, and living conditions [@davis2014; @garcia2017; @garcia2018; @stonich2000], we generally did not find an effect of population on child signaling behavior.
-Although we sought data from both English-speaking participants born on the island and Spanish-speaking immigrants from the mainland, our smaller sample of English-speaking Utilians may have limited our ability to detect real differences between the populations.
-It is possible that differences in wealth and status between the groups translate to differences in child signaling due to children facing different types adversity or weighting different types of investment differently.
-It is also likely that the two groups differ in how caretakers could respond to child signaling, and caretakers between the two populations may also differ in their cultural models of parenting.
-However, as previously mentioned, it is conceivable that household level differences have little effect on child signaling behavior.
-
-possessions
+Many models designed to explain signals of need (e.g., the begging of young birds) predict that offspring who are worse off or are receiving less investment have more to gain from signaling than those with more [@mock2011; @godfray1991; @caro2016a; @godfray2000] and multiple game theoretical models examining parent-offspring conflict have found that parental resources have a modest impact on the evolutionary stable value for offspring demand for resources [@bossan2013; @mock1997 [confirm]].
+In line with this reasoning, children in lower income households and riskier environments might benefit more from signaling effort if they experience greater levels of adversity or perceive themselves as disadvantaged compared to other children they may be competing with for status and desirable cooperative relationships.
+~~In addition to overall effort in eliciting need, resource availability may have also affect how children signal for more investment based on the finding that birds in stable and unusually favorable environments are more likely to compete with siblings through signals thought to reveal information about need, while those in relatively unpredictable and unusually poor environments have been found to compete more through signals thought to reveal information about quality [@caro2016a].~~
 
@@
 ::: {#stbl-relvars}
-
 |  |  |
-|----------------------|--------------------------------------------------|
+|------------------------|------------------------------------------------|
 | Variable | Survey question wording |
@@
 
+::: {#stbl-medical-record}
 |  |  |
-|------------------|------------------------------------------------------|
+|---------------------|---------------------------------------------------|
 | Variable | Survey question wording |
@@
 Variable names and question wording for measures about a child's health history.
-Both questions were asked using a Likert scale from 1-5 where 1 was labeled "Strongly disagree" and 5 was labeled "Strongly agree." {#stbl-medical-record}
+Both questions were asked using a Likert scale from 1-5 where 1 was labeled "Strongly disagree" and 5 was labeled "Strongly agree."
+:::
 
+::: {#stbl-healthvars}
 |  |  |
-|-----------------------|-------------------------------------------------|
+|------------------------|-----------------------------------------------|
 | Variable | Survey question wording |
@@
 
-: Variable names and question wording for measures about a child's current likelihood of resisting illness.
+Variable names and question wording for measures about a child's current likelihood of resisting illness.
 All three questions were asked using a Likert scale from 1-5 where 1 was labeled "Strongly disagree" and 5 was labeled "Strongly agree." The scores for IllnessAvoidance1 and IllnessAvoidance2 were inverted so that 5 corresponded with the lowest susceptibility to illness, i.e., greatest *illness avoidance*.
-{#stbl-healthvars}
+:::
 
+::: {#stbl-Food-Security}
 |  |  |
-|------------------|------------------------------------------------------|
+|---------------------|---------------------------------------------------|
 | Variable | Survey question wording |
@@
 
-: Variable names and question wording for the modified version of the USDA Food Security Questionnaire [@blumberg1999].
-All questions were asked using a Likert scale from 1-5 where 1 was labeled “Strongly disagree” and 5 was labeled “Strongly agree.” {#stbl-Food-Security}
+Variable names and question wording for the modified version of the USDA Food Security Questionnaire [@blumberg1999].
+All questions were asked using a Likert scale from 1-5 where 1 was labeled “Strongly disagree” and 5 was labeled “Strongly agree.”
+:::
 
+::: {#stbl-Home-Instability}
 |  |  |
-|--------------------|----------------------------------------------------|
+|----------------------|-------------------------------------------------|
 | Variable | Survey question wording |
@@
 
-: Variable names and question wording for the home instability questions.
-All questions were asked using a Likert scale from 1-5 where 1 was labeled “Disagree strongly” and 5 was labeled “Agree Strongly.” {#stbl-Home-Instability}
+Variable names and question wording for the home instability questions.
+All questions were asked using a Likert scale from 1-5 where 1 was labeled “Disagree strongly” and 5 was labeled “Agree Strongly.”
+:::
 
+::: {#stbl-Neighborhood-Quality}
 |  |  |
-|-----------------------|-------------------------------------------------|
+|------------------------|-----------------------------------------------|
 | Variable | Survey question wording |
@@
 
-: Variable names and question wording for the perceptions of neighborhood quality questions.
-All questions were asked using a Likert scale from 1-5 where 1 was labeled “Strongly disagree” and 5 was labeled “Strongly agree.” {#stbl-Neighborhood-Quality}
+Variable names and question wording for the perceptions of neighborhood quality questions.
+All questions were asked using a Likert scale from 1-5 where 1 was labeled “Strongly disagree” and 5 was labeled “Strongly agree.”
+:::
 
+::: {#stbl-response-valence}
 Positive: Give resources (e.g., toys, food, money)\
@@
 
-: List of options for research assistant to classify participant responses to the question of how they responded when their child was last sad or cried. {#stbl-response-valence}
+List of options for research assistant to classify participant responses to the question of how they responded when their child was last sad or cried.
+:::
+
+::: {#stbl-possession-list}
+Motorized vehicle (scooter, 4-wheeler, etc.) Bicylce Television (TV) with cable Smartphone Washing machine Refrigerator A home that you own Credit card
 
 List of possessions owned (yes or no).
-
-Motorized vehicle (scooter, 4-wheeler, etc.) Bicylce Television (TV) with cable Smartphone Washing machine Refrigerator A home that you own Credit card
+:::
 
@@
 :::
+
+## Coefficient plots
+
+::: {#sfig-plot-need-coefs}
+![](Figures/plot_need_coefs.svg)
+Effects of predictor variables on relative need between household children. Ordinal logistic regression model with elastic net penalties.
+:::
+
+::: {#sfig-plot-invest-coefs}
+![](Figures/plot_invest_coefs.svg)
+Effects of predictor variables on relative investment in household children. Ordinal logistic regression model with elastic net penalties.
+:::
+
+::: {#sfig-plot-caregiverresponse-coefs}
+
+![](Figures/plot_caregiverresponse_coefs.svg)
+Effects of predictor variables on caregiver responses to the last time a child was sad or cried. Ordinal logistic regression model with elastic net penalties.
+:::
+
+# Deleted sections
+
+The effect of age was smallest on child sadness.
+However, interpretation of this finding is more difficult due to the fact that, unlike crying and temper tantrums, sadness is thought to have important non-signaling functions including drawing one’s attention to a problem, motivating deeper analysis of one’s situation, and causing withdrawal from situations with low fitness prospects [@nesse2019; @nettle2009; @allen2003; @hagen2011].
+By asking about sadness as a whole, our measure was unable to disentangle information caretakers gleaned from cues from information that came from signals.

### 2024-10-30 dataprep.R

@@ -267,12 +267,18 @@ d2 <-
     LogIncome = log10(IncomeCategoryN),
     CaregiverAge = ifelse(is.na(CaregiverAge), mean(CaregiverAge, na.rm = TRUE), CaregiverAge), # imputing age for 1 caretaker where it is missing (affects 2 children). Also imputes caregiver ages for 7 other children. But these children do not have signaling data.
     StayAtHomeMom = ifelse(str_detect(CurrentJob, "home"), 1, 0),
+    StayAtHomeMomF = as.factor(case_when(
+      StayAtHomeMom == 0 ~ "No",
+      StayAtHomeMom == 1 ~ "Yes")),
     BodyFat = c(scale(TricepR) + scale(SubscapR)),
     RecentSignalSeverity = com1[as.character(Sad1)],
     RecentSignalCause = com2[as.character(Sad1)],
     AdultsHousework = as.numeric(AdultsHousework),
     #AdultsChildcare = as.numeric(AdultsChildcare),
     Neighborhood2 = as.numeric(Neighborhood == "Camponado/Campolancho"),
+    Neighborhood2F = as.factor(case_when(
+      Neighborhood2 == 0 ~ "No",
+      Neighborhood2 == 1 ~ "Yes")),
     NeighborhoodF = as.factor(Neighborhood),
     OtherChildrenHH = number_children2 - 1,
     Punishment2 = ifelse(Punishment == "Maybe", 0, Punishment),
@@ -336,7 +342,6 @@ modeldf <-
   dplyr::select(
     householdID,
     childHHid,
-    Sex,
     CryFreqN,
     SadFreqN,
     TantrumFreqN,
@@ -346,6 +351,7 @@ modeldf <-
     ChildAge,
     OtherChildrenHH,
     YoungerKids,
+    OtherChildAlloparentingFreqN,
     OlderGirls,
     OldestChild,
     LogIncome,
@@ -356,14 +362,19 @@ modeldf <-
     IllnessSusceptibility1,
     IllnessSusceptibility2,
     IllnessSusceptibility3,
+    MedicalProblemsMean,
     IllnessSusceptibilityMean,
+    AdultsChildcare,
+    AdultsHousework,
     SadFreqOF,
     CryFreqOF,
     TantrumFreqOF,
     ConflictFreqOF,
     AlloparentingFreq0,
+    NeighborhoodQuality,
+    HouseQuality,
     NumberOfChildren,
[…]

### 2024-10-30 dictionaries.R

@@ -2,15 +2,28 @@
 longform_dict <-
   c(
     ChildAge = "Child age (years)",
+    CaregiverAge = "Caregiver age (years)",
     SadFreqN = "Sad frequency (times per month)",
     CryFreqN = "Crying frequency (times per month)",
     TantrumFreqN = "Temper tantrum frequency (times per month)",
     AlloparentingFreqN = "Frequency of child's own alloparenting effort (times per month)",
+    OtherChildAlloparentingFreqN = "Frequency of other child alloparenting (times per month)",
     ConflictFreqN = "Frequency of conflict with focal caretaker (times per month)",
     SignalFreq = "Frequency of signals summed (times per month)",
     SignalCost = "Signal cost",
     SignalFreqMax = "Frequency of most common signal (times per month)",
-    IllnessSusceptibilityMean = "Illness avoidance index",
+    NumberOfChildren = "Children in household",
+    MeanChildRelatedness = "Child relatedness",
+    number_adults = "Number of adults in household",
+    AdultsChildcare = "Household adults who provide childcare",
+    AdultsHousework = "Household adults who provide, food, housework, and othe neccesities",
+    EducationLevelYears = "Caregiver education level (years)",
+    IncomeCategoryN = "Monthly household income (Lempira)",
+    FoodSecurity = "Household food insecurity",
+    NeighborhoodQuality = "Neighborhood quality index",
+    HouseQuality = "House quality index",
+    MedicalProblemsMean = "Child medical problems index",
+    IllnessSusceptibilityMean = "Child illness avoidance index",
     HeightZ = "Height Z-score (CDC)",
     WeightZ = "Weight Z-score (CDC)",
     BMIZ = "BMI Z-score (CDC)",
@@ -18,13 +31,30 @@ longform_dict <-
     SubscapR = "Subscapular skinfold residuals",
     TricepMean = "Tricep skinfold thickness (mm)",
     TricepR = "Tricep skinfold residuals",
-    BodyFat =  "Body fat composite", # child summary stats end here (1:18)
-    CaregiverAge = "Caregiver age (years)",
-    EducationLevelYears = "Caregiver education level (years)",
-    IncomeCategoryN = "Monthly household income (Lempira)",
-    number_adults = "Number of adults in household",
-    AdultsChildcare = "Household adults who provide childcare" # adults summary table (19:23)
+    BodyFat =  "Body fat composite",
[…]

### 2024-10-30 regularize.R

@@ -321,6 +321,8 @@ ordinal_plot <- function(fit, predictor, data, title){
       {{predictor}}, Negative:Positive
     ) |>
     pivot_longer(Negative:Positive)
+  # x_axis_name <- shortform_dict[deparse(substitute(predictor))]
+  # print(x_axis_name)
   ggplot(d, aes({{predictor}}, value, colour = name)) +
     geom_count(position = position_dodge(width = 0.1)) +
     geom_smooth(method='lm', se = F, position = position_dodge(width = 0.1)) +
@@ -388,6 +390,9 @@ colMeans(summary(out))
 # coef(out$fit, matrix = TRUE, whichLambda = 1)
 plot_need_coefs <- ggdotchart(coef(out$fit)[-c(1:2)])
 plot_need_coefs
+ggsave("Figures/plot_need_coefs.pdf", plot_need_coefs, width = 12, height = 12)
+ggsave("Figures/plot_need_coefs.svg", plot_need_coefs, width = 12, height = 12)
+
 plot_need_age <- ordinal_plot2(out$fit, ChildAge, SignalVars4, 'Relative need')
 plot_need_sad <- ordinal_plot2(out$fit, SadFreqN, SignalVars4, 'Relative need')
 
@@ -421,6 +426,9 @@ colMeans(summary(out))
 # coef(out$fit, matrix = TRUE, whichLambda = 1)
 plot_invest_coefs <- ggdotchart(coef(out$fit)[-c(1:2)])
 plot_invest_coefs
+ggsave("Figures/plot_invest_coefs.pdf", plot_invest_coefs, width = 12, height = 12)
+ggsave("Figures/plot_invest_coefs.svg", plot_invest_coefs, width = 12, height = 12)
+
 plot_invest_need <- ordinal_plot2(out$fit, RelativeNeed3, SignalVars5, 'Relative investment')
 ggsave("Figures/plot_invest_need.pdf", plot_invest_need, width = 12, height = 12)
 ggsave("Figures/plot_invest_need.svg", plot_invest_need, width = 12, height = 12)
@@ -463,6 +471,10 @@ summary(out)
 colMeans(summary(out))
 # coef(out$fit, matrix = TRUE, whichLambda = 1)
 plot_caregiverresponse_coefs <- ggdotchart(coef(out$fit)[-c(1:2)])
+plot_caregiverresponse_coefs
[…]

### 2024-11-01 Paper.qmd
Paper: Added abstract and conclusion. Edits to discussion and methods.
Regularize: Added coded for full correlation matrix.

@@
 
+# Abstract
+
+To gain support, children use signals to communicate their desires to parents.
+Infant signals of need, particularly infant cries, have been extensively studied in a range of populations.
+However, the full range of potential child signals of need, which extend beyond cries, has never been investigated in children of all ages.
+To address this, we used exploratory analyses on caregiver reports of child signaling data from xxx families (xx children) on Utila, Honduras.
+We found that child signaling decreased with age and neighborhood quality and was highly correlated with the frequency of conflict between children and caretakers.
+Consistent with signaling theory, children who were perceived as needier within the household were more likely to receive more investment.
+They were also more likely to receive a pro-child response when the cause of their last instance of signaling involved discomfort, pain, injury, or, illness, but less likely to receive a beneficial response when it involved prior punishment.
+Our results show there is much potential to the further study of sadness, crying, and temper tantrums among older children using signaling theory.
+
 # Introduction
@@
 Compared to chimpanzee mothers, mothers in hunter-gatherer societies have shorter interbirth intervals and produce offspring that require many more years of parental care [@davison2021; @davison2022].
-Unlike chimpanzee offspring, human offspring in ancestral populations therefore often grew up with multiple older and/or younger siblings who were potential competitors for parental investment and yet who could also provide (or require) allocare [@xx, @yy].
+Unlike chimpanzee offspring, human offspring in ancestral populations therefore often grew up with multiple older and/or younger siblings who were potential competitors for parental investment and yet who could also provide (or require) allocare [@helfrecht2016; @lawson2016; hagen2009].
 
@@
 Temper tantrums may also be adaptive strategies by which children can elicit more support from parents.
-They are common in children from 18 months to 4 years old [@potegal2003], and similar behaviors have also been reported in other young animals, often surrounding feeding or weaning conflict [@altmann1980; @goodall1986; @schaller1964].
-Temper tantrums vary in their severity, ranging from crying, shrieking, and screaming to violent bodily movements, aggression, and self-harm [@potegal2003].
-They have been reported in many societies with different subsistence systems, political organization, and norms surrounding anger and parenting [@albino1956; @banisalameh2021; @fouts2005; @henderson1966; @hilger1957; @konner1974; @ward1970], suggesting they may be a human universal.
+They are common in children from 18 months to 4 years old [@potegal2003] yet are understudied among older children [@koch2003] outside of studies examining disruptive mood regulation disorder [for one exception see @book].
+Temper tantrums vary in their severity, ranging from crying, shrieking, and screaming to violent bodily movements, aggression, and self-harm [@potegal2003], with similar behaviors having also been reported in other young animals, often surrounding feeding or weaning conflict [@altmann1980; @goodall1986; @schaller1964].
+Tantrums have been reported in many societies with different subsistence systems, political organization, and norms surrounding anger and parenting [@albino1956; @banisalameh2021; @fouts2005; @henderson1966; @hilger1957; @konner1974; @ward1970], suggesting they may be a human universal.
 As with the other proposed signals of need, they are capable of eliciting a wide range of responses, including pro-child responses, neutral responses (e.g., parents ignoring the child) [@hilger1957; @konner1974], and negative responses like abuse [@schmitt1987].
@@
 A major challenge when studying signals of need is the fact that such signaling is expected to 1) be more common in times when individuals can benefit from increased investment (e.g., when they are facing major adversity or when a fitness-relevant opportunity arises) and 2) result in benefits to the signaler on average.
-This hypothesized causal, bi-directional relationship between investment and signaling makes it difficult to isolate the effects of signaling behavior on investment in children, especially given the lack of prior studies on signaling by older children in low income settings.For this reason, all analyses of these unique data were exploratory and meant to lay the foundation for future studies of child signaling.
+This hypothesized causal, bi-directional relationship between investment and signaling makes it difficult to isolate the effects of signaling behavior on investment in children, especially given the lack of prior studies on signaling by older children in low income settings.
+For this reason, all analyses of these unique data were exploratory and meant to lay the foundation for future studies of child signaling.
 
@@
 
-XXXXXXXXCHIld RecruitmentXXXXXXXXXXX
+[no longer needed] XXXXXXXXCHIld RecruitmentXXXXXXXXXXX
 
@@
 All participants for the survey portion of the study met the inclusion criteria.
-Most adult participants were the biological mothers of one or more children in the study.
-All children in the study, whose ages ranged from 5-20, were cared for by one of the adult participants.
+Most ~~adult~~ participants were the biological mothers of one or more children in the study.
+~~All children in the study, whose ages ranged from 5-20, were cared for by one of the adult participants.~~
 
@@
 
-All adult participants provided informed consent.
+All ~~adult~~ participants provided informed consent.
 The surveys, which included many topics outside of child signaling, were designed to last about an hour, plus 20 minutes for each additional child.
-Adults were paid \$XXXXX for the survey portion of the study.
+~~Adults~~ Participants were paid \$XXXXX for the survey portion of the study.
 
-xxxxxxxxxxxxxChild consent infoxxxxxxxxxxxxxx.
+~~xxxxxxxxxxxxxChild consent infoxxxxxxxxxxxxxx.~~
 
@@
 The survey also included many demographic questions about the caretakers and their households.
-Those used in this study included the (1) ages of those in the household, (2) sexes of those in the household, (3) kinship ties within households, (4) partnership status of the primary caregiver, (5) number of adults who provide childcare within the household, (6) highest level of education of the primary caregiver, (7) immigration status (in regards to Utila), (8) monthly household income, (9) neighborhood of residence, (10) current job of the primary caretaker, and (11) ownership of various possesions (e.g., homes, washing machines, or cellphones; see SIXXXXXX for full list).
+Those used in this study included the (1) ages of those in the household, (2) sexes of those in the household, (3) kinship ties within households, (4) partnership status of the primary caregiver, (5) number of adults who provide childcare within the household, (6) highest level of education of the primary caregiver, (7) immigration status (in regards to Utila), (8) monthly household income, (9) neighborhood of residence, (10) current job of the primary caretaker, and (11) ownership of various possessions (e.g., homes, washing machines, or cellphones; see @stbl-possession-list for full list).
 These data allowed us to calculate the number of children (total, younger, and older) and adults within the household as well as the genetic relatedness between siblings.
-We also imputed data for caregiver age by setting the missing value to the mean.
-For the data used in analyses, this amounted to imputing data for one mother which affected two children.
+We also created a variable for residence in Camponado, a densely populated neighborhood on the island with relatively poor sanitation in which most individuals from the mainland live that is seen as the most dangerous and poorest part of Utila [@garcia2017].
 
-\[mention Campanado here and in ethnographic portion?\]
-
-\[I think this gets deleted or do we mention no findings worth reporting?\] \## Child anthropometrics
-
-xxxxxxxxxxDescribe camps brieflyxxxxxxxxxxxxxx - include dates and how we handled those.
-For cross-sectional analyses, we used the 2023 anthropometric values due to them being closer in time to the caretaker surveys.
-
-All anthropometric measurements were taken twice, in accordance with anthropometric standards [@lohman1988], and then averaged.
-Before being measured, participants were asked to remove their shoes and heavy articles of clothing.
-Height was measured to the nearest millimeter using an anthropometer, and weight was measured to the nearest 100 g with a digital scale (model name) that was used in a building with a concrete foundation.
-All participants were wearing lightweight clothing, so we made no adjustments for clothing weight.
-Grip strength was measured using a dynamometer (model name), with participants seated/standing?
-and told to xxxxxxxx.
-XXXX time was passed between the first and second measurements.
-
-Triceps and subscapular skinfold thicknesses were measured using skinfold calipers.
-The triceps measurement was taken at the midpoint of the upper arm.
-The subscapular measurement occurred immediately below the inferior angle of the scapula, being picked up diagonally at the natural fold line of the skin [@antóna].
-Flexed bicep circumference was measured perpendicular to the long axis of the upper arm.
-For this measure participants were asked to sit/stand?
-and position their upper arm perpendicular to their body with their elbow being held at a 90° angle, before flexing their clenched fist towards themselves as much as they could.
-
-We computed each child's height, weight, and BMI Z-scores with respect to the CDC child growth standards [@kuczmarski2002].
-
-For our independent variables of skinfold thickness, flexed bicep circumference, and grip strength on signaling behavior, we used the residuals from generalized additive models which modeled the effect of sex and age on each measurement using the mgcv package [@wood2011].
-The resulting variables represent how each child differed from the average child for their given sex and age.
-We also centered and scaled the skinfold thickness residuals before summing them to create a body fat index.
+In one instance where data for caregiver age was missing, we set the missing value to the mean. This affected data for two children.
 
@@
 
-<!-- [NOTE: Add categorical variables to the table; combine child and adult variables into 1 table] -->
-
 ```{r sumtab_combined}
@@
 
-kable(custom.summarize(modeldf, longform_dict[1:31])) %>% kable_styling()
+kable(custom.summarize(modeldf, longform_dict[1:23])) %>% kable_styling()
 ```
@@
 Correlation matrix of the frequencies of signals/states covered in the study with sex, child age, and the frequency of conflict between children and primary caretakers.
-For a more complete correlation matrix of study variables see SIxxxxxxx.
+For a more complete correlation matrix of study variables see @sfig-plot-full-corrmat.
 :::
@@
 
-The effects of particularly large risk or protective factors for child signaling are shown in @fig-signal-effects.
+The effects of particularly large risk or protective factors for child signaling are shown in @fig-signal-effects and @fig-alloparentingXsex.
 
@@
 Effects of selected predictors of child signaling frequencies when holding continuous variables at their mean and factor variables at their mode.
+:::
+
+::: {#fig-alloparentingXsex}
+![](Figures/plot_alloparentingXsex.svg)
+
+Effects of alloparenting frequency (per month) and signal cost and frequency of temper tantrums by sex.
 :::
@@
 
-*Negative association between age and child signaling (all signaling measures)*
+Caregivers reported a wide range of child signaling frequencies.
+About half of children were reported to cry or throw tantrums more than once a month and slightly less than 40% were reported to engage in them more than once a week.
+Although similarly common overall, tantrums more common than crying among older children and might be a under-appreciated way by which children might signal for increased support based on these findings and a longitudinal study of temper tantrums and otherchild behaviors which found 50% of boys 30% of girls engaged in temper tantrums at age 9 among a representative, non-clincial sample of 252 children in Berkeley, California [@@macfarlane1962].
 
-As seen in Western cultures [@zeifman2013; @potegal2003; @koch2003] (check for others), the frequencies of crying and temper tantrums decreased with age, as did the frequency of child sadness, summed signal frequency, and signal cost.
+As seen in Western cultures [@zeifman2013; @potegal2003; @koch2003; @macfarlane1962], the frequencies of crying and temper tantrums decreased with age, as did the frequency of child sadness, summed signal frequency, and signal cost.
 One possible reason for this is that as children mature, they are more capable of fulfilling their own desires, more able to ask others for what they want, and less vulnerable to threats.
@@
 For example, while crying may be costly among the youngest children through its energetic costs, these costs are likely insignificant among older children, who might instead face costs in the form of reputational harm from crying in some contexts but not others.
-In times when the older child's valuation of potential increased support is small, such costs may be prohibitive.
+In times when an older child's valuation of potential increased support is small, such costs may be prohibitive.
 However, in times when the child places more value on potential support, those same costs could allow for a child to credibly signal need if they are too high to be worth it for deceptive signaling.
-- Increased use of signals not mentioned here (.e.g, depression, suicality, others?)
-
-~~Another is that since older children require less investment to reach adulthood than younger children, current or future, they may experience less severe conflicts of interest with  parents.
-For this reason, less costly signals may be more likely to work to increase investment, all else being equal\[\[4\]\].~~
-
-*Positive association between conflict with parent and child signaling (all signaling measures)*
 
@@
 However, it is also likely that parents find some forms of child signaling aversive, and therefore see it as conflict in and of itself.
-
-*Sex and focal child alloparenting effort interact (positive for females, negative/flat for males)*
-
-\[still need the plot for this\]
 
@@
 It is also possible that our hypothesis about conflict over investment in the household is incorrect.
-
-*Positive associations between child age and signaling frequency and child neediness*
-
-\[Am I missing anything about the significance of this being a relative measure. Could faking drive up the costs in ways everyone signals similarly\]
 
@@
 
-*Parental responses to last signal of need*
-
 Parents were more likely to respond positively and less likely to respond negatively to the last time a child signaled when the cause involved discomfort, pain, injury, or illness than when it did not.
@@
 
-\[delete?\] *In most models, there is a negative association between education and child signaling?*
-
+\[delete this paragraph?\] 
 We found a negative association between caretakers’ education level (years in school) and many of our child signaling measures despite our models controlling for household income and partner status.
@@
 It is also possible that education influences parenting norms in ways that cause differences in child signaling behavior.
-
-*Demographic effects*
 
@@
 Monthly income was also negatively associated with crying frequency,~~ tantrum frequency (too small of a coefficient to mention here?),~~ summed signal frequency, and signal cost albeit with smaller coefficients.
-
-One reason for this might be that neighborhood quality affects the amount of adversity children face, thereby increasing their neediness.
 
@@
 
-Outside of child anthropometric measurements, all data come from caregiver reports, and therefore, reflect the biases of caretakers, whatever they may be.
+~~Outside of child anthropometric measurements,~~ All data come from caregiver reports, and therefore, reflect the biases of caretakers, whatever they may be.
 Due to the fact the data used here were a part of survey which covered many topics and often multiple children, we also faced a tradeoff between survey time and question depth.
@@
 Our questions about the last time the child signaled may have also been more effective if they focused on specific signals due to our methods combining crying, which is unambiguous as a signal, with sadness, which has both signaling and non-signaling functions.
+
+# Conclusion
+
+Children face the challenge of eliciting support from others in times of adversity and opportunity.
+The signaling strategies of Utilian children were highly correlated and engaged in by children of all ages.
+Consistent with signaling models which emphasize the role of conflicts of interest in favoring the use of costlier signals, 1) child signaling effort was positively correlated with the frequency of conflict with caretakers and 2) greater perceptions of child need predicted greater investment in a given child.
+Our results suggest there is much potential for signaling theory to help explain patterns of child crying and temper tantrums.
 
@@
 ::: {#stbl-possession-list}
-Motorized vehicle (scooter, 4-wheeler, etc.) Bicylce Television (TV) with cable Smartphone Washing machine Refrigerator A home that you own Credit card
+Motorized vehicle (scooter, 4-wheeler, etc.) Bicycle Television (TV) with cable Smartphone Washing machine Refrigerator A home that you own Credit card
 
@@
 
-::: {#sfig-effect-relativeneed}
-```{r, fig.width= 10}
-p_need_age_cost
-```
+::: {#sfig-plot-full-corrmat}
+![](Figures/plot_fullcorrmat.svg)
 
-The proportion of children perceived by caretakers as being more needy, as needy, or less needy than a caretaker's other children.
-A: Effect of child age on relative neediness B: Effect of signal cost on relative neediness.
-Outcome variable question wording: How much time, effort, and resources does this child need compared to your other children?
+Complete correlation matrix of study variables.
 :::
@@
 ![](Figures/plot_need_coefs.svg)
+
 Effects of predictor variables on relative need between household children. Ordinal logistic regression model with elastic net penalties.
@@
 ![](Figures/plot_invest_coefs.svg)
+
 Effects of predictor variables on relative investment in household children. Ordinal logistic regression model with elastic net penalties.
@@
 ![](Figures/plot_caregiverresponse_coefs.svg)
+
 Effects of predictor variables on caregiver responses to the last time a child was sad or cried. Ordinal logistic regression model with elastic net penalties.
@@
 
+### Child anthropometrics
+
+xxxxxxxxxxDescribe camps brieflyxxxxxxxxxxxxxx - include dates and how we handled those.
+For cross-sectional analyses, we used the 2023 anthropometric values due to them being closer in time to the caretaker surveys.
+
+All anthropometric measurements were taken twice, in accordance with anthropometric standards [@lohman1988], and then averaged.
+Before being measured, participants were asked to remove their shoes and heavy articles of clothing.
+Height was measured to the nearest millimeter using an anthropometer, and weight was measured to the nearest 100 g with a digital scale (model name) that was used in a building with a concrete foundation.
+All participants were wearing lightweight clothing, so we made no adjustments for clothing weight.
+Grip strength was measured using a dynamometer (model name), with participants seated/standing?
+and told to xxxxxxxx.
+XXXX time was passed between the first and second measurements.
+
+Triceps and subscapular skinfold thicknesses were measured using skinfold calipers.
+The triceps measurement was taken at the midpoint of the upper arm.
+The subscapular measurement occurred immediately below the inferior angle of the scapula, being picked up diagonally at the natural fold line of the skin [@antóna].
+Flexed bicep circumference was measured perpendicular to the long axis of the upper arm.
+For this measure participants were asked to sit/stand?
+and position their upper arm perpendicular to their body with their elbow being held at a 90° angle, before flexing their clenched fist towards themselves as much as they could.
+
+We computed each child's height, weight, and BMI Z-scores with respect to the CDC child growth standards [@kuczmarski2002].
+
+For our independent variables of skinfold thickness, flexed bicep circumference, and grip strength on signaling behavior, we used the residuals from generalized additive models which modeled the effect of sex and age on each measurement using the mgcv package [@wood2011].
+The resulting variables represent how each child differed from the average child for their given sex and age.
+We also centered and scaled the skinfold thickness residuals before summing them to create a body fat index.
+
+### Old discussion parts
+
 The effect of age was smallest on child sadness.

### 2024-11-01 dictionaries.R

@@ -99,7 +99,7 @@ shortform_dict <-
     FoodSecurity = "Food insecurity",
     Neighborhood2 = "Camponado",
     UserLanguage = "Primary language",
-    NumberOfChildren = " Number of children",
+    NumberOfChildren = "Number of children",
     ImmigrateUtila = "Immigrant",
     OlderGirls = "Number of older girls",
     HouseQuality = "House quality",
@@ -146,7 +146,8 @@ shortform_dict <-
     #"TricepR" = "TricepR",
     #"SubscapR" = "SubscapR",
     "BodyFatPercentageR" = "BodyFatPercentageR",
-    "FlexedR" = "FlexedR"
+    "FlexedR" = "FlexedR",
+    RelativeMaternalInvestment2 = "Relative investment (w/in household)"
   )
 
 # shortform_dict[c(1,2,9,10,11,12,14,15,16)]

### 2024-11-01 regularize.R

@@ -116,7 +116,7 @@ ggsave("Figures/signal_effects_plot.pdf", signal_effects_plot, width = 12, heigh
 ggsave("Figures/signal_effects_plot.svg", signal_effects_plot, width = 12, height = 9)
 
 
-# Alloparenting special case
+# Alloparenting special case (SignalCost)
 
 e <- SignalVars |>
   dplyr::select(-householdID, -childHHid, -c(SadFreqN:SignalFreq), -ConflictFreqN, -AlloparentingXsex) |>
@@ -140,6 +140,39 @@ plot_predictions(m_alloparent, condition = c("AlloparentingFreqN", "Sex"), newda
 
 plot_alloparenting_cost$layers[[1]]$aes_params$linewidth <- 2
 plot_alloparenting_cost
+
+# Alloparenting special case (TantrumFreq)
+
+e2 <- SignalVars |>
+  dplyr::select(-householdID, -childHHid, -c(SadFreqN:CryFreqN), -c(SignalFreq:SignalCost), -ConflictFreqN, -AlloparentingXsex) |>
+  relocate(AlloparentingFreqN, .after = TantrumFreqN)
+e2[-1] <- scale(e2[-1])
+
+f2 <- paste(names(e2[-c(1,2)]), collapse = " + ")
+f2 <- as.formula(paste("TantrumFreqN ~ ", "AlloparentingFreqN*Sex +", f2))
+
+m_alloparent2 <-
+  poisson_reg(mixture = 1, penalty = signalparams$out$TantrumFreqN1$lambda.min) |>
+  set_engine("glmnet", family = quasipoisson) |>
+  fit(f2, data = e2)
+
+plot_alloparenting_cost_tantrum <-
+  plot_predictions(m_alloparent2, condition = c("AlloparentingFreqN", "Sex"), newdata = e2) +
+  scale_color_binary(labels = c("Female", "Male")) +
+  ylim(0, NA) +
+  labs(x = "Alloparenting Frequency (standardized)", y = "Tantrums (per month)") +
+  theme_minimal(15)
+
+plot_alloparenting_cost_tantrum$layers[[1]]$aes_params$linewidth <- 2
+plot_alloparenting_cost_tantrum
+
+plot_alloparentingXsex <- plot_alloparenting_cost / plot_alloparenting_cost_tantrum +
+  plot_layout(guides = "collect", axes = "collect") &
+  theme_bw(15)
+plot_alloparentingXsex
+
[…]

### 2024-11-01 Paper.qmd
Paper: minor edits throughout.
Regularize: fixed mistake that led to Family conflict not being included in the model of caregiver response to last instance of child signaling.

@@
 Compared to chimpanzee mothers, mothers in hunter-gatherer societies have shorter interbirth intervals and produce offspring that require many more years of parental care [@davison2021; @davison2022].
-Unlike chimpanzee offspring, human offspring in ancestral populations therefore often grew up with multiple older and/or younger siblings who were potential competitors for parental investment and yet who could also provide (or require) allocare [@helfrecht2016; @lawson2016; hagen2009].
+Unlike chimpanzee offspring, human offspring in ancestral populations therefore often grew up with multiple older and/or younger siblings who were potential competitors for parental investment and yet who could also provide (or require) allocare [@helfrecht2016; @lawson2016; @hagen2009].
 
@@
 Older children might prefer to spend their time investing in romantic or other social relationships, or learning valuable skills, whereas parents might prefer them to care for their siblings, acquire or prepare food, or complete other household chores [@lowe2003; @syme2023a].
-Conflicts of interest over investment in the family also appear to extend to when a child starts having their own children [@kushnick2023; @emlen1995], with parents potentially benefiting from delaying the reproductive careers of certain offspring in order to gain more allocare for their other children [@flinn1989; @emlen1992] or speeding up a child’s departure from the household so as to provide more parental investment to other children [@kushnick2023].
+Conflicts of interest over investment in the family also appear to extend to when a child starts having their own children [@kushnick2023; @emlen1995], with parents potentially benefiting from delaying the reproductive careers of certain offspring in order to gain more allocare for their other children [@flinn1989; @emlen1992] or speeding up a child’s departure from the household in order to provide more parental investment to other children [@kushnick2023].
 
@@
 
-Throughout human evolutionary history, individuals often found themselves in situations in which they received fitness-relevant benefits from others that could not easily be replaced [@apicella2012; @apicella2019; @hill2011; @tooby1996].
+Saved for now: Throughout human evolutionary history, individuals often found themselves in situations in which they received fitness-relevant benefits from others that could not easily be replaced [@apicella2012; @apicella2019; @hill2011; @tooby1996].
 In these instances, natural selection is expected to have favored those who provided support to others if 1) such help meant one could expect to receive fitness benefits from those helped in the future and 2) the costs of providing help were outweighed by the benefits to the helper on average [@andré2010; @hamilton1964; @tooby1996; @trivers1971].
@@
 Crying is the most studied way children signal need.
-Crying is a human universal [@gracanin2018] with few suspected non-signaling functions outside of the role vocal cries might play in the development of vocalization [@wells2003] and the functions tears have in terms of cleaning, protecting, and lubricating the eye [@murube2009; @sznycer].
-Hunger, pain, anger, fear, and separation are all common causes of crying [@vingerhoets2013; @wells2003], although tear production also often happens with joyful events [@sznycer; @vingerhoets2013; @zickfeld].
+Crying is a human universal [@gracanin2018] with few suspected non-signaling functions outside of the role vocal cries might play in the development of vocalization [@wells2003] and the functions tears have in terms of cleaning, protecting, and lubricating the eye [@murube2009; @sznycer2023].
+Hunger, pain, anger, fear, and separation are all common causes of crying [@vingerhoets2013; @wells2003], although tear production also often happens with joyful events [@sznycer2023; @vingerhoets2013; @zickfeld].
 
@@
 For this reason, crying among older children might be more similar to adult crying, which can increase support and empathy but sometimes also reduces reputation and increases stigmatization [@cornelius2001; @cretser1982; @hendriks2006; @jesser1989; @vingerhoets2016].
-Such costs could be a means by which crying can still be a credible signal of need if they dissuade potential fakers from making the signal [@sznycer] when there are moderate conflicts of interest.
+Such costs could be a means by which crying can still be a credible signal of need if they dissuade potential fakers from making the signal [@sznycer2023] when there are moderate conflicts of interest.
 
@@
 
-The likelihood of positive, neutral, and negative responses to the last time a child became sad or cried.
-A: Effect of punishment being implicated in the cause (0 = no punishment, 1 = punishment.) B: Effect of discomfort, pain, injury, or illness being implicated in the cause (0 = discomfort, pain, injury, or illness, 1 = discomfort, pain, injury, or illness.) Legend key: -1 = negative response, 0 = neutral response, 1 = positive response (all from child's perspective).
+Effects of selected predictors of the likelihood of positive, neutral, and negative responses to the last time a child became sad or cried.
+1 = evidence of the cause in the caregiver report; 0 = no evidence.
+The endpoints of the lines were dodged slightly for greater visibility.
+Dot size represents xxxxxxx.
+See @sfig-pot-caregiver-punish for the effect of the cause involving punishment.
 :::
@@
 About half of children were reported to cry or throw tantrums more than once a month and slightly less than 40% were reported to engage in them more than once a week.
-Although similarly common overall, tantrums more common than crying among older children and might be a under-appreciated way by which children might signal for increased support based on these findings and a longitudinal study of temper tantrums and otherchild behaviors which found 50% of boys 30% of girls engaged in temper tantrums at age 9 among a representative, non-clincial sample of 252 children in Berkeley, California [@@macfarlane1962].
+Although similarly common overall, tantrums more common than crying among older children and might be an under-appreciated way by which children signal for increased support based on these findings and a longitudinal study of temper tantrums and other child behaviors which found 50% of boys 30% of girls engaged in temper tantrums at age 9 among a representative, non-clincial sample of 252 children in Berkeley, California [@@macfarlane1962].
 
@@
 In times when an older child's valuation of potential increased support is small, such costs may be prohibitive.
-However, in times when the child places more value on potential support, those same costs could allow for a child to credibly signal need if they are too high to be worth it for deceptive signaling.
+However, in times when a child places more value on potential support, those same costs could allow for a child to credibly signal need if they are too high to be worth it for deceptive signaling but still made up for by the benefits for credible signalers.
 
@@
 If this is the case, the positive correlation between signaling effort and conflict frequency may represent children's attempts at credibly signaling their value of additional investment or the costliness of punishment [@syme2019] to parents who would have good reason to doubt the trustworthiness of simple verbal requests (or the trustworthiness of facial sadness and crying for those resorting to temper tantrums).
-However, it is also likely that parents find some forms of child signaling aversive, and therefore see it as conflict in and of itself.
+It is also likely that parents find some forms of child signaling aversive, and therefore see it as conflict in and of itself.
 
 We found evidence of an interaction between how often a child cared for their siblings and the sex of the child in our lasso models examining signal frequencies and cost, whereby there was a positive relationship between the frequency and summed costliness of child signals of need and alloparenting effort for girls but no relationship or a slightly negative one for boys.
-In many societies, girls do more alloparenting than boys [@helfrecht2016; @weisner1977; @barry1971; @turke1988], something which matches findings from focus groups of Utilian mothers \[add info from an analysis here?\].
+In many societies, girls do more alloparenting than boys [@helfrecht2016; @weisner1977; @barry1971; @turke1988], something which matches findings from focus groups of Utilian mothers.
 If this difference was the outcome of caretakers pressuring girls to do more or resulted in the expectation of more alloparenting, the increase in signaling with alloparenting effort seen among girls may be the result of a strategy to signal the costliness of their efforts in order to invest more in relationships and status seeking behaviors outside the household.
@@
 Ideally, we could control for signaling behavior to test this.
-However, the bidirectional relationship between child signaling and conflict with caretakers prevents this (explain better).
+However, the bidirectional relationship between child signaling and conflict with caretakers prevents this.
 It is also possible that our hypothesis about conflict over investment in the household is incorrect.
@@
 One reason for this might be that costs of sadness, while minor short-term, can be significant long-term.
-In line with this reasoning, depression, which often involves an exteneded period of a lack of interest in one's normal activities and decreased investment in oneself and one's cooperative relationships, has been suggested to capable of credibly signal need in times of substantial conflicts of interest [@watson2002; @hagen2003].
+These costs are clearly seen in depression, which often involves an extended period of a lack of interest in one's normal activities and decreased investment in oneself and one's cooperative relationships and has been suggested to capable of credibly signal need in times of substantial conflicts of interest [@watson2002; @hagen2003].
 
@@
 As cues are not created for the purpose of communication, they are completely reliable if perceived and interpreted correctly.
-Therefore, if some cues are hard to impossible to fake, they may be more reliable than most signals.
+Therefore, if some cues are hard to or are impossible to fake, they may be more reliable than most signals.
 That said, it is likely that some symptoms of sadness are simultaneously the result of both signals and cues, something which may have been favored by natural selection if the threat of deception meant cue-level symptoms were insufficient in eliciting support without upregulation, something which has been suggested to be the case with various illness symptoms [@tiokhin2016; @steinkopf2015].
@@
 
-When the cause involved punishment, however, parents were more likely to respond negatively and less likely to respond positively.
+When the cause involved the child having made a transgression, punishment, or the loss of privleges or an item, however, parents were more likely to respond negatively and less likely to respond positively.
 In these circumstances, children may be signaling about the severity of a punishment [@syme2019].
@@
 
-Although not a significant relationship, the negative effect of the family conflict being a reported cause of signaling on the likelihood of a positive response may reflect tradeoffs within the family, whereby investment in one child cannot go to another.
+Parents were also more likely to respond negatively and less likely to respond positively to children when the cause involved family conflict.
+This might reflect tradeoffs within the family, whereby investment in one child cannot go to another.
 Caretakers may have also responded more negatively when the cause involved family conflict, if it involved a child imposing costs on other family members.
@@
 Despite the differences in wealth, job certainty, discrimination, and living conditions [@davis2014; @garcia2017; @garcia2018; @stonich2000] between recent immigrants from the mainland and native Utilians, primary language, immigration status, and residence in Camponado did not predict child signaling in our lasso regressions.
-However, neighborhood quality, which was strongly negaltively correlated with residence in Camponado, (check others) was preserved in the lasso regressions predicting every signaling measure but sadness and was among the variables with the largest negative coefficients.
+However, neighborhood quality, which was strongly negatively correlated with residence in Camponado, (check others) was preserved in the lasso regressions predicting every signaling measure but sadness and was among the variables with the largest negative coefficients.
 Monthly income was also negatively associated with crying frequency,~~ tantrum frequency (too small of a coefficient to mention here?),~~ summed signal frequency, and signal cost albeit with smaller coefficients.
@@
 
+::: {#sfig-pot-caregiver-punish}
+![](Figures/plot_caregiver_punish.svg)
+
+The likelihood of positive, neutral, and negative responses to the last time a child became sad or cried when punishment was implicated in the cause (0 = no punishment, 1 = punishment.)
+:::
+
 ## Coefficient plots

### 2024-11-01 dictionaries.R

@@ -147,7 +147,11 @@ shortform_dict <-
     #"SubscapR" = "SubscapR",
     "BodyFatPercentageR" = "BodyFatPercentageR",
     "FlexedR" = "FlexedR",
-    RelativeMaternalInvestment2 = "Relative investment (w/in household)"
+    RelativeMaternalInvestment2 = "Relative investment (w/in household)",
+    Family = "Family conflict",
+    TransgressionMade = "Transgression made",
+    LossOfPrivlegesOrItem = "Loss of privleges or item"
+
   )
 
 # shortform_dict[c(1,2,9,10,11,12,14,15,16)]

### 2024-11-01 regularize.R

@@ -506,9 +506,8 @@ causematrix <-
   ) |>
   dplyr::select(householdID, childHHid, Family:StatusConcerns) |>
   na.omit()
-
-cause_count <- map_int(causematrix[-c(1:5)], \(x) sum(x, na.rm = T))
-cause_vars <- names(cause_count[cause_count > 4])
+cause_count <- map_int(causematrix[-c(1:2)], \(x) sum(x, na.rm = T))
+cause_vars <- names(cause_count[cause_count > 5])
 
 SignalVars3 <-
   left_join(SignalVars, d2[c("householdID", "childHHid", "CaregiverResponse")]) |>
@@ -516,7 +515,7 @@ SignalVars3 <-
   relocate(CaregiverResponse, .after = childHHid) |>
   mutate(
     across(-c(1:3), as.numeric),
-    across(SadFreqN:StatusConcerns, \(x) c(scale(x)))
+    across(SadFreqN:SeparationAttentionSeeking, \(x) c(scale(x))) # changed StatusConcerns to SeparationAttentionSeeking
     ) |>
   na.omit()
 
@@ -538,12 +537,18 @@ plot_caregiverresponse_coefs
 ggsave("Figures/plot_caregiverresponse_coefs.pdf", plot_caregiverresponse_coefs, width = 12, height = 12)
 ggsave("Figures/plot_caregiverresponse_coefs.svg", plot_caregiverresponse_coefs, width = 12, height = 12)
 
+plot_caregiver_familyconflict <- ordinal_plot(out$fit, Family, data = SignalVars3, title = 'Caregiver response')
+plot_caregiver_trangsression <- ordinal_plot(out$fit, TransgressionMade, data = SignalVars3, title = 'Caregiver response')
+plot_caregiver_loss <- ordinal_plot(out$fit, LossOfPrivlegesOrItem, data = SignalVars3, title = 'Caregiver response')
 plot_caregiver_pain <- ordinal_plot(out$fit, DiscomfortPainInjuryIllness, data = SignalVars3, title = 'Caregiver response')
-plot_caregiver_punish <- ordinal_plot(out$fit, Punishment, data = SignalVars3, title = 'Caregiver response')
-plot_caregiver_response_combined <- plot_caregiver_punish / plot_caregiver_pain + ggtitle("") + plot_layout(guides = 'collect')
 
[…]

### 2024-11-05 Paper.qmd
Edited tantrum sections in paper (background and discussion).
Made changes to accomodate the name differences in the new causes dataframe.

@@
 
-Temper tantrums may also be adaptive strategies by which children can elicit more support from parents.
-They are common in children from 18 months to 4 years old [@potegal2003] yet are understudied among older children [@koch2003] outside of studies examining disruptive mood regulation disorder [for one exception see @book].
-Temper tantrums vary in their severity, ranging from crying, shrieking, and screaming to violent bodily movements, aggression, and self-harm [@potegal2003], with similar behaviors having also been reported in other young animals, often surrounding feeding or weaning conflict [@altmann1980; @goodall1986; @schaller1964].
-Tantrums have been reported in many societies with different subsistence systems, political organization, and norms surrounding anger and parenting [@albino1956; @banisalameh2021; @fouts2005; @henderson1966; @hilger1957; @konner1974; @ward1970], suggesting they may be a human universal.
-As with the other proposed signals of need, they are capable of eliciting a wide range of responses, including pro-child responses, neutral responses (e.g., parents ignoring the child) [@hilger1957; @konner1974], and negative responses like abuse [@schmitt1987].
+Temper tantrums may also be the result of adaptive strategies by which children can elicit more support from parents [@koch2003].
+They are common in children from 18 months to 4 years old [@potegal2003] but also exist among older children, albeit with decreased frequency [@macfarlane1962; @barbarin1999; @ward2013; @osterman2010; @goodenough1931; @bhatia1990].
+Temper tantrums vary in their severity and include behaviors ranging from crying, shrieking, and screaming to violent bodily movements, aggression, and self-harm [@potegal2003].
+Similar behaviors are also seen in other young animals, often surrounding feeding or weaning conflict [@altmann1980; @goodall1986; @schaller1964].
+Tantrums have been reported in many societies with different forms of subsistence systems, political organization, and norms surrounding anger and parenting [@albino1956; @banisalameh2021; @fouts2005; @henderson1966; @hilger1957; @konner1974; @ward1970; @johnson2003; @bhatia1990; @myers1988; @dubois @konner1972], suggesting they might be a human universal [@potegal2019].
+As with the other proposed signals of need, they are capable of eliciting a wide range of responses, including pro-child responses, neutral responses (e.g., parents ignoring the child) [@hilger1957; @konner1974; @johnson2003], and negative responses like abuse [@schmitt1987].
 
@@
 
-[no longer needed] XXXXXXXXCHIld RecruitmentXXXXXXXXXXX
+\[no longer needed\] XXXXXXXXCHIld RecruitmentXXXXXXXXXXX
 
@@
 |  |  |
-|-----------------------|-------------------------------------------------|
+|-------------------------|-----------------------------------------------|
 | Variable | Survey question wording |
@@
 ::: {#fig-mst}
-![](Figures/plot_mst_uw.svg)
+![](Figures/plot_mst_1e6_cost_freq.edited.png)
 
@@
 
-Cap to add.
+Frequencies of caregiver response to the last instance of child signaling by the cause of the signaling.
+Categories of causes were based on cluster analysis of more fine-grained cause variables.
+See sixxxxxx for cluster and sitabxxxxxxx for the frequencies of causes and caregiver responses.
 :::
@@
 Caregivers reported a wide range of child signaling frequencies.
-About half of children were reported to cry or throw tantrums more than once a month and slightly less than 40% were reported to engage in them more than once a week.
-Although similarly common overall, tantrums more common than crying among older children and might be an under-appreciated way by which children signal for increased support based on these findings and a longitudinal study of temper tantrums and other child behaviors which found 50% of boys 30% of girls engaged in temper tantrums at age 9 among a representative, non-clincial sample of 252 children in Berkeley, California [@@macfarlane1962].
+About half of children were reported to cry or throw tantrums more than once a month and almost 40% were reported to do so more than once a week [@fig-barplot_SignalFreq].
+Frequencies of sadness, crying, and tantrums were positively correlated with each other and with ~~family conflict~~ the frequency of conflict with the primary caretaker [@fig-signal_corrplot].
+They also loaded in the same direction on PC1 as indices of adversity, such as food insecurity, and in the opposite direction of indices of wealth and environmental quality, such as possessions, neighborhood quality, and income [@fig-PCA].
+Signaling decreased with child age, and increased with ~~family~~ conflict with their primary caretaker, which were the two variables most closely linked with signaling in the MST [@fig-mst].
+In our lasso regression models (which excluded family conflict due to the likely bidirectional relationship between signaling and perceived conflict), child age and high neighborhood quality were consistently negatively associated with signaling frequency, and household income, caregiver education, and number of adults helping with childcare generally were too; medical problems were a consistent, albeit small positive predictor [@fig-lasso; @fig-signal-effects].
 
-As seen in Western cultures [@zeifman2013; @potegal2003; @koch2003; @macfarlane1962], the frequencies of crying and temper tantrums decreased with age, as did the frequency of child sadness, summed signal frequency, and signal cost.
-One possible reason for this is that as children mature, they are more capable of fulfilling their own desires, more able to ask others for what they want, and less vulnerable to threats.
+These results are consistent with the hypothesis that sadness, crying, and tantrums are signals of need, and that children's need is higher in adverse environments and at younger ages.
+One possible reason for the negative association of signaling with age, which is also seen in more industrialized settings [@zeifman2013; @potegal2003; @koch2003; @macfarlane1962], is that as children mature, they are more capable of fulfilling their own desires, more able to ask others for what they want [@osterman2010], and less vulnerable to threats.
 The effects of age on signaling are also likely due to how age influences the costliness of signaling need.
-For example, while crying may be costly among the youngest children through its energetic costs, these costs are likely insignificant among older children, who might instead face costs in the form of reputational harm from crying in some contexts but not others.
-In times when an older child's valuation of potential increased support is small, such costs may be prohibitive.
+For example, while the energetic cost of crying might be costly for the youngest children, it is likely insignificant for older children.
+Older children might instead need to weigh the reputational harm from crying against the potential increase in social support.
+
+Consistent with signaling models that emphasize the role of conflicts of interest in favoring the use of costlier signals, caregiver perceptions of conflict and child signaling frequency were correlated [@fig-signal_corrplot], and family conflict was the most common cause of a child's recent signaling episode [@fig-effect_last_signal_all_responses].
+Although conflicts of interest do not always lead to overt behavioral conflict [@altmann1980; @bateson1994], it is likely many, if not most, conflicts are ultimately caused by conflicts of interest.
+Conflicts, in turn, were negatively associated with child age and higher neigborhood quality, and positively associated with medical problems and (oddly) caregiver education [@fig-conflict-effects].
+Finally, more frequent sadness was positively associated with perceived higher need relative to other children [@fig-effect_relativeneed], and higher perceived need was associated with higher relative investment [@fig-plot_relative_investment].
+Interestingly, older children were percieved to be needier than younger ones.
+These results generally support the hypothesis that the signals we investigated are costly, credible signals of need.
+
+\<!- If this is the case, the positive correlation between signaling effort and conflict frequency may represent children's attempts at credibly signaling their value of additional investment or the costliness of punishment [@syme2019] to parents who would have good reason to doubt the trustworthiness of simple verbal requests (or the trustworthiness of facial sadness and crying for those resorting to temper tantrums).
+It is also likely that parents find some forms of child signaling aversive, and therefore see it as conflict in and of itself.
+-\>
+
+\<!- In times when an older child's valuation of potential increased support is small, such costs may be prohibitive.
 However, in times when a child places more value on potential support, those same costs could allow for a child to credibly signal need if they are too high to be worth it for deceptive signaling but still made up for by the benefits for credible signalers.
+-\>
 
-Consistent with signaling models which emphasize the role of conflicts of interest in favoring the use of costlier signals, conflict and signaling behavior were highly correlated.
-Although conflicts of interest do not always lead to behavioral conflict [@altmann1980; @bateson1994], it is likely many, if not most, conflicts are ultimately caused by conflicts of interest, potentially making it a useful proxy for the underlying conflicts of interest that may increase the benefits of deception for children and the benefits of greater skepticism for parents\[\[5\]\].
-If this is the case, the positive correlation between signaling effort and conflict frequency may represent children's attempts at credibly signaling their value of additional investment or the costliness of punishment [@syme2019] to parents who would have good reason to doubt the trustworthiness of simple verbal requests (or the trustworthiness of facial sadness and crying for those resorting to temper tantrums).
-It is also likely that parents find some forms of child signaling aversive, and therefore see it as conflict in and of itself.
+Caregiver responses to signals were generally positive or neutral, although less so for situations involving family conflict or transgressions [@fig-effect_last_signal_all_responses; @fig-effect_last_signal_response].
 
-We found evidence of an interaction between how often a child cared for their siblings and the sex of the child in our lasso models examining signal frequencies and cost, whereby there was a positive relationship between the frequency and summed costliness of child signals of need and alloparenting effort for girls but no relationship or a slightly negative one for boys.
+Finally, based on previous research showing that family demands, such as caring for younger siblings, was an important source of parent-offspring conflict [@syme2019], we investigated the relationship between the frequency of alloparenting by children and their frequency of signaling.
+We found that higher frequencies of alloparenting by girls was associated with a higher frequency of tantrums and overall signaling costs, but this relationship was slightly negative for boys [@fig-alloparentingXsex].
 In many societies, girls do more alloparenting than boys [@helfrecht2016; @weisner1977; @barry1971; @turke1988], something which matches findings from focus groups of Utilian mothers.
 If this difference was the outcome of caretakers pressuring girls to do more or resulted in the expectation of more alloparenting, the increase in signaling with alloparenting effort seen among girls may be the result of a strategy to signal the costliness of their efforts in order to invest more in relationships and status seeking behaviors outside the household.
+
+--- Hagen stopped here ---
+
+\<!- We found evidence of an interaction between how often a child cared for their siblings and the sex of the child in our lasso models examining signal frequencies and cost, whereby there was a positive relationship between the frequency and summed costliness of child signals of need and alloparenting effort for girls but no relationship or a slightly negative one for boys.
+-\>
 
@@
 They may also react with scolding or further punishment.
-That said, our model predicts positive responses in `r signif(updated_out$estimate[3] * 100, 2)`% of cases (when all other predictors are held at their mean values), suggesting that children are successful in negotiating better outcomes at least some of the time.
+That said, our model predicts positive responses in `r #signif(updated_out$estimate[3] * 100, 2)`% of cases (when all other predictors are held at their mean values), suggesting that children are successful in negotiating better outcomes at least some of the time.
 
@@
 
-\[delete this paragraph?\] 
-We found a negative association between caretakers’ education level (years in school) and many of our child signaling measures despite our models controlling for household income and partner status.
+\[delete this paragraph?\] We found a negative association between caretakers’ education level (years in school) and many of our child signaling measures despite our models controlling for household income and partner status.
 One possible reason for this may be that more educated caretakers see their children less during the day due to greater ability to afford childcare or due to them having jobs that are less compatible with child care (test model of social support predicted by education).
@@
 However, neighborhood quality, which was strongly negatively correlated with residence in Camponado, (check others) was preserved in the lasso regressions predicting every signaling measure but sadness and was among the variables with the largest negative coefficients.
-Monthly income was also negatively associated with crying frequency,~~ tantrum frequency (too small of a coefficient to mention here?),~~ summed signal frequency, and signal cost albeit with smaller coefficients.
+Monthly income was also negatively associated with crying frequency,\~\~ tantrum frequency (too small of a coefficient to mention here?),\~\~ summed signal frequency, and signal cost albeit with smaller coefficients.
 
@@
 ~~In addition to overall effort in eliciting need, resource availability may have also affect how children signal for more investment based on the finding that birds in stable and unusually favorable environments are more likely to compete with siblings through signals thought to reveal information about need, while those in relatively unpredictable and unusually poor environments have been found to compete more through signals thought to reveal information about quality [@caro2016a].~~
+
+(temporarilly moved here) Tantrums were present among older children and might be an under-appreciated way by which children signal for increased support.
+Most studies of tantrums involve children under 4 years old or older children with disruptive mood dysregulation disorder, likely due to their greater prevalence in these populations (cite).
+However, multiple studies which included older children from non-clinical populations have found similar evidence of tantrums [@macfarlane1962; @barbarin1999; @ward2013; @osterman2010; @goodenough1931; @bhatia1990], as have ethnographic studies [@ward2013; @lessa1950; @dennis1940].
 
@@
 |  |  |
-|------------------------|------------------------------------------------|
+|-------------------------|----------------------------------------------|
 | Variable | Survey question wording |
@@
 |  |  |
-|---------------------|---------------------------------------------------|
+|-----------------------|------------------------------------------------|
 | Variable | Survey question wording |
@@
 |  |  |
-|------------------------|-----------------------------------------------|
+|--------------------------|----------------------------------------------|
 | Variable | Survey question wording |
@@
 |  |  |
-|----------------------|-------------------------------------------------|
+|------------------------|------------------------------------------------|
 | Variable | Survey question wording |
@@
 ::: {#sfig-plot-caregiverresponse-coefs}
-
 ![](Figures/plot_caregiverresponse_coefs.svg)

### 2024-11-05 analysisforpaper.R

@@ -240,13 +240,13 @@ ggplot(caregivers2, aes(y=Neighborhood, fill=ImmigrateUtila)) +
 
 # Correlation matrix of predictor and outcome variables
 
-maincormat <- d2[c("IllnessSusceptibilityMean", "EducationLevelYears", "OtherChildrenHH", "LogIncome", "ConflictFreqN", "number_adults", "PartnerStatus", "AlloparentingFreqN", "Sex", "ChildAge", "RelativeNeed3", "RelativeMaternalInvestment2", "SadFreqN", "CryFreqN", "TantrumFreqN", "SignalFreq", "SignalFreqMax", "SignalCost", "Family2")] |>
+maincormat <- d2[c("IllnessSusceptibilityMean", "EducationLevelYears", "OtherChildrenHH", "LogIncome", "ConflictFreqN", "number_adults", "PartnerStatus", "AlloparentingFreqN", "Sex", "ChildAge", "RelativeNeed3", "RelativeMaternalInvestment2", "SadFreqN", "CryFreqN", "TantrumFreqN", "SignalFreq", "SignalFreqMax", "SignalCost", "ConflictFamily2")] |>
   mutate(
     PartnerStatus = as.numeric(PartnerStatus),
     Sex = as.numeric(Sex),
     RelativeNeed3 = as.numeric(RelativeNeed3),
     RelativeMaternalInvestment2 = as.numeric(RelativeMaternalInvestment2),
-    Family2 = as.numeric(Family2),
+    ConflictFamily2 = as.numeric(ConflictFamily2),
   ) |>
   cor( use = "pairwise.complete.obs")
 
@@ -273,7 +273,7 @@ glmmTMBmodels <- tibble(
 
 polrNeedFormulas <- str_glue("RelativeNeed3 ~ {signals} + {std_predictors} + OtherChildrenHH + IllnessSusceptibilityMean")
 polrInvestFormula <- paste0("RelativeMaternalInvestment2 ~ ", std_predictors, "+ OtherChildrenHH + IllnessSusceptibilityMean + RelativeNeed3")
-polrResponseFormula <- paste0("CaregiverResponse ~ ", std_predictors, "+ OtherChildrenHH + IllnessSusceptibilityMean + UnwantedTask + Punishment2 + Family2 + DiscomfortPainInjuryIllness")
[…]

### 2024-11-05 dataprep.R

@@ -258,7 +258,7 @@ d <-
                           EducationLevel, EducationLevelYears, AdultsMoney, AdultsHousework, AdultsChildcare, AdultsNoChildcare, CaregiverAge,
                           number_children2, number_adults, UserLanguage, CurrentJob, contains("CaregiverMarital"), NeighborhoodQuality, HouseQuality,
                           contains("SocialSupport"), contains("FoodSecurity"), contains("HomeInstability_"), contains("Reality_")), by = "householdID") |>
-  left_join(dplyr::select(causes, -householdID, -childHHid, -ChildID, -Sad1), by = "uniqueID")
+  left_join(dplyr::select(causes, -householdID, -childHHid, -text), by = "uniqueID")
 
 # prepare household variables for analyses
 d2 <-
@@ -282,12 +282,12 @@ d2 <-
     NeighborhoodF = as.factor(Neighborhood),
     OtherChildrenHH = number_children2 - 1,
     Punishment2 = ifelse(Punishment == "Maybe", 0, Punishment),
-    Family2 = ifelse(Family == "Maybe", 0, Family),
-    OutsideFamily2 = ifelse(OutsideFamily == "Maybe", 0, OutsideFamily),
-    WantNeed2 = ifelse(WantNeed == "Want" | WantNeed == "Need", WantNeed, NA),
-    WantNeedBinary = case_when(
-      WantNeed2 == "Want" ~ 0,
-      WantNeed2 == "Need" ~ 1),
+    ConflictFamily2 = ifelse(ConflictFamily == "Maybe", 0, ConflictFamily),
+    ConflictOutsideFamily2 = ifelse(ConflictOutsideFamily == "Maybe", 0, ConflictOutsideFamily),
+    # WantNeed2 = ifelse(WantNeed == "Want" | WantNeed == "Need", WantNeed, NA),
+    # WantNeedBinary = case_when(
+    #   WantNeed2 == "Want" ~ 0,
+    #   WantNeed2 == "Need" ~ 1),
     DeathORIllnessInjuryHarmInOthers2 = ifelse(DeathORIllnessInjuryHarmInOthers == "Maybe", 0, DeathORIllnessInjuryHarmInOthers),
     HouseholdAdversity2 = ifelse(HouseholdAdversity == "Maybe", 0, HouseholdAdversity),
     ExplicitInvestmentDesired2 = ifelse(ExplicitInvestmentDesired == "Maybe", 0, ExplicitInvestmentDesired),

### 2024-11-05 dictionaries.R

@@ -148,7 +148,7 @@ shortform_dict <-
     "BodyFatPercentageR" = "BodyFatPercentageR",
     "FlexedR" = "FlexedR",
     RelativeMaternalInvestment2 = "Relative investment (w/in household)",
-    Family = "Family conflict",
+    ConflictFamily = "Family conflict",
     TransgressionMade = "Transgression made",
     LossOfPrivlegesOrItem = "Loss of privleges or item"
 

### 2024-11-05 regularize.R

@@ -501,10 +501,10 @@ ggsave("Figures/plot_invest_need.svg", plot_invest_need, width = 12, height = 12
 causematrix <-
   causes |>
   mutate(
-    across(Family:StatusConcerns, \(x) ifelse(x == "Maybe" | x == 'maybe', "0", x)),
-    across(Family:StatusConcerns, as.numeric)
+    across(ConflictFamily:StatusConcerns, \(x) ifelse(x == "Maybe" | x == 'maybe', "0", x)),
+    across(ConflictFamily:StatusConcerns, as.numeric)
   ) |>
-  dplyr::select(householdID, childHHid, Family:StatusConcerns) |>
+  dplyr::select(householdID, childHHid, ConflictFamily:StatusConcerns) |>
   na.omit()
 cause_count <- map_int(causematrix[-c(1:2)], \(x) sum(x, na.rm = T))
 cause_vars <- names(cause_count[cause_count > 5])
@@ -537,7 +537,7 @@ plot_caregiverresponse_coefs
 ggsave("Figures/plot_caregiverresponse_coefs.pdf", plot_caregiverresponse_coefs, width = 12, height = 12)
 ggsave("Figures/plot_caregiverresponse_coefs.svg", plot_caregiverresponse_coefs, width = 12, height = 12)
 
-plot_caregiver_familyconflict <- ordinal_plot(out$fit, Family, data = SignalVars3, title = 'Caregiver response')
+plot_caregiver_familyconflict <- ordinal_plot(out$fit, ConflictFamily, data = SignalVars3, title = 'Caregiver response')
 plot_caregiver_trangsression <- ordinal_plot(out$fit, TransgressionMade, data = SignalVars3, title = 'Caregiver response')
 plot_caregiver_loss <- ordinal_plot(out$fit, LossOfPrivlegesOrItem, data = SignalVars3, title = 'Caregiver response')
 plot_caregiver_pain <- ordinal_plot(out$fit, DiscomfortPainInjuryIllness, data = SignalVars3, title = 'Caregiver response')

### 2024-11-05 Paper.qmd
Commiting changes made in Zed (dicussion and added SI plots).

@@
     orcid: 0000-0002-7552-6942
+    corresponding: false
+    affiliation:
+      - name: Washington State University
+        department: Anthropology
+  - name: Jessica K. Hlay
+    orcid: 0000-0002-6139-5144
+    corresponding: false
+    affiliation:
+      - name: Boston University
+        department: Anthropology
+  - name: Izabel Rodriguez
+  - name: Kristen L. Syme
+    orcid: 0000-0002-5601-8897
+    corresponding: false
+    affiliation:
+      - name: Leiden University
+        department: Institute of Security and Global Affairs
+  - name: Aaron D. Blackwell
+    orcid: 000-0002-5871-9865
     corresponding: false
@@
       - name: Washington State University
+        department: Anthropology
+  - name: Carolyn R. Hodges-Simeon
+    orcid: 0000-0002-0691-9080
+    corresponding: false
+    affiliation:
+      - name: Boston University
         department: Anthropology
@@
 Categories of causes were based on cluster analysis of more fine-grained cause variables.
-See sixxxxxx for cluster and sitabxxxxxxx for the frequencies of causes and caregiver responses.
+See @sfig-cause-clustering for cluster dendrogram and @sfig-sumtab-causes for the frequencies of causes and caregiver responses.
 :::
@@
 
-Effects of selected predictors of the likelihood of positive, neutral, and negative responses to the last time a child became sad or cried.
+Effects of selected predictors of the likelihood of positive, neutral, and negative responses to the last time a child became sad or cried when holding continuous variables at their mean and factor variables at their mode.
 1 = evidence of the cause in the caregiver report; 0 = no evidence.
-The endpoints of the lines were dodged slightly for greater visibility.
-Dot size represents xxxxxxx.
-See @sfig-pot-caregiver-punish for the effect of the cause involving punishment.
+The endpoints of the lines are dodged slightly for greater visibility.
+Dot size represents model predictions for each participant given their values of predictor variables.
+See @sfig-plot-caregiver-punish for the effect of the cause involving punishment.
 :::
@@
 Frequencies of sadness, crying, and tantrums were positively correlated with each other and with ~~family conflict~~ the frequency of conflict with the primary caretaker [@fig-signal_corrplot].
-They also loaded in the same direction on PC1 as indices of adversity, such as food insecurity, and in the opposite direction of indices of wealth and environmental quality, such as possessions, neighborhood quality, and income [@fig-PCA].
+They also loaded on PC1 in the same direction as indices of adversity, such as food insecurity, and in the opposite direction of indices of wealth and environmental quality, such as possessions, neighborhood quality, and income [@fig-PCA].
 Signaling decreased with child age, and increased with ~~family~~ conflict with their primary caretaker, which were the two variables most closely linked with signaling in the MST [@fig-mst].
-In our lasso regression models (which excluded family conflict due to the likely bidirectional relationship between signaling and perceived conflict), child age and high neighborhood quality were consistently negatively associated with signaling frequency, and household income, caregiver education, and number of adults helping with childcare generally were too; medical problems were a consistent, albeit small positive predictor [@fig-lasso; @fig-signal-effects].
+In our lasso regression models (which excluded family conflict due to the likely bidirectional relationship between signaling and perceived conflict), child age and high neighborhood quality were consistently negatively associated with signaling frequency, as were household income, caregiver education, and number of adults helping with childcare.
+Medical problems were a consistent, albeit small positive predictor [@fig-lasso; @fig-signal-effects].
 
 These results are consistent with the hypothesis that sadness, crying, and tantrums are signals of need, and that children's need is higher in adverse environments and at younger ages.
-One possible reason for the negative association of signaling with age, which is also seen in more industrialized settings [@zeifman2013; @potegal2003; @koch2003; @macfarlane1962], is that as children mature, they are more capable of fulfilling their own desires, more able to ask others for what they want [@osterman2010], and less vulnerable to threats.
+One possible reason for the negative association of signaling with age, which is also seen in more industrialized settings [@zeifman2013; @potegal2003; @koch2003; @macfarlane1962], is that as children mature, they are more capable of fulfilling their own needs, more able to ask others for what they need [@osterman2010], and less vulnerable to threats.
 The effects of age on signaling are also likely due to how age influences the costliness of signaling need.
@@
 Conflicts, in turn, were negatively associated with child age and higher neigborhood quality, and positively associated with medical problems and (oddly) caregiver education [@fig-conflict-effects].
-Finally, more frequent sadness was positively associated with perceived higher need relative to other children [@fig-effect_relativeneed], and higher perceived need was associated with higher relative investment [@fig-plot_relative_investment].
-Interestingly, older children were percieved to be needier than younger ones.
-These results generally support the hypothesis that the signals we investigated are costly, credible signals of need.
 
-\<!- If this is the case, the positive correlation between signaling effort and conflict frequency may represent children's attempts at credibly signaling their value of additional investment or the costliness of punishment [@syme2019] to parents who would have good reason to doubt the trustworthiness of simple verbal requests (or the trustworthiness of facial sadness and crying for those resorting to temper tantrums).
+Caregiver responses to signals were generally positive or neutral, more so for situations involving injury and illness, and less so for situations involving family conflict, loss of privileges, and transgressions [@fig-effect_last_signal_all_responses; @fig-effect_last_signal_response].
+These patterns probably reflect a combination of confluences of interest vs.
+conflicts of interest between child and caretaker, and less caretaker uncertainty about need in cases of injury and illness.
+That said, our model predicts positive responses in `r #signif(updated_out$estimate[3] * 100, 2)`% of cases involving trangressions (when all other predictors are held at their mean values), suggesting that children are successful in negotiating better outcomes at least some of the time.
+
+More frequent sadness was positively associated with perceived higher need relative to other children [@fig-effect_relativeneed].
+Although crying and tantrums were not associated with higher perceived need, this was likely due to their high correlation with sadness and the propensity of lasso regression to retain only one of a set of correlated predictors [@zou2005].
+Higher perceived need was associated with higher relative investment [@fig-plot_relative_investment].
+Signaling behavior did not predict differences in relative caregiver investment beyond its effect on perceived need [@sfig-plot-invest-coefs], similar to an experimental vignette study that found that the likelihood of costly support was largely mediated by the effect of costly signaling on belief in need [@gaffney2022].
+Altogether, the foregoing results generally support the hypothesis that the signals we investigated are costly, credible signals of need.
+
+Interestingly, older children were perceived to be needier than younger ones [@@fig-effect_relativeneed].
+One likely reason for this might be that caretakers perceive the types of investment older children seek as more costly.
+In line with this reasoning, multiple caretakers mentioned that older kids need more money and advice and get into trouble more often.
+It might also be that caretakers see investing in older children as relatively costly as they seek to withdraw investment in order to invest in younger children.
+
+<!- If this is the case, the positive correlation between signaling effort and conflict frequency may represent children's attempts at credibly signaling their value of additional investment or the costliness of punishment [@syme2019] to parents who would have good reason to doubt the trustworthiness of simple verbal requests (or the trustworthiness of facial sadness and crying for those resorting to temper tantrums).
 It is also likely that parents find some forms of child signaling aversive, and therefore see it as conflict in and of itself.
--\>
+->
 
@@
 
-Caregiver responses to signals were generally positive or neutral, although less so for situations involving family conflict or transgressions [@fig-effect_last_signal_all_responses; @fig-effect_last_signal_response].
-
 Finally, based on previous research showing that family demands, such as caring for younger siblings, was an important source of parent-offspring conflict [@syme2019], we investigated the relationship between the frequency of alloparenting by children and their frequency of signaling.
 We found that higher frequencies of alloparenting by girls was associated with a higher frequency of tantrums and overall signaling costs, but this relationship was slightly negative for boys [@fig-alloparentingXsex].
-In many societies, girls do more alloparenting than boys [@helfrecht2016; @weisner1977; @barry1971; @turke1988], something which matches findings from focus groups of Utilian mothers.
-If this difference was the outcome of caretakers pressuring girls to do more or resulted in the expectation of more alloparenting, the increase in signaling with alloparenting effort seen among girls may be the result of a strategy to signal the costliness of their efforts in order to invest more in relationships and status seeking behaviors outside the household.
+In many societies, girls do more alloparenting than boys [@helfrecht2016; @weisner1977; @barry1971; @turke1988], which matches reports from our focus groups of Utilian mothers.
+It is plausible that older girls are pressured to do more care of their younger siblings, which in other studies was found to spark parent-offspring conflict [@lowe2003; syme2023a].
+If so, girls with excessive alloparenting responsibilities might be signaling their need for more time to invest in peer relationships and status seeking behaviors outside the household.
+Opposed to this interpretation, our study found a small negative assocation between girls' alloparenting and caregiver reports of conflict [@fig-lasso], although perhaps when girls do what caregivers want, caregivers report less conflict.
 
@@
 
-Evidence for conflicts similar to the one hypothesized here has been reported by multiple researchers working with the Chuukese.
+<!--Evidence for conflicts similar to the one hypothesized here has been reported by multiple researchers working with the Chuukese.
 Lowe found that both male and female children reported conflict with their families over their desires to invest financial resources in their own status enhancing opportunities rather than in the household [@lowe2003].
@@
 In response to such conflicts, participants reported using a wide range strategies as children including, deception, withholding cooperation, running away, and seeking the assistance of others.
-Through these means, participants were often able to resolve conflicts in their favor despite their lower status in the household and community.
+Through these means, participants were often able to resolve conflicts in their favor despite their lower status in the household and community.-->
 
-However, child alloparenting effort was negatively associated with caregiver reports of child conflict.
+<!--However, child alloparenting effort was negatively associated with caregiver reports of child conflict.
 One potential reason for this is since parents presumably value the child alloparenting within the family, they see their relationships with children who invest more in the household as less conflictual.
@@
 However, the bidirectional relationship between child signaling and conflict with caretakers prevents this.
-It is also possible that our hypothesis about conflict over investment in the household is incorrect.
+It is also possible that our hypothesis about conflict over investment in the household is incorrect.-->
 
-There was a positive association between the frequency of child sadness and caretaker perceptions of child neediness within the family.
+<!--There was a positive association between the frequency of child sadness and caretaker perceptions of child neediness within the family.
 However, the other signaling measures examined had coefficients set to 0 in the same lasso model despite the expectation that their greater costliness would make them more credible sources of information to parents.
 One reason for this might be that costs of sadness, while minor short-term, can be significant long-term.
-These costs are clearly seen in depression, which often involves an extended period of a lack of interest in one's normal activities and decreased investment in oneself and one's cooperative relationships and has been suggested to capable of credibly signal need in times of substantial conflicts of interest [@watson2002; @hagen2003].
+These costs are clearly seen in depression, which often involves an extended period of a lack of interest in one's normal activities and decreased investment in oneself and one's cooperative relationships and has been suggested to capable of credibly signal need in times of substantial conflicts of interest [@watson2002; @hagen2003].-->
 
-Another possibility is that parents are using cues from sadness (i.e., withdrawal and rumination) to evaluate the relative neediness of children.
+<!--Another possibility is that parents are using cues from sadness (i.e., withdrawal and rumination) to evaluate the relative neediness of children.
 As cues are not created for the purpose of communication, they are completely reliable if perceived and interpreted correctly.
 Therefore, if some cues are hard to or are impossible to fake, they may be more reliable than most signals.
-That said, it is likely that some symptoms of sadness are simultaneously the result of both signals and cues, something which may have been favored by natural selection if the threat of deception meant cue-level symptoms were insufficient in eliciting support without upregulation, something which has been suggested to be the case with various illness symptoms [@tiokhin2016; @steinkopf2015].
+That said, it is likely that some symptoms of sadness are simultaneously the result of both signals and cues, something which may have been favored by natural selection if the threat of deception meant cue-level symptoms were insufficient in eliciting support without upregulation, something which has been suggested to be the case with various illness symptoms [@tiokhin2016; @steinkopf2015].-->
 
-Caretakers perceived older children as needier, despite having more conflicts with younger children.
-One likely reason for this might be that caretakers percieve the types of investment older children seek as more costly.
+<!--Caretakers perceived older children as needier, despite having more conflicts with younger children.
+One likely reason for this might be that caretakers perceive the types of investment older children seek as more costly.
 In line with this reasoning, multiple caretakers mentioned that older kids need more money and advice and get into trouble more often during surveys.
-It may also be that caretakers see investing in older children as relatively costly as they seek to withdraw investment in order to invest in younger children.
+It may also be that caretakers see investing in older children as relatively costly as they seek to withdraw investment in order to invest in younger children.-->
 
-Signaling behavior did not predict differences in relative caregiver investment between children within households.
+<!--Signaling behavior did not predict differences in relative caregiver investment between children within households, beyond its effect on perceived need [@sfig-plot-invest-coefs].
 However, greater perceptions of need among caregivers were associated with increased likelihood of a child being reported to receive more investment in terms of time, effort, and resources relative to other children within the household.
@@
 In other words, much of the effect of signaling on levels of support is expected to be mediated by the change in the beliefs of others that one could benefit from increased support that it causes (e.g., that a child is truly in need due to adversity) rather than direct effects.
-In line with this reasoning, an experimental vignette study which found that depression and suicidality increased the likelihood of costly support from participants in times of uncertainty and substantial conflicts of interest also found that between 67% and 78% of the effect of signal type on the likelihood of costly support was mediated by the effect of signaling on belief in need [@gaffney2022].
+In line with this reasoning, an experimental vignette study which found that depression and suicidality increased the likelihood of costly support from participants in times of uncertainty and substantial conflicts of interest also found that between 67% and 78% of the effect of signal type on the likelihood of costly support was mediated by the effect of signaling on belief in need [@gaffney2022].-->
 
-Parents were more likely to respond positively and less likely to respond negatively to the last time a child signaled when the cause involved discomfort, pain, injury, or illness than when it did not.
+<!--Parents were more likely to respond positively and less likely to respond negatively to the last time a child signaled when the cause involved discomfort, pain, injury, or illness than when it did not.
 One potential causal explanation for this is that parents may often have information for sources outside of child signaling to corroborate a child's need (e.g., a fever, a diagnosis, or witness a child hurt themselves) so that there is less private information about a child's current state.
@@
 a skinned knee).
-It may also be that many of the typical positive responses to pain and illness are not very costly to caretakers (e.g., providing comfort), thereby increasing their frequency.
+It may also be that many of the typical positive responses to pain and illness are not very costly to caretakers (e.g., providing comfort), thereby increasing their frequency.-->
 
-When the cause involved the child having made a transgression, punishment, or the loss of privleges or an item, however, parents were more likely to respond negatively and less likely to respond positively.
+<--When the cause involved the child having made a transgression, punishment, or the loss of privleges or an item, however, parents were more likely to respond negatively and less likely to respond positively.
 In these circumstances, children may be signaling about the severity of a punishment [@syme2019].
@@
 They may also react with scolding or further punishment.
-That said, our model predicts positive responses in `r #signif(updated_out$estimate[3] * 100, 2)`% of cases (when all other predictors are held at their mean values), suggesting that children are successful in negotiating better outcomes at least some of the time.
+That said, our model predicts positive responses in `r #signif(updated_out$estimate[3] * 100, 2)`% of cases (when all other predictors are held at their mean values), suggesting that children are successful in negotiating better outcomes at least some of the time.-->
 
@@
 
-The main limitation of this study is that it is completely exploratory, something which results in a high risk of our models overfitting the data and indicating associations which might not exist in the real world.
-Our study was also predominantly cross-sectional, further limiting our ability to make causal inferences.
+The main limitation of this study is that it is completely exploratory: our data-driven analyses, which can reveal unexpected relationships and inspire new theory, also risk overfitting the data, suggesting associations that might not exist.
+Our study was also entirely cross-sectional, further limiting our ability to make causal inferences.
 
@@
 This could happen if there was selection bias so that those who participated differed from the rest of the population in how they interacted with their children or how they assessed, remembered, and reported child behavior.
+All data come from caregiver reports, and therefore, reflect the biases of caretakers, whatever those might be.
 
-~~Outside of child anthropometric measurements,~~ All data come from caregiver reports, and therefore, reflect the biases of caretakers, whatever they may be.
-Due to the fact the data used here were a part of survey which covered many topics and often multiple children, we also faced a tradeoff between survey time and question depth.
+<!--Due to the fact the data used here were a part of survey which covered many topics and often multiple children, we also faced a tradeoff between survey time and question depth.
 Asking adult participants about their relative investment in children in their household was likely not as effective as more in depth, questions focused on specific questions.
@@
 Our questions about the last time the child signaled may have also been more effective if they focused on specific signals due to our methods combining crying, which is unambiguous as a signal, with sadness, which has both signaling and non-signaling functions.
+-->
 
@@
 Children face the challenge of eliciting support from others in times of adversity and opportunity.
-The signaling strategies of Utilian children were highly correlated and engaged in by children of all ages.
-Consistent with signaling models which emphasize the role of conflicts of interest in favoring the use of costlier signals, 1) child signaling effort was positively correlated with the frequency of conflict with caretakers and 2) greater perceptions of child need predicted greater investment in a given child.
-Our results suggest there is much potential for signaling theory to help explain patterns of child crying and temper tantrums.
+The signaling strategies of Utilian children were highly correlated and engaged in by children of all ages, albeit more frequently by younger children and those living in more adverse conditions, who plausibly had greater need.
+Consistent with signaling models that emphasize the role of conflicts of interest in favoring the use of costlier signals, child signaling effort was positively correlated with the frequency of conflict with caretakers and with perceptions of need, and greater perceptions of child need predicted greater investment in a given child.
+Our results suggest that evolutionary theories of signaling can help explain patterns of child sadness, crying, and temper tantrums.
 
@@
 
-::: {#sfig-pot-caregiver-punish}
+::: {#sfig-cause-clustering}
+```{r}
+plot(cause_cluster_analysis)
+```
+
+Cluster dendrogram for the causes of child signaling. Method = ward.D2.
+:::
+
+::: {#sfig-sumtab-causes}
+```{r}
+c2 <- causes[5:21] |>
+    mutate(
+        across(everything(), \(x) ifelse(x == 0, "No", "Yes"))
+    )
+
+datasummary_skim(data = c2)
+# datasummary_skim(data = c2, type = "categorical") # seems to have the seem output but might be slower
+```
+
+Frequencies of various causes of the last time each child was sad or cried as reported by caretakers.
+:::
+
+::: {#sfig-plot-caregiver-punish}
 ![](Figures/plot_caregiver_punish.svg)
 
-The likelihood of positive, neutral, and negative responses to the last time a child became sad or cried when punishment was implicated in the cause (0 = no punishment, 1 = punishment.)
+The likelihood of positive, neutral, and negative responses to the last time a child became sad or cried when punishment was implicated in the cause when holding continuous variables at their mean and factor variables at their mode.
+1 = evidence of the cause involving punishment.
+0 = no evidence of the cause involving punishment.
+The endpoints of the lines are dodged slightly for greater visibility.
+Dot size represents model predictions for each participant given their values of predictor variables.
 :::

### 2024-11-05 dictionaries.R

@@ -150,7 +150,7 @@ shortform_dict <-
     RelativeMaternalInvestment2 = "Relative investment (w/in household)",
     ConflictFamily = "Family conflict",
     TransgressionMade = "Transgression made",
-    LossOfPrivlegesOrItem = "Loss of privleges or item"
+    LossOfPrivlegesOrItem = "Loss of privileges or item"
 
   )
 

### 2024-11-05 regularize.R

@@ -547,8 +547,8 @@ ggsave("Figures/plot_caregiver_response_combined.pdf", plot_caregiver_response_c
 ggsave("Figures/plot_caregiver_response_combined.svg", plot_caregiver_response_combined, width = 12, height = 12)
 
 plot_caregiver_punish <- ordinal_plot(out$fit, Punishment, data = SignalVars3, title = 'Caregiver response')
-ggsave("Figures/plot_caregiver_punish.pdf", plot_caregiver_response_combined, width = 12, height = 12)
-ggsave("Figures/plot_caregiver_punish.svg", plot_caregiver_response_combined, width = 12, height = 12)
+ggsave("Figures/plot_caregiver_punish.pdf", plot_caregiver_punish, width = 12, height = 12)
+ggsave("Figures/plot_caregiver_punish.svg", plot_caregiver_punish, width = 12, height = 12)
 
 # want to add relativeneed and relativeinvestment but code does not work
 

### 2024-11-06 Paper.qmd
Commiting final changes before circulating to co-authors.

@@
 format:
+  docx:
+    number-sections: true
   html:
@@
 
-To gain support, children use signals to communicate their desires to parents.
+To gain support, children use signals to communicate their needs and wants to parents.
 Infant signals of need, particularly infant cries, have been extensively studied in a range of populations.
-However, the full range of potential child signals of need, which extend beyond cries, has never been investigated in children of all ages.
-To address this, we used exploratory analyses on caregiver reports of child signaling data from xxx families (xx children) on Utila, Honduras.
-We found that child signaling decreased with age and neighborhood quality and was highly correlated with the frequency of conflict between children and caretakers.
-Consistent with signaling theory, children who were perceived as needier within the household were more likely to receive more investment.
-They were also more likely to receive a pro-child response when the cause of their last instance of signaling involved discomfort, pain, injury, or, illness, but less likely to receive a beneficial response when it involved prior punishment.
-Our results show there is much potential to the further study of sadness, crying, and temper tantrums among older children using signaling theory.
+However, the full range of potential child signals of need, which extend beyond cries, has rarely been investigated in a single study of children of all ages.
+To fill this gap, we collected mother and other primary caregiver reports of child signaling from `{r} nrow(caregivers)` families with `{r} nrow(modeldf)` children on Utila, a small island off the coast of Honduras.
+In exploratory analyses, we found that child signaling was common in both sexes and across all ages, although it decreased with age and neighborhood quality and increased with the frequency of conflict between children and caretakers.
+Consistent with signaling theory, children who were sad more frequently were perceived as needier within the household and were more likely to receive investment.
+Consistent with previous research, increased alloparenting effort was associated with more signaling by girls but less signaling by boys.
+Our results suggest that evolutionary theories of signaling can help explain patterns of child sadness, crying, and temper tantrums.
 
@@
 Unlike chimpanzee offspring, human offspring in ancestral populations therefore often grew up with multiple older and/or younger siblings who were potential competitors for parental investment and yet who could also provide (or require) allocare [@helfrecht2016; @lawson2016; @hagen2009].
+Juveniles in humans, unlike chimps, therefore evolved in families with multiple siblings simultaneously dependent on parents and others for essential care, likely selecting for credible signals of need.
 
@@
 
-For the same reason, siblings also often find themselves in conflicts of interest with one another, with each being expected to prefer more resources for themself than others.
+For the same reason, siblings also often find themselves in conflicts of interest with one another, with each expected to prefer more resources for himself or herself than others.
 Sibling competition is evident across taxa and can range from relatively benign forms like strong begging competition in chicks [@caro2016] to siblicide in birds, mammals, and sharks [@salmon2014].
@@
 Signals differ from cues in that signals were favored by natural selection due to their ability to alter the behavior or perceptions of others in ways that benefit the signaler, on average, whereas cues are information organisms use to influence their behavior or physiological state that did not evolve for communicative effect [@maynardsmith1995].
-Crying when injured is likely a signal of need[\[1\]](#_ftn1), for instance, whereas bleeding would be a cue of need.
+Crying when injured is likely a signal of need, for instance, whereas bleeding would be a cue of need.
 
@@
 As crying comes with physiological costs [@rao1997; @thureen1998], it could serve as a credible signal of vigor if these costs are too high for unhealthy infants to bear [@furlow1997; @lummaa1998; @soltis2004; @wells2003].
-Alternatively, crying might impose costs on caregivers by, e.g., attracting predators or interfering with important activities, motivating caregivers to invest more to stop the crying [so-called blackmail or bargaining models; @xx, @yy].
+Alternatively, crying might impose costs on caregivers by, e.g., attracting predators or interfering with important activities, motivating caregivers to invest more to stop the crying [so-called blackmail or bargaining models; @zahavi1977a; @eshel1991; @bergstrom1999 @kennedy2021].
 More generally, contact calls between parents and offspring have evolved in numerous species, so cries might simply be a means for infants to elicit attention from caregivers [@mehr2021].
@@
 
-In all forms of recruitment, adult participants were informed of the purpose of the study, what they would be asked to do, and the inclusion criteria, such as being 18 or older, having permanent residence on Utila, and having a child between 6 and 18 years old.
-
-\[no longer needed\] XXXXXXXXCHIld RecruitmentXXXXXXXXXXX
+In all forms of recruitment, participants were informed of the purpose of the study, what they would be asked to do, and the inclusion criteria, such as being 18 or older, having permanent residence on Utila, and having a child between 6 and 18 years old.
 
@@
 
-All participants for the survey portion of the study met the inclusion criteria.
-Most ~~adult~~ participants were the biological mothers of one or more children in the study.
-~~All children in the study, whose ages ranged from 5-20, were cared for by one of the adult participants.~~
+All participants for the survey portion of the study met the inclusion criteria. Most participants were the biological mothers of one or more children in the study.
 
@@
 
-All ~~adult~~ participants provided informed consent.
+All participants provided informed consent.
 The surveys, which included many topics outside of child signaling, were designed to last about an hour, plus 20 minutes for each additional child.
-~~Adults~~ Participants were paid \$XXXXX for the survey portion of the study.
+Participants were paid \$XXXXX for the survey portion of the study.
 
-~~xxxxxxxxxxxxxChild consent infoxxxxxxxxxxxxxx.~~
-
-This study was approved by the xxxxxxxIRBINFOxxxxxxxxxxxxx.
+This study was approved by the WSU xxxIRBINFOxxx and BU xxxIRBINFOxxx.
 
@@
 |  |  |
-|-------------------------|-----------------------------------------------|
+|--------------------------|----------------------------------------------|
 | Variable | Survey question wording |
@@
 Categories of causes were based on cluster analysis of more fine-grained cause variables.
-See @sfig-cause-clustering for cluster dendrogram and @sfig-sumtab-causes for the frequencies of causes and caregiver responses.
+See @sfig-cause-clustering for cluster dendrogram and @stbl-sumtab-causes for the frequencies of causes and caregiver responses.
 :::
@@
 Caregivers reported a wide range of child signaling frequencies.
-About half of children were reported to cry or throw tantrums more than once a month and almost 40% were reported to do so more than once a week [@fig-barplot_SignalFreq].
-Frequencies of sadness, crying, and tantrums were positively correlated with each other and with ~~family conflict~~ the frequency of conflict with the primary caretaker [@fig-signal_corrplot].
+About half of children were reported to cry or throw tantrums more than once a month, and almost 40% were reported to do so more than once a week [@fig-barplot_SignalFreq].
+Frequencies of sadness, crying, and tantrums were positively correlated with each other and with the frequency of conflict with the primary caretaker [@fig-signal_corrplot].
 They also loaded on PC1 in the same direction as indices of adversity, such as food insecurity, and in the opposite direction of indices of wealth and environmental quality, such as possessions, neighborhood quality, and income [@fig-PCA].
-Signaling decreased with child age, and increased with ~~family~~ conflict with their primary caretaker, which were the two variables most closely linked with signaling in the MST [@fig-mst].
+
+Signaling decreased with child age, and increased with conflict with their primary caretaker, which were the two variables most closely linked with signaling in the MST [@fig-mst].
 In our lasso regression models (which excluded family conflict due to the likely bidirectional relationship between signaling and perceived conflict), child age and high neighborhood quality were consistently negatively associated with signaling frequency, as were household income, caregiver education, and number of adults helping with childcare.
@@
 
-These results are consistent with the hypothesis that sadness, crying, and tantrums are signals of need, and that children's need is higher in adverse environments and at younger ages.
+Despite the differences in wealth, job certainty, discrimination, and living conditions between recent immigrants from the mainland and native Utilians [@davis2014; @garcia2017; @garcia2018; @stonich2000], primary language, immigration status, and residence in Camponado did not predict child signaling in our lasso regressions.
+These variables correlated with neighborhood quality, however, which was retained in our models, and lasso regression tends to retain only one of a set of correlated predictors [@zou2005].
+
+The foregoing results are consistent with the hypothesis that sadness, crying, and tantrums are signals of need, and that children's need is higher in adverse environments and at younger ages.
 One possible reason for the negative association of signaling with age, which is also seen in more industrialized settings [@zeifman2013; @potegal2003; @koch2003; @macfarlane1962], is that as children mature, they are more capable of fulfilling their own needs, more able to ask others for what they need [@osterman2010], and less vulnerable to threats.
@@
 Although conflicts of interest do not always lead to overt behavioral conflict [@altmann1980; @bateson1994], it is likely many, if not most, conflicts are ultimately caused by conflicts of interest.
-Conflicts, in turn, were negatively associated with child age and higher neigborhood quality, and positively associated with medical problems and (oddly) caregiver education [@fig-conflict-effects].
+Conflicts, in turn, were negatively associated with child age and higher neigborhood quality, and positively associated with male sex, medical problems, and (oddly) caregiver education [@fig-conflict-effects].
 
@@
 These patterns probably reflect a combination of confluences of interest vs.
-conflicts of interest between child and caretaker, and less caretaker uncertainty about need in cases of injury and illness.
-That said, our model predicts positive responses in `r #signif(updated_out$estimate[3] * 100, 2)`% of cases involving trangressions (when all other predictors are held at their mean values), suggesting that children are successful in negotiating better outcomes at least some of the time.
+conflicts of interest between child and caretaker, some of which might represent caretaker investment in some children more than others, and less caretaker uncertainty about need in cases of injury and illness.
+That said, our model predicts positive responses in 35% of cases involving trangressions, suggesting that children are successful in negotiating better outcomes at least some of the time.
 
@@
 
-Interestingly, older children were perceived to be needier than younger ones [@@fig-effect_relativeneed].
+Interestingly, older children were perceived to be needier than younger ones [\@ @fig-effect_relativeneed].
 One likely reason for this might be that caretakers perceive the types of investment older children seek as more costly.
 In line with this reasoning, multiple caretakers mentioned that older kids need more money and advice and get into trouble more often.
-It might also be that caretakers see investing in older children as relatively costly as they seek to withdraw investment in order to invest in younger children.
-
-<!- If this is the case, the positive correlation between signaling effort and conflict frequency may represent children's attempts at credibly signaling their value of additional investment or the costliness of punishment [@syme2019] to parents who would have good reason to doubt the trustworthiness of simple verbal requests (or the trustworthiness of facial sadness and crying for those resorting to temper tantrums).
-It is also likely that parents find some forms of child signaling aversive, and therefore see it as conflict in and of itself.
-->
-
-\<!- In times when an older child's valuation of potential increased support is small, such costs may be prohibitive.
-However, in times when a child places more value on potential support, those same costs could allow for a child to credibly signal need if they are too high to be worth it for deceptive signaling but still made up for by the benefits for credible signalers.
--\>
+It might also be that caretakers see investing in older children as relatively costly as they seek to redirect investment toward younger, needier children.
 
 Finally, based on previous research showing that family demands, such as caring for younger siblings, was an important source of parent-offspring conflict [@syme2019], we investigated the relationship between the frequency of alloparenting by children and their frequency of signaling.
-We found that higher frequencies of alloparenting by girls was associated with a higher frequency of tantrums and overall signaling costs, but this relationship was slightly negative for boys [@fig-alloparentingXsex].
+Although there was little evidence for sex differences in signaling overall (perhaps because sex differences typically emerge in adolescence, and our sample of adolescents was small), there was evidence that sex interacted with alloparenting effort.
+We found that higher frequencies of alloparenting by girls was associated with a higher frequency of tantrums and overall signaling costs, whereas for boys this relationship was slightly negative [@fig-alloparentingXsex].
 In many societies, girls do more alloparenting than boys [@helfrecht2016; @weisner1977; @barry1971; @turke1988], which matches reports from our focus groups of Utilian mothers.
@@
 
---- Hagen stopped here ---
-
-\<!- We found evidence of an interaction between how often a child cared for their siblings and the sex of the child in our lasso models examining signal frequencies and cost, whereby there was a positive relationship between the frequency and summed costliness of child signals of need and alloparenting effort for girls but no relationship or a slightly negative one for boys.
--\>
-
-<!--Evidence for conflicts similar to the one hypothesized here has been reported by multiple researchers working with the Chuukese.
-Lowe found that both male and female children reported conflict with their families over their desires to invest financial resources in their own status enhancing opportunities rather than in the household [@lowe2003].
-In an analysis of conflict among Chuukese-Americans, Syme and Hagen found that severe conflicts were more likely with greater numbers of siblings and when conflicts involved a child spending time away from home or evading family responsibilities [@syme2023a], findings which would be consistent with parents pressuring older children into investing more in the household among other explanations.
-In response to such conflicts, participants reported using a wide range strategies as children including, deception, withholding cooperation, running away, and seeking the assistance of others.
-Through these means, participants were often able to resolve conflicts in their favor despite their lower status in the household and community.-->
-
-<!--However, child alloparenting effort was negatively associated with caregiver reports of child conflict.
-One potential reason for this is since parents presumably value the child alloparenting within the family, they see their relationships with children who invest more in the household as less conflictual.
-In line with this reasoning, we found evidence of a child alloparenting effort by child age interaction, with our model predicting increased levels of conflict for older children with greater alloparenting \[if we say this we should report the results of that lasso too\].
-Ideally, we could control for signaling behavior to test this.
-However, the bidirectional relationship between child signaling and conflict with caretakers prevents this.
-It is also possible that our hypothesis about conflict over investment in the household is incorrect.-->
-
-<!--There was a positive association between the frequency of child sadness and caretaker perceptions of child neediness within the family.
-However, the other signaling measures examined had coefficients set to 0 in the same lasso model despite the expectation that their greater costliness would make them more credible sources of information to parents.
-One reason for this might be that costs of sadness, while minor short-term, can be significant long-term.
-These costs are clearly seen in depression, which often involves an extended period of a lack of interest in one's normal activities and decreased investment in oneself and one's cooperative relationships and has been suggested to capable of credibly signal need in times of substantial conflicts of interest [@watson2002; @hagen2003].-->
-
-<!--Another possibility is that parents are using cues from sadness (i.e., withdrawal and rumination) to evaluate the relative neediness of children.
-As cues are not created for the purpose of communication, they are completely reliable if perceived and interpreted correctly.
-Therefore, if some cues are hard to or are impossible to fake, they may be more reliable than most signals.
-That said, it is likely that some symptoms of sadness are simultaneously the result of both signals and cues, something which may have been favored by natural selection if the threat of deception meant cue-level symptoms were insufficient in eliciting support without upregulation, something which has been suggested to be the case with various illness symptoms [@tiokhin2016; @steinkopf2015].-->
-
-<!--Caretakers perceived older children as needier, despite having more conflicts with younger children.
-One likely reason for this might be that caretakers perceive the types of investment older children seek as more costly.
-In line with this reasoning, multiple caretakers mentioned that older kids need more money and advice and get into trouble more often during surveys.
-It may also be that caretakers see investing in older children as relatively costly as they seek to withdraw investment in order to invest in younger children.-->
-
-<!--Signaling behavior did not predict differences in relative caregiver investment between children within households, beyond its effect on perceived need [@sfig-plot-invest-coefs].
-However, greater perceptions of need among caregivers were associated with increased likelihood of a child being reported to receive more investment in terms of time, effort, and resources relative to other children within the household.
-As previously mentioned, credible signals are capable of leading to pro-signaler responses due to their ability to reveal private information to others who would be motivated to increase their investment in another should they have that information.
-In other words, much of the effect of signaling on levels of support is expected to be mediated by the change in the beliefs of others that one could benefit from increased support that it causes (e.g., that a child is truly in need due to adversity) rather than direct effects.
-In line with this reasoning, an experimental vignette study which found that depression and suicidality increased the likelihood of costly support from participants in times of uncertainty and substantial conflicts of interest also found that between 67% and 78% of the effect of signal type on the likelihood of costly support was mediated by the effect of signaling on belief in need [@gaffney2022].-->
-
-<!--Parents were more likely to respond positively and less likely to respond negatively to the last time a child signaled when the cause involved discomfort, pain, injury, or illness than when it did not.
-One potential causal explanation for this is that parents may often have information for sources outside of child signaling to corroborate a child's need (e.g., a fever, a diagnosis, or witness a child hurt themselves) so that there is less private information about a child's current state.
-Simultaneously, conflicts of interest may be less significant in these times due to either 1) both sides benefiting similarly from their child returning to normal (e.g., with an illness) or 2) the cause involving something which is minor and would quickly resolve itself (e.g.
-a skinned knee).
-It may also be that many of the typical positive responses to pain and illness are not very costly to caretakers (e.g., providing comfort), thereby increasing their frequency.-->
-
-<--When the cause involved the child having made a transgression, punishment, or the loss of privleges or an item, however, parents were more likely to respond negatively and less likely to respond positively.
-In these circumstances, children may be signaling about the severity of a punishment [@syme2019].
-However, parents may be reluctant to reduce punishments if the behavior that warranted the punishment was repeated or severe.
-They may also react with scolding or further punishment.
-That said, our model predicts positive responses in `r #signif(updated_out$estimate[3] * 100, 2)`% of cases (when all other predictors are held at their mean values), suggesting that children are successful in negotiating better outcomes at least some of the time.-->
-
-Parents were also more likely to respond negatively and less likely to respond positively to children when the cause involved family conflict.
-This might reflect tradeoffs within the family, whereby investment in one child cannot go to another.
-Caretakers may have also responded more negatively when the cause involved family conflict, if it involved a child imposing costs on other family members.
-
-\[delete this paragraph?\] We found a negative association between caretakers’ education level (years in school) and many of our child signaling measures despite our models controlling for household income and partner status.
-One possible reason for this may be that more educated caretakers see their children less during the day due to greater ability to afford childcare or due to them having jobs that are less compatible with child care (test model of social support predicted by education).
-Unfortunately, we do not have data on how much time caretakers spent with children to test this, and it seems equally likely that less educated parents may see their kids less often if they need to work more hours to support their families.
-Another possibility is that lower levels of education are associated with greater levels of adversity for both parents and children, something which appears to be the case for other measures of SES [@pepper2017] and would increase the benefits of signaling need for children and the difficulty of parents to respond.
-It is also possible that education influences parenting norms in ways that cause differences in child signaling behavior.
-
-Outside of the sex by alloparenting effort interaction, we found little evidence of sex differences in child signaling behavior.
-As adults, females cry more than males, a finding that has been reported across cultures [@vanhemert2011; @becht2001].
-One reason why our study may have failed to find significant effects of sex is that sex differences in the causes and reputational effects of signaling need are small or nonexistent until adolescence and young adulthood, with our small sample size for this age range preventing us from detecting a real effect should it exist.
-It is also possible that the sex difference seen in rates of crying in adults may reflect the lesser ability of females to bargain for better treatment through aggression or the threat of it when faced with adversity [@hagen2016].
-Such a sex difference may not exist among preadolescencents due to a lack of difference in strength or athletic ability before puberty [@handelsman2018] or due to their relative lack of formidability compared to adults.
-
-Despite the differences in wealth, job certainty, discrimination, and living conditions [@davis2014; @garcia2017; @garcia2018; @stonich2000] between recent immigrants from the mainland and native Utilians, primary language, immigration status, and residence in Camponado did not predict child signaling in our lasso regressions.
-However, neighborhood quality, which was strongly negatively correlated with residence in Camponado, (check others) was preserved in the lasso regressions predicting every signaling measure but sadness and was among the variables with the largest negative coefficients.
-Monthly income was also negatively associated with crying frequency,\~\~ tantrum frequency (too small of a coefficient to mention here?),\~\~ summed signal frequency, and signal cost albeit with smaller coefficients.
-
 Many models designed to explain signals of need (e.g., the begging of young birds) predict that offspring who are worse off or are receiving less investment have more to gain from signaling than those with more [@mock2011; @godfray1991; @caro2016a; @godfray2000] and multiple game theoretical models examining parent-offspring conflict have found that parental resources have a modest impact on the evolutionary stable value for offspring demand for resources [@bossan2013; @mock1997 [confirm]].
-In line with this reasoning, children in lower income households and riskier environments might benefit more from signaling effort if they experience greater levels of adversity or perceive themselves as disadvantaged compared to other children they may be competing with for status and desirable cooperative relationships.
-~~In addition to overall effort in eliciting need, resource availability may have also affect how children signal for more investment based on the finding that birds in stable and unusually favorable environments are more likely to compete with siblings through signals thought to reveal information about need, while those in relatively unpredictable and unusually poor environments have been found to compete more through signals thought to reveal information about quality [@caro2016a].~~
-
-(temporarilly moved here) Tantrums were present among older children and might be an under-appreciated way by which children signal for increased support.
-Most studies of tantrums involve children under 4 years old or older children with disruptive mood dysregulation disorder, likely due to their greater prevalence in these populations (cite).
-However, multiple studies which included older children from non-clinical populations have found similar evidence of tantrums [@macfarlane1962; @barbarin1999; @ward2013; @osterman2010; @goodenough1931; @bhatia1990], as have ethnographic studies [@ward2013; @lessa1950; @dennis1940].
+In line with this reasoning, children in lower income households and riskier environments might benefit more from signaling effort if they experience greater levels of adversity or perceive themselves as disadvantaged compared to other children they may be competing with for resources, status, and valuable social relationships.
 
@@
 
-<!--Due to the fact the data used here were a part of survey which covered many topics and often multiple children, we also faced a tradeoff between survey time and question depth.
-Asking adult participants about their relative investment in children in their household was likely not as effective as more in depth, questions focused on specific questions.
-Framing the question this way may have also increased the likelihood of social desirability bias [@grimm2010], which could lead adult participants to be be more likely to report equal investment across children.
-Our questions about the last time the child signaled may have also been more effective if they focused on specific signals due to our methods combining crying, which is unambiguous as a signal, with sadness, which has both signaling and non-signaling functions.
--->
-
 # Conclusion
@@
 
-**Footnotes**
-
-[\[1\]](#_ftnref1) We expect “signals of need” to actually be signals of resource valuation.
-However, we use “signal of need” here to distinguish signals that involve attempts at eliciting support through revealing private information about one’s vulnerabilities from the more commonly studied signals which reveal private information about one’s positive qualities (i.e., those related to partner choice and intrasexual competition).
-
-[\[2\]](#_ftnref2) Our use of “credible” signaling is largely synonymous with the more common term, honest signaling.
-However, we prefer the term credible signals since signals do not need to be completely honest for natural selection to have favored individuals who paid attention to them.
-
-[\[3\]](#_ftnref3) Researchers often distinguish between sadness and low mood based on the length of negative affect.
-We remain agnostic on how different sadness and low mood are from one another but use the terms researchers used when describing their research.
-
-[\[4\]](#_ftnref4) All else is often not equal when it comes to the costliness of investment as older children may often cost more (e.g., the costs education and food are often higher among older children).
-However, parents would still have to pay similar costs to younger children as they grew, something which natural selection has likely acted on when it comes to our adaptations involved with differential child investment.
-
-[\[5\]](#_ftnref5) Even in instances where conflicts begin due to mistakes, the initial conflict can lead to increased conflicts of interest.
-This may be seen with findings that individuals who accidentally harm another often grow to dislike them (cite), something which has been suggested to be adaptive to the accidental perpetrator as the unintended victim is likely to value than they would have before, potentially to the point of wishing them harm.
-
-[\[6\]](#_ftnref6) Despite the risk, Shostak notes there were no known recent incidents of harm to women who ran away to spend the night in the Bush [@shostak1981].
-
 # Supplementary Information
@@
 |  |  |
-|-------------------------|----------------------------------------------|
+|--------------------------|----------------------------------------------|
 | Variable | Survey question wording |
@@
 |  |  |
-|-----------------------|------------------------------------------------|
+|-------------------------|------------------------------------------------|
 | Variable | Survey question wording |
@@
 |  |  |
-|------------------------|------------------------------------------------|
+|-------------------------|-----------------------------------------------|
 | Variable | Survey question wording |
@@
 ::: {#stbl-possession-list}
-Motorized vehicle (scooter, 4-wheeler, etc.) Bicycle Television (TV) with cable Smartphone Washing machine Refrigerator A home that you own Credit card
+Motorized vehicle (scooter, 4-wheeler, etc.)\
+Bicycle\
+Television (TV) with cable\
+Smartphone\
+Washing machine\
+Refrigerator\
+A home that you own\
+Credit card
 
@@
 
-::: {#sfig-sumtab-causes}
+::: {#sfig-plot-caregiver-punish}
+![](Figures/plot_caregiver_punish.svg)
+
+The likelihood of positive, neutral, and negative responses to the last time a child became sad or cried when punishment was implicated in the cause when holding continuous variables at their mean and factor variables at their mode.
+1 = evidence of the cause involving punishment.
+0 = no evidence of the cause involving punishment.
+The endpoints of the lines are dodged slightly for greater visibility.
+Dot size represents model predictions for each participant given their values of predictor variables.
+:::
+
+## Tables
+
+::: {#stbl-sumtab-causes}
 ```{r}
@@
 Frequencies of various causes of the last time each child was sad or cried as reported by caretakers.
-:::
-
-::: {#sfig-plot-caregiver-punish}
-![](Figures/plot_caregiver_punish.svg)
-
-The likelihood of positive, neutral, and negative responses to the last time a child became sad or cried when punishment was implicated in the cause when holding continuous variables at their mean and factor variables at their mode.
-1 = evidence of the cause involving punishment.
-0 = no evidence of the cause involving punishment.
-The endpoints of the lines are dodged slightly for greater visibility.
-Dot size represents model predictions for each participant given their values of predictor variables.
 :::

### 2024-11-06 analysisforpaper.R

@@ -30,7 +30,6 @@ library(gratia)
 library(pvclust)
 library(nnet)
 library(skimr)
-library(pvclust)
 
 source("recode.R")
 source("dictionaries.R")
@@ -1584,8 +1583,8 @@ causematrix <-
 
 hagenheat(t(causematrix), hc_method = "ward.D2")
 
-out <- pvclust(causematrix, method.hclust = "ward.D2")
-x <- cutree(out$hclust, k = 3)
+cause_cluster_analysis <- pvclust(causematrix, method.hclust = "ward.D2")
+x <- cutree(cause_cluster_analysis $hclust, k = 3)
 
 causes2 <- causes |>
   mutate(

### 2024-11-06 regularize.R

@@ -384,8 +384,11 @@ ordinal_plot <- function(fit, predictor, data, title){
       {{predictor}}, Negative:Positive
     ) |>
     pivot_longer(Negative:Positive)
-  # x_axis_name <- shortform_dict[deparse(substitute(predictor))]
-  # print(x_axis_name)
+
+  dsum <- d |>
+    summarise(mean = mean(value), .by = c(name, {{predictor}}))
+  print(dsum)
+
   ggplot(d, aes({{predictor}}, value, colour = name)) +
     geom_count(position = position_dodge(width = 0.1)) +
     geom_smooth(method='lm', se = F, position = position_dodge(width = 0.1)) +
@@ -869,4 +872,3 @@ ggsave("Figures/plot_fullcorrmat.svg", plot_fullcorrmat, width = 12, height = 12
 # conflictparams <- expand_grid(Outcome = c("ConflictFreqN"), alpha = c(0, 1))
 # names(conflictparams$Outcome) <- str_c(conflictparams$Outcome, conflictparams$alpha)
 # conflictparams$out <- map2(conflictparams$Outcome, conflictparams$alpha, \(x, y) glmnet2(sv3, x, 7:ncol(sv3), alpha = y), .progress = T)
-

### 2024-11-28 Paper.qmd
Consolidating library calls and script sourcing. Also minor changes to markdown tables due to Rstudio visual mode

@@
 ```{r setup, include=FALSE}
-source("dictionaries.R", local = knitr::knit_global())
 source("analysisforpaper.R", local = knitr::knit_global())
-
 library(modelsummary)
-# source("regularize.R")
 ```
@@
 |  |  |
-|--------------------------|----------------------------------------------|
+|------------------|------------------------------------------------------|
 | Variable | Survey question wording |
@@
 |  |  |
-|--------------------------|----------------------------------------------|
+|----------------|--------------------------------------------------------|
 | Variable | Survey question wording |
@@
 |  |  |
-|-------------------------|------------------------------------------------|
+|---------------|---------------------------------------------------------|
 | Variable | Survey question wording |
@@
 |  |  |
-|-------------------------|------------------------------------------------|
+|-------------|-----------------------------------------------------------|
 | Variable | Survey question wording |
@@
 |  |  |
-|-------------------------|-----------------------------------------------|
+|------------------|------------------------------------------------------|
 | Variable | Survey question wording |
@@
 |  |  |
-|--------------------------|----------------------------------------------|
+|-------------------|-----------------------------------------------------|
 | Variable | Survey question wording |

### 2024-11-28 analysisforpaper.R

@@ -3,36 +3,42 @@
 # note: list is cleaned up but some may no longer used
 library(utiladata2023) # data package
 library(tidyverse)
-library(labelled)
-# library(lme4)
-library(effects)
-library(visreg)
-library(emmeans)
-library(car)
-library(haven)
-library(MASS)
-library(glmmTMB)
-library(sandwich)
-library(lmtest)
-library(ggcorrplot)
-library(marginaleffects)
-library(ordinal) # clmm # polr
-library(brms)
-library(hagenutils)
-library(localgrowth)
-library(knitr)
-library(kableExtra)
-library(modelsummary)
-library(ggalluvial)
-library(patchwork)
-library(mgcv)
-library(gratia)
-library(pvclust)
-library(nnet)
-library(skimr)
 
-source("recode.R")
-source("dictionaries.R")
+library(BDgraph)
+library(brms)
+library(car)
+library(easybgm)
+library(effects)
+library(emmeans)
+library(ggalluvial)
+library(ggcorrplot)
+library(ggraph)
+library(glmmTMB)
+library(glmnet)
+library(glmnetUtils)
+library(gratia)
+library(hagenutils)
+library(haven)
+library(kableExtra)
+library(knitr)
+library(labelled)
+library(lmtest)
+library(localgrowth)
+library(marginaleffects)
+library(MASS)
+library(mgcv)
+library(modelsummary)
+library(nnet)
+library(ordinal) # clmm # polr
+library(patchwork)
+library(poissonreg)
+library(pvclust)
+library(sandwich)
+library(skimr)
+library(tidygraph)
+library(tidymodels)
+library(visreg)
+
 source("dataprep.R")
 # source("regularize.R")
 

### 2024-11-28 dataprep.R

@@ -1,8 +1,3 @@
-library(utiladata2023) # data package
-library(tidyverse)
-library(labelled)
-library(mgcv)
-library(localgrowth)
 
 source("recode.R")
 source("dictionaries.R")

### 2024-11-28 regularize.R

@@ -1,19 +1,4 @@
 
-library(glmnet)
-library(glmnetUtils)
-library(hagenutils)
-library(patchwork)
-library(skimr)
-library(pvclust)
-library(ggcorrplot)
-library(tidymodels)
-library(poissonreg)
-library(marginaleffects)
-library(easybgm)
-library(tidygraph)
-library(ggraph)
-library(BDgraph)
-
 # Regularized regression --------------------------------------------------
 
 glmnet2 <- function(d, outcome, indices, alpha = 1, fam = 'quasipoisson'){

### 2024-11-30 Paper.qmd
Moved code to analysis.R. Added numbers_to_words

@@
 ```{r setup, include=FALSE}
-source("analysisforpaper.R", local = knitr::knit_global())
-library(modelsummary)
+source("analysis.R", local = knitr::knit_global())
 ```
@@
 
-`{r} dplyr::n_distinct(modeldf$householdID)` primary caretakers (females = `{r} table(caregivers$CaregiverSex)[1]`, males = `{r} table(caregivers$CaregiverSex)[2]`, non-binary = `{r} table(caregivers$CaregiverSex)[3]`) took the survey and provided signaling data for at least one child.
-`{r} nrow(modeldf)` children had frequency data for at least one signal, with `{r} nrow(modeldf_FULLSIG)` having data for all three.
+```{r}
+sex_table <- table(modeldf$Sex)
+```
+
+`{r} xfun::numbers_to_words(n_distinct(modeldf$householdID), cap=T)` primary caretakers (females = `{r} table(caregivers$CaregiverSex)[1]`, males = `{r} table(caregivers$CaregiverSex)[2]`) took the survey and provided signaling data for at least one child.
+`{r} xfun::numbers_to_words(nrow(modeldf), cap=T)` children had frequency data for at least one signal, with `{r} nrow(modeldf_FULLSIG)` having data for all three.
 Among those with at least one record of signaling frequency, `{r} sex_table[1]` were female and `{r} sex_table[2]` were male.
-Anthropometric and signaling data exist for `{r} nrow(modeldf_FULLANTHROPOMETRICS)` children, of whom `{r} sex_table2[1]` were female and `{r} sex_table2[2]` were male.
 For summary statistics of study variables see @tbl-sumtab_combined and @tbl-sumtab_factors.
@@
 
+```{r}
+rel_need_table <- table(modeldf$RelativeNeed3)
+```
+
 Primary caretakers rarely described a child as “much less” `{r} fmt_sum(modeldf$RelativeNeed == "Much less")` \[fix in word doc?\] or “much more” `{r} fmt_sum(modeldf$RelativeNeed == "Much more")` needy compared to their other children.
@@
 :::
+
+```{r}
+response_overlap <- table(modeldf$PositiveResponse, modeldf$NegativeResponse)
+```
 

### 2024-11-30 analysis.R

@@ -1,0 +1,208 @@
+library(utiladata2023) # data package
+library(tidyverse)
+library(modelsummary)
+library(labelled)
+library(ggcorrplot)
+library(pvclust)
+library(ggalluvial)
+library(hagenutils)
+library(knitr)
+library(kableExtra)
+
+source("dataprep.R")
+
+## main paper plots --------------------------------------------------------
+
+signalfreqdf <-
+  d2 |>
+  dplyr::select(
+    householdID,
+    childHHid,
+    uniqueID,
+    SadFreqOF,
+    CryFreqOF,
+    TantrumFreqOF,
+    SignalCost
+  ) |>
+  rename(
+    Sadness = SadFreqOF,
+    Crying = CryFreqOF,
+    Tantrums = TantrumFreqOF
+  )
+
+freq_short <- c(
+  "Never" = "0",
+  "Once a month or less" = "≤1",
+  "More than once a month but less than once a week" = "2-3",
+  "More than once a week but not daily" = "4-29",
+  "Daily" = "30",
+  "Multiple times per day" = ">30"
+)
+
+freq_short2 <- c(
+  "0" = "Low",
+  "≤1" = "Low",
+  "2-3" = "Medium",
+  "4-29" = "Medium",
+  "30" = "High",
+  ">30" = "High"
+)
+
+sfdf <-
+  signalfreqdf |>
+  mutate(
+    across(Sadness:Tantrums, \(x) factor(c(freq_short[x]), levels = c(freq_short)))
+  ) |>
+  na.omit()
+
+signaldf_long <-
+  sfdf |>
+  pivot_longer(cols = Sadness:Tantrums, names_to = "Signal", values_to = "Freq")  |>
+  mutate(
+    Signal = factor(Signal, levels = c("Sadness", "Crying", "Tantrums"))
+  ) |>
+  na.omit()
+
+barplot_SignalFreq <-
+  signaldf_long |>
+  group_by(Signal, Freq) |>
+  summarise(N = n()) |>
+  ggplot(aes(N, Signal, fill = fct_rev(Freq))) +
+  geom_col(position = "stack", color = "grey") +
+  scale_fill_viridis_d(option = "B", direction = -1) +
+  labs(x = "Number of children", y = "") +
+  guides(fill = guide_legend("Times per month:", nrow = 1, reverse = TRUE, position = "top")) +
+  theme_minimal(15)
+barplot_SignalFreq
+
+
[…]

### 2024-11-30 Paper.qmd
Moved PCA code into analysis.R

@@
 
-The anonymized data used in the study are available at https://github.com/grasshoppermouse/utiladata2023. Study code is available at https://github.com/michaelrgaffney/utilasignaling.
+The anonymized data used in the study are available at <https://github.com/grasshoppermouse/utiladata2023>. Study code is available at <https://github.com/michaelrgaffney/utilasignaling>.
 

### 2024-11-30 analysis.R

@@ -206,3 +206,30 @@ signal_alluvial_plot <-
   theme(axis.title.y = element_text(angle = 0, hjust = 1)) +
   scale_fill_viridis_d()
 signal_alluvial_plot
+
+# PCA ---------------------------------------------------------------------
+e <- mdf2 |>
+  dplyr::select(
+    - SignalFreq,
+    - SignalCost,
+    - SignalFreqMax,
+    - OldestChild,
+    - OlderGirls,
+    - PartnerStatus,
+    - CaregiverAge,
+    - YoungerKids,
+    - contains("Xsex")
+  )
+
+e <- set_names(e, shortform_dict[names(e)])
+# names(e) <- shortform_dict[names(e)]
+
+m <- prcomp(e, scale. = T) # Remove composite signaling vars
+plot_loadings <- pca_loadings_plot(m, 1:2) # + theme(legend.position = 'top', legend.title = element_blank())
+plot_biplot <- pca_biplot(m) + theme_minimal(15)
+
+plot_pca <- plot_loadings + plot_biplot + plot_layout(widths = c(1,2)) +
+  plot_annotation(tag_levels = "A")
+ggsave("Figures/plot_pca.svg", plot_pca)
+ggsave("Figures/plot_pca.pdf", plot_pca)
+

### 2024-11-30 analysis2.R
renamed regularize.R to analysis2.R

@@ -1,4 +1,6 @@
 
+library(glmnet)
+
 # Regularized regression --------------------------------------------------
 
 glmnet2 <- function(d, outcome, indices, alpha = 1, fam = 'quasipoisson'){

### 2024-11-30 analysis.R
Added missing code and libraries for PCA

@@ -1,13 +1,15 @@
 library(utiladata2023) # data package
 library(tidyverse)
+library(patchwork)
 library(modelsummary)
 library(labelled)
 library(ggcorrplot)
 library(pvclust)
 library(ggalluvial)
-library(hagenutils)
+library(hagenutils) # from github
 library(knitr)
 library(kableExtra)
+library(skimr)
 
 source("dataprep.R")
 
@@ -208,7 +210,38 @@ signal_alluvial_plot <-
 signal_alluvial_plot
 
 # PCA ---------------------------------------------------------------------
-e <- mdf2 |>
+
+out <- skim(modeldf)
+nms <- out$skim_variable[out$skim_type == 'numeric' & out$complete_rate > 0.9]
+nms <- c(nms, "Sex", "UserLanguage", "ImmigrateUtila", "PartnerStatus", "OldestChild")
+
+mdf2 <-
+  modeldf |>
+  dplyr::select(
+    all_of(nms),
+    -householdID,
+    -childHHid,
+    -NegativeResponse,
+    -PositiveResponse,
+    -IncomeCategoryN,
+    # -ConflictFreqN,
+    -OtherChildrenHH, # omit or keep?
+    -contains('IllnessSusceptibility')
+  ) |>
+  na.omit() |>
+  mutate(
+    Sex = ifelse(Sex == "Female", 0, 1),
+    UserLanguage = ifelse(UserLanguage == "EN", 0, 1),
+    ImmigrateUtila = ifelse(ImmigrateUtila == "No", 0, 1),
+    PartnerStatus = ifelse(PartnerStatus == "Unpartnered", 0, 1),
+    across(-c(1:6), \(x) c(scale(x))),
+    AlloparentingXsex = AlloparentingFreqN * Sex,
+    ChildAgeXsex = ChildAge * Sex,
+    OldestXsex = OldestChild * Sex
+  )
+
+e <-
+  mdf2 |>
   dplyr::select(
     - SignalFreq,
     - SignalCost,

### 2024-11-30 Paper.qmd
renamed regularize.R to analysis2.R
Make html file stand-alone

@@
     number-sections: true
+    embed-resources: true
 lightbox: true
@@
 ```{r setup, include=FALSE}
-source("analysis.R", local = knitr::knit_global())
+source("analysis.R")
+# source("analysis2.R") # slow!
 ```

### 2024-12-01 Paper.qmd
using caregiverSex vector instead of caregiver df, plus a bit of code reorganization

--- 2024-11-30 Paper.qmd
@@
         department: Anthropology
+  - name: Carolyn R. Hodges-Simeon
+    orcid: 0000-0002-0691-9080
+    corresponding: false
+    affiliation:
+      - name: Boston University
+        department: Anthropology
   - name: Edward H. Hagen
@@
       - name: Washington State University
-        department: Anthropology
-  - name: Carolyn R. Hodges-Simeon
-    orcid: 0000-0002-0691-9080
-    corresponding: false
-    affiliation:
-      - name: Boston University
         department: Anthropology
@@
 ```{r setup, include=FALSE}
-source("analysis.R", local = knitr::knit_global())
+source("recode.R")
+source("dictionaries.R")
+source("dataprep.R")
+# source("dataprepPublic.R")
+source("analysis.R")
+# source("analysis2.R") # slow!
 ```
@@
 However, the full range of potential child signals of need, which extend beyond cries, has rarely been investigated in a single study of children of all ages.
-To fill this gap, we collected mother and other primary caregiver reports of child signaling from `{r} nrow(caregivers)` families with `{r} nrow(modeldf)` children on Utila, a small island off the coast of Honduras.
+To fill this gap, we collected mother and other primary caregiver reports of child signaling from `{r} n_distinct(modeldf$householdID)` families with `{r} nrow(modeldf)` children on Utila, a small island off the coast of Honduras.
 In exploratory analyses, we found that child signaling was common in both sexes and across all ages, although it decreased with age and neighborhood quality and increased with the frequency of conflict between children and caretakers.
@@
 
-`{r} xfun::numbers_to_words(n_distinct(modeldf$householdID), cap=T)` primary caretakers (females = `{r} table(caregivers$CaregiverSex)[1]`, males = `{r} table(caregivers$CaregiverSex)[2]`) took the survey and provided signaling data for at least one child.
+`{r} xfun::numbers_to_words(n_distinct(modeldf$householdID), cap=T)` primary caretakers (females = `{r} table(caregiverSex)[1]`, males = `{r} table(caregiverSex)[2]`, missing sex = `{r} sum(is.na(caregiverSex))`) took the survey and provided signaling data for at least one child.
 `{r} xfun::numbers_to_words(nrow(modeldf), cap=T)` children had frequency data for at least one signal, with `{r} nrow(modeldf_FULLSIG)` having data for all three.

### 2024-12-01 analysis.R

@@ -1,4 +1,4 @@
-library(utiladata2023) # data package
+
 library(tidyverse)
 library(patchwork)
 library(modelsummary)
@@ -19,7 +19,9 @@ library(tidygraph)
 library(ggraph)
 library(ordinalNet)
 
-source("dataprep.R")
+modeldf_FULLSIG <-
+  modeldf |>
+  dplyr::filter(!is.na(CryFreqN) & !is.na(SadFreqN) & !is.na(TantrumFreqN))
 
 ## main paper plots --------------------------------------------------------
 

### 2024-12-01 analysis2.R

@@ -242,7 +242,7 @@ names(discrete_vars) <- names(discrete_vars0$discrete)
 # set.seed(456)
 # m3 <- easybgm(SignalVars[-c(1,2,3,4,5)], type = 'mixed', not_cont = discrete_vars[-c(1,2,3,4,5)], iter = 1000000, package = 'BDgraph', g.prior = 0.05)
 # save(m3, file = 'm3_1e6_cost_freq.rda')
-load("data/m3_1e6_cost_freq.rda")
+
 plot_network(m3, layout = "spring", vsize = 4, label.cex = 1, exc_prob = 0.9)
 plot_structure(m3, vsize = 4, label.cex = 1)
 plot_structure_probabilities(m3, as_BF = F)

### 2024-12-01 Paper.qmd
changed modeldf to utila_df for clarity

@@
 However, the full range of potential child signals of need, which extend beyond cries, has rarely been investigated in a single study of children of all ages.
-To fill this gap, we collected mother and other primary caregiver reports of child signaling from `{r} n_distinct(modeldf$householdID)` families with `{r} nrow(modeldf)` children on Utila, a small island off the coast of Honduras.
+To fill this gap, we collected mother and other primary caregiver reports of child signaling from `{r} n_distinct(utila_df$householdID)` families with `{r} nrow(utila_df)` children on Utila, a small island off the coast of Honduras.
 In exploratory analyses, we found that child signaling was common in both sexes and across all ages, although it decreased with age and neighborhood quality and increased with the frequency of conflict between children and caretakers.
@@
 
-This study of children aged `{r} paste(range(modeldf$ChildAge), collapse="-")` and their parents and other primary caretakers, was conducted in Utila, Honduras, one of the country’s small Bay Islands in the Western Caribbean.
+This study of children aged `{r} paste(range(utila_df$ChildAge), collapse="-")` and their parents and other primary caretakers, was conducted in Utila, Honduras, one of the country’s small Bay Islands in the Western Caribbean.
 Participants in the study were sampled from two populations who face different levels of economic and social adversity: 1) a long-standing English-speaking community primarily comprising descendants of British immigrants who arrived on the island in the 1800’s, and 2) a more recent community of Spanish-speakers, mostly from the mainland.
@@
 We reversed the scores on two measures so that a score of 5 corresponded with the lowest susceptibility to illness before we averaged the scores of these to create a mean *illness avoidance* variable.
-In the `{r} sum((!is.na(modeldf$IllnessSusceptibility2)) & is.na(modeldf$IllnessSusceptibility3))` instance in which we only had data for 2 of the measures, we still calculated the mean but used a numerator of 2.
+In the `{r} sum((!is.na(utila_df$IllnessSusceptibility2)) & is.na(utila_df$IllnessSusceptibility3))` instance in which we only had data for 2 of the measures, we still calculated the mean but used a numerator of 2.
 See @stbl-healthvars for the wording of these variables.
@@
 ```{r}
-sex_table <- table(modeldf$Sex)
+sex_table <- table(utila_df$Sex)
 ```
 
-`{r} xfun::numbers_to_words(n_distinct(modeldf$householdID), cap=T)` primary caretakers (females = `{r} table(caregiverSex)[1]`, males = `{r} table(caregiverSex)[2]`, missing sex = `{r} sum(is.na(caregiverSex))`) took the survey and provided signaling data for at least one child.
-`{r} xfun::numbers_to_words(nrow(modeldf), cap=T)` children had frequency data for at least one signal, with `{r} nrow(modeldf_FULLSIG)` having data for all three.
+`{r} xfun::numbers_to_words(n_distinct(utila_df$householdID), cap=T)` primary caretakers (females = `{r} table(caregiverSex)[1]`, males = `{r} table(caregiverSex)[2]`, missing sex = `{r} sum(is.na(caregiverSex))`) took the survey and provided signaling data for at least one child.
+`{r} xfun::numbers_to_words(nrow(utila_df), cap=T)` children had frequency data for at least one signal, with `{r} nrow(modeldf_FULLSIG)` having data for all three.
 Among those with at least one record of signaling frequency, `{r} sex_table[1]` were female and `{r} sex_table[2]` were male.
@@
 
-kable(custom.summarize(modeldf, longform_dict[1:23])) %>% kable_styling()
+kable(custom.summarize(utila_df, longform_dict[1:23])) %>% kable_styling()
 ```
@@
 # convert to factor to omit NAs in datasummary_skim
-dx <- modeldf |>
+dx <- utila_df |>
   dplyr::select(any_of(names(longform_dict))) |>
@@
 See @fig-barplot_SignalFreq and @fig-signal_corrplot.
-Caregivers reported that `{r} fmt_sum(modeldf$SadFreqN >= 8)` were sad at least once a week, `{r} fmt_sum(modeldf$CryFreqN >= 8)` cried at least once a week, and `{r} fmt_sum(modeldf$TantrumFreqN >= 8)` threw temper tantrums at least once a week.
-A small subset of children engaged in these signals/states at least daily; sadness: `{r} fmt_sum(modeldf$SadFreqN >= 30)`; crying: `{r} fmt_sum(modeldf$CryFreqN >= 30)`; and tantrums: `{r} fmt_sum(modeldf$TantrumFreqN >= 30)`.
-However, many children were reported to never engage in these signals/states; sadness: `{r} fmt_sum(modeldf$SadFreqN == 0)`; crying; `{r} fmt_sum(modeldf$CryFreqN == 0)`; and tantrums: `{r} fmt_sum(modeldf$TantrumFreqN == 0)`.
+Caregivers reported that `{r} fmt_sum(utila_df$SadFreqN >= 8)` were sad at least once a week, `{r} fmt_sum(utila_df$CryFreqN >= 8)` cried at least once a week, and `{r} fmt_sum(utila_df$TantrumFreqN >= 8)` threw temper tantrums at least once a week.
+A small subset of children engaged in these signals/states at least daily; sadness: `{r} fmt_sum(utila_df$SadFreqN >= 30)`; crying: `{r} fmt_sum(utila_df$CryFreqN >= 30)`; and tantrums: `{r} fmt_sum(utila_df$TantrumFreqN >= 30)`.
+However, many children were reported to never engage in these signals/states; sadness: `{r} fmt_sum(utila_df$SadFreqN == 0)`; crying; `{r} fmt_sum(utila_df$CryFreqN == 0)`; and tantrums: `{r} fmt_sum(utila_df$TantrumFreqN == 0)`.
 For a visualization of how signal frequencies varied within children see @sfig-signal_alluvial_plot.
 
-Conflict between children and caretakers was also common, with `{r} fmt_sum(modeldf$ConflictFreqN >= 8)` having conflicts with parents more than once a week.
-However, `{r} fmt_sum(modeldf$ConflictFreqN == 0)` were described as never having conflicts with parents.
+Conflict between children and caretakers was also common, with `{r} fmt_sum(utila_df$ConflictFreqN >= 8)` having conflicts with parents more than once a week.
+However, `{r} fmt_sum(utila_df$ConflictFreqN == 0)` were described as never having conflicts with parents.
 See @fig-barplot_ConflictFreq.
@@
 
-The frequency of child alloparenting effort, which previous research has linked to parent-child conflict, was bimodally distributed, with `{r} fmt_sum(modeldf$AlloparentingFreqN == 0)` reported to never engage in it (`r sum(na.omit(modeldf$NumberOfChildren==1))` instances were due to a child being an only child) and `{r} fmt_sum(modeldf$AlloparentingFreqN >= 30)` reported to help with childcare at least once a day.
+The frequency of child alloparenting effort, which previous research has linked to parent-child conflict, was bimodally distributed, with `{r} fmt_sum(utila_df$AlloparentingFreqN == 0)` reported to never engage in it (`r sum(na.omit(utila_df$NumberOfChildren==1))` instances were due to a child being an only child) and `{r} fmt_sum(utila_df$AlloparentingFreqN >= 30)` reported to help with childcare at least once a day.
 See @fig-barplot_AlloparentingFreq.
@@
 ```{r}
-rel_need_table <- table(modeldf$RelativeNeed3)
+rel_need_table <- table(utila_df$RelativeNeed3)
 ```
 
-Primary caretakers rarely described a child as “much less” `{r} fmt_sum(modeldf$RelativeNeed == "Much less")` \[fix in word doc?\] or “much more” `{r} fmt_sum(modeldf$RelativeNeed == "Much more")` needy compared to their other children.
+Primary caretakers rarely described a child as “much less” `{r} fmt_sum(utila_df$RelativeNeed == "Much less")` \[fix in word doc?\] or “much more” `{r} fmt_sum(utila_df$RelativeNeed == "Much more")` needy compared to their other children.
 Therefore, we collapsed “less” and “much less” into a “less” factor and “more” and “much more” into a “more” factor.
@@
 ```{r}
-response_overlap <- table(modeldf$PositiveResponse, modeldf$NegativeResponse)
+response_overlap <- table(utila_df$PositiveResponse, utila_df$NegativeResponse)
 ```

### 2024-12-01 analysis.R

@@ -20,13 +20,13 @@ library(ggraph)
 library(ordinalNet)
 
 modeldf_FULLSIG <-
-  modeldf |>
+  utila_df |>
   dplyr::filter(!is.na(CryFreqN) & !is.na(SadFreqN) & !is.na(TantrumFreqN))
 
 ## main paper plots --------------------------------------------------------
 
 signalfreqdf <-
-  modeldf |>
+  utila_df |>
   dplyr::select(
     householdID,
     childHHid,
@@ -88,7 +88,7 @@ barplot_SignalFreq <-
 barplot_SignalFreq
 
 
-signal_subset <- modeldf[c("ConflictFreqN", "Sex", "ChildAge", "SadFreqN", "CryFreqN", "TantrumFreqN", "SignalFreq", "SignalCost", "AlloparentingFreqN", "NeighborhoodQuality")]
+signal_subset <- utila_df[c("ConflictFreqN", "Sex", "ChildAge", "SadFreqN", "CryFreqN", "TantrumFreqN", "SignalFreq", "SignalCost", "AlloparentingFreqN", "NeighborhoodQuality")]
 names(signal_subset) <- shortform_dict[names(signal_subset)]
 
 signal_corrplot <-
@@ -104,7 +104,7 @@ signal_corrplot
 
 
 barplot_conflict <-
-  modeldf |>
+  utila_df |>
   mutate(
     ConflictFreq2 = factor(freq_short[ConflictFreqOF], levels = c(freq_short))
   )|>
@@ -118,7 +118,7 @@ barplot_conflict <-
 barplot_conflict
 
 barplot_alloparenting <-
-  modeldf |>
+  utila_df |>
   mutate(
     AlloparentingFreq02 = factor(freq_short[AlloparentingFreq0], levels = c(freq_short))
   )|>
@@ -157,7 +157,7 @@ causes2 <-
   )
 
 cause_response0 <-
-  modeldf |>
+  utila_df |>
   left_join(causes2[c("uniqueID", "CauseType")]) |>
   dplyr::filter(!is.na(CauseType), !is.na(CaregiverResponse), !is.na(ChildAge)) |>
   mutate(
@@ -221,12 +221,12 @@ signal_alluvial_plot
 
 # PCA ---------------------------------------------------------------------
 
-out <- skim(modeldf)
+out <- skim(utila_df)
[…]

### 2024-12-01 analysis2.R

@@ -391,7 +391,7 @@ ordinal_plot2 <- function(fit, predictor, data, title){
 # Perceived need
 
 SignalVars4 <-
-  left_join(SignalVars, modeldf[c("householdID", "childHHid", "RelativeNeed3")]) |>
+  left_join(SignalVars, utila_df[c("householdID", "childHHid", "RelativeNeed3")]) |>
   relocate(RelativeNeed3, .after = 'childHHid') |>
   mutate(
     across(-c(1:3), as.numeric),
@@ -427,7 +427,7 @@ ggsave("Figures/plot_need_combined.svg", plot_need_combined, width = 12, height 
 # Relative investment
 
 SignalVars5 <-
-  left_join(SignalVars, modeldf[c("householdID", "childHHid", "RelativeMaternalInvestment2", "RelativeNeed3")]) |>
+  left_join(SignalVars, utila_df[c("householdID", "childHHid", "RelativeMaternalInvestment2", "RelativeNeed3")]) |>
   relocate(RelativeMaternalInvestment2, .after = 'childHHid') |>
   mutate(
     across(-c(1:3), as.numeric),
@@ -471,7 +471,7 @@ cause_count <- map_int(causematrix[-c(1:2)], \(x) sum(x, na.rm = T))
 cause_vars <- names(cause_count[cause_count > 5])
 
 SignalVars3 <-
-  left_join(SignalVars, modeldf[c("householdID", "childHHid", "CaregiverResponse")]) |>
+  left_join(SignalVars, utila_df[c("householdID", "childHHid", "CaregiverResponse")]) |>
   left_join(causematrix[c("householdID", "childHHid", cause_vars)]) |>
   relocate(CaregiverResponse, .after = childHHid) |>
   mutate(

### 2024-12-01 dataprep.R

@@ -1,6 +1,7 @@
 
-source("recode.R")
-source("dictionaries.R")
+library(utiladata2023) # data package
+library(tidyverse)
+library(labelled)
 
 # functions used ----------------------------------------------------------
 
@@ -271,7 +272,7 @@ d2 <-
 # dataframe for plots -----------------------------------------------------
 
 # filters based on having at least one value of SadFreqN, CryFreqN, or TantrumFreqN
-modeldf <-
+utila_df <-
   d2 |>
   dplyr::filter(!(is.na(CryFreqN) & is.na(SadFreqN) & is.na(TantrumFreqN))) |>
   dplyr::select(
@@ -344,10 +345,6 @@ modeldf <-
     ChildID
   )
 
-modeldf_FULLSIG <-
-  modeldf |>
-  dplyr::filter(!is.na(CryFreqN) & !is.na(SadFreqN) & !is.na(TantrumFreqN))
-
 # dataframes for analysis -------------------------------------------------
 
 # new vars:
@@ -400,7 +397,7 @@ signal_vars <- c(
 )
 
 SignalVars <-
-  modeldf |>
+  utila_df |>
   dplyr::filter(!(is.na(CryFreqN) & is.na(SadFreqN) & is.na(TantrumFreqN) & is.na(ConflictFreqN))) |>
   dplyr::select(
     householdID,
@@ -421,3 +418,5 @@ SignalVars <-
     AlloparentingXsex = AlloparentingFreqN * Sex,
   ) |>
   na.omit()
+
+caregiverSex <- sort(caregivers$CaregiverSex[caregivers$householdID %in% utila_df$householdID], na.last = T)

### 2024-12-02 Paper.qmd
Working toward anonymized data

@@
 source("dictionaries.R")
-source("dataprep.R")
-# source("dataprepPublic.R")
+# source("dataprep.R")
+source("dataprepPublic.R")
 source("analysis.R")
-# source("analysis2.R") # slow!
+# source("analysis2.R") # Generates many plots. Slow!
 ```
@@
 `{r} xfun::numbers_to_words(n_distinct(utila_df$householdID), cap=T)` primary caretakers (females = `{r} table(caregiverSex)[1]`, males = `{r} table(caregiverSex)[2]`, missing sex = `{r} sum(is.na(caregiverSex))`) took the survey and provided signaling data for at least one child.
-`{r} xfun::numbers_to_words(nrow(utila_df), cap=T)` children had frequency data for at least one signal, with `{r} nrow(modeldf_FULLSIG)` having data for all three.
+`{r} xfun::numbers_to_words(nrow(utila_df), cap=T)` children had frequency data for at least one signal, with `{r} nrow(utila_FULLSIG)` having data for all three.
 Among those with at least one record of signaling frequency, `{r} sex_table[1]` were female and `{r} sex_table[2]` were male.

### 2024-12-02 StatisticalDisclosureControl.R

@@ -18,26 +18,6 @@
 # sharing that data publicly in a way that could link it to individual
 # participants.
 
-
-source('dataprep.R')
-library(sdcMicro)
-
-AddedVars <-
-  d2 |>
-  dplyr::filter(!(is.na(CryFreqN) & is.na(SadFreqN) & is.na(TantrumFreqN) & is.na(ConflictFreqN))) |>
-  dplyr::select(
-    householdID,
-    childHHid,
-    RelativeNeed3,
-    RelativeMaternalInvestment2,
-    ConflictFamily:StatusConcerns,
-    CaregiverResponse
-  )
-
-StudyVars <- SignalVars |>
-  dplyr::select(-AlloparentingXsex) |>
-  left_join(AddedVars, by = c('householdID', 'childHHid'))
-
 # Definitely
 # ChildAge
 # Sex
@@ -61,53 +41,37 @@ StudyVars <- SignalVars |>
 # HouseQuality
 # LogIncome (but category based?)
 
+# Cluster ages
+
+library(sdcMicro)
+
+discretize <- function(v, n){
+  q <- quantile(v, seq(0, 1, 1/n))[-1]
+  map_int(v, \(x) min(q[x<=q]))
+}
+
+utila_df <-
+  utila_df |>
+  mutate(
+    Neighborhood2 = ifelse(is.na(Neighborhood2), c(0,1), Neighborhood2),
+    ChildAge = discretize(ChildAge, 4)
+  ) |>
+  mutate(
+    NumberOfChildren = sample(NumberOfChildren, n()),
+    number_adults = sample(number_adults, n()),
+    .by = Neighborhood2
+  )
+
 sdcObj <- createSdcObj(
-  StudyVars,
-  keyVars = c("Sex", "Neighborhood2", "ImmigrateUtila", "UserLanguage"),
-  numVars = c("ChildAge", "NumberOfChildren", "number_adults")
+  utila_df,
+  keyVars = c("ChildAge", "Sex", "Neighborhood2", "ImmigrateUtila") #,
 )
 measure_risk(sdcObj)
-
-sdcObj <- createSdcObj(
-  StudyVars,
-  keyVars = c("Sex", "Neighborhood2", "ImmigrateUtila"),
-  numVars = c("ChildAge", "NumberOfChildren", "number_adults")
-)
-measure_risk(sdcObj)
-
-sdcObj <- createSdcObj(
-  StudyVars,
[…]

### 2024-12-02 analysis.R

@@ -19,9 +19,67 @@ library(tidygraph)
 library(ggraph)
 library(ordinalNet)
 
-modeldf_FULLSIG <-
+utila_FULLSIG <-
   utila_df |>
   dplyr::filter(!is.na(CryFreqN) & !is.na(SadFreqN) & !is.na(TantrumFreqN))
+
+signal_vars <- c(
+  ChildAge = "Younger children have more needs they cannot address on their own and face lower reputational costs to crying. Older children engage in costlier signals.",
+  Sex = "The sexes differ in the adversity they face, the forms of competetion they will engage in with peers, their ability to bargain for increased support outside of signals of need, and the reputational costs of signaling need. These differences may be small or nonexistent in young children.",
+  IllnessSusceptibilityMean = "Children who are sick more require more energy to make up for the costs of the illness*immune response interaction. They may also benefit from increased investment if this has the potential to increase immune function.",
+  MedicalProblemsMean = "Children who are sick more require more energy to make up for the costs of the illness*immune response interaction. They may also benefit from increased investment if this has the potential to improve their health or increase immune function. No pathogenic health issues may limit the ability of the child to satisfy their own needs, something which can lead to greater payoffs for signaling need.",
+  AlloparentingFreqN = "Children might prefer to spend time outside of the family (e.g., to invest in their embodied capital or to find cooperative partners and maintain those relationships) and signal to reduce caregiver pressure to alloparent.",
+  CaregiverAge = "Overtime, caretakers learn how to better satsify child needs, resist child bargaining, communicate the reasons for their actions. Older moms also have lower residual reproductive value, somethign which may result in greater valuation of investment in existing children compared to younger moms.",
+  StayAtHomeMom = "Stay at home mom's are capable of providing more investment to children and see them more often.",
+  EducationLevelYears = "Greater education leads to differences in caretaking behavior and mental models of caretaking while also increasing caregiver earning potential.",
+  AdultsChildcare = "More adults helping with child care can lead to greater potential for investment. This might lead to decreased payoffs to singaling need in some circumstances. However, it may also increase signaling behaviors in others due to the presence of more signal targerts.",
[…]

### 2024-12-02 dataprep.R

@@ -361,62 +361,4 @@ utila_df <-
 
 #HomeInstabilityMean:Children in unstable homes likely face more adversity and find themselves having to signal to new people or face the challenge of previous caretakers leaving or moving away from friends.X
 
-signal_vars <- c(
-  ChildAge = "Younger children have more needs they cannot address on their own and face lower reputational costs to crying. Older children engage in costlier signals.",
-  Sex = "The sexes differ in the adversity they face, the forms of competetion they will engage in with peers, their ability to bargain for increased support outside of signals of need, and the reputational costs of signaling need. These differences may be small or nonexistent in young children.",
-  IllnessSusceptibilityMean = "Children who are sick more require more energy to make up for the costs of the illness*immune response interaction. They may also benefit from increased investment if this has the potential to increase immune function.",
-  MedicalProblemsMean = "Children who are sick more require more energy to make up for the costs of the illness*immune response interaction. They may also benefit from increased investment if this has the potential to improve their health or increase immune function. No pathogenic health issues may limit the ability of the child to satisfy their own needs, something which can lead to greater payoffs for signaling need.",
-  AlloparentingFreqN = "Children might prefer to spend time outside of the family (e.g., to invest in their embodied capital or to find cooperative partners and maintain those relationships) and signal to reduce caregiver pressure to alloparent.",
-  CaregiverAge = "Overtime, caretakers learn how to better satsify child needs, resist child bargaining, communicate the reasons for their actions. Older moms also have lower residual reproductive value, somethign which may result in greater valuation of investment in existing children compared to younger moms.",
-  StayAtHomeMom = "Stay at home mom's are capable of providing more investment to children and see them more often.",
-  EducationLevelYears = "Greater education leads to differences in caretaking behavior and mental models of caretaking while also increasing caregiver earning potential.",
-  AdultsChildcare = "More adults helping with child care can lead to greater potential for investment. This might lead to decreased payoffs to singaling need in some circumstances. However, it may also increase signaling behaviors in others due to the presence of more signal targerts.",
-  number_adults = "More adults equals more targets for signaling.",
[…]

### 2024-12-04 Paper.qmd
Getting closer to anonymized data

@@
 ```{r setup, include=FALSE}
+
+if(!dir.exists("Figures")) dir.create("Figures")
+
 source("recode.R")
 source("dictionaries.R")
-# source("dataprep.R")
-source("dataprepPublic.R")
+
+# To use public data, comment out dataprep.R 
+# and uncomment dataprepPublic.R
+source("dataprep.R")
+# source("dataprepPublic.R")
+
 source("analysis.R")
-# source("analysis2.R") # Generates many plots. Slow!
+source("analysis2.R") # Generates many plots. Slow!
 ```
@@
 However, the full range of potential child signals of need, which extend beyond cries, has rarely been investigated in a single study of children of all ages.
-To fill this gap, we collected mother and other primary caregiver reports of child signaling from `{r} n_distinct(utila_df$householdID)` families with `{r} nrow(utila_df)` children on Utila, a small island off the coast of Honduras.
+To fill this gap, we collected mother and other primary caregiver reports of child signaling from `{r} n_distinct(householdIDs)` families with `{r} nrow(utila_df)` children on Utila, a small island off the coast of Honduras.
 In exploratory analyses, we found that child signaling was common in both sexes and across all ages, although it decreased with age and neighborhood quality and increased with the frequency of conflict between children and caretakers.
@@
 
-`{r} xfun::numbers_to_words(n_distinct(utila_df$householdID), cap=T)` primary caretakers (females = `{r} table(caregiverSex)[1]`, males = `{r} table(caregiverSex)[2]`, missing sex = `{r} sum(is.na(caregiverSex))`) took the survey and provided signaling data for at least one child.
+`{r} xfun::numbers_to_words(n_distinct(householdIDs), cap=T)` primary caretakers (females = `{r} table(caregiverSex)[1]`, males = `{r} table(caregiverSex)[2]`, missing sex = `{r} sum(is.na(caregiverSex))`) took the survey and provided signaling data for at least one child.
 `{r} xfun::numbers_to_words(nrow(utila_df), cap=T)` children had frequency data for at least one signal, with `{r} nrow(utila_FULLSIG)` having data for all three.

### 2024-12-04 StatisticalDisclosureControl.R

@@ -41,25 +41,45 @@
 # HouseQuality
 # LogIncome (but category based?)
 
-# Cluster ages
+# Run in a fresh R session
+
+source("recode.R")
+source("dictionaries.R")
+source("dataprep.R")
 
 library(sdcMicro)
 
+# Cluster ages
 discretize <- function(v, n){
   q <- quantile(v, seq(0, 1, 1/n))[-1]
   map_int(v, \(x) min(q[x<=q]))
 }
 
+new_hid <- sample(nrow(utila_df))
+names(new_hid) <- utila_df$householdID
+
 utila_df <-
   utila_df |>
   mutate(
+    householdID = new_hid,
+    childHHid = householdID,
+    uniqueID = householdID,
     Neighborhood2 = ifelse(is.na(Neighborhood2), c(0,1), Neighborhood2),
     ChildAge = discretize(ChildAge, 4)
   ) |>
+  arrange(householdID) |>
   mutate(
     NumberOfChildren = sample(NumberOfChildren, n()),
     number_adults = sample(number_adults, n()),
     .by = Neighborhood2
+  )
+
+causes <-
+  causes |>
+  mutate(
+    householdID = new_hid[householdID],
+    childHHid = householdID,
+    uniqueID = householdID
   )
 
 sdcObj <- createSdcObj(
@@ -73,5 +93,5 @@ plot(out)
 
 utila_df$ChildAge <- out@manipKeyVars$ChildAge
 
-save(utila_df, causes, caregiverSex, file = "data/utilasignalingData.rda")
+save(utila_df, causes, householdIDs, caregiverSex, file = "data/utilasignalingData.rda")
 

### 2024-12-04 analysis2.R

@@ -520,7 +520,7 @@ SignalVars6 <- SignalVars5 |>
 # code breaks when inserting 550:552 in here
 plot_fullcorrmat <- ggcorrplot(
   cor(
-    SignalVars6|>
+    SignalVars6 |>
       set_names(nm = shortform_dict[names(SignalVars6)])
     ,
     use = 'pairwise.complete.obs'),

### 2024-12-04 dataprep.R

@@ -3,6 +3,8 @@ library(utiladata2023) # data package
 library(tidyverse)
 library(labelled)
 
+load("data/m3_1e6_cost_freq.rda")
+
 # functions used ----------------------------------------------------------
 
 # function to invert the scoring on a Likert scale
@@ -327,6 +329,7 @@ utila_df <-
     PartnerStatus,
     RelativeNeed,
     RelativeMaternalInvestment,
+    RelativeMaternalInvestment2,
     Neighborhood2F,
     StayAtHomeMom,
     StayAtHomeMomF,
@@ -361,4 +364,5 @@ utila_df <-
 
 #HomeInstabilityMean:Children in unstable homes likely face more adversity and find themselves having to signal to new people or face the challenge of previous caretakers leaving or moving away from friends.X
 
+householdIDs <- sample(utila_df$householdID, length(utila_df$householdID))
 caregiverSex <- sort(caregivers$CaregiverSex[caregivers$householdID %in% utila_df$householdID], na.last = T)

### 2024-12-04 dictionaries.R

@@ -16,7 +16,7 @@ longform_dict <-
     MeanChildRelatedness = "Child relatedness",
     number_adults = "Number of adults in household",
     AdultsChildcare = "Household adults who provide childcare",
-    AdultsHousework = "Household adults who provide, food, housework, and othe neccesities",
+    AdultsHousework = "Household adults who provide, food, housework, and other neccesities",
     EducationLevelYears = "Caregiver education level (years)",
     IncomeCategoryN = "Monthly household income (Lempira)",
     FoodSecurity = "Household food insecurity",

### 2024-12-04 Paper.qmd
Incorporation of co-author comments and minor typo fixes. My final commit before the major restructure.

--- 2024-12-01 Paper.qmd
@@
   warning: false
+  keep-md: true
 format:
@@
     number-sections: true
+    prefer-html: true
   html:
@@
     number-sections: true
-    embed-resources: true
-lightbox: true
+lightbox: false
 editor: visual
@@
 ```{r setup, include=FALSE}
-source("recode.R")
-source("dictionaries.R")
-source("dataprep.R")
-# source("dataprepPublic.R")
-source("analysis.R")
-# source("analysis2.R") # slow!
+source("dictionaries.R", local = knitr::knit_global())
+source("analysisforpaper.R", local = knitr::knit_global())
+
+library(modelsummary)
+library(english)
+# source("regularize.R")
 ```
@@
 However, the full range of potential child signals of need, which extend beyond cries, has rarely been investigated in a single study of children of all ages.
-To fill this gap, we collected mother and other primary caregiver reports of child signaling from `{r} n_distinct(utila_df$householdID)` families with `{r} nrow(utila_df)` children on Utila, a small island off the coast of Honduras.
+To fill this gap, we collected mother and other primary caregiver reports of child signaling from `{r} nrow(caregivers)` families with `{r} nrow(modeldf)` children on Utila, a small island off the coast of Honduras.
 In exploratory analyses, we found that child signaling was common in both sexes and across all ages, although it decreased with age and neighborhood quality and increased with the frequency of conflict between children and caretakers.
@@
 Children’s potential to invest in their families can also cause conflicts with their parents.
-Older children might prefer to spend their time investing in romantic or other social relationships, or learning valuable skills, whereas parents might prefer them to care for their siblings, acquire or prepare food, or complete other household chores [@lowe2003; @syme2023a].
+Older children might prefer to spend their time investing in romantic or other social relationships, or learning valuable skills, whereas parents might prefer them to care for their siblings, acquire or prepare food, or complete other household chores [@lowe2003; @syme2023].
 Conflicts of interest over investment in the family also appear to extend to when a child starts having their own children [@kushnick2023; @emlen1995], with parents potentially benefiting from delaying the reproductive careers of certain offspring in order to gain more allocare for their other children [@flinn1989; @emlen1992] or speeding up a child’s departure from the household in order to provide more parental investment to other children [@kushnick2023].
@@
 Instead, caregivers are expected to weigh the costs and benefits of support against other options [@daly1987; @daly1983].
-Ancestrally, this likely often meant evaluating the viability of the child as well as how much a particular resource benefited one child relative to other children, current or future[@clutton-brock1991; @daly1983; @dawkins1976; @hagen1999].
+Ancestrally, this likely often meant evaluating the viability of the child as well as how much a particular resource benefited one child relative to other children, current or future [@clutton-brock1991; @daly1983; @dawkins1976; @hagen1999].
 The former is thought to have resulted in children having to demonstrate their vigor early in life [@soltis2004], while the latter is expected to result in context-specific support [@beaulieu2008; @frankenhuis2022].
@@
 Crying is the most studied way children signal need.
-Crying is a human universal [@gracanin2018] with few suspected non-signaling functions outside of the role vocal cries might play in the development of vocalization [@wells2003] and the functions tears have in terms of cleaning, protecting, and lubricating the eye [@murube2009; @sznycer2023].
+Crying is a human universal [@gracanin2018] with few suspected non-signaling functions outside of the role vocal cries might play in the development of vocalization [@wells2003] and the functions tears have in cleaning, protecting, and lubricating the eye [@murube2009; @sznycer2023].
 Hunger, pain, anger, fear, and separation are all common causes of crying [@vingerhoets2013; @wells2003], although tear production also often happens with joyful events [@sznycer2023; @vingerhoets2013; @zickfeld].
@@
 
-Sadness is a type of negative affect that is generally thought to arise from loss [@horwitz2017; @levenson1999; @smith1993], especially when one has little power to reverse or mitigate the loss on one’s own [@emanuel2023; @gunderson2021].
+Sadness is a type of negative affect that is generally thought to arise from loss [@horwitz2017; @levenson1999; @smith1993], especially when one has little power to reverse or mitigate the loss without help [@emanuel2023; @gunderson2021].
 Sadness and low mood, like other forms of psychological pain, probably evolved to motivate victims of adversity to shift their attention to the causes of adversity to mitigate its negative fitness consequences and to learn to solve or avoid future such adverse events [@andrews2009; @nesse1990; @thornhill1989].
@@
 Tantrums have been reported in many societies with different forms of subsistence systems, political organization, and norms surrounding anger and parenting [@albino1956; @banisalameh2021; @fouts2005; @henderson1966; @hilger1957; @konner1974; @ward1970; @johnson2003; @bhatia1990; @myers1988; @dubois @konner1972], suggesting they might be a human universal [@potegal2019].
-As with the other proposed signals of need, they are capable of eliciting a wide range of responses, including pro-child responses, neutral responses (e.g., parents ignoring the child) [@hilger1957; @konner1974; @johnson2003], and negative responses like abuse [@schmitt1987].
+As with the other proposed signals of need, they are capable of eliciting a wide range of responses, including pro-child responses (e.g., parents giving in to a child’s wishes), neutral responses [e.g., parents ignoring the child, @hilger1957; @konner1974; @johnson2003], and negative responses like abuse [@schmitt1987].
 
@@
 Similarly, young chimpanzees have been observed to beat their heads on the ground when refused access to breast milk [@goodall1986], and American white pelican chicks have been found to throw themselves on the ground, flail violently, and bite at their own wings when begging for food [@schaller1964].
-Such behaviors s could credibly signal need as they are likely relatively less costly to those who are truly in dire need compared to those who are not, thus deterring deceptive signaling [@hagen2008].
+Such behaviors could credibly signal need as they are likely relatively less costly to those who are truly in dire need compared to those who are not, thus deterring deceptive signaling [@hagen2008].
 
@@
 Although infant signals of need, particularly infant cries, have been extensively studied in many populations [@barr1991; @fouts2005; @konner1974; @kruger2010; @kushnick2006; @meehan2017; @müller2021; @tronick1987], the full range of potential child signals of need, which likely extend beyond cries, has never been investigated in children of all ages.
-In particular, how and when older children signal need has rarely been investigated [for one exception, see @syme2023a].
+In particular, how and when older children signal need has rarely been investigated [for one exception, see @syme2023].
 
@@
 
-This study of children aged `{r} paste(range(utila_df$ChildAge), collapse="-")` and their parents and other primary caretakers, was conducted in Utila, Honduras, one of the country’s small Bay Islands in the Western Caribbean.
-Participants in the study were sampled from two populations who face different levels of economic and social adversity: 1) a long-standing English-speaking community primarily comprising descendants of British immigrants who arrived on the island in the 1800’s, and 2) a more recent community of Spanish-speakers, mostly from the mainland.
-Compared to those born on Utila, immigrants from the mainland [@davis2014] are more likely to live in crowded and unsanitary conditions, face active discrimination [@garcia2017], and experience economic uncertainty due the predominance of low paying, seasonal jobs primarily related to the island’s diving tourism industry [@garcia2018; @stonich2000].
-Such inequality appears to be increasing in response to recent growth in the tourism industry, with the majority of wealth from tourism being enjoyed by a small portion of native English-speaking islanders and other Westerners [@davis2014].
+This study of children aged `{r} paste(range(modeldf$ChildAge), collapse="-")` and their parents and other primary caretakers, was conducted in Utila, Honduras, one of the country’s small Bay Islands in the Western Caribbean.
+Participants in the study were sampled from two populations who face different levels of economic and social adversity: 1) a long-standing English-speaking community primarily comprising descendants of British immigrants who arrived on the island in the 1800’s, and 2) a more recent community of Spanish-speakers, mostly from mainland Honduras.
+Compared to the former, the latter are more likely to live in crowded and suboptimal conditions, face active discrimination and increased stress [@garcia2017], and experience economic uncertainty due the predominance of low paying, seasonal jobs primarily related to the island’s diving tourism industry [@garcia2018; @stonich2000].
+Such inequality appears to be increasing in response to recent growth in the tourism industry, with the majority of wealth from tourism flowing to a small portion of native English-speaking islanders and other Westerners [@davis2014].
 
@@
 
-To inform development of our survey questionnaire on child signaling, focus groups of Utilan parents were held in July 2022.
+To inform development of our survey questionnaire on child signaling, focus groups of Utilian parents were held in July 2022.
 Parents indicated frequent sibling competition and various ways children might signal need.
@@
 JH recruited early participants with the help of primary informants and the use of snowball sampling.
-Participants were also recruited via Facebook posts and fliers which were sent home with children attending Utila's Methodist school.
+Participants were also recruited via Facebook posts.
 
@@
 The surveys, which included many topics outside of child signaling, were designed to last about an hour, plus 20 minutes for each additional child.
-Participants were paid \$XXXXX for the survey portion of the study.
+Participants were paid 350 lempiras (\$13.84 at the time of writing) for the survey portion of the study.
 
-This study was approved by the WSU xxxIRBINFOxxx and BU xxxIRBINFOxxx.
+This study was approved by the WSU and BU IRBs.
 
@@
 
-We used 6-point Likert scales to get caregiver reports on the frequencies of 1) various types of child signaling, such as sadness and crying, 2) conflict between the caretaker and each participating child, 3) and child allocare of other children in their household.
+We used six-point Likert scales to get caregiver reports on the frequencies of 1) various types of child signaling, such as sadness and crying, 2) conflict between the caretaker and each participating child, 3) and child allocare of other children in their household.
 See @tbl-freqvars for exact wording.
@@
 |  |  |
-|------------------|------------------------------------------------------|
+|--------------------------|----------------------------------------------|
 | Variable | Survey question wording |
@@
 
-We used 6-point Likert scales for parental reports of 1) the relative neediness of children within the household, and 2) the relative investment they provide to each participating child compared to the rest of their children.
+We used six-point Likert scales for parental reports of 1) the relative neediness of children within the household, and 2) the relative investment they provide to each participating child compared to the rest of their children.
 Both questions were framed need as involving time, effort, and resources.
@@
 We then created a *caregiver response* variable, which was coded as 1 for a positive response (e.g., helping child), 0 for a neutral response (e.g., not responding), and -1 for a negative response (e.g., punishment; all from the child's perspective).
-MG and KS also coded the causes for the presence or absence of 17 elements (e.g., whether the caused involved the child was asked to do an *unwanted task*, *punishment*, *family conflict*, and/or *discomfort, pain, injury, or illness*).
+MG and KS also coded the causes for the presence or absence of 17 elements (e.g., whether the cause involved *punishment*, *family conflict*, and/or *discomfort, pain, injury, or illness*).
 
@@
 Caretakers were asked about the frequency of their child missing school and their child's history of medical problems in general terms (i.e., the extent to which they experienced a "variety of medical issues").
-Both involved the use of five point Likert scales, the scores of which were averaged for use in analyses.
+Both involved the use of five-point Likert scales, the scores of which were averaged for use in analyses.
 See @stbl-medical-record for the wording of these variables.
 
-Caretakers were also asked three questions using a five point Likert scale about their child's susceptibility to illness.
+Caretakers were also asked three questions using a five-point Likert scale about their child's susceptibility to illness.
 We reversed the scores on two measures so that a score of 5 corresponded with the lowest susceptibility to illness before we averaged the scores of these to create a mean *illness avoidance* variable.
-In the `{r} sum((!is.na(utila_df$IllnessSusceptibility2)) & is.na(utila_df$IllnessSusceptibility3))` instance in which we only had data for 2 of the measures, we still calculated the mean but used a numerator of 2.
+In the `{r} sum((!is.na(modeldf$IllnessSusceptibility2)) & is.na(modeldf$IllnessSusceptibility3))` instance in which we only had data for two of the measures, we still calculated the mean but used a numerator of 2.
 See @stbl-healthvars for the wording of these variables.
 
-### Caretaker perceptions of house quality, neighborhood quality, home instability, and food security
+### Caretaker perceptions of home sanitation, neighborhood quality, home instability, and food security
 
-Affirmative caregiver responses to the following questions about their home were summed to create a home quality index: having an indoor bathroom, having a bathroom that is not shared with anyone outside of the household, having a flush toilet, having running water to wash hands in the bathroom, and having a home which did not flood in the last year.
+Affirmative caregiver responses to the following questions about their home were summed to create a home sanitation index: having an indoor bathroom, having a bathroom that is not shared with anyone outside of the household, having a flush toilet, having running water to wash hands in the bathroom, and having a home which did not flood in the last year.
 
-Caregiver perceptions of neighborhood quality were measured using ten questions relating to the costs and benefits associated with living in one's neighborhood.
-The mean of these values was used in analyses as a neighborhood quality index.
+Caregiver perceptions of neighborhood quality were measured using ten questions relating to the safety and suitability of their neighborhood for raising children in addition to questions explicitly asking about neighborhood quality, e.g., "Sometimes I hear about illegal drug activity in my neighborhood." The mean of these values was used in analyses as a neighborhood quality index.
 Levels of home instability were measured with four questions covering the degree to which the caregiver's family had/has to move homes and whether people move in and out of their homes frequently.
-For wording of the these items, which used five point Likert scales, see @stbl-Neighborhood-Quality and @stbl-Home-Instability.
+For wording of the these items, which used five-point Likert scales, see @stbl-Neighborhood-Quality and @stbl-Home-Instability.
 
-We used a modified version of the USDA Food Security Questionnaire [@blumberg1999] to measure food insecurity, with the most significant changes being converting of the measures to a five point Likert scale and averaging them for a food insecurity index.
+We used a modified version of the USDA Food Security Questionnaire [@blumberg1999] to measure food insecurity, with the most significant changes being converting the measures to a five-point Likert scale and averaging them for a food insecurity index.
 See @stbl-Food-Security for the wording of these variables.
@@
 These data allowed us to calculate the number of children (total, younger, and older) and adults within the household as well as the genetic relatedness between siblings.
-We also created a variable for residence in Camponado, a densely populated neighborhood on the island with relatively poor sanitation in which most individuals from the mainland live that is seen as the most dangerous and poorest part of Utila [@garcia2017].
+We also created a variable for residence in Camponado, a low-SES, densely populated neighborhood on the island with less sanitation access mostly comprising individuals from the mainland [@garcia2017].
 
@@
 
-The overall structure of our data was analyzed with principle components analysis (PCA), Bayesian graphical models (BGM), and minimum spanning trees (MST).
-PCA is a linear transformation of the data to a new orthonormal coordinate system in which the the first principle component (PC1) corresponds to the axis of maximum variance, the second principle component (PC2) corresponds to the axis perpendicular to the first of the maximum of the remaining variance, and so forth.
+The overall structure of our data was analyzed with principal components analysis (PCA), Bayesian graphical models (BGM), and minimum spanning trees (MST).
+PCA is a linear transformation of the data to a new orthonormal coordinate system in which the the first principal component (PC1) corresponds to the axis of maximum variance, the second principal component (PC2) corresponds to the axis perpendicular to the first of the maximum of the remaining variance, and so forth.
 Among other uses, PCA identifies the linear combinations of variables that account for most of the variance, thereby providing a lower dimensional representation of high dimensional data.
-In BGA, variables are nodes in an undirected graph, and edges represent their conditional dependencies.
+In BGM, variables are nodes in an undirected graph, and edges represent their conditional dependencies.
 Specifically, an edge indicates that two variables are dependent, conditional on the other variables; edge weights indicate the strength and sign of the relationship.
@@
 
-The anonymized data used in the study are available at <https://github.com/grasshoppermouse/utiladata2023>. Study code is available at <https://github.com/michaelrgaffney/utilasignaling>.
+The anonymized data used in the study are available at https://github.com/grasshoppermouse/utiladata2023. Study code is available at https://github.com/michaelrgaffney/utilasignaling.
 
@@
 ```{r}
-sex_table <- table(utila_df$Sex)
+# primary_caregivers <- str_to_sentence(english::english(sum(caregivers$HasSignalingData)))
+
+tbl <- table(caregivers$HasSignalingData, caregivers$CaregiverSex)
+primary_caregivers <- tbl[2,1] + tbl[2,2] + tbl[2,3]
 ```
 
-`{r} xfun::numbers_to_words(n_distinct(utila_df$householdID), cap=T)` primary caretakers (females = `{r} table(caregiverSex)[1]`, males = `{r} table(caregiverSex)[2]`, missing sex = `{r} sum(is.na(caregiverSex))`) took the survey and provided signaling data for at least one child.
-`{r} xfun::numbers_to_words(nrow(utila_df), cap=T)` children had frequency data for at least one signal, with `{r} nrow(modeldf_FULLSIG)` having data for all three.
+`{r} primary_caregivers` primary caregivers (females = `{r} tbl['TRUE', 'Female']`, males = `{r} tbl['TRUE', 'Male']`, non-binary = `{r} tbl['TRUE', 'Non-binary / third gender']`) took the survey and provided signaling data for at least one child.
+`{r} nrow(modeldf)` children had frequency data for at least one signal, with `{r} nrow(modeldf_FULLSIG)` having data for all three.
 Among those with at least one record of signaling frequency, `{r} sex_table[1]` were female and `{r} sex_table[2]` were male.
@@
 
-kable(custom.summarize(utila_df, longform_dict[1:23])) %>% kable_styling()
+kable(custom.summarize(modeldf, longform_dict[1:23])) #%>% kable_styling()
 ```
@@
 # convert to factor to omit NAs in datasummary_skim
-dx <- utila_df |>
+dx <- modeldf |>
   dplyr::select(any_of(names(longform_dict))) |>
@@
 ::: {#fig-PCA}
-![](Figures/plot_pca.svg)
+![](Figures/plot_pca.png)
 
@@
 See @fig-barplot_SignalFreq and @fig-signal_corrplot.
-Caregivers reported that `{r} fmt_sum(utila_df$SadFreqN >= 8)` were sad at least once a week, `{r} fmt_sum(utila_df$CryFreqN >= 8)` cried at least once a week, and `{r} fmt_sum(utila_df$TantrumFreqN >= 8)` threw temper tantrums at least once a week.
-A small subset of children engaged in these signals/states at least daily; sadness: `{r} fmt_sum(utila_df$SadFreqN >= 30)`; crying: `{r} fmt_sum(utila_df$CryFreqN >= 30)`; and tantrums: `{r} fmt_sum(utila_df$TantrumFreqN >= 30)`.
-However, many children were reported to never engage in these signals/states; sadness: `{r} fmt_sum(utila_df$SadFreqN == 0)`; crying; `{r} fmt_sum(utila_df$CryFreqN == 0)`; and tantrums: `{r} fmt_sum(utila_df$TantrumFreqN == 0)`.
+Caregivers reported that `{r} fmt_sum(modeldf$SadFreqN >= 8)` were sad at least once a week, `{r} fmt_sum(modeldf$CryFreqN >= 8)` cried at least once a week, and `{r} fmt_sum(modeldf$TantrumFreqN >= 8)` threw temper tantrums at least once a week.
+A small subset of children engaged in these signals/states at least daily; sadness: `{r} fmt_sum(modeldf$SadFreqN >= 30)`; crying: `{r} fmt_sum(modeldf$CryFreqN >= 30)`; and tantrums: `{r} fmt_sum(modeldf$TantrumFreqN >= 30)`.
+However, many children were reported to never engage in these signals/states; sadness: `{r} fmt_sum(modeldf$SadFreqN == 0)`; crying; `{r} fmt_sum(modeldf$CryFreqN == 0)`; and tantrums: `{r} fmt_sum(modeldf$TantrumFreqN == 0)`.
 For a visualization of how signal frequencies varied within children see @sfig-signal_alluvial_plot.
 
-Conflict between children and caretakers was also common, with `{r} fmt_sum(utila_df$ConflictFreqN >= 8)` having conflicts with parents more than once a week.
-However, `{r} fmt_sum(utila_df$ConflictFreqN == 0)` were described as never having conflicts with parents.
+Conflict between children and caretakers was also common, with `{r} fmt_sum(modeldf$ConflictFreqN >= 8)` having conflicts with parents more than once a week.
+However, `{r} fmt_sum(modeldf$ConflictFreqN == 0)` were described as never having conflicts with parents.
 See @fig-barplot_ConflictFreq.
@@
 
-The frequency of child alloparenting effort, which previous research has linked to parent-child conflict, was bimodally distributed, with `{r} fmt_sum(utila_df$AlloparentingFreqN == 0)` reported to never engage in it (`r sum(na.omit(utila_df$NumberOfChildren==1))` instances were due to a child being an only child) and `{r} fmt_sum(utila_df$AlloparentingFreqN >= 30)` reported to help with childcare at least once a day.
+The frequency of child alloparenting effort, which previous research has linked to parent-child conflict, was bimodally distributed, with `{r} fmt_sum(modeldf$AlloparentingFreqN == 0)` reported to never engage in it (`r sum(na.omit(modeldf$NumberOfChildren==1))` instances were due to a child being an only child) and `{r} fmt_sum(modeldf$AlloparentingFreqN >= 30)` reported to help with childcare at least once a day.
 See @fig-barplot_AlloparentingFreq.
@@
 ::: {#fig-lasso}
-![](Figures/plot_lasso.svg)
+![](Figures/plot_lasso.png)
 
@@
 ::: {#fig-signal-effects}
-![](Figures/signal_effects_plot.svg)
+![](Figures/signal_effects_plot.png)
 
@@
 ::: {#fig-alloparentingXsex}
-![](Figures/plot_alloparentingXsex.svg)
+![](Figures/plot_alloparentingXsex.png)
 
@@
 ::: {#fig-conflict-effects}
-![](Figures/conflict_effects_plot.svg)
+![](Figures/conflict_effects_plot.png)
 
@@
 
-```{r}
-rel_need_table <- table(utila_df$RelativeNeed3)
-```
-
-Primary caretakers rarely described a child as “much less” `{r} fmt_sum(utila_df$RelativeNeed == "Much less")` \[fix in word doc?\] or “much more” `{r} fmt_sum(utila_df$RelativeNeed == "Much more")` needy compared to their other children.
+Primary caretakers rarely described a child as “much less” `{r} fmt_sum(modeldf$RelativeNeed == "Much less")` \[fix in word doc?\] or “much more” `{r} fmt_sum(modeldf$RelativeNeed == "Much more")` needy compared to their other children.
 Therefore, we collapsed “less” and “much less” into a “less” factor and “more” and “much more” into a “more” factor.
@@
 ::: {#fig-effect_relativeneed}
-![](Figures/plot_need_combined.svg)
+![](Figures/plot_need_combined.png)
 
@@
 ::: {#fig-plot_relative_investment}
-![](Figures/plot_invest_need.svg)
+![](Figures/plot_invest_need.png)
 
@@
 
-```{r}
-response_overlap <- table(utila_df$PositiveResponse, utila_df$NegativeResponse)
-```
-
 There were `{r} response_overlap[1,1]` instances in which a caregiver's response involved acts which were both negative and positive from the child's perspective.
@@
 Our models found no evidence of relationships between, child age, child sex, the number of children in the household, and log family income and the valence of parental reponses (positive, neutral, or negative) from the child's perspective.
-However, primary caretakers were more likely to respond positively towards a child when they attributed the cause of signaling to discomfort, pain, injury, or illness, and less likely to respond positively when they attributed the cause of signaling to child punishment (any statistic report here?).
+However, primary caretakers were more likely to respond positively towards a child when they attributed the cause of signaling to discomfort, pain, injury, or illness, and less likely to respond positively when they attributed the cause of signaling to family conflict, child punishment, child trangressions, or a child losing an item or privleges.
 See @fig-effect_last_signal_response.
@@
 ::: {#fig-effect_last_signal_response}
-![](Figures/plot_caregiver_response_combined.svg)
+![](Figures/plot_caregiver_response_combined.png)
 
@@
 
-Despite the differences in wealth, job certainty, discrimination, and living conditions between recent immigrants from the mainland and native Utilians [@davis2014; @garcia2017; @garcia2018; @stonich2000], primary language, immigration status, and residence in Camponado did not predict child signaling in our lasso regressions.
+Despite the differences in wealth, job certainty, discrimination, and living conditions between those from the mainland and Utilians [@davis2014; @garcia2017; @garcia2018; @stonich2000], primary language, immigration status, and residence in Camponado did not predict child signaling in our lasso regressions.
 These variables correlated with neighborhood quality, however, which was retained in our models, and lasso regression tends to retain only one of a set of correlated predictors [@zou2005].
@@
 For example, while the energetic cost of crying might be costly for the youngest children, it is likely insignificant for older children.
-Older children might instead need to weigh the reputational harm from crying against the potential increase in social support.
+Older children might instead need to weigh the reputational harm from crying [@citexxx] against the potential increase in social support.
 
@@
 Although conflicts of interest do not always lead to overt behavioral conflict [@altmann1980; @bateson1994], it is likely many, if not most, conflicts are ultimately caused by conflicts of interest.
-Conflicts, in turn, were negatively associated with child age and higher neigborhood quality, and positively associated with male sex, medical problems, and (oddly) caregiver education [@fig-conflict-effects].
+Conflicts, in turn, were negatively associated with child age and higher neighborhood quality, and positively associated with male sex, medical problems, and (oddly) caregiver education [@fig-conflict-effects].
 
@@
 conflicts of interest between child and caretaker, some of which might represent caretaker investment in some children more than others, and less caretaker uncertainty about need in cases of injury and illness.
-That said, our model predicts positive responses in 35% of cases involving trangressions, suggesting that children are successful in negotiating better outcomes at least some of the time.
+That said, our model predicts positive responses in 35% of cases involving transgressions, suggesting that children are successful in negotiating better outcomes at least some of the time.
 
@@
 
-Interestingly, older children were perceived to be needier than younger ones [\@ @fig-effect_relativeneed].
+Interestingly, older children were perceived to be needier than younger ones [@fig-effect_relativeneed].
 One likely reason for this might be that caretakers perceive the types of investment older children seek as more costly.
@@
 
-Finally, based on previous research showing that family demands, such as caring for younger siblings, was an important source of parent-offspring conflict [@syme2019], we investigated the relationship between the frequency of alloparenting by children and their frequency of signaling.
+Finally, based on previous research showing that family demands, such as caring for younger siblings, was an important source of parent-offspring conflict [@syme2023], we investigated the relationship between the frequency of alloparenting by children and their frequency of signaling.
 Although there was little evidence for sex differences in signaling overall (perhaps because sex differences typically emerge in adolescence, and our sample of adolescents was small), there was evidence that sex interacted with alloparenting effort.
@@
 In many societies, girls do more alloparenting than boys [@helfrecht2016; @weisner1977; @barry1971; @turke1988], which matches reports from our focus groups of Utilian mothers.
-It is plausible that older girls are pressured to do more care of their younger siblings, which in other studies was found to spark parent-offspring conflict [@lowe2003; syme2023a].
+It is plausible that older girls are pressured to do more care of their younger siblings, in line with other studies finding an association between household responsibilities, like caregiving, and parent-offspring conflict [@lowe2003; @syme2023].
 If so, girls with excessive alloparenting responsibilities might be signaling their need for more time to invest in peer relationships and status seeking behaviors outside the household.
-Opposed to this interpretation, our study found a small negative assocation between girls' alloparenting and caregiver reports of conflict [@fig-lasso], although perhaps when girls do what caregivers want, caregivers report less conflict.
+Opposed to this interpretation, our study found a small negative association between girls' alloparenting and caregiver reports of conflict [@fig-lasso], although plausibly when girls do what caregivers want, caregivers report less conflict.
 
 Many models designed to explain signals of need (e.g., the begging of young birds) predict that offspring who are worse off or are receiving less investment have more to gain from signaling than those with more [@mock2011; @godfray1991; @caro2016a; @godfray2000] and multiple game theoretical models examining parent-offspring conflict have found that parental resources have a modest impact on the evolutionary stable value for offspring demand for resources [@bossan2013; @mock1997 [confirm]].
-In line with this reasoning, children in lower income households and riskier environments might benefit more from signaling effort if they experience greater levels of adversity or perceive themselves as disadvantaged compared to other children they may be competing with for resources, status, and valuable social relationships.
+In line with this reasoning, children in lower income households and riskier environments might benefit more from signaling effort if they experience greater levels of adversity or perceive themselves as disadvantaged compared to other children.
 
@@
 Our study was also entirely cross-sectional, further limiting our ability to make causal inferences.
+Furthermore, our penalized regressions included many variables, some of which might have been colliders, inducing spurious associations between outcome and predictor variables.
 
@@
 Children face the challenge of eliciting support from others in times of adversity and opportunity.
-The signaling strategies of Utilian children were highly correlated and engaged in by children of all ages, albeit more frequently by younger children and those living in more adverse conditions, who plausibly had greater need.
+On the island of Utila, child signaling strategies were highly correlated and engaged in by children of all ages, albeit more frequently by younger children and those living in more adverse conditions, who plausibly had greater need.
 Consistent with signaling models that emphasize the role of conflicts of interest in favoring the use of costlier signals, child signaling effort was positively correlated with the frequency of conflict with caretakers and with perceptions of need, and greater perceptions of child need predicted greater investment in a given child.
@@
 |  |  |
-|----------------|--------------------------------------------------------|
+|--------------------------|----------------------------------------------|
 | Variable | Survey question wording |
@@
 |  |  |
-|---------------|---------------------------------------------------------|
+|-------------------------|------------------------------------------------|
 | Variable | Survey question wording |
@@
 |  |  |
-|-------------|-----------------------------------------------------------|
+|-------------------------|------------------------------------------------|
 | Variable | Survey question wording |
@@
 |  |  |
-|------------------|------------------------------------------------------|
+|-------------------------|-----------------------------------------------|
 | Variable | Survey question wording |
@@
 |  |  |
-|-------------------|-----------------------------------------------------|
+|--------------------------|----------------------------------------------|
 | Variable | Survey question wording |
@@
 
-List of options for research assistant to classify participant responses to the question of how they responded when their child was last sad or cried.
+List of options for research assistants to classify participant responses to the question of how they responded when their child was last sad or cried.
 :::
@@
 ::: {#sfig-plot-full-corrmat}
-![](Figures/plot_fullcorrmat.svg)
+![](Figures/plot_fullcorrmat.png)
 
@@
 ::: {#sfig-plot-caregiver-punish}
-![](Figures/plot_caregiver_punish.svg)
+![](Figures/plot_caregiver_punish.png)
 
@@
 ::: {#sfig-plot-need-coefs}
-![](Figures/plot_need_coefs.svg)
+![](Figures/plot_need_coefs.png)
 
@@
 ::: {#sfig-plot-invest-coefs}
-![](Figures/plot_invest_coefs.svg)
+![](Figures/plot_invest_coefs.png)
 
@@
 ::: {#sfig-plot-caregiverresponse-coefs}
-![](Figures/plot_caregiverresponse_coefs.svg)
+![](Figures/plot_caregiverresponse_coefs.png)
 
@@
 :::
-
-# Deleted sections
-
-### Child anthropometrics
-
-xxxxxxxxxxDescribe camps brieflyxxxxxxxxxxxxxx - include dates and how we handled those.
-For cross-sectional analyses, we used the 2023 anthropometric values due to them being closer in time to the caretaker surveys.
-
-All anthropometric measurements were taken twice, in accordance with anthropometric standards [@lohman1988], and then averaged.
-Before being measured, participants were asked to remove their shoes and heavy articles of clothing.
-Height was measured to the nearest millimeter using an anthropometer, and weight was measured to the nearest 100 g with a digital scale (model name) that was used in a building with a concrete foundation.
-All participants were wearing lightweight clothing, so we made no adjustments for clothing weight.
-Grip strength was measured using a dynamometer (model name), with participants seated/standing?
-and told to xxxxxxxx.
-XXXX time was passed between the first and second measurements.
-
-Triceps and subscapular skinfold thicknesses were measured using skinfold calipers.
-The triceps measurement was taken at the midpoint of the upper arm.
-The subscapular measurement occurred immediately below the inferior angle of the scapula, being picked up diagonally at the natural fold line of the skin [@antóna].
-Flexed bicep circumference was measured perpendicular to the long axis of the upper arm.
-For this measure participants were asked to sit/stand?
-and position their upper arm perpendicular to their body with their elbow being held at a 90° angle, before flexing their clenched fist towards themselves as much as they could.
-
-We computed each child's height, weight, and BMI Z-scores with respect to the CDC child growth standards [@kuczmarski2002].
-
-For our independent variables of skinfold thickness, flexed bicep circumference, and grip strength on signaling behavior, we used the residuals from generalized additive models which modeled the effect of sex and age on each measurement using the mgcv package [@wood2011].
-The resulting variables represent how each child differed from the average child for their given sex and age.
-We also centered and scaled the skinfold thickness residuals before summing them to create a body fat index.
-
-### Old discussion parts
-
-The effect of age was smallest on child sadness.
-However, interpretation of this finding is more difficult due to the fact that, unlike crying and temper tantrums, sadness is thought to have important non-signaling functions including drawing one’s attention to a problem, motivating deeper analysis of one’s situation, and causing withdrawal from situations with low fitness prospects [@nesse2019; @nettle2009; @allen2003; @hagen2011].
-By asking about sadness as a whole, our measure was unable to disentangle information caretakers gleaned from cues from information that came from signals.

### 2024-12-04 dictionaries.R

@@ -16,12 +16,12 @@ longform_dict <-
     MeanChildRelatedness = "Child relatedness",
     number_adults = "Number of adults in household",
     AdultsChildcare = "Household adults who provide childcare",
-    AdultsHousework = "Household adults who provide, food, housework, and othe neccesities",
+    AdultsHousework = "Household adults who provide, food, housework, and other neccesities",
     EducationLevelYears = "Caregiver education level (years)",
     IncomeCategoryN = "Monthly household income (Lempira)",
     FoodSecurity = "Household food insecurity",
     NeighborhoodQuality = "Neighborhood quality index",
-    HouseQuality = "House quality index",
+    HouseQuality = "House sanitation index",
     MedicalProblemsMean = "Child medical problems index",
     IllnessSusceptibilityMean = "Child illness avoidance index",
     HeightZ = "Height Z-score (CDC)",

### 2024-12-04 regularize.R

@@ -75,7 +75,7 @@ plot_lasso <-
   plot_layout(axis_titles = 'collect', ncol = 2)
 plot_lasso
 ggsave("Figures/plot_lasso.pdf", plot_lasso, width = 12, height = 12)
-ggsave("Figures/plot_lasso.svg", plot_lasso, width = 12, height = 12)
+ggsave("Figures/plot_lasso.png", plot_lasso, width = 12, height = 12)
 
 plotpredictions2 <- function(params, cond, title = NULL, plot = T){
 
@@ -113,7 +113,7 @@ signal_effects_plot &
   scico::scale_color_scico_d(palette = 'roma', begin = 0.1, end = 0.8)
 
 ggsave("Figures/signal_effects_plot.pdf", signal_effects_plot, width = 12, height = 9)
-ggsave("Figures/signal_effects_plot.svg", signal_effects_plot, width = 12, height = 9)
+ggsave("Figures/signal_effects_plot.png", signal_effects_plot, width = 12, height = 9)
 
 
 # Alloparenting special case (SignalCost)
@@ -172,7 +172,7 @@ plot_alloparentingXsex <- plot_alloparenting_cost / plot_alloparenting_cost_tant
 plot_alloparentingXsex
 
 ggsave("Figures/plot_alloparentingXsex.pdf", plot_alloparentingXsex, width = 12, height = 9)
-ggsave("Figures/plot_alloparentingXsex.svg", plot_alloparentingXsex, width = 12, height = 9)
+ggsave("Figures/plot_alloparentingXsex.png", plot_alloparentingXsex, width = 12, height = 9)
 
 # Conflict effects plot
 
@@ -202,7 +202,7 @@ conflict_effects_plot <-
 conflict_effects_plot
 
 ggsave("Figures/conflict_effects_plot.pdf", conflict_effects_plot, width = 12, height = 9)
-ggsave("Figures/conflict_effects_plot.svg", conflict_effects_plot, width = 12, height = 9)
+ggsave("Figures/conflict_effects_plot.png", conflict_effects_plot, width = 12, height = 9)
 
 # Need to plot this by sex
 # out <- plotpredictions2(signalparams, c('AlloparentingXsex', 'Sex'), plot = F)
@@ -230,7 +230,7 @@ plot_pca <- plot_loadings1 + plot_biplot1 + plot_layout(widths = c(1,2))
 plot_pca
 
[…]

### 2024-12-04 Paper.qmd

@@
     number-sections: true
-lightbox: false
+    embed-resources: true
+lightbox: true
 editor: visual
@@
 ```{r setup, include=FALSE}
-source("dictionaries.R", local = knitr::knit_global())
-source("analysisforpaper.R", local = knitr::knit_global())
+if(!dir.exists("Figures")) dir.create("Figures")
 
-library(modelsummary)
-library(english)
-# source("regularize.R")
+source("recode.R")
+source("dictionaries.R")
+
+# To use public data, comment out dataprep.R
+# and uncomment dataprepPublic.R
+source("dataprep.R")
+# source("dataprepPublic.R")
+
+source("analysis.R")
+source("analysis2.R") # Generates many plots. Slow!
 ```
@@
 However, the full range of potential child signals of need, which extend beyond cries, has rarely been investigated in a single study of children of all ages.
-To fill this gap, we collected mother and other primary caregiver reports of child signaling from `{r} nrow(caregivers)` families with `{r} nrow(modeldf)` children on Utila, a small island off the coast of Honduras.
+To fill this gap, we collected mother and other primary caregiver reports of child signaling from `{r} n_distinct(householdIDs)` families with `{r} nrow(utila_df)` children on Utila, a small island off the coast of Honduras.
 In exploratory analyses, we found that child signaling was common in both sexes and across all ages, although it decreased with age and neighborhood quality and increased with the frequency of conflict between children and caretakers.
@@
 
-This study of children aged `{r} paste(range(modeldf$ChildAge), collapse="-")` and their parents and other primary caretakers, was conducted in Utila, Honduras, one of the country’s small Bay Islands in the Western Caribbean.
+This study of children aged `{r} paste(range(utila_df$ChildAge), collapse="-")` and their parents and other primary caretakers, was conducted in Utila, Honduras, one of the country’s small Bay Islands in the Western Caribbean.
 Participants in the study were sampled from two populations who face different levels of economic and social adversity: 1) a long-standing English-speaking community primarily comprising descendants of British immigrants who arrived on the island in the 1800’s, and 2) a more recent community of Spanish-speakers, mostly from mainland Honduras.
@@
 
-Caretakers were also asked three questions using a five-point Likert scale about their child's susceptibility to illness.
+Caretakers were also asked three questions using a five point Likert scale about their child's susceptibility to illness.
 We reversed the scores on two measures so that a score of 5 corresponded with the lowest susceptibility to illness before we averaged the scores of these to create a mean *illness avoidance* variable.
-In the `{r} sum((!is.na(modeldf$IllnessSusceptibility2)) & is.na(modeldf$IllnessSusceptibility3))` instance in which we only had data for two of the measures, we still calculated the mean but used a numerator of 2.
+In the `{r} sum((!is.na(utila_df$IllnessSusceptibility2)) & is.na(utila_df$IllnessSusceptibility3))` instance in which we only had data for 2 of the measures, we still calculated the mean but used a numerator of 2.
 See @stbl-healthvars for the wording of these variables.
@@
 ```{r}
-# primary_caregivers <- str_to_sentence(english::english(sum(caregivers$HasSignalingData)))
-
-tbl <- table(caregivers$HasSignalingData, caregivers$CaregiverSex)
-primary_caregivers <- tbl[2,1] + tbl[2,2] + tbl[2,3]
+sex_table <- table(utila_df$Sex)
 ```
 
-`{r} primary_caregivers` primary caregivers (females = `{r} tbl['TRUE', 'Female']`, males = `{r} tbl['TRUE', 'Male']`, non-binary = `{r} tbl['TRUE', 'Non-binary / third gender']`) took the survey and provided signaling data for at least one child.
-`{r} nrow(modeldf)` children had frequency data for at least one signal, with `{r} nrow(modeldf_FULLSIG)` having data for all three.
+`{r} xfun::numbers_to_words(n_distinct(householdIDs), cap=T)` primary caretakers (females = `{r} table(caregiverSex)[1]`, males = `{r} table(caregiverSex)[2]`, missing sex = `{r} sum(is.na(caregiverSex))`) took the survey and provided signaling data for at least one child.
+`{r} xfun::numbers_to_words(nrow(utila_df), cap=T)` children had frequency data for at least one signal, with `{r} nrow(utila_FULLSIG)` having data for all three.
 Among those with at least one record of signaling frequency, `{r} sex_table[1]` were female and `{r} sex_table[2]` were male.
@@
 
-kable(custom.summarize(modeldf, longform_dict[1:23])) #%>% kable_styling()
+kable(custom.summarize(utila_df, longform_dict[1:23])) %>% kable_styling()
 ```
@@
 
-Primary caretakers rarely described a child as “much less” `{r} fmt_sum(modeldf$RelativeNeed == "Much less")` \[fix in word doc?\] or “much more” `{r} fmt_sum(modeldf$RelativeNeed == "Much more")` needy compared to their other children.
+```{r}
+rel_need_table <- table(utila_df$RelativeNeed3)
+```
+
+Primary caretakers rarely described a child as “much less” `{r} fmt_sum(utila_df$RelativeNeed == "Much less")` \[fix in word doc?\] or “much more” `{r} fmt_sum(utila_df$RelativeNeed == "Much more")` needy compared to their other children.
 Therefore, we collapsed “less” and “much less” into a “less” factor and “more” and “much more” into a “more” factor.
@@
 :::
+
+```{r}
+response_overlap <- table(utila_df$PositiveResponse, utila_df$NegativeResponse)
+```
 

### 2024-12-05 Paper.qmd
Added sessionInfo() at the end
deleted old text
removed editor: visual

@@
 lightbox: true
-editor: visual
 tbl-cap-location: bottom
@@
 :::
+
+```{r}
+sessionInfo()
+```
+
+# References

### 2024-12-05 analysis2.R
simplified code for using public data

@@ -242,6 +242,7 @@ names(discrete_vars) <- names(discrete_vars0$discrete)
 # set.seed(456)
 # m3 <- easybgm(SignalVars[-c(1,2,3,4,5)], type = 'mixed', not_cont = discrete_vars[-c(1,2,3,4,5)], iter = 1000000, package = 'BDgraph', g.prior = 0.05)
 # save(m3, file = 'm3_1e6_cost_freq.rda')
+load("data/m3_1e6_cost_freq.rda")
 
 plot_network(m3, layout = "spring", vsize = 4, label.cex = 1, exc_prob = 0.9)
 plot_structure(m3, vsize = 4, label.cex = 1)

### 2024-12-05 dataprepPublic.R

@@ -1,4 +1,0 @@
-
-load("data/utilasignalingPublicData.rda")
-load("data/m3_1e6_cost_freq.rda")
-

### 2024-12-05 Paper.qmd
simplified code for using public data
reducing dependencies

@@
 # To use public data, comment out dataprep.R
-# and uncomment dataprepPublic.R
+# and uncomment load statement
 source("dataprep.R")
-# source("dataprepPublic.R")
+# load("data/utilasignalingPublicData.rda")
 
 source("analysis.R")
-source("analysis2.R") # Generates many plots. Slow!
+# source("analysis2.R") # Generates many plots. Slow!
 ```
@@
 
-kable(custom.summarize(utila_df, longform_dict[1:23])) %>% kable_styling()
+utila_df |> 
+  dplyr::select(any_of(names(utila_df[names(longform_dict[1:23])]))) |> 
+  setNames(longform_dict[1:23]) |> 
+  datasummary_skim(fun_numeric = list(N = \(x) length(na.omit(x)), Range = \(x) str_glue("{min(x, na.rm=T)}-{max(x, na.rm=T)}"), Mean = Mean, SD = SD, Histogram = function(x) ""))
 ```

### 2024-12-05 analysis.R

@@ -1,14 +1,11 @@
 
-library(tidyverse)
+library(ggplot2)
+library(forcats)
 library(patchwork)
 library(modelsummary)
-library(labelled)
 library(ggcorrplot)
 library(pvclust)
 library(ggalluvial)
-library(hagenutils) # from github
-library(knitr)
-library(kableExtra)
 library(skimr)
 library(glmnet)
 library(tidymodels)
@@ -18,6 +15,8 @@ library(easybgm)
 library(tidygraph)
 library(ggraph)
 library(ordinalNet)
+
+source("pca_plot_functions.R")
 
 utila_FULLSIG <-
   utila_df |>

### 2024-12-05 dataprep.R

@@ -1,6 +1,8 @@
 
-library(utiladata2023) # data package
-library(tidyverse)
+# library(utiladata2023) # data package
+library(dplyr)
+library(purrr)
+library(stringr)
 library(labelled)
 
 load("data/m3_1e6_cost_freq.rda")

### 2024-12-05 pca_plot_functions.R

@@ -1,0 +1,89 @@
+pca_loadings_plot <- function(obj, components = 1:3, sortby = 1, threshold = 0, reverse=NULL){
+  require(dplyr)
+  if (!all(reverse %in% components)) stop("reverse vector must be a subset of components vector")
+  pca_summ <- summary(obj)
+  comp_nms <- colnames(pca_summ$importance)
+  varprop <- pca_summ$importance[2,components] # Get % variance
+  nms <- comp_nms[components]
+  varprop <- paste0(nms, ' (', round(varprop*100, 1), '%)')
+  names(varprop) <- nms
+
+  pcs <- colnames(obj$rotation)[components]
+  loadings <-
+    obj$rotation[,components, drop = F] %*% diag(obj$sdev[components]) %>%
+    data.frame %>%
+    setNames(pcs)
+
+  if(!is.null(reverse)) loadings[reverse] <- -loadings[reverse]
+
+  loadings <- loadings[rowSums(abs(loadings)>threshold)>0,]
+
+  loadings %>%
+    dplyr::mutate(Variable = forcats::fct_reorder(rownames(.), loadings[, sortby])) %>%
+    tidyr::gather(key = PC, value = Loading, -Variable) %>%
+    dplyr::mutate(PC = varprop[PC]) %>%
+    ggplot2::ggplot() +
+    ggplot2::geom_segment(ggplot2::aes(x=0, xend=Loading, y=Variable, yend=Variable, colour = Loading), size=2) +
+    ggplot2::geom_point(ggplot2::aes(x=Loading, y=Variable, colour = Loading), size=3) +
+    ggplot2::scale_color_gradient2(low = viridisLite::magma(11)[8], mid = 'white', 'high' = viridisLite::magma(11)[4]) +
+    ggplot2::geom_vline(xintercept = 0) +
+    ggplot2::facet_wrap(~PC) +
+    ggplot2::labs(x = '\nStandardized loadings', y = '') +
+    ggplot2::theme_bw(15)
+}
+
+pca_biplot <- function(obj, components = c(1,2), data = NULL, threshold = 0, reverse=NULL, label_size = 5, geom_point=T) {
+
+  if (class(obj) != 'prcomp') stop('obj class must be prcomp')
+  if (length(components) != 2 | mode(components) != 'numeric') stop('components is not a numeric vector of length 2')
[…]

### 2024-12-06 analysis.R
Adjusted and combined figures, ggsave.

@@ -154,11 +154,12 @@ signal_corrplot <-
     `Child sex` = as.numeric(`Child sex`)
   ) |>
   cor( use = "pairwise.complete.obs") |>
-  ggcorrplot(hc.order = TRUE, hc.method = "ward.D",lab = TRUE, lab_col = "black", lab_size = 4.5) +
+  ggcorrplot(hc.order = TRUE, hc.method = "ward.D",lab = TRUE, lab_col = "black", lab_size = 4.5, tl.cex = 16) +
   scico::scale_fill_scico(palette = "vik", midpoint = 0, begin = .1, end = .9, limits = c(-1, 1)) +
   guides(fill = guide_colorbar(title = "Correlation coefficients"))
 signal_corrplot
-
+ggsave("Figures/signal_corrplot.pdf", width = 12, height = 9)
+ggsave("Figures/signal_corrplot.png", width = 12, height = 9)
 
 barplot_conflict <-
   utila_df |>
@@ -168,6 +169,7 @@ barplot_conflict <-
   filter_at(vars(ConflictFreqOF),all_vars(!is.na(.))) |>
   ggplot(aes(x=ConflictFreq2, fill= Sex)) +
   geom_bar(position = "stack") +
+  ylim(0, 150) +
   labs(title = "Conflict") +
   scale_fill_viridis_d(option = "B", begin = 0.3, end = 0.7) +
   labs(x = "Frequency of conflict (per month)", y = "Number of children") +
@@ -182,11 +184,18 @@ barplot_alloparenting <-
   filter_at(vars(AlloparentingFreq0),all_vars(!is.na(.))) |>
   ggplot(aes(x= AlloparentingFreq02, fill = Sex)) +
   geom_bar(position = "stack") +
+  ylim(0, 150) +
   labs(title = "Alloparenting") +
   scale_fill_viridis_d(option = "B", begin = 0.3, end = 0.7) +
   labs(x = "Frequency of child alloparenting (per month)", y = "Number of children") +
   theme_minimal(15)
 barplot_alloparenting
+
+plot_conflict_allo <-
+  (barplot_conflict + barplot_alloparenting) + plot_layout(axes = 'collect_y', axis_titles = "collect_y", guides = 'collect')
+plot_conflict_allo
+ggsave("Figures/plot_conflict_allo.pdf", width = 12, height = 9)
+ggsave("Figures/plot_conflict_allo.png", width = 12, height = 9)
 
[…]

### 2024-12-06 analysis2.R

@@ -117,7 +117,7 @@ m_alloparent <-
   fit(f, data = e)
 
 plot_alloparenting_cost <-
-plot_predictions(m_alloparent, condition = c("AlloparentingFreqN", "Sex"), newdata = e) +
+  plot_predictions(m_alloparent, condition = c("AlloparentingFreqN", "Sex"), newdata = e) +
   scale_color_binary(labels = c("Female", "Male")) +
   ylim(0, NA) +
   labs(x = "Alloparenting Frequency (standardized)", y = "Signal cost") +

### 2024-12-07 Paper.qmd
Adjusted and combined figures, ggsave.

@@
 source("analysis.R")
-# source("analysis2.R") # Generates many plots. Slow!
+source("analysis2.R") # Generates many plots. Slow!
 ```
@@
 Unlike chimpanzee offspring, human offspring in ancestral populations therefore often grew up with multiple older and/or younger siblings who were potential competitors for parental investment and yet who could also provide (or require) allocare [@helfrecht2016; @lawson2016; @hagen2009].
-Juveniles in humans, unlike chimps, therefore evolved in families with multiple siblings simultaneously dependent on parents and others for essential care, likely selecting for credible signals of need.
+Juveniles in humans, unlike chimps, therefore evolved in families with multiple siblings simultaneously dependent on parents and others for essential care, likely intensifying selection for credible signals of need [@leonard2000; @harper1986; @neuenschwander2003].
 
@@
 However, `{r} fmt_sum(utila_df$ConflictFreqN == 0)` were described as never having conflicts with parents.
-See @fig-barplot_ConflictFreq.
+See @fig-barplot-conflict-allo.
 Conflict was also associated with higher signaling frequencies.
@@
 
-::: {#fig-barplot_ConflictFreq}
+::: {#fig-barplot-conflict-allo}
 ```{r}
-barplot_conflict
+plot_conflict_allo
 ```
 
-Frequency of conflict (per month) between child and primary caregiver as reported by primary caretaker.
-Color corresponds to sex of the child.
-Frequency key: 0 corresponds to "never," \<=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and \>30 corresponds to "multiple times per day."
-:::
-
-::: {#fig-barplot_AlloparentingFreq}
-```{r}
-barplot_alloparenting
-```
-
-Frequency of a child's alloparenting effort (per month) as reported by primary caretaker.
+**Left**: Frequency of conflict (per month) between child and primary caregiver as reported by primary caretaker.
+**Right**: Frequency of a child's alloparenting effort (per month) as reported by primary caretaker.
 Color corresponds to sex of the child.
@@
 
-Using lasso regression, we explored all potential predictors of each child signal frequency, as well as of conflict frequency. See @fig-lasso.
+Using lasso regression, we explored all potential predictors of each child signal frequency, as well as of conflict frequency.
+Child age was a strong negative predictor of all frequencies, and *Neighborhood quality* was too, albeit less so for *Crying* and not for *Sadness*.
+*Medical problems* was a weak positive predictor for all frequencies.
+Possession of a bicycle or washing machine was a positive predictor of *Signaling Frequency" and *Signal Cost*, whereas some other possessions were negative predictors of various frequencies.
+Possession of a TV with cable was the strongest negative predictor of *Sadness*, for example.
+Male sex was associated with higher *Conflict frequency* with the primary caretaker.
+See @fig-lasso.
 
@@
 
-The effects of particularly large risk or protective factors for child signaling are shown in @fig-signal-effects and @fig-alloparentingXsex.
+The effects of particularly large and consistent risk or protective factors for child signaling identified by the lasso regressions (@fig-lasso) are shown in @fig-signal-effects and @fig-alloparentingXsex.
 
@@
 :::
+
+In general, older child age and better neighborhoods were associated with less conflict and less signaling, whereas medical problems were associated with more conflict and more signaling.
+Alloparenting frequency was positively associated with signaling for girls, and negatively for boys.
 
@@
 For example, while the energetic cost of crying might be costly for the youngest children, it is likely insignificant for older children.
-Older children might instead need to weigh the reputational harm from crying [@citexxx] against the potential increase in social support.
+Older children might instead need to weigh the potential social costs of crying or otherwise expressing sadness [@glasberg1981; @glasberg1982; @zeman1996] against the potential increase in social support.
 

### 2024-12-07 Paper.qmd
Minor edit in Results

@@
 In general, older child age and better neighborhoods were associated with less conflict and less signaling, whereas medical problems were associated with more conflict and more signaling.
-Alloparenting frequency was positively associated with signaling for girls, and negatively for boys.
+There was little evidence for a sex difference in signaling, except that alloparenting frequency was positively associated with signaling for girls, and negatively for boys.
+Male sex was associated with more conflict.
 

### 2024-12-07 Paper.qmd
Changed "[" to "(" for figure refs

@@
 source("analysis.R")
-source("analysis2.R") # Generates many plots. Slow!
+# source("analysis2.R") # Generates many plots. Slow!
 ```
@@
 Caregivers reported a wide range of child signaling frequencies.
-About half of children were reported to cry or throw tantrums more than once a month, and almost 40% were reported to do so more than once a week [@fig-barplot_SignalFreq].
-Frequencies of sadness, crying, and tantrums were positively correlated with each other and with the frequency of conflict with the primary caretaker [@fig-signal_corrplot].
-They also loaded on PC1 in the same direction as indices of adversity, such as food insecurity, and in the opposite direction of indices of wealth and environmental quality, such as possessions, neighborhood quality, and income [@fig-PCA].
+About half of children were reported to cry or throw tantrums more than once a month, and almost 40% were reported to do so more than once a week (@fig-barplot_SignalFreq).
+Frequencies of sadness, crying, and tantrums were positively correlated with each other and with the frequency of conflict with the primary caretaker (@fig-signal_corrplot).
+They also loaded on PC1 in the same direction as indices of adversity, such as food insecurity, and in the opposite direction of indices of wealth and environmental quality, such as possessions, neighborhood quality, and income (@fig-PCA).
 
-Signaling decreased with child age, and increased with conflict with their primary caretaker, which were the two variables most closely linked with signaling in the MST [@fig-mst].
+Signaling decreased with child age, and increased with conflict with their primary caretaker, which were the two variables most closely linked with signaling in the MST (@fig-mst).
 In our lasso regression models (which excluded family conflict due to the likely bidirectional relationship between signaling and perceived conflict), child age and high neighborhood quality were consistently negatively associated with signaling frequency, as were household income, caregiver education, and number of adults helping with childcare.
-Medical problems were a consistent, albeit small positive predictor [@fig-lasso; @fig-signal-effects].
+Medical problems were a consistent, albeit small positive predictor (@fig-lasso; @fig-signal-effects).
 
@@
 
-Consistent with signaling models that emphasize the role of conflicts of interest in favoring the use of costlier signals, caregiver perceptions of conflict and child signaling frequency were correlated [@fig-signal_corrplot], and family conflict was the most common cause of a child's recent signaling episode [@fig-effect_last_signal_all_responses].
+Consistent with signaling models that emphasize the role of conflicts of interest in favoring the use of costlier signals, caregiver perceptions of conflict and child signaling frequency were correlated (@fig-signal_corrplot), and family conflict was the most common cause of a child's recent signaling episode (@fig-effect_last_signal_all_responses).
 Although conflicts of interest do not always lead to overt behavioral conflict [@altmann1980; @bateson1994], it is likely many, if not most, conflicts are ultimately caused by conflicts of interest.
-Conflicts, in turn, were negatively associated with child age and higher neighborhood quality, and positively associated with male sex, medical problems, and (oddly) caregiver education [@fig-conflict-effects].
+Conflicts, in turn, were negatively associated with child age and higher neighborhood quality, and positively associated with male sex, medical problems, and (oddly) caregiver education (@fig-conflict-effects).
 
-Caregiver responses to signals were generally positive or neutral, more so for situations involving injury and illness, and less so for situations involving family conflict, loss of privileges, and transgressions [@fig-effect_last_signal_all_responses; @fig-effect_last_signal_response].
+Caregiver responses to signals were generally positive or neutral, more so for situations involving injury and illness, and less so for situations involving family conflict, loss of privileges, and transgressions (@fig-effect_last_signal_all_responses; @fig-effect_last_signal_response).
 These patterns probably reflect a combination of confluences of interest vs.
@@
 
-More frequent sadness was positively associated with perceived higher need relative to other children [@fig-effect_relativeneed].
+More frequent sadness was positively associated with perceived higher need relative to other children (@fig-effect_relativeneed).
 Although crying and tantrums were not associated with higher perceived need, this was likely due to their high correlation with sadness and the propensity of lasso regression to retain only one of a set of correlated predictors [@zou2005].
-Higher perceived need was associated with higher relative investment [@fig-plot_relative_investment].
+Higher perceived need was associated with higher relative investment (@fig-plot_relative_investment).
 Signaling behavior did not predict differences in relative caregiver investment beyond its effect on perceived need [@sfig-plot-invest-coefs], similar to an experimental vignette study that found that the likelihood of costly support was largely mediated by the effect of costly signaling on belief in need [@gaffney2022].
@@
 
-Interestingly, older children were perceived to be needier than younger ones [@fig-effect_relativeneed].
+Interestingly, older children were perceived to be needier than younger ones (@fig-effect_relativeneed).
 One likely reason for this might be that caretakers perceive the types of investment older children seek as more costly.
@@
 Although there was little evidence for sex differences in signaling overall (perhaps because sex differences typically emerge in adolescence, and our sample of adolescents was small), there was evidence that sex interacted with alloparenting effort.
-We found that higher frequencies of alloparenting by girls was associated with a higher frequency of tantrums and overall signaling costs, whereas for boys this relationship was slightly negative [@fig-alloparentingXsex].
+We found that higher frequencies of alloparenting by girls was associated with a higher frequency of tantrums and overall signaling costs, whereas for boys this relationship was slightly negative (@fig-alloparentingXsex).
 In many societies, girls do more alloparenting than boys [@helfrecht2016; @weisner1977; @barry1971; @turke1988], which matches reports from our focus groups of Utilian mothers.
@@
 If so, girls with excessive alloparenting responsibilities might be signaling their need for more time to invest in peer relationships and status seeking behaviors outside the household.
-Opposed to this interpretation, our study found a small negative association between girls' alloparenting and caregiver reports of conflict [@fig-lasso], although plausibly when girls do what caregivers want, caregivers report less conflict.
+Opposed to this interpretation, our study found a small negative association between girls' alloparenting and caregiver reports of conflict (@fig-lasso), although plausibly when girls do what caregivers want, caregivers report less conflict.
 

### 2024-12-09 Paper.qmd
Minor fix
Minor paper edits.

@@
 For older children, the relatively insignificant energetic costs of crying might not be enough to ensure it is a credible signal when there are conflicts of interest.
-For this reason, crying among older children might be more similar to adult crying, which can increase support and empathy but sometimes also reduces reputation and increases stigmatization [@cornelius2001; @cretser1982; @hendriks2006; @jesser1989; @vingerhoets2016].
+For this reason, crying among older children might be more similar to adult crying, which can increase support and empathy but sometimes comes with social costs [@cornelius2001; @cretser1982; @hendriks2006; @jesser1989; @vingerhoets2016; @glasberg1981; @glasberg1982; @zeman1996].
 Such costs could be a means by which crying can still be a credible signal of need if they dissuade potential fakers from making the signal [@sznycer2023] when there are moderate conflicts of interest.
@@
 Both questions were framed need as involving time, effort, and resources.
-See Table @stbl-relvars for exact wording.
+See @stbl-relvars for exact wording.
 

### 2024-12-10 Paper.qmd
preparing to use public data; tweaks to pca plots

--- 2024-12-07 Paper.qmd
@@
 
+library(dplyr)
+library(purrr)
+library(stringr)
+library(labelled)
+
+load("data/m3_1e6_cost_freq.rda")
+
 source("recode.R")
@@
 
-# To use public data, comment out dataprep.R
-# and uncomment load statement
+# To use public data, comment out this line
 source("dataprep.R")
+
+# And uncomment this line
 # load("data/utilasignalingPublicData.rda")

### 2024-12-10 StatisticalDisclosureControl.R

@@ -44,6 +44,11 @@
 # Run in a fresh R session
 
 #+ message=F, warning=F
+library(dplyr)
+library(purrr)
+library(stringr)
+library(labelled)
+
 source("recode.R")
 source("dictionaries.R")
 source("dataprep.R")
@@ -72,10 +77,10 @@ utila_df <-
   arrange(householdID) |>
   # These variables are randomized within the two neighborhoods
   mutate(
-    NumberOfChildren = sample(NumberOfChildren, n()),
-    number_adults = sample(number_adults, n()),
-    UserLanguage = sample(UserLanguage, n()),
-    .by = Neighborhood2
+    NumberOfChildren = mean(NumberOfChildren, na.rm=T), #sample(NumberOfChildren, n()),
+    number_adults = mean(number_adults, na.rm=T), #sample(number_adults, n()),
+    UserLanguage = sample(UserLanguage, 1), #n()),
+    # .by = Neighborhood2
   )
 
 # householdIDs are set to the new ones generated above

### 2024-12-10 analysis.R

@@ -318,6 +318,7 @@ mdf2 <-
     OldestXsex = OldestChild * Sex
   )
 
+# Remove composite signaling vars
 e <-
   mdf2 |>
   dplyr::select(
@@ -333,9 +334,8 @@ e <-
   )
 
 e <- set_names(e, shortform_dict[names(e)])
-# names(e) <- shortform_dict[names(e)]
 
-m <- prcomp(e, scale. = T) # Remove composite signaling vars
+m <- prcomp(e, scale. = T)
 plot_loadings <- pca_loadings_plot(m, 1:2) # + theme(legend.position = 'top', legend.title = element_blank())
 plot_biplot <- pca_biplot(m, threshold = 0.2) + theme_minimal(15)
 

### 2024-12-10 dataprep.R

@@ -1,11 +1,5 @@
 
 library(utiladata2023) # data package
-library(dplyr)
-library(purrr)
-library(stringr)
-library(labelled)
-
-load("data/m3_1e6_cost_freq.rda")
 
 # functions used ----------------------------------------------------------
 
@@ -195,7 +189,7 @@ d <-
     MedicalProblemsMean = mean2(SchoolAges1, SchoolAges2, na.rm = TRUE) # 1 child has data for only 1
   ) |>
   ungroup() |>
-  left_join(dplyr::select(caregivers, householdID, CPRatio, Neighborhood, ImmigrateUtila, IncomeCategory, IncomeCategoryN,
+  left_join(dplyr::select(remove_labels(caregivers), householdID, CPRatio, Neighborhood, ImmigrateUtila, IncomeCategory, IncomeCategoryN,
                           EducationLevel, EducationLevelYears, AdultsMoney, AdultsHousework, AdultsChildcare, AdultsNoChildcare, CaregiverAge,
                           number_children2, number_adults, UserLanguage, CurrentJob, contains("CaregiverMarital"), NeighborhoodQuality, HouseQuality,
                           contains("SocialSupport"), contains("FoodSecurity"), contains("HomeInstability_"), contains("Reality_")), by = "householdID") |>

### 2024-12-10 dictionaries.R

@@ -91,7 +91,7 @@ shortform_dict <-
     YoungerKids = "Younger children",
     AdultsNoChildcare = "Adults that do not provide childcare",
     AdultsChildcare = "Adults that provide childcare",
-    OtherChildAlloparentingFreqN = "Other household child alloparenting freq.",
+    OtherChildAlloparentingFreqN = "Other child alloparenting freq.",
     HeightMean_2024 = "2024 height (cm)",
     HeightMean_2023 = "2023 height (cm)",
     WeightMean_KG_2024 = "2024 weight (kg)",

### 2024-12-10 pca_plot_functions.R

@@ -34,7 +34,7 @@ pca_loadings_plot <- function(obj, components = 1:3, sortby = 1, threshold = 0, 
     ggplot2::theme_bw(15)
 }
 
-pca_biplot <- function(obj, components = c(1,2), data = NULL, threshold = 0, reverse=NULL, label_size = 5, geom_point=T) {
+pca_biplot <- function(obj, components = c(1,2), data = NULL, threshold = 0, reverse=NULL, label_size = 5, geom_point=T, force = 1, force_pull = 1) {
 
   if (class(obj) != 'prcomp') stop('obj class must be prcomp')
   if (length(components) != 2 | mode(components) != 'numeric') stop('components is not a numeric vector of length 2')
@@ -70,7 +70,7 @@ pca_biplot <- function(obj, components = c(1,2), data = NULL, threshold = 0, rev
     labs(x = pct_var[1], y = pct_var[2])
 
   if (geom_point){
-    plot <- plot + geom_point()
+    plot <- plot + geom_point(color = "black")
   }
 
   datapc <- data.frame(varnames=rownames(obj$rotation), obj$rotation[,components])
@@ -85,9 +85,9 @@ pca_biplot <- function(obj, components = c(1,2), data = NULL, threshold = 0, rev
   )
   plot +
     coord_equal() +
-    ggrepel::geom_text_repel(data=datapc, aes(x=v1, y=v2, label=varnames), size = label_size, color="red", max.overlaps = Inf) +
-    geom_segment(data=datapc, aes(x=0, y=0, xend=v1, yend=v2), arrow=arrow(length=unit(0.2,"cm")), alpha=0.75, color="red") +
-    theme_minimal()
+    geom_segment(data=datapc, aes(x=0, y=0, xend=v1, yend=v2), arrow=arrow(length=unit(0.2,"cm")), linewidth=1, alpha=0.75, color="red") +
+    ggrepel::geom_text_repel(data=datapc, aes(x=v1, y=v2, label=varnames), size = label_size, color="blue", max.overlaps = Inf, force = force, force_pull = force_pull) +
+    theme_minimal(20)
 }
 
 ggdotchart <- function(v, threshold = NULL){

### 2024-12-10 analysis.R
Old PCA removed from analysis.

@@ -286,61 +286,3 @@ signal_alluvial_plot <-
   theme(axis.title.y = element_text(angle = 0, hjust = 1)) +
   scale_fill_viridis_d()
 signal_alluvial_plot
-
-# PCA ---------------------------------------------------------------------
-
-out <- skim(utila_df)
-nms <- out$skim_variable[out$skim_type == 'numeric' & out$complete_rate > 0.9]
-nms <- c(nms, "Sex", "UserLanguage", "ImmigrateUtila", "PartnerStatus", "OldestChild")
-
-mdf2 <-
-  utila_df |>
-  dplyr::select(
-    all_of(nms),
-    -householdID,
-    -childHHid,
-    -NegativeResponse,
-    -PositiveResponse,
-    -IncomeCategoryN,
-    # -ConflictFreqN,
-    -OtherChildrenHH, # omit or keep?
-    -contains('IllnessSusceptibility')
-  ) |>
-  na.omit() |>
-  mutate(
-    Sex = ifelse(Sex == "Female", 0, 1),
-    UserLanguage = ifelse(UserLanguage == "EN", 0, 1),
-    ImmigrateUtila = ifelse(ImmigrateUtila == "No", 0, 1),
-    PartnerStatus = ifelse(PartnerStatus == "Unpartnered", 0, 1),
-    across(-c(1:6), \(x) c(scale(x))),
-    AlloparentingXsex = AlloparentingFreqN * Sex,
-    ChildAgeXsex = ChildAge * Sex,
-    OldestXsex = OldestChild * Sex
-  )
-
-e <-
-  mdf2 |>
-  dplyr::select(
-    - SignalFreq,
-    - SignalCost,
-    - SignalFreqMax,
-    - OldestChild,
-    - OlderGirls,
-    - PartnerStatus,
-    - CaregiverAge,
-    - YoungerKids,
-    - contains("Xsex")
-  )
-
-e <- set_names(e, shortform_dict[names(e)])
-# names(e) <- shortform_dict[names(e)]
-
-m <- prcomp(e, scale. = T) # Remove composite signaling vars
-plot_loadings <- pca_loadings_plot(m, 1:2) # + theme(legend.position = 'top', legend.title = element_blank())
-plot_biplot <- pca_biplot(m, threshold = 0.2) + theme_minimal(15)
-
-plot_pca <- plot_loadings + plot_biplot + plot_layout(widths = c(1,2)) +
[…]

### 2024-12-11 Paper.qmd
keep-md: false; make html default

@@
   warning: false
-  keep-md: true
+  keep-md: false
 format:
-  docx:
-    number-sections: true
-    prefer-html: true
   html:
@@
     embed-resources: true
+  docx:
+    number-sections: true
+    prefer-html: true
 lightbox: true
@@
 # To use public data, comment out this line
-source("dataprep.R")
+# source("dataprep.R")
 
 # And uncomment this line
-# load("data/utilasignalingPublicData.rda")
+load("data/utilasignalingPublicData.rda")
 

### 2025-04-08 Paper.qmd
Committing submitted version prior to revising.

--- 2024-12-10 Paper.qmd
@@
 source("analysis.R")
-source("analysis2.R") # Generates many plots. Slow!
+#source("analysis2.R") # Generates many plots. Slow!
 ```
@@
 
-Humans are a uniquely interdependent species among the living primates [@tomasello2012] in that one’s fitness prospects depend greatly on a network of close kin and other social allies [@apicella2019].
+Humans are a uniquely interdependent species among the living primates [@tomasello2012] in that both sex's fitness prospects depend greatly on a network of close kin and other social allies [@apicella2019].
 This need for others is most intense with children, who are born highly altricial and require many years of investment to develop to the point where they can obtain enough food to survive on their own and develop the social skills needed to succeed in their communities [@kaplan2009; @rosenberg2021; @schniter2015].
 Much of this investment comes from mothers and fathers [@kramer2018].
-However, grandparents, aunts, uncles, and non-relatives also provide significant investment in children in many societies [@kramer2018; @sear2008], which was likely favored due to a combination of inclusive fitness benefits, reciprocity, and reputation management [@gurven2000] + (Schniter Preprint (permission needed).
+However, grandparents, aunts, uncles, and non-relatives also provide significant investment in children in many societies [@kramer2018; @sear2008], which was likely favored due to a combination of inclusive fitness benefits, reciprocity, and reputation management [@gurven2000; @schniter2023].
 Children also often serve as alloparents, helping to raise their younger siblings via direct care, food collection and preparation, and other family tasks [@hrdy2009; @meehan2009; @sear2008] despite still requiring substantial amounts of investment themselves [@kramer2018a].
@@
 Children often have needs and wants, where needs involve avoiding potential fitness costs, such as poor nutrition, injuries, or threats to status, and wants involve acquiring potential fitness benefits, such as highly nutritious foods and valuable artifacts.
-Needs and wants are often not apparent to caregivers, however, i.e., they are private information, and hence children must signal them to receive help.
+Needs and wants (hereafter needs) are often not apparent to caregivers, however, i.e., they are private information, and hence children must signal them to receive help.
 Signals differ from cues in that signals were favored by natural selection due to their ability to alter the behavior or perceptions of others in ways that benefit the signaler, on average, whereas cues are information organisms use to influence their behavior or physiological state that did not evolve for communicative effect [@maynardsmith1995].
@@
 
-However, even credible signals of needs or wants might not result in a beneficial response because belief in a child’s need is not the only factor determining the form and amount of support given by caregivers.
+However, even credible signals of need might not result in a beneficial response because belief in a child’s need is not the only factor determining the form and amount of support given by caregivers.
 Instead, caregivers are expected to weigh the costs and benefits of support against other options [@daly1987; @daly1983].
@@
 
-The anonymized data used in the study are available at <https://github.com/grasshoppermouse/utiladata2023>. Study code is available at <https://github.com/michaelrgaffney/utilasignaling>.
+Study code is available at <https://github.com/michaelrgaffney/utilasignaling>.
 
@@
 
+```{r sumtab_adult_cont}
+#| label: tbl-sumtab_adult_cont
+#| tbl-cap: Summary statistics for continuous variables.
+
+utila_df_caregiver |>
+  dplyr::select(any_of(names(utila_df_caregiver[names(longform_dict_adult_cont)]))) |>
+  setNames(longform_dict_adult_cont) |>
+  datasummary_skim(fun_numeric = list(N = \(x) length(na.omit(x)), Range = \(x) str_glue("{min(x, na.rm=T)}-{max(x, na.rm=T)}"), Mean = Mean, SD = SD, Histogram = function(x) ""))
+```
+
 ```{r sumtab_factors}
@@
 # convert to factor to omit NAs in datasummary_skim
-dx <- utila_df |>
+# dx <- utila_df |>
+#   dplyr::select(any_of(names(longform_dict))) |>
+#   mutate(
+#     across(starts_with(c("Lifestyle", "ImmigrateUtila")), as.factor)
+#   )
+
+dx <- utila_df_caregiver |>
   dplyr::select(any_of(names(longform_dict))) |>
@@
 
-Minimum spanning tree of child signaling and variables thought to cause differences in signaling strategies.
-Blue edges = negative \[do we want more detail here?
-e.g., partial correlations or w/e the proper Bayesian way to word that is\] correlations between variables.
-Red edges = positive correlations between variables.
-Edge lengths represent \[check that this is right and if more specificity is needed\] distances between variables.
-Line weights represent(1 - correlation coefficient)
+Minimum spanning tree derived from a Bayesian graphical model of child signaling and variables thought to cause differences in signaling strategies.
+Purple edges represent negative partial correlations between variables.
+Orange edges represent positive partial correlations between variables.
+Line weights represent partial correlation coefficients.
+Blue text represents our interpretation of the main clusters.
 :::
@@
 
-Effects of alloparenting frequency (per month) and signal cost and frequency of temper tantrums by sex.
+Effects of alloparenting frequency (per month) and signal cost and frequency of temper tantrums by sex when holding continuous variables at their mean and factor variables at their mode.
 :::
@@
 
-The proportion of children perceived by caretakers as being more needy, as needy, or less needy than a caretaker's other children.
+The proportion of children perceived by caretakers as being more needy, as needy, or less needy than a caretaker's other children when holding continuous variables at their mean and factor variables at their mode.
 A: Effect of child age on relative neediness B: Effect child sadness frequency on relative neediness.
@@
 
-The effect of perceived need (in terms of time, effort, and resources) relative investment within the household.
+The effect of perceived need (in terms of time, effort, and resources) relative investment within the household when holding continuous variables at their mean and factor variables at their mode.
 Outcome variable question wording: How much time, effort, and resources do you put into this child compared to your other children?
@@
 
-Cluster dendrogram for the causes of child signaling. Method = ward.D2.
+Hierarchical cluster analysis dendrogram for the causes of child signaling. Method: Euclidean distances between variables agglomerated with the ward.D2 algorithm.
 :::

### 2025-04-08 analysis.R

@@ -140,9 +140,10 @@ barplot_SignalFreq <-
   geom_col(position = "stack", color = "grey") +
   scale_fill_viridis_d(option = "B", direction = -1) +
   labs(x = "Number of children", y = "") +
-  guides(fill = guide_legend("Times per month:", nrow = 1, reverse = TRUE, position = "top")) +
+  guides(fill = guide_legend("Times/month:", nrow = 1, reverse = TRUE, position = "top")) +
   theme_minimal(15)
 barplot_SignalFreq
+ggsave("Figures/barplot_SignalFreq.png", barplot_SignalFreq)
 
 
 signal_subset <- utila_df[c("ConflictFreqN", "Sex", "ChildAge", "SadFreqN", "CryFreqN", "TantrumFreqN", "SignalFreq", "SignalCost", "AlloparentingFreqN", "NeighborhoodQuality")]

### 2025-04-08 analysis2.R

@@ -419,7 +419,7 @@ ggsave("Figures/plot_need_coefs.png", plot_need_coefs, width = 12, height = 12)
 plot_need_age <- ordinal_plot2(out$fit, ChildAge, SignalVars4, 'Relative need')
 plot_need_sad <- ordinal_plot2(out$fit, SadFreqN, SignalVars4, 'Relative need')
 
-plot_need_combined <- plot_need_age / plot_need_sad + ggtitle("") + plot_layout(guides = 'collect') + plot_annotation(tag_levels = "A")
+plot_need_combined <- plot_need_age / plot_need_sad + ggtitle("") + plot_layout(guides = 'collect') + plot_annotation(tag_levels = "a")
 ggsave("Figures/plot_need_combined.pdf", plot_need_combined, width = 12, height = 12)
 ggsave("Figures/plot_need_combined.png", plot_need_combined, width = 12, height = 12)
 

### 2025-04-08 analysis.R
Switched to using private data, and also only the upper triangle of the corr matrix.

@@ -155,7 +155,7 @@ signal_corrplot <-
     `Child sex` = as.numeric(`Child sex`)
   ) |>
   cor( use = "pairwise.complete.obs") |>
-  ggcorrplot(hc.order = TRUE, hc.method = "ward.D",lab = TRUE, lab_col = "black", lab_size = 4.5, tl.cex = 16) +
+  ggcorrplot(type = "upper", hc.order = TRUE, hc.method = "ward.D",lab = TRUE, lab_col = "black", lab_size = 4.5, tl.cex = 16) +
   scico::scale_fill_scico(palette = "vik", midpoint = 0, begin = .1, end = .9, limits = c(-1, 1)) +
   guides(fill = guide_colorbar(title = "Correlation coefficients"))
 signal_corrplot

### 2025-04-14 analysis.R
corplot: upper triangle only. Many code formatting changes

@@ -1,4 +1,3 @@
-
 library(ggplot2)
 library(forcats)
 library(patchwork)
@@ -59,11 +58,22 @@ signal_vars <- c(
 
 SignalVars <-
   utila_df |>
-  dplyr::filter(!(is.na(CryFreqN) & is.na(SadFreqN) & is.na(TantrumFreqN) & is.na(ConflictFreqN))) |>
+  dplyr::filter(
+    !(is.na(CryFreqN) &
+    is.na(SadFreqN) &
+    is.na(TantrumFreqN) &
+    is.na(ConflictFreqN))
+  ) |>
   dplyr::select(
     householdID,
     childHHid,
-    SadFreqN, CryFreqN, TantrumFreqN, SignalFreq, SignalCost, ConflictFreqN, MeanChildRelatedness, # Fullsibs, Halfsibs, Stepsibs,
+    SadFreqN,
+    CryFreqN,
+    TantrumFreqN,
+    SignalFreq,
+    SignalCost,
+    ConflictFreqN,
+    MeanChildRelatedness, # Fullsibs, Halfsibs, Stepsibs,
     all_of(names(signal_vars))
     # OnlyChild, Only children do not have to compete for attention or other forms of investment with existing children. They still may be motivated to signal for more investment which parents might prefer to devote to future children.X
     # OldestChild, Does not seem to add much beyond age + number of children.X
@@ -78,7 +88,12 @@ SignalVars <-
     # PartnerStatus = ifelse(PartnerStatus == "Unpartnered", 0, 1),
     AlloparentingXsex = AlloparentingFreqN * Sex,
   ) |>
-  na.omit()
+  na.omit() |>
+  mutate(
+    OtherKidsSignalCost = sum(SignalCost) - SignalCost,
+    OtherKidsConflict = sum(ConflictFreqN) - ConflictFreqN,
+    .by = householdID
+  )
 
 ## main paper plots --------------------------------------------------------
 
@@ -120,13 +135,20 @@ freq_short2 <- c(
 sfdf <-
   signalfreqdf |>
   mutate(
-    across(Sadness:Tantrums, \(x) factor(c(freq_short[x]), levels = c(freq_short)))
+    across(
+      Sadness:Tantrums,
+      \(x) factor(c(freq_short[x]), levels = c(freq_short))
+    )
   ) |>
   na.omit()
 
 signaldf_long <-
   sfdf |>
[…]

### 2025-04-14 Paper.qmd
Switched to using private data, and also only the upper triangle of the corr matrix.
Initial edits in response to reviewers and bib clean up.

@@
 # To use public data, comment out this line
-# source("dataprep.R")
+source("dataprep.R")
 
 # And uncomment this line
-load("data/utilasignalingPublicData.rda")
+#load("data/utilasignalingPublicData.rda")
 
@@
 Juveniles in humans, unlike chimps, therefore evolved in families with multiple siblings simultaneously dependent on parents and others for essential care, likely intensifying selection for credible signals of need [@leonard2000; @harper1986; @neuenschwander2003].
+
+Although infant signals of need, particularly infant cries, have been extensively studied in many populations [@barr1991; @fouts2005; @konner1974; @kruger2010; @kushnick2006; @meehan2017; @müller2021; @tronick1987], the full range of potential child signals of need, which likely extend beyond cries, has never been investigated in children of all ages.
+In particular, how and when older children signal need has rarely been investigated [for one exception, see @syme2023].
+The aim of this study was to help fill these gaps.
 
@@
 Children’s potential to invest in their families can also cause conflicts with their parents.
-Older children might prefer to spend their time investing in romantic or other social relationships, or learning valuable skills, whereas parents might prefer them to care for their siblings, acquire or prepare food, or complete other household chores [@lowe2003; @syme2023].
-Conflicts of interest over investment in the family also appear to extend to when a child starts having their own children [@kushnick2023; @emlen1995], with parents potentially benefiting from delaying the reproductive careers of certain offspring in order to gain more allocare for their other children [@flinn1989; @emlen1992] or speeding up a child’s departure from the household in order to provide more parental investment to other children [@kushnick2023].
+Older children can provide greater allocare, but might prefer to spend their time investing in romantic or other social relationships, or learning valuable skills, whereas parents might prefer them to care for their siblings, acquire or prepare food, or complete other household chores [@lowe2003; @syme2023].
+Conflicts of interest over investment in the family can even involve adult children who are starting to have their own children [@kushnick2023; @emlen1995].
+Parents can potentially benefit from delaying the reproductive careers of certain offspring in order to gain more allocare for their other children [@flinn1989; @emlen1992] or speeding up a child’s departure from the household in order to provide more parental investment to other children [@kushnick2023].
 
@@
 
-When conflicts of interest over investment exist, there may be much to gain from deception (e.g., crying despite the lack of need) so that one could receive the benefits that come from being seen as needy [@zahavi1993].
+When the fitness interests of signalers and receivers are aligned, i.e., there are no conflicts of interest, child signals of need are expected to be low cost, what @maynardsmith1995 refer to as "minimal signals," those whose cost is no greater than that needed to transmit the information effectively.
+In other words, although minimal signals might involve some cost (e.g., energy for a loud scream), they do not require any additional mechanism to enforce honesty.
+
+When conflicts of interest over investment exist, however, there may be much to gain from deception (e.g., crying despite the lack of need) so that one could receive the benefits that come from being seen as needy [@zahavi1993].
 Unchecked, deceptive signals could result in receivers ignoring all signals.
-
-Credible signals can nevertheless evolve and be maintained in multiple ways.
-Signals can be trusted in instances when there are no conflicts of interest over what information is conveyed [@guilford1995; @maynardsmith1995].
-They can also arise when signals cannot be faked (i.e., indexical signals).
+Credible signals can nevertheless evolve and be maintained in multiple ways, such as when signals cannot be faked (i.e., indexical signals).
 Credible signals can also evolve when the benefits of sending a signal outweigh the costs for individuals with the underlying quality or state, but otherwise the costs outweigh the benefits [@spence1973; @zahavi1975; @chisausky2023].
-Such signals are often termed “costly”, although what is costly is deception, not necessarily signaling per se [for discussion, see @chisausky2023].
+Such signals are often termed “costly,” although what is costly is deception, not necessarily signaling per se [for discussion, see @chisausky2023].^[Even when a mechanism is needed to ensure honesty, it is possible for signal cost to be zero at equilibrium because it is the potential cost of deceptive signaling, not honest signaling, that ensures honesty [@chisausky2023].]
 
 In the context of child signals of need, a child has access to information about their internal state, which is unavailable to others, that allows them to estimate their valuation of additional investment.
-By signaling in a credible way, the child reveals their resource valuation to potential caretakers who may be skeptical of the child’s claims due to the common conflicts of interest described above.
+When there are conflicts of interest between children and their caregivers, children might need to use signals whose costs outweigh the benefits of additional care for those not in need, but would result in a net benefit to those with genuine needs.
+In doing so, the child reveals their resource valuation to potential caretakers who may be skeptical of the child’s claims.
 With the information revealed by the signal, caretakers should generally be more likely to provide additional support due to their fitness interest in the child’s wellbeing.
@@
 Instead, caregivers are expected to weigh the costs and benefits of support against other options [@daly1987; @daly1983].
-Ancestrally, this likely often meant evaluating the viability of the child as well as how much a particular resource benefited one child relative to other children, current or future [@clutton-brock1991; @daly1983; @dawkins1976; @hagen1999].
-The former is thought to have resulted in children having to demonstrate their vigor early in life [@soltis2004], while the latter is expected to result in context-specific support [@beaulieu2008; @frankenhuis2022].
-For example, one unit of a resource is often expected to be worth more to a newborn than an adolescent, favoring that investment flowing to the newborn all else being equal.
-However, if resources are in short supply, parents may benefit more from investing more in older children as they are closer to reaching adulthood and have higher reproductive value [@daly1983].
-Additionally, a parent may also decide against investing in a child, despite deeming that investment worthwhile, if time is not right (e.g., if a mother preferred to breastfeed during rest rather than foraging) [@altmann1980].
+Ancestrally, this likely often meant evaluating the viability of the child, which favored selection for child signals of vigor [@soltis2004].
+In general, selection favors allocation of parental investment to current and future offspring that equalizes marginal fitness returns to the parent [@clutton-brock1991; @daly1983; @dawkins1976; @hagen1999; @beaulieu2008; @frankenhuis2022; @daly1983].
+
+Additionally, a parent may also decide against investing in a child, despite deeming that investment worthwhile, if time is not right (e.g., if a mother preferred to breastfeed during rest rather than when foraging).
 When this is the case, outward conflicts may exaggerate the extent of a conflict of interest and involve parents teaching children ways to behave that benefit them both [@altmann1980].
@@
 
-Saved for now: Although differing in their details, a common theme is that the indirect fitness costs signalers incur from imposing costs on their that parents and siblings through begging (be them energetic costs, survival costs, or reduced likelihood of future siblings) creates a floor to possible deception but falls short of preventing it [@daly1990; @wells2003].
+### Sadness
 
-Saved for now: Throughout human evolutionary history, individuals often found themselves in situations in which they received fitness-relevant benefits from others that could not easily be replaced [@apicella2012; @apicella2019; @hill2011; @tooby1996].
-In these instances, natural selection is expected to have favored those who provided support to others if 1) such help meant one could expect to receive fitness benefits from those helped in the future and 2) the costs of providing help were outweighed by the benefits to the helper on average [@andré2010; @hamilton1964; @tooby1996; @trivers1971].
+Sadness is an internal affective state that also has signaling functions in the form of facial expressions and other behavioral manifestations.
+It is generally thought to arise from loss [@horwitz2017; @levenson1999; @smith1993], especially when one has little power to reverse or mitigate the loss without help [@emanuel2023; @gunderson2021].
+Sadness and low mood, like other forms of psychological pain, probably evolved to motivate victims of adversity to shift their attention to the causes of adversity to mitigate its negative fitness consequences and to learn to solve or avoid future such adverse events [@andrews2009; @nesse1990; @thornhill1989].
+Low mood, for example, is often thought to be adaptive in situations when activity is dangerous or unlikely to result in benefits, with anhedonia, pessimism, and fatigue all helping reduce behavior that may not be worth the costs or that may cause additional problems [@nesse2019; @nettle2009].
+It may also encourage the reappraisal of one’s circumstances after new information comes to light that one’s current strategy is not working [@nesse2009].
+
+Through the reduction in behavior and attention in one’s usual interests that can accompany sadness and low-mood [@nettle2009], one’s relatives and other social partners would likely gain some indication, in the form of cues, that the sad individual was experiencing adversity and could benefit from increased support.
+In addition to this cue-based information, sadness also results in facial expressions and crying, both of which probably evolved to signal need [@balsters2013; @bowlby1980; @reed2017; @darwin1872].
+Such expressions seem to be important to conveying this information as is seen by responses to those who cannot produce certain elements of facial expressions like tear production [@hasson2009; @nesse2019] or certain facial movements [@bogart2012; @bogart2010].
+We conceptualize infrequent and transient sadness as a minimal signal of need.
+Frequent or prolonged sadness, however, is likely a form of withdrawal or depression, both of which might be effective costly signals of need when there are conflicts between the child and their caregivers [@gaffney2022;@syme2023].
 
@@
 Similarly, !Kung San mothers frequently respond to crying infants by providing comfort and breastfeeding access, with comforting behaviors being far more common than behaviors restricting infants [@kruger2010].
-Although this study found mothers were the most common responders, siblings, fathers, and other adults frequently stepped in so that 88% of crying bouts were responded to, with the majority of the cries ignored lasting under ten seconds.
+Although @kruger2010 found mothers were the most common responders, siblings, fathers, and other adults frequently stepped in so that 88% of crying bouts were responded to, with the majority of the cries ignored lasting under ten seconds.
 Similar responsiveness has also been reported among the Aka [@hewlett1998], the Mbendjele [@chaudhary2024], the Efe [@tronick1987], and the Ache, among whom young children have been described as “win\[ning\] most conflicts with parents simply by crying and whining” [@kaplan1987].
-Studies examining the potential for long-term benefits are rare and difficult to achieve; however, during a time of high infant mortality in a sample of 13 Massai infants, those with easy temperaments were more likely to die (5/7) than those with difficult temperaments (1/6), an outcome that is consistent with signals eliciting beneficial treatment [@devries1987].
+Studies examining the potential for long-term benefits are rare and difficult to achieve.
+However, during a time of high infant mortality in a sample of 13 Massai infants, those with easy temperaments were more likely to die (5/7) than those with difficult temperaments (1/6), an outcome that is consistent with signals eliciting beneficial treatment [@devries1987].
 
@@
 
-Although widely accepted as a signal, there is no consensus on why emotional crying might elicit positive responses.
-As the costs and benefits of signaling and gaining investment change with age, it is likely the strategic logic of crying changes too.
-It could be vital for newborns to signal health and vigor to reduce the risk of neglect or infanticide experienced by children with poor survival prospects [@daly1984; @soltis2004].
+Although it is widely accepted that crying is a signal, there is no consensus on if or when it should be conceptualized as a minimal signal, a costly signal, an indexical signal, or a deceptive (manipulative) signal.
+Contact calls between parents and offspring have evolved in numerous species, so cries might simply be a means for infants to elicit attention from caregivers [@mehr2021], which would be type of minimal signal.
+Demonstrating vigor and health early in life was likely vital for much of our evolutionary history due to the risk of neglect or infanticide experienced by children with poor survival prospects [@daly1984; @soltis2004].
 As crying comes with physiological costs [@rao1997; @thureen1998], it could serve as a credible signal of vigor if these costs are too high for unhealthy infants to bear [@furlow1997; @lummaa1998; @soltis2004; @wells2003].
+Children with severe medical conditions also produce acoustically distinct cries, suggesting that some elements of crying are indexical [@soltis2004].
 Alternatively, crying might impose costs on caregivers by, e.g., attracting predators or interfering with important activities, motivating caregivers to invest more to stop the crying [so-called blackmail or bargaining models; @zahavi1977a; @eshel1991; @bergstrom1999 @kennedy2021].
-More generally, contact calls between parents and offspring have evolved in numerous species, so cries might simply be a means for infants to elicit attention from caregivers [@mehr2021].
+
 For older children, the relatively insignificant energetic costs of crying might not be enough to ensure it is a credible signal when there are conflicts of interest.
 For this reason, crying among older children might be more similar to adult crying, which can increase support and empathy but sometimes comes with social costs [@cornelius2001; @cretser1982; @hendriks2006; @jesser1989; @vingerhoets2016; @glasberg1981; @glasberg1982; @zeman1996].
-Such costs could be a means by which crying can still be a credible signal of need if they dissuade potential fakers from making the signal [@sznycer2023] when there are moderate conflicts of interest.
-
-### Sadness
-
-Sadness is a type of negative affect that is generally thought to arise from loss [@horwitz2017; @levenson1999; @smith1993], especially when one has little power to reverse or mitigate the loss without help [@emanuel2023; @gunderson2021].
-Sadness and low mood, like other forms of psychological pain, probably evolved to motivate victims of adversity to shift their attention to the causes of adversity to mitigate its negative fitness consequences and to learn to solve or avoid future such adverse events [@andrews2009; @nesse1990; @thornhill1989].
-Low mood, for example, is often thought to be adaptive in situations when activity is dangerous or unlikely to result in benefits, with anhedonia, pessimism, and fatigue all helping reduce behavior that may not be worth the costs or that may cause additional problems [@nesse2019; @nettle2009].
-It may also encourage the reappraisal of one’s circumstances after new information comes to light that one’s current strategy is not working [@nesse2009].
-
-Through the reduction in behavior and attention in one’s usual interests that can accompany sadness and low-mood [@nettle2009], one’s relatives and other social partners would likely gain some indication, in the form of cues, that the sad individual was experiencing adversity and could benefit from increased support.
-In addition to this cue-based information, sadness also results in facial expressions and crying, both of which probably evolved to indicate need [@balsters2013; @bowlby1980; @reed2017; @darwin1872].
-Such expressions seem to be important to conveying this information as is seen by responses to those who cannot produce certain elements of facial expressions like tear production [@hasson2009; @nesse2019] or certain facial movements [@bogart2012; @bogart2010].
+Such costs could be a means by which crying can still be a credible signal of need if they dissuade potential fakers from making the signal [@sznycer2023] when there are conflicts of interest.
 
@@
 Temper tantrums may also be the result of adaptive strategies by which children can elicit more support from parents [@koch2003].
-They are common in children from 18 months to 4 years old [@potegal2003] but also exist among older children, albeit with decreased frequency [@macfarlane1962; @barbarin1999; @ward2013; @osterman2010; @goodenough1931; @bhatia1990].
+They are common in children from 18 months to 4 years old [@potegal2003] but also exist among older children, albeit with decreased frequency [@macfarlane1962; @barbarin1999; @ward1970; @osterman2010; @goodenough1931; @bhatia1990].
 Temper tantrums vary in their severity and include behaviors ranging from crying, shrieking, and screaming to violent bodily movements, aggression, and self-harm [@potegal2003].
@@
 Similarly, young chimpanzees have been observed to beat their heads on the ground when refused access to breast milk [@goodall1986], and American white pelican chicks have been found to throw themselves on the ground, flail violently, and bite at their own wings when begging for food [@schaller1964].
-Such behaviors could credibly signal need as they are likely relatively less costly to those who are truly in dire need compared to those who are not, thus deterring deceptive signaling [@hagen2008].
+Such behaviors could credibly signal need if the benefits of care for those who are truly in need outweigh the signal costs, whereas for those who are not in need, the benefits of care do not outweigh the signal costs, thus deterring deceptive signaling [@hagen2008].
 
@@
 
-Although infant signals of need, particularly infant cries, have been extensively studied in many populations [@barr1991; @fouts2005; @konner1974; @kruger2010; @kushnick2006; @meehan2017; @müller2021; @tronick1987], the full range of potential child signals of need, which likely extend beyond cries, has never been investigated in children of all ages.
-In particular, how and when older children signal need has rarely been investigated [for one exception, see @syme2023].
+This study aimed to establish the natural history of child signaling in Utila, Honduras, by measuring various indices of adversity, children's conflicts with caregivers, the frequencies of common child signals of need, and various types of caregiver responses.
+From these, we used signaling theory to explore 1) the causes of child signals of need, 2) how and why signal costs might vary, 3) how caregivers respond to indications of child need, 4) how signaling behavior changes with age and sex, 5) and if there were differences in child signaling and parental responses in two subpopulations that face different levels of social and economic adversity.
 
-This study aimed to establish the natural history of child signaling in Utila, Honduras, by quantifying the frequencies of common child signals of need and various types of caregiver responses.
-From these, we used signaling theory to explore 1) the causes of child signals of need, 2) how and why signal costs might vary, 3) how parents respond to indications of child need, 4) how signaling behavior changes with age and sex, 5) and if there were differences in child signaling and parental responses in two subpopulations that face different levels of social and economic adversity.
-More specifically, we hypothesized that children have evolved strategies to tailor their signals of need to 1) their valuation of a given resource, 2) the presence of existing information about a child's needs and wants (i.e., potential to benefit from a resource) and 3) the extent of conflicts of interest with caretakers, which should influence signal cost.
-Such strategies can lead to some combination of verbal requests for more support, facial sadness, crying, withdrawal, temper tantrums and self-harm, all of which may be effective in certain circumstances and are sometimes successful in eliciting more support from caretakers, despite a child's low status and formidability.
+More specifically, we hypothesized that children have evolved strategies to tailor their signals of need to 1) their valuation of a given resource, 2) the presence of existing information about a child's needs and wants (i.e., potential to benefit from a resource) and 3) the extent of conflicts of interest with caretakers, which should influence the decision to use minimal signals or signals whose costs would deter deception.
+Such strategies can lead to some combination of verbal requests for more support and facial sadness, which we consider to be minimal signals, or withdrawal, temper tantrums and self-harm, which we consider "costly" signals.
+Crying might be considered a "minimal" signal in some circumstances, such as when a child needs a loud auditory signal to attract attention on a noisy playground, but a "costly" signal in others, such as when signaling vigor.
+Similarly, infrequent sadness would be a minimal signal, but frequent sadness might be better conceptualized as costly withdrawal or depression.
+Each of these might be effective in certain circumstances and are sometimes successful in eliciting more support from caretakers.
 
@@
 
-### Frequency of child signaling, alloparenting contributions, and conflict with caretaker
+### Child signaling frequency and cost, alloparenting contributions, and conflicts with caregiver
 
@@
 For this reason, we gave children an alloparenting frequency of “never” if they were the only child within a household.
+We also calculated the summed frequency of child alloparenting contributions by other children within the household.
 
@@
 | Running Away Frequency | How often does your child leave the house without telling you when they get mad? |
-| Parent/Child Conflict Frequency | How often do you have conflicts with your child? |
+| Caregiver/Child Conflict Frequency | How often do you have conflicts with your child? |
 | Child Alloparenting Frequency | How often does your child take care of their other siblings? |
@@
 *Signal cost* was the sum of the frequencies of these three signals weighted by their relative costliness (i.e., *Signal cost* = *frequency of sadness* + 2 $\times$ *frequency of crying* + 3 $\times$ *frequency of temper tantrums*).
-
-We also calculated the summed frequency of child alloparenting contributions by other children within the household.
+Although we treated *signal cost* as a continuous variable, we conceptualized low signal cost values (e.g., infrequent sadness or crying) as more likely to be minimal signals, and high signal cost values (e.g., frequent crying or temper tantrums) as more likely to be "costly" signals.
 
@@
 
-We used six-point Likert scales for parental reports of 1) the relative neediness of children within the household, and 2) the relative investment they provide to each participating child compared to the rest of their children.
+We used six-point Likert scales for caretaker reports of 1) the relative neediness of children within the household, and 2) the relative investment they provide to each participating child compared to the rest of their children.
 Both questions were framed need as involving time, effort, and resources.
 See @stbl-relvars for exact wording.
-
-### Causes of, and responses to, the last instance of child signaling of need
-
-We asked caretakers about the last time each participating child was sad or cried.
-Caretakers provided a free response of the reason(s) for the emotional display, and their responses to it.
-The researcher then binned responses into at least one of 31 categories which included positive, negative, and neutral responses (from the child's perspective).
-For the full list of options see @stbl-response-valence.
-We then created a *caregiver response* variable, which was coded as 1 for a positive response (e.g., helping child), 0 for a neutral response (e.g., not responding), and -1 for a negative response (e.g., punishment; all from the child's perspective).
-MG and KS also coded the causes for the presence or absence of 17 elements (e.g., whether the cause involved *punishment*, *family conflict*, and/or *discomfort, pain, injury, or illness*).
 
@@
 See @stbl-Food-Security for the wording of these variables.
+
+### Causes of, and responses to, the last instance of child signaling of need
+
+We asked caretakers about the last time each participating child was sad or cried.
+Caretakers provided a free response of the reason(s) for the emotional display, and their responses to it.
+The researcher then binned responses into at least one of 31 categories which included positive, negative, and neutral responses (from the child's perspective).
+For the full list of options see @stbl-response-valence.
+We then created a *caregiver response* variable, which was coded as 1 for a positive response (e.g., helping child), 0 for a neutral response (e.g., not responding), and -1 for a negative response (e.g., punishment; all from the child's perspective).
+MG and KS also coded the causes for the presence or absence of 17 elements (e.g., whether the cause involved *punishment*, *family conflict*, and/or *discomfort, pain, injury, or illness*).
 
@@
 
-A PCA of primary study variables found that sociodemographic and ecological variables loaded primarily on PC1, whereas the child signaling, conflict, and child age variables loaded primarily on PC2.
-Child signaling variables modestly load on PC1 as well in a direction suggesting that greater signaling covaries with more adverse environments.
+A PCA of primary study variables found that the child signaling variables and the sociodemographic and ecological variables loaded primarily on PC1, in a direction suggesting that greater signaling covaried with more adverse environments.
+Household composition variables loaded primarily on PC2, covarying with child signaling variables in a direction suggesting that larger households covaried with less signaling.
 See @fig-PCA.
@@
 Child sadness, crying, and temper tantrums were common and highly correlated with one another.
-See @fig-barplot_SignalFreq and @fig-signal_corrplot.
+See @fig-barplot-SignalFreq and @fig-signal-corrplot.
 Caregivers reported that `{r} fmt_sum(utila_df$SadFreqN >= 8)` were sad at least once a week, `{r} fmt_sum(utila_df$CryFreqN >= 8)` cried at least once a week, and `{r} fmt_sum(utila_df$TantrumFreqN >= 8)` threw temper tantrums at least once a week.
@@
 However, many children were reported to never engage in these signals/states; sadness: `{r} fmt_sum(utila_df$SadFreqN == 0)`; crying; `{r} fmt_sum(utila_df$CryFreqN == 0)`; and tantrums: `{r} fmt_sum(utila_df$TantrumFreqN == 0)`.
-For a visualization of how signal frequencies varied within children see @sfig-signal_alluvial_plot.
+For a visualization of how signal frequencies varied within children see @sfig-signal-alluvial-plot.
 
@@
 However, `{r} fmt_sum(utila_df$ConflictFreqN == 0)` were described as never having conflicts with parents.
-See @fig-barplot-conflict-allo.
+See @sfig-barplot-conflict-allo.
 Conflict was also associated with higher signaling frequencies.
-See @fig-PCA and @fig-signal_corrplot.
+See @fig-PCA and @fig-signal-corrplot.
 
 The frequency of child alloparenting effort, which previous research has linked to parent-child conflict, was bimodally distributed, with `{r} fmt_sum(utila_df$AlloparentingFreqN == 0)` reported to never engage in it (`r sum(na.omit(utila_df$NumberOfChildren==1))` instances were due to a child being an only child) and `{r} fmt_sum(utila_df$AlloparentingFreqN >= 30)` reported to help with childcare at least once a day.
-See @fig-barplot_AlloparentingFreq.
+See @sfig-barplot-conflict-allo.
 
-::: {#fig-barplot_SignalFreq}
+::: {#fig-barplot-SignalFreq}
 ```{r}
@@
 
-::: {#fig-signal_corrplot}
+::: {#fig-signal-corrplot}
 ```{r}
@@
 For a more complete correlation matrix of study variables see @sfig-plot-full-corrmat.
-:::
-
-::: {#fig-barplot-conflict-allo}
-```{r}
-plot_conflict_allo
-```
-
-**Left**: Frequency of conflict (per month) between child and primary caregiver as reported by primary caretaker.
-**Right**: Frequency of a child's alloparenting effort (per month) as reported by primary caretaker.
-Color corresponds to sex of the child.
-Frequency key: 0 corresponds to "never," \<=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and \>30 corresponds to "multiple times per day."
 :::
@@
 
-The effects of particularly large and consistent risk or protective factors for child signaling identified by the lasso regressions (@fig-lasso) are shown in @fig-signal-effects and @fig-alloparentingXsex.
+The effects of particularly large and consistent risk or protective factors for child signaling identified by the lasso regressions (@fig-lasso) are shown in @fig-signal-effects and @sfig-alloparentingXsex.
 
@@
 Effects of selected predictors of child signaling frequencies when holding continuous variables at their mean and factor variables at their mode.
-:::
-
-::: {#fig-alloparentingXsex}
-![](Figures/plot_alloparentingXsex.png)
-
-Effects of alloparenting frequency (per month) and signal cost and frequency of temper tantrums by sex when holding continuous variables at their mean and factor variables at their mode.
 :::
@@
 
-There were positive relationships between perceived relative neediness within the family and both child age and the frequency of child sadness. See @fig-effect_relativeneed.
+There were positive relationships between perceived relative neediness within the family and both child age and the frequency of child sadness. See @fig-effect-relativeneed.
 
-::: {#fig-effect_relativeneed}
+::: {#fig-effect-relativeneed}
 ![](Figures/plot_need_combined.png)
@@
 Consistent with signaling theory, our model predicts a greater proportion of caretakers investing more in children who they deem needier and a greater proportion of caretakers investing less in children they deem less needy, controlling for the age of the child, and household demographics.
-See @fig-plot_relative_investment.
+See @fig-plot-relative-investment.
 
-::: {#fig-plot_relative_investment}
+::: {#fig-plot-relative-investment}
 ![](Figures/plot_invest_need.png)
@@
 
-::: {#fig-effect_last_signal_all_responses}
+::: {#fig-effect-last-signal-al-responses}
 ```{r}
@@
 However, primary caretakers were more likely to respond positively towards a child when they attributed the cause of signaling to discomfort, pain, injury, or illness, and less likely to respond positively when they attributed the cause of signaling to family conflict, child punishment, child trangressions, or a child losing an item or privleges.
-See @fig-effect_last_signal_response.
+See @fig-effect-last-signal-response.
 
-::: {#fig-effect_last_signal_response}
+::: {#fig-effect-last-signal-response}
 ![](Figures/plot_caregiver_response_combined.png)
@@
 Caregivers reported a wide range of child signaling frequencies.
-About half of children were reported to cry or throw tantrums more than once a month, and almost 40% were reported to do so more than once a week (@fig-barplot_SignalFreq).
-Frequencies of sadness, crying, and tantrums were positively correlated with each other and with the frequency of conflict with the primary caretaker (@fig-signal_corrplot).
-They also loaded on PC1 in the same direction as indices of adversity, such as food insecurity, and in the opposite direction of indices of wealth and environmental quality, such as possessions, neighborhood quality, and income (@fig-PCA).
+About half of children were reported to cry or throw tantrums more than once a month, and almost 40% were reported to do so more than once a week (@fig-barplot-SignalFreq).
 
-Signaling decreased with child age, and increased with conflict with their primary caretaker, which were the two variables most closely linked with signaling in the MST (@fig-mst).
+Our bivariate correlations, PCA, MST, and lasso regression analyses all provided a consistent pattern of results: child age was a strong negative predictor of signaling frequency and cost; conflict with the primary caretaker was a positive predictor of signaling frequency and cost; and multiple indices of adversity, such as low neighborhood quality, low household income, and medical problems, were positive predictors of signaling frequency and cost (@fig-mst, @fig-PCA, @fig-signal-corrplot, @fig-lasso, @fig-signal-effects).
+
+Contrary to expectations, immigration status, residence in Camponado, and primary language (Spanish or English as proxies for cultural differences), did not predict child signaling in our lasso regressions, despite the differences in wealth, job certainty, discrimination, and living conditions between those from the mainland and Utilians [@davis2014; @garcia2017; @garcia2018; @stonich2000].
+These variables correlated with neighborhood quality, however, which was retained in our models, and lasso regression tends to retain only one of a set of correlated predictors [@zou2005].
+
+Consistent with signaling models that emphasize the role of conflicts of interest in favoring the use of costlier signals, caregiver perceptions of conflict and child signaling frequency were correlated (@fig-signal-corrplot), and family conflict was the most common cause of a child's recent signaling episode (@fig-effect-last-signal-al-responses).
+Conflicts, in turn, were negatively associated with child age and higher neighborhood quality, and positively associated with male sex, medical problems, and (oddly) caregiver education (@fig-conflict-effects).
+Although conflicts of interest do not always lead to overt behavioral conflict [@altmann1980; @bateson1994], it is likely many, if not most, conflicts are ultimately caused by conflicts of interest.
+
+Consistent with the need signaling hypothesis, more frequent sadness was positively associated with perceived higher need relative to other children (@fig-effect-relativeneed).
+Higher perceived need, in turn, was associated with higher relative investment (@fig-plot-relative-investment).
+Crying and tantrums were not associated with higher perceived need, likely due to their high correlation with sadness and the propensity of lasso regression to retain only one of a set of correlated predictors [@zou2005].
+Signaling behavior did not predict differences in relative caregiver investment beyond its effect on perceived need [@sfig-plot-invest-coefs], similar to an experimental vignette study that found that the likelihood of costly support was largely mediated by the effect of costly signaling on belief in need [@gaffney2022].
+
+Interestingly, older children were perceived to be needier than younger ones (@fig-effect-relativeneed).
+One likely reason for this might be that caretakers perceive the types of investment older children seek as more costly.
+In line with this reasoning, multiple caretakers mentioned that older kids need more money and advice and get into trouble more often.
+It might also be that caretakers see investing in older children as relatively costly when they seek to redirect investment to younger children.
+
+Caregiver responses to signals of need were generally positive or neutral, more so for situations involving injury and illness, and less so for situations involving family conflict, loss of privileges, and transgressions (@fig-effect-last-signal-al-responses; @fig-effect-last-signal-response).
+These patterns probably reflect a combination of shared interests vs.
+conflicts of interest between child and caretaker, some of which might represent caretaker investment in some children more than others, and greater caretaker certainty about need in cases of injury and illness.
+That said, our model predicts positive responses in 35% of cases involving transgressions, suggesting that children are successful in negotiating better outcomes at least some of the time.
+
+The foregoing results are consistent with the hypothesis that sadness, crying, and tantrums are signals of need, that children's need is higher in adverse environments and at younger ages, that costly signals rather than minimal ones are often needed to ensure credibility when there are conflicts of interest between caregivers and children, and that signals of need are effective at increasing caregiver perceptions of need that ultimately lead to increased investment in children.
+
+<!-- Signaling decreased with child age, and increased with conflict with their primary caretaker, which were the two variables most closely linked with signaling in the MST (@fig-mst).
 In our lasso regression models (which excluded family conflict due to the likely bidirectional relationship between signaling and perceived conflict), child age and high neighborhood quality were consistently negatively associated with signaling frequency, as were household income, caregiver education, and number of adults helping with childcare.
 Medical problems were a consistent, albeit small positive predictor (@fig-lasso; @fig-signal-effects).
+-->
 
-Despite the differences in wealth, job certainty, discrimination, and living conditions between those from the mainland and Utilians [@davis2014; @garcia2017; @garcia2018; @stonich2000], primary language, immigration status, and residence in Camponado did not predict child signaling in our lasso regressions.
-These variables correlated with neighborhood quality, however, which was retained in our models, and lasso regression tends to retain only one of a set of correlated predictors [@zou2005].
-
-The foregoing results are consistent with the hypothesis that sadness, crying, and tantrums are signals of need, and that children's need is higher in adverse environments and at younger ages.
 One possible reason for the negative association of signaling with age, which is also seen in more industrialized settings [@zeifman2013; @potegal2003; @koch2003; @macfarlane1962], is that as children mature, they are more capable of fulfilling their own needs, more able to ask others for what they need [@osterman2010], and less vulnerable to threats.
@@
 
-Consistent with signaling models that emphasize the role of conflicts of interest in favoring the use of costlier signals, caregiver perceptions of conflict and child signaling frequency were correlated (@fig-signal_corrplot), and family conflict was the most common cause of a child's recent signaling episode (@fig-effect_last_signal_all_responses).
-Although conflicts of interest do not always lead to overt behavioral conflict [@altmann1980; @bateson1994], it is likely many, if not most, conflicts are ultimately caused by conflicts of interest.
-Conflicts, in turn, were negatively associated with child age and higher neighborhood quality, and positively associated with male sex, medical problems, and (oddly) caregiver education (@fig-conflict-effects).
-
-Caregiver responses to signals were generally positive or neutral, more so for situations involving injury and illness, and less so for situations involving family conflict, loss of privileges, and transgressions (@fig-effect_last_signal_all_responses; @fig-effect_last_signal_response).
-These patterns probably reflect a combination of confluences of interest vs.
-conflicts of interest between child and caretaker, some of which might represent caretaker investment in some children more than others, and less caretaker uncertainty about need in cases of injury and illness.
-That said, our model predicts positive responses in 35% of cases involving transgressions, suggesting that children are successful in negotiating better outcomes at least some of the time.
-
-More frequent sadness was positively associated with perceived higher need relative to other children (@fig-effect_relativeneed).
-Although crying and tantrums were not associated with higher perceived need, this was likely due to their high correlation with sadness and the propensity of lasso regression to retain only one of a set of correlated predictors [@zou2005].
-Higher perceived need was associated with higher relative investment (@fig-plot_relative_investment).
-Signaling behavior did not predict differences in relative caregiver investment beyond its effect on perceived need [@sfig-plot-invest-coefs], similar to an experimental vignette study that found that the likelihood of costly support was largely mediated by the effect of costly signaling on belief in need [@gaffney2022].
-Altogether, the foregoing results generally support the hypothesis that the signals we investigated are costly, credible signals of need.
-
-Interestingly, older children were perceived to be needier than younger ones (@fig-effect_relativeneed).
-One likely reason for this might be that caretakers perceive the types of investment older children seek as more costly.
-In line with this reasoning, multiple caretakers mentioned that older kids need more money and advice and get into trouble more often.
-It might also be that caretakers see investing in older children as relatively costly as they seek to redirect investment toward younger, needier children.
-
 Finally, based on previous research showing that family demands, such as caring for younger siblings, was an important source of parent-offspring conflict [@syme2023], we investigated the relationship between the frequency of alloparenting by children and their frequency of signaling.
 Although there was little evidence for sex differences in signaling overall (perhaps because sex differences typically emerge in adolescence, and our sample of adolescents was small), there was evidence that sex interacted with alloparenting effort.
-We found that higher frequencies of alloparenting by girls was associated with a higher frequency of tantrums and overall signaling costs, whereas for boys this relationship was slightly negative (@fig-alloparentingXsex).
+We found that higher frequencies of alloparenting by girls was associated with a higher frequency of tantrums and overall signaling costs, whereas for boys this relationship was slightly negative (@sfig-alloparentingXsex).
 In many societies, girls do more alloparenting than boys [@helfrecht2016; @weisner1977; @barry1971; @turke1988], which matches reports from our focus groups of Utilian mothers.
@@
 
-Many models designed to explain signals of need (e.g., the begging of young birds) predict that offspring who are worse off or are receiving less investment have more to gain from signaling than those with more [@mock2011; @godfray1991; @caro2016a; @godfray2000] and multiple game theoretical models examining parent-offspring conflict have found that parental resources have a modest impact on the evolutionary stable value for offspring demand for resources [@bossan2013; @mock1997 [confirm]].
+Many models designed to explain signals of need (e.g., the begging of young birds) predict that offspring who are worse off or are receiving less investment have more to gain from signaling than those with more [@mock2011; @godfray1991; @caro2016a; @godfray2000] and multiple game theoretical models examining parent-offspring conflict have found that parental resources have a modest impact on the evolutionary stable value for offspring demand for resources [@bossan2013; @mock1997].
 In line with this reasoning, children in lower income households and riskier environments might benefit more from signaling effort if they experience greater levels of adversity or perceive themselves as disadvantaged compared to other children.
@@
 
-The main limitation of this study is that it is completely exploratory: our data-driven analyses, which can reveal unexpected relationships and inspire new theory, also risk overfitting the data, suggesting associations that might not exist.
-Our study was also entirely cross-sectional, further limiting our ability to make causal inferences.
+The main limitation of this study is that it was exploratory: our data-driven analyses, which can reveal unexpected relationships and inspire new theory, also risked overfitting the data, suggesting associations that might not exist.
+Our study was also cross-sectional, further limiting our ability to make causal inferences.
+In particular, there is likely a bidirectional relationship between child signaling and caregivers' perceptions of conflict that our study design could not disentangle.
 Furthermore, our penalized regressions included many variables, some of which might have been colliders, inducing spurious associations between outcome and predictor variables.
+We also assumed *a priori* that, at least in young children, infrequent sadness and crying were minimal signals not requiring a costly honesty-enforcing mechanism, whereas frequent sadness and crying, as well as tantrums, were costly signals.
 
@@
 All data come from caregiver reports, and therefore, reflect the biases of caretakers, whatever those might be.
+For example, caregiver reports of their own responses to recent signals of need might be positively biased.
+In addition, questions about neighborhood quality lacked a reference, so some caregivers might have compared their neighborhoods to mainland neighborhoods and others to island neighborhoods.
 
@@
 Children face the challenge of eliciting support from others in times of adversity and opportunity.
-On the island of Utila, child signaling strategies were highly correlated and engaged in by children of all ages, albeit more frequently by younger children and those living in more adverse conditions, who plausibly had greater need.
+On the island of Utila, child signaling strategies were highly correlated with each other, and were exhibted by children of all ages, albeit more frequently by younger children and those living in more adverse conditions, who plausibly had greater need.
 Consistent with signaling models that emphasize the role of conflicts of interest in favoring the use of costlier signals, child signaling effort was positively correlated with the frequency of conflict with caretakers and with perceptions of need, and greater perceptions of child need predicted greater investment in a given child.
@@
 
-::: {#sfig-signal_alluvial_plot}
+::: {#sfig-signal-alluvial-plot}
 ```{r}
@@
 Frequency key: Low includes "never" and "once a month or less," medium includes "more than once a month but less than once a week" and "more than once a week but not daily," high includes "daily," and "multiple times per day."
+:::
+
+::: {#sfig-barplot-conflict-allo}
+```{r}
+plot_conflict_allo
+```
+
+**Left**: Frequency of conflict (per month) between child and primary caregiver as reported by primary caretaker.
+**Right**: Frequency of a child's alloparenting effort (per month) as reported by primary caretaker.
+Color corresponds to sex of the child.
+Frequency key: 0 corresponds to "never," \<=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and \>30 corresponds to "multiple times per day."
+:::
+
+::: {#sfig-alloparentingXsex}
+![](Figures/plot_alloparentingXsex.png)
+
+Effects of alloparenting frequency (per month) and signal cost and frequency of temper tantrums by sex when holding continuous variables at their mean and factor variables at their mode.
 :::

### 2025-04-18 Paper.qmd

@@
 # And uncomment this line
-#load("data/utilasignalingPublicData.rda")
+# load("data/utilasignalingPublicData.rda")
 
@@
 Consistent with signaling theory, children who were sad more frequently were perceived as needier within the household and were more likely to receive investment.
-Consistent with previous research, increased alloparenting effort was associated with more signaling by girls but less signaling by boys.
+Caregivers were less likely to respond positively in situations of family conflict or child transgressions, and more likely for injuries and illnesses.
 Our results suggest that evolutionary theories of signaling can help explain patterns of child sadness, crying, and temper tantrums.
@@
 
-Although infant signals of need, particularly infant cries, have been extensively studied in many populations [@barr1991; @fouts2005; @konner1974; @kruger2010; @kushnick2006; @meehan2017; @müller2021; @tronick1987], the full range of potential child signals of need, which likely extend beyond cries, has never been investigated in children of all ages.
+Although infant signals of need, particularly infant cries, have been extensively studied in many populations [@barr1991; @fouts2005; @konner1972; @kruger2010; @kushnick2006; @meehan2017; @müller2021; @tronick1987], the full range of potential child signals of need, which likely extend beyond cries, has never been investigated in children of all ages.
 In particular, how and when older children signal need has rarely been investigated [for one exception, see @syme2023].
@@
 It is generally thought to arise from loss [@horwitz2017; @levenson1999; @smith1993], especially when one has little power to reverse or mitigate the loss without help [@emanuel2023; @gunderson2021].
-Sadness and low mood, like other forms of psychological pain, probably evolved to motivate victims of adversity to shift their attention to the causes of adversity to mitigate its negative fitness consequences and to learn to solve or avoid future such adverse events [@andrews2009; @nesse1990; @thornhill1989].
+Sadness and low mood, like other forms of psychological pain, probably evolved to motivate victims of adversity to shift their attention to the causes of adversity to mitigate its negative fitness consequences and to learn to solve or avoid similar events in the future [@andrews2009; @nesse1990; @thornhill1989].
 Low mood, for example, is often thought to be adaptive in situations when activity is dangerous or unlikely to result in benefits, with anhedonia, pessimism, and fatigue all helping reduce behavior that may not be worth the costs or that may cause additional problems [@nesse2019; @nettle2009].
@@
 We conceptualize infrequent and transient sadness as a minimal signal of need.
-Frequent or prolonged sadness, however, is likely a form of withdrawal or depression, both of which might be effective costly signals of need when there are conflicts between the child and their caregivers [@gaffney2022;@syme2023].
+Frequent or prolonged sadness, however, is likely a form of withdrawal or depression, both of which might be effective costly signals of need when there are conflicts between the child and their caregivers [@gaffney2022; @syme2023].
 
@@
 For this reason, crying among older children might be more similar to adult crying, which can increase support and empathy but sometimes comes with social costs [@cornelius2001; @cretser1982; @hendriks2006; @jesser1989; @vingerhoets2016; @glasberg1981; @glasberg1982; @zeman1996].
-Such costs could be a means by which crying can still be a credible signal of need if they dissuade potential fakers from making the signal [@sznycer2023] when there are conflicts of interest.
+Such costs could be a means by which crying can still be a credible signal of need if they dissuade potential fakers from making the signal when there are conflicts of interest [@sznycer2023].
 
@@
 Temper tantrums may also be the result of adaptive strategies by which children can elicit more support from parents [@koch2003].
-They are common in children from 18 months to 4 years old [@potegal2003] but also exist among older children, albeit with decreased frequency [@macfarlane1962; @barbarin1999; @ward1970; @osterman2010; @goodenough1931; @bhatia1990].
+They are common in children from 18 months to 4 years old [@potegal2003] but are also exhibited by older children, albeit with decreased frequency [@macfarlane1954; @barbarin1999; @ward1970; @osterman2010; @goodenough1931; @bhatia1990].
 Temper tantrums vary in their severity and include behaviors ranging from crying, shrieking, and screaming to violent bodily movements, aggression, and self-harm [@potegal2003].
 Similar behaviors are also seen in other young animals, often surrounding feeding or weaning conflict [@altmann1980; @goodall1986; @schaller1964].
-Tantrums have been reported in many societies with different forms of subsistence systems, political organization, and norms surrounding anger and parenting [@albino1956; @banisalameh2021; @fouts2005; @henderson1966; @hilger1957; @konner1974; @ward1970; @johnson2003; @bhatia1990; @myers1988; @dubois @konner1972], suggesting they might be a human universal [@potegal2019].
-As with the other proposed signals of need, they are capable of eliciting a wide range of responses, including pro-child responses (e.g., parents giving in to a child’s wishes), neutral responses [e.g., parents ignoring the child, @hilger1957; @konner1974; @johnson2003], and negative responses like abuse [@schmitt1987].
+Tantrums have been reported in many societies with different forms of subsistence systems, political organization, and norms surrounding anger and parenting [@albino1956; @banisalameh2021; @fouts2005; @henderson1966; @hilger1957; @konner1972; @ward1970; @johnson2003; @bhatia1990; @myers1988; @dubois], suggesting they might be a human universal [@potegal2019].
+As with the other proposed signals of need, they are capable of eliciting a wide range of responses, including pro-child responses (e.g., parents giving in to a child’s wishes), neutral responses [e.g., parents ignoring the child, @hilger1957; @konner1972; @johnson2003], and negative responses like abuse [@schmitt1987].
 
-As with the vocalizations that come with crying, temper tantrums might motivate caretakers to upregulate their support of the signaler due to the energetic costs of tantrums or the increased risk of predation stemming from the noisy display, both of which might result in credible signals if they are costly enough to deter faking [@lummaa1998].
-In addition, tantrums also impose direct costs on caretakers who may be less able to complete important tasks [@altmann1980] like finding or preparing food or taking care of other children.
-
-Another possibility is that temper tantrums carry the risk or implication of self-harm.
+Sadness is often not accompanied by anger, suggesting little conflict with caregivers, and its costs are low when not prolonged.
+Temper tantrums, in constrast, are closely linked to anger [@potegal2003], suggesting conflict with caregivers, and have high costs even when the tantrum is brief.
+Thus, unlike sadness, temper tantrums might be useful when there are substantial and immediate gains to be had from a credible signal of need.
+The costs of temper tantrums could be energetic, or could come from the risk or implication of self-harm [@lummaa1998].
 For example, Utilian mothers reported in focus groups that some temper tantrums involved children banging their heads on things when they do not get what they want, behaviors that have been reported in other populations [@belden2008; @burrows1953; @daniels2012; @dennis1940; @vandenakker2022].
@@
 Such behaviors could credibly signal need if the benefits of care for those who are truly in need outweigh the signal costs, whereas for those who are not in need, the benefits of care do not outweigh the signal costs, thus deterring deceptive signaling [@hagen2008].
+
+Alternatively, because tantrums impose direct costs on caretakers by interfering with important tasks [@altmann1980] like finding or preparing food or taking care of other children, they might be more akin to a strategy to coerce caretakers into providing more support.
 
@@
 
-Participants for this study were recruited through convenience sampling.
+Participant caregivers for this study were recruited through convenience sampling.
 JH recruited early participants with the help of primary informants and the use of snowball sampling.
@@
 
-All participants for the survey portion of the study met the inclusion criteria. Most participants were the biological mothers of one or more children in the study.
+`{r} xfun::numbers_to_words(n_distinct(householdIDs), cap=T)` primary caretakers provided data on `{r} nrow(utila_df)` children.
+All participants for the survey portion of the study met the inclusion criteria.
+Most participants were the biological mothers of one or more children in the study.
+Several participants were caring for children younger than 6 or older than 18.
+We collected caregiver reports of signaling for all children 5-20 years old.
 
@@
 | Tantrum Frequency | How often does your child throw temper tantrums? (e.g., scream, throws things, or throws themselves on the floor) |
-| Running Away Frequency | How often does your child leave the house without telling you when they get mad? |
 | Caregiver/Child Conflict Frequency | How often do you have conflicts with your child? |
@@
 Those used in this study included the (1) ages of those in the household, (2) sexes of those in the household, (3) kinship ties within households, (4) partnership status of the primary caregiver, (5) number of adults who provide childcare within the household, (6) highest level of education of the primary caregiver, (7) immigration status (in regards to Utila), (8) monthly household income, (9) neighborhood of residence, (10) current job of the primary caretaker, and (11) ownership of various possessions (e.g., homes, washing machines, or cellphones; see @stbl-possession-list for full list).
-These data allowed us to calculate the number of children (total, younger, and older) and adults within the household as well as the genetic relatedness between siblings.
-We also created a variable for residence in Camponado, a low-SES, densely populated neighborhood on the island with less sanitation access mostly comprising individuals from the mainland [@garcia2017].
+These data allowed us to calculate the number of children (total, younger, and older) and adults within the household as well as the genetic relatedness between siblings.^[Because we wanted a relatedness value for all ego children, including those without siblings, this measure included the focal child (r=1) in calculations of average child relatedness within households.] We also created a variable for residence in Camponado, a low-SES, densely populated neighborhood on the island with less sanitation access mostly comprising individuals from the mainland [@garcia2017].
 
@@
 
+Although child age was negatively associated with all signaling frequencies, a general linear mixed effects negative binomial model showed that the negative association was strongest for *Crying* and for males (@sfig-signaltype).
+
 The effects of particularly large risk or protective factors for conflict frequency are shown in @fig-conflict-effects.
@@
 
+<!-- [One dense paragraph on how what we have learned relates to what has been reported empirically in the past]
+- Note: we did not mention declines with age in the introduction
+- Vingerhoets' lab and their colleagues maintain crying is understudied with older children
+
+[Move first sentence down to 468 and broaden? Or at least broaden? MacFarlane's study of tantrums among Berkeley: declines from 5 to 12]
+
+The cries and fussing of young infants increases until 6 weeks, peaking in some populations at over two hours a day, and then declines steadily until 12-16 weeks [@vermillet2022; @barr1991].
+-->
+
 Caregivers reported a wide range of child signaling frequencies.
-About half of children were reported to cry or throw tantrums more than once a month, and almost 40% were reported to do so more than once a week (@fig-barplot-SignalFreq).
+About half of children were reported to cry or throw tantrums more than once a month, and almost 40% were reported to do so more than once a week (@fig-barplot-SignalFreq), with higher frequencies in younger children, lower frequencies in older children, and particularly steep age-related declines for crying and for all types of signaling by males (@sfig-signaltype).
 
@@
 
-Consistent with signaling models that emphasize the role of conflicts of interest in favoring the use of costlier signals, caregiver perceptions of conflict and child signaling frequency were correlated (@fig-signal-corrplot), and family conflict was the most common cause of a child's recent signaling episode (@fig-effect-last-signal-al-responses).
+Consistent with signaling models that emphasize the role of conflicts of interest in favoring the use of signals with honesty-enforcing mechanisms (e.g., "costly" signals), caregiver perceptions of conflict and child signaling frequency were positively correlated (@fig-signal-corrplot), and family conflict was the most common cause of a child's recent signaling episode (@fig-effect-last-signal-al-responses).
 Conflicts, in turn, were negatively associated with child age and higher neighborhood quality, and positively associated with male sex, medical problems, and (oddly) caregiver education (@fig-conflict-effects).
@@
 
-Caregiver responses to signals of need were generally positive or neutral, more so for situations involving injury and illness, and less so for situations involving family conflict, loss of privileges, and transgressions (@fig-effect-last-signal-al-responses; @fig-effect-last-signal-response).
+Caregiver responses to signals of need were generally positive or neutral, more so for situations involving injury and illness, and less so for situations involving family conflict, loss of privileges, and transgressions, which had greater proportions of negative responses (@fig-effect-last-signal-al-responses; @fig-effect-last-signal-response).
 These patterns probably reflect a combination of shared interests vs.
 conflicts of interest between child and caretaker, some of which might represent caretaker investment in some children more than others, and greater caretaker certainty about need in cases of injury and illness.
+Negative responses, e.g., punishment, can be used to change future child behavior that is costly to the caregiver (e.g., harming siblings or wasting resources), and possibly also to signal the costliness of providing the desired support [@hagen2024; @gaffney2022].
 That said, our model predicts positive responses in 35% of cases involving transgressions, suggesting that children are successful in negotiating better outcomes at least some of the time.
 
-The foregoing results are consistent with the hypothesis that sadness, crying, and tantrums are signals of need, that children's need is higher in adverse environments and at younger ages, that costly signals rather than minimal ones are often needed to ensure credibility when there are conflicts of interest between caregivers and children, and that signals of need are effective at increasing caregiver perceptions of need that ultimately lead to increased investment in children.
+Our results parallel those of previous studies of infants and young children that found that, outside of peaks in infancy and early childhood, the frequencies of crying and temper tantrums decrease with age [@macfarlane1954; @ward1970; @osterman2010; @vermillet2022; @barr1991; @potegal2003; @bhatia1990].
+Studies examining crying and tantrums in later childhood and adolensce are rare, especially for crying [@rottenberg2012; @zeifman2013].
+Our findings suggest the frequency of crying continues to decrease into childhood, dropping precipitously as children grow older.
+The frequency of temper tantrums also declined, as did the frequency of child sadness, albeit at a much slower rate (@fig-lasso, @fig-signal-effects).
+Our results also align with studies examining responses to the crying and fussing of infants that have shown that they often lead to increased support [e.g., breastfeeding, attention, or maintaining proximity; @lin2012; @kushnick2006; @kruger2010; @hewlett1998; @chaudhary2024; @tronick1987], but can also lead  to negative responses like abuse [@zeevenhooven2018; @kurth2011; @howard2006; @soltis2004].
+
+Our results are consistent with the hypothesis that sadness, crying, and tantrums are signals of need, that children's need is higher in adverse environments and at younger ages, that costly signals rather than minimal ones are often needed to ensure credibility when there are conflicts of interest between caregivers and children, and that signals of need are effective at increasing caregiver perceptions of need that ultimately lead to increased investment in children.
+
+Many models designed to explain signals of need (e.g., the begging of young birds) predict that offspring who are worse off or are receiving less investment have more to gain from signaling than those with more [@mock2011; @godfray1991; @caro2016a; @godfray2000] and multiple game theoretical models examining parent-offspring conflict have found that parental resources have a modest impact on the evolutionary stable value for offspring demand for resources [@bossan2013; @mock1997].
+In line with this reasoning, children in lower income households and riskier environments might benefit more from signaling effort if they experience greater levels of adversity or perceive themselves as disadvantaged compared to other children.
 
@@
 
-One possible reason for the negative association of signaling with age, which is also seen in more industrialized settings [@zeifman2013; @potegal2003; @koch2003; @macfarlane1962], is that as children mature, they are more capable of fulfilling their own needs, more able to ask others for what they need [@osterman2010], and less vulnerable to threats.
+One possible reason for the negative association of signaling with age, which is also seen in more industrialized settings [@zeifman2013; @potegal2003; @koch2003; @macfarlane1954], is that as children mature, they are more capable of fulfilling their own needs, more able to ask others for what they need [@osterman2010], and less vulnerable to threats.
 The effects of age on signaling are also likely due to how age influences the costliness of signaling need.
 For example, while the energetic cost of crying might be costly for the youngest children, it is likely insignificant for older children.
-Older children might instead need to weigh the potential social costs of crying or otherwise expressing sadness [@glasberg1981; @glasberg1982; @zeman1996] against the potential increase in social support.
+Older children might instead need to weigh the potential social costs of crying or otherwise expressing sadness [@glasberg1981; @glasberg1982; @zeman1996] against the potential increase in social support, which is consistent with our result that the age-related decline in signaling was steepest for crying.
+The age-related decline in signaling was also generally steeper for males than females [].
 
-Finally, based on previous research showing that family demands, such as caring for younger siblings, was an important source of parent-offspring conflict [@syme2023], we investigated the relationship between the frequency of alloparenting by children and their frequency of signaling.
+Previous research has shown family demands, such as caring for younger siblings (alloparenting), is an important source of parent-offspring conflict [@syme2023].
 Although there was little evidence for sex differences in signaling overall (perhaps because sex differences typically emerge in adolescence, and our sample of adolescents was small), there was evidence that sex interacted with alloparenting effort.
@@
 
-Many models designed to explain signals of need (e.g., the begging of young birds) predict that offspring who are worse off or are receiving less investment have more to gain from signaling than those with more [@mock2011; @godfray1991; @caro2016a; @godfray2000] and multiple game theoretical models examining parent-offspring conflict have found that parental resources have a modest impact on the evolutionary stable value for offspring demand for resources [@bossan2013; @mock1997].
-In line with this reasoning, children in lower income households and riskier environments might benefit more from signaling effort if they experience greater levels of adversity or perceive themselves as disadvantaged compared to other children.
+## Limitations
 
-## Limitations
+Although this study was broader than most in that it examined sadness, crying, and temper tantrums across much of childhood and adolescence, it does not cover all the ways children might signal need.
+Pained facial and bodily expressions [@vigil2014; @craig2009; @cano2010; @schiavenato2010], self-harm outside of temper tantrums [@hagen2008; @nock2008], and suicidality [@syme2016] have all been suggested to credibly signal need in times of conflicts of interest and may be viable strategies for children, at least at certain ages.
+Running away has received similar treatment, especially in instances when it comes with substantial risk [@shostak1981] or is seen as a suicide-threat [@fischer1950; @syme2023].
+For this reason, we also collected data on running away.
+However, mothers on Utila reported that running away on the island often involved children visiting neighbors or aunts and was much less risky than it would be on mainland Honduras, with a few mothers mentioning in focus groups that they warned their children about the difference in danger that came with running away on the mainland.
+Despite the relative safety of Utila, young children may still face risks out on their own due to proximity to the ocean and the presence of dogs, motorcycles, golf carts, and tuk-tuks (auto rickshaws) on the narrow roads typical of much of populated parts of the island.
+Due to uncertainty on how to conceptualize the risks involved, we did not include running away in our analyses.
 
@@
 
+## Signal frequency vs. type
+
+::: {#sfig-signaltype}
+```{r}
+plot(allEffects(m_freq_signal))
+```
+
+Effects plots of a negative binomial model of signal frequency as a function of signal type by child age and neighborhood quality.
+:::
+
 ```{r}

### 2025-04-22 Paper.qmd
Minor bib and paper edits.

@@
 Children with severe medical conditions also produce acoustically distinct cries, suggesting that some elements of crying are indexical [@soltis2004].
-Alternatively, crying might impose costs on caregivers by, e.g., attracting predators or interfering with important activities, motivating caregivers to invest more to stop the crying [so-called blackmail or bargaining models; @zahavi1977a; @eshel1991; @bergstrom1999 @kennedy2021].
+Alternatively, crying might impose costs on caregivers by, e.g., attracting predators or interfering with important activities, motivating caregivers to invest more to stop the crying [so-called blackmail or bargaining models; @zahavi1977a; @lummaa1998; @eshel1991; @bergstrom1999 @kennedy2021].
 
@@
 
-<!-- [One dense paragraph on how what we have learned relates to what has been reported empirically in the past]
-- Note: we did not mention declines with age in the introduction
-- Vingerhoets' lab and their colleagues maintain crying is understudied with older children
-
-[Move first sentence down to 468 and broaden? Or at least broaden? MacFarlane's study of tantrums among Berkeley: declines from 5 to 12]
-
-The cries and fussing of young infants increases until 6 weeks, peaking in some populations at over two hours a day, and then declines steadily until 12-16 weeks [@vermillet2022; @barr1991].
--->
-
 Caregivers reported a wide range of child signaling frequencies.
@@
 Although conflicts of interest do not always lead to overt behavioral conflict [@altmann1980; @bateson1994], it is likely many, if not most, conflicts are ultimately caused by conflicts of interest.
+
+["perceptions of need" instead of "perceived higher"?]
 
@@
 The age-related decline in signaling was also generally steeper for males than females [].
+
+[new additions above too repetive?]
 
@@
 
+::: {#sfig-signaltype}
+```{r}
+plot(allEffects(m_freq_signal))
+```
+
+Effects plots of a negative binomial model of signal frequency as a function of signal type by child age and neighborhood quality.
+:::
+
 ::: {#sfig-cause-clustering}
@@
 
-## Signal frequency vs. type
-
-::: {#sfig-signaltype}
-```{r}
-plot(allEffects(m_freq_signal))
-```
-
-Effects plots of a negative binomial model of signal frequency as a function of signal type by child age and neighborhood quality.
-:::
-
 ```{r}

### 2025-04-22 Paper.qmd
Minor paper edits.

@@
 To gain support, children use signals to communicate their needs and wants to parents.
-Infant signals of need, particularly infant cries, have been extensively studied in a range of populations.
+Infant signals of need, particularly infant cries, have been extensively studied in diverse populations.
 However, the full range of potential child signals of need, which extend beyond cries, has rarely been investigated in a single study of children of all ages.
-To fill this gap, we collected mother and other primary caregiver reports of child signaling from `{r} n_distinct(householdIDs)` families with `{r} nrow(utila_df)` children on Utila, a small island off the coast of Honduras.
+To help fill this gap, we collected mother and other primary caregiver reports of three common types of child signaling from `{r} n_distinct(householdIDs)` families with `{r} nrow(utila_df)` children on Utila, a small island off the coast of Honduras.
 In exploratory analyses, we found that child signaling was common in both sexes and across all ages, although it decreased with age and neighborhood quality and increased with the frequency of conflict between children and caretakers.
@@
 In particular, how and when older children signal need has rarely been investigated [for one exception, see @syme2023].
-The aim of this study was to help fill these gaps.
+The aim of this study was to help fill these gaps by investigating three common types of child signals.
 
@@
 
-Humans are a uniquely interdependent species among the living primates [@tomasello2012] in that both sex's fitness prospects depend greatly on a network of close kin and other social allies [@apicella2019].
+Humans are a uniquely interdependent species among the living primates [@tomasello2012] in that both sexes' fitness prospects depend greatly on a network of close kin and other social allies [@apicella2019].
 This need for others is most intense with children, who are born highly altricial and require many years of investment to develop to the point where they can obtain enough food to survive on their own and develop the social skills needed to succeed in their communities [@kaplan2009; @rosenberg2021; @schniter2015].
@@
 
-Sadness is often not accompanied by anger, suggesting little conflict with caregivers, and its costs are low when not prolonged.
-Temper tantrums, in constrast, are closely linked to anger [@potegal2003], suggesting conflict with caregivers, and have high costs even when the tantrum is brief.
-Thus, unlike sadness, temper tantrums might be useful when there are substantial and immediate gains to be had from a credible signal of need.
+Temper tantrums are closely linked to anger [@potegal2003], suggesting conflict with caregivers, and have high costs even when the tantrum is brief.
+Thus, temper tantrums might be useful when there are substantial and immediate gains to be had from a credible signal of need.
+This distinguishes them from sadness, and even crying, which are often unrelated to conflicts and have low costs when short-lived.
+
 The costs of temper tantrums could be energetic, or could come from the risk or implication of self-harm [@lummaa1998].
@@
 
-This study aimed to establish the natural history of child signaling in Utila, Honduras, by measuring various indices of adversity, children's conflicts with caregivers, the frequencies of common child signals of need, and various types of caregiver responses.
+This study aimed to establish the natural history of child signaling in Utila, Honduras, by measuring various indices of adversity, children's conflicts with caregivers, the frequencies of three common child signals of need, and various types of caregiver responses.
 From these, we used signaling theory to explore 1) the causes of child signals of need, 2) how and why signal costs might vary, 3) how caregivers respond to indications of child need, 4) how signaling behavior changes with age and sex, 5) and if there were differences in child signaling and parental responses in two subpopulations that face different levels of social and economic adversity.
@@
 Older children might instead need to weigh the potential social costs of crying or otherwise expressing sadness [@glasberg1981; @glasberg1982; @zeman1996] against the potential increase in social support, which is consistent with our result that the age-related decline in signaling was steepest for crying.
-The age-related decline in signaling was also generally steeper for males than females [].
-
-[new additions above too repetive?]
+The age-related decline in signaling was also generally steeper for males than females [@sfig-signaltype].
 
@@
 
-Although this study was broader than most in that it examined sadness, crying, and temper tantrums across much of childhood and adolescence, it does not cover all the ways children might signal need.
+Although this study was broader than most in that it examined sadness, crying, and temper tantrums across much of childhood and adolescence, it did not cover all the ways that caregivers might infer need, such as cues of injury.
+It also did not investigate all potential signals of need.
 Pained facial and bodily expressions [@vigil2014; @craig2009; @cano2010; @schiavenato2010], self-harm outside of temper tantrums [@hagen2008; @nock2008], and suicidality [@syme2016] have all been suggested to credibly signal need in times of conflicts of interest and may be viable strategies for children, at least at certain ages.

### 2025-04-22 Paper.qmd
Edits after full read through.

@@
 In particular, how and when older children signal need has rarely been investigated [for one exception, see @syme2023].
-The aim of this study was to help fill these gaps by investigating three common types of child signals.
+The aim of this study was to help fill these gaps by investigating three common types of signals of need in children and adolescents.
 
@@
 
-Although the fitness prospects of children and their familial caretakers are largely intertwined, evolutionary theory predicts that disagreements about the amount of investment from parents and other close kin will be ubiquitous [@fouts2005; @godfray1995; @godfray1992; @trivers1974; @bossan2013] because each child shares more genes through descent with themselves (100%) than with their full siblings (50% on average), half-siblings (25% on average), and stepsiblings (0%).
+Although the fitness prospects of children and their familial caretakers are largely intertwined, evolutionary theory predicts that disagreements about the amount of investment from parents and other close kin will be ubiquitous [@fouts2005; @godfray1995; @godfray1992; @trivers1974; @bossan2013] because each child shares more genes with themselves (100%) than (through recent common descent) with their full siblings (50% on average), half-siblings (25% on average), and stepsiblings (0%).
 Due to this asymmetry, children are often expected to attempt to elicit more investment from parents than is optimal from the perspective of the parent, while parents are expected to attempt to invest less than is optimal for each child in order to invest in their other children, whom parents are expected to value identically, all else being equal [@trivers1974; @bossan2013].
 
-For the same reason, siblings also often find themselves in conflicts of interest with one another, with each expected to prefer more resources for himself or herself than others.
+For the same reason, siblings have both confluences and conflicts of interest with one another, with each expected to prefer more resources for himself or herself than others.
 Sibling competition is evident across taxa and can range from relatively benign forms like strong begging competition in chicks [@caro2016] to siblicide in birds, mammals, and sharks [@salmon2014].
@@
 
-Humans are a uniquely interdependent species among the living primates [@tomasello2012] in that both sexes' fitness prospects depend greatly on a network of close kin and other social allies [@apicella2019].
+Humans are a uniquely interdependent species among the living primates [@tomasello2012] in that both sexes' fitness prospects depend greatly on a network of close kin and other social allies [@apicella2019; @syme2024].
 This need for others is most intense with children, who are born highly altricial and require many years of investment to develop to the point where they can obtain enough food to survive on their own and develop the social skills needed to succeed in their communities [@kaplan2009; @rosenberg2021; @schniter2015].
@@
 
-When the fitness interests of signalers and receivers are aligned, i.e., there are no conflicts of interest, child signals of need are expected to be low cost, what @maynardsmith1995 refer to as "minimal signals," those whose cost is no greater than that needed to transmit the information effectively.
+When the fitness interests of signalers and receivers are aligned, i.e., there are no conflicts of interest, child signals of need are expected to be what @maynardsmith1995 refer to as "minimal signals," signals whose cost is no greater than that needed to transmit the information effectively.
 In other words, although minimal signals might involve some cost (e.g., energy for a loud scream), they do not require any additional mechanism to enforce honesty.
@@
 With the information revealed by the signal, caretakers should generally be more likely to provide additional support due to their fitness interest in the child’s wellbeing.
-When this is the case, all parties are better off than they would be without the signal, with caretakers benefitting from the revelation of fitness-relevant information and the child benefitting from the increased likelihood of a beneficial response.
 
 However, even credible signals of need might not result in a beneficial response because belief in a child’s need is not the only factor determining the form and amount of support given by caregivers.
-Instead, caregivers are expected to weigh the costs and benefits of support against other options [@daly1987; @daly1983].
-Ancestrally, this likely often meant evaluating the viability of the child, which favored selection for child signals of vigor [@soltis2004].
-In general, selection favors allocation of parental investment to current and future offspring that equalizes marginal fitness returns to the parent [@clutton-brock1991; @daly1983; @dawkins1976; @hagen1999; @beaulieu2008; @frankenhuis2022; @daly1983].
+Instead, caregivers are expected to weigh the costs and benefits of support, which would take into account the reproductive value of current and future offspring, against other options [@daly1987; @daly1983], with selection favoring the allocation of parental investment to current and future offspring that equalizes marginal fitness returns to the parent [@clutton-brock1991; @daly1983; @dawkins1976; @hagen1999; @beaulieu2008; @frankenhuis2022; @daly1983].
 
-Additionally, a parent may also decide against investing in a child, despite deeming that investment worthwhile, if time is not right (e.g., if a mother preferred to breastfeed during rest rather than when foraging).
-When this is the case, outward conflicts may exaggerate the extent of a conflict of interest and involve parents teaching children ways to behave that benefit them both [@altmann1980].
+Additionally, a parent may also decide against investing in a child, despite deeming that investment worthwhile, if the time is not right (e.g., if a mother preferred to breastfeed during rest rather than when foraging).
+When this is the case, outward conflicts might not reflect genuine conflicts of interest and instead involve parents teaching children ways to behave that benefit them both [@altmann1980].
 Finally, caregivers often have better information about the costs and benefits of various items and actions, and therefore can make more beneficial decisions for a child than the child can herself (e.g., preventing a child from playing with matches).
@@
 Such expressions seem to be important to conveying this information as is seen by responses to those who cannot produce certain elements of facial expressions like tear production [@hasson2009; @nesse2019] or certain facial movements [@bogart2012; @bogart2010].
-We conceptualize infrequent and transient sadness as a minimal signal of need.
+We conceptualize infrequent and transient sadness signals as minimal signals of need [per @maynardsmith1995].
 Frequent or prolonged sadness, however, is likely a form of withdrawal or depression, both of which might be effective costly signals of need when there are conflicts between the child and their caregivers [@gaffney2022; @syme2023].
@@
 Although it is widely accepted that crying is a signal, there is no consensus on if or when it should be conceptualized as a minimal signal, a costly signal, an indexical signal, or a deceptive (manipulative) signal.
-Contact calls between parents and offspring have evolved in numerous species, so cries might simply be a means for infants to elicit attention from caregivers [@mehr2021], which would be type of minimal signal.
+Contact calls between parents and offspring have evolved in numerous species, so cries might simply be a means for infants to elicit attention from caregivers [@mehr2021], which would be a type of minimal signal.
 Demonstrating vigor and health early in life was likely vital for much of our evolutionary history due to the risk of neglect or infanticide experienced by children with poor survival prospects [@daly1984; @soltis2004].
@@
 Children with severe medical conditions also produce acoustically distinct cries, suggesting that some elements of crying are indexical [@soltis2004].
-Alternatively, crying might impose costs on caregivers by, e.g., attracting predators or interfering with important activities, motivating caregivers to invest more to stop the crying [so-called blackmail or bargaining models; @zahavi1977a; @lummaa1998; @eshel1991; @bergstrom1999 @kennedy2021].
+Alternatively, crying might impose costs on caregivers by, e.g., attracting predators or interfering with important activities, motivating caregivers to invest more to stop the crying [so-called blackmail or bargaining models, @zahavi1977a; @lummaa1998; @eshel1991; @bergstrom1999; @kennedy2021].
 
@@
 Such strategies can lead to some combination of verbal requests for more support and facial sadness, which we consider to be minimal signals, or withdrawal, temper tantrums and self-harm, which we consider "costly" signals.
-Crying might be considered a "minimal" signal in some circumstances, such as when a child needs a loud auditory signal to attract attention on a noisy playground, but a "costly" signal in others, such as when signaling vigor.
+Crying might be considered a "minimal" signal in some circumstances, such as when a child needs a loud auditory signal to attract attention on a noisy playground, but a "costly" signal in others, such as lengthy bouts that incur physiological costs in infants signaling vigor and social costs in older children signaling need.
 Similarly, infrequent sadness would be a minimal signal, but frequent sadness might be better conceptualized as costly withdrawal or depression.
-Each of these might be effective in certain circumstances and are sometimes successful in eliciting more support from caretakers.
 
@@
 
-This study of children aged `{r} paste(range(utila_df$ChildAge), collapse="-")` and their parents and other primary caretakers, was conducted in Utila, Honduras, one of the country’s small Bay Islands in the Western Caribbean.
-Participants in the study were sampled from two populations who face different levels of economic and social adversity: 1) a long-standing English-speaking community primarily comprising descendants of British immigrants who arrived on the island in the 1800’s, and 2) a more recent community of Spanish-speakers, mostly from mainland Honduras.
+This study of children aged `{r} paste(range(utila_df$ChildAge), collapse="-")` and their parents and other primary caretakers was conducted in Utila, Honduras, one of the country’s small Bay Islands in the Western Caribbean.
+Participants in the study were sampled from two populations that face different levels of economic and social adversity: 1) a long-standing English-speaking community primarily comprising descendants of British immigrants who arrived on the island in the 1800’s, and 2) a more recent community of Spanish-speakers, mostly from mainland Honduras.
 Compared to the former, the latter are more likely to live in crowded and suboptimal conditions, face active discrimination and increased stress [@garcia2017], and experience economic uncertainty due the predominance of low paying, seasonal jobs primarily related to the island’s diving tourism industry [@garcia2018; @stonich2000].
@@
 Both Spanish-speaking and English-speaking moms reported that fighting between siblings was extremely common, with fights over toys, gifts, and phone time being the most reported causes.
-In both populations, crying and various forms of withdrawal were listed as the most common responses to a child not receiving investment or support, followed by temper tantrums and making messes in their home.
-Spanish-speaking mothers reported that refusing to do homework was a common tactic, while English-speaking mothers emphasized running away and self-harm (e.g., a child slamming their head on the ground) to a greater degree.
+In both populations, crying and various forms of sadness and withdrawal were listed as the most common responses to a child not receiving investment or support, followed by temper tantrums.
 In response to these hypothesized signals, discipline was common.
@@
 See @tbl-freqvars for exact wording.
-Caretakers were often not asked about child alloparenting behavior if a child was an only child.
-For this reason, we gave children an alloparenting frequency of “never” if they were the only child within a household.
+We gave children an alloparenting frequency of “never” if they were the only child within a household.
 We also calculated the summed frequency of child alloparenting contributions by other children within the household.
@@
 All questions had the following labels attached to each point on its Likert scale: Never; Once a month or less; More than once a month but less than once a week; More than once a week but not daily; Daily; Multiple times per day.
-For analyses, we converted these categorical variables to numeric frequencies per month as follows: never = 0; once a month or less = 1; more than once a month = 3; more than once a week = 8; daily = 30; multiple times per day = 60.
-{#tbl-freqvars}
+For analyses, we converted these categorical variables to numeric frequencies per month as follows: never = 0; once a month or less = 1; more than once a month = 3; more than once a week = 8; daily = 30; multiple times per day = 60 {#tbl-freqvars}
 
@@
 We reversed the scores on two measures so that a score of 5 corresponded with the lowest susceptibility to illness before we averaged the scores of these to create a mean *illness avoidance* variable.
-In the `{r} sum((!is.na(utila_df$IllnessSusceptibility2)) & is.na(utila_df$IllnessSusceptibility3))` instance in which we only had data for 2 of the measures, we still calculated the mean but used a numerator of 2.
+In the `{r} sum((!is.na(utila_df$IllnessSusceptibility2)) & is.na(utila_df$IllnessSusceptibility3))` instance in which we only had data for two of the measures, we still calculated the mean but used a numerator of 2.
 See @stbl-healthvars for the wording of these variables.
@@
 
-A)  Loadings of PC1 and PC2 on each variable. B) Biplot map of PC1 and PC2. Each point is one child. Arrows represent the loadings of PC1 and PC2 on each variable.
+a) Loadings of each variable on PC1 and PC2.
+b) Biplot of PC1 and PC2.
+Each point is one child.
+Arrows represent the loadings of each variable on PC1 and PC2.
+Variables with small loading factors not displayed
 :::
@@
 Line weights represent partial correlation coefficients.
-Blue text represents our interpretation of the main clusters.
+Blue text represents our interpretation of the main clusters
 :::
@@
 However, many children were reported to never engage in these signals/states; sadness: `{r} fmt_sum(utila_df$SadFreqN == 0)`; crying; `{r} fmt_sum(utila_df$CryFreqN == 0)`; and tantrums: `{r} fmt_sum(utila_df$TantrumFreqN == 0)`.
-For a visualization of how signal frequencies varied within children see @sfig-signal-alluvial-plot.
+For a visualization of how signal frequencies varied within children, see @sfig-signal-alluvial-plot.
 
@@
 Frequencies of child sadness, crying, and temper tantrums per month as reported by the primary caretaker.
-Frequency key: 0 corresponds to "never," \<=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and \>30 corresponds to "multiple times per day."
+Frequency key: 0 corresponds to "never," \<=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and \>30 corresponds to "multiple times per day"
 :::
@@
 Correlation matrix of the frequencies of signals/states covered in the study with sex, child age, and the frequency of conflict between children and primary caretakers.
-For a more complete correlation matrix of study variables see @sfig-plot-full-corrmat.
+For a more complete correlation matrix of study variables see @sfig-plot-full-corrmat
 :::
@@
 
-Lasso regression models of each child signaling frequency and conflict frequency. Variables whose coefficient values were shrunk to 0 are not displayed. Models fit using quasi-Poisson regression.
+Lasso regression models of each child signaling frequency and conflict frequency. Variables whose coefficient values were shrunk to 0 are not displayed. Models fit using quasi-Poisson regression
 :::
@@
 
-Effects of selected predictors of child signaling frequencies when holding continuous variables at their mean and factor variables at their mode.
+Effects of selected predictors of child signaling frequencies when holding continuous variables at their mean and factor variables at their mode
 :::
@@
 
-Effects of selected predictors of child-parent conflict frequencies when holding continuous variables at their mean and factor variables at their mode.
+Effects of selected predictors of child-parent conflict frequencies when holding continuous variables at their mean and factor variables at their mode
 :::
@@
 The proportion of children perceived by caretakers as being more needy, as needy, or less needy than a caretaker's other children when holding continuous variables at their mean and factor variables at their mode.
-A: Effect of child age on relative neediness B: Effect child sadness frequency on relative neediness.
+A: Effect of child age on relative neediness B: Effect of child sadness frequency on relative neediness.
 Outcome variable question wording: How much time, effort, and resources does this child need compared to your other children?
@@
 Categories of causes were based on cluster analysis of more fine-grained cause variables.
-See @sfig-cause-clustering for cluster dendrogram and @stbl-sumtab-causes for the frequencies of causes and caregiver responses.
+See @sfig-cause-clustering for cluster dendrogram and @stbl-sumtab-causes for the frequencies of causes and caregiver responses
 :::
@@
 Dot size represents model predictions for each participant given their values of predictor variables.
-See @sfig-plot-caregiver-punish for the effect of the cause involving punishment.
+See @sfig-plot-caregiver-punish for the effect of the cause involving punishment
 :::
@@
 
-Our bivariate correlations, PCA, MST, and lasso regression analyses all provided a consistent pattern of results: child age was a strong negative predictor of signaling frequency and cost; conflict with the primary caretaker was a positive predictor of signaling frequency and cost; and multiple indices of adversity, such as low neighborhood quality, low household income, and medical problems, were positive predictors of signaling frequency and cost (@fig-mst, @fig-PCA, @fig-signal-corrplot, @fig-lasso, @fig-signal-effects).
+Our bivariate correlations, PCA, MST, and lasso regression analyses all provided a consistent pattern of results: child age was a strong negative predictor of signaling frequency and cost; conflict with the primary caretaker was a positive predictor of signaling frequency and cost; and multiple indices of adversity, such as low neighborhood quality, low household income, and medical problems, were positive predictors of signaling frequency and cost (@fig-PCA,, @fig-mst, @fig-signal-corrplot, @fig-lasso, @fig-signal-effects).
 
@@
 
-["perceptions of need" instead of "perceived higher"?]
-
-Consistent with the need signaling hypothesis, more frequent sadness was positively associated with perceived higher need relative to other children (@fig-effect-relativeneed).
+Consistent with the need signaling hypothesis, more frequent sadness was positively associated with perceptions of relative child need within the household (@fig-effect-relativeneed).
 Higher perceived need, in turn, was associated with higher relative investment (@fig-plot-relative-investment).
@@
 The frequency of temper tantrums also declined, as did the frequency of child sadness, albeit at a much slower rate (@fig-lasso, @fig-signal-effects).
-Our results also align with studies examining responses to the crying and fussing of infants that have shown that they often lead to increased support [e.g., breastfeeding, attention, or maintaining proximity; @lin2012; @kushnick2006; @kruger2010; @hewlett1998; @chaudhary2024; @tronick1987], but can also lead  to negative responses like abuse [@zeevenhooven2018; @kurth2011; @howard2006; @soltis2004].
+Our results also align with studies examining responses to the crying and fussing of infants that have shown that they often lead to increased support [e.g., breastfeeding, attention, or maintaining proximity, @lin2012; @kushnick2006; @kruger2010; @hewlett1998; @chaudhary2024; @tronick1987], but can also lead  to negative responses like abuse [@zeevenhooven2018; @kurth2011; @howard2006; @soltis2004].
 
@@
 
+<!-- Move paragrpah below to 461? -->
 One possible reason for the negative association of signaling with age, which is also seen in more industrialized settings [@zeifman2013; @potegal2003; @koch2003; @macfarlane1954], is that as children mature, they are more capable of fulfilling their own needs, more able to ask others for what they need [@osterman2010], and less vulnerable to threats.
@@
 Older children might instead need to weigh the potential social costs of crying or otherwise expressing sadness [@glasberg1981; @glasberg1982; @zeman1996] against the potential increase in social support, which is consistent with our result that the age-related decline in signaling was steepest for crying.
-The age-related decline in signaling was also generally steeper for males than females [@sfig-signaltype].
+The age-related decline in signaling was also generally steeper for males than females (@sfig-signaltype).
 
@@
 It also did not investigate all potential signals of need.
+For example, Spanish-speaking mothers reported in focus groups that refusing to do homework was a common tactic.
 Pained facial and bodily expressions [@vigil2014; @craig2009; @cano2010; @schiavenato2010], self-harm outside of temper tantrums [@hagen2008; @nock2008], and suicidality [@syme2016] have all been suggested to credibly signal need in times of conflicts of interest and may be viable strategies for children, at least at certain ages.
 Running away has received similar treatment, especially in instances when it comes with substantial risk [@shostak1981] or is seen as a suicide-threat [@fischer1950; @syme2023].
-For this reason, we also collected data on running away.
-However, mothers on Utila reported that running away on the island often involved children visiting neighbors or aunts and was much less risky than it would be on mainland Honduras, with a few mothers mentioning in focus groups that they warned their children about the difference in danger that came with running away on the mainland.
+English-speaking mothers in our focus groups emphasized running away and self-harm (e.g., a child slamming their head on the ground).
+However, this often involved children visiting neighbors or aunts and was much less risky than it would be on mainland Honduras.
 Despite the relative safety of Utila, young children may still face risks out on their own due to proximity to the ocean and the presence of dogs, motorcycles, golf carts, and tuk-tuks (auto rickshaws) on the narrow roads typical of much of populated parts of the island.
@@
 
-The main limitation of this study is that it was exploratory: our data-driven analyses, which can reveal unexpected relationships and inspire new theory, also risked overfitting the data, suggesting associations that might not exist.
-Our study was also cross-sectional, further limiting our ability to make causal inferences.
+The main limitation of this study is that all data come from caregiver reports, and therefore, reflect the biases of caretakers, whatever those might be.
+For example, caregiver reports of their own responses to recent signals of need might be positively biased.
+In addition, questions about neighborhood quality lacked a reference, so some caregivers might have compared their neighborhoods to mainland neighborhoods and others to island neighborhoods.
+
+Another important limitation is that this study was exploratory: such data-driven analyses can reveal unexpected relationships and inspire new theory but also risk overfitting the data, suggesting associations that might not exist.Our study was also cross-sectional, further limiting our ability to make causal inferences.
 In particular, there is likely a bidirectional relationship between child signaling and caregivers' perceptions of conflict that our study design could not disentangle.
@@
 This could happen if there was selection bias so that those who participated differed from the rest of the population in how they interacted with their children or how they assessed, remembered, and reported child behavior.
-All data come from caregiver reports, and therefore, reflect the biases of caretakers, whatever those might be.
-For example, caregiver reports of their own responses to recent signals of need might be positively biased.
-In addition, questions about neighborhood quality lacked a reference, so some caregivers might have compared their neighborhoods to mainland neighborhoods and others to island neighborhoods.
 

### 2025-05-12 Paper.qmd
Paper edits based on co-author feedback.

@@
         department: Institute of Security and Global Affairs
+  - name: Steven A. Arnocky
+    orcid: 0000-0003-3812-1871
+    corresponding: false
+    affiliation:
+      - name: Nipissing University
+        department: Psychology
   - name: Aaron D. Blackwell
@@
 Consistent with signaling theory, children who were sad more frequently were perceived as needier within the household and were more likely to receive investment.
-Caregivers were less likely to respond positively in situations of family conflict or child transgressions, and more likely for injuries and illnesses.
+Caregivers were less likely to respond positively in situations of family conflict or child transgressions, and more likely for injuries and illness.
 Our results suggest that evolutionary theories of signaling can help explain patterns of child sadness, crying, and temper tantrums.
@@
 
-Compared to chimpanzee mothers, mothers in hunter-gatherer societies have shorter interbirth intervals and produce offspring that require many more years of parental care [@davison2021; @davison2022].
+Compared to chimpanzee mothers, human mothers in hunter-gatherer societies have shorter interbirth intervals and produce offspring that require many more years of parental care [@davison2021; @davison2022].
 Unlike chimpanzee offspring, human offspring in ancestral populations therefore often grew up with multiple older and/or younger siblings who were potential competitors for parental investment and yet who could also provide (or require) allocare [@helfrecht2016; @lawson2016; @hagen2009].
@@
 
-While parent-offspring conflict and sibling competition are likely ubiquitous, the resources being contested will vary across human groups.
+Whereas parent-offspring conflict and sibling competition are likely ubiquitous, the resources being contested will vary across human groups.
 In relatively egalitarian and food constrained populations, conflict over resources is likely to be centered on food, which probably explains the negative impacts siblings have on child age-specific heights and weights among the Shuar of Ecuador [@hagen2009], although similar effects have also been found in industrialized populations [@lawson2008].
@@
 Conflicts of interest over investment in the family can even involve adult children who are starting to have their own children [@kushnick2023; @emlen1995].
-Parents can potentially benefit from delaying the reproductive careers of certain offspring in order to gain more allocare for their other children [@flinn1989; @emlen1992] or speeding up a child’s departure from the household in order to provide more parental investment to other children [@kushnick2023].
+Parents can potentially benefit from delaying the reproductive careers of certain offspring to gain more allocare for their other children [@flinn1989; @emlen1992] or speeding up a child’s departure from the household to provide more parental investment to other children [@kushnick2023].
 
@@
 Children often have needs and wants, where needs involve avoiding potential fitness costs, such as poor nutrition, injuries, or threats to status, and wants involve acquiring potential fitness benefits, such as highly nutritious foods and valuable artifacts.
-Needs and wants (hereafter needs) are often not apparent to caregivers, however, i.e., they are private information, and hence children must signal them to receive help.
+Needs and wants (hereafter needs) are often not apparent to caregivers, i.e., they are private information, and hence children must signal them to receive help.
 Signals differ from cues in that signals were favored by natural selection due to their ability to alter the behavior or perceptions of others in ways that benefit the signaler, on average, whereas cues are information organisms use to influence their behavior or physiological state that did not evolve for communicative effect [@maynardsmith1995].
@@
 Sadness and low mood, like other forms of psychological pain, probably evolved to motivate victims of adversity to shift their attention to the causes of adversity to mitigate its negative fitness consequences and to learn to solve or avoid similar events in the future [@andrews2009; @nesse1990; @thornhill1989].
-Low mood, for example, is often thought to be adaptive in situations when activity is dangerous or unlikely to result in benefits, with anhedonia, pessimism, and fatigue all helping reduce behavior that may not be worth the costs or that may cause additional problems [@nesse2019; @nettle2009].
+Low mood, for example, is often thought to be adaptive in situations when activity is dangerous or unlikely to result in benefits, with anhedonia, pessimism, and fatigue all helping reduce behavior that may not be worth the costs or that may cause additional problems [@nesse2019; @nettle2009; @trimmer2015].
 It may also encourage the reappraisal of one’s circumstances after new information comes to light that one’s current strategy is not working [@nesse2009].
@@
 In addition to this cue-based information, sadness also results in facial expressions and crying, both of which probably evolved to signal need [@balsters2013; @bowlby1980; @reed2017; @darwin1872].
-Such expressions seem to be important to conveying this information as is seen by responses to those who cannot produce certain elements of facial expressions like tear production [@hasson2009; @nesse2019] or certain facial movements [@bogart2012; @bogart2010].
+Such expressions seem to be important to conveying this information as is seen by the skeptical responses to those who cannot produce certain elements of facial expressions like tear production [@hasson2009; @nesse2019] or certain facial movements [@bogart2012; @bogart2010].
 We conceptualize infrequent and transient sadness signals as minimal signals of need [per @maynardsmith1995].
@@
 Similarly, !Kung San mothers frequently respond to crying infants by providing comfort and breastfeeding access, with comforting behaviors being far more common than behaviors restricting infants [@kruger2010].
-Although @kruger2010 found mothers were the most common responders, siblings, fathers, and other adults frequently stepped in so that 88% of crying bouts were responded to, with the majority of the cries ignored lasting under ten seconds.
-Similar responsiveness has also been reported among the Aka [@hewlett1998], the Mbendjele [@chaudhary2024], the Efe [@tronick1987], and the Ache, among whom young children have been described as “win\[ning\] most conflicts with parents simply by crying and whining” [@kaplan1987].
+Although @kruger2010 found mothers were the most common responders, siblings, fathers, and other adults frequently stepped in so that 88% of crying bouts were responded to, with most of the cries ignored lasting under ten seconds.
+Similar responsiveness has also been reported among the Aka [@hewlett1998], the Mbendjele [@chaudhary2024], the Efe [@tronick1987], and the Aché, among whom young children have been described as “win\[ning\] most conflicts with parents simply by crying and whining” [@kaplan1987].
 Studies examining the potential for long-term benefits are rare and difficult to achieve.
@@
 It has been found to increase parental distress, the likelihood of maternal depression, and the child’s risk of maltreatment [@frodi1985; @fujiwara2011; @howard2006; @lin2012; @soltis2004].
-For example, in a sample of 3,259 parents of Dutch infants under 6 months old, 5.6% self-reported smothering, slapping, or shaking their babies in response to their crying, outcomes that were more likely among parents who were single, unemployed, or deemed their infant’s crying to be excessive [@reijneveld2004].
+For example, in a sample of 3,259 Dutch parents with infants under 6 months old, 5.6% self-reported smothering, slapping, or shaking their babies in response to their crying, outcomes that were more likely among parents who were single, unemployed, or deemed their infant’s crying to be excessive [@reijneveld2004].
 Excessive infant crying, i.e., colic, has also been found to be a risk factor for multiple negative outcomes [@zeevenhooven2018], including abuse, shaken baby syndrome [@kurth2011], and earlier cessation of breastfeeding [@howard2006].
@@
 
-The costs of temper tantrums could be energetic, or could come from the risk or implication of self-harm [@lummaa1998].
+The costs of temper tantrums could be energetic or come from the risk or implication of self-harm [@lummaa1998].
 For example, Utilian mothers reported in focus groups that some temper tantrums involved children banging their heads on things when they do not get what they want, behaviors that have been reported in other populations [@belden2008; @burrows1953; @daniels2012; @dennis1940; @vandenakker2022].
@@
 A major challenge when studying signals of need is the fact that such signaling is expected to 1) be more common in times when individuals can benefit from increased investment (e.g., when they are facing major adversity or when a fitness-relevant opportunity arises) and 2) result in benefits to the signaler on average.
-This hypothesized causal, bi-directional relationship between investment and signaling makes it difficult to isolate the effects of signaling behavior on investment in children, especially given the lack of prior studies on signaling by older children in low income settings.
+This hypothesized causal, bi-directional relationship between investment and signaling makes it difficult to isolate the effects of signaling behavior on investment in children, especially given the lack of prior studies on signaling by older children in low-income settings.
 For this reason, all analyses of these unique data were exploratory and meant to lay the foundation for future studies of child signaling.
@@
 
-To inform development of our survey questionnaire on child signaling, focus groups of Utilian parents were held in July 2022.
+To inform development of our survey questionnaire on child signaling, focus groups of parents of families living in Utila were held in July 2022.
 Parents indicated frequent sibling competition and various ways children might signal need.
@@
 
-We used six-point Likert scales to get caregiver reports on the frequencies of 1) various types of child signaling, such as sadness and crying, 2) conflict between the caretaker and each participating child, 3) and child allocare of other children in their household.
+We used six-point Likert-type scales to get caregiver reports on the frequencies of 1) various types of child signaling, such as sadness and crying, 2) conflict between the caretaker and each participating child, 3) and child allocare of other children in their household.
 See @tbl-freqvars for exact wording.
@@
 Caretakers were asked about the frequency of their child missing school and their child's history of medical problems in general terms (i.e., the extent to which they experienced a "variety of medical issues").
-Both involved the use of five-point Likert scales, the scores of which were averaged for use in analyses.
+Both questions were slightly modified from @hill2016 and involved the use of five-point Likert scales, the scores of which were averaged for use in analyses.
 See @stbl-medical-record for the wording of these variables.
 
-Caretakers were also asked three questions using a five point Likert scale about their child's susceptibility to illness.
+Caretakers were also asked three questions using a five-point Likert scale about their child's susceptibility to illness, all of which were slightly modified versions of questions from the Perceived Vulnerability to Disease Scale [@duncan2009].
 We reversed the scores on two measures so that a score of 5 corresponded with the lowest susceptibility to illness before we averaged the scores of these to create a mean *illness avoidance* variable.
@@
 
-Caregiver perceptions of neighborhood quality were measured using ten questions relating to the safety and suitability of their neighborhood for raising children in addition to questions explicitly asking about neighborhood quality, e.g., "Sometimes I hear about illegal drug activity in my neighborhood." The mean of these values was used in analyses as a neighborhood quality index.
+Caregiver perceptions of neighborhood quality were measured using ten questions modified from the Neighborhood Characteristics Questionnaire [@mcguire1997].
+These questions covered the safety and suitability of their neighborhood for raising children and included questions explicitly asking about neighborhood quality, e.g., "Sometimes I hear about illegal drug activity in my neighborhood." The mean of these values was used in analyses as a neighborhood quality index.
 Levels of home instability were measured with four questions covering the degree to which the caregiver's family had/has to move homes and whether people move in and out of their homes frequently.
-For wording of the these items, which used five-point Likert scales, see @stbl-Neighborhood-Quality and @stbl-Home-Instability.
+For wording of these items, which used five-point Likert scales, see @stbl-Neighborhood-Quality and @stbl-Home-Instability.
 
@@
 The overall structure of our data was analyzed with principal components analysis (PCA), Bayesian graphical models (BGM), and minimum spanning trees (MST).
-PCA is a linear transformation of the data to a new orthonormal coordinate system in which the the first principal component (PC1) corresponds to the axis of maximum variance, the second principal component (PC2) corresponds to the axis perpendicular to the first of the maximum of the remaining variance, and so forth.
+PCA is a linear transformation of the data to a new orthonormal coordinate system in which the first principal component (PC1) corresponds to the axis of maximum variance, the second principal component (PC2) corresponds to the axis perpendicular to the first of the maximum of the remaining variance, and so forth.
 Among other uses, PCA identifies the linear combinations of variables that account for most of the variance, thereby providing a lower dimensional representation of high dimensional data.
@@
 
-Elasticnet regression models of signaling and conflict frequency, which were overdispered count data, were fit using the glmnet package [@friedman2010; @wurm2021] and the quasipoisson error family.
+Elasticnet regression models of signaling and conflict frequency, which were overdispersed count data, were fit using the glmnet package [@friedman2010; @wurm2021] and the quasipoisson error family.
 Predictors included demographic, household, health, and neighborhood variables that might cause signaling behavior or conflict between children and caretakers.
@@
 Line weights represent partial correlation coefficients.
+Red dots indicate the two variables (child age and conflict) that are closest to the signaling variables.
 Blue text represents our interpretation of the main clusters
@@
 Therefore, we collapsed “less” and “much less” into a “less” factor and “more” and “much more” into a “more” factor.
-The end result was `{r} rel_need_table[1]` less needy children, `{r} rel_need_table[2]` children who were as needy as their siblings, and `{r} rel_need_table[3]` more needy children.
+The result was `{r} rel_need_table[1]` less needy children, `{r} rel_need_table[2]` children who were as needy as their siblings, and `{r} rel_need_table[3]` more needy children.
 
@@
 The proportion of children perceived by caretakers as being more needy, as needy, or less needy than a caretaker's other children when holding continuous variables at their mean and factor variables at their mode.
-A: Effect of child age on relative neediness B: Effect of child sadness frequency on relative neediness.
+A: Effect of child age on the probability of relative neediness B: Effect of child sadness frequency on the probability of relative neediness.
 Outcome variable question wording: How much time, effort, and resources does this child need compared to your other children?
@@
 
-Signals are capable of eliciting better treatment because the information they provide alters the behavior of signal targets in ways that benefit the signaler.
+Signals can elicit better treatment because the information they provide alters the behavior of signal targets in ways that benefit the signaler.
 For this reason, we tested the association between caretaker perceptions of relative child need within the household and the caretaker's relative investment within the household.
@@
 
-The effect of perceived need (in terms of time, effort, and resources) relative investment within the household when holding continuous variables at their mean and factor variables at their mode.
+The effect of perceived need (in terms of time, effort, and resources) on the probability of relative investment within the household when holding continuous variables at their mean and factor variables at their mode.
 Outcome variable question wording: How much time, effort, and resources do you put into this child compared to your other children?
@@
 
-Our models found no evidence of relationships between, child age, child sex, the number of children in the household, and log family income and the valence of parental reponses (positive, neutral, or negative) from the child's perspective.
-However, primary caretakers were more likely to respond positively towards a child when they attributed the cause of signaling to discomfort, pain, injury, or illness, and less likely to respond positively when they attributed the cause of signaling to family conflict, child punishment, child trangressions, or a child losing an item or privleges.
+Our models found no evidence of relationships between, child age, child sex, the number of children in the household, and log family income and the valence of parental responses (positive, neutral, or negative) from the child's perspective.
+However, primary caretakers were more likely to respond positively towards a child when they attributed the cause of signaling to discomfort, pain, injury, or illness, and less likely to respond positively when they attributed the cause of signaling to family conflict, child punishment, child transgressions, or a child losing an item or privileges.
 See @fig-effect-last-signal-response.
@@
 
-Effects of selected predictors of the likelihood of positive, neutral, and negative responses to the last time a child became sad or cried when holding continuous variables at their mean and factor variables at their mode.
+Effects of selected predictors of the probability of positive, neutral, and negative responses to the last time a child became sad or cried when holding continuous variables at their mean and factor variables at their mode.
 1 = evidence of the cause in the caregiver report; 0 = no evidence.
@@
 
-Our bivariate correlations, PCA, MST, and lasso regression analyses all provided a consistent pattern of results: child age was a strong negative predictor of signaling frequency and cost; conflict with the primary caretaker was a positive predictor of signaling frequency and cost; and multiple indices of adversity, such as low neighborhood quality, low household income, and medical problems, were positive predictors of signaling frequency and cost (@fig-PCA,, @fig-mst, @fig-signal-corrplot, @fig-lasso, @fig-signal-effects).
+Our bivariate correlations, PCA, MST, and lasso regression analyses all provided a consistent pattern of results: child age was a strong negative predictor of signaling frequency and cost; conflict with the primary caretaker was a positive predictor of signaling frequency and cost; and multiple indices of adversity, such as low neighborhood quality, low household income, and medical problems, were positive predictors of signaling frequency and cost (@fig-PCA, @fig-mst, @fig-signal-corrplot, @fig-lasso, @fig-signal-effects).
 
@@
 In line with this reasoning, multiple caretakers mentioned that older kids need more money and advice and get into trouble more often.
-It might also be that caretakers see investing in older children as relatively costly when they seek to redirect investment to younger children.
+It might also be that caretakers see investing in older children as relatively costly when they seek to redirect investment to younger children, whose needs might be perceived to be more legitimate.
 
@@
 Our results parallel those of previous studies of infants and young children that found that, outside of peaks in infancy and early childhood, the frequencies of crying and temper tantrums decrease with age [@macfarlane1954; @ward1970; @osterman2010; @vermillet2022; @barr1991; @potegal2003; @bhatia1990].
-Studies examining crying and tantrums in later childhood and adolensce are rare, especially for crying [@rottenberg2012; @zeifman2013].
+Studies examining crying and tantrums in later childhood and adolescence are rare, especially for crying [@rottenberg2012; @zeifman2013].
 Our findings suggest the frequency of crying continues to decrease into childhood, dropping precipitously as children grow older.
@@
 <!-- Move paragrpah below to 461? -->
-One possible reason for the negative association of signaling with age, which is also seen in more industrialized settings [@zeifman2013; @potegal2003; @koch2003; @macfarlane1954], is that as children mature, they are more capable of fulfilling their own needs, more able to ask others for what they need [@osterman2010], and less vulnerable to threats.
+One possible reason for the negative association of signaling with age, which is also seen in more industrialized settings [@zeifman2013; @potegal2003; @koch2003; @macfarlane1954], is that as children mature, they are more capable of fulfilling their own needs, more able to verbalize what they need [@osterman2010], and less vulnerable to threats.
 The effects of age on signaling are also likely due to how age influences the costliness of signaling need.
-For example, while the energetic cost of crying might be costly for the youngest children, it is likely insignificant for older children.
+For example, although the energetic cost of crying might be costly for the youngest children, it is likely insignificant for older children.
 Older children might instead need to weigh the potential social costs of crying or otherwise expressing sadness [@glasberg1981; @glasberg1982; @zeman1996] against the potential increase in social support, which is consistent with our result that the age-related decline in signaling was steepest for crying.
@@
 We found that higher frequencies of alloparenting by girls was associated with a higher frequency of tantrums and overall signaling costs, whereas for boys this relationship was slightly negative (@sfig-alloparentingXsex).
-In many societies, girls do more alloparenting than boys [@helfrecht2016; @weisner1977; @barry1971; @turke1988], which matches reports from our focus groups of Utilian mothers.
+In many societies, girls do more alloparenting than boys [@helfrecht2016; @weisner1977; @barry1971; @turke1988], which matches reports from our focus groups of mothers.
 It is plausible that older girls are pressured to do more care of their younger siblings, in line with other studies finding an association between household responsibilities, like caregiving, and parent-offspring conflict [@lowe2003; @syme2023].
 If so, girls with excessive alloparenting responsibilities might be signaling their need for more time to invest in peer relationships and status seeking behaviors outside the household.
-Opposed to this interpretation, our study found a small negative association between girls' alloparenting and caregiver reports of conflict (@fig-lasso), although plausibly when girls do what caregivers want, caregivers report less conflict.
+Although our study found a small negative association between girls' alloparenting and caregiver reports of conflict (@fig-lasso), when girls do what caregivers want, caregivers plausibly report less conflict.
 
@@
 Running away has received similar treatment, especially in instances when it comes with substantial risk [@shostak1981] or is seen as a suicide-threat [@fischer1950; @syme2023].
-English-speaking mothers in our focus groups emphasized running away and self-harm (e.g., a child slamming their head on the ground).
-However, this often involved children visiting neighbors or aunts and was much less risky than it would be on mainland Honduras.
-Despite the relative safety of Utila, young children may still face risks out on their own due to proximity to the ocean and the presence of dogs, motorcycles, golf carts, and tuk-tuks (auto rickshaws) on the narrow roads typical of much of populated parts of the island.
+English-speaking mothers in our focus groups emphasized self-harm (e.g., a child slamming their head on the ground) and running away.
+However, in Utila running away often involves children visiting neighbors or aunts and is much less risky than it would be in many parts of mainland Honduras [@davila2024].
+Despite the relative safety of Utila, young children may still face risks out on their own due to proximity to the ocean and the presence of dogs, motorcycles, golf carts, and tuk-tuks (auto rickshaws) on the narrow roads typical of much of the populated parts of the island.
 Due to uncertainty on how to conceptualize the risks involved, we did not include running away in our analyses.
 
-The main limitation of this study is that all data come from caregiver reports, and therefore, reflect the biases of caretakers, whatever those might be.
-For example, caregiver reports of their own responses to recent signals of need might be positively biased.
+The main limitation of this study is self-report bias: all data come from caregiver reports, and therefore, reflect the biases of caretakers, whatever those might be.
+For example, caregivers might have reported more investment in children or fairer distribtions of resources than was actually the case.
 In addition, questions about neighborhood quality lacked a reference, so some caregivers might have compared their neighborhoods to mainland neighborhoods and others to island neighborhoods.
 
-Another important limitation is that this study was exploratory: such data-driven analyses can reveal unexpected relationships and inspire new theory but also risk overfitting the data, suggesting associations that might not exist.Our study was also cross-sectional, further limiting our ability to make causal inferences.
+Another important limitation is that this study was exploratory: such data-driven analyses can reveal unexpected relationships and inspire new theory but also risk overfitting the data, suggesting associations that might not exist.
+Our study was also cross-sectional, further limiting our ability to make causal inferences.
 In particular, there is likely a bidirectional relationship between child signaling and caregivers' perceptions of conflict that our study design could not disentangle.
@@
 Children face the challenge of eliciting support from others in times of adversity and opportunity.
-On the island of Utila, child signaling strategies were highly correlated with each other, and were exhibted by children of all ages, albeit more frequently by younger children and those living in more adverse conditions, who plausibly had greater need.
+On the island of Utila, child signaling strategies were highly correlated with each other, and were exhibited by children of all ages, albeit more frequently by younger children and those living in more adverse conditions, who plausibly had greater need.
 Consistent with signaling models that emphasize the role of conflicts of interest in favoring the use of costlier signals, child signaling effort was positively correlated with the frequency of conflict with caretakers and with perceptions of need, and greater perceptions of child need predicted greater investment in a given child.
@@
 
+::: {#sfig-plot-full-corrmat}
+![](Figures/plot_fullcorrmat.png)
+
+Complete correlation matrix of study variables.
+:::
+
 ::: {#sfig-alloparentingXsex}
@@
 
-::: {#sfig-plot-full-corrmat}
-![](Figures/plot_fullcorrmat.png)
-
-Complete correlation matrix of study variables.
-:::
-
 ::: {#sfig-signaltype}
 ```{r}
-plot(allEffects(m_freq_signal))
+#| fig-height: 8
+# plot(allEffects(m_freq_signal))
+plot(Effect(c("Signal", "ChildAge"), mod=m_freq_signal), cols = 3)
+plot(Effect(c("Signal", "NeighborhoodQuality"), mod=m_freq_signal), cols = 3)
+plot(Effect(c("ChildAge", "Sex"), mod=m_freq_signal), cols = 2)
 ```

### 2025-05-23 Paper.qmd
Minor edits

@@
 ```{r setup, include=FALSE}
-if(!dir.exists("Figures")) dir.create("Figures")
+if (!dir.exists("Figures")) dir.create("Figures")
 
@@
 source("analysis.R")
-#source("analysis2.R") # Generates many plots. Slow!
+# source("analysis2.R") # Generates many plots. Slow!
 ```
@@
 In the context of child signals of need, a child has access to information about their internal state, which is unavailable to others, that allows them to estimate their valuation of additional investment.
-When there are conflicts of interest between children and their caregivers, children might need to use signals whose costs outweigh the benefits of additional care for those not in need, but would result in a net benefit to those with genuine needs.
+When there are conflicts of interest between children and their caregivers, children might need to use signals whose costs outweigh the benefits of additional care for those not in need, e.g., because increased caregiver investment is less valuable to them and/or the opportunity costs of signaling are higher, but would result in a net benefit to those with genuine needs.
 In doing so, the child reveals their resource valuation to potential caretakers who may be skeptical of the child’s claims.
@@
 We conceptualize infrequent and transient sadness signals as minimal signals of need [per @maynardsmith1995].
-Frequent or prolonged sadness, however, is likely a form of withdrawal or depression, both of which might be effective costly signals of need when there are conflicts between the child and their caregivers [@gaffney2022; @syme2023].
+Frequent or prolonged sadness, however, is likely a form of withdrawal or depression, both of which might be effective costly signals of need when there are conflicts between the child and their caregivers [for theory and evidence, see @gaffney2022; @syme2023].
 
@@
   setNames(longform_dict[1:23]) |>
-  datasummary_skim(fun_numeric = list(N = \(x) length(na.omit(x)), Range = \(x) str_glue("{min(x, na.rm=T)}-{max(x, na.rm=T)}"), Mean = Mean, SD = SD, Histogram = function(x) ""))
+  datasummary_skim(
+    fun_numeric = list(
+      N = \(x) length(na.omit(x)),
+      Range = \(x) str_glue("{min(x, na.rm=T)}-{max(x, na.rm=T)}"),
+      Mean = Mean,
+      SD = SD,
+      Histogram = function(x) ""
+    )
+  )
 ```
@@
 utila_df_caregiver |>
-  dplyr::select(any_of(names(utila_df_caregiver[names(longform_dict_adult_cont)]))) |>
+  dplyr::select(any_of(names(utila_df_caregiver[names(
+    longform_dict_adult_cont
+  )]))) |>
   setNames(longform_dict_adult_cont) |>
-  datasummary_skim(fun_numeric = list(N = \(x) length(na.omit(x)), Range = \(x) str_glue("{min(x, na.rm=T)}-{max(x, na.rm=T)}"), Mean = Mean, SD = SD, Histogram = function(x) ""))
+  datasummary_skim(
+    fun_numeric = list(
+      N = \(x) length(na.omit(x)),
+      Range = \(x) str_glue("{min(x, na.rm=T)}-{max(x, na.rm=T)}"),
+      Mean = Mean,
+      SD = SD,
+      Histogram = function(x) ""
+    )
+  )
 ```
@@
 ```{r}
-fmt_sum <- function(v){
+fmt_sum <- function(v) {
   v <- na.omit(v)
@@
   thesum <- sum(v)
-  thepct <- 100*signif(thesum/n, 2)
+  thepct <- 100 * signif(thesum / n, 2)
   str_glue("{thesum} children ({thepct}%)")
@@
 ```{r}
-barplot_SignalFreq + guides(fill = guide_legend(title = 'Times/month', nrow = 1, position = 'top', reverse = T))
+barplot_SignalFreq +
+  guides(
+    fill = guide_legend(
+      title = 'Times/month',
+      nrow = 1,
+      position = 'top',
+      reverse = T
+    )
+  )
 ```

### 2025-05-23 analysis.R

@@ -172,9 +172,9 @@ barplot_SignalFreq <-
       position = "top"
     )
   ) +
-  theme_minimal(15)
+  theme_minimal(20)
 barplot_SignalFreq
-ggsave("Figures/barplot_SignalFreq.png", barplot_SignalFreq)
+ggsave("Figures/barplot_SignalFreq.svg", barplot_SignalFreq)
 
 
 signal_subset <- utila_df[c(
@@ -215,8 +215,9 @@ signal_corrplot <-
   ) +
   guides(fill = guide_colorbar(title = "Correlation coefficients"))
 signal_corrplot
-ggsave("Figures/signal_corrplot.pdf", width = 12, height = 9)
-ggsave("Figures/signal_corrplot.png", width = 12, height = 9)
+ggsave("Figures/signal_corrplot.pdf", signal_corrplot, width = 12, height = 9)
+ggsave("Figures/signal_corrplot.png", signal_corrplot, width = 12, height = 9)
+ggsave("Figures/signal_corrplot.svg", signal_corrplot, width = 12, height = 9)
 
 barplot_conflict <-
   utila_df |>

### 2025-05-23 analysis2.R

@@ -402,15 +402,15 @@ SignalVars4 <-
   relocate(RelativeNeed3, .after = 'childHHid') |>
   mutate(
     across(-c(1:3), as.numeric),
-    across(SadFreqN:AlloparentingXsex, \(x) c(scale(x)))
+    across(SadFreqN:OtherKidsConflict, \(x) c(scale(x)))
   ) |>
   na.omit()
 
 out <- ordinalNetCV(
-  as.matrix(SignalVars4[-c(1:3)]),
+  as.matrix(SignalVars4[-c(1:3, 45)]),
   SignalVars4[[3]],
   standardize = F,
-  alpha = 1.0,
+  alpha = 1,
   family = "cumulative",
   link = "logit",
   lambdaMinRatio = 1e-04,
@@ -419,7 +419,9 @@ out <- ordinalNetCV(
 summary(out)
 colMeans(summary(out))
 # coef(out$fit, matrix = TRUE, whichLambda = 1)
-plot_need_coefs <- ggdotchart(coef(out$fit)[-c(1:2)])
+m_ordinalcv1coefs <- coef(out$fit)[-c(1:2)]
+names(m_ordinalcv1coefs) <- shortform_dict[names(m_ordinalcv1coefs)]
+plot_need_coefs <- ggdotchart(m_ordinalcv1coefs)
 plot_need_coefs
 ggsave("Figures/plot_need_coefs.pdf", plot_need_coefs, width = 12, height = 12)
 ggsave("Figures/plot_need_coefs.png", plot_need_coefs, width = 12, height = 12)
@@ -485,7 +487,7 @@ SignalVars3 <-
   relocate(CaregiverResponse, .after = childHHid) |>
   mutate(
     across(-c(1:3), as.numeric),
-    across(SadFreqN:SeparationAttentionSeeking, \(x) c(scale(x))) # changed StatusConcerns to SeparationAttentionSeeking
+    across(SadFreqN:StatusConcerns, \(x) c(scale(x))) # changed StatusConcerns to SeparationAttentionSeeking
     ) |>
   na.omit()
 

### 2025-05-23 Paper.qmd
Changes made for knitting and plot making (mainly updated dictionaries).

--- 2025-05-12 Paper.qmd
@@
 
+set.seed(51300)
+
 load("data/m3_1e6_cost_freq.rda")
@@
 source("analysis.R")
-#source("analysis2.R") # Generates many plots. Slow!
+source("analysis2.R") # Generates many plots. Slow!
 ```
@@
 
-From these count variables we created three composite signaling variables to be used as outcome variables.
+From these count variables we created two composite signaling variables.
 The *frequency of signaling* was the sum of the frequencies of child sadness, crying, and temper tantrums.
-*Maximum signal frequency* was the frequency of the most common of these three signals.
 *Signal cost* was the sum of the frequencies of these three signals weighted by their relative costliness (i.e., *Signal cost* = *frequency of sadness* + 2 $\times$ *frequency of crying* + 3 $\times$ *frequency of temper tantrums*).
 Although we treated *signal cost* as a continuous variable, we conceptualized low signal cost values (e.g., infrequent sadness or crying) as more likely to be minimal signals, and high signal cost values (e.g., frequent crying or temper tantrums) as more likely to be "costly" signals.
+For each child we also summed the frequencies of conflict with caretakers and signaling cost for all other children in the household, termed *Other children conflict* and *Other children signal cost*, respectively.
 
@@
 
-utila_df |>
-  dplyr::select(any_of(names(utila_df[names(longform_dict[1:23])]))) |>
-  setNames(longform_dict[1:23]) |>
+utila_df2 |>
+  dplyr::select(any_of(names(utila_df2[names(longform_dict[1:25])]))) |>
+  setNames(longform_dict[1:25]) |>
   datasummary_skim(fun_numeric = list(N = \(x) length(na.omit(x)), Range = \(x) str_glue("{min(x, na.rm=T)}-{max(x, na.rm=T)}"), Mean = Mean, SD = SD, Histogram = function(x) ""))
@@
 The proportion of children perceived by caretakers as being more needy, as needy, or less needy than a caretaker's other children when holding continuous variables at their mean and factor variables at their mode.
-A: Effect of child age on the probability of relative neediness B: Effect of child sadness frequency on the probability of relative neediness.
+a: Effect of child age on the probability of relative neediness.
+b: Effect of child sadness frequency on the probability of relative neediness.
 Outcome variable question wording: How much time, effort, and resources does this child need compared to your other children?
@@
 
-Contrary to expectations, immigration status, residence in Camponado, and primary language (Spanish or English as proxies for cultural differences), did not predict child signaling in our lasso regressions, despite the differences in wealth, job certainty, discrimination, and living conditions between those from the mainland and Utilians [@davis2014; @garcia2017; @garcia2018; @stonich2000].
+Contrary to expectations, immigration status, residence in Camponado, and primary language (Spanish or English as proxies for cultural differences), did not predict child signaling in our lasso regressions, despite the differences in wealth, job certainty, discrimination, and living conditions between those from the mainland and English speaking Utilians [@davis2014; @garcia2017; @garcia2018; @stonich2000].
 These variables correlated with neighborhood quality, however, which was retained in our models, and lasso regression tends to retain only one of a set of correlated predictors [@zou2005].
+Additionally, families from Camponado appeared to be willing to care for children outside their immediate families, which might have reduced children's needs.
 
@@
 
-Caregiver responses to signals of need were generally positive or neutral, more so for situations involving injury and illness, and less so for situations involving family conflict, loss of privileges, and transgressions, which had greater proportions of negative responses (@fig-effect-last-signal-al-responses; @fig-effect-last-signal-response).
+Keeping in mind possible biases, caregiver self-reported responses to signals of need were generally positive or neutral, more so for situations involving injury and illness, and less so for situations involving family conflict, loss of privileges, and transgressions, which had greater proportions of negative responses (@fig-effect-last-signal-al-responses; @fig-effect-last-signal-response).
 These patterns probably reflect a combination of shared interests vs.

### 2025-05-23 analysis2.R

@@ -402,11 +402,11 @@ SignalVars4 <-
   relocate(RelativeNeed3, .after = 'childHHid') |>
   mutate(
     across(-c(1:3), as.numeric),
-    across(SadFreqN:AlloparentingXsex, \(x) c(scale(x)))
+    across(SadFreqN:OtherKidsConflict, \(x) c(scale(x)))
   ) |>
   na.omit()
 
-out <- ordinalNetCV(
+m_ordinalcv1 <- ordinalNetCV(
   as.matrix(SignalVars4[-c(1:3)]),
   SignalVars4[[3]],
   standardize = F,
@@ -416,16 +416,18 @@ out <- ordinalNetCV(
   lambdaMinRatio = 1e-04,
   printProgress = T
 )
-summary(out)
-colMeans(summary(out))
-# coef(out$fit, matrix = TRUE, whichLambda = 1)
-plot_need_coefs <- ggdotchart(coef(out$fit)[-c(1:2)])
+summary(m_ordinalcv1)
+colMeans(summary(m_ordinalcv1))
+# coef(m_ordinalcv1$fit, matrix = TRUE, whichLambda = 1)
+m_ordinalcv1coefs <- coef(m_ordinalcv1$fit)[-c(1:2)]
+names(m_ordinalcv1coefs) <- shortform_dict[names(m_ordinalcv1coefs)]
+plot_need_coefs <- ggdotchart(m_ordinalcv1coefs)
 plot_need_coefs
 ggsave("Figures/plot_need_coefs.pdf", plot_need_coefs, width = 12, height = 12)
 ggsave("Figures/plot_need_coefs.png", plot_need_coefs, width = 12, height = 12)
 
-plot_need_age <- ordinal_plot2(out$fit, ChildAge, SignalVars4, 'Relative need', ylabel = "Probability of relative need")
-plot_need_sad <- ordinal_plot2(out$fit, SadFreqN, SignalVars4, 'Relative need', ylabel = "Probability of relative need")
+plot_need_age <- ordinal_plot2(m_ordinalcv1$fit, ChildAge, SignalVars4, 'Relative need', ylabel = "Probability of relative need")
+plot_need_sad <- ordinal_plot2(m_ordinalcv1$fit, SadFreqN, SignalVars4, 'Relative need', ylabel = "Probability of relative need")
 
 plot_need_combined <- plot_need_age / plot_need_sad + ggtitle("") + plot_layout(guides = 'collect') + plot_annotation(tag_levels = "a")
[…]
